### 2012-10-25 paper.tex
initial paper

\documentclass[12pt,preprint]{aastex}
% for \sout
\usepackage{ulem}
% makes sure \em{} is italic rather than underlined (corrects ulem from line above)
\normalem

% for the red MarginPars
\usepackage{color}

% some extra math symbols
\usepackage{mathtools}

\newcommand{\rhocutoff}{\rho_\mathrm{cutoff}}
\newcommand{\rhoanelastic}{\rho_\mathrm{anelastic}}

\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
\newcommand{\Tcutoff}{T_\mathrm{cutoff}}

% MarginPars
\setlength{\marginparwidth}{0.75in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\newcommand{\msolar}{\mathrm{M}_\odot}

\begin{document}

%==========================================================================
% Title
%==========================================================================
\title{WD Merger Notes}

\shorttitle{WD Mergers. I. Methodology}
\shortauthors{Max}

\author{TBD}
%==========================================================================
% Abstract
%==========================================================================
\begin{abstract}
Notes on WD mergers
\end{abstract}
\keywords{hydrodynamics - methods: numerical - nuclear reactions, 
          nucleosynthesis, abundances - supernovae: general - white dwarfs}

%==========================================================================
% Introduction
%==========================================================================
\section{Introduction}

%==========================================================================
% Numerical Methodology
%==========================================================================
\section{Numerical Methodology}\label{sec:Numerical Methodology}

%==========================================================================
% Test Problems
%==========================================================================
\section{Test Problems}\label{Sec:Tests}

%==========================================================================
% Conclusions
%==========================================================================
\section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}

\acknowledgments

\clearpage

%\bibliographystyle{apj}
%\bibliography{ws}

\end{document}

### 2012-11-06 paper.tex
Added test subsections
Filled out gravitational free fall section, added directory with results for this simulation

@@
 %==========================================================================
-\title{WD Merger Notes}
+\title{Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Validation}
 
-\shorttitle{WD Mergers. I. Methodology}
+\shorttitle{DWD Mergers. I. Methodology}
 \shortauthors{Max}
@@
 
+\subsection{Maintaining Hydrostatic Equilibrium}\label{Sec:HSE}
+
+\subsection{Gravitational Free Fall}\label{Sec:Gravitational Free Fall}
+
+A simple test to verify the Poisson solver implemented by CASTRO is the case of gravitational free fall.
+In this setup, two stars, each individually in an equilibrium state, are placed on the computational grid, separated by an initial distance $r_0$ along the $x$ axis with zero initial velocity.
+We choose stars of masses $0.6$ and $0.8\, M_\odot$, with the lower mass star on the left side (e.g.
+$x < 0$) of the domain and the higher mass star on the right, such that their center of mass coincides with the center of the domain.
+Gravitationally, the stars may be treated as point masses until the point of contact, so the equation of motion governing the distance $r$ between their centers of mass is that of simple free fall:
+\begin{equation}
+  \ddot{r}(t) = - \frac{GM}{r},
+\end{equation}
+where $G$ is the gravitational constant and $M$ is the total mass of the system. This differential equation has a closed-form solution for the evolution time as a function of separation:
+\begin{equation}
+  t(r) = \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\,\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right]. \label{analyticalFreeFall}
+\end{equation}
+We determine the initial separation to be consistent with the simulation performed in Section \ref{Sec:Kepler}; that is, we select an initial orbital period and use Kepler's third law to calculate the radius of a circular orbit with that period.
+We select an initial orbital period of 200 s for this simulation, so that the separation is
+\[
+  r_0 = 5.73 \times 10^{9}\ \text{cm}.
+\]
+We chose a relatively low resolution simulation to demonstrate the capabilities of CASTRO even while using modest resources.
+The computational grid is covered by a coarse grid of $48^3$ zones, with two levels of refinement above the coarse grid.
+Each refined grid carries an increase in resolution by a factor of 4 relative to the coarser grid below it.
+CASTRO initially assigns $93\%$ of the domain to be covered by the intermediate resolution grids, and $0.05\%$ of the domain to be covered by the finest resolution grids.
+
+The analytical result in Equation \ref{analyticalFreeFall} determines the total elapsed free-fall time,
+\[
+  t_{\text{ff}} = \frac{\pi}{2} \sqrt{\frac{r_0^3}{2GM}}.
+\]
+The physical radius of each white dwarf is roughly $10\%$ of the initial separation, so we consider the evolution terminated when the radial separation reaches that value (at that point, the stars will no longer be in free-fall due to physical contact).
+Since the elapsed time goes roughly as the square root of the distance for small $r$, we consider the evolution terminated when $t = 0.99\, t_{\text{ff}}$.
+The results of our simulation are shown in Figure \ref{Fig:Free Fall}.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=2.0]{freefall/plot_freefall}
+  \caption{Time evolution of two initially stationary white dwarfs, mutually attracted to each other by the gravitational force.
+The horizontal axis gives the separation of the white dwarfs, scaled to the initial separation, and the vertical axis gives the elapsed time of the simulation, scaled to the total elapsed time necessarily.
+The solid curve shows the analytical result, calculated from Newtonian mechanics, and the circles show the samples from the time evolution with CASTRO.
+The agreement between theory and simulation is excellent, even for a simulation with modest resolution.}
+  \label{Fig:Free Fall}
+\end{figure}
+
+\subsection{Keplerian Orbit}\label{Sec:Kepler}
+
 %==========================================================================

### 2012-12-05 paper.tex
ppm changes

@@
 
+We use the Castro code as described in \citet{castro}.  For all the runs,
+the PPM reconstruction is done, using the original limiters \citep{ppm}.
+The original implementation of the PPM prediction in Castro takes the
+form:
+\begin{equation}
+q_{i+1/2,L}^{n+1/2} = q_i -
+   \sum_{\nu;\lambda^{(\nu)}\ge 0} l^{(\nu)} \cdot \left [
+        q_i - \mathcal{I}_+(\sigma_\nu)
+       \right ] r^{(\nu)}
+\end{equation}
+where $q$ is the vector of primitive variables, with $q_i$
+representing the average in the cell, $l$ and $r$ are the left and
+right eigenvectors with eigenvalue $\lambda$.  The sum is over all the
+waves that result from the characteristic structure of the problem,
+but designed such that only waves moving toward the interface
+contribute to the interface value, $q_{i+1/2,L}^{n+1/2}$.  Finally,
+$\mathcal{I}_+(\sigma_\nu)$ is the jump carried to the right interface
+of the cell by the wave $\nu$, constructed by averaging under the
+parabolic interpolant in the cell all information that can reach the
+interface over the timestep.  $\sigma$ here is the fraction of the
+zone crossed by the wave denoted by $\nu$.  The reader is referred to
+\citet{ppmunsplit} for further defaults.
+
+The presence of $q_i$ in this expression serves as a reference state.
+The idea is that the error in the characteristic projection of the
+jump (due to the nonlinearity) is minimized if we pick a suitable
+reference state, since we only care about the jumps carried to the
+interface over the timestep.  The original Castro implementation used
+the cell-average quantity.  Experiments found that this is ill-behaved
+in the presence of strong shocks (especially at low resolution). For
+the present work, we switch the reference state to the jump carried by the 
+fastest wave moving toward the interface, $\mathcal{I}_+(\sigma_3)$ in this
+case, where the $3$ here means the wave associated with the $u + c$ eigenvalue.
+This is in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
+expression appear as:
+\begin{equation}
+q_{i+1/2,L}^{n+1/2} = \mathcal{I}_+(\sigma_3) -
+   \sum_{\nu;\lambda^{(\nu)}\ge 0} l^{(\nu)} \cdot \left [
+        \mathcal{I}_+(\sigma_3)  - \mathcal{I}_+(\sigma_\nu)
+       \right ] r^{(\nu)}
+\end{equation}
+
+boundary conditions
+
+initial models
+
+AMR
+
 %==========================================================================
@@
 
-A simple test to verify the Poisson solver implemented by CASTRO is the case of gravitational free fall.
-In this setup, two stars, each individually in an equilibrium state, are placed on the computational grid, separated by an initial distance $r_0$ along the $x$ axis with zero initial velocity.
-We choose stars of masses $0.6$ and $0.8\, M_\odot$, with the lower mass star on the left side (e.g.
-$x < 0$) of the domain and the higher mass star on the right, such that their center of mass coincides with the center of the domain.
-Gravitationally, the stars may be treated as point masses until the point of contact, so the equation of motion governing the distance $r$ between their centers of mass is that of simple free fall:
+A simple test to verify the Poisson solver implemented by CASTRO is
+the case of gravitational free fall. In this setup, two stars, each
+individually in an equilibrium state, are placed on the computational
+grid, separated by an initial distance $r_0$ along the $x$ axis with
+zero initial velocity. We choose stars of masses $0.6$ and $0.8\,
+M_\odot$, with the lower mass star on the left side (e.g. $x < 0$) of
+the domain and the higher mass star on the right, such that their
+center of mass coincides with the center of the
+domain. Gravitationally, the stars may be treated as point masses
+until the point of contact, so the equation of motion governing the
+distance $r$ between their centers of mass is that of simple free
+fall:
 \begin{equation}
@@
 \end{equation}
-where $G$ is the gravitational constant and $M$ is the total mass of the system. This differential equation has a closed-form solution for the evolution time as a function of separation:
+where $G$ is the gravitational constant and $M$ is the total mass of
+the system. This differential equation has a closed-form solution for
+the evolution time as a function of separation:
 \begin{equation}
   t(r) = \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\,\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right]. \label{analyticalFreeFall}
-\end{equation}
-We determine the initial separation to be consistent with the simulation performed in Section \ref{Sec:Kepler}; that is, we select an initial orbital period and use Kepler's third law to calculate the radius of a circular orbit with that period.
-We select an initial orbital period of 200 s for this simulation, so that the separation is
+\end{equation} \MarginPar{is there a reference for this? or did you derive it?}
+We determine the initial separation to be consistent with the
+simulation performed in Section \ref{Sec:Kepler}; that is, we select
+an initial orbital period and use Kepler's third law to calculate the
+radius of a circular orbit with that period. We select an initial
+orbital period of 200 s for this simulation, so that the separation is
 \[
@@
 \]
-We chose a relatively low resolution simulation to demonstrate the capabilities of CASTRO even while using modest resources.
-The computational grid is covered by a coarse grid of $48^3$ zones, with two levels of refinement above the coarse grid.
-Each refined grid carries an increase in resolution by a factor of 4 relative to the coarser grid below it.
-CASTRO initially assigns $93\%$ of the domain to be covered by the intermediate resolution grids, and $0.05\%$ of the domain to be covered by the finest resolution grids.
+We chose a relatively low resolution simulation to demonstrate the
+capabilities of CASTRO even while using modest resources. The
+computational grid is covered by a coarse grid of $48^3$ zones, with
+two levels of refinement above the coarse grid. Each refined grid
+carries an increase in resolution by a factor of 4 relative to the
+coarser grid below it. CASTRO initially assigns $93\%$ of the domain
+to be covered by the intermediate resolution grids, and $0.05\%$ of
+the domain to be covered by the finest resolution grids.
 
-The analytical result in Equation \ref{analyticalFreeFall} determines the total elapsed free-fall time,
+The analytical result in Equation \ref{analyticalFreeFall} determines
+the total elapsed free-fall time,
 \[
@@
 \]
-The physical radius of each white dwarf is roughly $10\%$ of the initial separation, so we consider the evolution terminated when the radial separation reaches that value (at that point, the stars will no longer be in free-fall due to physical contact).
-Since the elapsed time goes roughly as the square root of the distance for small $r$, we consider the evolution terminated when $t = 0.99\, t_{\text{ff}}$.
-The results of our simulation are shown in Figure \ref{Fig:Free Fall}.
+The physical radius of each white dwarf is roughly $10\%$ of the
+initial separation, so we consider the evolution terminated when the
+radial separation reaches that value (at that point, the stars will no
+longer be in free-fall due to physical contact). Since the elapsed
+time goes roughly as the square root of the distance for small $r$, we
+consider the evolution terminated when $t = 0.99\, t_{\text{ff}}$. The
+results of our simulation are shown in Figure \ref{Fig:Free Fall}.
 
@@
   \includegraphics[scale=2.0]{freefall/plot_freefall}
-  \caption{Time evolution of two initially stationary white dwarfs, mutually attracted to each other by the gravitational force.
-The horizontal axis gives the separation of the white dwarfs, scaled to the initial separation, and the vertical axis gives the elapsed time of the simulation, scaled to the total elapsed time necessarily.
-The solid curve shows the analytical result, calculated from Newtonian mechanics, and the circles show the samples from the time evolution with CASTRO.
-The agreement between theory and simulation is excellent, even for a simulation with modest resolution.}
+  \caption{Time evolution of two initially stationary white dwarfs,
+    mutually attracted to each other by the gravitational force. The
+    horizontal axis gives the separation of the white dwarfs, scaled
+    to the initial separation, and the vertical axis gives the elapsed
+    time of the simulation, scaled to the total elapsed time
+    necessarily. The solid curve shows the analytical result,
+    calculated from Newtonian mechanics, and the circles show the
+    samples from the time evolution with CASTRO. The agreement between
+    theory and simulation is excellent, even for a simulation with
+    modest resolution.}
   \label{Fig:Free Fall}
@@
 
-%\bibliographystyle{apj}
-%\bibliography{ws}
+\bibliographystyle{apj}
+\bibliography{ws}
 

### 2012-12-05 paper.tex
tweak

@@
 parabolic interpolant in the cell all information that can reach the
-interface over the timestep.  $\sigma$ here is the fraction of the
-zone crossed by the wave denoted by $\nu$.  The reader is referred to
-\citet{ppmunsplit} for further defaults.
+interface over the timestep, with $\sigma$ representing the fraction
+of the zone crossed by the wave denoted by $\nu$ during the timestep.
+The reader is referred to \citet{ppmunsplit} for further defaults.
 

### 2012-12-05 paper.tex
new ref

@@
 reference state, since we only care about the jumps carried to the
-interface over the timestep.  The original Castro implementation used
-the cell-average quantity.  Experiments found that this is ill-behaved
-in the presence of strong shocks (especially at low resolution). For
-the present work, we switch the reference state to the jump carried by the 
-fastest wave moving toward the interface, $\mathcal{I}_+(\sigma_3)$ in this
-case, where the $3$ here means the wave associated with the $u + c$ eigenvalue.
-This is in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
+interface over the timestep \citep{colellaglaz1985}.  The original
+Castro implementation used the cell-average quantity.  Experiments
+found that this is ill-behaved in the presence of strong shocks
+(especially at low resolution). For the present work, we switch the
+reference state to the jump carried by the fastest wave moving toward
+the interface, $\mathcal{I}_+(\sigma_3)$ in this case, where the $3$
+here means the wave associated with the $u + c$ eigenvalue.  This is
+in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
 expression appear as:

### 2012-12-05 paper.tex
document gravity correction

@@
 
+We also modify slightly the correction done to the gravitational
+source terms after integrating the system in time.  In the original
+Castro paper (Eqs. 19 and 20), the state after the conservative update
+due to fluxes and external sources (other than gravity) was
+represented with the superscript `$(2,\star)$'.  The gravitational
+sources were then corrected, making them centered in time, yielding
+the final state at the end of the hydrodynamics step, indicated with
+the superscript `$(2)$'.  This appeared as:
+\begin{eqnarray}
+(\rho {\bf u})^{(2)} &=& (\rho {\bf u})^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf g})^{(2,\star)} - 
+                              (\rho {\bf g})^{(1)} \right ]\\
+(\rho E)^{(2)} &=& (\rho E)^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf u\cdot g})^{(2,\star)} - 
+                              (\rho {\bf u\cdot g})^{(1)} \right ]
+\end{eqnarray}
+Here we make the subtle change of using the corrected ${\bf u}$ in the
+correction to the energy.  So our new energy correction equation appears as:
+\begin{equation}
+(\rho E)^{(2)} = (\rho E)^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf u\cdot g})^{(2)} - 
+                              (\rho {\bf u\cdot g})^{(1)} \right ]
+\end{equation}
+Both of these are second-order accurate, but we've found that the latter
+is slightly better for conserving energy.
+
 boundary conditions
@@
 AMR
+
+rotation
 

### 2012-12-10 paper.tex
Updated some of the free-fall section.

@@
   t(r) = \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\,\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right]. \label{analyticalFreeFall}
-\end{equation} \MarginPar{is there a reference for this? or did you derive it?}
+\end{equation}\MarginPar{This result is derived in the freefall directory; I have edited the text to make it clear how I arrived at this result.}
+This result can be derived by recognizing that
+\[
+  t(r) = \int_{r_0}^r \frac{dr}{v(r)}
+\]
+and inserting the velocity as a function of radial separation,
+\[
+  v(r) = \sqrt{\frac{2GM}{r_0}\left(\frac{r_0}{r} - 1\right)}.
+\]
 We determine the initial separation to be consistent with the
@@
 consider the evolution terminated when $t = 0.99\, t_{\text{ff}}$. The
-results of our simulation are shown in Figure \ref{Fig:Free Fall}.
+results of our simulation are shown in Figure \ref{Fig:Free Fall}. The
+positions of the two stars are determined by calculating the center of
+mass of the right $(x > 0)$ and left $(x < 0)$ sides of the domain at 
+the end of each time step, and treating the centers of mass as the 
+respective location of the two stars.
 

### 2012-12-20 paper.tex
add a plot showing the effect of the reference state change
Changed numerical values for freefall simulation.

@@
   t(r) = \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\,\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right]. \label{analyticalFreeFall}
-\end{equation}\MarginPar{This result is derived in the freefall directory; I have edited the text to make it clear how I arrived at this result.}
+\end{equation}\MarginPar{This result is derived in the freefall directory.}
 This result can be derived by recognizing that
@@
 radius of a circular orbit with that period. We select an initial
-orbital period of 200 s for this simulation, so that the separation is
+orbital period of 100 s for this simulation, so that the separation is
 \[
-  r_0 = 5.73 \times 10^{9}\ \text{cm}.
+  r_0 = 3.61 \times 10^{9}\ \text{cm}.
 \]
@@
 
+\subsection{Keplerian Orbit}\label{Sec:Kepler}
+
+%==========================================================================
+% Conclusions
+%==========================================================================
+\section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}
+
+\acknowledgments
+
+\clearpage
+
+\bibliographystyle{apj}
+\bibliography{ws}
+
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=1.0]{reference}
+  \caption{Solution to Sod's problem with the original reference state and the
+    new reference state, as compared to the exact solution.  We note that the
+    new reference state shows a slightly sharper shock and contact, but also has
+    a dip at the tail of the rarefaction.}
+  \label{Fig:sod}
+\end{figure}
+
+\clearpage
+
 \begin{figure}
@@
 
-\subsection{Keplerian Orbit}\label{Sec:Kepler}
-
-%==========================================================================
-% Conclusions
-%==========================================================================
-\section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}
-
-\acknowledgments
-
-\clearpage
-
-\bibliographystyle{apj}
-\bibliography{ws}
-
 \end{document}

### 2012-12-21 paper.tex
more updates / corrections

@@
 q_{i+1/2,L}^{n+1/2} = q_i -
-   \sum_{\nu;\lambda^{(\nu)}\ge 0} l^{(\nu)} \cdot \left [
-        q_i - \mathcal{I}_+(\sigma_\nu)
-       \right ] r^{(\nu)}
+   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
+        q_i - \mathcal{I}_+^{(\nu)}(q_i)
+       \right ] r_i^{(\nu)}
 \end{equation}
 where $q$ is the vector of primitive variables, with $q_i$
-representing the average in the cell, $l$ and $r$ are the left and
-right eigenvectors with eigenvalue $\lambda$.  The sum is over all the
-waves that result from the characteristic structure of the problem,
-but designed such that only waves moving toward the interface
-contribute to the interface value, $q_{i+1/2,L}^{n+1/2}$.  Finally,
-$\mathcal{I}_+(\sigma_\nu)$ is the jump carried to the right interface
-of the cell by the wave $\nu$, constructed by averaging under the
-parabolic interpolant in the cell all information that can reach the
-interface over the timestep, with $\sigma$ representing the fraction
-of the zone crossed by the wave denoted by $\nu$ during the timestep.
-The reader is referred to \citet{ppmunsplit} for further defaults.
+representing the average in the cell, $l^{(\nu)}$ and $r^{(\nu)}$ are
+the left and right eigenvectors with eigenvalue $\lambda^{(\nu)}$,
+with $\nu$ the index of the characteristic wave of the system.  The
+sum is over all the waves that result from the characteristic
+structure of the problem, but designed such that only waves moving
+toward the interface contribute to the interface value,
+$q_{i+1/2,L}^{n+1/2}$.  Finally, $\mathcal{I}_+^{(\nu)}(q)$ is the
+average under the parabolic profile of quantity $q$ of all the information
+that can reach the right interface of the zone $i$ as carried by the wave
+$\nu$.   The reader is referred to
+\citet{ppmunsplit} for further defaults.
 
@@
 (especially at low resolution). For the present work, we switch the
-reference state to the jump carried by the fastest wave moving toward
-the interface, $\mathcal{I}_+(\sigma_3)$ in this case, where the $3$
-here means the wave associated with the $u + c$ eigenvalue.  This is
+reference state to
+\begin{equation}
+\tilde{q}_L = \left \{ \begin{array}{cc}
+       \mathcal{I}_+^{(+)}(q_i) & \mathrm{if~} u + c > 0 \\
+       q_i                    & \mathrm{otherwise}
+\end{array}
+\right .
+\end{equation}
+where the $(+)$ superscript here means the fastest wave moving to the right
+(the $u+c$ eigenvalue).   This is simply the average under the largest
+portion of the parabolic profile that could possible reach the interface 
+over the timestep.  This is
 in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
@@
 \begin{equation}
-q_{i+1/2,L}^{n+1/2} = \mathcal{I}_+(\sigma_3) -
-   \sum_{\nu;\lambda^{(\nu)}\ge 0} l^{(\nu)} \cdot \left [
-        \mathcal{I}_+(\sigma_3)  - \mathcal{I}_+(\sigma_\nu)
-       \right ] r^{(\nu)}
+q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
+   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
+        \tilde{q_i}  - \mathcal{I}^{(\nu)}_+(q_i)
+       \right ] r_i^{(\nu)}
 \end{equation}
+This change can be used in the public version of Castro by setting
+{\tt castro.ppm\_reference = 1}.
 
@@
 Both of these are second-order accurate, but we've found that the latter
-is slightly better for conserving energy.
+is slightly better for conserving energy.  This change can be used in the
+public version of Castro by setting {\tt castro.grav\_source\_type = 2}.
 

### 2012-12-21 paper.tex
more

@@
 This change can be used in the public version of Castro by setting
-{\tt castro.ppm\_reference = 1}.
+{\tt castro.ppm\_reference = 1}.  Figure~\ref{Fig:sod} shows the
+solution to the Sod problem using the old and new reference states.
+We see that the shock and contact are slightly sharper with this new
+reference state, but also there is a slight dip at the tail of the
+rarefaction with the new method.
 
@@
   \includegraphics[scale=1.0]{reference}
-  \caption{Solution to Sod's problem with the original reference state and the
-    new reference state, as compared to the exact solution.  We note that the
-    new reference state shows a slightly sharper shock and contact, but also has
-    a dip at the tail of the rarefaction.}
-  \label{Fig:sod}
+  \caption{ \label{Fig:sod} Solution to Sod's problem with the original
+    reference state and the new reference state, as compared to the
+    exact solution.  We note that the new reference state shows a
+    slightly sharper shock and contact, but also has a dip at the tail
+    of the rarefaction.}
 \end{figure}

### 2012-12-28 paper.tex
more numerical stuff

@@
 
-boundary conditions
+We note a few additional differences between the original PPM
+implementation of \citet{ppm} and Castro.  In the original
+implementation, the gravitational source term was included in the
+characteristic projection.  Here, we follow \citet{ppmunsplit} and add
+it afterwards.  In our experiments, projecting the source term did not
+make a noticable difference in the energy conservation of the
+solution.  Finally, the PPM paper uses an edge-centered gravitational
+acceleration in the prediction, whereas we use a cell-centered one.
+Both are second-order accurate.  In the unsplit implementation, there
+is an additional freedom that comes into play---do we add the source
+terms to the normal predictor before or after using it to construct
+the transverse flux differences?  Our experiments showed no
+substantial difference with either variant.  We add the sources only
+at the final construction of the interface states.
+
+The boundary conditions on the hyperbolic system are simply
+zero-gradient.  Using AMR, we make the coarse grid very large, so as
+to place the boundaries far from the region of interest.  For the
+gravity solve, we would like isolated boundary conditions.  Here we
+experiment with using two monopoles (one for each star).  The
+individual potentials are summed and this is evaluated on the physical
+boundary to set Dirichlet conditions for the Poisson solve.
 

### 2012-12-28 paper.tex
more organization

@@
 %==========================================================================
-\title{Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Validation}
+\title{Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code 
+       Verification}
 
@@
 \begin{abstract}
-Notes on WD mergers
+We describe our numerical methodology for modeling double white dwarf
+systems with the AMR hydrodynamics code Castro.
+
 \end{abstract}
-\keywords{hydrodynamics - methods: numerical - nuclear reactions, 
-          nucleosynthesis, abundances - supernovae: general - white dwarfs}
+\keywords{hydrodynamics - methods: numerical - supernovae: general - white dwarfs}
 
@@
 
-We use the Castro code as described in \citet{castro}.  For all the runs,
-the PPM reconstruction is done, using the original limiters \citep{ppm}.
-The original implementation of the PPM prediction in Castro takes the
-form:
+\subsection{Hydrodynamics and Gravity}
+
+We use the Castro code as described in \citet{castro}.  For all the
+runs, the PPM reconstruction is done, using the original limiters for
+the parabolic profiles \citep{ppm}.  We modify the prediction of the
+interface states slightly.  The original implementation of the PPM
+prediction in Castro takes the form:
 \begin{equation}
@@
 
-initial models
-
 AMR
@@
 rotation
+
+\subsection{Initial Models and Microphysics}
+
+initial models
+
+EOS
 
@@
 
+should show:
+
+effect of resolution
+
+inertial vs. rotating frame
+
+effect of ppm\_reference and gravity update type
+
+effect of boundary conditions (size of domain and James vs.\ double monopole)
+
 %==========================================================================

### 2013-05-07 paper.tex
add acks
add performance outline
more on flattening

@@
 \end{equation}
+The final change is that now we can no longer simply multiply
+$\tilde{q_i} - \mathcal{I}^{(\nu)}_+(q_i)$ by the flattening
+parameter, $\chi$, but instead blend the traced state with the cell
+centered state as:
+  \begin{equation}
+  q_{i+1/2,\{L,R\}}^{n+1/2} \leftarrow (1 - \chi) q_i + \chi q_{i+1/2,\{L,R\}}^{n+1/2}
+  \end{equation}
+This is the same prescription used in the \cite{ppm}.
+
 This change can be used in the public version of Castro by setting
@@
 %==========================================================================
+% Performance
+%==========================================================================
+\section{Performance}\label{Sec:Performance}
+
+strong scaling plot
+
+space filling curve?
+
+OMP vs. MPI?
+
+breakdown by component?
+
+%==========================================================================
 % Conclusions
@@
 \acknowledgments
+
+This research was supported by NSF award AST-1211563.
+This research used resources of the National Energy Research
+Scientific Computing Center, which is supported by the Office of
+Science of the U.S. Department of Energy under Contract
+No. DE-AC02-05CH11231.
 

### 2013-05-08 paper.tex
add figs
add some more gravity

@@
 
+\subsection{Gravity}
+
+To properly describe the gravitational potential for the system, we need 
+isolated boundary conditions.  The adaptive gravity solve in Castro 
+is described in \cite{castro}, and consists of both level and composite
+solves, and a final synchronization at the end.  To implement isolated
+boundary conditions, we make the simplying assumption that only level
+0 grids will touch the boundary of our computational domain.
+
+Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$
+on a 2-level grid.
+
 %==========================================================================
@@
   \centering
+  \includegraphics[width=3.1in]{nested2}\hspace{1em}
+  \includegraphics[width=3.1in]{nested3}
+  \caption{\label{fig:2levgrav} Schematic showing the solve for $\phi$ for a 2-level grid.}
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+  \centering
   \includegraphics[scale=2.0]{freefall/plot_freefall}

### 2013-05-11 paper.tex
change to default

@@
 
-This change can be used in the public version of Castro by setting
-{\tt castro.ppm\_reference = 1}.  Figure~\ref{Fig:sod} shows the
+This change is now the default in the public version of Castro.
+Figure~\ref{Fig:sod} shows the
 solution to the Sod problem using the old and new reference states.
@@
 To properly describe the gravitational potential for the system, we need 
-isolated boundary conditions.  The adaptive gravity solve in Castro 
-is described in \cite{castro}, and consists of both level and composite
-solves, and a final synchronization at the end.  To implement isolated
-boundary conditions, we make the simplying assumption that only level
-0 grids will touch the boundary of our computational domain.
+isolated boundary conditions.
 

### 2013-05-11 paper.tex
typo in equation
more updates

--- 2013-05-08 paper.tex
@@
    \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
-        \tilde{q_i}  - \mathcal{I}^{(\nu)}_+(q_i)
+        \tilde{q}_L  - \mathcal{I}^{(\nu)}_+(q_i)
        \right ] r_i^{(\nu)}
@@
 implementation of \citet{ppm} and Castro.  In the original
-implementation, the gravitational source term was included in the
-characteristic projection.  Here, we follow \citet{ppmunsplit} and add
+implementation, the gravitational was reconstructed as a parabola,
+and this was traced under to find the forcing that affects the interface
+for each wave.   Here, we follow \citet{ppmunsplit} and add   \MarginPar{we have a version that does the tracing---are we using it?  But there is a $\Delta t/2$ difference still}
 it afterwards.  In our experiments, projecting the source term did not
 make a noticable difference in the energy conservation of the
-solution.  Finally, the PPM paper uses an edge-centered gravitational
-acceleration in the prediction, whereas we use a cell-centered one.
-Both are second-order accurate.  In the unsplit implementation, there
+solution.  In the unsplit implementation, there
 is an additional freedom that comes into play---do we add the source
@@
 zero-gradient.  Using AMR, we make the coarse grid very large, so as
-to place the boundaries far from the region of interest.  For the
-gravity solve, we would like isolated boundary conditions.  Here we
-experiment with using two monopoles (one for each star).  The
-individual potentials are summed and this is evaluated on the physical
-boundary to set Dirichlet conditions for the Poisson solve.
+to place the boundaries far from the region of interest.  We further
+make the restriction that only level-0 (coarse) grids can touch the
+domain boundary.
 
@@
 \subsection{Gravity}
+
+For the
+gravity solve, we would like isolated boundary conditions.  Here we
+experiment with using two monopoles (one for each star).  The
+individual potentials are summed and this is evaluated on the physical
+boundary to set Dirichlet conditions for the Poisson solve.
 

### 2013-05-22 paper.tex
Added subsection on the multipole BCs for the Poisson equation.

@@
 
-For the
-gravity solve, we would like isolated boundary conditions.  Here we
-experiment with using two monopoles (one for each star).  The
-individual potentials are summed and this is evaluated on the physical
-boundary to set Dirichlet conditions for the Poisson solve.
+We solve the Poisson equation for self-gravity for our problem,
+\begin{equation}
+  \nabla^2 \Phi(\mathbf{x}) = -4\pi G\, \rho(\mathbf{x}),
+\end{equation}
+and we note that the sign convention used in CASTRO is opposite to that commonly seen in the physics literature.
+The solution of this equation in CASTRO is described in \cite{castro}, and consists of both level and composite solves, and a final synchronization at the end.
 
-To properly describe the gravitational potential for the system, we need 
-isolated boundary conditions.  The adaptive gravity solve in Castro 
-is described in \cite{castro}, and consists of both level and composite
-solves, and a final synchronization at the end.  To implement isolated
-boundary conditions, we make the simplying assumption that only level
-0 grids will touch the boundary of our computational domain.
+The white dwarf merger problem falls into a class of problems with ``isolated'' boundary conditions, where we assume that all of the mass in the system is enclosed within the physical domain.
+Analytically, the Poisson equation is customarily solved by assuming that the potential vanishes as one gets very far away from the region of non-zero mass density.
+However, on a finite computational domain this assumption does not hold in general: there will be some non-zero value of the gravitational potential on the boundaries of the domain.
+In principle, the Dirichlet boundary values can be computed by doing a direct sum of the mass distribution for each zone outside the domain:
+\begin{equation}
+  \left. \Phi(\mathbf{x}) \right|_{\text{boundary}} = \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.
+\end{equation}
+If there are $N$ zones per spatial dimension, then there are $\mathcal{O}(N^2)$ boundary zones, and each boundary zone requires a sum over $N^3$ zones, so the direct computation of the boundary conditions is an $\mathcal{O}(N^5)$ operation.
+However, in a typical simulation we place the boundaries of the domain far away from the region containing most of the mass, so that they do not contaminate the hydrodynamics in the region of interest at the center of the domain.
+As a result, a common approach is to make some approximation for the mass distribution on the domain that results in a more tractable summation scheme.
 
-Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$
-on a 2-level grid.
+The original release of CASTRO featured a monopole prescription, where the boundary values were computed by summing up all the mass on the domain and treating it as a point source at the domain center.
+Newton's shell theorem shows that this is exact for a spherically symmetric mass distribution, and indeed this is generally a sufficient approximation for single-star Type Ia supernova simulations that employ self-gravity (ref).
+However, for a mass distribution with significant departures from spherical symmetry, this assumption may fail to produce accurate boundary values.
+Furthermore, since the Poisson equation is an elliptic partial differential equation, incorrect boundary values can instantaneously pollute the entire solution.
+We describe here new features we have added to CASTRO to ameliorate this issue.
+
+\subsubsection{Multipole Expansion}
+
+The most natural extension of the monopole prescription is to include higher-order multipole moments.
+If the entire mass distribution is enclosed, then the potential can be expanded as a series of spherical harmonics:
+\begin{equation}
+  \Phi(\mathbf{x}) = \sum_{l=0}^{\infty}\sum_{m=-l}^{l} \frac{4\pi}{2l + 1} q_{lm} \frac{Y_{lm}(\theta,\phi)}{r^{l+1}}, \label{spherical_harmonic_expansion}
+\end{equation}
+where the origin of the coordinate system is taken to be at the center of the computational domain.
+The multipole moments can be calculated by expanding the Green's function for the Poisson equation as a series spherical harmonics, which yields
+\begin{equation}
+  q_{lm} = \int Y^*_{lm}(\theta^\prime, \phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3x^\prime. \label{multipole_moments_original}
+\end{equation}
+Some simplification of Equation \ref{spherical_harmonic_expansion} can be achieved by using the addition theorem for spherical harmonics:
+\begin{align}
+  &\frac{4\pi}{2l+1} \sum_{m=-l}^{l} Y^*_{lm}(\theta^\prime,\phi^\prime)\, Y_{lm}(\theta, \phi) = P_l(\text{cos}\, \theta) P_l(\text{cos}\, \theta^\prime) \notag \\
+  &\ \ + 2 \sum_{m=1}^{l} \frac{(l-m)!}{(l+m)!} P_{l}^{m}(\text{cos}\, \theta)\, P_{l}^{m}(\text{cos}\, \theta^\prime)\, \left[\text{cos}(m\phi)\, \text{cos}(m\phi^\prime) + \text{sin}(m\phi)\, \text{sin}(m\phi^\prime)\right].
+\end{align}
+Here the $P_{l}^{m}$ are the associated Legendre polynomials and the $P_l$ are the Legendre polynomials.
+After some algebraic simplification, the potential outside of the mass distribution can be written in the following way:
+\begin{equation}
+  \Phi(\mathbf{x}) \approx \sum_{l=0}^{l_{\text{max}}} \left[Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} + \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right].
+\end{equation}
+The modified multipole moments are:
+\begin{align}
+  Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime) d^3 x^\prime \\
+  Q_{lm}^{(C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
+  Q_{lm}^{(S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime.
+\end{align}
+Our strategy for the multipole boundary conditions, then, is to pick some value $l_{\text{max}}$ that is of sufficiently high order to capture the distribution of mass on the grid, evaluate the discretized analog of the modified multipole moments for $0 \leq l \leq l_{\text{max}}$ and $1 \leq m \leq l$, and then directly compute the value of the potential on all of the boundary zones.
+This is ultimately an $\mathcal{O}(N^3)$ operation, the same order as the monopole approximation, and the wall time required to calculate the boundary conditions will depend on the chosen value of $l_{\text{max}}$.
+
+\subsubsection{James Algorithm}
+
+Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$ on a 2-level grid.
 

### 2013-05-22 paper.tex
Minor changes to multipole section, and changed name of hydro section.

@@
 
-\subsection{Hydrodynamics and Gravity}
+\subsection{Hydrodynamics}
 
@@
 \begin{equation}
-  \left. \Phi(\mathbf{x}) \right|_{\text{boundary}} = \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.
+  \left. \Phi(\mathbf{x}) \right|_{\text{boundary}} = G\, \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.
 \end{equation}
@@
 \begin{align}
-  Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime) d^3 x^\prime \\
+  Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
   Q_{lm}^{(C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\

### 2013-08-02 paper.tex
move hydro changes to the appendix

@@
 
-We use the Castro code as described in \citet{castro}.  For all the
-runs, the PPM reconstruction is done, using the original limiters for
-the parabolic profiles \citep{ppm}.  We modify the prediction of the
-interface states slightly.  The original implementation of the PPM
-prediction in Castro takes the form:
-\begin{equation}
-q_{i+1/2,L}^{n+1/2} = q_i -
-   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
-        q_i - \mathcal{I}_+^{(\nu)}(q_i)
-       \right ] r_i^{(\nu)}
-\end{equation}
-where $q$ is the vector of primitive variables, with $q_i$
-representing the average in the cell, $l^{(\nu)}$ and $r^{(\nu)}$ are
-the left and right eigenvectors with eigenvalue $\lambda^{(\nu)}$,
-with $\nu$ the index of the characteristic wave of the system.  The
-sum is over all the waves that result from the characteristic
-structure of the problem, but designed such that only waves moving
-toward the interface contribute to the interface value,
-$q_{i+1/2,L}^{n+1/2}$.  Finally, $\mathcal{I}_+^{(\nu)}(q)$ is the
-average under the parabolic profile of quantity $q$ of all the information
-that can reach the right interface of the zone $i$ as carried by the wave
-$\nu$.   The reader is referred to
-\citet{ppmunsplit} for further defaults.
-
-The presence of $q_i$ in this expression serves as a reference state.
-The idea is that the error in the characteristic projection of the
-jump (due to the nonlinearity) is minimized if we pick a suitable
-reference state, since we only care about the jumps carried to the
-interface over the timestep \citep{colellaglaz1985}.  The original
-Castro implementation used the cell-average quantity.  Experiments
-found that this is ill-behaved in the presence of strong shocks
-(especially at low resolution). For the present work, we switch the
-reference state to
-\begin{equation}
-\tilde{q}_L = \left \{ \begin{array}{cc}
-       \mathcal{I}_+^{(+)}(q_i) & \mathrm{if~} u + c > 0 \\
-       q_i                    & \mathrm{otherwise}
-\end{array}
-\right .
-\end{equation}
-where the $(+)$ superscript here means the fastest wave moving to the right
-(the $u+c$ eigenvalue).   This is simply the average under the largest
-portion of the parabolic profile that could possible reach the interface 
-over the timestep.  This is
-in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
-expression appear as:
-\begin{equation}
-q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
-   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
-        \tilde{q}_L  - \mathcal{I}^{(\nu)}_+(q_i)
-       \right ] r_i^{(\nu)}
-\end{equation}
-The final change is that now we can no longer simply multiply
-$\tilde{q_i} - \mathcal{I}^{(\nu)}_+(q_i)$ by the flattening
-parameter, $\chi$, but instead blend the traced state with the cell
-centered state as:
-  \begin{equation}
-  q_{i+1/2,\{L,R\}}^{n+1/2} \leftarrow (1 - \chi) q_i + \chi q_{i+1/2,\{L,R\}}^{n+1/2}
-  \end{equation}
-This is the same prescription used in the \cite{ppm}.
-
-This change is now the default in the public version of Castro.
-Figure~\ref{Fig:sod} shows the
-solution to the Sod problem using the old and new reference states.
-We see that the shock and contact are slightly sharper with this new
-reference state, but also there is a slight dip at the tail of the
-rarefaction with the new method.
-
-We also modify slightly the correction done to the gravitational
-source terms after integrating the system in time.  In the original
-Castro paper (Eqs. 19 and 20), the state after the conservative update
-due to fluxes and external sources (other than gravity) was
-represented with the superscript `$(2,\star)$'.  The gravitational
-sources were then corrected, making them centered in time, yielding
-the final state at the end of the hydrodynamics step, indicated with
-the superscript `$(2)$'.  This appeared as:
-\begin{eqnarray}
-(\rho {\bf u})^{(2)} &=& (\rho {\bf u})^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf g})^{(2,\star)} - 
-                              (\rho {\bf g})^{(1)} \right ]\\
-(\rho E)^{(2)} &=& (\rho E)^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf u\cdot g})^{(2,\star)} - 
-                              (\rho {\bf u\cdot g})^{(1)} \right ]
-\end{eqnarray}
-Here we make the subtle change of using the corrected ${\bf u}$ in the
-correction to the energy.  So our new energy correction equation appears as:
-\begin{equation}
-(\rho E)^{(2)} = (\rho E)^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf u\cdot g})^{(2)} - 
-                              (\rho {\bf u\cdot g})^{(1)} \right ]
-\end{equation}
-Both of these are second-order accurate, but we've found that the latter
-is slightly better for conserving energy.  This change can be used in the
-public version of Castro by setting {\tt castro.grav\_source\_type = 2}.
-
-We note a few additional differences between the original PPM
-implementation of \citet{ppm} and Castro.  In the original
-implementation, the gravitational was reconstructed as a parabola,
-and this was traced under to find the forcing that affects the interface
-for each wave.   Here, we follow \citet{ppmunsplit} and add   \MarginPar{we have a version that does the tracing---are we using it?  But there is a $\Delta t/2$ difference still}
-it afterwards.  In our experiments, projecting the source term did not
-make a noticable difference in the energy conservation of the
-solution.  In the unsplit implementation, there
-is an additional freedom that comes into play---do we add the source
-terms to the normal predictor before or after using it to construct
-the transverse flux differences?  Our experiments showed no
-substantial difference with either variant.  We add the sources only
-at the final construction of the interface states.
+We use the unsplit PPM solver in Castro to advance the hydrodynamics
+system in time.  A number of changes were made to the solver, which are 
+detailed in the Appendix.  These changes bring the algorithm more in 
+line with that of \cite{ppm}.
 
@@
 \clearpage
+\appendix
+
+\section{Castro hydrodynamics changes}
+
+Summary of changes:
+\begin{itemize}
+\item New reference state/flattening fix: helps energy conservation a
+  lot, also fixes some undershoot/overshoots
+
+\item New gravity update type: fixes some energy conservation
+
+\item temperature-based PPM: fixes temperature floors
+
+\item changing I's to use limit of the parabola instead of the
+  cell-center: not sure if this does much
+
+\item Colella \& Glaz Riemann solver: does a much better job with
+  temperature in strong shocks
+
+\item Tracing of gravity: seems to help with the material at the edge
+  of the star
+\end{itemize}
+
+We use the Castro code as described in \citet{castro}.  For all the
+runs, the PPM reconstruction is done, using the original limiters for
+the parabolic profiles \citep{ppm}.  We modify the prediction of the
+interface states slightly.  The original implementation of the PPM
+prediction in Castro takes the form:
+\begin{equation}
+q_{i+1/2,L}^{n+1/2} = q_i -
+   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
+        q_i - \mathcal{I}_+^{(\nu)}(q_i)
+       \right ] r_i^{(\nu)}
+\end{equation}
+where $q$ is the vector of primitive variables, with $q_i$
+representing the average in the cell, $l^{(\nu)}$ and $r^{(\nu)}$ are
+the left and right eigenvectors with eigenvalue $\lambda^{(\nu)}$,
+with $\nu$ the index of the characteristic wave of the system.  The
+sum is over all the waves that result from the characteristic
+structure of the problem, but designed such that only waves moving
+toward the interface contribute to the interface value,
+$q_{i+1/2,L}^{n+1/2}$.  Finally, $\mathcal{I}_+^{(\nu)}(q)$ is the
+average under the parabolic profile of quantity $q$ of all the information
+that can reach the right interface of the zone $i$ as carried by the wave
+$\nu$.   The reader is referred to
+\citet{ppmunsplit} for further defaults.
+
+The presence of $q_i$ in this expression serves as a reference state.
+The idea is that the error in the characteristic projection of the
+jump (due to the nonlinearity) is minimized if we pick a suitable
+reference state, since we only care about the jumps carried to the
+interface over the timestep \citep{colellaglaz1985}.  The original
+Castro implementation used the cell-average quantity.  Experiments
+found that this is ill-behaved in the presence of strong shocks
+(especially at low resolution). For the present work, we switch the
+reference state to
+\begin{equation}
+\tilde{q}_L = \left \{ \begin{array}{cc}
+       \mathcal{I}_+^{(+)}(q_i) & \mathrm{if~} u + c > 0 \\
+       q_i                    & \mathrm{otherwise}
+\end{array}
+\right .
+\end{equation}
+where the $(+)$ superscript here means the fastest wave moving to the right
+(the $u+c$ eigenvalue).   This is simply the average under the largest
+portion of the parabolic profile that could possible reach the interface 
+over the timestep.  This is
+in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
+expression appear as:
+\begin{equation}
+q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
+   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
+        \tilde{q}_L  - \mathcal{I}^{(\nu)}_+(q_i)
+       \right ] r_i^{(\nu)}
+\end{equation}
+The final change is that now we can no longer simply multiply
+$\tilde{q_i} - \mathcal{I}^{(\nu)}_+(q_i)$ by the flattening
+parameter, $\chi$, but instead blend the traced state with the cell
+centered state as:
+  \begin{equation}
+  q_{i+1/2,\{L,R\}}^{n+1/2} \leftarrow (1 - \chi) q_i + \chi q_{i+1/2,\{L,R\}}^{n+1/2}
+  \end{equation}
+This is the same prescription used in the \cite{ppm}.
+
+This change is now the default in the public version of Castro.
+Figure~\ref{Fig:sod} shows the
+solution to the Sod problem using the old and new reference states.
+We see that the shock and contact are slightly sharper with this new
+reference state, but also there is a slight dip at the tail of the
+rarefaction with the new method.
+
+We also modify slightly the correction done to the gravitational
+source terms after integrating the system in time.  In the original
+Castro paper (Eqs. 19 and 20), the state after the conservative update
+due to fluxes and external sources (other than gravity) was
+represented with the superscript `$(2,\star)$'.  The gravitational
+sources were then corrected, making them centered in time, yielding
+the final state at the end of the hydrodynamics step, indicated with
+the superscript `$(2)$'.  This appeared as:
+\begin{eqnarray}
+(\rho {\bf u})^{(2)} &=& (\rho {\bf u})^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf g})^{(2,\star)} - 
+                              (\rho {\bf g})^{(1)} \right ]\\
+(\rho E)^{(2)} &=& (\rho E)^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf u\cdot g})^{(2,\star)} - 
+                              (\rho {\bf u\cdot g})^{(1)} \right ]
+\end{eqnarray}
+Here we make the subtle change of using the corrected ${\bf u}$ in the
+correction to the energy.  So our new energy correction equation appears as:
+\begin{equation}
+(\rho E)^{(2)} = (\rho E)^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf u\cdot g})^{(2)} - 
+                              (\rho {\bf u\cdot g})^{(1)} \right ]
+\end{equation}
+Both of these are second-order accurate, but we've found that the latter
+is slightly better for conserving energy.  This change can be used in the
+public version of Castro by setting {\tt castro.grav\_source\_type = 2}.
+
+We note a few additional differences between the original PPM
+implementation of \citet{ppm} and Castro.  In the original
+implementation, the gravitational was reconstructed as a parabola,
+and this was traced under to find the forcing that affects the interface
+for each wave.   Here, we follow \citet{ppmunsplit} and add   \MarginPar{we have a version that does the tracing---are we using it?  But there is a $\Delta t/2$ difference still}
+it afterwards.  In our experiments, projecting the source term did not
+make a noticable difference in the energy conservation of the
+solution.  In the unsplit implementation, there
+is an additional freedom that comes into play---do we add the source
+terms to the normal predictor before or after using it to construct
+the transverse flux differences?  Our experiments showed no
+substantial difference with either variant.  We add the sources only
+at the final construction of the interface states.
+
+\clearpage
 

### 2013-08-02 paper.tex
initial model stuff + move hydro to appendix

@@
 
-initial models
+We generate initial models by integrating the equation of hydrostatic
+equilibrium, taking the temperature and composition to be constant,
+and using the general stellar equation of state.  This results in
+a single nonlinear equation to find the density in a zone given the
+conditions in the zone beneath it:
+\begin{equation}
+\frac{p(\rho_{i+1}) - p_i}{\Delta x} = \frac{1}{2} (\rho_i + \rho_{i+1}) g_{i+1/2}
+\end{equation}
+This is solved via Newton iteration to give $\rho_{i+1}$.  Here, $g_{i+1/2}$
+is the gravitational acceleration at the interface between zones $i$ and $i+1$,
+found by simply adding up all the mass from zones $1, \ldots, i$ to get the
+enclosed mass, $M_{i+1/2}$, and then $g_{i+1/2} = GM_{i+1/2}/r_{i+1/2}^2$.
 
-EOS
+To start the integration off, we need a central density.  We guess at
+this, and then iterate over the entire integration procedure to find
+the central density needed to yield the desired total mass.  Finally,
+we generate the initial model at the same resolution as the finest
+grid in our simulation.
+
+EOS (no Coulomb corrections)
 

### 2013-08-02 paper.tex
add more ack

@@
 
-This research was supported by NSF award AST-1211563.
-This research used resources of the National Energy Research
-Scientific Computing Center, which is supported by the Office of
-Science of the U.S. Department of Energy under Contract
-No. DE-AC02-05CH11231.
+This research was supported by NSF award AST-1211563.  This research
+used resources of the National Energy Research Scientific Computing
+Center, which is supported by the Office of Science of the
+U.S. Department of Energy under Contract No. DE-AC02-05CH11231.  An
+award of computer time was provided by the Innovative and Novel
+Computational Impact on Theory and Experiment (INCITE) program.  This
+research used resources of the Oak Ridge Leadership Computing Facility
+located in the Oak Ridge National Laboratory, which is supported by
+the Office of Science of the Department of Energy under Contract
+DE-AC05-00OR22725.
 

### 2013-08-02 paper.tex
fix sign error

@@
 found by simply adding up all the mass from zones $1, \ldots, i$ to get the
-enclosed mass, $M_{i+1/2}$, and then $g_{i+1/2} = GM_{i+1/2}/r_{i+1/2}^2$.
+enclosed mass, $M_{i+1/2}$, and then $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$.
 

### 2013-08-02 paper.tex
more methods

@@
 
-\subsection{Initial Models and Microphysics}
+\subsection{Microphysics}
+
+EOS (no Coulomb corrections)
+
+\subsection{Initial Models}
 
@@
 
-EOS (no Coulomb corrections)
+We map the 1-d model onto the 3-d Cartesian grid by taking density,
+temperature, and composition as the independent variables,
+interpolating these to the cell-centers, and then calling the equation
+of state to initialize the remaining terms.  The velocity is taken to
+be zero in the rotating frame initially.
 

### 2013-08-22 paper.tex
more items of change
more Castro change documentating

@@
 \newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
+
+\newcommand{\evm}{{(-)}}
+\newcommand{\evz}{{(\circ)}}
+\newcommand{\evp}{{(+)}}
+\newcommand{\enu}{{(\nu)}}
 
@@
   of the star
+
+\item implicit update of the rotation terms?  (I coded this up once, but
+  it is not currently in the code)
 \end{itemize}
@@
 \citet{ppmunsplit} for further defaults.
+
+\subsection{Reference States}
 
@@
 
-This change is now the default in the public version of Castro.
+This change is now the default in the public version of Castro,
+and can be controlled with the parameter {\tt castro.ppm\_reference}.
 Figure~\ref{Fig:sod} shows the
@@
 rarefaction with the new method.
+
+\subsection{Gravity Source Correction}
 
@@
 \end{equation}
-Both of these are second-order accurate, but we've found that the latter
-is slightly better for conserving energy.  This change can be used in the
-public version of Castro by setting {\tt castro.grav\_source\_type = 2}.
+Both of these are second-order accurate, but we've found that the
+latter is slightly better for conserving energy.  This change is now
+the default in the public version of Castro (it can be controlled by
+{\tt castro.grav\_source\_type}.
+
+\subsection{Gravity Tracing}
 
 We note a few additional differences between the original PPM
-implementation of \citet{ppm} and Castro.  In the original
-implementation, the gravitational was reconstructed as a parabola,
-and this was traced under to find the forcing that affects the interface
-for each wave.   Here, we follow \citet{ppmunsplit} and add   \MarginPar{we have a version that does the tracing---are we using it?  But there is a $\Delta t/2$ difference still}
-it afterwards.  In our experiments, projecting the source term did not
-make a noticable difference in the energy conservation of the
-solution.  In the unsplit implementation, there
-is an additional freedom that comes into play---do we add the source
-terms to the normal predictor before or after using it to construct
-the transverse flux differences?  Our experiments showed no
-substantial difference with either variant.  We add the sources only
-at the final construction of the interface states.
+implementation of \citet{ppm} and Castro.  In the original PPM
+implementation, the gravitational was reconstructed as a parabola, and
+this was traced under to find the forcing that affects the interface
+for each wave.  Castro followed \citet{ppmunsplit} which instead adds
+$(\Delta t/2)g$ to the interface states at the end of the
+reconstruction.  In the current implementation, we do the original
+parabolic reconstruction and characteristic tracing of the gravitation
+source.  Our system with the source appears as:
+\begin{equation}
+q_t + A(q) q_x = G
+\end{equation}
+where $G = (0, g, 0)^T$---i.e. the gravitational source only affects
+$u$, not $\rho$ or $p$.  Note that in the PPM paper, they put $G$ on 
+the lefthand side of the primitive variable equation, so our signs are
+opposite.  Our projections are now:
+\begin{equation}
+\sum_{\nu; \lambda^\enu \ge 0}l^\enu \cdot (\tilde{q} - \mathcal{I}^\enu_+(q) - \tfrac{\Delta t}{2} G) r^\enu
+\end{equation}
+for the left state, and
+\begin{equation}
+\sum_{\nu; \lambda^\enu \le 0} l^\enu \cdot (\tilde{q} - \mathcal{I}^\enu_-(q) - \tfrac{\Delta t}{2} G) r^\enu 
+\end{equation}
+for the right state.  Since $G$ is only non-zero for velocity, only
+the velocity changes.  Writing out the sum (and performing the vector products), we
+get:
+\begin{eqnarray}
+u_{i+1/2,L}^{n+1/2} =
+   \tilde{u}_+ 
+  &-& \frac{1}{2} \left [
+      \left (\tilde{u}_+ - \mathcal{I}_+^\evm(u) - \frac{\Delta t}{2} \mathcal{I}^\evm_+(g) \right ) - 
+       \frac{\tilde{p}_+ - \mathcal{I}_+^\evm(p)}{C} \right ] \nonumber \\
+  &-& \frac{1}{2} \left [
+      \left (\tilde{u}_+ - \mathcal{I}_+^\evp(u) - \frac{\Delta t}{2} \mathcal{I}^\evp_+(g) \right ) +
+       \frac{\tilde{p}_+ - \mathcal{I}_+^\evp(p)}{C} \right ]
+\end{eqnarray}
+These differ from the expression in the PPM paper, where $\Delta t G$,
+not $\Delta t/2 G$ is used in the projection, however we believe that
+the factor of $1/2$ is correct.  To see this, notice that if both
+waves are moving toward the interface, then the source term that is
+added to the interface state is $(\Delta t/4) (\mathcal{I}_+^\evm(g) +
+\mathcal{I}_+^\evp(g))$ for the left state, which reduces to $(\Delta
+t/2) g$ for constant g---this matches the result from Taylor
+expanding to the interface at the half-time (as in \citealt{ppmunsplit}).
+
+There is one additional effect of this change---now the gravitational
+source is seen by all Riemann solves (including the transverse solves)
+whereas previously it was only added to the final unsplit interface
+states.  Both methods are second-order accurate.
+
+The tracing of gravity can be controlled in the public version of Castro
+with the parmeter {\tt castro.ppm\_trace\_grav}.
 

### 2013-08-22 paper.tex
more on reference states

@@
 \begin{equation}
+\label{eq:refchoice}
 \tilde{q}_L = \left \{ \begin{array}{cc}
@@
 \begin{equation}
+\label{eq:ppmstatel}
 q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
@@
 rarefaction with the new method.
+
+It is instructive to look at the effect of the reference state choice.
+First, consider one of the variables present in one-dimensional flow
+(density, velocity in the normal direction, and pressure).  Here we
+our reference state is as in Eq.~\ref{eq:refchoice} and we have.
+Ignoring flattening, if there are no waves moving toward our
+interface, then Eq.~\ref{eq:ppmstatel} reduces to:
+\begin{equation}
+q_{i+1/2,L}^{n+1/2} = \tilde{q}_L = q_i
+\end{equation}
+If instead only the fastest wave is moving toward the interface, then
+only the term corresponding to the fastest wave in the sum will be
+added in Eq.~\ref{eq:ppmstatel}, but our choice of reference state makes
+that term 0 by design, and our interface state is:
+\begin{equation}
+q_{i+1/2,L}^{n+1/2} = \tilde{q}_L = \mathcal{I}_+^{+}(q_i)
+\end{equation}
+This is the desired behavior for each of these cases.
+
+The details of the reference state for passively advected quantities 
+(which includes the transverse velocities in the 1-d tracing) is not
+typically discussed.  If we use the same idea of the reference state as
+in Eq.~\ref{eq:refchoice}, and consider a quantity $\xi$ which should only
+jump across the contact, then our interface state becomes:
+\begin{equation}
+\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L -
+  \underbrace{l_i^\evz \cdot \left [
+        \tilde{\xi}_L  - \mathcal{I}^\evz_+(\xi_i)
+       \right ] r_i^\evz}_{\text{only if~$u \ge 0$}}
+\end{equation}
+Again, ignoring flattening, if $u \ge 0$, then we have
+\begin{equation}
+\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L -
+  \left (\tilde{\xi}_L  - \mathcal{I}^\evz_+(\xi_i) \right ) = \mathcal{I}^\evz_+(\xi_i)
+\end{equation}
+(where we used the fact that the eigenvectors are normalized to 1 and
+don't mix in any other states when dealing with passive terms).  This
+is the expected behavior---we see a state that is traced only by the
+contact wave.  However, if $u < 0$ but $u + c \ge 0$, then we instead get:
+\begin{equation}
+\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L = \mathcal{I}^\evp_+(\xi_i)
+\end{equation}
+Here we used the same definition of the reference state and see that our
+interface state sees the profile traced under the fastest wave, not the
+contact.  This is not the correct behavior for a passively-advected
+quantity.
+
+The fix for passively-advected quantities is to simply ignore the 
+idea of a reference state and just test on the speed of the contact
+itself, setting:
+\begin{equation}
+\xi_{i+1/2,L}^{n+1/2} = \left \{ \begin{array}{cc}
+       \mathcal{I}_+^\evz(\xi_i) & \mathrm{if~} u  > 0 \\
+       \xi_i                    & \mathrm{otherwise}
+\end{array}
+\right .
+\end{equation}
 

### 2013-08-26 paper.tex
add description of the implicit rotation update

@@
 
-\subsection{Gravity Source Correction}
-
-We also modify slightly the correction done to the gravitational
-source terms after integrating the system in time.  In the original
-Castro paper (Eqs. 19 and 20), the state after the conservative update
-due to fluxes and external sources (other than gravity) was
-represented with the superscript `$(2,\star)$'.  The gravitational
-sources were then corrected, making them centered in time, yielding
-the final state at the end of the hydrodynamics step, indicated with
-the superscript `$(2)$'.  This appeared as:
-\begin{eqnarray}
-(\rho {\bf u})^{(2)} &=& (\rho {\bf u})^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf g})^{(2,\star)} - 
-                              (\rho {\bf g})^{(1)} \right ]\\
-(\rho E)^{(2)} &=& (\rho E)^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf u\cdot g})^{(2,\star)} - 
-                              (\rho {\bf u\cdot g})^{(1)} \right ]
-\end{eqnarray}
-Here we make the subtle change of using the corrected ${\bf u}$ in the
-correction to the energy.  So our new energy correction equation appears as:
-\begin{equation}
-(\rho E)^{(2)} = (\rho E)^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf u\cdot g})^{(2)} - 
-                              (\rho {\bf u\cdot g})^{(1)} \right ]
-\end{equation}
-Both of these are second-order accurate, but we've found that the
-latter is slightly better for conserving energy.  This change is now
-the default in the public version of Castro (it can be controlled by
-{\tt castro.grav\_source\_type}.
-
 \subsection{Gravity Tracing}
@@
 
+\subsection{Gravity Source Correction}
+
+We also modify slightly the correction done to the gravitational
+source terms after integrating the system in time.  In the original
+Castro paper (Eqs. 19 and 20), the state after the conservative update
+due to fluxes and external sources (other than gravity) was
+represented with the superscript `$(2,\star)$'.  The gravitational
+sources were then corrected, making them centered in time, yielding
+the final state at the end of the hydrodynamics step, indicated with
+the superscript `$(2)$'.  This appeared as:
+\begin{eqnarray}
+(\rho {\bf U})^{(2)} &=& (\rho {\bf U})^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf g})^{(2,\star)} - 
+                              (\rho {\bf g})^{(1)} \right ]\\
+(\rho E)^{(2)} &=& (\rho E)^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf U\cdot g})^{(2,\star)} - 
+                              (\rho {\bf U\cdot g})^{(1)} \right ]
+\end{eqnarray}
+Here we make the subtle change of using the corrected ${\bf u}$ in the
+correction to the energy.  So our new energy correction equation appears as:
+\begin{equation}
+(\rho E)^{(2)} = (\rho E)^{(2,\star)} +
+   \frac{\Delta t}{2} \left [ (\rho {\bf U\cdot g})^{(2)} - 
+                              (\rho {\bf U\cdot g})^{(1)} \right ]
+\end{equation}
+Both of these are second-order accurate, but we've found that the
+latter is slightly better for conserving energy.  This change is now
+the default in the public version of Castro (it can be controlled by
+{\tt castro.grav\_source\_type}.
+
+Rotation is handled similarly to gravity during the conservative update.
+The momentum equation, with rotation sources appears as:
+\begin{equation}
+\frac{\partial \rho {\bf U}}{\partial t} + \nabla \cdot (\rho {\bf U U} + p) = 
+   \rho {\bf g} - 2 \rho {\bf \Omega \times U} 
+                - \rho {\bf \Omega \times (\Omega \ times U)}
+\end{equation}
+we take the rotation axis to be in the $z$-direction, ${\bf \Omega} =
+\Omega \hat{\bf k}$.  Working out the cross-products, the rotation
+forces appear as
+\begin{equation}
+{\bf F}_\mathrm{rot} = \rho \left (
+    \begin{array}{c}
+     \phantom{-}2 \Omega v + \Omega^2 x \\
+              - 2 \Omega u + \Omega^2 y \\
+              0 
+    \end{array}
+  \right )
+\end{equation}
+We want to time-center this force in the final conservative update, giving
+updates for the $u$ and $v$ components of the velocity as:
+\begin{eqnarray}
+\frac{(\rho u)^{n+1} - (\rho u)^n}{\Delta t} + A^{(x),n+1/2} &=&
+   \frac{1}{2} \left [ (\rho {\bf g}\cdot \hat{\bf i})^n
+                      + (\rho {\bf g}\cdot \hat{\bf i})^{n+1} \right ] \nonumber\\&+&
+   \frac{1}{2} \left [ (2\rho \Omega v)^n + (2\rho \Omega v)^{n+1} \right ] \nonumber \\&+&
+   \frac{1}{2} \left [ (\rho \Omega^2 x)^n + (\rho \Omega^2 x)^{n+1} \right ] \\
+\frac{(\rho v)^{n+1} - (\rho v)^n}{\Delta t} + A^{(y),n+1/2} &=&
+   \frac{1}{2} \left [ (\rho {\bf g}\cdot \hat{\bf j})^n +
+                       (\rho {\bf g}\cdot \hat{\bf j})^{n+1} \right ] \nonumber \\ &+&
+   \frac{1}{2} \left [ (-2\rho \Omega u)^n + (-2\rho \Omega u)^{n+1} \right ] \nonumber \\ &+&
+   \frac{1}{2} \left [ (\rho \Omega^2 y)^n + (\rho \Omega^2 y)^{n+1} \right ]
+\end{eqnarray}
+where here the explicit advective term from the Godunov method is represented as $A$ and is already time-centered.
+In the Castro algorithm, a full $\Delta t$ of the old-time-level source is first added and later we correct this by subtracting off $\Delta t/2$ of the old-time-level term and adding $\Delta t/2$ of the new-time-level source.
+This correction appears as:
+\begin{eqnarray}
+(\rho u)^{n+1} = (\rho u)^\prime 
+   &+& \underbrace{
+      \frac{\Delta t}{2} \left [ (\rho {\bf g}\cdot \hat{\bf i})^{n+1}
+                               - (\rho {\bf g}\cdot \hat{\bf i})^{n} \right ]
+    }_{\equiv \Delta^{g,x}} \nonumber \\
+   &+& \frac{\Delta t}{2} \left [ (2\rho \Omega v)^{n+1} 
+                                - (2\rho \Omega v)^{n} \right ] \nonumber \\
+   &+& \underbrace{
+      \frac{\Delta t}{2} \left [ (\rho \Omega^2 x)^{n+1} 
+                               - (\rho \Omega^2 x)^{n} \right ]
+    }_{\equiv \Delta^{c,x}} \\
+%
+(\rho v)^{n+1} = (\rho v)^\prime 
+   &+& \underbrace{
+      \frac{\Delta t}{2} \left [ (\rho {\bf g}\cdot \hat{\bf j})^{n+1}
+                               - (\rho {\bf g}\cdot \hat{\bf j})^{n} \right ]
+    }_{\equiv \Delta^{g,y}} \nonumber \\
+   &+& \frac{\Delta t}{2} \left [ (-2\rho \Omega u)^{n+1} 
+                                - (-2\rho \Omega u)^{n} \right ] \nonumber \\
+   &+& \underbrace{
+      \frac{\Delta t}{2} \left [ (\rho \Omega^2 y)^{n+1} 
+                               - (\rho \Omega^2 y)^{n} \right ]
+    }_{\equiv \Delta^{c,y}} \\
+\end{eqnarray}
+For compactness, we define the following:
+\begin{eqnarray}
+S_x &=& (\rho u)^\prime + \Delta^{g,x} + \Delta^{c,x} 
+        - \frac{\Delta t}{2} (2 \rho \Omega v)^n \\
+%
+S_y &=& (\rho v)^\prime + \Delta^{g,y} + \Delta^{c,y} 
+        + \frac{\Delta t}{2} (2 \rho \Omega u)^n 
+\end{eqnarray}
+and our updates are now a coupled implicit system for the updates of
+$u$ and $v$:
+\begin{eqnarray}
+(\rho u)^{n+1} = S_x + \frac{\Delta t}{2}(2\rho \Omega v)^{n+1} \label{eq:rhouupdate}\\
+(\rho v)^{n+1} = S_y - \frac{\Delta t}{2}(2\rho \Omega u)^{n+1} 
+\end{eqnarray}
+This can be solved analytically:
+\begin{equation}
+(\rho v)^{n+1} = \frac{S_y - \Delta t \Omega S_x}{1 + \Delta t^2 \Omega^2}
+\end{equation}
+and then we can find $(\rho u)^{n+1}$ from Eq.~\ref{eq:rhouupdate}.
 \clearpage

### 2014-04-30 paper.tex
typo
Some updates for wording and clarity for the gravitational boundary conditions.

@@
 \end{equation}
-and we note that the sign convention used in CASTRO is opposite to that commonly seen in the physics literature.
+where $\Phi$ is the gravitational potential, $G$ is the gravitational constant, and $\rho$ is the mass density.
+We note that the sign convention used in CASTRO is opposite to that commonly seen in the physics literature, so that $\Phi$ is positive everywhere.
 The solution of this equation in CASTRO is described in \cite{castro}, and consists of both level and composite solves, and a final synchronization at the end.
 
-The white dwarf merger problem falls into a class of problems with ``isolated'' boundary conditions, where we assume that all of the mass in the system is enclosed within the physical domain.
-Analytically, the Poisson equation is customarily solved by assuming that the potential vanishes as one gets very far away from the region of non-zero mass density.
-However, on a finite computational domain this assumption does not hold in general: there will be some non-zero value of the gravitational potential on the boundaries of the domain.
-In principle, the Dirichlet boundary values can be computed by doing a direct sum of the mass distribution for each zone outside the domain:
+Analytic solutions to the Poisson equation customarily assume that the potential vanishes at large distances from the region of non-zero density.
+However, on a finite computational domain it is usually not possible to have the edges of the domain be far enough away that the potential can be taken to be zero there.
+Solving the Poisson equation therefore requires knowledge of the values of the potential on the edges of the computational domain.
+In principle, they can be computed by doing a direct sum over the mass distribution inside the domain, where the mass in each zone is treated as a point source:
 \begin{equation}
-  \left. \Phi(\mathbf{x}) \right|_{\text{boundary}} = G\, \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.
+  \Phi_{\text{lmn}} = G\, \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x}_{\text{lmn}} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.
 \end{equation}
-If there are $N$ zones per spatial dimension, then there are $\mathcal{O}(N^2)$ boundary zones, and each boundary zone requires a sum over $N^3$ zones, so the direct computation of the boundary conditions is an $\mathcal{O}(N^5)$ operation.
-However, in a typical simulation we place the boundaries of the domain far away from the region containing most of the mass, so that they do not contaminate the hydrodynamics in the region of interest at the center of the domain.
-As a result, a common approach is to make some approximation for the mass distribution on the domain that results in a more tractable summation scheme.
+Here (i, j, k) are the indices of cells inside the domain, and (l, m, n) are the indices of boundary locations.
+$\Delta V$ is the volume of the zone.
+We have implemented this as an option\footnote{It is controlled with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in CASTRO.
+If there are $N$ zones per spatial dimension, then there are $6 N^2$ boundary zones, and each boundary zone requires a sum over $N^3$ zones, so the direct computation of the boundary conditions scales as $N^5$.
+This method is expensive enough that it is not used for hydrodynamics simulations (though it is useful for comparison to the approximate solutions described below).
 
-The original release of CASTRO featured a monopole prescription, where the boundary values were computed by summing up all the mass on the domain and treating it as a point source at the domain center.
-Newton's shell theorem shows that this is exact for a spherically symmetric mass distribution, and indeed this is generally a sufficient approximation for single-star Type Ia supernova simulations that employ self-gravity (ref).
-However, for a mass distribution with significant departures from spherical symmetry, this assumption may fail to produce accurate boundary values.
-Furthermore, since the Poisson equation is an elliptic partial differential equation, incorrect boundary values can instantaneously pollute the entire solution.
-We describe here new features we have added to CASTRO to ameliorate this issue.
+In a typical simulation we place the boundaries of the domain far enough away from the region containing most of the mass that some method of approximation to this direct summation is justified.
+Many approaches exist in the literature.
+The original release of CASTRO featured the crudest possible approximation: a monopole prescription, where the boundary values were computed by summing up all the mass on the domain and treating it as a point source at the domain center.
+This is exactly correct only for a spherically symmetric mass distribution, and indeed this is often a sufficient approximation for single-star Type Ia supernova simulations that employ self-gravity (ref).
+However, for a problem like that of a binary star system with significant departures from spherical symmetry, this assumption fails to produce accurate boundary values.
+This results in a significant drift of the center of the mass of the system over time.
+We describe here two new features we have added to CASTRO to address this issue.
 
@@
 The most natural extension of the monopole prescription is to include higher-order multipole moments.
-If the entire mass distribution is enclosed, then the potential can be expanded as a series of spherical harmonics:
+If the entire mass distribution is enclosed, then the potential can be expanded in a series of spherical harmonics $Y_{lm}$:
 \begin{equation}
@@
 \end{equation}
-where the origin of the coordinate system is taken to be at the center of the computational domain.
-The multipole moments can be calculated by expanding the Green's function for the Poisson equation as a series spherical harmonics, which yields
+where $q_{lm}$ are the so-called multipole moments.
+The origin of the coordinate system is taken to be the center of the computational domain, and $r$ is the distance to the origin.
+The multipole moments can be calculated by expanding the Green's function for the Poisson equation as a series of spherical harmonics.
+\begin{comment}
+, which yields
 \begin{equation}
@@
 \end{equation}
-Some simplification of Equation \ref{spherical_harmonic_expansion} can be achieved by using the addition theorem for spherical harmonics:
+\end{comment}
+After some algebraic simplification of Equation \ref{spherical_harmonic_expansion} using the addition theorem for spherical harmonics,
+\begin{comment}
 \begin{align}
@@
 \end{align}
-Here the $P_{l}^{m}$ are the associated Legendre polynomials and the $P_l$ are the Legendre polynomials.
-After some algebraic simplification, the potential outside of the mass distribution can be written in the following way:
+\end{comment}
+the potential outside of the mass distribution can be written as:
 \begin{equation}
-  \Phi(\mathbf{x}) \approx \sum_{l=0}^{l_{\text{max}}} \left[Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} + \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right].
+  \Phi(\mathbf{x}) = \sum_{l=0}^{\infty} \left[Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} + \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right].\label{multipole_potential}
 \end{equation}
-The modified multipole moments are:
+$P_l(x)$ are the Legendre polynomials and $P_{l}^{m}(x)$ are the associated Legendre polynomials. The modified multipole moments are:
 \begin{align}
@@
 \end{align}
-Our strategy for the multipole boundary conditions, then, is to pick some value $l_{\text{max}}$ that is of sufficiently high order to capture the distribution of mass on the grid, evaluate the discretized analog of the modified multipole moments for $0 \leq l \leq l_{\text{max}}$ and $1 \leq m \leq l$, and then directly compute the value of the potential on all of the boundary zones.
-This is ultimately an $\mathcal{O}(N^3)$ operation, the same order as the monopole approximation, and the wall time required to calculate the boundary conditions will depend on the chosen value of $l_{\text{max}}$.
+This approach becomes computationally feasible when we cut off the outer summation in Equation \ref{multipole_potential} at some finite value of $l_{\text{max}}$.
+If it is of sufficiently high order, we will accurately capture the distribution of mass on the grid.
+In practice we first evaluate the discretized analog of the modified multipole moments for $0 \leq l \leq l_{\text{max}}$ and $1 \leq m \leq l$, an operation that scales as $N^3$.
+We then directly compute the value of the potential on all of the $6N^2$ boundary zones.
+Since the multipole moments only need to be calculated once, the full operation scales only as $N^3$.
+The amount of time required to calculate the boundary conditions will be directly related to the chosen value of $l_{\text{max}}$, so there is a trade-off between computational expense and accuracy of the result.
+We find that for the early stages of the evolution of the binary system, $l_\text{max} \approx 6$ is sufficient.
 
-\subsubsection{James Algorithm}
+\subsubsection{Isolated Potential}
 
-Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$ on a 2-level grid.
+%Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$ on a 2-level grid.
 
@@
    \rho {\bf g} - 2 \rho {\bf \Omega \times U} 
-                - \rho {\bf \Omega \times (\Omega \ times U)}
+                - \rho {\bf \Omega \times (\Omega \times U)}
 \end{equation}
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=3.1in]{nested2}\hspace{1em}
-  \includegraphics[width=3.1in]{nested3}
-  \caption{\label{fig:2levgrav} Schematic showing the solve for $\phi$ for a 2-level grid.}
-\end{figure}
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=3.1in]{nested2}\hspace{1em}
+%  \includegraphics[width=3.1in]{nested3}
+%  \caption{\label{fig:2levgrav} Schematic showing the solve for $\phi$ for a 2-level grid.}
+%\end{figure}
 

### 2014-04-30 paper.tex
Some starting information on the James isolated boundary condition method. More to be added later.

@@
 \usepackage{mathtools}
+
+% allows Greek symbols to be bold
+\usepackage{bm}
 
@@
 \begin{equation}
-  \Phi_{\text{lmn}} = G\, \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x}_{\text{lmn}} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.
+  \Phi_{\text{lmn}} = G\, \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x}_{\text{lmn}} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.\label{direct_sum}
 \end{equation}
@@
 %Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$ on a 2-level grid.
+A clever way to evaluate Equation \ref{direct_sum} for the boundary values of the potential was described by \cite{james77}.
+In electrostatics a commonly encountered boundary-value problem is that of a positive charge distribution inside of a grounded conducting box.
+The potential vanishes on the surface of the box.
+For this to happen, a collection of negative charges must form on the outer surface of the box, to compensate the charge enclosed inside the box.
+This can be described by a surface charge density $\sigma$.
+This construction is useful in determining what the potential would be at the location of the surface when the box is removed.
+We will describe the potential outside of a density distribution that is non-zero only in some finite region as ``isolated'', and denote it by $\Phi$.
+This is the quantity we want to solve for.
+The potential in the case where the surface is grounded and conducting is called the ``homogeneous'' potential, and will be denoted by $\phi$.
+The homogeneous potential is the sum of the isolated potential and also the potential due to the surface charges $\Psi$:
+\begin{equation*}
+  \phi = \Phi + \Psi.
+\end{equation*}
+In order for $\phi$ to vanish at the surface, $\Phi$ and $\Psi$ must be of opposite sign, which makes sense since they are generated by opposite-sign charges.
+Rearranging this equation for the isolated potential yields
+\begin{equation}
+  \Phi = \phi - \Psi.
+\end{equation}
+Therefore a strategy for finding $\Phi$ is to solve the Poisson equation in the case where the boundary conditions are homogeneous and obtain $\phi$, and subtract from it the $\Psi$ that would be generated by the necessary compensating surface charges $\sigma$.
+All of this can be done for the gravitational version of Poisson's equation.
+Even though we do not commonly speak of negative gravitational charges, they are simply a mathematical tool for obtaining the correct potential.
+In general,
+\begin{equation}
+  \sigma = \frac{1}{4\pi G} {\bm{\nabla}} \phi \cdot \hat{\mathbf{n}},
+\end{equation}
+where $\hat{\mathbf{n}}$ is the unit vector perpendicular to (and pointing away from) the surface.
+On the computational domain, using an appropriate discretized representation of the ${\bm{\nabla}}$ operator, we calculate $\sigma$ at every boundary location.
+We can then find the resulting potential by summing over all of the surface charges:
+\begin{equation}
+  \Psi_{\text{l,m,n}} = \sum_{\text{l$^\prime$,m$^\prime$,n$^\prime$}} \frac{G\sigma_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}}{\left|\mathbf{x}_{\text{l,m,n}} - \mathbf{x}_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}\right|} dA_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}
+\end{equation}
+
+%In general,
+%\begin{equation}
+%  \sigma = 4\pi G
 
@@
 \bibliographystyle{apj}
-\bibliography{ws}
+\bibliography{refs}
 

### 2014-05-02 paper.tex
Some updates to the James approach, and a simplification of the method.

@@
 \end{equation}
-$P_l(x)$ are the Legendre polynomials and $P_{l}^{m}(x)$ are the associated Legendre polynomials. The modified multipole moments are:
+$P_l(x)$ are the Legendre polynomials and $P_{l}^{m}(x)$ are the associated Legendre polynomials.
+$Q_l^{(0)}$ and $Q_{lm}^{(C,S)}$ are variants of the multipole moments that involve integrals of $P_l$ and $P_l^m$, respectively, over the computational domain.
+\begin{comment}
 \begin{align}
@@
 \end{align}
+\end{comment}
 This approach becomes computationally feasible when we cut off the outer summation in Equation \ref{multipole_potential} at some finite value of $l_{\text{max}}$.
@@
 %Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$ on a 2-level grid.
-A clever way to evaluate Equation \ref{direct_sum} for the boundary values of the potential was described by \cite{james77}.
+A clever way to evaluate Equation \ref{direct_sum} exactly for the boundary values of the potential was described by \cite{james77}, and we have implemented a method inspired by that work here.
 In electrostatics a commonly encountered boundary-value problem is that of a positive charge distribution inside of a grounded conducting box.
@@
 \end{equation}
-Therefore a strategy for finding $\Phi$ is to solve the Poisson equation in the case where the boundary conditions are homogeneous and obtain $\phi$, and subtract from it the $\Psi$ that would be generated by the necessary compensating surface charges $\sigma$.
+Therefore a strategy for finding $\Phi$ is: solve the Poisson equation with homogeneous boundary conditions for $\phi$; calculate the necessary surface charges $\sigma$; find the resulting $\Psi$ on the boundary of the domain; solve the Laplace equation with inhomogeneous boundary conditions for $\Psi$ inside the domain; and then take $\Phi = \phi - \Psi$.
+
+We slightly simplify this approach by observing that, since $\phi$ is by construction zero on the boundary, the value of $\Psi$ everywhere on the boundary must be equal in magnitude and opposite in sign to $\Phi$.
+Therefore the problem of finding the boundary values for $\Phi$ is reduced to solving the Poisson equation for $\phi$, constructing the surface charges $\sigma$, and then performing the appropriate integration of $-\sigma$ to get the boundary values.
+We then proceed with the Poisson solve as normal.
+
 All of this can be done for the gravitational version of Poisson's equation.
@@
 where $\hat{\mathbf{n}}$ is the unit vector perpendicular to (and pointing away from) the surface.
-On the computational domain, using an appropriate discretized representation of the ${\bm{\nabla}}$ operator, we calculate $\sigma$ at every boundary location.
-We can then find the resulting potential by summing over all of the surface charges:
+On the computational domain, using an appropriate discretized representation of the ${\bm{\nabla}}$ operator\footnote{The gradient of $\phi$ is discontinuous across the surface; it is zero just outside the surface, and non-zero just inside the surface.
+Therefore instead of using a difference centered on the boundary edge, we take a one-sided difference using the grid points just inside the domain.
+In order for the discretization to remain second-order accurate, data from the two closest grid points are used.}, we calculate $\sigma$ at every boundary location.
+We can then find the resulting potential on the boundary by summing over all of the surface charges:
 \begin{equation}
@@
 \end{equation}
+Here (l, m, n) refer only to boundary locations. The self-contribution from a zone must be excluded in this sum, since the denominator would diverge there.
 
-%In general,
-%\begin{equation}
-%  \sigma = 4\pi G
+As described so far, the method is not especially attractive: calculation of the boundary values requires an additional Poisson solve and a sum that scales as $N^4$.
+Furthermore, the latter sum faces problems with balancing of the computational work: only processors that have data on the boundary contribute to the summation, while the others remain idle.
+This does not compare well with the $N^3$ scaling of the multipole approach.
 

### 2014-07-07 paper.tex
Initial description of the Helmholtz EOS.

@@
 
-EOS (no Coulomb corrections)
+The equation of state (EOS) for our simulations is the Helmholtz EOS \citep{timmes_swesty:2000}.
+This models an electron-positron gas of arbitrary relativity and degeneracy over a wide range of temperatures and densities.
+Thermodynamic quantities are calculated as derivatives of the Helmholtz free energy, and the values are interpolated from a table.
+The natural variables of the Helmholtz free energy are temperature and density, and calling the EOS is simplest in this form.
+However, in hydrodynamics we often have the density, and internal energy as independent variables, and we want to obtain the temperature, pressure, and other quantities.
+To do this, we employ a Newton-Raphson iteration over the temperature (given some sufficient starting guess) until we find the temperature that corresponds to the desired internal energy.
+Sometimes this process fails to converge and the iterative value approaches zero.
+In these cases we employ a ``floor'' that limits how low the temperature can go (typically $10^4$ or $10^5$ K).
+There is a choice here how to proceed: we can either assign this floor value to the temperature and let that zone be thermodynamically inconsistent (the original behavior in Castro), or we can adjust the internal energy to be thermodynamically consistent with the temperature, at the cost of violating energy conservation.
+We have found in test problems of one-dimensional shocks that the latter yields more accurate results, so we employ the latter method.
+
+Reactions?
 
@@
 where $\hat{\mathbf{n}}$ is the unit vector perpendicular to (and pointing away from) the surface.
-On the computational domain, using an appropriate discretized representation of the ${\bm{\nabla}}$ operator\footnote{The gradient of $\phi$ is discontinuous across the surface; it is zero just outside the surface, and non-zero just inside the surface.
+On the computational domain, using an appropriate discretized representation of the ${\bm{\nabla}}$ operator, we calculate $\sigma$ at every boundary location.
+This requires some care.
+In CASTRO, the boundaries are on cell edges a distance $dx / 2$ away from the outermost domain location.
+It is therefore tempting to use a central difference that uses the ghost zone on one side and the outermost zone inside the domain in the other.
+This is normally second-order accurate.
+However, by construction, the gradient of $\phi$ is discontinuous across that surface; it is zero just outside the surface, and non-zero just inside the surface.
+This discontinuity is what establishes the non-zero surface mass density, similar to how a jump in the electric field is associated with a surface charge density in electrostatics.
+Therefore it is inappropriate to use the central difference method.
+Instead, a one-sided difference is required, using only the data interior to the surface.
+For the discretization to be second-order accurate, data from the two nearest cells must be used.
+
  Therefore instead of using a difference centered on the boundary edge, we take a one-sided difference using the grid points just inside the domain.
-In order for the discretization to remain second-order accurate, data from the two closest grid points are used.}, we calculate $\sigma$ at every boundary location.
+In order for the discretization to remain second-order accurate, data from the two closest grid points are used.}
+
 We can then find the resulting potential on the boundary by summing over all of the surface charges:

### 2014-07-16 paper.tex
Some more details on the current state of the isolated BCs.

@@
 \begin{equation}
-  \sigma = \frac{1}{4\pi G} {\bm{\nabla}} \phi \cdot \hat{\mathbf{n}},
+  \sigma = \frac{1}{4\pi G} {\bm{\nabla}} \phi \cdot \hat{\mathbf{n}},\label{eq-isolated-sigma}
 \end{equation}
@@
 For the discretization to be second-order accurate, data from the two nearest cells must be used.
-
- Therefore instead of using a difference centered on the boundary edge, we take a one-sided difference using the grid points just inside the domain.
-In order for the discretization to remain second-order accurate, data from the two closest grid points are used.}
+Generally, $\left.\partial f/\partial x\right|_{\rm i} = (-3 f_{\rm i} + 4 f_{\rm i+1} - f_{\rm i+2}) / (2 \Delta x)$ is a second-order accurate forward difference to the derivative of a function $f$ at zone index i; however, we are indexing from the domain edge, which does not coincide with a cell center.
+We can use this formula on our finite volume grid if we let $\Delta x \to \Delta x / 2$ (here we assume that the index 0 corresponds to the left-most cell center):
+\begin{equation*}
+  \left.\frac{\partial \phi}{\partial x}\right|_{-1/2} = \frac{-3 \phi_{-1/2} + 4 \phi_{0} - \phi_{1/2}}{2(\Delta x / 2)} + \mathcal{O}(\Delta x^2).
+\end{equation*}
+To proceed, we note that $\phi_{-1/2} = 0$ by construction, and $\phi_{1/2} = (\phi_{1} + \phi_{0}) / 2$ to second order, so:
+\begin{equation}
+  \left.\frac{\partial \phi}{\partial x}\right|_{-1/2} = \frac{7\phi_{0} - \phi_{1}}{2\Delta x} + \mathcal{O}(\Delta x^2).
+\end{equation}
+A similar formula holds for the one-sided (backward) difference at the opposite boundary.
+Once this gradient has been calculated, we can calculate surface masses using Equation \ref{eq-isolated-sigma} (recalling to negate the result to obtain positive masses, following the above discussion).
+For example, for a mass on the lower surface of constant $x$,
+\begin{equation}
+  M_{-1/2} = -\sigma_{-1/2} \Delta A = \frac{1}{4\pi G} \frac{7\phi_{0} - \phi_{1}}{2\Delta x}\Delta y \Delta z.
+\end{equation}
+According to Gauss' law, the total mass induced on the surface should be equal to the mass enclosed inside the boundary.
+We have verified directly that the sum of the mass added on the surface approaches the total mass inside the domain at a second-order rate of convergence.
 
@@
 \begin{equation}
-  \Psi_{\text{l,m,n}} = \sum_{\text{l$^\prime$,m$^\prime$,n$^\prime$}} \frac{G\sigma_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}}{\left|\mathbf{x}_{\text{l,m,n}} - \mathbf{x}_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}\right|} dA_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}
+  \Psi_{\text{l,m,n}} = \sum_{\text{l$^\prime$,m$^\prime$,n$^\prime$}} g_{\text{l-l$^\prime$},\text{m-m$^\prime$},\text{n-n$^\prime$}} M_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}
+%= \sum_{\text{l$^\prime$,m$^\prime$,n$^\prime$}} \frac{GM_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}}{\left|\mathbf{x}_{\text{l,m,n}} - \mathbf{x}_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}\right|}
 \end{equation}
-Here (l, m, n) refer only to boundary locations. The self-contribution from a zone must be excluded in this sum, since the denominator would diverge there.
+Here (l, m, n) refer only to boundary locations, and $g_{\text{a},\text{b},\text{c}}$ is the Green's function for the Poisson equation.
+At large distances, this is simply $G / r$, where $G$ is the gravitational constant and $r$ is the distance.
+However, this analytical expression is not correct for the discretized Laplacian, and the corresponding Green's function is a series in the odd powers of $1/r$.
+We use the expression derived by \citet{burkhart:1997}, only retaining the terms up to $r^{-5}$.
+Also, this expression nominally diverges at $r = 0$, which physically corresponds to the self-contribution to the potential from a zone.
+However, this singularity is integrable: the cell does have a contribution to its own potential.
+A numerical evaluation of the integral of the Green's function over a cubic region yields $\phi(r = 0) = 2.38\, G M / \Delta x$, assuming $\Delta x = \Delta y = \Delta z$, and so our Green's function at the origin is $g_{0,0,0} = 2.38\, G / \Delta x$.
+This is similar to how other authors treat this singularity, for example \citet{hockney_eastwood}.
+However, they simply set the Green's function as $G / \Delta x$, that is with unit prefactor, without further explanation.
 

### 2014-07-31 paper.tex
Minor updates.

@@
 \subsection{Initial Models}
+\label{sec:initial_models}
 
@@
 
+In this section we describe a series of test problems that couple the hydrodynamics, gravity, and EOS modules.
+
 \subsection{Maintaining Hydrostatic Equilibrium}\label{Sec:HSE}
+
+In Section \ref{sec:initial_models} we describe the process by which we generate initial stellar models.
+While the 1D models are in hydrostatic equilibrium to within a small error, interpolation onto the 3D Cartesian grid will introduce perturbations into the solution \citep{zingale:2002}.
+Although we ensure that the initial models are generated with the same equation of state and are as well resolved as our finest grid, there will still be a hydrodynamical error associated with the fact that the rectangular grid cannot faithfully represent a spherical star.
+Additionally, the gravitational potential obtained by the multigrid solver will differ slightly from the one assumed by the initial model, and the operator splitting between the gravity and hydrodynamics should also result in small errors.
+As a result, we expect that the star will oscillate slightly about an equilibrium point, but that the amplitude of this oscillation should decrease with increasing resolution.
 

### 2014-08-04 paper.tex
Got started on the introduction.

@@
 \section{Introduction}
+
+Type Ia supernovae (SNe Ia) are currently some of the most exciting events to study in astrophysics.
+These bright, brief pulses of light in the distant universe have led to a number of important discoveries in recent years, including the discovery of the accelerated expansion of the universe \citep{perlmutter1999,riess1998}.
+However, their origin is shrouded in mystery.
+It has long been expected that these events arise from the thermonuclear explosions of white dwarfs \citep{hoyle_fowler:1960}, but the cause of these explosions is uncertain.
+In particular, it is not clear what process causes the temperatures in these white dwarfs to become hot enough for explosive burning of its constituent nuclei.
+The model favored initially by the community was the so-called single-degenerate (SD) model.
+Accretion of material from a companion star such as a red giant would cause the star to approach the Chandrasekhar mass, and in doing so the temperature and density in the center would become sufficient for thermonuclear fusion to proceed.
+A review of the observational evidence and theoretical modelling that formed this belief can be found in \citet{hillebrandtniemeyer2000}.
 

### 2014-08-15 paper.tex
More detailed info on computational resources, including XSEDE cite.
Started a discussion of the history of double degenerate modelling.

@@
 A review of the observational evidence and theoretical modelling that formed this belief can be found in \citet{hillebrandtniemeyer2000}.
+
+% History of single-degenerate model (one paragraph)
+
+% Observational evidence that disfavors SD model (one paragraph)
+
+% Early theoretical history of double-degenerate model (one paragraph)
+
+% Observational evidence that favors DD model (two-three paragraphs)
+
+The history of double degenerate theory can broadly be described as a series of three major paradigm shifts.
+The first attempts to model the results of the merger process came in the 1980s.
+\cite{nomotoiben:1985} demonstrated that if the secondary overflowed its Roche lobe and its mass accreted onto the primary near the Eddington rate, that off-center carbon ignition would occur.
+\cite{saionomoto:1985} tracked the evolution of the flame and found that it propagated quiescently into the center, converting the carbon-oxygen white dwarf into an oxygen-neon-magnesium white dwarf.
+This would then be followed by collapse into a neutron star -- a result with significantly different observational properties compared to a SN Ia.
+This scenario, termed accretion-induced collapse, would be avoided only if the accretion rate were well below the Eddington rate.
+\cite{tutukov_yungelson:1979} observed that this could happen if the mass loss from the secondary was higher than the Eddington rate and thus the accreted material formed an accretion disc, which might rain down on the primary more slowly.
+The main finding was that double degenerate systems would not obviously lead to Type Ia supernovae.
+
+The first three-dimensional simulations of double degenerate systems were performed by \citet{benz:1990}, who used the smoothed particle hydrodynamics (SPH) method to simulate the merger process.
+They found that if the lower-mass secondary was close enough to the primary to begin mass transfer on a dynamical time scale, the secondary completely disrupted and formed a hot envelope on the primary, with an accretion disk surrounding the core and envelope.
+They noted that the envelope temperature was hot enough to generate carbon fusion but neglected reactions in their simulation.
+Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007} corroborated this result.
+This validation of the prediction of \cite{tutukov_yungelson:1979} did not instill confidence in the community.
+\cite{mochkovitch_livio:1990} and \cite{livio:2000} observed that turbulent viscosity would be sufficiently large for angular momentum to be removed from the disk at a rate high enough to generate these troublesome accretion timescales.
+Based on all of this evidence, the review of \cite{hillebrandtniemeyer2000} argued that the model was only viable if the accretion-induced collapse problem could be avoided.
+Later work looked more carefully at the initial conditions for the simulation and reached the conclusion that mass transfer from the secondary would occur over many orbits, instead of disrupting completely over the span of a few orbits as originally suggested (ref).
+This may be a way to avoid the off-center carbon ignition (ref).
+
+The most recent shift in perspective on this problem started with \cite{pakmor:2010}.
+This group used the SPH method to study the merger of equal-mass ($0.9\ M_\odot$) white dwarfs and found that a hotspot was generated near the surface of the primary white dwarf.
+They argued that this region had a temperature and density sufficient to trigger a thermonuclear detonation.
+They propagated this detonation throughout the system and found that it would observationally appear as a subluminous Type Ia supernova.
+This was the first time any simulation successfully reproduced at least some characteristics of a SN Ia.
+\cite{pakmor:2011} tried a few different mass combinations and found empirically that this would hold as long as the secondary was at least 80\% as massive as the primary.
+\cite{pakmor:2012} followed this with a study of a $1.1\ M_\odot$ and $0.9\ M_\odot$ pair, with more careful attention to the initial conditions of the simulation, and found that they could reproduce the observational characteristics of normal SNe Ia.
+\cite{pakmor:2013} refined this approach by adding a thin helium shell on the primary, and found that this robustly led to a detonation of the white dwarf.
+They argued that this framework could explain most of the observed types of SNe Ia because it allowed for both carbon-oxygen and helium white dwarfs to be the secondary, allowing for a wide range of possible outcomes.
+
+% Discuss some other recent work (Guillochon et al. 2010, van Kerkwijk et al. 2012)
+
+% Motivation: a series of papers to verify or disprove the prompt/violent merger hypothesis using a grid-based code
 
@@
 DE-AC05-00OR22725.
+Project AST006 supported use of the ORNL/Titan resource.
+This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number ACI-1053575.
+Project AST100037 supported use of the resources NICS/Kraken and NICS/Darter.
 

### 2014-08-26 paper.tex
Added quite a bit of DD theory to the introduction section.

@@
 The first three-dimensional simulations of double degenerate systems were performed by \citet{benz:1990}, who used the smoothed particle hydrodynamics (SPH) method to simulate the merger process.
-They found that if the lower-mass secondary was close enough to the primary to begin mass transfer on a dynamical time scale, the secondary completely disrupted and formed a hot envelope on the primary, with an accretion disk surrounding the core and envelope.
+They found that if the lower-mass secondary was close enough to the primary to begin mass transfer on a dynamical time scale, the secondary completely disrupted and formed a hot envelope on the primary, with a centrifugally-supported accretion disk surrounding the core and envelope.
 They noted that the envelope temperature was hot enough to generate carbon fusion but neglected reactions in their simulation.
-Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007} corroborated this result.
+Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009} corroborated this result.
 This validation of the prediction of \cite{tutukov_yungelson:1979} did not instill confidence in the community.
@@
 Based on all of this evidence, the review of \cite{hillebrandtniemeyer2000} argued that the model was only viable if the accretion-induced collapse problem could be avoided.
-Later work looked more carefully at the initial conditions for the simulation and reached the conclusion that mass transfer from the secondary would occur over many orbits, instead of disrupting completely over the span of a few orbits as originally suggested (ref).
-This may be a way to avoid the off-center carbon ignition (ref).
+Later work by \cite{shen:2012} and \cite{schwab:2012} used a more detailed treatment of the viscous transport in the outer regions of the remnant and found that while the centrifugally supported envelope would be converted into hot envelope material on a viscous timescale, their simulations still led to off-center carbon burning.
+\cite{vankerkwijk:2010} argued that equal-mass mergers would lead to the conditions necessary for carbon detonation in the center of the merged object, but \cite{shen:2012} questioned this for similar reasons related to how viscous transport would convert rotational motion into pressure support.
+\cite{zhu:2013} followed this with an expanded parameter space study and argued that many of their carbon-oxygen systems had the potential to detonate.
+The study of the long-term evolution of the remnants is thus still an open subject of research.
 
-The most recent shift in perspective on this problem started with \cite{pakmor:2010}.
-This group used the SPH method to study the merger of equal-mass ($0.9\ M_\odot$) white dwarfs and found that a hotspot was generated near the surface of the primary white dwarf.
+The most recent shift in perspective on this problem started with two series of papers appearing at roughly the same time.
+One began with \cite{pakmor:2010} and \cite{pakmor:2011}.
+This group used the SPH method to study the merger of equal-mass ($0.9\ M_\odot$) carbon-oxygen white dwarfs and found that a hotspot was generated near the surface of the primary white dwarf.
 They argued that this region had a temperature and density sufficient to trigger a thermonuclear detonation.
@@
 \cite{pakmor:2011} tried a few different mass combinations and found empirically that this would hold as long as the secondary was at least 80\% as massive as the primary.
-\cite{pakmor:2012} followed this with a study of a $1.1\ M_\odot$ and $0.9\ M_\odot$ pair, with more careful attention to the initial conditions of the simulation, and found that they could reproduce the observational characteristics of normal SNe Ia.
-\cite{pakmor:2013} refined this approach by adding a thin helium shell on the primary, and found that this robustly led to a detonation of the white dwarf.
-They argued that this framework could explain most of the observed types of SNe Ia because it allowed for both carbon-oxygen and helium white dwarfs to be the secondary, allowing for a wide range of possible outcomes.
+These events, where the merger process resulted in the detonation of the system during the merger coalescence -- avoiding the much longer time-scale evolution -- were termed ``violent'' mergers.
 
-% Discuss some other recent work (Guillochon et al. 2010, van Kerkwijk et al. 2012)
+Around the same time, however, \cite{guillochon:2010} and \cite{dan:2011} came to significantly different conclusions.
+They pointed out that the previously mentioned simulations generally all shared a significant drawback, which was that their initial conditions were not carefully constructed.
+\cite{motl:2002}, \cite{dsouza:2006}, and \cite{motl:2007} (the first three-dimensional grid-based simulations of white dwarf mergers) pioneered the study of looking at the long-term dynamical evolution of merger scenarios after constructing exact equilibrium initial conditions.
+In contrast to the other studies, they found that paying close attention to the accuracy of the initial conditions of the simulation has important effects.
+In particular, earlier work placed the stars too close together and ignored the effects of tidal forces on changing the shape of the secondary, leading to the merger happening artificially too quickly.
+When the initial conditions are constructed in exact equilibrium, the system can be stable for tens of orbital periods, substantially changing the character of the mass transfer phase.
+However, one limitation of this work is that the authors used a polytropic equation of state and thus could not consider nuclear reactions.
+\cite{guillochon:2010} and \cite{dan:2011} improved on this using a realistic equation of state, a nuclear reaction network, and a similar approach to the equilibrium initial conditions, and found substantial agreement with the idea that mass transfer occurs in a stable manner over tens of orbital periods.
+They also found that, assuming the material accreted onto the surface of the primary was primarily helium, explosive surface detonations would occur as a result of accretion stream instabilities during the mass transfer phase prior to the full merger.
+This could trigger a double-detonation explosion and thus perhaps a SN Ia.
+
+The most recent developments give some areas of agreement and some of remaining tension.
+\cite{pakmor:2012} performed a merger scenario with better treatment of the initial conditions and also found that the merger process happened over more than ten orbits.
+Nevertheless, they still found that a carbon-oxygen detonation would occur, in line with their earlier results.
+On the other hand, \cite{dan:2012} and \cite{dan:2014} performed a large sweep of the parameter space for merger pairs and found that pure carbon-oxygen systems would generally not lead to detonations and violent mergers.
+They did find that for systems containing helium, many of them would detonate and potentially lead to SNe Ia, either through the aforementioned instabilities in the accretion stream, or during the contact phase, similar to the violent carbon-oxygen mergers.
+\cite{pakmor:2013} added a thin helium shell on their primary white dwarf, and found that this robustly led to a detonation of the white dwarf.
+Thus the preliminary consensus seems to be that systems containing helium could robustly lead to events resembling SNe Ia, while there is still debate about larger-mass systems containing only carbon and oxygen.
+
+% Still need Raskin/Moll/et al. papers
 

### 2014-08-26 paper.tex
Added a couple papers by Raskin/Moll/Kasen/etc. and thus concluded the first draft of the DD theory history section.

@@
 They noted that the envelope temperature was hot enough to generate carbon fusion but neglected reactions in their simulation.
-Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009} corroborated this result.
+Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009,raskin:2012} corroborated this result.
 This validation of the prediction of \cite{tutukov_yungelson:1979} did not instill confidence in the community.
@@
 
-The most recent developments give some areas of agreement and some of remaining tension.
-\cite{pakmor:2012} performed a merger scenario with better treatment of the initial conditions and also found that the merger process happened over more than ten orbits.
+The most recent developments give some areas of agreement and some of remaining uncertainty.
+\cite{pakmor:2012} performed a merger scenario with a $1.1\ M_\odot$ and $0.9\ M_\odot$ setup, with better treatment of the initial conditions, and also found that the merger process happened over more than ten orbits.
 Nevertheless, they still found that a carbon-oxygen detonation would occur, in line with their earlier results.
-On the other hand, \cite{dan:2012} and \cite{dan:2014} performed a large sweep of the parameter space for merger pairs and found that pure carbon-oxygen systems would generally not lead to detonations and violent mergers.
+\cite{moll:2014} was also able to find a detonation in a similarly massive system.
+\cite{dan:2012} and \cite{dan:2014} performed a large sweep of the parameter space for merger pairs and found that pure carbon-oxygen systems would generally not lead to detonations and violent mergers except for the most massive systems.
 They did find that for systems containing helium, many of them would detonate and potentially lead to SNe Ia, either through the aforementioned instabilities in the accretion stream, or during the contact phase, similar to the violent carbon-oxygen mergers.
 \cite{pakmor:2013} added a thin helium shell on their primary white dwarf, and found that this robustly led to a detonation of the white dwarf.
-Thus the preliminary consensus seems to be that systems containing helium could robustly lead to events resembling SNe Ia, while there is still debate about larger-mass systems containing only carbon and oxygen.
+Thus the preliminary consensus seems to be that systems containing helium could robustly lead to events resembling SNe Ia, as well as very massive carbon-oxygen binaries.
 

### 2014-09-12 paper.tex
Finished first draft of introduction.

@@
 In particular, it is not clear what process causes the temperatures in these white dwarfs to become hot enough for explosive burning of its constituent nuclei.
-The model favored initially by the community was the so-called single-degenerate (SD) model.
+The model favored initially by the community was the so-called single-degenerate (SD) model \citep{whelan_iben:1973}.
 Accretion of material from a companion star such as a red giant would cause the star to approach the Chandrasekhar mass, and in doing so the temperature and density in the center would become sufficient for thermonuclear fusion to proceed.
-A review of the observational evidence and theoretical modelling that formed this belief can be found in \citet{hillebrandtniemeyer2000}.
+However, in recent years many alternative progenitor models have been discussed.
+A leading candidate for explaining some or most of these explosions is the double-degenerate model, in which two white dwarfs merge and the merged object reaches the conditions necessary for a thermonuclear ignition \citep{ibentutukov:1984,webbink:1984}.
+Another is the double detonation scenario, where accretion of material onto a sub-Chandrasekhar white dwarf would lead to a detonation inside the accreted envelope, and this would send a compressional wave into the core of the star that would trigger a secondary detonation.
+A recent review of the progenitor models can be found in \citet{hillebrandt:2013}.
 
-% History of single-degenerate model (one paragraph)
-
-% Observational evidence that disfavors SD model (one paragraph)
-
-% Early theoretical history of double-degenerate model (one paragraph)
-
-% Observational evidence that favors DD model (two-three paragraphs)
+There are several observational reasons why double-degenerate systems are a promising progenitor system for at least a substantial fraction of normal SNe Ia.
+No conclusive evidence exists for a surviving companion star of a SN Ia; this is naturally explained by the DD model because both WDs are destroyed in the merger process.
+Similarly, pre-explosion images of the SN Ia systems have never clearly turned up a companion star, and in some cases a large fraction of the SD parameter space is excluded.
+Additionally, not enough progenitor systems are seen for the SD case to match the observed local SN Ia rate, whereas the number of white dwarf binaries may be sufficient to account for this rate.
+Finally, the DD model can naturally explain the fact that many SNe Ia are observed to occur at very long delay times after the stars were formed, since the progenitor systems only become active once both stars have evolved off the main sequence.
+A thorough review of the observational evidence about SNe Ia and further discussion of these ideas can be found in \cite{maoz:2014}.
 
@@
 \cite{pakmor:2013} added a thin helium shell on their primary white dwarf, and found that this robustly led to a detonation of the white dwarf.
-Thus the preliminary consensus seems to be that systems containing helium could robustly lead to events resembling SNe Ia, as well as very massive carbon-oxygen binaries.
+Thus at least a preliminary agreement may be that systems containing helium could robustly lead to events resembling SNe Ia, as well as very massive carbon-oxygen binaries.
 
-% Still need Raskin/Moll/et al. papers
-
-% Motivation: a series of papers to verify or disprove the prompt/violent merger hypothesis using a grid-based code
+This is the first in a series of papers designed to address these outstanding theoretical issues for white dwarf mergers.
+This work will discuss the verification of our hydrodynamics code for simulating these events.
+Later efforts will look at the initial conditions of the system, the robustness with which a hotspot is found from which a detonation could occur, and the importance of the initial white dwarf models, which should be more sophisticated than simple carbon-oxygen mixtures and in principle should use results from modern stellar evolution calculations.
+Section \ref{sec:Numerical Methodology} describes our code and why it can provide useful results compared to other methodologies used for this problem.
+Section \ref{sec:Tests} discusses a few test problems that we use to demonstrate that our code accurately solves the equations of fluid dynamics.
+Section \ref{sec:Performance} demonstrates that the software scales well for supercomputer applications.
+Finally, Section \ref{sec:Conclusions and Discussion} recaps what we have shown and highlights some of the future work we plan to do.
 
@@
 %==========================================================================
-\section{Test Problems}\label{Sec:Tests}
+\section{Test Problems}\label{sec:Tests}
 
@@
 %==========================================================================
-\section{Performance}\label{Sec:Performance}
+\section{Performance}\label{sec:Performance}
 
@@
 %==========================================================================
-\section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}
+\section{Conclusions and Discussion}\label{sec:Conclusions and Discussion}
 

### 2014-09-27 paper.tex
Added a start for the bulk of what we need for the hydrodynamics subsection.

@@
 
-We use the unsplit PPM solver in Castro to advance the hydrodynamics
-system in time.  A number of changes were made to the solver, which are 
-detailed in the Appendix.  These changes bring the algorithm more in 
-line with that of \cite{ppm}.
+To study the white dwarf merger problem, we use the grid-based hydrodynamics code CASTRO \citep{castro}.
+CASTRO solves the Euler equations, along with the inclusion of optional modules for gravity, nuclear reactions and thermodynamics.
+We direct the reader to the original code paper for a description of CASTRO's approach to solving the equations of hydrodynamics.
+In this work, we report only on the changes we have made to the code since its original release, for the purpose of approaching this problem.
+CASTRO is based on the BoxLib adaptive-mesh refinement (AMR) framework (ref), which represents fluid data on a mesh where regions of interest have higher spatial resolution.
+CASTRO is highly parallel and is designed for large-scale use on modern supercomputers; see Section \ref{sec:Performance} for information on how CASTRO performs for our problem.
 
-The boundary conditions on the hyperbolic system are simply
-zero-gradient.  Using AMR, we make the coarse grid very large, so as
-to place the boundaries far from the region of interest.  We further
-make the restriction that only level-0 (coarse) grids can touch the
-domain boundary.
+We use the unsplit piecewise-parabolic method (PPM) solver in CASTRO to advance the hydrodynamics system in time \citep{ppmunsplit}.
+A number of changes were made to the solver, which are detailed in the Appendix.
+These changes bring the algorithm more in line with that of \cite{ppm}.
+CASTRO as originally released featured a slightly modified version of the higher resolution limiters of \cite{colella_sekora:2008}, which can be accessed in the code using \texttt{castro.ppm\_type = 2}.
+The advantage of this limiter is that it preserves physical extrema rather than clipping them off as in the original approach of \cite{ppm}.
+However, we found these limiters to be unsatisfactory for our problem.
+There are many regions in our problem with large density gradients (such as the interface between the star's atmosphere and the ambient gas outside of it) and in these regions the algorithm can yield negative densities.
+This often results from the limiters interpreting these gradients as being true minima.
+As a result, we use the original limiter, which is strictly monotonicity preserving in the parabolic profiles it generates; this is activated with \texttt{castro.ppm\_type = 1}.
 
-AMR
+A related issue that required a code improvement was that in cases of large density gradients such as the edge of a star, it is possible to generate negative densities in zones even with the more strongly limited PPM.
+This can occur if a region of large density is moving away from an ambient zone at relatively large speeds; then the net density flux in the ambient zones can be large enough to unphysically drag the density below zero.
+In practice, this will occur at the trailing edge of a star that is moving across a grid.
+In such a situation, there are two main approaches one could take: either explicitly introduce a positivity-guaranteeing diffusive flux, or reset the characteristics of the affected zone.
+We choose the latter approach.
+Even though it is non-conservative, it preserves a characteristic we value, which is to keep the edge of the stars relatively sharp, as they physically should be.
+Since the mass of the affected zones is typically already fairly low, this should not seriously affect the energy conservation properties of our simulation.
+Similarly, it is also possible in this physical situation for very large velocities to be generated in the material behind the star.
+The momentum transferred to the ambient zone is close to the momentum of the stellar material, but this is of much higher density and so the velocity of the ambient material must become very large to compensate.
+While this may be physically viable in certain circumstances, it poses a problem for our code, which uses subcycling in time (see below).
+The timestep is fixed at the beginning of the set of subcycled timesteps, and if in the middle of the cycle a velocity is generated that is so large that it violates the CFL criterion for this timestep, we cannot guarantee the stability of the resulting solution since we cannot alter the timestep.
+(Nor would we want to, because in some cases because the velocity becomes spuriously large and would drag down the global timestep.) Therefore we also insert a check at the end of every hydrodynamics timestep that resets the velocity and/or sound-speed (indirectly, using the temperature) of any zone that violates the CFL criterion given the timestep we chose at the beginning of the cycle.
+In practice, we find that this is mainly necessary only for simulations in the inertial reference frame of the white dwarf merger, which we do not normally do for the reasons stated in Section \ref{Sec:Kepler}.
 
-rotation
+The boundary conditions on the hyperbolic system are simply zero-gradient zones that allow material to flow directly out of the domain. Using AMR, we make the coarse grid very large, so as
+to place the boundaries far from the region of interest. This ensures that any boundary effects do not pollute the inner region where the stars will eventually make contact.  We further
+make the restriction that only level-0 (coarse) grids can touch the domain boundary.
+
+CASTRO's approach to adaptive mesh refinement, based on its underlying BoxLib framework, is to refine zones based on certain user-specified criteria that tag regions of interest for higher spatial resolution.
+Data is represented on one of a number of AMR levels, where each level corresponds to a set of zones at the same resolution, which covers a subset of the domain covered by the level immediately below it.
+We typically call the level 0 grid the \textit{coarse} grid, which has the lowest spatial resolution.
+Each finer, higher-level grid has a higher resolution than the grid below it by some integer factor $N$, which is restricted to be $N = 2\ \text{or}\ 4$ in the code.
+The zones are strictly contained within the rectangular extent of the underlying coarser zones (the code is not restricted to representing only Cartesian geometries, but we use a Cartesian mesh with uniform spacing in each dimension for the present study).
+For the time evolution of the AMR system we use subcycling, where each AMR level is advanced at a different timestep and a correction step is applied at the end to synchronize the various levels.
+Generally the number of subcycled timesteps is equal to the jump in refinement between levels, so for example on a grid with three levels and two jumps of four in refinement, the level 2 zones will have 16 times higher spatial resolution than the coarse grid and there will be 16 level 2 timesteps per level 0 timestep.
+
+For the evolution of binary systems, it is most natural to evolve the two stars in a frame that is co-rotating at the same period as the orbital period.
+CASTRO has the ability to evolve systems in a rotating reference frame, as discussed in the original code paper.
+Source terms corresponding to the Coriolis and centrifugal force terms are added to the momentum and energy equations.
+In this frame, the stars essentially remain stationary in their original positions due to the centrifugal force supporting against the gravitational attraction, and will remain this way as long as significant mass transfer does not occur.
+\cite{swesty00} demonstrated (in the context of neutron star mergers) that conservation of angular momentum is much easier to obtain in the rotating reference frame than in an inertial frame in which stars advect large amounts of material around the domain.
+We wish to emphasize that although it is commonly stated in the literature that grid-based codes poorly conserve angular momentum, this is not generally true.
+When the resolution is sufficiently high and a rotating reference frame is employed, excellent conservation properties can result, as demonstrated in Section \ref{Sec:Kepler}.
+We note that as the stars begin to coalesce, the rotating reference frame will no longer provide a good approximation to the spatial motion of the stars and then they will begin to significantly move around the domain.
+This is not necessarily problematic because the most important feature of the rotating frame is that it helps ensure that the initial coalescence is not the result of spurious numerical loss of angular momentum.
+When significant mass transfer sets in and evolution proceeds on a dynamical timescale, the conservation properties may be slightly worse but angular momentum conservation is also less important.
 

### 2014-11-10 paper.tex
Remove section on James algorithm since it's not going to appear in this paper.

@@
 
+\subsubsection{Boundary Conditions}
+
 Analytic solutions to the Poisson equation customarily assume that the potential vanishes at large distances from the region of non-zero density.
@@
 If there are $N$ zones per spatial dimension, then there are $6 N^2$ boundary zones, and each boundary zone requires a sum over $N^3$ zones, so the direct computation of the boundary conditions scales as $N^5$.
-This method is expensive enough that it is not used for hydrodynamics simulations (though it is useful for comparison to the approximate solutions described below).
+This method is expensive enough that it is not used for hydrodynamics simulations (though it is useful for comparison to approximate solutions).
 
@@
 This results in a significant drift of the center of the mass of the system over time.
-We describe here two new features we have added to CASTRO to address this issue.
-
-\subsubsection{Multipole Expansion}
 
@@
 We find that for the early stages of the evolution of the binary system, $l_\text{max} \approx 6$ is sufficient.
-
-\subsubsection{Isolated Potential}
-
-%Figure~\ref{fig:2levgrav} shows the sequence of updates for $\phi$ on a 2-level grid.
-A clever way to evaluate Equation \ref{direct_sum} exactly for the boundary values of the potential was described by \cite{james77}, and we have implemented a method inspired by that work here.
-In electrostatics a commonly encountered boundary-value problem is that of a positive charge distribution inside of a grounded conducting box.
-The potential vanishes on the surface of the box.
-For this to happen, a collection of negative charges must form on the outer surface of the box, to compensate the charge enclosed inside the box.
-This can be described by a surface charge density $\sigma$.
-This construction is useful in determining what the potential would be at the location of the surface when the box is removed.
-We will describe the potential outside of a density distribution that is non-zero only in some finite region as ``isolated'', and denote it by $\Phi$.
-This is the quantity we want to solve for.
-The potential in the case where the surface is grounded and conducting is called the ``homogeneous'' potential, and will be denoted by $\phi$.
-The homogeneous potential is the sum of the isolated potential and also the potential due to the surface charges $\Psi$:
-\begin{equation*}
-  \phi = \Phi + \Psi.
-\end{equation*}
-In order for $\phi$ to vanish at the surface, $\Phi$ and $\Psi$ must be of opposite sign, which makes sense since they are generated by opposite-sign charges.
-Rearranging this equation for the isolated potential yields
-\begin{equation}
-  \Phi = \phi - \Psi.
-\end{equation}
-Therefore a strategy for finding $\Phi$ is: solve the Poisson equation with homogeneous boundary conditions for $\phi$; calculate the necessary surface charges $\sigma$; find the resulting $\Psi$ on the boundary of the domain; solve the Laplace equation with inhomogeneous boundary conditions for $\Psi$ inside the domain; and then take $\Phi = \phi - \Psi$.
-
-We slightly simplify this approach by observing that, since $\phi$ is by construction zero on the boundary, the value of $\Psi$ everywhere on the boundary must be equal in magnitude and opposite in sign to $\Phi$.
-Therefore the problem of finding the boundary values for $\Phi$ is reduced to solving the Poisson equation for $\phi$, constructing the surface charges $\sigma$, and then performing the appropriate integration of $-\sigma$ to get the boundary values.
-We then proceed with the Poisson solve as normal.
-
-All of this can be done for the gravitational version of Poisson's equation.
-Even though we do not commonly speak of negative gravitational charges, they are simply a mathematical tool for obtaining the correct potential.
-In general,
-\begin{equation}
-  \sigma = \frac{1}{4\pi G} {\bm{\nabla}} \phi \cdot \hat{\mathbf{n}},\label{eq-isolated-sigma}
-\end{equation}
-where $\hat{\mathbf{n}}$ is the unit vector perpendicular to (and pointing away from) the surface.
-On the computational domain, using an appropriate discretized representation of the ${\bm{\nabla}}$ operator, we calculate $\sigma$ at every boundary location.
-This requires some care.
-In CASTRO, the boundaries are on cell edges a distance $dx / 2$ away from the outermost domain location.
-It is therefore tempting to use a central difference that uses the ghost zone on one side and the outermost zone inside the domain in the other.
-This is normally second-order accurate.
-However, by construction, the gradient of $\phi$ is discontinuous across that surface; it is zero just outside the surface, and non-zero just inside the surface.
-This discontinuity is what establishes the non-zero surface mass density, similar to how a jump in the electric field is associated with a surface charge density in electrostatics.
-Therefore it is inappropriate to use the central difference method.
-Instead, a one-sided difference is required, using only the data interior to the surface.
-For the discretization to be second-order accurate, data from the two nearest cells must be used.
-Generally, $\left.\partial f/\partial x\right|_{\rm i} = (-3 f_{\rm i} + 4 f_{\rm i+1} - f_{\rm i+2}) / (2 \Delta x)$ is a second-order accurate forward difference to the derivative of a function $f$ at zone index i; however, we are indexing from the domain edge, which does not coincide with a cell center.
-We can use this formula on our finite volume grid if we let $\Delta x \to \Delta x / 2$ (here we assume that the index 0 corresponds to the left-most cell center):
-\begin{equation*}
-  \left.\frac{\partial \phi}{\partial x}\right|_{-1/2} = \frac{-3 \phi_{-1/2} + 4 \phi_{0} - \phi_{1/2}}{2(\Delta x / 2)} + \mathcal{O}(\Delta x^2).
-\end{equation*}
-To proceed, we note that $\phi_{-1/2} = 0$ by construction, and $\phi_{1/2} = (\phi_{1} + \phi_{0}) / 2$ to second order, so:
-\begin{equation}
-  \left.\frac{\partial \phi}{\partial x}\right|_{-1/2} = \frac{7\phi_{0} - \phi_{1}}{2\Delta x} + \mathcal{O}(\Delta x^2).
-\end{equation}
-A similar formula holds for the one-sided (backward) difference at the opposite boundary.
-Once this gradient has been calculated, we can calculate surface masses using Equation \ref{eq-isolated-sigma} (recalling to negate the result to obtain positive masses, following the above discussion).
-For example, for a mass on the lower surface of constant $x$,
-\begin{equation}
-  M_{-1/2} = -\sigma_{-1/2} \Delta A = \frac{1}{4\pi G} \frac{7\phi_{0} - \phi_{1}}{2\Delta x}\Delta y \Delta z.
-\end{equation}
-According to Gauss' law, the total mass induced on the surface should be equal to the mass enclosed inside the boundary.
-We have verified directly that the sum of the mass added on the surface approaches the total mass inside the domain at a second-order rate of convergence.
-
-We can then find the resulting potential on the boundary by summing over all of the surface charges:
-\begin{equation}
-  \Psi_{\text{l,m,n}} = \sum_{\text{l$^\prime$,m$^\prime$,n$^\prime$}} g_{\text{l-l$^\prime$},\text{m-m$^\prime$},\text{n-n$^\prime$}} M_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}
-%= \sum_{\text{l$^\prime$,m$^\prime$,n$^\prime$}} \frac{GM_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}}{\left|\mathbf{x}_{\text{l,m,n}} - \mathbf{x}_{\text{l$^\prime$,m$^\prime$,n$^\prime$}}\right|}
-\end{equation}
-Here (l, m, n) refer only to boundary locations, and $g_{\text{a},\text{b},\text{c}}$ is the Green's function for the Poisson equation.
-At large distances, this is simply $G / r$, where $G$ is the gravitational constant and $r$ is the distance.
-However, this analytical expression is not correct for the discretized Laplacian, and the corresponding Green's function is a series in the odd powers of $1/r$.
-We use the expression derived by \citet{burkhart:1997}, only retaining the terms up to $r^{-5}$.
-Also, this expression nominally diverges at $r = 0$, which physically corresponds to the self-contribution to the potential from a zone.
-However, this singularity is integrable: the cell does have a contribution to its own potential.
-A numerical evaluation of the integral of the Green's function over a cubic region yields $\phi(r = 0) = 2.38\, G M / \Delta x$, assuming $\Delta x = \Delta y = \Delta z$, and so our Green's function at the origin is $g_{0,0,0} = 2.38\, G / \Delta x$.
-This is similar to how other authors treat this singularity, for example \citet{hockney_eastwood}.
-However, they simply set the Green's function as $G / \Delta x$, that is with unit prefactor, without further explanation.
-
-As described so far, the method is not especially attractive: calculation of the boundary values requires an additional Poisson solve and a sum that scales as $N^4$.
-Furthermore, the latter sum faces problems with balancing of the computational work: only processors that have data on the boundary contribute to the summation, while the others remain idle.
-This does not compare well with the $N^3$ scaling of the multipole approach.
 
@@
 Project AST100037 supported use of the resources NICS/Kraken and NICS/Darter.
+
+%TODO: add blue waters citation
 

### 2014-12-20 papers/wdmerger_I/freefall/freefall_solution.tex
Move all wdmerger_I paper files into the papers/wdmerger_I subdirectory.

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[margin=1.0in]{geometry}
\usepackage{parskip}

\begin{document}

  \begin{center}
    \textbf{Solution of the Two-Body Free Fall Problem}\\
    \textbf{November 5, 2012}
  \end{center}

  Suppose two particles are initially at rest, at locations $x_A$ and $x_B$. The particles have masses $M_A$ and $M_B$, respectively. Here I provide the analytical solution for the free fall. The center of mass of this system stays stationaary in this problem, since there are no external forces. Therefore, if I denote by $r(t)$ the present separation of the two particles, the equation of motion is
  \begin{equation}
    \ddot{r}(t) = - \frac{GM}{r^2},
  \end{equation}
  where $M$ is the total mass of the system $M = M_A + M_B$ and the initial condition is $r_0 = x_{A0} + x_{B0}$.

  This can be solved by recognizing that
  \begin{equation}
    \frac{d^2 r}{dt^2} = \frac{dv}{dt} = v(r) \frac{dv(r)}{dr}
  \end{equation}
  where $v(r)$ is the instantaneous velocity at particle separation $r(t)$. The equation of motion is then separable:
  \begin{align}
    v\, dv &= -GM \frac{dr}{r^2} \\
    \Rightarrow \frac{1}{2}v(r)^2 &= -GM\int_{r_0}^r \frac{dr}{r^2} \\
                                  &= GM\left[ \frac{1}{r} - \frac{1}{r_0} \right] \\
    \Rightarrow v(r) &= \sqrt{\frac{2GM}{r_0} \left[\frac{r_0}{r} - 1\right]}.
  \end{align}
  Since $v(r) = dr / dt$, this equation is also separable:
  \begin{align}
    t(r) = \sqrt{\frac{r_0}{2GM}}\int_{r_0}^r \frac{dr}{\sqrt{\frac{r_0}{r} - 1}}.
  \end{align}
  Here I use the variable substitution
  \begin{equation}
    u(r) = \frac{r}{r_0}; \quad u(r_0) = 1; \quad du = \frac{1}{r_0} dr.
  \end{equation}
  Then the integral takes the form
  \begin{align}
    t(r) &= \sqrt{\frac{r_0^3}{2GM}}\int_{1}^{r/r_0} \frac{du}{\sqrt{\frac{1}{u}-1}}.
  \end{align}
  Here a trigonometric substitution makes sense:
  \begin{equation}
    u(\theta) = \text{cos}^2(\theta); \quad \theta(u = 1) = 0; \quad du = 2\,\text{cos}(\theta)\, \text{sin}(\theta)\, d\theta
  \end{equation}
  Then, using the trigonometric identity
  \[
    \frac{1}{\text{cos}^2(\theta)} - 1 = \text{tan}^2(\theta),
  \]
  the integral gets transformed into the final form
  \begin{align}
    t(r) &= 2\sqrt{\frac{r_0^3}{2GM}} \int_{0}^{\theta(r)} d\theta\, \frac{\text{sin}(\theta)\, \text{cos}(\theta)}{\text{tan}(\theta)} \\
         &= 2\sqrt{\frac{r_0^3}{2GM}} \int_{0}^{\theta(r)} d\theta\, \text{cos}^2(\theta) \\
         &= \sqrt{\frac{r_0^3}{2GM}} \bigg[ \theta + \text{sin}(\theta)\, \text{cos}(\theta) \bigg]_0^{\theta(r)} \\
         &= \sqrt{\frac{r_0^3}{2GM}} \bigg[ \theta + \text{cos}(\theta)\, \sqrt{1 - \text{cos}^2(\theta)} \bigg]_0^{\theta(r)} \\
         &= \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right].
  \end{align}

\end{document}

### 2014-12-21 papers/wdmerger_I/paper.tex
Move all wdmerger_I paper files into the papers/wdmerger_I subdirectory.
Addition of a numerical test section on Galilean invariance, as well as a proper introduction to the numerical test section itself.

--- 2014-11-10 paper.tex
@@
 (Nor would we want to, because in some cases because the velocity becomes spuriously large and would drag down the global timestep.) Therefore we also insert a check at the end of every hydrodynamics timestep that resets the velocity and/or sound-speed (indirectly, using the temperature) of any zone that violates the CFL criterion given the timestep we chose at the beginning of the cycle.
-In practice, we find that this is mainly necessary only for simulations in the inertial reference frame of the white dwarf merger, which we do not normally do for the reasons stated in Section \ref{Sec:Kepler}.
+In practice, we find that this is mainly necessary only for simulations in the inertial reference frame of the white dwarf merger, which we do not normally do for the reasons stated in Section \ref{sec:Kepler}.
 
@@
 We wish to emphasize that although it is commonly stated in the literature that grid-based codes poorly conserve angular momentum, this is not generally true.
-When the resolution is sufficiently high and a rotating reference frame is employed, excellent conservation properties can result, as demonstrated in Section \ref{Sec:Kepler}.
+When the resolution is sufficiently high and a rotating reference frame is employed, excellent conservation properties can result, as demonstrated in Section \ref{sec:Kepler}.
 We note that as the stars begin to coalesce, the rotating reference frame will no longer provide a good approximation to the spatial motion of the stars and then they will begin to significantly move around the domain.
@@
 \subsection{Gravity}
+\label{sec:gravity}
 
@@
 %==========================================================================
-\section{Test Problems}\label{sec:Tests}
+\section{Numerical Test Problems}\label{sec:Tests}
 
-In this section we describe a series of test problems that couple the hydrodynamics, gravity, and EOS modules.
+Merger simulations face a number of numerical difficulties that are not present in single-degenerate Type Ia and core-collapse supernova simulations.
+In Section \ref{sec:gravity}, we discussed how the lack of spherical symmetry necessitates a careful look at the gravity solver.
+There are also hydrodynamical issues: the merger process will result in substantial motion of stellar material across the grid.
+This bulk motion presents an opportunity for advection errors to build up, and is only partially mitigated by evolving the white dwarfs in a co-rotating frame.
+It is therefore important to be be aware of the behavior of the code in such circumstances.
+The behavior of CASTRO for many standard hydrodynamics test problems was detailed in the original code paper \citep{castro}, and in the interest of brevity we will not repeat them here.
+Instead, we focus on a subset of problems that highlight the special difficulties introduced in merger simulations.
+These problems couple the hydrodynamics, gravity and equation of state modules.
+We observe that while in most non-trivial three-dimensional problems this creates a complexity that makes it impossible to determine exact analytical solutions, it is straightforward to devise problems for which certain global properties should obey simple, expected behaviors.
+Where possible, these should be quantified and a convergence study performed.
+We recommend that all simulation codes should check their performance on such problems to ensure their reliability when coupling the various physics modules necessary to build a complete and realistic simulation.
 
-\subsection{Maintaining Hydrostatic Equilibrium}\label{Sec:HSE}
+\subsection{Maintaining Hydrostatic Equilibrium}\label{sec:HSE}
 
@@
 
-\subsection{Gravitational Free Fall}\label{Sec:Gravitational Free Fall}
+This problem was studied in the first CASTRO paper, but is worth revisiting here.
+A single star explosion simulation may only last a couple of seconds, and the CASTRO paper studied the behavior of the star after one second of evolution.
+However, the dynamical timescale of a typical carbon-oxygen white dwarf is on the order of 1--10 seconds.
+Additionally, a binary orbit is typically on the order of 10--100 seconds when a merger simulation starts, and with equilibrium initial conditions the system may survive for tens of orbits before the secondary is disrupted.
+When this does happen, we want to be confident that it was because of the dynamics of the merger process and not because of an instability in an individual star.
+Our goal here is thus to populate a single star onto our three-dimensional coordinate grid and evolve it for a period of time long enough to assess whether the star is truly stable, and to probe how the size of deviation from equilibrium is affected by grid resolution.
+
+\subsection{Gravitational Free Fall}\label{sec:Gravitational Free Fall}
 
@@
 We determine the initial separation to be consistent with the
-simulation performed in Section \ref{Sec:Kepler}; that is, we select
+simulation performed in Section \ref{sec:Kepler}; that is, we select
 an initial orbital period and use Kepler's third law to calculate the
 radius of a circular orbit with that period. We select an initial
-orbital period of 100 s for this simulation, so that the separation is
+orbital period of $T = 100$ s for this simulation, so that the separation is
 \[
@@
 \[
-  t_{\text{ff}} = \frac{\pi}{2} \sqrt{\frac{r_0^3}{2GM}}.
+  t_{\text{ff}} = \frac{\pi}{2} \sqrt{\frac{r_0^3}{2GM}} = \frac{T}{4\sqrt{2}}.
 \]
@@
 
-\subsection{Keplerian Orbit}\label{Sec:Kepler}
+\subsection{Galilean Invariance}\label{sec:galileo}
+
+It is often stated in the literature that Eulerian methods for hydrodynamics with grids fixed in space do not obey the Galilean invariance of the underlying Euler equations, so that simulations moving at a uniform bulk velocity will appear different than an equivalent stationary simulation.
+If true, this means that codes such as CASTRO may be fundamentally inappropriate for merger problems (though any errors should be greatly diminished in a reference frame that rotates with the initial binary orbit).
+Recently, this has come up in two ways which are of note for us in the present study.
+We will explain these situations and run tests to determine whether this actually is a significant concern.
+
+\citet{arepo} performed a Kelvin-Helmholtz instability test and showed that (at low resolution) a fixed-grid code failed to develop the expected fluid instability when the whole fluid was moving at a strongly supersonic uniform velocity.
+This contrasted with the results of the moving-mesh code AREPO being presented in that study, which demonstrated Galilean invariance even at large bulk velocities.
+If verified, this claim would have important consequences on how much we can trust the ability of CASTRO to test the violent merger progenitor model.
+Shearing between the material flowing out of the secondary and material near the surface of the primary may trigger fluid instabilities that play an important role in the evolution of that gas, which is the site of the initial detonation in the prompt explosion model.
+\citet{guillochon:2010} showed for their simulation that Kelvin-Helmholtz instabilities produced this way may raise the temperature of the accreting material enough to ignite a detonation.
+Therefore if we are not correctly reproducing the characteristics of the Kelvin-Helmholtz instability in the case where there is significant mass motion on the grid, we cannot be confident that a detonation (or lack thereof) is not numerically seeded.
+However, \citet{robertson:2010} observe that Galilean invariance of simulation results for the Euler equations occurs only because of truncation error in the discretization of the fluid equations.
+This takes the form of a numerical diffusion term which is dependent on velocity (and also resolution).
+The advantage of a moving-mesh code is that the mesh everywhere moves with the local flow velocity, which substantially reduces the numerical diffusion.
+\citeauthor{robertson:2010} argue that the differences seen between the moving-mesh and fixed-grid code are caused by the interaction of this numerical diffusion with small-scale instabilities (which may be physical or numerical) which couple with and fundamentally alter the large-scale modes.
+Small-scale instabilities are seeded in this problem by the choice of a sharp initial discontinuity between the fluids.
+Crucially though, \citeauthor{robertson:2010} point out that this problem does not converge with resolution and so it is not possible to know the correct behavior of this problem.
+As such, we do not know whether the small-scale modes found in AREPO are real, so the problem is not useful in formally discriminating between methodologies.
+They instead propose an alternate test with a smoother initial contact.
+This converges to the same solution qualitatively in both the stationary and bulk velocity cases, indicating that the code does generally maintain Galilean invariance (to some specified error that depends on resolution and the uniform flow speed).
+We will see whether we can reproduce this result, as well as a result by \cite{wadsley:2008}, who used the FLASH code to simulate a hot bubble subject to mixing by the Kelvin-Helmholtz instability, and also found that the mixing was affected by a uniform bulk velocity.
+
+A related question is whether our code reliably simulates the bulk motion of the stars across the grid, and whether such bulk motion affects the stability of the star.
+This concern is prompted by the study of \cite{tasker:2008}, who studied the effect of uniform translation on the stability of a spherically symmetric model for a galaxy cluster.
+They compared the radial profile of the cluster at initialization and after a period of time evolution.
+Using FLASH and ENZO, they found that a static cluster retains its shape at high enough resolution, while uniform translation of the cluster causes mixing of the core material due to numerical diffusion which results in an underestimation of the core's true density.
+The SPH codes they used did a better job maintaining the core density.
+We will perform a variant of this test using white dwarf models.
+
+\subsubsection{Kelvin-Helmholtz Instability}\label{sec:khi}
+
+\subsubsection{Hot Bubble}\label{sec:hot_bubble}
+
+\subsubsection{Moving Star}\label{sed:moving_star}
+
+\subsection{Keplerian Orbit}\label{sec:Kepler}
 
@@
 
-effect of boundary conditions (size of domain and James vs.\ double monopole)
+effect of boundary conditions (size of domain and possibly multipole dependence)
 

### 2014-12-21 papers/wdmerger_I/paper.tex
Added a subsection on our uniform sphere and cube convergence test for the Poisson solver.

@@
 (Nor would we want to, because in some cases because the velocity becomes spuriously large and would drag down the global timestep.) Therefore we also insert a check at the end of every hydrodynamics timestep that resets the velocity and/or sound-speed (indirectly, using the temperature) of any zone that violates the CFL criterion given the timestep we chose at the beginning of the cycle.
-In practice, we find that this is mainly necessary only for simulations in the inertial reference frame of the white dwarf merger, which we do not normally do for the reasons stated in Section \ref{sec:Kepler}.
+In practice, we find that this is mainly necessary only for simulations in the inertial reference frame of the white dwarf merger, which we do not normally do for the reasons stated in Section \ref{sec:kepler}.
 
@@
 We wish to emphasize that although it is commonly stated in the literature that grid-based codes poorly conserve angular momentum, this is not generally true.
-When the resolution is sufficiently high and a rotating reference frame is employed, excellent conservation properties can result, as demonstrated in Section \ref{sec:Kepler}.
+When the resolution is sufficiently high and a rotating reference frame is employed, excellent conservation properties can result, as demonstrated in Section \ref{sec:kepler}.
 We note that as the stars begin to coalesce, the rotating reference frame will no longer provide a good approximation to the spatial motion of the stars and then they will begin to significantly move around the domain.
@@
 
-\subsubsection{Boundary Conditions}
+\subsubsection{Coupling to Hydrodynamics}\label{sec:gravity_hydro_coupling}
+
+\subsubsection{Boundary Conditions}\label{sec:gravity_boundary_conditions}
 
@@
 $\Delta V$ is the volume of the zone.
-We have implemented this as an option\footnote{It is controlled with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in CASTRO.
+We have implemented this as an option \footnote{It is controlled with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in CASTRO.
 If there are $N$ zones per spatial dimension, then there are $6 N^2$ boundary zones, and each boundary zone requires a sum over $N^3$ zones, so the direct computation of the boundary conditions scales as $N^5$.
@@
 The original release of CASTRO featured the crudest possible approximation: a monopole prescription, where the boundary values were computed by summing up all the mass on the domain and treating it as a point source at the domain center.
-This is exactly correct only for a spherically symmetric mass distribution, and indeed this is often a sufficient approximation for single-star Type Ia supernova simulations that employ self-gravity (ref).
+This is exactly correct only for a spherically symmetric mass distribution, and therefore is best suited for problems like single-star Type Ia supernova simulations that employ self-gravity.
 However, for a problem like that of a binary star system with significant departures from spherical symmetry, this assumption fails to produce accurate boundary values.
-This results in a significant drift of the center of the mass of the system over time.
+This results in a significant drift of the center of the mass of the system over time, which will be discussed in Section \ref{sec:kepler}.
 
@@
 We then directly compute the value of the potential on all of the $6N^2$ boundary zones.
-Since the multipole moments only need to be calculated once, the full operation scales only as $N^3$.
+Since the multipole moments only need to be calculated once per Poisson solve, the full operation scales only as $N^3$.
 The amount of time required to calculate the boundary conditions will be directly related to the chosen value of $l_{\text{max}}$, so there is a trade-off between computational expense and accuracy of the result.
-We find that for the early stages of the evolution of the binary system, $l_\text{max} \approx 6$ is sufficient.
+
+% Insert multipole vs. exact calculation here
+
+\subsubsection{Convergence Testing}\label{sec:gravity_convergence_testing}
+
+Since the results of a merger simulation depend strongly on gravity, it is important to check whether proper numerical convergence is achieved for the Poisson solver.
+To do so, we created a simple test that inserts a sphere of radius $R$ and uniform mass density $\rho$ onto our grid, and used CASTRO to calculate the gravitational potential $\phi$ of this setup.
+We ensure that $R$ is an integer multiple of the grid spacing, and the center of the sphere is at the origin.
+The problem domain for our simulations is $[-1.6, 1.6]^3$, and we take $R = 0.5$ and $\rho = 10^3$.
+The zones with $r > R$ are filled with an ambient material of very low density.
+We run this problem at multiple resolutions corresponding to jumps by a factor of two.
+For comparison, at each grid point we evaluate the analytical potential of a uniform sphere, which can be easily determined using Gauss' law:
+\begin{equation}
+  \phi_{\text{sphere}}(r) = \frac{GM}{r} \times \begin{cases} (3R^2 - r^2)/(2 r^2) & r \leq R \\ 1 & r > R \end{cases},\label{eq:sphere-analytical}
+\end{equation}
+where $M = 4\pi R^3 / 3$ is the mass of the sphere. Defining the $L^2$ norm of a field $f$ as
+\begin{equation}
+  L^2(f) = \left(\sum_{i,j,k} dx\, dy\, dz\, f_{ijk}^2\right)^{1/2},
+\end{equation}
+we measure the numerical error by calculating the $L^2$ norm of the error and normalizing it by the $L^2$ norm of the analytical solution:
+\begin{equation}
+  \text{Error} = \frac{L^2(\phi - \phi_{\text{sphere}})}{L^2(\phi_{\text{sphere}})}.
+\end{equation}
+
+\begin{figure}
+  \label{fig:gravity_convergence}
+\end{figure}
+
+The results of this test are plotted in Figure \ref{fig:gravity_convergence}.
+We find that convergence is actually substantially better than second-order, and that the difference is largest at low resolutions.
+The explanation for this is that we are attempting to model a spherical object on a rectangular grid.
+At very low resolution, the object does not look very spherical.
+As the resolution is increased, the total amount of mass on the grid, as well as its location, will change as the sphere fills out.
+This means we are combining the true accuracy bonus from increased resolution with the artificial accuracy bonus from getting closer to solving the problem we are supposed to be solving.
+
+We can solve one of these two sources of error by evaluating Equation \ref{eq:sphere-analytical} with a mass $M$ that corresponds to the amount of mass actually on the grid, defined as the sum of the density multiplied by the volume of each cell for all cells that have greater than ambient density.
+The resolution study for this case is also plotted in Figure \ref{fig:gravity_convergence}.
+We still obtain convergence better than second-order, indicating that we still have the geometrical problem.
+
+The only way to fully eliminate this effect is to test a problem that does not change with resolution.
+The obvious companion is a cube of uniform density $\rho$, where now $R$ is half of the side length of the cube.
+At each resolution we use the same $R$ as for the sphere, which ensures that the cube always fills exactly the same fraction of the domain and thus has the same mass, so the only improvement comes from better sampling at higher resolution.
+The potential for this object has been worked out analytically by \cite{waldvogel:1976} (see also a similar result in \cite{hummer:1996}).
+The potential is in Equation 15 of that paper, though the last term is missing a factor of $1/2$, which destroys the symmetry.
+Inserting this missing factor and performing a simple coordinate transformation so that the center of the cube is at the origin, the potential is
+\begin{align}
+  \phi_{\text{cube}}(x,y,z) &= G\rho\sum_{i,j,k=0}^1\left(x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right) + y_j z_k\, \text{tanh}^{-1}\left(\frac{x_i}{r_{ijk}}\right) + z_k x_i\, \text{tanh}^{-1}\left(\frac{y_j}{r_{ijk}}\right) \right.\notag \\
+  &\left.
+- \frac{x_i^2}{2}\,\text{tan}^{-1}\left(\frac{y_j z_k}{x_i r_{ijk}}\right) - \frac{y_j^2}{2}\,\text{tan}^{-1}\left(\frac{z_k x_i}{y_j r_{ijk}}\right) - \frac{z_k^2}{2}\,\text{tan}^{-1}\left(\frac{x_i y_j}{z_k r_{ijk}}\right)\right)
+\end{align}
+where $x_0 = R + x$, $x_1 = R - x$, $y_0 = R + y$, $y_1 = R-y$, $z_0 = R+z$, $z_1 = R-z$, and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$.
+In \texttt{FORTRAN} and \texttt{C}, the inverse hyperbolic tangent is \texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not} \texttt{atan2}).
+This formula is valid both inside and outside the cube.
+The normalized $L^2$ error norm for this problem is also shown in Figure \ref{fig:gravity_convergence}, and indeed we see near-perfect second-order scaling.
+
+The main lesson here is that in a convergence study, it is important to ensure that the physical problem does not change with resolution.
+Since in the case of spherical objects on rectangular grids the effect is to artificially boost convergence with resolution, in a simulation with spherical objects like stars one can envision a scenario of being fooled into believing apparently good convergence results that are simply a convolution of artificially high gravitational convergence and poor convergence in the hydrodynamics.
+A convergence study in this case is only fully valid if there is reason to be confident that this effect is negligible compared to other factors.
+
+Another factor that may affect convergence testing is easily seen in Figure \ref{fig:gravity_convergence}, where at the highest resolution the error flattens out.
+This occurs because of a limitation in the accuracy of the multigrid solver.
+For problems of this size we find empirically that the multigrid solver does not converge for error tolerances smaller than about $10^{-11}$, which enforces a lower limit on the total error.
+This is also something to be aware of in assessing convergence for our full simulations.
+
+\subsection{Rotation}\label{sec:rotation}
 
@@
 We determine the initial separation to be consistent with the
-simulation performed in Section \ref{sec:Kepler}; that is, we select
+simulation performed in Section \ref{sec:kepler}; that is, we select
 an initial orbital period and use Kepler's third law to calculate the
@@
 
-\subsection{Keplerian Orbit}\label{sec:Kepler}
+\subsection{Keplerian Orbit}\label{sec:kepler}
 

### 2014-12-21 papers/wdmerger_I/paper.tex
Move the subsection on initial conditions to its own full section.

@@
 
-For the evolution of binary systems, it is most natural to evolve the two stars in a frame that is co-rotating at the same period as the orbital period.
-CASTRO has the ability to evolve systems in a rotating reference frame, as discussed in the original code paper.
-Source terms corresponding to the Coriolis and centrifugal force terms are added to the momentum and energy equations.
-In this frame, the stars essentially remain stationary in their original positions due to the centrifugal force supporting against the gravitational attraction, and will remain this way as long as significant mass transfer does not occur.
-\cite{swesty00} demonstrated (in the context of neutron star mergers) that conservation of angular momentum is much easier to obtain in the rotating reference frame than in an inertial frame in which stars advect large amounts of material around the domain.
-We wish to emphasize that although it is commonly stated in the literature that grid-based codes poorly conserve angular momentum, this is not generally true.
-When the resolution is sufficiently high and a rotating reference frame is employed, excellent conservation properties can result, as demonstrated in Section \ref{sec:kepler}.
-We note that as the stars begin to coalesce, the rotating reference frame will no longer provide a good approximation to the spatial motion of the stars and then they will begin to significantly move around the domain.
-This is not necessarily problematic because the most important feature of the rotating frame is that it helps ensure that the initial coalescence is not the result of spurious numerical loss of angular momentum.
-When significant mass transfer sets in and evolution proceeds on a dynamical timescale, the conservation properties may be slightly worse but angular momentum conservation is also less important.
-
 \subsection{Microphysics}
@@
 Reactions?
-
-\subsection{Initial Models}
-\label{sec:initial_models}
-
-We generate initial models by integrating the equation of hydrostatic
-equilibrium, taking the temperature and composition to be constant,
-and using the general stellar equation of state.  This results in
-a single nonlinear equation to find the density in a zone given the
-conditions in the zone beneath it:
-\begin{equation}
-\frac{p(\rho_{i+1}) - p_i}{\Delta x} = \frac{1}{2} (\rho_i + \rho_{i+1}) g_{i+1/2}
-\end{equation}
-This is solved via Newton iteration to give $\rho_{i+1}$.  Here, $g_{i+1/2}$
-is the gravitational acceleration at the interface between zones $i$ and $i+1$,
-found by simply adding up all the mass from zones $1, \ldots, i$ to get the
-enclosed mass, $M_{i+1/2}$, and then $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$.
-
-To start the integration off, we need a central density.  We guess at
-this, and then iterate over the entire integration procedure to find
-the central density needed to yield the desired total mass.  Finally,
-we generate the initial model at the same resolution as the finest
-grid in our simulation.
-
-We map the 1-d model onto the 3-d Cartesian grid by taking density,
-temperature, and composition as the independent variables,
-interpolating these to the cell-centers, and then calling the equation
-of state to initialize the remaining terms.  The velocity is taken to
-be zero in the rotating frame initially.
 
@@
 
+For the evolution of binary systems, it is most natural to evolve the two stars in a frame that is co-rotating at the same period as the orbital period.
+CASTRO has the ability to evolve systems in a rotating reference frame, as discussed in the original code paper.
+Source terms corresponding to the Coriolis and centrifugal force terms are added to the momentum and energy equations.
+In this frame, the stars essentially remain stationary in their original positions due to the centrifugal force supporting against the gravitational attraction, and will remain this way as long as significant mass transfer does not occur.
+\cite{swc:2000} demonstrated (in the context of neutron star mergers) that conservation of angular momentum is much easier to obtain in the rotating reference frame than in an inertial frame in which stars advect large amounts of material around the domain.
+We wish to emphasize that although it is commonly stated in the literature that grid-based codes poorly conserve angular momentum, this is not generally true.
+When the resolution is sufficiently high and a rotating reference frame is employed, excellent conservation properties can result, as demonstrated in Section \ref{sec:kepler}.
+We note that as the stars begin to coalesce, the rotating reference frame will no longer provide a good approximation to the spatial motion of the stars and then they will begin to significantly move around the domain.
+This is not necessarily problematic because the most important feature of the rotating frame is that it helps ensure that the initial coalescence is not the result of spurious numerical loss of angular momentum.
+When significant mass transfer sets in and evolution proceeds on a dynamical timescale, the conservation properties may be slightly worse but angular momentum conservation is also less important.
+
 %==========================================================================
-% Test Problems
+% Initial Models and Problem Setup
+%==========================================================================
+
+\section{Initial Models and Problem Setup}
+\label{sec:initial_models}
+
+At the start of any full simulation, we generate initial model white dwarfs by integrating the equation of hydrostatic equilibrium, taking the temperature and composition to be constant, and using the general stellar equation of state.
+This results in a single non-linear equation to find the density in a zone given the conditions in the zone beneath it:
+\begin{equation}
+\frac{p_{i+1} - p_i}{\Delta x} = \frac{1}{2} (\rho_i + \rho_{i+1}) g_{i+1/2}.
+\end{equation}
+This equation is a function of $\rho_{i+1}$ only since the pressure is uniquely determined by the density in this case.
+$\rho_i and p_i$ are known,  $g_{i+1/2}$ is the gravitational acceleration at the interface between zones $i$ and $i+1$, found by simply adding up all the mass from zones $1, \ldots, i$ to get the enclosed mass, $M_{i+1/2}$, and then setting $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$.
+We solve this equation for $\rho_{i+1}$ using a Newton-Raphson iteration.
+
+We desire to specify the mass of the white dwarf, as well as its temperature and composition.
+To start the integration off, we therefore need to guess at a central density.
+We then do a secant iteration over the entire integration procedure to find the central density needed to yield the desired total mass.
+We note that the grid spacing $\Delta x$ is chosen such that it is equal to the zone size on the finest level of our problem, which is known in advance since we specify the domain size and the maximum number of possible levels of refinement.
+
+We map the 1D model onto the 3D Cartesian grid by taking density, temperature, and composition as the independent variables, interpolating these to the cell centers, and then calling the equation of state to initialize the remaining terms.
+The interpolation process divides each zone into $n_{\text{sub}}$ sub-zones of equal volume for the purpose of sampling the 1D model, and the sub-zones are added together to obtain the full zone's characteristics.
+This sub-grid-scale interpolation is useful near the edge of the star, where the density falls off rapidly with radius.
+
+For a single star simulation, the star is simply placed at the center of the computational domain, which we take to be the origin.
+For a binary star simulation, we take as parameters the mass of the two white dwarfs and the initial orbital period $T$.
+Using Kepler's third law and assuming a circular orbit, we can then work out the orbital separation $a$:
+\begin{equation}
+  a = \left(\frac{4\pi^2 T^2}{GM}\right)^{1/3}.
+\end{equation}
+Here $M = M_P + M_S$ is the total mass of the system, where $M_P$ is the specified \textit{primary} mass and $M_S$ is the specified \textit{secondary} mass.
+The primary WD will always be on the left side of the computational domain for our simulations, and is more massive than the secondary.
+This reflects the usual terminology in the literature where the primary WD is the accretor and the secondary is the donor.
+The center of mass is located at the center of the computational domain, and the stars lie along the $x$ axis, so that the primary's center of mass is located at $x = -(M_S / M)\, a$ and the secondary's center of mass is located at $x = (M_P / M)\, a$.
+
+The initial velocity is taken to be zero in if we are in the reference frame that rotates with the WDs, and if we are in the inertial frame the velocity is set equal to the rigid rotation rate corresponding to the specified period $T$.
+
+We do not attempt to enforce equilibrium with an additional relaxation step.
+This will be an important part of future work in this series, as numerous groups working on binary evolution \citep{swc:2000,motl:2002,rosswog:2004,dan:2011,pakmor:2012:gadget} have commented on the importance of equilibrium initial conditions in determining the evolution of the system.
+In particular, we plan to compare the self-consistent field method (which finds an equilibrium solution \textit{a priori}) and the method of \cite{rosswog:2004}, which starts with our method and then attempts to guide the evolution toward the equilibrium state using suitably defined damping terms.
+However, these issues are not considered for the present study, as we are not yet attempting to study the merger process.
+
+%==========================================================================
+% Numerical Test Problems
 %==========================================================================

### 2014-12-22 papers/wdmerger_I/paper.tex
Kepler's third law was inverted.
Added a description of the problem setup for the KH tests to the paper.

@@
 \begin{equation}
-  a = \left(\frac{4\pi^2 T^2}{GM}\right)^{1/3}.
+  a = \left(\frac{GM T^2}{4\pi^2}\right)^{1/3}.
 \end{equation}
@@
 \subsubsection{Kelvin-Helmholtz Instability}\label{sec:khi}
+
+Following \cite{robertson:2010}, we set up a Kelvin-Helmholtz test in the following way.
+The problem domain runs from 0 to 1 in both the $x$ and $y$ directions.
+Though this is a two-dimensional test, we run CASTRO in 3D and replicate the problem along 8 cells in the $z$ direction; the physical size of these zones is determined so that $dz = dx = dy$, a requirement at present in CASTRO.
+The problem involves a fluid slab of density $\rho_1 = 2.0$ traveling rightward in the $x$-direction at velocity $v_1 = 0.5$, sandwiched by a fluid of density $\rho_2 = 1.0$ traveling leftward at velocity $v_2 = -0.5$.
+The density gradient is in the $y$ direction, so this creates a velocity shear along the interface between the fluids.
+The density and velocity distribution on the computational domain are given by:
+\begin{align}
+  \rho &= \rho_1 + R(y)\left[\rho_2 - \rho_1\right] \\
+  v_x  &= v_1 + R(y)\left[v_2 - v_1\right] \\
+  v_y  &= v_{\text{bulk}} + v^\prime
+\end{align}
+Here $R(y)$ is a ramp function that describes the transition between the two fluids, while $v_{\text{bulk}}$ is the bulk motion of the fluid in the $y$ direction and $v^\prime$ is the velocity perturbation that seeds the instability.
+The problem will be established for two sets of initial conditions (ICs), which we follow \citeauthor{robertson:2010} in calling ICs A and B.
+They differ in their ramp function ($R_A$ and $R_B$ respectively), as well as the initial perturbation ($v^\prime_A$ and $v^\prime_B$ respectively), and the frequency of the perturbation ($n_A = 4$ and $n_B = 2$):
+\begin{align}
+  R_A &= \begin{cases} 0 & |y - 0.5| > 0.25 \\ 1 & |y - 0.5| < 0.25 \end{cases} \\
+  R_B &= \Big\{\left[1 + \text{exp}(-2(y-0.25)/\Delta_y)\right]\left[1 + \text{exp}(2(y-0.75)/\Delta_y)\right]\Big\}^{-1} \\
+  v^\prime_A &= w_0\, \text{sin}\left(n_A\, \pi\, x\right) \left\{\text{exp}\left[-\frac{(y-0.25)^2}{2\sigma^2}\right] + \text{exp}\left[-\frac{(y-0.75)^2}{2\sigma^2}\right]\right\} \\
+  v^\prime_B &= w_0\, \text{sin}\left(n_B\, \pi\, x\right).
+\end{align}
+Here $w_0 = 0.1$ is the scale of the velocity perturbation, $\sigma = 0.05/\sqrt{2}$ controls the width of the Gaussian for IC A, and $\Delta_y$ is the transition distance scale for the smooth ramp of IC B.
+The pressure everywhere is set to $p = 2.5$, and we run this with a gamma-law equation of state set to $\gamma = 5/3$.
+Plotfiles are generated every 0.01 seconds, and the problem is run until $t = 2$.
+
+We run the problem for $v_\text{bulk} = [0, 1, 3, 10, 30, 100]$, and for each set of initial conditions run the problem at resolutions of $64^2$, $128^2$, $256^2$, $512^2$.
+In addition, for each initial condition we run a simulation at the very high resolution of $2560^2$ for the stationary problem only.
+This serves as a reference solution to gauge the extent to which the bulk flow affects the development of the fluid instability.
 

### 2015-01-01 papers/wdmerger_I/paper.tex
Add DD acronym

@@
 However, in recent years many alternative progenitor models have been discussed.
-A leading candidate for explaining some or most of these explosions is the double-degenerate model, in which two white dwarfs merge and the merged object reaches the conditions necessary for a thermonuclear ignition \citep{ibentutukov:1984,webbink:1984}.
+A leading candidate for explaining some or most of these explosions is the double-degenerate (DD) model, in which two white dwarfs merge and the merged object reaches the conditions necessary for a thermonuclear ignition \citep{ibentutukov:1984,webbink:1984}.
 Another is the double detonation scenario, where accretion of material onto a sub-Chandrasekhar white dwarf would lead to a detonation inside the accreted envelope, and this would send a compressional wave into the core of the star that would trigger a secondary detonation.

### 2015-01-01 papers/wdmerger_I/paper.tex
Some more introduction cleanup

@@
 There are several observational reasons why double-degenerate systems are a promising progenitor system for at least a substantial fraction of normal SNe Ia.
-No conclusive evidence exists for a surviving companion star of a SN Ia; this is naturally explained by the DD model because both WDs are destroyed in the merger process.
-Similarly, pre-explosion images of the SN Ia systems have never clearly turned up a companion star, and in some cases a large fraction of the SD parameter space is excluded.
+No conclusive evidence exists for a surviving companion star of a SN Ia; this is naturally explained by the DD model because both WDs are likely to be destroyed in the merger process.
+Similarly, pre-explosion images of the SN Ia systems have never clearly turned up a companion star, and in some cases a large fraction of the parameter space for the nature of the companion star is excluded.
 Additionally, not enough progenitor systems are seen for the SD case to match the observed local SN Ia rate, whereas the number of white dwarf binaries may be sufficient to account for this rate.
@@
 The first attempts to model the results of the merger process came in the 1980s.
-\cite{nomotoiben:1985} demonstrated that if the secondary overflowed its Roche lobe and its mass accreted onto the primary near the Eddington rate, that off-center carbon ignition would occur.
+\cite{nomotoiben:1985} demonstrated that off-center carbon ignition would occur in the more massive white dwarf as it accreted mass near the Eddington rate from the less massive white dwarf overflowing its Roche lobe.
 \cite{saionomoto:1985} tracked the evolution of the flame and found that it propagated quiescently into the center, converting the carbon-oxygen white dwarf into an oxygen-neon-magnesium white dwarf.
@@
 The first three-dimensional simulations of double degenerate systems were performed by \citet{benz:1990}, who used the smoothed particle hydrodynamics (SPH) method to simulate the merger process.
-They found that if the lower-mass secondary was close enough to the primary to begin mass transfer on a dynamical time scale, the secondary completely disrupted and formed a hot envelope on the primary, with a centrifugally-supported accretion disk surrounding the core and envelope.
-They noted that the envelope temperature was hot enough to generate carbon fusion but neglected reactions in their simulation.
+They found that if the lower-mass star (generally called the ``secondary'') was close enough to the more massive star (the ``primary'') to begin mass transfer on a dynamical time scale, the secondary completely disrupted and formed a hot envelope around the primary, with a centrifugally-supported accretion disk surrounding the core and envelope.
+They noted that the envelope temperature was hot enough to generate carbon fusion but neglected nuclear reactions in their simulation.
 Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009,raskin:2012} corroborated this result.
 This validation of the prediction of \cite{tutukov_yungelson:1979} did not instill confidence in the community.
-\cite{mochkovitch_livio:1990} and \cite{livio:2000} observed that turbulent viscosity would be sufficiently large for angular momentum to be removed from the disk at a rate high enough to generate these troublesome accretion timescales.
+\cite{mochkovitch_livio:1990} and \cite{livio:2000} observed that turbulent viscosity would be sufficiently large for angular momentum to be removed from the disk at a rate high enough to generate the troublesome accretion timescales.
 Based on all of this evidence, the review of \cite{hillebrandtniemeyer2000} argued that the model was only viable if the accretion-induced collapse problem could be avoided.
@@
 They propagated this detonation throughout the system and found that it would observationally appear as a subluminous Type Ia supernova.
-This was the first time any simulation successfully reproduced at least some characteristics of a SN Ia.
+This was the first time a DD simulation successfully reproduced at least some characteristics of a SN Ia.
 \cite{pakmor:2011} tried a few different mass combinations and found empirically that this would hold as long as the secondary was at least 80\% as massive as the primary.
@@
 When the initial conditions are constructed in exact equilibrium, the system can be stable for tens of orbital periods, substantially changing the character of the mass transfer phase.
-However, one limitation of this work is that the authors used a polytropic equation of state and thus could not consider nuclear reactions.
+However, one limitation of this series of studies is that the authors used a polytropic equation of state and thus could not consider nuclear reactions.
 \cite{guillochon:2010} and \cite{dan:2011} improved on this using a realistic equation of state, a nuclear reaction network, and a similar approach to the equilibrium initial conditions, and found substantial agreement with the idea that mass transfer occurs in a stable manner over tens of orbital periods.
@@
 
-The most recent developments give some areas of agreement and some of remaining uncertainty.
+The very latest developments give some areas of agreement and some of remaining uncertainty.
 \cite{pakmor:2012} performed a merger scenario with a $1.1\ M_\odot$ and $0.9\ M_\odot$ setup, with better treatment of the initial conditions, and also found that the merger process happened over more than ten orbits.

### 2015-01-01 papers/wdmerger_I/paper.tex
Added a reason for why we need a grid-based code to the introduction.

@@
 They pointed out that the previously mentioned simulations generally all shared a significant drawback, which was that their initial conditions were not carefully constructed.
-\cite{motl:2002}, \cite{dsouza:2006}, and \cite{motl:2007} (the first three-dimensional grid-based simulations of white dwarf mergers) pioneered the study of looking at the long-term dynamical evolution of merger scenarios after constructing exact equilibrium initial conditions.
+\cite{motl:2002}, \cite{dsouza:2006}, and \cite{motl:2007} (the first three-dimensional grid-based simulations of mass transfer in binary white dwarf systems) pioneered the study of looking at the long-term dynamical evolution of binary white dwarf systems after constructing exact equilibrium initial conditions.
 In contrast to the other studies, they found that paying close attention to the accuracy of the initial conditions of the simulation has important effects.
@@
 Thus at least a preliminary agreement may be that systems containing helium could robustly lead to events resembling SNe Ia, as well as very massive carbon-oxygen binaries.
+
+Given the above, why is another approach using a different simulation code warranted?
+First and foremost, reproducibility of the results across simulation codes and algorithms is important for gauging confidence in this result.
+With only a couple of exceptions, most of the existing results that study the viability of the DD progenitor as an explanation for Type Ia supernovae (that is, including a temperature-dependent equation of state and nuclear reactions) have used the SPH method.
+SPH codes have a number of features which do aid them in the study of these systems, such as excellent conservation of angular momentum.
+However, the question of whether a prompt detonation in a merger actually happens depends in detail on the nature of the gas at the interface between the two stars, which is at much lower density than the rest of the stellar material.
+The SPH codes use uniform mass particles, so their effective resolution is \textit{lowest} at the stellar surface.
+In contrast, a grid-based code with adaptive mesh refinement can zoom in on the regions where hotspots will develop, while also maintaining high enough resolution in the high-density regions to adequately capture the large-scale mass transfer dynamics.
+However, there are also outstanding questions of convergence in SPH (e.g.
+\cite{zhu_SPH:2014}) and whether the method correctly captures fluid instabilities.
+This is an important question for white dwarf mergers because of the likely importance small-scale instabilities will have on the evolution of the low-density gas at the primary's surface.
+The pioneering work of \cite{agertz:2007} compared grid and SPH codes and found some important differences.
+Most relevant for this discussion is that the SPH codes could not adequately handle mixing from the Kelvin-Helmholtz instability in the test they propose.
+\cite{price:2008} attributed this to the fact that the standard SPH evolution equations do not have a mechanism for capturing discontinuities in internal energy, and showed that the addition of an artificial thermal conductivity can dramatically improve the ability of the SPH codes to exhibit this instability.
+There have since been a number of papers discussing this issue, but to our knowledge none of these improvements have yet been incorporated into an SPH model of a WD merger.
+Another reason to be skeptical is that the detonation itself was inserted manually in \cite{pakmor:2010} based on small-scale detonation simulations.
+A self-consistent evolution of a thermonuclear detonation, if it existed, would significantly build confidence in the progenitor model.
 

### 2015-01-01 papers/wdmerger_I/paper.tex
Slight cleanup of hydrodynamics section, plus addition of the Euler equations.

@@
 
+To study the white dwarf merger problem, we use the grid-based hydrodynamics code CASTRO \citep{castro}.
+CASTRO solves the Euler equations, along with the inclusion of optional modules for gravity, nuclear reactions and thermodynamics.
+CASTRO is based on the BoxLib adaptive-mesh refinement (AMR) framework \citep{rendleman:2000}, which represents fluid data on a mesh where regions of interest have higher spatial resolution.
+CASTRO is highly parallel and is designed for large-scale use on modern supercomputers; see Section \ref{sec:Performance} for information on how CASTRO performs for our problem.
+The next few subsections describe our approach to each of the physics components used in this work.
+We direct the reader to the original code paper for a full description of CASTRO's approach to solving the equations of hydrodynamics.
+In this work, we report mainly on the changes we have made to the code since its original release, for the purpose of approaching this problem.
+
 \subsection{Hydrodynamics}
 
-To study the white dwarf merger problem, we use the grid-based hydrodynamics code CASTRO \citep{castro}.
-CASTRO solves the Euler equations, along with the inclusion of optional modules for gravity, nuclear reactions and thermodynamics.
-We direct the reader to the original code paper for a description of CASTRO's approach to solving the equations of hydrodynamics.
-In this work, we report only on the changes we have made to the code since its original release, for the purpose of approaching this problem.
-CASTRO is based on the BoxLib adaptive-mesh refinement (AMR) framework (ref), which represents fluid data on a mesh where regions of interest have higher spatial resolution.
-CASTRO is highly parallel and is designed for large-scale use on modern supercomputers; see Section \ref{sec:Performance} for information on how CASTRO performs for our problem.
+The Euler equations for hydrodynamics (in the absence of source terms) in conservative form are:
+\begin{align}
+  \frac{\partial \rho}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}) \\
+  \frac{\partial \rho \mathbf{u}}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}\mathbf{u}) - \bm{\nabla}p \\
+  \frac{\partial \rho E}{\partial t} &= -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u}).
+\end{align}
+Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity vector, $p$ is the pressure, and $E = \rho \mathbf{u}^2 / 2 + e$ is the total energy density where $e$ is the internal (thermal) energy density.
 
@@
 
-The boundary conditions on the hyperbolic system are simply zero-gradient zones that allow material to flow directly out of the domain. Using AMR, we make the coarse grid very large, so as
-to place the boundaries far from the region of interest. This ensures that any boundary effects do not pollute the inner region where the stars will eventually make contact.  We further
-make the restriction that only level-0 (coarse) grids can touch the domain boundary.
-
 CASTRO's approach to adaptive mesh refinement, based on its underlying BoxLib framework, is to refine zones based on certain user-specified criteria that tag regions of interest for higher spatial resolution.
@@
 Generally the number of subcycled timesteps is equal to the jump in refinement between levels, so for example on a grid with three levels and two jumps of four in refinement, the level 2 zones will have 16 times higher spatial resolution than the coarse grid and there will be 16 level 2 timesteps per level 0 timestep.
+
+The boundary conditions on the hyperbolic system are simply zero-gradient zones that allow material to flow directly out of the domain.
+Using AMR, we make the coarse grid large enough that the boundaries are relatively far from the region of interest.
+This ensures that any boundary effects do not pollute the inner region where the stars will eventually make contact.
+We further make the restriction that refined grids cannot reach the domain boundary.
 

### 2015-01-01 papers/wdmerger_I/paper.tex
Slight clarification of how SPH handles the KH instability.

@@
 Most relevant for this discussion is that the SPH codes could not adequately handle mixing from the Kelvin-Helmholtz instability in the test they propose.
-\cite{price:2008} attributed this to the fact that the standard SPH evolution equations do not have a mechanism for capturing discontinuities in internal energy, and showed that the addition of an artificial thermal conductivity can dramatically improve the ability of the SPH codes to exhibit this instability.
-There have since been a number of papers discussing this issue, but to our knowledge none of these improvements have yet been incorporated into an SPH model of a WD merger.
+As pointed out by \cite{price:2008}, this is not a result of SPH being inherently unable to model this instability, but instead it is attributed to the fact that the standard SPH evolution equations do not have a mechanism for capturing discontinuities in internal energy.
+\citeauthor{price:2008} showed that the addition of an artificial thermal conductivity can dramatically improve the ability of the SPH codes to exhibit this instability.
+There have since been a number of other papers discussing this issue, but to our knowledge none of these improvements have yet been incorporated into an SPH model of a WD merger.
 Another reason to be skeptical is that the detonation itself was inserted manually in \cite{pakmor:2010} based on small-scale detonation simulations.

### 2015-01-01 papers/wdmerger_I/paper.tex
Added sentence on initial model section to introduction.
Added a description of the three major methods we're considering for the gravitational coupling to the hydrodynamics.

@@
 Section \ref{sec:Numerical Methodology} describes our code and why it can provide useful results compared to other methodologies used for this problem.
+Section \ref{sec:initial_models} describes the method we use for setting up a binary white dwarf simulation.
 Section \ref{sec:Tests} discusses a few test problems that we use to demonstrate that our code accurately solves the equations of fluid dynamics.
@@
 \begin{align}
-  \frac{\partial \rho}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}) \\
-  \frac{\partial \rho \mathbf{u}}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}\mathbf{u}) - \bm{\nabla}p \\
-  \frac{\partial \rho E}{\partial t} &= -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u}).
+  \frac{\partial \rho}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}) \label{eq:euler_density}\\
+  \frac{\partial \rho \mathbf{u}}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}\mathbf{u}) - \bm{\nabla}p \label{eq:euler_momentum}\\
+  \frac{\partial \rho E}{\partial t} &= -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u}). \label{eq:euler_energy}
 \end{align}
-Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity vector, $p$ is the pressure, and $E = \rho \mathbf{u}^2 / 2 + e$ is the total energy density where $e$ is the internal (thermal) energy density.
+Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the total energy density where $e$ is the internal (thermal) energy density.
 
@@
 \subsubsection{Coupling to Hydrodynamics}\label{sec:gravity_hydro_coupling}
+
+The effect of gravity on the hydrodynamical evolution is typically incorporated by the use of a source term for the momentum and energy equations. The integral form of the momentum source term is
+\begin{equation}
+  \int \rho \mathbf{g}\, dV
+\end{equation}
+and for the energy source term it is
+\begin{equation}
+  \int \rho \mathbf{u}\cdot\mathbf{g}\, dV \label{eq:cell_center_gravity_source}.
+\end{equation}
+In most hydrodynamics codes these are discretized as $\rho\, \mathbf{g}\, \Delta V$ and $\rho\, \mathbf{u}\,\cdot\mathbf{g}\, \Delta V$, respectively, where $\Delta V$ is the zone volume and $\rho$, $\mathbf{u}$, and $\mathbf{g}$ are evaluated at the zone center.
+As observed by \cite{arepo}, the essential problem with this approach is that fluid movements in a grid code are calculated using fluxes at zone boundaries, while the standard gravity source term is an estimate of the flux using the zone center quantities.
+This can lead to a striking mismatch when the gradient across a zone is large.
+We have found that this is especially important at the edge of a star, where the large density gradient means that the momentum and energy fluxes can transition sharply even across half of a zone width.
+Another problem is that in a merger calculation in which two stars are moving on a grid, large quantities of material are being advected, so even a slight error in the calculation for the high-density material can manifest itself as a long-term error in energy conservation.
+This becomes apparent in Section \ref{sec:kepler}.
+So while a closer look at this may be unwarranted for situations like a stationary single star central explosion calculation, a detailed discussion is certainly called for in the case of an orbiting binary system where getting the gravity right is absolutely necessary for fully trusting the long-term evolution of the system.
+
+\citet[Chapter 4]{shu:1992} observes that it is possible to describe the source term for the momentum equation by taking the divergence of a gravitational stress tensor,
+\begin{equation}
+  G_{ij} = -\frac{1}{4\pi G}\left(g_i g_j - \frac{1}{2}|\mathbf{g}|^2\delta_{ij}\right).
+\end{equation}
+The momentum equations are then written explicitly in conservative form when discretized: just as with the momentum stress tensor, divergences are calculated using discretized fluxes at zone boundaries, and the flux at any zone boundary is added to one cell and subtracted from another, so that the total momentum stays constant to within numerical roundoff error (with the possible exception of domain boundaries).
+This method works because, from the Poisson equation, $\rho$ can be written as proportional to the divergence of $\mathbf{g}$.
+We note that in order for the method to be reliable, the expression for all components of $\mathbf{g}$ must be correctly calculated from $\phi$ on zone boundaries so that the total divergence does indeed recover the density.
+This is straightforward for the component normal to the zone interface, but requires averaging over zones in the perpendicular direction to obtain the transverse components.
+
+Dealing with the energy equation is more difficult.
+The central challenge is to write down a form of the discretized energy equation that explicitly conserves total energy when coupled to gravity.
+When gravity is included, the total energy is
+\begin{equation}
+  \rho E_{\text{tot}} = E_G - \frac{1}{2}\rho\phi = \frac{1}{2}\rho \mathbf{u}^2 + \rho e - \frac{1}{2}\rho\phi,
+\end{equation}
+where $E_G$ is the gas energy from the pure hydrodynamics equation, and the factor of $1/2$ in the gravitational energy term is necessary for simulations with self-gravity to prevent double-counting of interactions.
+Recently, three different approaches have been proposed for achieving this.
+We will briefly examine these and then describe the approach that makes the most sense for us.
+
+\citet{marcello:2012} use the mass continuity equation to rewrite the standard source term from Equation \ref{eq:cell_center_gravity_source}.
+When this is done, the resulting evolution equation for the total energy is:
+\begin{equation}
+  \frac{\partial (\rho E_{\text{tot}})}{\partial t} = -\bm{\nabla}\cdot\left(\rho \mathbf{u} E_{\text{tot}} + p\mathbf{u} - \frac{1}{2}\rho\mathbf{u}\phi\right) + \frac{1}{2}\phi\frac{\partial \rho}{\partial t} - \frac{1}{2}\rho\frac{\partial \phi}{\partial t}.
+\end{equation}
+In practice, we would still actually evolve the total gas energy $E_G$, and so when discretized to second order in time from time level $n$ to time level $n+1$ the gas energy equation would look like
+\begin{align}
+  (\rho E_G)^{n+1} &= (\rho E_G)^{n} + \left(\frac{1}{2} (\rho \phi)^{n+1} - \frac{1}{2} (\rho \phi)^{n}\right) - \Delta t\, \bm{\nabla}\cdot\left(\rho \mathbf{u} E_{\text{tot}} + p\mathbf{u} - \frac{1}{2}\rho\mathbf{u}\phi\right) \notag \\
+   &+ \frac{\Delta t}{2}\left(\phi \frac{\partial \rho}{\partial t}\right)^{n+1/2} - \frac{\Delta t}{2} \left(\rho \frac{\partial \phi}{\partial t}\right)^{n+1/2}.
+\end{align}
+It is not immediately obvious that this is explicitly conservative for the total energy, due to the last two terms on the right-hand side, but written in discrete form the potential is linked to the density in such a way that there is indeed a perfect cancellation of terms between adjacent cells (see Equation 88 in \citeauthor{marcello:2012}).
+Doing it this way requires a prescription for calculating $\phi^{n+1/2}$, the potential at the time halfway between the new and old state, as well as $\dot{\phi}^{n+1/2}$, the time derivative of the potential at that time.
+To second order in time, these can be calculated as $(\phi^{n+1} + \phi^{n})/2$ and $(\phi^{n+1}-\phi^{n})/\Delta t$ respectively.
+So, in practice what we do is to first solve the Poisson equation using $\rho^n$ at the beginning of a time step; then, evolve the pure hydrodynamics equations without the gravitational source to obtain $\rho^{n+1}$; then, solve the Poisson equation for $\phi^{n+1}$; and then to finally calculate the gravitational source terms and add them in.
+An alternative approach is to use the predictor-corrector scheme that CASTRO features, where the time level $n$ source term is included in the hydrodynamics update and then left out of the final update.
+In either case, the time level $n$ calculation of $\mathbf{g}$ is included the PPM reconstruction of the edge states.
+
+\citet{jiang:2013} start from the motivation of extending to the energy equation the flux-based formalism that allows explicit conservation for the momentum equation.
+They determine that the energy is perfectly conserved if we define a gravitational flux $\mathbf{F_g}$ and take its divergence for the energy equation:
+\begin{align}
+  \mathbf{F_g} &= \frac{1}{8\pi G}\left(\phi\bm{\nabla}\dot{\phi} - \dot{\phi}\bm{\nabla}\phi\right) + \rho\mathbf{u}\phi \label{eq:jiang_flux}\\
+  \frac{\partial (\rho E_G)}{\partial t} &= -\bm{\nabla} \cdot \left(\rho\mathbf{u}E_G + p\mathbf{u} + \mathbf{F_g}\right)
+\end{align}
+This is analytically equivalent to the approach by \citet{marcello:2012}, which can be seen by taking the divergence of the flux and using the Poisson equation.
+It would be justified to use this form simply because it is cleaner to write down, but there is a more important benefit to using an explicit flux based formula: in AMR, conservation between levels can only happen if all fluid movements are written down using fluxes, because it is those fluxes that are adjusted at coarse-fine interfaces to ensure conservation.
+In this form we can simply add the gravitational source term to the hydrodynamical fluxes and use the existing AMR algorithm to guarantee conservation between levels.
+We note that \citeauthor{jiang:2013} say that this method requires the solution of an additional Poisson equation to obtain the intermediate time gravitational potential, but as we pointed out above, this can be done to second order simply by taking a centered difference using the existing Poisson solves, so no extra Poisson solve is necessary.
+
+This flux-based formalism has a significant drawback for our problem: unlike \citeauthor{jiang:2013}, we use open hydrodynamic boundary conditions, not periodic boundary conditions.
+Mass, momentum and energy can be lost (or gained) across the domain boundary.
+This is not a serious source of error in the standard cell-centered discretization, because the source term is proportional to $\rho$ and the density near the domain boundary is generally filled with a very low density ambient medium, so the size of the source term is always small.
+However, the gravitational flux in Equation \ref{eq:jiang_flux} has terms that are not explicitly proportional to density.
+Consequently, the flux lost across the domain boundaries can be significant (though this effect diminishes with increasing resolution).
+A related problem is that at the edge of a star, the gravitational flux can be so large that the gas energy becomes negative after an update.
+This effect is greatly diminished if the flux is instead proportional to density, as in the standard method.
+This is not a major problem for ideal or polytropic gases in pure hydrodynamics problems, but in our simulations we want the internal energy (and thus the temperature) to always be reliable so that nuclear reactions occur at the correct rates.
+This is especially sensitive at the edges of the stars, where the burning is expected to occur in a violent merger scneario.
+
+A possible compromise approach is discussed in \cite{arepo}.
+We start with the integral formulation of the energy source term (Equation \ref{eq:cell_center_gravity_source}), but instead of discretizing at the cell center, we use Gauss' theorem to evaluate the source term at the zone boundaries (see Equations 91--96 of that paper; we exclude the term related to the motion of the zones since our grids are fixed).
+This results in a gravitational update for zone $i$ written as
+\begin{equation}
+  \Delta E^{\text{grav}}_{i} = -\frac{1}{2}\sum_{j} \Delta m_{ij}\, \mathbf{r}_{ij} \cdot \bm{\nabla}\phi_i \approx -\frac{1}{2}\sum_{j} \Delta m_{ij}\left(\phi_i - \phi_j\right),
+\end{equation}
+where the sum is over all $j$ zones that share a boundary with zone $i$.
+Here $\mathbf{r}_{ij}$ is the radius vector joining zones $i$ and $j$ and $\Delta m_{ij}$ is the mass exchanged between zones $i$ and $j$ in the hydrodynamics update, and is equal to $\Delta t A_{ij} F_\rho^{ij}$, where $A_{ij}$ is the area of the zone boundary and $F_{\rho}^{ij}$ is the mass flux from the mass continuity equation.
+It is straightforward to see that the second version is explicitly conservative.
+The benefit of this form is that the mass exchanged between zones near the domain boundary will always be small in absolute terms since these zones are at very low density.
+Thus the energy lost through the boundary will be of comparable size to the standard method, but the gravity update should be more accurate since the change in gravitational energy now tracks the explicit motion of mass on the grid.
+
+% Now rewrite this so that it is in flux form
 

### 2015-01-07 job_scripts/run_utils.sh
Have the run script create a plots directory for storing analysis output.

@@ -145,3 +145,10 @@ if [ ! -d $results_dir ]; then
   mkdir $results_dir
 fi
 
+# Create the plots directory, for saving output
+
+plots_dir="plots"
+
+if [ ! -d $plots_dir ]; then
+    mkdir $plots_dir
+fi

### 2015-01-07 papers/wdmerger_I/paper.tex

@@
 This equation is a function of $\rho_{i+1}$ only since the pressure is uniquely determined by the density in this case.
-$\rho_i and p_i$ are known,  $g_{i+1/2}$ is the gravitational acceleration at the interface between zones $i$ and $i+1$, found by simply adding up all the mass from zones $1, \ldots, i$ to get the enclosed mass, $M_{i+1/2}$, and then setting $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$.
+$\rho_i$ and $p_i$ are known,  $g_{i+1/2}$ is the gravitational acceleration at the interface between zones $i$ and $i+1$, found by simply adding up all the mass from zones $1, \ldots, i$ to get the enclosed mass, $M_{i+1/2}$, and then setting $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$.
 We solve this equation for $\rho_{i+1}$ using a Newton-Raphson iteration.

### 2015-01-08 papers/wdmerger_I/paper.tex
Energy density -> specific energy

@@
 \end{align}
-Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the total energy density where $e$ is the internal (thermal) energy density.
+Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the total specific energy, where $e$ is the internal (thermal) specific energy.
 

### 2015-01-08 papers/wdmerger_I/paper.tex
Initial attempt at an author list
Make sure that all equations have numbers.

@@
 
-\author{TBD}
+\author{Max P. Katz, Michael Zingale, Alan C. Calder, F. Douglas Swesty, Ann S. Almgren, Weiqun Zhang}
 %==========================================================================
@@
   t(r) = \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\,\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right]. \label{analyticalFreeFall}
-\end{equation}\MarginPar{This result is derived in the freefall directory.}
+\end{equation}
 This result can be derived by recognizing that
-\[
+\begin{equation}
   t(r) = \int_{r_0}^r \frac{dr}{v(r)}
-\]
+\end{equation}
 and inserting the velocity as a function of radial separation,
-\[
+\begin{equation}
   v(r) = \sqrt{\frac{2GM}{r_0}\left(\frac{r_0}{r} - 1\right)}.
-\]
+\end{equation}
 We determine the initial separation to be consistent with the
@@
 orbital period of $T = 100$ s for this simulation, so that the separation is 
-\[
-  r_0 = 3.61 \times 10^{9}\ \text{cm}.
-\]
+$r_0 = 3.61 \times 10^{9}\ \text{cm}$.
 We chose a relatively low resolution simulation to demonstrate the
@@
 the total elapsed free-fall time,
-\[
+\begin{equation}
   t_{\text{ff}} = \frac{\pi}{2} \sqrt{\frac{r_0^3}{2GM}} = \frac{T}{4\sqrt{2}}.
-\]
+\end{equation}
 The physical radius of each white dwarf is roughly $10\%$ of the

### 2015-01-15 papers/wdmerger_I/paper.tex
a little appendix work

@@
 
-We use the unsplit piecewise-parabolic method (PPM) solver in CASTRO to advance the hydrodynamics system in time \citep{ppmunsplit}.
-A number of changes were made to the solver, which are detailed in the Appendix.
-These changes bring the algorithm more in line with that of \cite{ppm}.
-CASTRO as originally released featured a slightly modified version of the higher resolution limiters of \cite{colella_sekora:2008}, which can be accessed in the code using \texttt{castro.ppm\_type = 2}.
-The advantage of this limiter is that it preserves physical extrema rather than clipping them off as in the original approach of \cite{ppm}.
-However, we found these limiters to be unsatisfactory for our problem.
-There are many regions in our problem with large density gradients (such as the interface between the star's atmosphere and the ambient gas outside of it) and in these regions the algorithm can yield negative densities.
-This often results from the limiters interpreting these gradients as being true minima.
-As a result, we use the original limiter, which is strictly monotonicity preserving in the parabolic profiles it generates; this is activated with \texttt{castro.ppm\_type = 1}.
+We use the unsplit piecewise-parabolic method (PPM) solver in CASTRO
+to advance the hydrodynamics system in time \citep{ppmunsplit}.  A
+number of changes were made to the solver, which are detailed in the
+Appendix.  These changes bring the algorithm more in line with that of
+\cite{ppm}. CASTRO as originally released featured a slightly modified
+version of the higher resolution limiters of
+\cite{colella_sekora:2008}, which can be accessed in the code using
+\texttt{castro.ppm\_type = 2}. The advantage of this limiter is that
+it preserves physical extrema rather than clipping them off as in the
+original approach of \cite{ppm}. However, we found these limiters to
+be unsatisfactory for our problem. There are many regions in our
+problem with large density gradients (such as the interface between
+the star's atmosphere and the ambient gas outside of it) and in these
+regions the algorithm can yield negative densities. This often results
+from the limiters interpreting these gradients as being true
+minima. As a result, we use the original limiter, which is strictly
+monotonicity preserving in the parabolic profiles it generates; this
+is activated with \texttt{castro.ppm\_type = 1}.
 
@@
 
+The basic PPM algorithm in CASTRO has undergone a number of changes since
+the original code paper \citep{castro}.  A discussion of the pure
+hydrodynamics changes was given in \citet{zingalekatz:2015}.  Here
+we discuss the changes that affect multispecies flow and source
+terms.
+
 Summary of changes:
 \begin{itemize}
-\item New reference state/flattening fix: helps energy conservation a
-  lot, also fixes some undershoot/overshoots
-
 \item New gravity update type: fixes some energy conservation
-
-\item temperature-based PPM: fixes temperature floors
 

### 2015-01-15 papers/wdmerger_I/paper.tex
no changes execpt for line wrapping

@@
 
-Type Ia supernovae (SNe Ia) are currently some of the most exciting events to study in astrophysics.
-These bright, brief pulses of light in the distant universe have led to a number of important discoveries in recent years, including the discovery of the accelerated expansion of the universe \citep{perlmutter1999,riess1998}.
-However, their origin is shrouded in mystery.
-It has long been expected that these events arise from the thermonuclear explosions of white dwarfs \citep{hoyle_fowler:1960}, but the cause of these explosions is uncertain.
-In particular, it is not clear what process causes the temperatures in these white dwarfs to become hot enough for explosive burning of its constituent nuclei.
-The model favored initially by the community was the so-called single-degenerate (SD) model \citep{whelan_iben:1973}.
-Accretion of material from a companion star such as a red giant would cause the star to approach the Chandrasekhar mass, and in doing so the temperature and density in the center would become sufficient for thermonuclear fusion to proceed.
-However, in recent years many alternative progenitor models have been discussed.
-A leading candidate for explaining some or most of these explosions is the double-degenerate (DD) model, in which two white dwarfs merge and the merged object reaches the conditions necessary for a thermonuclear ignition \citep{ibentutukov:1984,webbink:1984}.
-Another is the double detonation scenario, where accretion of material onto a sub-Chandrasekhar white dwarf would lead to a detonation inside the accreted envelope, and this would send a compressional wave into the core of the star that would trigger a secondary detonation.
-A recent review of the progenitor models can be found in \citet{hillebrandt:2013}.
-
-There are several observational reasons why double-degenerate systems are a promising progenitor system for at least a substantial fraction of normal SNe Ia.
-No conclusive evidence exists for a surviving companion star of a SN Ia; this is naturally explained by the DD model because both WDs are likely to be destroyed in the merger process.
-Similarly, pre-explosion images of the SN Ia systems have never clearly turned up a companion star, and in some cases a large fraction of the parameter space for the nature of the companion star is excluded.
-Additionally, not enough progenitor systems are seen for the SD case to match the observed local SN Ia rate, whereas the number of white dwarf binaries may be sufficient to account for this rate.
-Finally, the DD model can naturally explain the fact that many SNe Ia are observed to occur at very long delay times after the stars were formed, since the progenitor systems only become active once both stars have evolved off the main sequence.
-A thorough review of the observational evidence about SNe Ia and further discussion of these ideas can be found in \cite{maoz:2014}.
-
-The history of double degenerate theory can broadly be described as a series of three major paradigm shifts.
-The first attempts to model the results of the merger process came in the 1980s.
-\cite{nomotoiben:1985} demonstrated that off-center carbon ignition would occur in the more massive white dwarf as it accreted mass near the Eddington rate from the less massive white dwarf overflowing its Roche lobe.
-\cite{saionomoto:1985} tracked the evolution of the flame and found that it propagated quiescently into the center, converting the carbon-oxygen white dwarf into an oxygen-neon-magnesium white dwarf.
-This would then be followed by collapse into a neutron star -- a result with significantly different observational properties compared to a SN Ia.
-This scenario, termed accretion-induced collapse, would be avoided only if the accretion rate were well below the Eddington rate.
-\cite{tutukov_yungelson:1979} observed that this could happen if the mass loss from the secondary was higher than the Eddington rate and thus the accreted material formed an accretion disc, which might rain down on the primary more slowly.
-The main finding was that double degenerate systems would not obviously lead to Type Ia supernovae.
-
-The first three-dimensional simulations of double degenerate systems were performed by \citet{benz:1990}, who used the smoothed particle hydrodynamics (SPH) method to simulate the merger process.
-They found that if the lower-mass star (generally called the ``secondary'') was close enough to the more massive star (the ``primary'') to begin mass transfer on a dynamical time scale, the secondary completely disrupted and formed a hot envelope around the primary, with a centrifugally-supported accretion disk surrounding the core and envelope.
-They noted that the envelope temperature was hot enough to generate carbon fusion but neglected nuclear reactions in their simulation.
-Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009,raskin:2012} corroborated this result.
-This validation of the prediction of \cite{tutukov_yungelson:1979} did not instill confidence in the community.
-\cite{mochkovitch_livio:1990} and \cite{livio:2000} observed that turbulent viscosity would be sufficiently large for angular momentum to be removed from the disk at a rate high enough to generate the troublesome accretion timescales.
-Based on all of this evidence, the review of \cite{hillebrandtniemeyer2000} argued that the model was only viable if the accretion-induced collapse problem could be avoided.
-Later work by \cite{shen:2012} and \cite{schwab:2012} used a more detailed treatment of the viscous transport in the outer regions of the remnant and found that while the centrifugally supported envelope would be converted into hot envelope material on a viscous timescale, their simulations still led to off-center carbon burning.
-\cite{vankerkwijk:2010} argued that equal-mass mergers would lead to the conditions necessary for carbon detonation in the center of the merged object, but \cite{shen:2012} questioned this for similar reasons related to how viscous transport would convert rotational motion into pressure support.
-\cite{zhu:2013} followed this with an expanded parameter space study and argued that many of their carbon-oxygen systems had the potential to detonate.
-The study of the long-term evolution of the remnants is thus still an open subject of research.
-
-The most recent shift in perspective on this problem started with two series of papers appearing at roughly the same time.
-One began with \cite{pakmor:2010} and \cite{pakmor:2011}.
-This group used the SPH method to study the merger of equal-mass ($0.9\ M_\odot$) carbon-oxygen white dwarfs and found that a hotspot was generated near the surface of the primary white dwarf.
-They argued that this region had a temperature and density sufficient to trigger a thermonuclear detonation.
-They propagated this detonation throughout the system and found that it would observationally appear as a subluminous Type Ia supernova.
-This was the first time a DD simulation successfully reproduced at least some characteristics of a SN Ia.
-\cite{pakmor:2011} tried a few different mass combinations and found empirically that this would hold as long as the secondary was at least 80\% as massive as the primary.
-These events, where the merger process resulted in the detonation of the system during the merger coalescence -- avoiding the much longer time-scale evolution -- were termed ``violent'' mergers.
-
-Around the same time, however, \cite{guillochon:2010} and \cite{dan:2011} came to significantly different conclusions.
-They pointed out that the previously mentioned simulations generally all shared a significant drawback, which was that their initial conditions were not carefully constructed.
-\cite{motl:2002}, \cite{dsouza:2006}, and \cite{motl:2007} (the first three-dimensional grid-based simulations of mass transfer in binary white dwarf systems) pioneered the study of looking at the long-term dynamical evolution of binary white dwarf systems after constructing exact equilibrium initial conditions.
-In contrast to the other studies, they found that paying close attention to the accuracy of the initial conditions of the simulation has important effects.
-In particular, earlier work placed the stars too close together and ignored the effects of tidal forces on changing the shape of the secondary, leading to the merger happening artificially too quickly.
-When the initial conditions are constructed in exact equilibrium, the system can be stable for tens of orbital periods, substantially changing the character of the mass transfer phase.
-However, one limitation of this series of studies is that the authors used a polytropic equation of state and thus could not consider nuclear reactions.
-\cite{guillochon:2010} and \cite{dan:2011} improved on this using a realistic equation of state, a nuclear reaction network, and a similar approach to the equilibrium initial conditions, and found substantial agreement with the idea that mass transfer occurs in a stable manner over tens of orbital periods.
-They also found that, assuming the material accreted onto the surface of the primary was primarily helium, explosive surface detonations would occur as a result of accretion stream instabilities during the mass transfer phase prior to the full merger.
-This could trigger a double-detonation explosion and thus perhaps a SN Ia.
-
-The very latest developments give some areas of agreement and some of remaining uncertainty.
-\cite{pakmor:2012} performed a merger scenario with a $1.1\ M_\odot$ and $0.9\ M_\odot$ setup, with better treatment of the initial conditions, and also found that the merger process happened over more than ten orbits.
-Nevertheless, they still found that a carbon-oxygen detonation would occur, in line with their earlier results.
-\cite{moll:2014} was also able to find a detonation in a similarly massive system.
-\cite{dan:2012} and \cite{dan:2014} performed a large sweep of the parameter space for merger pairs and found that pure carbon-oxygen systems would generally not lead to detonations and violent mergers except for the most massive systems.
-They did find that for systems containing helium, many of them would detonate and potentially lead to SNe Ia, either through the aforementioned instabilities in the accretion stream, or during the contact phase, similar to the violent carbon-oxygen mergers.
-\cite{pakmor:2013} added a thin helium shell on their primary white dwarf, and found that this robustly led to a detonation of the white dwarf.
-Thus at least a preliminary agreement may be that systems containing helium could robustly lead to events resembling SNe Ia, as well as very massive carbon-oxygen binaries.
-
-Given the above, why is another approach using a different simulation code warranted?
-First and foremost, reproducibility of the results across simulation codes and algorithms is important for gauging confidence in this result.
-With only a couple of exceptions, most of the existing results that study the viability of the DD progenitor as an explanation for Type Ia supernovae (that is, including a temperature-dependent equation of state and nuclear reactions) have used the SPH method.
-SPH codes have a number of features which do aid them in the study of these systems, such as excellent conservation of angular momentum.
-However, the question of whether a prompt detonation in a merger actually happens depends in detail on the nature of the gas at the interface between the two stars, which is at much lower density than the rest of the stellar material.
-The SPH codes use uniform mass particles, so their effective resolution is \textit{lowest} at the stellar surface.
-In contrast, a grid-based code with adaptive mesh refinement can zoom in on the regions where hotspots will develop, while also maintaining high enough resolution in the high-density regions to adequately capture the large-scale mass transfer dynamics.
-However, there are also outstanding questions of convergence in SPH (e.g.
-\cite{zhu_SPH:2014}) and whether the method correctly captures fluid instabilities.
-This is an important question for white dwarf mergers because of the likely importance small-scale instabilities will have on the evolution of the low-density gas at the primary's surface.
-The pioneering work of \cite{agertz:2007} compared grid and SPH codes and found some important differences.
-Most relevant for this discussion is that the SPH codes could not adequately handle mixing from the Kelvin-Helmholtz instability in the test they propose.
-As pointed out by \cite{price:2008}, this is not a result of SPH being inherently unable to model this instability, but instead it is attributed to the fact that the standard SPH evolution equations do not have a mechanism for capturing discontinuities in internal energy.
-\citeauthor{price:2008} showed that the addition of an artificial thermal conductivity can dramatically improve the ability of the SPH codes to exhibit this instability.
-There have since been a number of other papers discussing this issue, but to our knowledge none of these improvements have yet been incorporated into an SPH model of a WD merger.
-Another reason to be skeptical is that the detonation itself was inserted manually in \cite{pakmor:2010} based on small-scale detonation simulations.
-A self-consistent evolution of a thermonuclear detonation, if it existed, would significantly build confidence in the progenitor model.
-
-This is the first in a series of papers designed to address these outstanding theoretical issues for white dwarf mergers.
-This work will discuss the verification of our hydrodynamics code for simulating these events.
-Later efforts will look at the initial conditions of the system, the robustness with which a hotspot is found from which a detonation could occur, and the importance of the initial white dwarf models, which should be more sophisticated than simple carbon-oxygen mixtures and in principle should use results from modern stellar evolution calculations.
-Section \ref{sec:Numerical Methodology} describes our code and why it can provide useful results compared to other methodologies used for this problem.
-Section \ref{sec:initial_models} describes the method we use for setting up a binary white dwarf simulation.
-Section \ref{sec:Tests} discusses a few test problems that we use to demonstrate that our code accurately solves the equations of fluid dynamics.
-Section \ref{sec:Performance} demonstrates that the software scales well for supercomputer applications.
-Finally, Section \ref{sec:Conclusions and Discussion} recaps what we have shown and highlights some of the future work we plan to do.
+Type Ia supernovae (SNe Ia) are currently some of the most exciting
+events to study in astrophysics. These bright, brief pulses of light
+in the distant universe have led to a number of important discoveries
+in recent years, including the discovery of the accelerated expansion
+of the universe \citep{perlmutter1999,riess1998}. However, their
+origin is shrouded in mystery. It has long been expected that these
+events arise from the thermonuclear explosions of white dwarfs
+\citep{hoyle_fowler:1960}, but the cause of these explosions is
+uncertain. In particular, it is not clear what process causes the
+temperatures in these white dwarfs to become hot enough for explosive
+burning of its constituent nuclei. The model favored initially by the
+community was the so-called single-degenerate (SD) model
+\citep{whelan_iben:1973}. Accretion of material from a companion star
+such as a red giant would cause the star to approach the Chandrasekhar
+mass, and in doing so the temperature and density in the center would
+become sufficient for thermonuclear fusion to proceed. However, in
+recent years many alternative progenitor models have been discussed. A
+leading candidate for explaining some or most of these explosions is
+the double-degenerate (DD) model, in which two white dwarfs merge and
+the merged object reaches the conditions necessary for a thermonuclear
+ignition \citep{ibentutukov:1984,webbink:1984}. Another is the double
+detonation scenario, where accretion of material onto a
+sub-Chandrasekhar white dwarf would lead to a detonation inside the
+accreted envelope, and this would send a compressional wave into the
+core of the star that would trigger a secondary detonation. A recent
+review of the progenitor models can be found in
+\citet{hillebrandt:2013}.
+
+There are several observational reasons why double-degenerate systems
+are a promising progenitor system for at least a substantial fraction
+of normal SNe Ia. No conclusive evidence exists for a surviving
+companion star of a SN Ia; this is naturally explained by the DD model
+because both WDs are likely to be destroyed in the merger
+process. Similarly, pre-explosion images of the SN Ia systems have
+never clearly turned up a companion star, and in some cases a large
+fraction of the parameter space for the nature of the companion star
+is excluded. Additionally, not enough progenitor systems are seen for
+the SD case to match the observed local SN Ia rate, whereas the number
+of white dwarf binaries may be sufficient to account for this
+rate. Finally, the DD model can naturally explain the fact that many
+SNe Ia are observed to occur at very long delay times after the stars
+were formed, since the progenitor systems only become active once both
+stars have evolved off the main sequence. A thorough review of the
+observational evidence about SNe Ia and further discussion of these
+ideas can be found in \cite{maoz:2014}.
+
+The history of double degenerate theory can broadly be described as a
+series of three major paradigm shifts. The first attempts to model the
+results of the merger process came in the
+1980s. \cite{nomotoiben:1985} demonstrated that off-center carbon
+ignition would occur in the more massive white dwarf as it accreted
+mass near the Eddington rate from the less massive white dwarf
+overflowing its Roche lobe. \cite{saionomoto:1985} tracked the
+evolution of the flame and found that it propagated quiescently into
+the center, converting the carbon-oxygen white dwarf into an
+oxygen-neon-magnesium white dwarf. This would then be followed by
+collapse into a neutron star -- a result with significantly different
+observational properties compared to a SN Ia. This scenario, termed
+accretion-induced collapse, would be avoided only if the accretion
+rate were well below the Eddington rate. \cite{tutukov_yungelson:1979}
+observed that this could happen if the mass loss from the secondary
+was higher than the Eddington rate and thus the accreted material
+formed an accretion disc, which might rain down on the primary more
+slowly. The main finding was that double degenerate systems would not
+obviously lead to Type Ia supernovae.
+
+The first three-dimensional simulations of double degenerate systems
+were performed by \citet{benz:1990}, who used the smoothed particle
+hydrodynamics (SPH) method to simulate the merger process. They found
+that if the lower-mass star (generally called the ``secondary'') was
+close enough to the more massive star (the ``primary'') to begin mass
+transfer on a dynamical time scale, the secondary completely disrupted
+and formed a hot envelope around the primary, with a
+centrifugally-supported accretion disk surrounding the core and
+envelope. They noted that the envelope temperature was hot enough to
+generate carbon fusion but neglected nuclear reactions in their
+simulation. Later simulations of similar setups
+\citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009,raskin:2012}
+corroborated this result. This validation of the prediction of
+\cite{tutukov_yungelson:1979} did not instill confidence in the
+community. \cite{mochkovitch_livio:1990} and \cite{livio:2000}
+observed that turbulent viscosity would be sufficiently large for
+angular momentum to be removed from the disk at a rate high enough to
+generate the troublesome accretion timescales. Based on all of this
+evidence, the review of \cite{hillebrandtniemeyer2000} argued that the
+model was only viable if the accretion-induced collapse problem could
+be avoided. Later work by \cite{shen:2012} and \cite{schwab:2012} used
+a more detailed treatment of the viscous transport in the outer
+regions of the remnant and found that while the centrifugally
+supported envelope would be converted into hot envelope material on a
+viscous timescale, their simulations still led to off-center carbon
+burning. \cite{vankerkwijk:2010} argued that equal-mass mergers would
+lead to the conditions necessary for carbon detonation in the center
+of the merged object, but \cite{shen:2012} questioned this for similar
+reasons related to how viscous transport would convert rotational
+motion into pressure support. \cite{zhu:2013} followed this with an
+expanded parameter space study and argued that many of their
+carbon-oxygen systems had the potential to detonate. The study of the
+long-term evolution of the remnants is thus still an open subject of
+research.
+
+The most recent shift in perspective on this problem started with two
+series of papers appearing at roughly the same time. One began with
+\cite{pakmor:2010} and \cite{pakmor:2011}. This group used the SPH
+method to study the merger of equal-mass ($0.9\ M_\odot$)
+carbon-oxygen white dwarfs and found that a hotspot was generated near
+the surface of the primary white dwarf. They argued that this region
+had a temperature and density sufficient to trigger a thermonuclear
+detonation. They propagated this detonation throughout the system and
+found that it would observationally appear as a subluminous Type Ia
+supernova. This was the first time a DD simulation successfully
+reproduced at least some characteristics of a SN
+Ia. \cite{pakmor:2011} tried a few different mass combinations and
+found empirically that this would hold as long as the secondary was at
+least 80\% as massive as the primary. These events, where the merger
+process resulted in the detonation of the system during the merger
+coalescence -- avoiding the much longer time-scale evolution -- were
+termed ``violent'' mergers.
+
+Around the same time, however, \cite{guillochon:2010} and
+\cite{dan:2011} came to significantly different conclusions. They
+pointed out that the previously mentioned simulations generally all
+shared a significant drawback, which was that their initial conditions
+were not carefully constructed. \cite{motl:2002}, \cite{dsouza:2006},
+and \cite{motl:2007} (the first three-dimensional grid-based
+simulations of mass transfer in binary white dwarf systems) pioneered
+the study of looking at the long-term dynamical evolution of binary
+white dwarf systems after constructing exact equilibrium initial
+conditions. In contrast to the other studies, they found that paying
+close attention to the accuracy of the initial conditions of the
+simulation has important effects. In particular, earlier work placed
+the stars too close together and ignored the effects of tidal forces
+on changing the shape of the secondary, leading to the merger
+happening artificially too quickly. When the initial conditions are
+constructed in exact equilibrium, the system can be stable for tens of
+orbital periods, substantially changing the character of the mass
+transfer phase. However, one limitation of this series of studies is
+that the authors used a polytropic equation of state and thus could
+not consider nuclear reactions. \cite{guillochon:2010} and
+\cite{dan:2011} improved on this using a realistic equation of state,
+a nuclear reaction network, and a similar approach to the equilibrium
+initial conditions, and found substantial agreement with the idea that
+mass transfer occurs in a stable manner over tens of orbital
+periods. They also found that, assuming the material accreted onto the
+surface of the primary was primarily helium, explosive surface
+detonations would occur as a result of accretion stream instabilities
+during the mass transfer phase prior to the full merger. This could
+trigger a double-detonation explosion and thus perhaps a SN Ia.
+
+The very latest developments give some areas of agreement and some of
+remaining uncertainty. \cite{pakmor:2012} performed a merger scenario
+with a $1.1\ M_\odot$ and $0.9\ M_\odot$ setup, with better treatment
+of the initial conditions, and also found that the merger process
+happened over more than ten orbits. Nevertheless, they still found
+that a carbon-oxygen detonation would occur, in line with their
+earlier results. \cite{moll:2014} was also able to find a detonation
+in a similarly massive system. \cite{dan:2012} and \cite{dan:2014}
+performed a large sweep of the parameter space for merger pairs and
+found that pure carbon-oxygen systems would generally not lead to
+detonations and violent mergers except for the most massive
+systems. They did find that for systems containing helium, many of
+them would detonate and potentially lead to SNe Ia, either through the
+aforementioned instabilities in the accretion stream, or during the
+contact phase, similar to the violent carbon-oxygen
+mergers. \cite{pakmor:2013} added a thin helium shell on their primary
+white dwarf, and found that this robustly led to a detonation of the
+white dwarf. Thus at least a preliminary agreement may be that systems
+containing helium could robustly lead to events resembling SNe Ia, as
+well as very massive carbon-oxygen binaries.
+
+Given the above, why is another approach using a different simulation
+code warranted? First and foremost, reproducibility of the results
+across simulation codes and algorithms is important for gauging
+confidence in this result. With only a couple of exceptions, most of
+the existing results that study the viability of the DD progenitor as
+an explanation for Type Ia supernovae (that is, including a
+temperature-dependent equation of state and nuclear reactions) have
+used the SPH method. SPH codes have a number of features which do aid
+them in the study of these systems, such as excellent conservation of
+angular momentum. However, the question of whether a prompt detonation
+in a merger actually happens depends in detail on the nature of the
+gas at the interface between the two stars, which is at much lower
+density than the rest of the stellar material. The SPH codes use
+uniform mass particles, so their effective resolution is
+\textit{lowest} at the stellar surface. In contrast, a grid-based code
+with adaptive mesh refinement can zoom in on the regions where
+hotspots will develop, while also maintaining high enough resolution
+in the high-density regions to adequately capture the large-scale mass
+transfer dynamics. However, there are also outstanding questions of
+convergence in SPH (e.g. \cite{zhu_SPH:2014}) and whether the method
+correctly captures fluid instabilities. This is an important question
+for white dwarf mergers because of the likely importance small-scale
+instabilities will have on the evolution of the low-density gas at the
+primary's surface. The pioneering work of \cite{agertz:2007} compared
+grid and SPH codes and found some important differences. Most relevant
+for this discussion is that the SPH codes could not adequately handle
+mixing from the Kelvin-Helmholtz instability in the test they
+propose. As pointed out by \cite{price:2008}, this is not a result of
+SPH being inherently unable to model this instability, but instead it
+is attributed to the fact that the standard SPH evolution equations do
+not have a mechanism for capturing discontinuities in internal
+energy. \citeauthor{price:2008} showed that the addition of an
+artificial thermal conductivity can dramatically improve the ability
+of the SPH codes to exhibit this instability. There have since been a
+number of other papers discussing this issue, but to our knowledge
+none of these improvements have yet been incorporated into an SPH
+model of a WD merger. Another reason to be skeptical is that the
+detonation itself was inserted manually in \cite{pakmor:2010} based on
+small-scale detonation simulations. A self-consistent evolution of a
+thermonuclear detonation, if it existed, would significantly build
+confidence in the progenitor model.
+
+This is the first in a series of papers designed to address these
+outstanding theoretical issues for white dwarf mergers. This work will
+discuss the verification of our hydrodynamics code for simulating
+these events. Later efforts will look at the initial conditions of the
+system, the robustness with which a hotspot is found from which a
+detonation could occur, and the importance of the initial white dwarf
+models, which should be more sophisticated than simple carbon-oxygen
+mixtures and in principle should use results from modern stellar
+evolution calculations. Section \ref{sec:Numerical Methodology}
+describes our code and why it can provide useful results compared to
+other methodologies used for this problem. Section
+\ref{sec:initial_models} describes the method we use for setting up a
+binary white dwarf simulation. Section \ref{sec:Tests} discusses a few
+test problems that we use to demonstrate that our code accurately
+solves the equations of fluid dynamics. Section \ref{sec:Performance}
+demonstrates that the software scales well for supercomputer
+applications. Finally, Section \ref{sec:Conclusions and Discussion}
+recaps what we have shown and highlights some of the future work we
+plan to do.
 
@@
 
-To study the white dwarf merger problem, we use the grid-based hydrodynamics code CASTRO \citep{castro}.
-CASTRO solves the Euler equations, along with the inclusion of optional modules for gravity, nuclear reactions and thermodynamics.
-CASTRO is based on the BoxLib adaptive-mesh refinement (AMR) framework \citep{rendleman:2000}, which represents fluid data on a mesh where regions of interest have higher spatial resolution.
-CASTRO is highly parallel and is designed for large-scale use on modern supercomputers; see Section \ref{sec:Performance} for information on how CASTRO performs for our problem.
-The next few subsections describe our approach to each of the physics components used in this work.
-We direct the reader to the original code paper for a full description of CASTRO's approach to solving the equations of hydrodynamics.
-In this work, we report mainly on the changes we have made to the code since its original release, for the purpose of approaching this problem.
+To study the white dwarf merger problem, we use the grid-based
+hydrodynamics code CASTRO \citep{castro}. CASTRO solves the Euler
+equations, along with the inclusion of optional modules for gravity,
+nuclear reactions and thermodynamics. CASTRO is based on the BoxLib
+adaptive-mesh refinement (AMR) framework \citep{rendleman:2000}, which
+represents fluid data on a mesh where regions of interest have higher
+spatial resolution. CASTRO is highly parallel and is designed for
+large-scale use on modern supercomputers; see Section
+\ref{sec:Performance} for information on how CASTRO performs for our
+problem. The next few subsections describe our approach to each of the
+physics components used in this work. We direct the reader to the
+original code paper for a full description of CASTRO's approach to
+solving the equations of hydrodynamics. In this work, we report mainly
+on the changes we have made to the code since its original release,
+for the purpose of approaching this problem.
 
@@
 \end{align}
-Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the total specific energy, where $e$ is the internal (thermal) specific energy.
+Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity
+vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the
+total specific energy, where $e$ is the internal (thermal) specific
+energy.
 
@@
 
-A related issue that required a code improvement was that in cases of large density gradients such as the edge of a star, it is possible to generate negative densities in zones even with the more strongly limited PPM.
-This can occur if a region of large density is moving away from an ambient zone at relatively large speeds; then the net density flux in the ambient zones can be large enough to unphysically drag the density below zero.
-In practice, this will occur at the trailing edge of a star that is moving across a grid.
-In such a situation, there are two main approaches one could take: either explicitly introduce a positivity-guaranteeing diffusive flux, or reset the characteristics of the affected zone.
-We choose the latter approach.
-Even though it is non-conservative, it preserves a characteristic we value, which is to keep the edge of the stars relatively sharp, as they physically should be.
-Since the mass of the affected zones is typically already fairly low, this should not seriously affect the energy conservation properties of our simulation.
-Similarly, it is also possible in this physical situation for very large velocities to be generated in the material behind the star.
-The momentum transferred to the ambient zone is close to the momentum of the stellar material, but this is of much higher density and so the velocity of the ambient material must become very large to compensate.
-While this may be physically viable in certain circumstances, it poses a problem for our code, which uses subcycling in time (see below).
-The timestep is fixed at the beginning of the set of subcycled timesteps, and if in the middle of the cycle a velocity is generated that is so large that it violates the CFL criterion for this timestep, we cannot guarantee the stability of the resulting solution since we cannot alter the timestep.
-(Nor would we want to, because in some cases because the velocity becomes spuriously large and would drag down the global timestep.) Therefore we also insert a check at the end of every hydrodynamics timestep that resets the velocity and/or sound-speed (indirectly, using the temperature) of any zone that violates the CFL criterion given the timestep we chose at the beginning of the cycle.
-In practice, we find that this is mainly necessary only for simulations in the inertial reference frame of the white dwarf merger, which we do not normally do for the reasons stated in Section \ref{sec:kepler}.
-
-CASTRO's approach to adaptive mesh refinement, based on its underlying BoxLib framework, is to refine zones based on certain user-specified criteria that tag regions of interest for higher spatial resolution.
-Data is represented on one of a number of AMR levels, where each level corresponds to a set of zones at the same resolution, which covers a subset of the domain covered by the level immediately below it.
-We typically call the level 0 grid the \textit{coarse} grid, which has the lowest spatial resolution.
-Each finer, higher-level grid has a higher resolution than the grid below it by some integer factor $N$, which is restricted to be $N = 2\ \text{or}\ 4$ in the code.
-The zones are strictly contained within the rectangular extent of the underlying coarser zones (the code is not restricted to representing only Cartesian geometries, but we use a Cartesian mesh with uniform spacing in each dimension for the present study).
-For the time evolution of the AMR system we use subcycling, where each AMR level is advanced at a different timestep and a correction step is applied at the end to synchronize the various levels.
-Generally the number of subcycled timesteps is equal to the jump in refinement between levels, so for example on a grid with three levels and two jumps of four in refinement, the level 2 zones will have 16 times higher spatial resolution than the coarse grid and there will be 16 level 2 timesteps per level 0 timestep.
-
-The boundary conditions on the hyperbolic system are simply zero-gradient zones that allow material to flow directly out of the domain.
-Using AMR, we make the coarse grid large enough that the boundaries are relatively far from the region of interest.
-This ensures that any boundary effects do not pollute the inner region where the stars will eventually make contact.
-We further make the restriction that refined grids cannot reach the domain boundary.
+A related issue that required a code improvement was that in cases of
+large density gradients such as the edge of a star, it is possible to
+generate negative densities in zones even with the more strongly
+limited PPM. This can occur if a region of large density is moving
+away from an ambient zone at relatively large speeds; then the net
+density flux in the ambient zones can be large enough to unphysically
+drag the density below zero. In practice, this will occur at the
+trailing edge of a star that is moving across a grid. In such a
+situation, there are two main approaches one could take: either
+explicitly introduce a positivity-guaranteeing diffusive flux, or
+reset the characteristics of the affected zone. We choose the latter
+approach. Even though it is non-conservative, it preserves a
+characteristic we value, which is to keep the edge of the stars
+relatively sharp, as they physically should be. Since the mass of the
+affected zones is typically already fairly low, this should not
+seriously affect the energy conservation properties of our
+simulation. Similarly, it is also possible in this physical situation
+for very large velocities to be generated in the material behind the
+star. The momentum transferred to the ambient zone is close to the
+momentum of the stellar material, but this is of much higher density
+and so the velocity of the ambient material must become very large to
+compensate. While this may be physically viable in certain
+circumstances, it poses a problem for our code, which uses subcycling
+in time (see below). The timestep is fixed at the beginning of the set
+of subcycled timesteps, and if in the middle of the cycle a velocity
+is generated that is so large that it violates the CFL criterion for
+this timestep, we cannot guarantee the stability of the resulting
+solution since we cannot alter the timestep. (Nor would we want to,
+because in some cases because the velocity becomes spuriously large
+and would drag down the global timestep.) Therefore we also insert a
+check at the end of every hydrodynamics timestep that resets the
+velocity and/or sound-speed (indirectly, using the temperature) of any
+zone that violates the CFL criterion given the timestep we chose at
+the beginning of the cycle. In practice, we find that this is mainly
+necessary only for simulations in the inertial reference frame of the
+white dwarf merger, which we do not normally do for the reasons stated
+in Section \ref{sec:kepler}.
+
+CASTRO's approach to adaptive mesh refinement, based on its underlying
+BoxLib framework, is to refine zones based on certain user-specified
+criteria that tag regions of interest for higher spatial
+resolution. Data is represented on one of a number of AMR levels,
+where each level corresponds to a set of zones at the same resolution,
+which covers a subset of the domain covered by the level immediately
+below it. We typically call the level 0 grid the \textit{coarse} grid,
+which has the lowest spatial resolution. Each finer, higher-level grid
+has a higher resolution than the grid below it by some integer factor
+$N$, which is restricted to be $N = 2\ \text{or}\ 4$ in the code. The
+zones are strictly contained within the rectangular extent of the
+underlying coarser zones (the code is not restricted to representing
+only Cartesian geometries, but we use a Cartesian mesh with uniform
+spacing in each dimension for the present study). For the time
+evolution of the AMR system we use subcycling, where each AMR level is
+advanced at a different timestep and a correction step is applied at
+the end to synchronize the various levels. Generally the number of
+subcycled timesteps is equal to the jump in refinement between levels,
+so for example on a grid with three levels and two jumps of four in
+refinement, the level 2 zones will have 16 times higher spatial
+resolution than the coarse grid and there will be 16 level 2 timesteps
+per level 0 timestep.
+
+The boundary conditions on the hyperbolic system are simply
+zero-gradient zones that allow material to flow directly out of the
+domain. Using AMR, we make the coarse grid large enough that the
+boundaries are relatively far from the region of interest. This
+ensures that any boundary effects do not pollute the inner region
+where the stars will eventually make contact.  We further make the
+restriction that refined grids cannot reach the domain boundary.
 
@@
 
-The equation of state (EOS) for our simulations is the Helmholtz EOS \citep{timmes_swesty:2000}.
-This models an electron-positron gas of arbitrary relativity and degeneracy over a wide range of temperatures and densities.
-Thermodynamic quantities are calculated as derivatives of the Helmholtz free energy, and the values are interpolated from a table.
-The natural variables of the Helmholtz free energy are temperature and density, and calling the EOS is simplest in this form.
-However, in hydrodynamics we often have the density, and internal energy as independent variables, and we want to obtain the temperature, pressure, and other quantities.
-To do this, we employ a Newton-Raphson iteration over the temperature (given some sufficient starting guess) until we find the temperature that corresponds to the desired internal energy.
-Sometimes this process fails to converge and the iterative value approaches zero.
-In these cases we employ a ``floor'' that limits how low the temperature can go (typically $10^4$ or $10^5$ K).
-There is a choice here how to proceed: we can either assign this floor value to the temperature and let that zone be thermodynamically inconsistent (the original behavior in Castro), or we can adjust the internal energy to be thermodynamically consistent with the temperature, at the cost of violating energy conservation.
-We have found in test problems of one-dimensional shocks that the latter yields more accurate results, so we employ the latter method.
+The equation of state (EOS) for our simulations is the Helmholtz EOS
+\citep{timmes_swesty:2000}. This models an electron-positron gas of
+arbitrary relativity and degeneracy over a wide range of temperatures
+and densities. Thermodynamic quantities are calculated as derivatives
+of the Helmholtz free energy, and the values are interpolated from a
+table. The natural variables of the Helmholtz free energy are
+temperature and density, and calling the EOS is simplest in this
+form. However, in hydrodynamics we often have the density, and
+internal energy as independent variables, and we want to obtain the
+temperature, pressure, and other quantities. To do this, we employ a
+Newton-Raphson iteration over the temperature (given some sufficient
+starting guess) until we find the temperature that corresponds to the
+desired internal energy. Sometimes this process fails to converge and
+the iterative value approaches zero. In these cases we employ a
+``floor'' that limits how low the temperature can go (typically $10^4$
+or $10^5$ K). There is a choice here how to proceed: we can either
+assign this floor value to the temperature and let that zone be
+thermodynamically inconsistent (the original behavior in Castro), or
+we can adjust the internal energy to be thermodynamically consistent
+with the temperature, at the cost of violating energy conservation. We
+have found in test problems of one-dimensional shocks that the latter
+yields more accurate results, so we employ the latter method.
 
@@
 \end{equation}
-where $\Phi$ is the gravitational potential, $G$ is the gravitational constant, and $\rho$ is the mass density.
-We note that the sign convention used in CASTRO is opposite to that commonly seen in the physics literature, so that $\Phi$ is positive everywhere.
-The solution of this equation in CASTRO is described in \cite{castro}, and consists of both level and composite solves, and a final synchronization at the end.
+where $\Phi$ is the gravitational potential, $G$ is the gravitational
+constant, and $\rho$ is the mass density. We note that the sign
+convention used in CASTRO is opposite to that commonly seen in the
+physics literature, so that $\Phi$ is positive everywhere. The
+solution of this equation in CASTRO is described in \cite{castro}, and
+consists of both level and composite solves, and a final
+synchronization at the end.
 
@@
 
-The effect of gravity on the hydrodynamical evolution is typically incorporated by the use of a source term for the momentum and energy equations. The integral form of the momentum source term is
+The effect of gravity on the hydrodynamical evolution is typically
+incorporated by the use of a source term for the momentum and energy
+equations. The integral form of the momentum source term is
 \begin{equation}
@@
 \end{equation}
-In most hydrodynamics codes these are discretized as $\rho\, \mathbf{g}\, \Delta V$ and $\rho\, \mathbf{u}\,\cdot\mathbf{g}\, \Delta V$, respectively, where $\Delta V$ is the zone volume and $\rho$, $\mathbf{u}$, and $\mathbf{g}$ are evaluated at the zone center.
-As observed by \cite{arepo}, the essential problem with this approach is that fluid movements in a grid code are calculated using fluxes at zone boundaries, while the standard gravity source term is an estimate of the flux using the zone center quantities.
-This can lead to a striking mismatch when the gradient across a zone is large.
-We have found that this is especially important at the edge of a star, where the large density gradient means that the momentum and energy fluxes can transition sharply even across half of a zone width.
-Another problem is that in a merger calculation in which two stars are moving on a grid, large quantities of material are being advected, so even a slight error in the calculation for the high-density material can manifest itself as a long-term error in energy conservation.
-This becomes apparent in Section \ref{sec:kepler}.
-So while a closer look at this may be unwarranted for situations like a stationary single star central explosion calculation, a detailed discussion is certainly called for in the case of an orbiting binary system where getting the gravity right is absolutely necessary for fully trusting the long-term evolution of the system.
+In most hydrodynamics codes these are discretized as $\rho\,
+\mathbf{g}\, \Delta V$ and $\rho\, \mathbf{u}\,\cdot\mathbf{g}\,
+\Delta V$, respectively, where $\Delta V$ is the zone volume and
+$\rho$, $\mathbf{u}$, and $\mathbf{g}$ are evaluated at the zone
+center. As observed by \cite{arepo}, the essential problem with this
+approach is that fluid movements in a grid code are calculated using
+fluxes at zone boundaries, while the standard gravity source term is
+an estimate of the flux using the zone center quantities. This can
+lead to a striking mismatch when the gradient across a zone is
+large. We have found that this is especially important at the edge of
+a star, where the large density gradient means that the momentum and
+energy fluxes can transition sharply even across half of a zone
+width. Another problem is that in a merger calculation in which two
+stars are moving on a grid, large quantities of material are being
+advected, so even a slight error in the calculation for the
+high-density material can manifest itself as a long-term error in
+energy conservation. This becomes apparent in Section
+\ref{sec:kepler}. So while a closer look at this may be unwarranted
+for situations like a stationary single star central explosion
+calculation, a detailed discussion is certainly called for in the case
+of an orbiting binary system where getting the gravity right is
+absolutely necessary for fully trusting the long-term evolution of the
+system.
 
@@
 \end{equation}
-The momentum equations are then written explicitly in conservative form when discretized: just as with the momentum stress tensor, divergences are calculated using discretized fluxes at zone boundaries, and the flux at any zone boundary is added to one cell and subtracted from another, so that the total momentum stays constant to within numerical roundoff error (with the possible exception of domain boundaries).
-This method works because, from the Poisson equation, $\rho$ can be written as proportional to the divergence of $\mathbf{g}$.
-We note that in order for the method to be reliable, the expression for all components of $\mathbf{g}$ must be correctly calculated from $\phi$ on zone boundaries so that the total divergence does indeed recover the density.
-This is straightforward for the component normal to the zone interface, but requires averaging over zones in the perpendicular direction to obtain the transverse components.
-
-Dealing with the energy equation is more difficult.
-The central challenge is to write down a form of the discretized energy equation that explicitly conserves total energy when coupled to gravity.
-When gravity is included, the total energy is
+The momentum equations are then written explicitly in conservative
+form when discretized: just as with the momentum stress tensor,
+divergences are calculated using discretized fluxes at zone
+boundaries, and the flux at any zone boundary is added to one cell and
+subtracted from another, so that the total momentum stays constant to
+within numerical roundoff error (with the possible exception of domain
+boundaries). This method works because, from the Poisson equation,
+$\rho$ can be written as proportional to the divergence of
+$\mathbf{g}$. We note that in order for the method to be reliable, the
+expression for all components of $\mathbf{g}$ must be correctly
+calculated from $\phi$ on zone boundaries so that the total divergence
+does indeed recover the density. This is straightforward for the
+component normal to the zone interface, but requires averaging over
+zones in the perpendicular direction to obtain the transverse
+components.
+
+Dealing with the energy equation is more difficult. The central
+challenge is to write down a form of the discretized energy equation
+that explicitly conserves total energy when coupled to gravity. When
+gravity is included, the total energy is
 \begin{equation}
@@
 \end{equation}
-where $E_G$ is the gas energy from the pure hydrodynamics equation, and the factor of $1/2$ in the gravitational energy term is necessary for simulations with self-gravity to prevent double-counting of interactions.
-Recently, three different approaches have been proposed for achieving this.
-We will briefly examine these and then describe the approach that makes the most sense for us.
-
-\citet{marcello:2012} use the mass continuity equation to rewrite the standard source term from Equation \ref{eq:cell_center_gravity_source}.
-When this is done, the resulting evolution equation for the total energy is:
+where $E_G$ is the gas energy from the pure hydrodynamics equation,
+and the factor of $1/2$ in the gravitational energy term is necessary
+for simulations with self-gravity to prevent double-counting of
+interactions. Recently, three different approaches have been proposed
+for achieving this. We will briefly examine these and then describe
+the approach that makes the most sense for us.
+
+\citet{marcello:2012} use the mass continuity equation to rewrite the
+standard source term from Equation
+\ref{eq:cell_center_gravity_source}. When this is done, the resulting
+evolution equation for the total energy is:
 \begin{equation}
@@
 \end{equation}
-In practice, we would still actually evolve the total gas energy $E_G$, and so when discretized to second order in time from time level $n$ to time level $n+1$ the gas energy equation would look like
+In practice, we would still actually evolve the total gas energy
+$E_G$, and so when discretized to second order in time from time level
+$n$ to time level $n+1$ the gas energy equation would look like
 \begin{align}
@@
 \end{align}
-It is not immediately obvious that this is explicitly conservative for the total energy, due to the last two terms on the right-hand side, but written in discrete form the potential is linked to the density in such a way that there is indeed a perfect cancellation of terms between adjacent cells (see Equation 88 in \citeauthor{marcello:2012}).
-Doing it this way requires a prescription for calculating $\phi^{n+1/2}$, the potential at the time halfway between the new and old state, as well as $\dot{\phi}^{n+1/2}$, the time derivative of the potential at that time.
-To second order in time, these can be calculated as $(\phi^{n+1} + \phi^{n})/2$ and $(\phi^{n+1}-\phi^{n})/\Delta t$ respectively.
-So, in practice what we do is to first solve the Poisson equation using $\rho^n$ at the beginning of a time step; then, evolve the pure hydrodynamics equations without the gravitational source to obtain $\rho^{n+1}$; then, solve the Poisson equation for $\phi^{n+1}$; and then to finally calculate the gravitational source terms and add them in.
-An alternative approach is to use the predictor-corrector scheme that CASTRO features, where the time level $n$ source term is included in the hydrodynamics update and then left out of the final update.
-In either case, the time level $n$ calculation of $\mathbf{g}$ is included the PPM reconstruction of the edge states.
-
-\citet{jiang:2013} start from the motivation of extending to the energy equation the flux-based formalism that allows explicit conservation for the momentum equation.
-They determine that the energy is perfectly conserved if we define a gravitational flux $\mathbf{F_g}$ and take its divergence for the energy equation:
+It is not immediately obvious that this is explicitly conservative for
+the total energy, due to the last two terms on the right-hand side,
+but written in discrete form the potential is linked to the density in
+such a way that there is indeed a perfect cancellation of terms
+between adjacent cells (see Equation 88 in
+\citeauthor{marcello:2012}). Doing it this way requires a prescription
+for calculating $\phi^{n+1/2}$, the potential at the time halfway
+between the new and old state, as well as $\dot{\phi}^{n+1/2}$, the
+time derivative of the potential at that time. To second order in
+time, these can be calculated as $(\phi^{n+1} + \phi^{n})/2$ and
+$(\phi^{n+1}-\phi^{n})/\Delta t$ respectively. So, in practice what we
+do is to first solve the Poisson equation using $\rho^n$ at the
+beginning of a time step; then, evolve the pure hydrodynamics
+equations without the gravitational source to obtain $\rho^{n+1}$;
+then, solve the Poisson equation for $\phi^{n+1}$; and then to finally
+calculate the gravitational source terms and add them in. An
+alternative approach is to use the predictor-corrector scheme that
+CASTRO features, where the time level $n$ source term is included in
+the hydrodynamics update and then left out of the final update. In
+either case, the time level $n$ calculation of $\mathbf{g}$ is
+included the PPM reconstruction of the edge states.
+
+\citet{jiang:2013} start from the motivation of extending to the
+energy equation the flux-based formalism that allows explicit
+conservation for the momentum equation. They determine that the energy
+is perfectly conserved if we define a gravitational flux
+$\mathbf{F_g}$ and take its divergence for the energy equation:
 \begin{align}
@@
 \end{align}
-This is analytically equivalent to the approach by \citet{marcello:2012}, which can be seen by taking the divergence of the flux and using the Poisson equation.
-It would be justified to use this form simply because it is cleaner to write down, but there is a more important benefit to using an explicit flux based formula: in AMR, conservation between levels can only happen if all fluid movements are written down using fluxes, because it is those fluxes that are adjusted at coarse-fine interfaces to ensure conservation.
-In this form we can simply add the gravitational source term to the hydrodynamical fluxes and use the existing AMR algorithm to guarantee conservation between levels.
-We note that \citeauthor{jiang:2013} say that this method requires the solution of an additional Poisson equation to obtain the intermediate time gravitational potential, but as we pointed out above, this can be done to second order simply by taking a centered difference using the existing Poisson solves, so no extra Poisson solve is necessary.
-
-This flux-based formalism has a significant drawback for our problem: unlike \citeauthor{jiang:2013}, we use open hydrodynamic boundary conditions, not periodic boundary conditions.
-Mass, momentum and energy can be lost (or gained) across the domain boundary.
-This is not a serious source of error in the standard cell-centered discretization, because the source term is proportional to $\rho$ and the density near the domain boundary is generally filled with a very low density ambient medium, so the size of the source term is always small.
-However, the gravitational flux in Equation \ref{eq:jiang_flux} has terms that are not explicitly proportional to density.
-Consequently, the flux lost across the domain boundaries can be significant (though this effect diminishes with increasing resolution).
-A related problem is that at the edge of a star, the gravitational flux can be so large that the gas energy becomes negative after an update.
-This effect is greatly diminished if the flux is instead proportional to density, as in the standard method.
-This is not a major problem for ideal or polytropic gases in pure hydrodynamics problems, but in our simulations we want the internal energy (and thus the temperature) to always be reliable so that nuclear reactions occur at the correct rates.
-This is especially sensitive at the edges of the stars, where the burning is expected to occur in a violent merger scneario.
-
-A possible compromise approach is discussed in \cite{arepo}.
-We start with the integral formulation of the energy source term (Equation \ref{eq:cell_center_gravity_source}), but instead of discretizing at the cell center, we use Gauss' theorem to evaluate the source term at the zone boundaries (see Equations 91--96 of that paper; we exclude the term related to the motion of the zones since our grids are fixed).
-This results in a gravitational update for zone $i$ written as
+This is analytically equivalent to the approach by
+\citet{marcello:2012}, which can be seen by taking the divergence of
+the flux and using the Poisson equation. It would be justified to use
+this form simply because it is cleaner to write down, but there is a
+more important benefit to using an explicit flux based formula: in
+AMR, conservation between levels can only happen if all fluid
+movements are written down using fluxes, because it is those fluxes
+that are adjusted at coarse-fine interfaces to ensure conservation. In
+this form we can simply add the gravitational source term to the
+hydrodynamical fluxes and use the existing AMR algorithm to guarantee
+conservation between levels. We note that \citeauthor{jiang:2013} say
+that this method requires the solution of an additional Poisson
+equation to obtain the intermediate time gravitational potential, but
+as we pointed out above, this can be done to second order simply by
+taking a centered difference using the existing Poisson solves, so no
+extra Poisson solve is necessary.
+
+This flux-based formalism has a significant drawback for our problem:
+unlike \citeauthor{jiang:2013}, we use open hydrodynamic boundary
+conditions, not periodic boundary conditions. Mass, momentum and
+energy can be lost (or gained) across the domain boundary. This is not
+a serious source of error in the standard cell-centered
+discretization, because the source term is proportional to $\rho$ and
+the density near the domain boundary is generally filled with a very
+low density ambient medium, so the size of the source term is always
+small. However, the gravitational flux in Equation \ref{eq:jiang_flux}
+has terms that are not explicitly proportional to
+density. Consequently, the flux lost across the domain boundaries can
+be significant (though this effect diminishes with increasing
+resolution). A related problem is that at the edge of a star, the
+gravitational flux can be so large that the gas energy becomes
+negative after an update. This effect is greatly diminished if the
+flux is instead proportional to density, as in the standard
+method. This is not a major problem for ideal or polytropic gases in
+pure hydrodynamics problems, but in our simulations we want the
+internal energy (and thus the temperature) to always be reliable so
+that nuclear reactions occur at the correct rates. This is especially
+sensitive at the edges of the stars, where the burning is expected to
+occur in a violent merger scneario.
+
+A possible compromise approach is discussed in \cite{arepo}. We start
+with the integral formulation of the energy source term (Equation
+\ref{eq:cell_center_gravity_source}), but instead of discretizing at
+the cell center, we use Gauss' theorem to evaluate the source term at
+the zone boundaries (see Equations 91--96 of that paper; we exclude
+the term related to the motion of the zones since our grids are
+fixed). This results in a gravitational update for zone $i$ written as
 \begin{equation}
@@
 \end{equation}
-where the sum is over all $j$ zones that share a boundary with zone $i$.
-Here $\mathbf{r}_{ij}$ is the radius vector joining zones $i$ and $j$ and $\Delta m_{ij}$ is the mass exchanged between zones $i$ and $j$ in the hydrodynamics update, and is equal to $\Delta t A_{ij} F_\rho^{ij}$, where $A_{ij}$ is the area of the zone boundary and $F_{\rho}^{ij}$ is the mass flux from the mass continuity equation.
-It is straightforward to see that the second version is explicitly conservative.
-The benefit of this form is that the mass exchanged between zones near the domain boundary will always be small in absolute terms since these zones are at very low density.
-Thus the energy lost through the boundary will be of comparable size to the standard method, but the gravity update should be more accurate since the change in gravitational energy now tracks the explicit motion of mass on the grid.
+where the sum is over all $j$ zones that share a boundary with zone
+$i$. Here $\mathbf{r}_{ij}$ is the radius vector joining zones $i$ and
+$j$ and $\Delta m_{ij}$ is the mass exchanged between zones $i$ and
+$j$ in the hydrodynamics update, and is equal to $\Delta t A_{ij}
+F_\rho^{ij}$, where $A_{ij}$ is the area of the zone boundary and
+$F_{\rho}^{ij}$ is the mass flux from the mass continuity equation. It
+is straightforward to see that the second version is explicitly
+conservative. The benefit of this form is that the mass exchanged
+between zones near the domain boundary will always be small in
+absolute terms since these zones are at very low density. Thus the
+energy lost through the boundary will be of comparable size to the
+standard method, but the gravity update should be more accurate since
+the change in gravitational energy now tracks the explicit motion of
+mass on the grid.
 
@@
 
-Analytic solutions to the Poisson equation customarily assume that the potential vanishes at large distances from the region of non-zero density.
-However, on a finite computational domain it is usually not possible to have the edges of the domain be far enough away that the potential can be taken to be zero there.
-Solving the Poisson equation therefore requires knowledge of the values of the potential on the edges of the computational domain.
-In principle, they can be computed by doing a direct sum over the mass distribution inside the domain, where the mass in each zone is treated as a point source:
+Analytic solutions to the Poisson equation customarily assume that the
+potential vanishes at large distances from the region of non-zero
+density. However, on a finite computational domain it is usually not
+possible to have the edges of the domain be far enough away that the
+potential can be taken to be zero there. Solving the Poisson equation
+therefore requires knowledge of the values of the potential on the
+edges of the computational domain. In principle, they can be computed
+by doing a direct sum over the mass distribution inside the domain,
+where the mass in each zone is treated as a point source:
 \begin{equation}
@@
 \end{equation}
-Here (i, j, k) are the indices of cells inside the domain, and (l, m, n) are the indices of boundary locations.
-$\Delta V$ is the volume of the zone.
-We have implemented this as an option \footnote{It is controlled with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in CASTRO.
-If there are $N$ zones per spatial dimension, then there are $6 N^2$ boundary zones, and each boundary zone requires a sum over $N^3$ zones, so the direct computation of the boundary conditions scales as $N^5$.
-This method is expensive enough that it is not used for hydrodynamics simulations (though it is useful for comparison to approximate solutions).
-
-In a typical simulation we place the boundaries of the domain far enough away from the region containing most of the mass that some method of approximation to this direct summation is justified.
-Many approaches exist in the literature.
-The original release of CASTRO featured the crudest possible approximation: a monopole prescription, where the boundary values were computed by summing up all the mass on the domain and treating it as a point source at the domain center.
-This is exactly correct only for a spherically symmetric mass distribution, and therefore is best suited for problems like single-star Type Ia supernova simulations that employ self-gravity.
-However, for a problem like that of a binary star system with significant departures from spherical symmetry, this assumption fails to produce accurate boundary values.
-This results in a significant drift of the center of the mass of the system over time, which will be discussed in Section \ref{sec:kepler}.
-
-The most natural extension of the monopole prescription is to include higher-order multipole moments.
-If the entire mass distribution is enclosed, then the potential can be expanded in a series of spherical harmonics $Y_{lm}$:
+Here (i, j, k) are the indices of cells inside the domain, and (l, m,
+n) are the indices of boundary locations. $\Delta V$ is the volume of
+the zone. We have implemented this as an option \footnote{It is
+  controlled with the \texttt{gravity.direct\_sum\_bcs} input
+  parameter.} in CASTRO. If there are $N$ zones per spatial dimension,
+then there are $6 N^2$ boundary zones, and each boundary zone requires
+a sum over $N^3$ zones, so the direct computation of the boundary
+conditions scales as $N^5$.  This method is expensive enough that it
+is not used for hydrodynamics simulations (though it is useful for
+comparison to approximate solutions).
+
+In a typical simulation we place the boundaries of the domain far
+enough away from the region containing most of the mass that some
+method of approximation to this direct summation is justified. Many
+approaches exist in the literature. The original release of CASTRO
+featured the crudest possible approximation: a monopole prescription,
+where the boundary values were computed by summing up all the mass on
+the domain and treating it as a point source at the domain
+center. This is exactly correct only for a spherically symmetric mass
+distribution, and therefore is best suited for problems like
+single-star Type Ia supernova simulations that employ
+self-gravity. However, for a problem like that of a binary star system
+with significant departures from spherical symmetry, this assumption
+fails to produce accurate boundary values. This results in a
+significant drift of the center of the mass of the system over time,
+which will be discussed in Section \ref{sec:kepler}.
+
+The most natural extension of the monopole prescription is to include
+higher-order multipole moments. If the entire mass distribution is
+enclosed, then the potential can be expanded in a series of spherical
+harmonics $Y_{lm}$:
 \begin{equation}
@@
 \end{equation}
-where $q_{lm}$ are the so-called multipole moments.
-The origin of the coordinate system is taken to be the center of the computational domain, and $r$ is the distance to the origin.
-The multipole moments can be calculated by expanding the Green's function for the Poisson equation as a series of spherical harmonics.
+where $q_{lm}$ are the so-called multipole moments. The origin of the
+coordinate system is taken to be the center of the computational
+domain, and $r$ is the distance to the origin. The multipole moments
+can be calculated by expanding the Green's function for the Poisson
+equation as a series of spherical harmonics.
 \begin{comment}
@@
 \end{comment}
-After some algebraic simplification of Equation \ref{spherical_harmonic_expansion} using the addition theorem for spherical harmonics,
+After some algebraic simplification of Equation
+\ref{spherical_harmonic_expansion} using the addition theorem for
+spherical harmonics,
 \begin{comment}
@@
 \end{comment}
-This approach becomes computationally feasible when we cut off the outer summation in Equation \ref{multipole_potential} at some finite value of $l_{\text{max}}$.
-If it is of sufficiently high order, we will accurately capture the distribution of mass on the grid.
-In practice we first evaluate the discretized analog of the modified multipole moments for $0 \leq l \leq l_{\text{max}}$ and $1 \leq m \leq l$, an operation that scales as $N^3$.
-We then directly compute the value of the potential on all of the $6N^2$ boundary zones.
-Since the multipole moments only need to be calculated once per Poisson solve, the full operation scales only as $N^3$.
-The amount of time required to calculate the boundary conditions will be directly related to the chosen value of $l_{\text{max}}$, so there is a trade-off between computational expense and accuracy of the result.
+This approach becomes computationally feasible when we cut off the
+outer summation in Equation \ref{multipole_potential} at some finite
+value of $l_{\text{max}}$. If it is of sufficiently high order, we
+will accurately capture the distribution of mass on the grid. In
+practice we first evaluate the discretized analog of the modified
+multipole moments for $0 \leq l \leq l_{\text{max}}$ and $1 \leq m
+\leq l$, an operation that scales as $N^3$. We then directly compute
+the value of the potential on all of the $6N^2$ boundary zones. Since
+the multipole moments only need to be calculated once per Poisson
+solve, the full operation scales only as $N^3$. The amount of time
+required to calculate the boundary conditions will be directly related
+to the chosen value of $l_{\text{max}}$, so there is a trade-off
+between computational expense and accuracy of the result.
 
@@
 
-Since the results of a merger simulation depend strongly on gravity, it is important to check whether proper numerical convergence is achieved for the Poisson solver.
-To do so, we created a simple test that inserts a sphere of radius $R$ and uniform mass density $\rho$ onto our grid, and used CASTRO to calculate the gravitational potential $\phi$ of this setup.
-We ensure that $R$ is an integer multiple of the grid spacing, and the center of the sphere is at the origin.
-The problem domain for our simulations is $[-1.6, 1.6]^3$, and we take $R = 0.5$ and $\rho = 10^3$.
-The zones with $r > R$ are filled with an ambient material of very low density.
-We run this problem at multiple resolutions corresponding to jumps by a factor of two.
-For comparison, at each grid point we evaluate the analytical potential of a uniform sphere, which can be easily determined using Gauss' law:
+Since the results of a merger simulation depend strongly on gravity,
+it is important to check whether proper numerical convergence is
+achieved for the Poisson solver. To do so, we created a simple test
+that inserts a sphere of radius $R$ and uniform mass density $\rho$
+onto our grid, and used CASTRO to calculate the gravitational
+potential $\phi$ of this setup. We ensure that $R$ is an integer
+multiple of the grid spacing, and the center of the sphere is at the
+origin. The problem domain for our simulations is $[-1.6, 1.6]^3$, and
+we take $R = 0.5$ and $\rho = 10^3$. The zones with $r > R$ are filled
+with an ambient material of very low density. We run this problem at
+multiple resolutions corresponding to jumps by a factor of two. For
+comparison, at each grid point we evaluate the analytical potential of
+a uniform sphere, which can be easily determined using Gauss' law:
 \begin{equation}
@@
 
-The results of this test are plotted in Figure \ref{fig:gravity_convergence}.
-We find that convergence is actually substantially better than second-order, and that the difference is largest at low resolutions.
-The explanation for this is that we are attempting to model a spherical object on a rectangular grid.
-At very low resolution, the object does not look very spherical.
-As the resolution is increased, the total amount of mass on the grid, as well as its location, will change as the sphere fills out.
-This means we are combining the true accuracy bonus from increased resolution with the artificial accuracy bonus from getting closer to solving the problem we are supposed to be solving.
-
-We can solve one of these two sources of error by evaluating Equation \ref{eq:sphere-analytical} with a mass $M$ that corresponds to the amount of mass actually on the grid, defined as the sum of the density multiplied by the volume of each cell for all cells that have greater than ambient density.
-The resolution study for this case is also plotted in Figure \ref{fig:gravity_convergence}.
-We still obtain convergence better than second-order, indicating that we still have the geometrical problem.
-
-The only way to fully eliminate this effect is to test a problem that does not change with resolution.
-The obvious companion is a cube of uniform density $\rho$, where now $R$ is half of the side length of the cube.
-At each resolution we use the same $R$ as for the sphere, which ensures that the cube always fills exactly the same fraction of the domain and thus has the same mass, so the only improvement comes from better sampling at higher resolution.
-The potential for this object has been worked out analytically by \cite{waldvogel:1976} (see also a similar result in \cite{hummer:1996}).
-The potential is in Equation 15 of that paper, though the last term is missing a factor of $1/2$, which destroys the symmetry.
-Inserting this missing factor and performing a simple coordinate transformation so that the center of the cube is at the origin, the potential is
+The results of this test are plotted in Figure
+\ref{fig:gravity_convergence}. We find that convergence is actually
+substantially better than second-order, and that the difference is
+largest at low resolutions. The explanation for this is that we are
+attempting to model a spherical object on a rectangular grid. At very
+low resolution, the object does not look very spherical. As the
+resolution is increased, the total amount of mass on the grid, as well
+as its location, will change as the sphere fills out. This means we
+are combining the true accuracy bonus from increased resolution with
+the artificial accuracy bonus from getting closer to solving the
+problem we are supposed to be solving.
+
+We can solve one of these two sources of error by evaluating Equation
+\ref{eq:sphere-analytical} with a mass $M$ that corresponds to the
+amount of mass actually on the grid, defined as the sum of the density
+multiplied by the volume of each cell for all cells that have greater
+than ambient density. The resolution study for this case is also
+plotted in Figure \ref{fig:gravity_convergence}. We still obtain
+convergence better than second-order, indicating that we still have
+the geometrical problem.
+
+The only way to fully eliminate this effect is to test a problem that
+does not change with resolution. The obvious companion is a cube of
+uniform density $\rho$, where now $R$ is half of the side length of
+the cube. At each resolution we use the same $R$ as for the sphere,
+which ensures that the cube always fills exactly the same fraction of
+the domain and thus has the same mass, so the only improvement comes
+from better sampling at higher resolution. The potential for this
+object has been worked out analytically by \cite{waldvogel:1976} (see
+also a similar result in \cite{hummer:1996}). The potential is in
+Equation 15 of that paper, though the last term is missing a factor of
+$1/2$, which destroys the symmetry. Inserting this missing factor and
+performing a simple coordinate transformation so that the center of
+the cube is at the origin, the potential is
 \begin{align}
@@
 \end{align}
-where $x_0 = R + x$, $x_1 = R - x$, $y_0 = R + y$, $y_1 = R-y$, $z_0 = R+z$, $z_1 = R-z$, and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$.
-In \texttt{FORTRAN} and \texttt{C}, the inverse hyperbolic tangent is \texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not} \texttt{atan2}).
-This formula is valid both inside and outside the cube.
-The normalized $L^2$ error norm for this problem is also shown in Figure \ref{fig:gravity_convergence}, and indeed we see near-perfect second-order scaling.
-
-The main lesson here is that in a convergence study, it is important to ensure that the physical problem does not change with resolution.
-Since in the case of spherical objects on rectangular grids the effect is to artificially boost convergence with resolution, in a simulation with spherical objects like stars one can envision a scenario of being fooled into believing apparently good convergence results that are simply a convolution of artificially high gravitational convergence and poor convergence in the hydrodynamics.
-A convergence study in this case is only fully valid if there is reason to be confident that this effect is negligible compared to other factors.
-
-Another factor that may affect convergence testing is easily seen in Figure \ref{fig:gravity_convergence}, where at the highest resolution the error flattens out.
-This occurs because of a limitation in the accuracy of the multigrid solver.
-For problems of this size we find empirically that the multigrid solver does not converge for error tolerances smaller than about $10^{-11}$, which enforces a lower limit on the total error.
-This is also something to be aware of in assessing convergence for our full simulations.
+where $x_0 = R + x$, $x_1 = R - x$, $y_0 = R + y$, $y_1 = R-y$, $z_0 =
+R+z$, $z_1 = R-z$, and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
+\texttt{FORTRAN} and \texttt{C}, the inverse hyperbolic tangent is
+\texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not}
+\texttt{atan2}). This formula is valid both inside and outside the
+cube. The normalized $L^2$ error norm for this problem is also shown
+in Figure \ref{fig:gravity_convergence}, and indeed we see
+near-perfect second-order scaling.
+
+The main lesson here is that in a convergence study, it is important
+to ensure that the physical problem does not change with
+resolution. Since in the case of spherical objects on rectangular
+grids the effect is to artificially boost convergence with resolution,
+in a simulation with spherical objects like stars one can envision a
+scenario of being fooled into believing apparently good convergence
+results that are simply a convolution of artificially high
+gravitational convergence and poor convergence in the hydrodynamics. A
+convergence study in this case is only fully valid if there is reason
+to be confident that this effect is negligible compared to other
+factors.
+
+Another factor that may affect convergence testing is easily seen in
+Figure \ref{fig:gravity_convergence}, where at the highest resolution
+the error flattens out. This occurs because of a limitation in the
+accuracy of the multigrid solver. For problems of this size we find
+empirically that the multigrid solver does not converge for error
+tolerances smaller than about $10^{-11}$, which enforces a lower limit
+on the total error. This is also something to be aware of in assessing
+convergence for our full simulations.
 
@@
 
-For the evolution of binary systems, it is most natural to evolve the two stars in a frame that is co-rotating at the same period as the orbital period.
-CASTRO has the ability to evolve systems in a rotating reference frame, as discussed in the original code paper.
-Source terms corresponding to the Coriolis and centrifugal force terms are added to the momentum and energy equations.
-In this frame, the stars essentially remain stationary in their original positions due to the centrifugal force supporting against the gravitational attraction, and will remain this way as long as significant mass transfer does not occur.
-\cite{swc:2000} demonstrated (in the context of neutron star mergers) that conservation of angular momentum is much easier to obtain in the rotating reference frame than in an inertial frame in which stars advect large amounts of material around the domain.
-We wish to emphasize that although it is commonly stated in the literature that grid-based codes poorly conserve angular momentum, this is not generally true.
-When the resolution is sufficiently high and a rotating reference frame is employed, excellent conservation properties can result, as demonstrated in Section \ref{sec:kepler}.
-We note that as the stars begin to coalesce, the rotating reference frame will no longer provide a good approximation to the spatial motion of the stars and then they will begin to significantly move around the domain.
-This is not necessarily problematic because the most important feature of the rotating frame is that it helps ensure that the initial coalescence is not the result of spurious numerical loss of angular momentum.
-When significant mass transfer sets in and evolution proceeds on a dynamical timescale, the conservation properties may be slightly worse but angular momentum conservation is also less important.
+For the evolution of binary systems, it is most natural to evolve the
+two stars in a frame that is co-rotating at the same period as the
+orbital period. CASTRO has the ability to evolve systems in a rotating
+reference frame, as discussed in the original code paper. Source terms
+corresponding to the Coriolis and centrifugal force terms are added to
+the momentum and energy equations. In this frame, the stars
+essentially remain stationary in their original positions due to the
+centrifugal force supporting against the gravitational attraction, and
+will remain this way as long as significant mass transfer does not
+occur. \cite{swc:2000} demonstrated (in the context of neutron star
+mergers) that conservation of angular momentum is much easier to
+obtain in the rotating reference frame than in an inertial frame in
+which stars advect large amounts of material around the domain. We
+wish to emphasize that although it is commonly stated in the
+literature that grid-based codes poorly conserve angular momentum,
+this is not generally true. When the resolution is sufficiently high
+and a rotating reference frame is employed, excellent conservation
+properties can result, as demonstrated in Section \ref{sec:kepler}. We
+note that as the stars begin to coalesce, the rotating reference frame
+will no longer provide a good approximation to the spatial motion of
+the stars and then they will begin to significantly move around the
+domain. This is not necessarily problematic because the most important
+feature of the rotating frame is that it helps ensure that the initial
+coalescence is not the result of spurious numerical loss of angular
+momentum. When significant mass transfer sets in and evolution
+proceeds on a dynamical timescale, the conservation properties may be
+slightly worse but angular momentum conservation is also less
+important.
 
@@
 
-At the start of any full simulation, we generate initial model white dwarfs by integrating the equation of hydrostatic equilibrium, taking the temperature and composition to be constant, and using the general stellar equation of state.
-This results in a single non-linear equation to find the density in a zone given the conditions in the zone beneath it:
+At the start of any full simulation, we generate initial model white
+dwarfs by integrating the equation of hydrostatic equilibrium, taking
+the temperature and composition to be constant, and using the general
+stellar equation of state.  This results in a single non-linear
+equation to find the density in a zone given the conditions in the
+zone beneath it:
 \begin{equation}
@@
 \end{equation}
-This equation is a function of $\rho_{i+1}$ only since the pressure is uniquely determined by the density in this case.
-$\rho_i$ and $p_i$ are known,  $g_{i+1/2}$ is the gravitational acceleration at the interface between zones $i$ and $i+1$, found by simply adding up all the mass from zones $1, \ldots, i$ to get the enclosed mass, $M_{i+1/2}$, and then setting $g_{i+1/2} = -GM_{i+1/2}/r_{i+1/2}^2$.
-We solve this equation for $\rho_{i+1}$ using a Newton-Raphson iteration.
-
-We desire to specify the mass of the white dwarf, as well as its temperature and composition.
-To start the integration off, we therefore need to guess at a central density.
-We then do a secant iteration over the entire integration procedure to find the central density needed to yield the desired total mass.
-We note that the grid spacing $\Delta x$ is chosen such that it is equal to the zone size on the finest level of our problem, which is known in advance since we specify the domain size and the maximum number of possible levels of refinement.
-
-We map the 1D model onto the 3D Cartesian grid by taking density, temperature, and composition as the independent variables, interpolating these to the cell centers, and then calling the equation of state to initialize the remaining terms.
-The interpolation process divides each zone into $n_{\text{sub}}$ sub-zones of equal volume for the purpose of sampling the 1D model, and the sub-zones are added together to obtain the full zone's characteristics.
-This sub-grid-scale interpolation is useful near the edge of the star, where the density falls off rapidly with radius.
-
-For a single star simulation, the star is simply placed at the center of the computational domain, which we take to be the origin.
-For a binary star simulation, we take as parameters the mass of the two white dwarfs and the initial orbital period $T$.
-Using Kepler's third law and assuming a circular orbit, we can then work out the orbital separation $a$:
+This equation is a function of $\rho_{i+1}$ only since the pressure is
+uniquely determined by the density in this case. $\rho_i$ and $p_i$
+are known, $g_{i+1/2}$ is the gravitational acceleration at the
+interface between zones $i$ and $i+1$, found by simply adding up all
+the mass from zones $1, \ldots, i$ to get the enclosed mass,
+$M_{i+1/2}$, and then setting $g_{i+1/2} =
+-GM_{i+1/2}/r_{i+1/2}^2$. We solve this equation for $\rho_{i+1}$
+using a Newton-Raphson iteration.
+
+We desire to specify the mass of the white dwarf, as well as its
+temperature and composition. To start the integration off, we
+therefore need to guess at a central density.  We then do a secant
+iteration over the entire integration procedure to find the central
+density needed to yield the desired total mass.  We note that the grid
+spacing $\Delta x$ is chosen such that it is equal to the zone size on
+the finest level of our problem, which is known in advance since we
+specify the domain size and the maximum number of possible levels of
+refinement.
+
+We map the 1D model onto the 3D Cartesian grid by taking density,
+temperature, and composition as the independent variables,
+interpolating these to the cell centers, and then calling the equation
+of state to initialize the remaining terms.  The interpolation process
+divides each zone into $n_{\text{sub}}$ sub-zones of equal volume for
+the purpose of sampling the 1D model, and the sub-zones are added
+together to obtain the full zone's characteristics. This
+sub-grid-scale interpolation is useful near the edge of the star,
+where the density falls off rapidly with radius.
+
+For a single star simulation, the star is simply placed at the center
+of the computational domain, which we take to be the origin. For a
+binary star simulation, we take as parameters the mass of the two
+white dwarfs and the initial orbital period $T$. Using Kepler's third
+law and assuming a circular orbit, we can then work out the orbital
+separation $a$:
 \begin{equation}
@@
 \end{equation}
-Here $M = M_P + M_S$ is the total mass of the system, where $M_P$ is the specified \textit{primary} mass and $M_S$ is the specified \textit{secondary} mass.
-The primary WD will always be on the left side of the computational domain for our simulations, and is more massive than the secondary.
-This reflects the usual terminology in the literature where the primary WD is the accretor and the secondary is the donor.
-The center of mass is located at the center of the computational domain, and the stars lie along the $x$ axis, so that the primary's center of mass is located at $x = -(M_S / M)\, a$ and the secondary's center of mass is located at $x = (M_P / M)\, a$.
-
-The initial velocity is taken to be zero in if we are in the reference frame that rotates with the WDs, and if we are in the inertial frame the velocity is set equal to the rigid rotation rate corresponding to the specified period $T$.
-
-We do not attempt to enforce equilibrium with an additional relaxation step.
-This will be an important part of future work in this series, as numerous groups working on binary evolution \citep{swc:2000,motl:2002,rosswog:2004,dan:2011,pakmor:2012:gadget} have commented on the importance of equilibrium initial conditions in determining the evolution of the system.
-In particular, we plan to compare the self-consistent field method (which finds an equilibrium solution \textit{a priori}) and the method of \cite{rosswog:2004}, which starts with our method and then attempts to guide the evolution toward the equilibrium state using suitably defined damping terms.
-However, these issues are not considered for the present study, as we are not yet attempting to study the merger process.
+Here $M = M_P + M_S$ is the total mass of the system, where $M_P$ is
+the specified \textit{primary} mass and $M_S$ is the specified
+\textit{secondary} mass. The primary WD will always be on the left
+side of the computational domain for our simulations, and is more
+massive than the secondary. This reflects the usual terminology in the
+literature where the primary WD is the accretor and the secondary is
+the donor. The center of mass is located at the center of the
+computational domain, and the stars lie along the $x$ axis, so that
+the primary's center of mass is located at $x = -(M_S / M)\, a$ and
+the secondary's center of mass is located at $x = (M_P / M)\, a$.
+
+The initial velocity is taken to be zero in if we are in the reference
+frame that rotates with the WDs, and if we are in the inertial frame
+the velocity is set equal to the rigid rotation rate corresponding to
+the specified period $T$.
+
+We do not attempt to enforce equilibrium with an additional relaxation
+step. This will be an important part of future work in this series, as
+numerous groups working on binary evolution
+\citep{swc:2000,motl:2002,rosswog:2004,dan:2011,pakmor:2012:gadget}
+have commented on the importance of equilibrium initial conditions in
+determining the evolution of the system. In particular, we plan to
+compare the self-consistent field method (which finds an equilibrium
+solution \textit{a priori}) and the method of \cite{rosswog:2004},
+which starts with our method and then attempts to guide the evolution
+toward the equilibrium state using suitably defined damping
+terms. However, these issues are not considered for the present study,
+as we are not yet attempting to study the merger process.
 
@@
 
-Merger simulations face a number of numerical difficulties that are not present in single-degenerate Type Ia and core-collapse supernova simulations.
-In Section \ref{sec:gravity}, we discussed how the lack of spherical symmetry necessitates a careful look at the gravity solver.
-There are also hydrodynamical issues: the merger process will result in substantial motion of stellar material across the grid.
-This bulk motion presents an opportunity for advection errors to build up, and is only partially mitigated by evolving the white dwarfs in a co-rotating frame.
-It is therefore important to be be aware of the behavior of the code in such circumstances.
-The behavior of CASTRO for many standard hydrodynamics test problems was detailed in the original code paper \citep{castro}, and in the interest of brevity we will not repeat them here.
-Instead, we focus on a subset of problems that highlight the special difficulties introduced in merger simulations.
-These problems couple the hydrodynamics, gravity and equation of state modules.
-We observe that while in most non-trivial three-dimensional problems this creates a complexity that makes it impossible to determine exact analytical solutions, it is straightforward to devise problems for which certain global properties should obey simple, expected behaviors.
-Where possible, these should be quantified and a convergence study performed.
-We recommend that all simulation codes should check their performance on such problems to ensure their reliability when coupling the various physics modules necessary to build a complete and realistic simulation.
+Merger simulations face a number of numerical difficulties that are
+not present in single-degenerate Type Ia and core-collapse supernova
+simulations. In Section \ref{sec:gravity}, we discussed how the lack
+of spherical symmetry necessitates a careful look at the gravity
+solver. There are also hydrodynamical issues: the merger process will
+result in substantial motion of stellar material across the grid. This
+bulk motion presents an opportunity for advection errors to build up,
+and is only partially mitigated by evolving the white dwarfs in a
+co-rotating frame. It is therefore important to be be aware of the
+behavior of the code in such circumstances. The behavior of CASTRO for
+many standard hydrodynamics test problems was detailed in the original
+code paper \citep{castro}, and in the interest of brevity we will not
+repeat them here. Instead, we focus on a subset of problems that
+highlight the special difficulties introduced in merger
+simulations. These problems couple the hydrodynamics, gravity and
+equation of state modules. We observe that while in most non-trivial
+three-dimensional problems this creates a complexity that makes it
+impossible to determine exact analytical solutions, it is
+straightforward to devise problems for which certain global properties
+should obey simple, expected behaviors. Where possible, these should
+be quantified and a convergence study performed. We recommend that all
+simulation codes should check their performance on such problems to
+ensure their reliability when coupling the various physics modules
+necessary to build a complete and realistic simulation.
 
@@
 
-In Section \ref{sec:initial_models} we describe the process by which we generate initial stellar models.
-While the 1D models are in hydrostatic equilibrium to within a small error, interpolation onto the 3D Cartesian grid will introduce perturbations into the solution \citep{zingale:2002}.
-Although we ensure that the initial models are generated with the same equation of state and are as well resolved as our finest grid, there will still be a hydrodynamical error associated with the fact that the rectangular grid cannot faithfully represent a spherical star.
-Additionally, the gravitational potential obtained by the multigrid solver will differ slightly from the one assumed by the initial model, and the operator splitting between the gravity and hydrodynamics should also result in small errors.
-As a result, we expect that the star will oscillate slightly about an equilibrium point, but that the amplitude of this oscillation should decrease with increasing resolution.
-
-This problem was studied in the first CASTRO paper, but is worth revisiting here.
-A single star explosion simulation may only last a couple of seconds, and the CASTRO paper studied the behavior of the star after one second of evolution.
-However, the dynamical timescale of a typical carbon-oxygen white dwarf is on the order of 1--10 seconds.
-Additionally, a binary orbit is typically on the order of 10--100 seconds when a merger simulation starts, and with equilibrium initial conditions the system may survive for tens of orbits before the secondary is disrupted.
-When this does happen, we want to be confident that it was because of the dynamics of the merger process and not because of an instability in an individual star.
-Our goal here is thus to populate a single star onto our three-dimensional coordinate grid and evolve it for a period of time long enough to assess whether the star is truly stable, and to probe how the size of deviation from equilibrium is affected by grid resolution.
+In Section \ref{sec:initial_models} we describe the process by which
+we generate initial stellar models. While the 1D models are in
+hydrostatic equilibrium to within a small error, interpolation onto
+the 3D Cartesian grid will introduce perturbations into the solution
+\citep{zingale:2002}. Although we ensure that the initial models are
+generated with the same equation of state and are as well resolved as
+our finest grid, there will still be a hydrodynamical error associated
+with the fact that the rectangular grid cannot faithfully represent a
+spherical star. Additionally, the gravitational potential obtained by
+the multigrid solver will differ slightly from the one assumed by the
+initial model, and the operator splitting between the gravity and
+hydrodynamics should also result in small errors. As a result, we
+expect that the star will oscillate slightly about an equilibrium
+point, but that the amplitude of this oscillation should decrease with
+increasing resolution.
+
+This problem was studied in the first CASTRO paper, but is worth
+revisiting here. A single star explosion simulation may only last a
+couple of seconds, and the CASTRO paper studied the behavior of the
+star after one second of evolution. However, the dynamical timescale
+of a typical carbon-oxygen white dwarf is on the order of 1--10
+seconds. Additionally, a binary orbit is typically on the order of
+10--100 seconds when a merger simulation starts, and with equilibrium
+initial conditions the system may survive for tens of orbits before
+the secondary is disrupted. When this does happen, we want to be
+confident that it was because of the dynamics of the merger process
+and not because of an instability in an individual star. Our goal here
+is thus to populate a single star onto our three-dimensional
+coordinate grid and evolve it for a period of time long enough to
+assess whether the star is truly stable, and to probe how the size of
+deviation from equilibrium is affected by grid resolution.
 
@@
 
-It is often stated in the literature that Eulerian methods for hydrodynamics with grids fixed in space do not obey the Galilean invariance of the underlying Euler equations, so that simulations moving at a uniform bulk velocity will appear different than an equivalent stationary simulation.
-If true, this means that codes such as CASTRO may be fundamentally inappropriate for merger problems (though any errors should be greatly diminished in a reference frame that rotates with the initial binary orbit).
-Recently, this has come up in two ways which are of note for us in the present study.
-We will explain these situations and run tests to determine whether this actually is a significant concern.
-
-\citet{arepo} performed a Kelvin-Helmholtz instability test and showed that (at low resolution) a fixed-grid code failed to develop the expected fluid instability when the whole fluid was moving at a strongly supersonic uniform velocity.
-This contrasted with the results of the moving-mesh code AREPO being presented in that study, which demonstrated Galilean invariance even at large bulk velocities.
-If verified, this claim would have important consequences on how much we can trust the ability of CASTRO to test the violent merger progenitor model.
-Shearing between the material flowing out of the secondary and material near the surface of the primary may trigger fluid instabilities that play an important role in the evolution of that gas, which is the site of the initial detonation in the prompt explosion model.
-\citet{guillochon:2010} showed for their simulation that Kelvin-Helmholtz instabilities produced this way may raise the temperature of the accreting material enough to ignite a detonation.
-Therefore if we are not correctly reproducing the characteristics of the Kelvin-Helmholtz instability in the case where there is significant mass motion on the grid, we cannot be confident that a detonation (or lack thereof) is not numerically seeded.
-However, \citet{robertson:2010} observe that Galilean invariance of simulation results for the Euler equations occurs only because of truncation error in the discretization of the fluid equations.
-This takes the form of a numerical diffusion term which is dependent on velocity (and also resolution).
-The advantage of a moving-mesh code is that the mesh everywhere moves with the local flow velocity, which substantially reduces the numerical diffusion.
-\citeauthor{robertson:2010} argue that the differences seen between the moving-mesh and fixed-grid code are caused by the interaction of this numerical diffusion with small-scale instabilities (which may be physical or numerical) which couple with and fundamentally alter the large-scale modes.
-Small-scale instabilities are seeded in this problem by the choice of a sharp initial discontinuity between the fluids.
-Crucially though, \citeauthor{robertson:2010} point out that this problem does not converge with resolution and so it is not possible to know the correct behavior of this problem.
-As such, we do not know whether the small-scale modes found in AREPO are real, so the problem is not useful in formally discriminating between methodologies.
-They instead propose an alternate test with a smoother initial contact.
-This converges to the same solution qualitatively in both the stationary and bulk velocity cases, indicating that the code does generally maintain Galilean invariance (to some specified error that depends on resolution and the uniform flow speed).
-We will see whether we can reproduce this result, as well as a result by \cite{wadsley:2008}, who used the FLASH code to simulate a hot bubble subject to mixing by the Kelvin-Helmholtz instability, and also found that the mixing was affected by a uniform bulk velocity.
-
-A related question is whether our code reliably simulates the bulk motion of the stars across the grid, and whether such bulk motion affects the stability of the star.
-This concern is prompted by the study of \cite{tasker:2008}, who studied the effect of uniform translation on the stability of a spherically symmetric model for a galaxy cluster.
-They compared the radial profile of the cluster at initialization and after a period of time evolution.
-Using FLASH and ENZO, they found that a static cluster retains its shape at high enough resolution, while uniform translation of the cluster causes mixing of the core material due to numerical diffusion which results in an underestimation of the core's true density.
-The SPH codes they used did a better job maintaining the core density.
-We will perform a variant of this test using white dwarf models.
+It is often stated in the literature that Eulerian methods for
+hydrodynamics with grids fixed in space do not obey the Galilean
+invariance of the underlying Euler equations, so that simulations
+moving at a uniform bulk velocity will appear different than an
+equivalent stationary simulation. If true, this means that codes such
+as CASTRO may be fundamentally inappropriate for merger problems
+(though any errors should be greatly diminished in a reference frame
+that rotates with the initial binary orbit). Recently, this has come
+up in two ways which are of note for us in the present study. We will
+explain these situations and run tests to determine whether this
+actually is a significant concern.
+
+\citet{arepo} performed a Kelvin-Helmholtz instability test and showed
+that (at low resolution) a fixed-grid code failed to develop the
+expected fluid instability when the whole fluid was moving at a
+strongly supersonic uniform velocity. This contrasted with the results
+of the moving-mesh code AREPO being presented in that study, which
+demonstrated Galilean invariance even at large bulk velocities. If
+verified, this claim would have important consequences on how much we
+can trust the ability of CASTRO to test the violent merger progenitor
+model. Shearing between the material flowing out of the secondary and
+material near the surface of the primary may trigger fluid
+instabilities that play an important role in the evolution of that
+gas, which is the site of the initial detonation in the prompt
+explosion model. \citet{guillochon:2010} showed for their simulation
+that Kelvin-Helmholtz instabilities produced this way may raise the
+temperature of the accreting material enough to ignite a
+detonation. Therefore if we are not correctly reproducing the
+characteristics of the Kelvin-Helmholtz instability in the case where
+there is significant mass motion on the grid, we cannot be confident
+that a detonation (or lack thereof) is not numerically
+seeded. However, \citet{robertson:2010} observe that Galilean
+invariance of simulation results for the Euler equations occurs only
+because of truncation error in the discretization of the fluid
+equations. This takes the form of a numerical diffusion term which is
+dependent on velocity (and also resolution). The advantage of a
+moving-mesh code is that the mesh everywhere moves with the local flow
+velocity, which substantially reduces the numerical
+diffusion. \citeauthor{robertson:2010} argue that the differences seen
+between the moving-mesh and fixed-grid code are caused by the
+interaction of this numerical diffusion with small-scale instabilities
+(which may be physical or numerical) which couple with and
+fundamentally alter the large-scale modes. Small-scale instabilities
+are seeded in this problem by the choice of a sharp initial
+discontinuity between the fluids. Crucially though,
+\citeauthor{robertson:2010} point out that this problem does not
+converge with resolution and so it is not possible to know the correct
+behavior of this problem. As such, we do not know whether the
+small-scale modes found in AREPO are real, so the problem is not
+useful in formally discriminating between methodologies. They instead
+propose an alternate test with a smoother initial contact. This
+converges to the same solution qualitatively in both the stationary
+and bulk velocity cases, indicating that the code does generally
+maintain Galilean invariance (to some specified error that depends on
+resolution and the uniform flow speed).  We will see whether we can
+reproduce this result, as well as a result by \cite{wadsley:2008}, who
+used the FLASH code to simulate a hot bubble subject to mixing by the
+Kelvin-Helmholtz instability, and also found that the mixing was
+affected by a uniform bulk velocity.
+
+A related question is whether our code reliably simulates the bulk
+motion of the stars across the grid, and whether such bulk motion
+affects the stability of the star. This concern is prompted by the
+study of \cite{tasker:2008}, who studied the effect of uniform
+translation on the stability of a spherically symmetric model for a
+galaxy cluster. They compared the radial profile of the cluster at
+initialization and after a period of time evolution. Using FLASH and
+ENZO, they found that a static cluster retains its shape at high
+enough resolution, while uniform translation of the cluster causes
+mixing of the core material due to numerical diffusion which results
+in an underestimation of the core's true density. The SPH codes they
+used did a better job maintaining the core density. We will perform a
+variant of this test using white dwarf models.
 
@@
 
-Following \cite{robertson:2010}, we set up a Kelvin-Helmholtz test in the following way.
-The problem domain runs from 0 to 1 in both the $x$ and $y$ directions.
-Though this is a two-dimensional test, we run CASTRO in 3D and replicate the problem along 8 cells in the $z$ direction; the physical size of these zones is determined so that $dz = dx = dy$, a requirement at present in CASTRO.
-The problem involves a fluid slab of density $\rho_1 = 2.0$ traveling rightward in the $x$-direction at velocity $v_1 = 0.5$, sandwiched by a fluid of density $\rho_2 = 1.0$ traveling leftward at velocity $v_2 = -0.5$.
-The density gradient is in the $y$ direction, so this creates a velocity shear along the interface between the fluids.
-The density and velocity distribution on the computational domain are given by:
+Following \cite{robertson:2010}, we set up a Kelvin-Helmholtz test in
+the following way. The problem domain runs from 0 to 1 in both the $x$
+and $y$ directions. Though this is a two-dimensional test, we run
+CASTRO in 3D and replicate the problem along 8 cells in the $z$
+direction; the physical size of these zones is determined so that $dz
+= dx = dy$, a requirement at present in CASTRO. The problem involves a
+fluid slab of density $\rho_1 = 2.0$ traveling rightward in the
+$x$-direction at velocity $v_1 = 0.5$, sandwiched by a fluid of
+density $\rho_2 = 1.0$ traveling leftward at velocity $v_2 =
+-0.5$. The density gradient is in the $y$ direction, so this creates a
+velocity shear along the interface between the fluids. The density and
+velocity distribution on the computational domain are given by:
 \begin{align}
@@
 \end{align}
-Here $R(y)$ is a ramp function that describes the transition between the two fluids, while $v_{\text{bulk}}$ is the bulk motion of the fluid in the $y$ direction and $v^\prime$ is the velocity perturbation that seeds the instability.
-The problem will be established for two sets of initial conditions (ICs), which we follow \citeauthor{robertson:2010} in calling ICs A and B.
-They differ in their ramp function ($R_A$ and $R_B$ respectively), as well as the initial perturbation ($v^\prime_A$ and $v^\prime_B$ respectively), and the frequency of the perturbation ($n_A = 4$ and $n_B = 2$):
+Here $R(y)$ is a ramp function that describes the transition between
+the two fluids, while $v_{\text{bulk}}$ is the bulk motion of the
+fluid in the $y$ direction and $v^\prime$ is the velocity perturbation
+that seeds the instability. The problem will be established for two
+sets of initial conditions (ICs), which we follow
+\citeauthor{robertson:2010} in calling ICs A and B. They differ in
+their ramp function ($R_A$ and $R_B$ respectively), as well as the
+initial perturbation ($v^\prime_A$ and $v^\prime_B$ respectively), and
+the frequency of the perturbation ($n_A = 4$ and $n_B = 2$):
 \begin{align}
@@
 \end{align}
-Here $w_0 = 0.1$ is the scale of the velocity perturbation, $\sigma = 0.05/\sqrt{2}$ controls the width of the Gaussian for IC A, and $\Delta_y$ is the transition distance scale for the smooth ramp of IC B.
-The pressure everywhere is set to $p = 2.5$, and we run this with a gamma-law equation of state set to $\gamma = 5/3$.
-Plotfiles are generated every 0.01 seconds, and the problem is run until $t = 2$.
-
-We run the problem for $v_\text{bulk} = [0, 1, 3, 10, 30, 100]$, and for each set of initial conditions run the problem at resolutions of $64^2$, $128^2$, $256^2$, $512^2$.
-In addition, for each initial condition we run a simulation at the very high resolution of $2560^2$ for the stationary problem only.
-This serves as a reference solution to gauge the extent to which the bulk flow affects the development of the fluid instability.
+Here $w_0 = 0.1$ is the scale of the velocity perturbation, $\sigma =
+0.05/\sqrt{2}$ controls the width of the Gaussian for IC A, and
+$\Delta_y$ is the transition distance scale for the smooth ramp of IC
+B. The pressure everywhere is set to $p = 2.5$, and we run this with a
+gamma-law equation of state set to $\gamma = 5/3$. Plotfiles are
+generated every 0.01 seconds, and the problem is run until $t = 2$.
+
+We run the problem for $v_\text{bulk} = [0, 1, 3, 10, 30, 100]$, and
+for each set of initial conditions run the problem at resolutions of
+$64^2$, $128^2$, $256^2$, $512^2$. In addition, for each initial
+condition we run a simulation at the very high resolution of $2560^2$
+for the stationary problem only. This serves as a reference solution
+to gauge the extent to which the bulk flow affects the development of
+the fluid instability.
 
@@
 \end{eqnarray}
-where here the explicit advective term from the Godunov method is represented as $A$ and is already time-centered.
-In the Castro algorithm, a full $\Delta t$ of the old-time-level source is first added and later we correct this by subtracting off $\Delta t/2$ of the old-time-level term and adding $\Delta t/2$ of the new-time-level source.
-This correction appears as:
+where here the explicit advective term from the Godunov method is
+represented as $A$ and is already time-centered.  In the Castro
+algorithm, a full $\Delta t$ of the old-time-level source is first
+added and later we correct this by subtracting off $\Delta t/2$ of the
+old-time-level term and adding $\Delta t/2$ of the new-time-level
+source.  This correction appears as:
 \begin{eqnarray}

### 2015-01-15 papers/wdmerger_I/paper.tex
minor corrections

@@
 total specific energy, where $e$ is the internal (thermal) specific
-energy.
+energy (energy per unit mass).
 
@@
 \end{equation}
-Here (i, j, k) are the indices of cells inside the domain, and (l, m,
+Here $(i, j, k)$ are the indices of cells inside the domain, and (l, m,
 n) are the indices of boundary locations. $\Delta V$ is the volume of
@@
 \end{equation}
-where $M = 4\pi R^3 / 3$ is the mass of the sphere. Defining the $L^2$ norm of a field $f$ as
+where $M = 4\pi R^3 / 3$ is the mass of the sphere. Defining the $L^2$
+norm of a field $f$ as \MarginPar{We usually use $\| \ldots \|_2$ for a norm}
 \begin{equation}
@@
 from better sampling at higher resolution. The potential for this
-object has been worked out analytically by \cite{waldvogel:1976} (see
-also a similar result in \cite{hummer:1996}). The potential is in
+object has been worked out analytically by \citet{waldvogel:1976} (see
+also a similar result in \citealt{hummer:1996}). The potential is in
 Equation 15 of that paper, though the last term is missing a factor of
@@
 This equation is a function of $\rho_{i+1}$ only since the pressure is
-uniquely determined by the density in this case. $\rho_i$ and $p_i$
+uniquely determined by the density in this case. Here, $\rho_i$ and $p_i$
 are known, $g_{i+1/2}$ is the gravitational acceleration at the

### 2015-01-15 papers/wdmerger_I/paper.tex
more TeX

@@
 \begin{equation}
-  \Phi_{\text{lmn}} = G\, \sum_{\text{i, j, k}} \frac{G \rho_{\text{ijk}}}{|\mathbf{x}_{\text{lmn}} - \mathbf{x}_{\text{ijk}}|}\, \Delta V_{\text{ijk}}.\label{direct_sum}
-\end{equation}
-Here $(i, j, k)$ are the indices of cells inside the domain, and (l, m,
-n) are the indices of boundary locations. $\Delta V$ is the volume of
-the zone. We have implemented this as an option \footnote{It is
-  controlled with the \texttt{gravity.direct\_sum\_bcs} input
-  parameter.} in CASTRO. If there are $N$ zones per spatial dimension,
-then there are $6 N^2$ boundary zones, and each boundary zone requires
-a sum over $N^3$ zones, so the direct computation of the boundary
-conditions scales as $N^5$.  This method is expensive enough that it
-is not used for hydrodynamics simulations (though it is useful for
-comparison to approximate solutions).
+  \Phi_{{lmn}} = G\, \sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{direct_sum}
+\end{equation}
+Here $(i, j, k)$ are the indices of cells inside the domain, and $(l,
+m, n)$ \MarginPar{indices should be in normal math style, not roman}
+are the indices of boundary locations. $\Delta V$ is the volume of the
+zone. We have implemented this as an option\footnote{It is controlled
+  with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in
+CASTRO. If there are $N$ zones per spatial dimension, then there are
+$6 N^2$ boundary zones, and each boundary zone requires a sum over
+$N^3$ zones, so the direct computation of the boundary conditions
+scales as $\mathcal{O}(N^5)$.  This method is expensive enough that it is not used
+for hydrodynamics simulations (though it is useful for comparison to
+approximate solutions).
 

### 2015-01-15 papers/wdmerger_I/paper.tex
another note

@@
   \Phi_{{lmn}} = G\, \sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{direct_sum}
-\end{equation}
+\end{equation} \MarginPar{should there really be 2 $G$'s?}
 Here $(i, j, k)$ are the indices of cells inside the domain, and $(l,

### 2015-01-15 papers/wdmerger_I/paper.tex
cite -> citealt
another tex-ism

@@
 transfer dynamics. However, there are also outstanding questions of
-convergence in SPH (e.g. \cite{zhu_SPH:2014}) and whether the method
+convergence in SPH (e.g.\ \citealt{zhu_SPH:2014}) and whether the method
 correctly captures fluid instabilities. This is an important question

### 2015-01-16 papers/wdmerger_I/paper.tex
some appendix updates

@@
 
-The basic PPM algorithm in CASTRO has undergone a number of changes since
-the original code paper \citep{castro}.  A discussion of the pure
-hydrodynamics changes was given in \citet{zingalekatz:2015}.  Here
-we discuss the changes that affect multispecies flow and source
-terms.
+The basic PPM algorithm in CASTRO has undergone a number of changes
+since the original code paper \citep{castro}.  A discussion of the
+pure hydrodynamics changes along with verification of CASTRO when
+using the stellar equation of state was given in
+\citet{zingalekatz:2015}.  Here we discuss the changes that affect
+multispecies flow and source terms.
 
@@
 
+\subsection{Reference States}
+
 We use the Castro code as described in \citet{castro}.  For all the
 runs, the PPM reconstruction is done, using the original limiters for
-the parabolic profiles \citep{ppm}.  We modify the prediction of the
-interface states slightly.  The original implementation of the PPM
-prediction in Castro takes the form:
-\begin{equation}
-q_{i+1/2,L}^{n+1/2} = q_i -
+the parabolic profiles \citep{ppm}.  The prediction of the interface
+states appears as:
+\begin{equation}
+\label{eq:ppmstatel}
+q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
    \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
-        q_i - \mathcal{I}_+^{(\nu)}(q_i)
+        \tilde{q}_L  - \mathcal{I}^{(\nu)}_+(q_i)
        \right ] r_i^{(\nu)}
 \end{equation}
-where $q$ is the vector of primitive variables, with $q_i$
-representing the average in the cell, $l^{(\nu)}$ and $r^{(\nu)}$ are
-the left and right eigenvectors with eigenvalue $\lambda^{(\nu)}$,
-with $\nu$ the index of the characteristic wave of the system.  The
-sum is over all the waves that result from the characteristic
-structure of the problem, but designed such that only waves moving
-toward the interface contribute to the interface value,
-$q_{i+1/2,L}^{n+1/2}$.  Finally, $\mathcal{I}_+^{(\nu)}(q)$ is the
-average under the parabolic profile of quantity $q$ of all the information
-that can reach the right interface of the zone $i$ as carried by the wave
-$\nu$.   The reader is referred to
+where $q$ is the vector of primitive variables, $l_i^{(\nu)}$ and
+$r_i^{(\nu)}$ are the left and right eigenvectors with eigenvalue
+$\lambda_i^{(\nu)}$, with $\nu$ the index of the characteristic wave of
+the system.  The sum is over all the waves that result from the
+characteristic structure of the problem, but designed such that only
+waves moving toward the interface contribute to the interface value,
+$q_{i+1/2,L}^{n+1/2}$.  The reference state, $\tilde{q}_L$ is
+chosen to minimize the work of the characteristic projection.
+Finally, $\mathcal{I}_+^{(\nu)}(q)$ is the
+average under the parabolic profile of quantity $q$ of all the
+information that can reach the right interface of the zone $i$ as
+carried by the wave $\nu$.  The reader is referred to
 \citet{ppmunsplit} for further defaults.
 
-\subsection{Reference States}
-
-The presence of $q_i$ in this expression serves as a reference state.
-The idea is that the error in the characteristic projection of the
-jump (due to the nonlinearity) is minimized if we pick a suitable
-reference state, since we only care about the jumps carried to the
-interface over the timestep \citep{colellaglaz1985}.  The original
-Castro implementation used the cell-average quantity.  Experiments
-found that this is ill-behaved in the presence of strong shocks
-(especially at low resolution). For the present work, we switch the
-reference state to
+Since the original CASTRO paper, the reference state implementation
+has been switched to:
 \begin{equation}
@@
 over the timestep.  This is
-in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
-expression appear as:
-\begin{equation}
-\label{eq:ppmstatel}
-q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
-   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
-        \tilde{q}_L  - \mathcal{I}^{(\nu)}_+(q_i)
-       \right ] r_i^{(\nu)}
-\end{equation}
-The final change is that now we can no longer simply multiply
-$\tilde{q_i} - \mathcal{I}^{(\nu)}_+(q_i)$ by the flattening
-parameter, $\chi$, but instead blend the traced state with the cell
-centered state as:
-  \begin{equation}
-  q_{i+1/2,\{L,R\}}^{n+1/2} \leftarrow (1 - \chi) q_i + \chi q_{i+1/2,\{L,R\}}^{n+1/2}
-  \end{equation}
-This is the same prescription used in the \cite{ppm}.
-
-This change is now the default in the public version of Castro,
-and can be controlled with the parameter {\tt castro.ppm\_reference}.
-Figure~\ref{Fig:sod} shows the
-solution to the Sod problem using the old and new reference states.
-We see that the shock and contact are slightly sharper with this new
-reference state, but also there is a slight dip at the tail of the
-rarefaction with the new method.
-
-It is instructive to look at the effect of the reference state choice.
+in agreement with \citet{ppmunsplit} (eq. 90).  The flattening
+in the original CASTRO paper has also been updated as discussed
+in \citet{zingalekatz:2015}.
+
+We comment on the choice of reference state for passively-advected
+quantities (like $X_k$ or the transverse velocity).
 First, consider one of the variables present in one-dimensional flow

### 2015-01-28 papers/wdmerger_I/paper.tex
Use the emulateapj class for this preprint.

@@
-\documentclass[12pt,preprint]{aastex}
+%\documentclass[12pt,preprint]{aastex}
+\documentclass{emulateapj}
 % for \sout
@@
 \usepackage{bm}
+
+% allows comment sections
+\usepackage{verbatim}
 
@@
 \shorttitle{DWD Mergers. I. Methodology}
-\shortauthors{Max}
+\shortauthors{Katz et al. (2015)}
 
@@
 \begin{align}
-  (\rho E_G)^{n+1} &= (\rho E_G)^{n} + \left(\frac{1}{2} (\rho \phi)^{n+1} - \frac{1}{2} (\rho \phi)^{n}\right) - \Delta t\, \bm{\nabla}\cdot\left(\rho \mathbf{u} E_{\text{tot}} + p\mathbf{u} - \frac{1}{2}\rho\mathbf{u}\phi\right) \notag \\
+  (\rho E_G)^{n+1} &= (\rho E_G)^{n} + \left(\frac{1}{2} (\rho \phi)^{n+1} - \frac{1}{2} (\rho \phi)^{n}\right) \notag \\
+   &- \Delta t\, \bm{\nabla}\cdot\left(\rho \mathbf{u} E_{\text{tot}} + p\mathbf{u} - \frac{1}{2}\rho\mathbf{u}\phi\right) \notag \\
    &+ \frac{\Delta t}{2}\left(\phi \frac{\partial \rho}{\partial t}\right)^{n+1/2} - \frac{\Delta t}{2} \left(\rho \frac{\partial \phi}{\partial t}\right)^{n+1/2}.
@@
 \begin{equation}
-  \Phi_{{lmn}} = G\, \sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{direct_sum}
-\end{equation} \MarginPar{should there really be 2 $G$'s?}
-Here $(i, j, k)$ are the indices of cells inside the domain, and $(l,
-m, n)$ \MarginPar{indices should be in normal math style, not roman}
+  \Phi_{{lmn}} = \sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{direct_sum}
+\end{equation}
+Here $(i, j, k)$ are the indices of cells inside the domain, and $(l,m, n)$ 
 are the indices of boundary locations. $\Delta V$ is the volume of the
@@
 the potential outside of the mass distribution can be written as:
-\begin{equation}
-  \Phi(\mathbf{x}) = \sum_{l=0}^{\infty} \left[Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} + \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right].\label{multipole_potential}
-\end{equation}
+\begin{align}
+  \Phi(\mathbf{x}) &= \sum_{l=0}^{\infty} \left\{Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} \right. \notag \\
+    &+ \left. \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right\}.\label{multipole_potential}
+\end{align}
 $P_l(x)$ are the Legendre polynomials and $P_{l}^{m}(x)$ are the associated Legendre polynomials.
@@
 where $M = 4\pi R^3 / 3$ is the mass of the sphere. Defining the $L^2$
-norm of a field $f$ as \MarginPar{We usually use $\| \ldots \|_2$ for a norm}
-\begin{equation}
-  L^2(f) = \left(\sum_{i,j,k} dx\, dy\, dz\, f_{ijk}^2\right)^{1/2},
+norm of a field $f$ as
+\begin{equation}
+  \| f \|_2 = \left(\sum_{i,j,k} dx\, dy\, dz\, f_{ijk}^2\right)^{1/2},
 \end{equation}
@@
 \begin{equation}
-  \text{Error} = \frac{L^2(\phi - \phi_{\text{sphere}})}{L^2(\phi_{\text{sphere}})}.
+  \text{Error} = \frac{\|\phi - \phi_{\text{sphere}}\|_2}{\|\phi_{\text{sphere}}\|_2}.
 \end{equation}
@@
 \begin{align}
-  \phi_{\text{cube}}(x,y,z) &= G\rho\sum_{i,j,k=0}^1\left(x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right) + y_j z_k\, \text{tanh}^{-1}\left(\frac{x_i}{r_{ijk}}\right) + z_k x_i\, \text{tanh}^{-1}\left(\frac{y_j}{r_{ijk}}\right) \right.\notag \\
-  &\left.
-- \frac{x_i^2}{2}\,\text{tan}^{-1}\left(\frac{y_j z_k}{x_i r_{ijk}}\right) - \frac{y_j^2}{2}\,\text{tan}^{-1}\left(\frac{z_k x_i}{y_j r_{ijk}}\right) - \frac{z_k^2}{2}\,\text{tan}^{-1}\left(\frac{x_i y_j}{z_k r_{ijk}}\right)\right)
+  \phi_{\text{cube}}(x,y,z) &= G\rho\sum_{i,j,k=0}^1\left[x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right)\right. \notag \\
+  &+ \left. y_j z_k\, \text{tanh}^{-1}\left(\frac{x_i}{r_{ijk}}\right) + z_k x_i\, \text{tanh}^{-1}\left(\frac{y_j}{r_{ijk}}\right) \right.\notag \\
+  &\left. - \frac{x_i^2}{2}\,\text{tan}^{-1}\left(\frac{y_j z_k}{x_i r_{ijk}}\right) - \frac{y_j^2}{2}\,\text{tan}^{-1}\left(\frac{z_k x_i}{y_j r_{ijk}}\right) \right. \notag \\
+  &- \left. \frac{z_k^2}{2}\,\text{tan}^{-1}\left(\frac{x_i y_j}{z_k r_{ijk}}\right)\right]
 \end{align}
@@
 the domain to be covered by the finest resolution grids.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=1.0]{freefall/plot_freefall}
+  \caption{Time evolution of two initially stationary white dwarfs,
+    mutually attracted to each other by the gravitational force. The
+    horizontal axis gives the separation of the white dwarfs, scaled
+    to the initial separation, and the vertical axis gives the elapsed
+    time of the simulation, scaled to the total elapsed time
+    necessarily. The solid curve shows the analytical result,
+    calculated from Newtonian mechanics, and the circles show the
+    samples from the time evolution with CASTRO. The agreement between
+    theory and simulation is excellent, even for a simulation with
+    modest resolution.}
+  \label{Fig:Free Fall}
+\end{figure}
 
@@
   R_A &= \begin{cases} 0 & |y - 0.5| > 0.25 \\ 1 & |y - 0.5| < 0.25 \end{cases} \\
-  R_B &= \Big\{\left[1 + \text{exp}(-2(y-0.25)/\Delta_y)\right]\left[1 + \text{exp}(2(y-0.75)/\Delta_y)\right]\Big\}^{-1} \\
-  v^\prime_A &= w_0\, \text{sin}\left(n_A\, \pi\, x\right) \left\{\text{exp}\left[-\frac{(y-0.25)^2}{2\sigma^2}\right] + \text{exp}\left[-\frac{(y-0.75)^2}{2\sigma^2}\right]\right\} \\
+  R_B &= \Big\{\left[1 + e^{-2(y-0.25)/\Delta_y}\right]\left[1 + e^{2(y-0.75)/\Delta_y}\right]\Big\}^{-1} \\
+  v^\prime_A &= w_0\, \text{sin}\left(n_A\, \pi\, x\right) \left\{e^{-(y-0.25)^2 / 2\sigma^2} + e^{-(y-0.75)^2/2\sigma^2}\right\} \\
   v^\prime_B &= w_0\, \text{sin}\left(n_B\, \pi\, x\right).
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[scale=2.0]{freefall/plot_freefall}
-  \caption{Time evolution of two initially stationary white dwarfs,
-    mutually attracted to each other by the gravitational force. The
-    horizontal axis gives the separation of the white dwarfs, scaled
-    to the initial separation, and the vertical axis gives the elapsed
-    time of the simulation, scaled to the total elapsed time
-    necessarily. The solid curve shows the analytical result,
-    calculated from Newtonian mechanics, and the circles show the
-    samples from the time evolution with CASTRO. The agreement between
-    theory and simulation is excellent, even for a simulation with
-    modest resolution.}
-  \label{Fig:Free Fall}
-\end{figure}
-
 \end{document}

### 2015-01-28 papers/wdmerger_I/paper.tex
Rewrite of the gravity source term section.

@@
 incorporated by the use of a source term for the momentum and energy
-equations. The integral form of the momentum source term is
-\begin{equation}
-  \int \rho \mathbf{g}\, dV
+equations. In a finite volume methodology, the momentum source term appears in 
+integral form as
+\begin{equation}
+  \left.\frac{\partial (\rho \mathbf{u})}{\partial t}\right|_{\text{grav}} = \frac{1}{\Delta V} \int \rho \mathbf{g}\, dV
 \end{equation}
@@
 \begin{equation}
-  \int \rho \mathbf{u}\cdot\mathbf{g}\, dV \label{eq:cell_center_gravity_source}.
-\end{equation}
+  \left.\frac{\partial (\rho E)}{\partial t}\right|_{\text{grav}} = \frac{1}{\Delta V} \int \rho \mathbf{u}\cdot\mathbf{g}\, dV \label{eq:cell_center_gravity_source}.
+\end{equation}
+Here $\Delta V$ is the cell's volume.
 In most hydrodynamics codes these are discretized as $\rho\,
-\mathbf{g}\, \Delta V$ and $\rho\, \mathbf{u}\,\cdot\mathbf{g}\,
-\Delta V$, respectively, where $\Delta V$ is the zone volume and
-$\rho$, $\mathbf{u}$, and $\mathbf{g}$ are evaluated at the zone
-center. As observed by \cite{arepo}, the essential problem with this
-approach is that fluid movements in a grid code are calculated using
+\mathbf{g}$ and $\rho\, \mathbf{u}\,\cdot\mathbf{g}$, respectively, 
+where $\rho$, $\mathbf{u}$, and $\mathbf{g}$ 
+are evaluated at the zone
+center.
+
+There are two ways that this source term enters the hydrodynamics update.
+First, during the hydrodynamics update, we alter the edge states that enter
+into the determination of the fluxes. To second order in space and time, 
+this can be done using the cell-centered
+source term evaluated at time-level $n$. The details of this are described in 
+the appendix. Second, after the hydrodynamics step, we add the time-centered source term
+to the state. First we will describe how we do this for the momentum,
+and then we will describe our approach for the energy. This discussion will be somewhat detailed.
+However, we believe that the attention is necessary because of the importance of accuracy
+in the gravitational source terms for our problem. The stability of the white dwarf binary
+system is dependent in large part upon accurate coupling of the hydrodynamics and gravity;
+an error in this approach could lead to, for example, a spurious mass transfer episode
+that might lead us to very different conclusions about the long term stability of such a system.
+Such considerations are generally unimportant for spherically-symmetric single star calculations,
+but are of the utmost importance in a simulation where the global gravitational field can change 
+quite significantly over the course of the simulation.
+
+In a system with self-gravity, total momentum is conserved if the spatial domain
+includes all of the mass of the system. This must be the 
+case because each mass element exerts an equal and opposite gravitational force 
+on every other mass element. However, the standard approach does not necessarily
+guarantee that momentum is conserved numerically. We cannot represent a vacuum state 
+in our code, so there is a small but non-zero density on the edge of the grid. 
+This allows momentum to leak out of the domain even if the gravitational source term 
+is written in an explicitly conservative manner. To see this, one can write the 
+density in terms of the potential and then consider its spatial discretization. For simplicity,
+we will consider one spatial dimension.
+\begin{align}
+  -\rho_{i}  \frac{d\Phi_{i}}{dx} &= -\frac{1}{4\pi G} \frac{d^2\Phi_i}{dx^2} \frac{d \Phi}{dx} \notag \\
+  &= -\frac{1}{4\pi G} \left[\frac{\Phi_{i-1} - 2 \Phi_{i} + \Phi_{i+1}}{\Delta x^2}\right] \left[ \frac{\Phi_{i+1} - \Phi_{i-1}}{2\Delta x} \right] \notag \\
+  &= -\frac{1}{8\pi G \Delta x^3} \left[ \Phi_{i+1}^2 - \Phi_{i-1}^2 - 2\Phi_i\left(\Phi_{i+1} - \Phi_{i-1}\right) \right] \label{eq:momentum_discretization}
+\end{align}
+It is easy to verify that adding the source terms two cells to the left
+and right of this zone will result in complete cancellation of the source terms.
+However, if the potential if non-zero outside of the domain, then there will be
+momentum lost or gained from the grid, which will be encapsulated in the ghost cells
+just outside the domain. In addition, when we replace the Laplacian above by the full
+three-dimensional stencil including the $y$ and $z$ derivatives, and depending on the
+discretization these will not be cancelled at all. This latter problem can be resolved by
+writing the momentum update in an explicitly conservative way.
+
+\citet[Chapter 4]{shu:1992} observes that it is possible to describe the source term 
+for the momentum equation by taking the divergence of a gravitational stress tensor,
+\begin{equation}
+  G_{ij} = -\frac{1}{4\pi G}\left(g_i g_j - \frac{1}{2}|\mathbf{g}|^2\delta_{ij}\right).
+\end{equation}
+The momentum equations are then written explicitly in conservative form.
+The flux at any zone boundary is added to one cell and
+subtracted from another, so that the total momentum in the domain interior stays constant to
+within numerical roundoff error. This result can be derived by analytically recasting 
+Equation \ref{eq:momentum_discretization}. In the continuum limit, the two momentum
+formulations are identical. Thus the latter has been advocated by, for example, 
+\cite{jiang:2013} for the ATHENA code. However, in a finite discretization 
+the divergence of the gravitational acceleration is no longer guaranteed to equal
+the zone density. In particular, we find that the mixing of the gravitational components
+means that the truncation error in the gravitational field can lead to large errors
+that imply a density much different than the zone's actual density. This is especially
+problematic in a simulation with a low-density ambient medium, where even a small error 
+in the momentum update can lead to large changes in a zone's momentum. By continuing to explicitly
+use the cell density in the momentum update, we can forgo this possibility: the size of the update
+will always be suitably small if the zone's density is small. Thus for our simulations
+we continue to use the standard source term for the momentum.
+
+Time centering of this source term is done in CASTRO using a predictor-corrector approach.
+At the start of a coarse grid timestep, we solve the gravitational potential for the density $\rho^n$.
+We then add to the momenta a prediction of the source term that is first-order accurate in time, 
+$\Delta t\, \rho^n\, \mathbf{g}^n$. After the hydrodynamics update, we recalculate
+the gravitational potential based on the new density, $\rho^n$, and then add 
+$-(\Delta t/2) \rho^n \mathbf{g}^n + (\Delta t/2) \rho^{n+1} \mathbf{g}^{n+1}$ to the momenta.
+
+Dealing with the energy equation is more difficult. The central
+challenge is to write down a form of the discretized energy equation
+that explicitly conserves total energy when coupled to gravity. When
+gravity is included, the total energy is
+\begin{equation}
+  \rho E_{\text{tot}} = \rho E - \frac{1}{2}\rho\Phi = \frac{1}{2}\rho \mathbf{u}^2 + \rho e - \frac{1}{2}\rho\Phi, \label{eq:total_energy_gravity}
+\end{equation}
+where $\rho E$ is the total gas energy from the pure hydrodynamics equation,
+and the factor of $1/2$ in the gravitational energy term is necessary
+for simulations with self-gravity to prevent double-counting of
+interactions. The standard approach has been to note that the total energy update
+can be written down by taking the dot product of the velocity and the momentum source term,
+so that the time-centered energy update is
+\begin{equation}
+  \left.\Delta (\rho E)\right|_\text{grav} = \frac{\Delta t}{2} \left[ \rho^n \mathbf{u}^n \cdot \mathbf{g}^n + \rho^{n+1} \mathbf{u}^{n+1} \cdot \mathbf{g}^{n+1}\right].
+\end{equation}
+However, this does not conserve the total energy of Equation \ref{eq:total_energy_gravity}.
+In particular, we notice in our simulations a long-term secular change in the gravitational potential energy
+that is not accounted for in the gas energy. One solution to this problem has been offered by 
+\cite{jiang:2013} and \cite{marcello:2012}. They propose to evolve the total energy instead of the gas energy 
+in the hydrodynamics equations. When this is done, one can write down an explicitly conservative 
+formulation:
+\begin{equation}
+  \frac{\partial \rho E_{\text{tot}}}{\partial t} = -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u} + \mathbf{F_g}).
+\end{equation}
+Here $\mathbf{F_g}$ is a flux term which explicitly accounts for the change in the total energy due to self-gravity. 
+By using Equation \ref{eq:cell_center_gravity_source} and the continuity equation, it is straightforward to show that
+\begin{equation}
+  {\bm\nabla} \cdot \mathbf{F_g} = -\rho \mathbf{u} \cdot \mathbf{g} - \frac{\partial}{\partial t}\left(\frac{1}{2}\rho\Phi\right). \label{eq:gradient_F_g}
+\end{equation}
+From here one can use the continuity equation to explicitly calculate the flux $\mathbf{F_g}$, as in \cite{jiang:2013}:
+\begin{equation}
+  \mathbf{F_g} = \frac{1}{8\pi G}\left(\Phi\bm{\nabla}\dot{\Phi} - \dot{\Phi}\bm{\nabla}\Phi\right) + \rho\mathbf{u}\Phi \label{eq:jiang_flux}.
+\end{equation}
+Here the dots refer to time derivatives. This has some attractive features. 
+In particular, for a simulation with AMR, a flux-based approach
+can quite naturally be accommodated in a way that conserves energy across levels, when fluxes
+at interfaces between coarse and fine grids are matched. We note in passing that although
+\citeauthor{jiang:2013} claim that this approach requires a Poisson solve for the time derivative of the potential,
+$\dot{\Phi}$, this is not needed as in practice we will already have the potential at time-levels
+$n$ and $n+1$, and to second order in time, $\dot{\Phi}^{n+1/2} = (\Phi^{n+1} - \Phi^{n}) / \Delta t.$
+
+Now, the first term on the right-hand side of Equation \ref{eq:gradient_F_g} is the old source term. The second term simply 
+accounts for the time rate of change of the gravitational potential energy. But since
+\begin{equation}
+  \frac{\partial(\rho E_{\text{tot}})}{\partial t} = \frac{\partial (\rho E)}{\partial t} + \frac{\partial (1/2) \rho \Phi}{\partial t},
+\end{equation}
+this does not actually add any information, since we can subtract off the time derivative 
+of the potential energy and be back where we started. This demonstrates that the problem
+with the original approach lies in the numerical discretization; if this is done correctly,
+we can still properly conserve total energy.
+
+Before pursuing that, we wish to point out the reason why we prefer not to use this 
+approach of solving a total energy equation. The expression for the flux $\mathbf{F_g}$
+given in \cite{jiang:2013} includes terms that are proportional to $\Phi$ and ${\bm{\nabla}} \Phi$,
+but that are not explicitly proportional to the zone density $\rho$. Thus the approach suffers
+from the same problems as the conservative momentum formulation discussed above. Truncation errors
+in the gravitational field will result in an approximation to the density that can be quite 
+different from the actual density, especially in low-density regions. In practice,
+we have found this to be such a serious problem that this approach is not feasible for our problem.
+The formulation of \cite{marcello:2012} does not suffer from the same problem. However, a 
+remaining problem is that in this formalism, the internal energy
+is calculated by subtracting the potential energy and the kinetic energy from the 
+total energy. Due to numerical discretization errors, however, this can result in a serious 
+error in the internal energy. This problem can be eased by the dual energy formalism,
+but not completely. This means that we will potentially have very large errors in a 
+zone's temperature, which can have serious repercussions in the reacting flows we will eventually consider,
+as many reaction rates are very sensitive to temperature. Therefore our goal is instead to consider an approach
+that is explicitly \textit{non}-conservative for the gas energy $(\rho E)$, but which
+instead conserves the total energy $(\rho E_{\text{tot}})$ over the computational domain 
+(subject, of course, to small errors due to energy leaking from the domain boundary).
+
+The question of how to do this was resolved by \cite{arepo}. 
+The essential problem with the standard approach is that 
+fluid movements in a grid code are calculated using
 fluxes at zone boundaries, while the standard gravity source term is
@@
 energy conservation. This becomes apparent in Section
-\ref{sec:kepler}. So while a closer look at this may be unwarranted
-for situations like a stationary single star central explosion
-calculation, a detailed discussion is certainly called for in the case
-of an orbiting binary system where getting the gravity right is
-absolutely necessary for fully trusting the long-term evolution of the
-system.
-
-\citet[Chapter 4]{shu:1992} observes that it is possible to describe the source term for the momentum equation by taking the divergence of a gravitational stress tensor,
-\begin{equation}
-  G_{ij} = -\frac{1}{4\pi G}\left(g_i g_j - \frac{1}{2}|\mathbf{g}|^2\delta_{ij}\right).
-\end{equation}
-The momentum equations are then written explicitly in conservative
-form when discretized: just as with the momentum stress tensor,
-divergences are calculated using discretized fluxes at zone
-boundaries, and the flux at any zone boundary is added to one cell and
-subtracted from another, so that the total momentum stays constant to
-within numerical roundoff error (with the possible exception of domain
-boundaries). This method works because, from the Poisson equation,
-$\rho$ can be written as proportional to the divergence of
-$\mathbf{g}$. We note that in order for the method to be reliable, the
-expression for all components of $\mathbf{g}$ must be correctly
-calculated from $\phi$ on zone boundaries so that the total divergence
-does indeed recover the density. This is straightforward for the
-component normal to the zone interface, but requires averaging over
-zones in the perpendicular direction to obtain the transverse
-components.
-
-Dealing with the energy equation is more difficult. The central
-challenge is to write down a form of the discretized energy equation
-that explicitly conserves total energy when coupled to gravity. When
-gravity is included, the total energy is
-\begin{equation}
-  \rho E_{\text{tot}} = E_G - \frac{1}{2}\rho\phi = \frac{1}{2}\rho \mathbf{u}^2 + \rho e - \frac{1}{2}\rho\phi,
-\end{equation}
-where $E_G$ is the gas energy from the pure hydrodynamics equation,
-and the factor of $1/2$ in the gravitational energy term is necessary
-for simulations with self-gravity to prevent double-counting of
-interactions. Recently, three different approaches have been proposed
-for achieving this. We will briefly examine these and then describe
-the approach that makes the most sense for us.
-
-\citet{marcello:2012} use the mass continuity equation to rewrite the
-standard source term from Equation
-\ref{eq:cell_center_gravity_source}. When this is done, the resulting
-evolution equation for the total energy is:
-\begin{equation}
-  \frac{\partial (\rho E_{\text{tot}})}{\partial t} = -\bm{\nabla}\cdot\left(\rho \mathbf{u} E_{\text{tot}} + p\mathbf{u} - \frac{1}{2}\rho\mathbf{u}\phi\right) + \frac{1}{2}\phi\frac{\partial \rho}{\partial t} - \frac{1}{2}\rho\frac{\partial \phi}{\partial t}.
-\end{equation}
-In practice, we would still actually evolve the total gas energy
-$E_G$, and so when discretized to second order in time from time level
-$n$ to time level $n+1$ the gas energy equation would look like
+\ref{sec:kepler}.
+
+We start with the integral formulation of the energy source term (Equation
+\ref{eq:cell_center_gravity_source}), but do not discretize it at
+the cell center.  Instead, we recognize that $\mathbf{u} = {\bm\nabla}\left[\mathbf{r}\cdot\mathbf{u}\right]$,
+where $\mathbf{r}$ is the position vector originating from the zone center.
+Then, since in the cell-centered approach we assume that the density, velocity,
+and gravitational acceleration are constant across the zone, we can include all terms
+inside the gradient operator. This allows us to use the divergence theorem to evaluate
+the quantities on the zone faces. To be second-order accurate in time, the fields must 
+all be evaluated at time-level $n+1/2$, but for clarity we will neglect the superscript.
 \begin{align}
-  (\rho E_G)^{n+1} &= (\rho E_G)^{n} + \left(\frac{1}{2} (\rho \phi)^{n+1} - \frac{1}{2} (\rho \phi)^{n}\right) \notag \\
-   &- \Delta t\, \bm{\nabla}\cdot\left(\rho \mathbf{u} E_{\text{tot}} + p\mathbf{u} - \frac{1}{2}\rho\mathbf{u}\phi\right) \notag \\
-   &+ \frac{\Delta t}{2}\left(\phi \frac{\partial \rho}{\partial t}\right)^{n+1/2} - \frac{\Delta t}{2} \left(\rho \frac{\partial \phi}{\partial t}\right)^{n+1/2}.
+  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{\Delta t}{\Delta V} \int \rho \mathbf{u}\cdot\mathbf{g}\, dV_i \notag \\
+     &= \frac{\Delta t}{\Delta V} \int \rho\, {\bm \nabla} \left[\mathbf{r} \cdot \mathbf{u}\right]\cdot\mathbf{g}\, dV_i \notag \\
+     &= \frac{\Delta t}{\Delta V} \int {\bm \nabla} \cdot \left[ \rho\, (\mathbf{r} \cdot \mathbf{u})\, \mathbf{g}\right]\, dV_i \notag \\
+     &= \frac{\Delta t}{\Delta V} \int \rho\, (\mathbf{r} \cdot \mathbf{u})\, \mathbf{g} \cdot \mathbf{dS}_i.
 \end{align}
-It is not immediately obvious that this is explicitly conservative for
-the total energy, due to the last two terms on the right-hand side,
-but written in discrete form the potential is linked to the density in
-such a way that there is indeed a perfect cancellation of terms
-between adjacent cells (see Equation 88 in
-\citeauthor{marcello:2012}). Doing it this way requires a prescription
-for calculating $\phi^{n+1/2}$, the potential at the time halfway
-between the new and old state, as well as $\dot{\phi}^{n+1/2}$, the
-time derivative of the potential at that time. To second order in
-time, these can be calculated as $(\phi^{n+1} + \phi^{n})/2$ and
-$(\phi^{n+1}-\phi^{n})/\Delta t$ respectively. So, in practice what we
-do is to first solve the Poisson equation using $\rho^n$ at the
-beginning of a time step; then, evolve the pure hydrodynamics
-equations without the gravitational source to obtain $\rho^{n+1}$;
-then, solve the Poisson equation for $\phi^{n+1}$; and then to finally
-calculate the gravitational source terms and add them in. An
-alternative approach is to use the predictor-corrector scheme that
-CASTRO features, where the time level $n$ source term is included in
-the hydrodynamics update and then left out of the final update. In
-either case, the time level $n$ calculation of $\mathbf{g}$ is
-included the PPM reconstruction of the edge states.
-
-\citet{jiang:2013} start from the motivation of extending to the
-energy equation the flux-based formalism that allows explicit
-conservation for the momentum equation. They determine that the energy
-is perfectly conserved if we define a gravitational flux
-$\mathbf{F_g}$ and take its divergence for the energy equation:
+Here $\mathbf{dS}$ is a surface element whose vector points normal to
+and away from the surface. In one dimension, this expression evaluates to
 \begin{align}
-  \mathbf{F_g} &= \frac{1}{8\pi G}\left(\phi\bm{\nabla}\dot{\phi} - \dot{\phi}\bm{\nabla}\phi\right) + \rho\mathbf{u}\phi \label{eq:jiang_flux}\\
-  \frac{\partial (\rho E_G)}{\partial t} &= -\bm{\nabla} \cdot \left(\rho\mathbf{u}E_G + p\mathbf{u} + \mathbf{F_g}\right)
+  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{\Delta t}{\Delta V} \rho_{i-1/2}\, \frac{\Delta x}{2} u_{x_{i-1/2}} g_{x_{i-1/2}} \Delta A_{i-1/2} \notag \\
+             &+ \frac{\Delta t}{\Delta V} \rho_{i+1/2}\, \frac{\Delta x}{2} u_{x_{i+1/2}} g_{x_{i+1/2}} \Delta A_{i+1/2} \notag \\
+             &= \frac{1}{2} \rho_{i-1/2} u_{x_{i-1/2}} g_{x_{i-1/2}} \notag \\
+             &+ \frac{1}{2} \rho_{i+1/2} u_{x_{i+1/2}} g_{x_{i+1/2}}.\label{eq:edge_centered_gravity_source}
 \end{align}
-This is analytically equivalent to the approach by
-\citet{marcello:2012}, which can be seen by taking the divergence of
-the flux and using the Poisson equation. It would be justified to use
-this form simply because it is cleaner to write down, but there is a
-more important benefit to using an explicit flux based formula: in
-AMR, conservation between levels can only happen if all fluid
-movements are written down using fluxes, because it is those fluxes
-that are adjusted at coarse-fine interfaces to ensure conservation. In
-this form we can simply add the gravitational source term to the
-hydrodynamical fluxes and use the existing AMR algorithm to guarantee
-conservation between levels. We note that \citeauthor{jiang:2013} say
-that this method requires the solution of an additional Poisson
-equation to obtain the intermediate time gravitational potential, but
-as we pointed out above, this can be done to second order simply by
-taking a centered difference using the existing Poisson solves, so no
-extra Poisson solve is necessary.
-
-This flux-based formalism has a significant drawback for our problem:
-unlike \citeauthor{jiang:2013}, we use open hydrodynamic boundary
-conditions, not periodic boundary conditions. Mass, momentum and
-energy can be lost (or gained) across the domain boundary. This is not
-a serious source of error in the standard cell-centered
-discretization, because the source term is proportional to $\rho$ and
-the density near the domain boundary is generally filled with a very
-low density ambient medium, so the size of the source term is always
-small. However, the gravitational flux in Equation \ref{eq:jiang_flux}
-has terms that are not explicitly proportional to
-density. Consequently, the flux lost across the domain boundaries can
-be significant (though this effect diminishes with increasing
-resolution). A related problem is that at the edge of a star, the
-gravitational flux can be so large that the gas energy becomes
-negative after an update. This effect is greatly diminished if the
-flux is instead proportional to density, as in the standard
-method. This is not a major problem for ideal or polytropic gases in
-pure hydrodynamics problems, but in our simulations we want the
-internal energy (and thus the temperature) to always be reliable so
-that nuclear reactions occur at the correct rates. This is especially
-sensitive at the edges of the stars, where the burning is expected to
-occur in a violent merger scneario.
-
-A possible compromise approach is discussed in \cite{arepo}. We start
-with the integral formulation of the energy source term (Equation
-\ref{eq:cell_center_gravity_source}), but instead of discretizing at
-the cell center, we use Gauss' theorem to evaluate the source term at
-the zone boundaries (see Equations 91--96 of that paper; we exclude
-the term related to the motion of the zones since our grids are
-fixed). This results in a gravitational update for zone $i$ written as
-\begin{equation}
-  \Delta E^{\text{grav}}_{i} = -\frac{1}{2}\sum_{j} \Delta m_{ij}\, \mathbf{r}_{ij} \cdot \bm{\nabla}\phi_i \approx -\frac{1}{2}\sum_{j} \Delta m_{ij}\left(\phi_i - \phi_j\right),
-\end{equation}
-where the sum is over all $j$ zones that share a boundary with zone
-$i$. Here $\mathbf{r}_{ij}$ is the radius vector joining zones $i$ and
-$j$ and $\Delta m_{ij}$ is the mass exchanged between zones $i$ and
-$j$ in the hydrodynamics update, and is equal to $\Delta t A_{ij}
-F_\rho^{ij}$, where $A_{ij}$ is the area of the zone boundary and
-$F_{\rho}^{ij}$ is the mass flux from the mass continuity equation. It
-is straightforward to see that the second version is explicitly
-conservative. The benefit of this form is that the mass exchanged
-between zones near the domain boundary will always be small in
-absolute terms since these zones are at very low density. Thus the
-energy lost through the boundary will be of comparable size to the
-standard method, but the gravity update should be more accurate since
-the change in gravitational energy now tracks the explicit motion of
-mass on the grid.
-
-% Now rewrite this so that it is in flux form
+(We note that this differs slightly from Equation 94 in \cite{arepo} in that
+we evaluate the gravitational acceleration at the zone face rather than the zone center.)
+In the limit of constant fields, this expression simply averages out to the standard 
+source term at the zone center. However, when the fields vary this can give very different results.
+
+As pointed out by \citeauthor{arepo}, $(\rho u)$ evaluated at the faces at time level $n+1/2$ is the mass flux,
+so $(\rho u)_{i-1/2}\, \Delta t\, \Delta A = \Delta m_{i-1/2}$ is the mass exchanged between zones $i$ and $i-1$.
+Additionally, to second order $g_{x_{i-1/2}} \Delta x = \Phi_{i} - \Phi_{i-1}$. Therefore the energy gained by the zone is
+\begin{align}
+  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{1}{2} \Delta m_{i-1/2} \left(\Phi_{i} - \Phi_{i-1}\right) \notag \\
+     &+ \frac{1}{2} \Delta m_{i+1/2} \left(\Phi_{i+1} - \Phi_{i}\right)
+\end{align}
+This expression says that the energy gained by a zone is due to a mass element moving through a potential gradient
+between two zones. Now, the key observation is that this does not have the telescoping property of a conservative flux;
+instead, the \textit{same} amount of gravitational potential energy is gained by zones $i$ and $i-1$ by the mass flux 
+between them. When this source term is summed up over all zones, it will exactly compensate for the change in 
+gravitational potential energy, except for any energy lost through the boundaries. The extension of this energy change 
+to multiple dimensions simply means that we include the mass fluxes and potential gradients across the $y$ and $z$ faces.
+We have found that in practice this method for conserving the total energy is the least susceptible to deleterious effects
+from numerical discretization errors. However, it does mean that there can be sharp changes in the total energy of a zone
+in regions near a white dwarf's surface, because the mass fluxes into the ambient medium can be quite large compared to
+the mass of the zone. This effect diminishes at higher resolutions as the gradients become better resolved.
 
@@
 onto our grid, and used CASTRO to calculate the gravitational
-potential $\phi$ of this setup. We ensure that $R$ is an integer
+potential $\Phi$ of this setup. We ensure that $R$ is an integer
 multiple of the grid spacing, and the center of the sphere is at the
@@
 \begin{equation}
-  \phi_{\text{sphere}}(r) = \frac{GM}{r} \times \begin{cases} (3R^2 - r^2)/(2 r^2) & r \leq R \\ 1 & r > R \end{cases},\label{eq:sphere-analytical}
+  \Phi_{\text{sphere}}(r) = \frac{GM}{r} \times \begin{cases} (3R^2 - r^2)/(2 r^2) & r \leq R \\ 1 & r > R \end{cases},\label{eq:sphere-analytical}
 \end{equation}
@@
 \begin{equation}
-  \text{Error} = \frac{\|\phi - \phi_{\text{sphere}}\|_2}{\|\phi_{\text{sphere}}\|_2}.
+  \text{Error} = \frac{\|\Phi - \Phi_{\text{sphere}}\|_2}{\|\Phi_{\text{sphere}}\|_2}.
 \end{equation}
@@
 \begin{align}
-  \phi_{\text{cube}}(x,y,z) &= G\rho\sum_{i,j,k=0}^1\left[x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right)\right. \notag \\
+  \Phi_{\text{cube}}(x,y,z) &= G\rho\sum_{i,j,k=0}^1\left[x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right)\right. \notag \\
   &+ \left. y_j z_k\, \text{tanh}^{-1}\left(\frac{x_i}{r_{ijk}}\right) + z_k x_i\, \text{tanh}^{-1}\left(\frac{y_j}{r_{ijk}}\right) \right.\notag \\
@@
 The tracing of gravity can be controlled in the public version of Castro
-with the parmeter {\tt castro.ppm\_trace\_grav}.
-
-\subsection{Gravity Source Correction}
-
-We also modify slightly the correction done to the gravitational
-source terms after integrating the system in time.  In the original
-Castro paper (Eqs. 19 and 20), the state after the conservative update
-due to fluxes and external sources (other than gravity) was
-represented with the superscript `$(2,\star)$'.  The gravitational
-sources were then corrected, making them centered in time, yielding
-the final state at the end of the hydrodynamics step, indicated with
-the superscript `$(2)$'.  This appeared as:
-\begin{eqnarray}
-(\rho {\bf U})^{(2)} &=& (\rho {\bf U})^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf g})^{(2,\star)} - 
-                              (\rho {\bf g})^{(1)} \right ]\\
-(\rho E)^{(2)} &=& (\rho E)^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf U\cdot g})^{(2,\star)} - 
-                              (\rho {\bf U\cdot g})^{(1)} \right ]
-\end{eqnarray}
-Here we make the subtle change of using the corrected ${\bf u}$ in the
-correction to the energy.  So our new energy correction equation appears as:
-\begin{equation}
-(\rho E)^{(2)} = (\rho E)^{(2,\star)} +
-   \frac{\Delta t}{2} \left [ (\rho {\bf U\cdot g})^{(2)} - 
-                              (\rho {\bf U\cdot g})^{(1)} \right ]
-\end{equation}
-Both of these are second-order accurate, but we've found that the
-latter is slightly better for conserving energy.  This change is now
-the default in the public version of Castro (it can be controlled by
-{\tt castro.grav\_source\_type}.
-
-Rotation is handled similarly to gravity during the conservative update.
-The momentum equation, with rotation sources appears as:
-\begin{equation}
-\frac{\partial \rho {\bf U}}{\partial t} + \nabla \cdot (\rho {\bf U U} + p) = 
-   \rho {\bf g} - 2 \rho {\bf \Omega \times U} 
-                - \rho {\bf \Omega \times (\Omega \times U)}
-\end{equation}
-we take the rotation axis to be in the $z$-direction, ${\bf \Omega} =
-\Omega \hat{\bf k}$.  Working out the cross-products, the rotation
-forces appear as
-\begin{equation}
-{\bf F}_\mathrm{rot} = \rho \left (
-    \begin{array}{c}
-     \phantom{-}2 \Omega v + \Omega^2 x \\
-              - 2 \Omega u + \Omega^2 y \\
-              0 
-    \end{array}
-  \right )
-\end{equation}
-We want to time-center this force in the final conservative update, giving
-updates for the $u$ and $v$ components of the velocity as:
-\begin{eqnarray}
-\frac{(\rho u)^{n+1} - (\rho u)^n}{\Delta t} + A^{(x),n+1/2} &=&
-   \frac{1}{2} \left [ (\rho {\bf g}\cdot \hat{\bf i})^n
-                      + (\rho {\bf g}\cdot \hat{\bf i})^{n+1} \right ] \nonumber\\&+&
-   \frac{1}{2} \left [ (2\rho \Omega v)^n + (2\rho \Omega v)^{n+1} \right ] \nonumber \\&+&
-   \frac{1}{2} \left [ (\rho \Omega^2 x)^n + (\rho \Omega^2 x)^{n+1} \right ] \\
-\frac{(\rho v)^{n+1} - (\rho v)^n}{\Delta t} + A^{(y),n+1/2} &=&
-   \frac{1}{2} \left [ (\rho {\bf g}\cdot \hat{\bf j})^n +
-                       (\rho {\bf g}\cdot \hat{\bf j})^{n+1} \right ] \nonumber \\ &+&
-   \frac{1}{2} \left [ (-2\rho \Omega u)^n + (-2\rho \Omega u)^{n+1} \right ] \nonumber \\ &+&
-   \frac{1}{2} \left [ (\rho \Omega^2 y)^n + (\rho \Omega^2 y)^{n+1} \right ]
-\end{eqnarray}
-where here the explicit advective term from the Godunov method is
-represented as $A$ and is already time-centered.  In the Castro
-algorithm, a full $\Delta t$ of the old-time-level source is first
-added and later we correct this by subtracting off $\Delta t/2$ of the
-old-time-level term and adding $\Delta t/2$ of the new-time-level
-source.  This correction appears as:
-\begin{eqnarray}
-(\rho u)^{n+1} = (\rho u)^\prime 
-   &+& \underbrace{
-      \frac{\Delta t}{2} \left [ (\rho {\bf g}\cdot \hat{\bf i})^{n+1}
-                               - (\rho {\bf g}\cdot \hat{\bf i})^{n} \right ]
-    }_{\equiv \Delta^{g,x}} \nonumber \\
-   &+& \frac{\Delta t}{2} \left [ (2\rho \Omega v)^{n+1} 
-                                - (2\rho \Omega v)^{n} \right ] \nonumber \\
-   &+& \underbrace{
-      \frac{\Delta t}{2} \left [ (\rho \Omega^2 x)^{n+1} 
-                               - (\rho \Omega^2 x)^{n} \right ]
-    }_{\equiv \Delta^{c,x}} \\
-%
-(\rho v)^{n+1} = (\rho v)^\prime 
-   &+& \underbrace{
-      \frac{\Delta t}{2} \left [ (\rho {\bf g}\cdot \hat{\bf j})^{n+1}
-                               - (\rho {\bf g}\cdot \hat{\bf j})^{n} \right ]
-    }_{\equiv \Delta^{g,y}} \nonumber \\
-   &+& \frac{\Delta t}{2} \left [ (-2\rho \Omega u)^{n+1} 
-                                - (-2\rho \Omega u)^{n} \right ] \nonumber \\
-   &+& \underbrace{
-      \frac{\Delta t}{2} \left [ (\rho \Omega^2 y)^{n+1} 
-                               - (\rho \Omega^2 y)^{n} \right ]
-    }_{\equiv \Delta^{c,y}} \\
-\end{eqnarray}
-For compactness, we define the following:
-\begin{eqnarray}
-S_x &=& (\rho u)^\prime + \Delta^{g,x} + \Delta^{c,x} 
-        - \frac{\Delta t}{2} (2 \rho \Omega v)^n \\
-%
-S_y &=& (\rho v)^\prime + \Delta^{g,y} + \Delta^{c,y} 
-        + \frac{\Delta t}{2} (2 \rho \Omega u)^n 
-\end{eqnarray}
-and our updates are now a coupled implicit system for the updates of
-$u$ and $v$:
-\begin{eqnarray}
-(\rho u)^{n+1} = S_x + \frac{\Delta t}{2}(2\rho \Omega v)^{n+1} \label{eq:rhouupdate}\\
-(\rho v)^{n+1} = S_y - \frac{\Delta t}{2}(2\rho \Omega u)^{n+1} 
-\end{eqnarray}
-This can be solved analytically:
-\begin{equation}
-(\rho v)^{n+1} = \frac{S_y - \Delta t \Omega S_x}{1 + \Delta t^2 \Omega^2}
-\end{equation}
-and then we can find $(\rho u)^{n+1}$ from Eq.~\ref{eq:rhouupdate}.
-\clearpage
+with the parameter {\tt castro.ppm\_trace\_grav}.
 

### 2015-01-28 papers/wdmerger_I/paper.tex
Add description to the paper of the rotation coupling to the hydrodynamics.

@@
 two stars in a frame that is co-rotating at the same period as the
-orbital period. CASTRO has the ability to evolve systems in a rotating
-reference frame, as discussed in the original code paper. Source terms
-corresponding to the Coriolis and centrifugal force terms are added to
-the momentum and energy equations. In this frame, the stars
-essentially remain stationary in their original positions due to the
+orbital period. Since the publication of the original code paper, CASTRO 
+now has the ability to evolve systems in a rotating reference frame. 
+Source terms corresponding to the Coriolis and centrifugal 
+force terms are added to the momentum and energy equations. In this frame, 
+the stars essentially remain stationary in their original positions due to the
 centrifugal force supporting against the gravitational attraction, and
@@
 literature that grid-based codes poorly conserve angular momentum,
-this is not generally true. When the resolution is sufficiently high
+it is only generally true that grid-based codes do not exactly conserve 
+angular momentum when the equations are written in conservative form
+for linear momentum. (See also \cite{motl:2002} for an example of how 
+to evolve the hydrodynamics equations for angular momentum.) 
+The extent to which angular momentum conservation is violated
+will be a function of the resolution. When this is sufficiently high
 and a rotating reference frame is employed, excellent conservation
-properties can result, as demonstrated in Section \ref{sec:kepler}. We
-note that as the stars begin to coalesce, the rotating reference frame
+properties can result, as demonstrated in Section \ref{sec:kepler}. 
+In particular, at reasonable resolution for a binary orbit our code 
+conserves angular momentum well enough that this no longer becomes 
+a reason to use a Lagrangian method such as smoothed-particle hydrodynamics.
+We note that as the stars begin to coalesce, the rotating reference frame
 will no longer provide a good approximation to the spatial motion of
@@
 important.
+
+In a rotating reference frame with angular frequency vector $\bm{\omega}$, the momentum equation becomes:
+\begin{equation}
+  \left.\frac{\partial(\rho \mathbf{u})}{\partial t}\right|_{\text{rot}} = -2\, {\bm\omega} \times (\rho\mathbf{u}) - \rho {\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right).
+\end{equation}
+Here $\mathbf{r}$ is the position vector with respect to the origin. Typically we choose $\bm{\omega} = (0, 0, 2\pi / T)^T$,
+with the rotation axis coincident with the $z$ axis at $x = y = 0$.
+$T$ is the rotation period, which is the most natural quantity to specify
+for a rotating stellar system. As described in the appendix, we include this source term
+in the edge state prediction in a way that is analogous to the gravity source.
+We evaluate all quantities at cell centers and use the same predictor-corrector 
+approach that we use for the gravity source terms to the momentum equations.
+
+The update to the energy equation can be determined by taking the dot product of the velocity
+with the momentum source terms. The Coriolis term vanishes identically, as it must
+because the Coriolis term does no work on the fluid. The update from the centrifugal force becomes
+\begin{equation}
+  \left.\frac{\partial(\rho E)}{\partial t}\right|_{\text{rot}} = \frac{1}{\Delta V}\int \rho \mathbf{u} \cdot \mathbf{f}_R\, dV,
+\end{equation}
+with $\mathbf{f}_R \equiv  -{\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right)$. 
+This expression is identical to the gravity source under the interchange of $\mathbf{g}$ with $\mathbf{f}_R$.
+As observed by \cite{marcello:2012}, we can similarly write down a rotational potential,
+\begin{equation}
+  \Phi_R = -\frac{1}{2} \left| {\bm\omega} \times \mathbf{r} \right|^2.
+\end{equation}
+Therefore we can retrace the steps of Section \ref{sec:gravity_hydro_coupling} to obtain
+an update to $(\rho E)$ that conserves the total energy in the presence of rotation,
+\begin{equation}
+  (\rho E_{\text{tot}}) = (\rho E) + \frac{1}{2} \rho \Phi + \rho \Phi_R.
+\end{equation}
+In one dimension, this would be
+\begin{align}
+  \left.\Delta(\rho E_i)\right|_{\text{rot}} &= \frac{1}{2} \Delta m_{i-1/2} (\Phi_{R_{i}} - \Phi_{R_{i-1}}) \notag \\
+  &+ \frac{1}{2} \Delta m_{i+1/2} (\Phi_{R_{i+1}} - \Phi_{R_{i}}).
+\end{align}
+As with gravity, in three dimensions we also include the mass fluxes through all six cell faces.
 
@@
 
-We use the Castro code as described in \citet{castro}.  For all the
-runs, the PPM reconstruction is done, using the original limiters for
+For all the runs, the PPM reconstruction is done using the original limiters for
 the parabolic profiles \citep{ppm}.  The prediction of the interface
@@
 
-\subsection{Gravity Tracing}
+\subsection{Gravity and Rotation Tracing}
 
 We note a few additional differences between the original PPM
-implementation of \citet{ppm} and Castro.  In the original PPM
-implementation, the gravitational was reconstructed as a parabola, and
+implementation of \citet{ppm} and CASTRO.  In the original PPM
+implementation, the gravitational acceleration was reconstructed as a parabola, and
 this was traced under to find the forcing that affects the interface
-for each wave.  Castro followed \citet{ppmunsplit} which instead adds
+for each wave.  CASTRO followed \citet{ppmunsplit} which instead adds
 $(\Delta t/2)g$ to the interface states at the end of the
@@
 
-The tracing of gravity can be controlled in the public version of Castro
+The tracing of gravity can be controlled in the public version of CASTRO
 with the parameter {\tt castro.ppm\_trace\_grav}.
+
+Similarly, we also trace under the centrifugal and Coriolis forces to capture
+the effects of rotation on the edge states. For the primitive variables, the rotation
+source term is $F_R = (0, f_R, 0)^T$, where $f_R = -2\, {\bm{\omega}} \times \mathbf{u} - {\bm\omega}\times \left({\bm\omega} \times \mathbf{r}\right).$
+Just like the gravity source, the rotation source terms only update the velocities. Therefore 
+the rotation source terms can be included in a manner identical to that of the gravity source
+terms, with the replacement of $G$ with $F_R$ and $g$ with $f_R$ in the above. This is controlled in CASTRO
+with the parameter {\tt castro.ppm\_trace\_rot}.
 

### 2015-01-28 documentation/userguide.tex
Create a user's guide file that is compiled in LaTeX.

\documentclass[12pt]{book}

\usepackage[margin=1.0in]{geometry}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{epsfig}

\usepackage{natbib}

% chapter title styles
\usepackage[Sonny]{fncychap}
\ChNameVar{\LARGE}
\ChTitleVar{\LARGE\sl}

% part page style see
% http://tex.stackexchange.com/questions/6609/problems-with-part-labels-using-titlesec
\usepackage{titlesec}

\titleformat{\part}[display]
   {\Huge\filcenter}
   {{\partname{}} \thepart}
   {0em}
   {\hrule}

% hyperlinks -- load after fncychap
\usepackage{hyperref}

% color package
\usepackage[usenames]{color}

% number subsubsections and put them in the TOC
\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}

% custom hrule for title page
\newcommand{\HRule}{\rule{\linewidth}{0.125mm}}

% short table of contents
\usepackage{shorttoc}

% spacing in the table of contents
\usepackage[titles]{tocloft}

\setlength{\cftbeforechapskip}{2ex}
\setlength{\cftbeforesecskip}{0.25ex}

% don't put a header on blank pages, see
% http://www.latex-community.org/forum/viewtopic.php?f=4&p=51559
% change ``plain'' to ``empty'' to eliminate the page number
\makeatletter
\renewcommand*\cleardoublepage{\clearpage\if@twoside
\ifodd\c@page\else
\hbox{}
\thispagestyle{empty}
\newpage
\if@twocolumn\hbox{}\newpage\fi\fi\fi}
\makeatother

% don't make the chapter/section headings uppercase.  See the fancyhdr
% documentation (section 9)
\usepackage{fancyhdr}
\renewcommand{\chaptermark}[1]{%
 \markboth{\chaptername
\ \thechapter.\ #1}{}}

\renewcommand{\sectionmark}[1]{\markright{\thesection---#1}}

% skip a bit of space between paragraphs, to enhance readability
\usepackage{parskip}

% special fraction
\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

% codes
\newcommand{\castro}{{\sf Castro}}
\newcommand{\boxlib}{{\sf BoxLib}}
\newcommand{\yt}{{\sf yt}}

%------------------------------------------------------------------------------
\begin{document}

\frontmatter

\begin{titlepage}
\begin{center}
\ \\[3in]
{\sf \Huge WDMERGER}

\begin{minipage}{5.5in}
\HRule\\[2mm]
\centering
{\Large \em A software package for simulating white dwarf mergers with CASTRO}

\HRule
\end{minipage}

\ \\[1 in]
{\sf \huge User's Guide}

\vfill

{\large \today}
\end{center}

\end{titlepage}

\shorttoc{Chapter Listing}{0}

\setcounter{tocdepth}{2}
\tableofcontents

\clearpage

\addcontentsline{toc}{chapter}{preface}

\clearpage

\mainmatter

\chapter{Getting Started}

Welcome to the user's guide for the wdmerger software package. This software is designed to
simulate binary white dwarf systems, and is intended to provide useful information on the 
viability of mergers of white dwarfs as a progenitor for Type Ia supernovae. It is a set of 
source files designed to use the CASTRO code \citep{castro}.

%------------------------------------------------------------------------------
\backmatter

\bibliographystyle{../papers/apj}
\bibliography{../papers/refs}

\end{document}

### 2015-01-28 documentation/userguide.tex
Include required software in the documentation.

@@
 
+This software requires a few other software packages to be installed before it can be used.
+The science software includes \href{https://ccse.lbl.gov/BoxLib/}{BoxLib} and 
+\href{https://ccse.lbl.gov/Downloads/downloadCASTRO.html}{CASTRO}. This software is written 
+in C++ and Fortran, and compilers for each are needed. The gcc compilers are sufficient.
+In addition, it relies on several utilities that are 
+commonly distributed on Linux systems: \texttt{bash}, \texttt{GNU bc}, \texttt{GNU make}, 
+\texttt{batch}, \texttt{GNU sed}, and \texttt{GNU awk}. Analysis is mainly performed in 
+\href{https://www.python.org/}{python}, with the \href{http://matplotlib.org/}{matplotlib} 
+library and the \href{http://yt-project.org/}{yt} code being used for visualization. 
+If python and yt are not installed on your system, you will still be able to compile and 
+run this code, but you will need some other mechanism for analyzing the BoxLib output data.
+We have some routines for this purpose available in Fortran and may be able to provide them
+upon request.
+
 %------------------------------------------------------------------------------

### 2015-01-30 papers/wdmerger_I/paper.tex
Move the bibliography file up to the main papers subdirectory so that it can be included in any file that needs references.
Do the same for the apj bibliography style file.
Add discussion to paper of starting timestep size.

@@
 as we are not yet attempting to study the merger process.
+
+As a consequence of starting in a non-equilibrium setting, there are 
+large density and pressure gradients near the white dwarf surfaces
+that result in large amounts of mass flowing into the ambient
+medium. This can result in spurious non-physical consequences such as 
+the total density or energy going negative in a zone. To compensate 
+for this, we start the simulation with a timestep that is a few orders 
+of magnitude smaller than that required by the CFL criterion, and allow
+the timestep to increase by 1\% each timestep so that the timestep reaches 
+its maximum allowed by the velocities on the grid over a span of approximately 
+1000 timesteps. This allows the ambient medium just outside the white dwarf
+to come closer to equilibrium with the surface without having 
+discontinuous jumps in the density or energy. For all simulations, 
+the maximum timestep is set to be equal to one-half of the CFL limit.
 
@@
 
-\bibliographystyle{apj}
-\bibliography{refs}
+\bibliographystyle{../apj}
+\bibliography{../refs}
 

### 2015-01-30 papers/wdmerger_I/paper.tex
First draft of an abstract.

@@
 \begin{abstract}
-We describe our numerical methodology for modeling double white dwarf
-systems with the AMR hydrodynamics code Castro.
+The Type Ia supernova progenitor problem is one of the most perplexing and 
+exciting problems in astrophysics, requiring detailed numerical modelling to 
+complement the observations of these explosions. One possibility that has 
+merited recent theoretical attention is the white dwarf merger scenario.
+This can naturally explain many of the observed characteristics of 
+Type Ia supernovae, as well as their distribution in time and space.
+However, to date there have been few fully self-consistent simulations 
+of binary white dwarf systems using mesh-based hydrodynamics, 
+relative to other methods. This is the first paper in a series designed to 
+describe simulations of these systems with the adaptive-mesh refinement 
+hydrodynamics code CASTRO. In this paper we describe our numerical 
+methodology. CASTRO solves the Euler equations of hydrodynamics 
+and the Poisson equation for self-gravity. Standard techniques for 
+coupling gravitation and rotation forces to the hydrodynamics do 
+not do an adequate job conserving the total energy of the system, 
+but recent advances in the literature have made advances on this 
+problem and we discuss our implementation here. We also run an 
+extensive set of test problems to verify that our software can sufficiently
+model a system where large amounts of mass are advected on the computational 
+domain over very long timescales. Future papers in this series will describe
+our treatment of the initial conditions of these systems, and will 
+examine the early phases of the merger to determine its viability
+for triggering a thermonuclear detonation.
 

### 2015-01-30 documentation/userguide.tex
Added a description of the test problems to the user's guide.

@@
 
+This software assumes that you have set several environment variables:
+\begin{itemize}
+  \item \texttt{BOXLIB\_HOME}, the location of the BoxLib code's top-level directory
+  \item \texttt{CASTRO\_DIR}, the location of the CASTRO code's top-level directory
+  \item \texttt{WDMERGER\_HOME}, the location of this software's top-level directory
+\end{itemize}
+
+In addition, you should include in your \texttt{PYTHONPATH} the location of the 
+\texttt{analysis} subdirectory. In \texttt{bash}, include the following in your 
+\texttt{.bashrc} profile file:
+
+\texttt{export PYTHONPATH=\$PYTHONPATH:\$WDMERGER\_HOME/analysis}
+
+\chapter{Code Structure}
+
+There are several top-level directories in the \texttt{wdmerger} project. Here we will describe
+the contents of each. There are also a \texttt{LICENSE.txt} file containing the software license,
+and a \texttt{README.md} file describing how to get started.
+
+The \texttt{documentation} directory contains this user's guide and is intended to provide information
+on how to use and run this software.
+
+The \texttt{source} directory contains all of the Fortran routines needed to build a CASTRO 
+executable file. It is not normally built directly, but you can type \texttt{make} in that directory
+to build a vanilla executable built with the default settings.
+
+The \texttt{analysis} directory contains the Python analysis scripts used for analyzing output
+data (both in BoxLib and plaintext format).
+
+The \texttt{tests} directory includes problem setups that are used for testing 
+the hydrodynamics code CASTRO and exercises a wide range of its capabilities.
+
+The \texttt{job\_scripts} directory contains utilities that are used for building and running
+problem setups on various machine architectures.
+
+The \texttt{papers} directory includes the LaTeX source code for the journal articles
+that are being submitted as a result of this project.
+
+\chapter{Test Problems}
+
+The \texttt{tests} directory contains a number of example problems that we use
+to assess the reliability of CASTRO in performing hydrodynamics simulations 
+with self-gravity. This section includes a list of the existing problems 
+as well as the instructions for building and running a problem.
+
+\begin{itemize}
+  \item The \texttt{gravity} subdirectory includes problems that test the Poisson 
+self-gravity module in CASTRO.
+  \begin{itemize}
+    \item The \texttt{boundary\_condition\_comparison} test checks the validity of our 
+multipole approximation for the boundary conditions on the Poisson equation. It runs
+the gravity solve for several different choices of the number of expansion coefficients
+and compares the resulting gravitational potential to the exact result.
+    \item The \texttt{uniform\_cube\_sphere} test is designed to check the convergence
+properties of the Poisson solver. It loads either a sphere or cube of uniform density
+onto the computational domain and then calculates the potential at several different
+spatial resolutions to check if the potential converges.
+  \end{itemize}
+  \item The \texttt{hydro} subdirectory tests includes tests that use the hydrodynamics 
+module, including many with self-gravity.
+  \begin{itemize}
+    \item The \texttt{circular\_orbit} test loads two white dwarfs on a grid and then orbits
+them for a large number of periods, to test conservation of linear momentum,
+angular momentum and energy convergence.
+    \item The \texttt{evrard\_collapse} test runs the test problem described originally by 
+\cite{evrard:1988} and whose results for a grid-based code can be seen in \cite{arepo}. A spherical
+distribution of gas obeying a gamma-law equation of state is loaded onto a grid with negligible 
+pressure. The gas collapses, rebounds with a shock, and then settles into hydrodynamic equilibrium.
+It is a good test of the energy conservation properties of the code.
+    \item The \texttt{freefall} test loads two white dwarfs onto a grid and allows them to fall towards
+each other under their mutual gravitational influence. We check to see whether the timescale for them
+colliding matches the analytical expectation for two point masses.
+    \item The \texttt{kelvin\_helmoltz} test runs several varieties of a 2D shearing fluid that exhibits the 
+Kelvin-Helmholtz instability. The relevant papers are \cite{robertson:2010}, \cite{arepo}, and \cite{mcnally:2012}.
+    \item The \texttt{single\_star\_hse} test loads a single white dwarf on a grid and checks how well 
+hydrostatic equilibrium is maintained.
+  \end{itemize}
+  \item The \texttt{performance} subdirectory examines the scaling and performance properties of CASTRO.
+  \begin{itemize}
+    \item The \texttt{strong\_scaling} test runs the same problem with a varying number of processors and 
+checks the scaling properties.
+    \item The \texttt{weak\_scaling} test scales the size of the problem with the number of processors and
+checks the scaling properties.
+  \end{itemize}
+\end{itemize}
+
+Each test directory is set up with a uniform pattern. The \texttt{source/} subdirectory includes all of the source code
+needed to build the test. The \texttt{makefile} builds the executable and creates
+the inputs files needed for the simulation. The \texttt{run\_test.sh} script executes the test by creating a subdirectory
+for each parameter set and then submitting it with a batch submission script (for a vanilla Linux installation,
+this will use the \texttt{batch} utility; for a limited number of clusters and supercomputers, 
+there will be recommended options and specialized submission scripts). The \texttt{analysis.py} Python script
+is used to analyze the output and generature figures. In summary, to run a test, perform the following series of commands:
+\begin{itemize}
+  \item \texttt{make}
+  \item \texttt{bash run\_test.sh}
+  \item \texttt{python analysis.py} (after the execution has completed)
+\end{itemize}
+
 %------------------------------------------------------------------------------

### 2015-01-30 documentation/userguide.tex
Add list of contributors to the documentation.

@@
 
-\shorttoc{Chapter Listing}{0}
+%\shorttoc{Chapter Listing}{0}
 
-\setcounter{tocdepth}{2}
+%\setcounter{tocdepth}{2}
 \tableofcontents
-
-\clearpage
-
-\addcontentsline{toc}{chapter}{preface}
 
@@
 
+\chapter{Contributors}
+
+This project was started as the Ph.D. dissertation project for Max Katz at Stony Brook University. 
+Michael Zingale is Max's primary thesis advisor. Correspondence and requests for information 
+about this software should be directed to them; their e-mail addresses are maximilian.katz@stonybrook.edu
+and michael.zingale@stonybrook.edu, respectively. Alan Calder and Doug Swesty at Stony Brook are 
+also co-investigators on this project. The CASTRO code was developed primarily at the 
+Lawrence Berkeley National Laboratory. Staff there that have contributed directly and 
+indirectly to this project include Ann Almgren, John Bell, Weiqun Zhang, Andy Nonaka, and Vince Beckner.
+
+We also gratefully acknowledge the assistance of Noel Scudder and Platon Karpov, 
+Stony Brook University undergraduates who helped with running simulations and 
+developing the visualization software based on yt. Adam Jacobs at Stony Brook also provided 
+assistance in various aspects of deploying our codes on supercomputing resources.
+
 %------------------------------------------------------------------------------

### 2015-01-31 papers/wdmerger_I/paper.tex
Flip the sign of the Poisson equation. Also, a brief summary to the gravity-hydro coupling section.

@@
 \begin{equation}
-  \nabla^2 \Phi(\mathbf{x}) = -4\pi G\, \rho(\mathbf{x}),
+  \nabla^2 \Phi(\mathbf{x}) = 4\pi G\, \rho(\mathbf{x}),
 \end{equation}
 where $\Phi$ is the gravitational potential, $G$ is the gravitational
-constant, and $\rho$ is the mass density. We note that the sign
-convention used in CASTRO is opposite to that commonly seen in the
-physics literature, so that $\Phi$ is positive everywhere. The
-solution of this equation in CASTRO is described in \cite{castro}, and
+constant, and $\rho$ is the mass density.\footnote{In the CASTRO code, the 
+right-hand side is negated and therefore $\Phi$ is positive. We use the 
+sign convention that is typical for astrophysics in this paper. 
+When $\Phi$ appears in the code it is negated to compensate for this.} 
+The solution of this equation in CASTRO is described in \cite{castro}, and
 consists of both level and composite solves, and a final
@@
 \begin{equation}
-  \rho E_{\text{tot}} = \rho E - \frac{1}{2}\rho\Phi = \frac{1}{2}\rho \mathbf{u}^2 + \rho e - \frac{1}{2}\rho\Phi, \label{eq:total_energy_gravity}
+  \rho E_{\text{tot}} = \rho E + \frac{1}{2}\rho\Phi = \frac{1}{2}\rho \mathbf{u}^2 + \rho e + \frac{1}{2}\rho\Phi, \label{eq:total_energy_gravity}
 \end{equation}
@@
 \ref{eq:cell_center_gravity_source}), but do not discretize it at
-the cell center.  Instead, we recognize that $\mathbf{u} = {\bm\nabla}\left[\mathbf{r}\cdot\mathbf{u}\right]$,
+the cell center.  Instead, we recognize that in the limit that the velocity is 
+constant, $\mathbf{u} = {\bm\nabla}\left[\mathbf{r}\cdot\mathbf{u}\right]$,
 where $\mathbf{r}$ is the position vector originating from the zone center.
@@
 (We note that this differs slightly from Equation 94 in \cite{arepo} in that
-we evaluate the gravitational acceleration at the zone face rather than the zone center.)
+we evaluate the gravitational acceleration at the zone face rather than the zone center.
+We are free to do this since we assumed a uniform gravitational field to begin with.)
 In the limit of constant fields, this expression simply averages out to the standard 
@@
 
-As pointed out by \citeauthor{arepo}, $(\rho u)$ evaluated at the faces at time level $n+1/2$ is the mass flux,
+As pointed out by \citeauthor{arepo}, $(\rho u)$ evaluated at the faces at time level $n+1/2$ is the mass flux
+that is obtained from the Riemann solver in the hydrodynamics step,
 so $(\rho u)_{i-1/2}\, \Delta t\, \Delta A = \Delta m_{i-1/2}$ is the mass exchanged between zones $i$ and $i-1$.
@@
 
+Our strategy for coupling the gravity to the hydrodynamics on a uniform grid can thus be summarized as follows. 
+At the beginning of a timestep, we solve the Poisson equation given the density $\rho^n$. We then solve the 
+Euler equations and advance the system in time by $\Delta t$ using zone-centered fluxes at time-level $n+1/2$. 
+We add $\Delta t \rho^n \mathbf{g}^n$ to the momentum equations. We then solve the Poisson equation again, given
+the density $\rho^{n+1}$. We then add $-(\Delta t/2) \rho^{n} \mathbf{g}^{n} + (\Delta t / 2) \rho^{n+1} \mathbf{g}^{n+1}$
+to the momentum equations, time-centering the source term. We simultaneously add $\sum_{j} -(1/2) \Delta m_{j} (\Phi - \Phi_j)$
+to the zone's energy, where the index $j$ runs over all adjacent zones. This conserves the total energy on the grid
+to roundoff error, modulo any losses from an open domain boundary.
+
 \subsubsection{Boundary Conditions}\label{sec:gravity_boundary_conditions}
@@
 \begin{equation}
-  \Phi_{{lmn}} = \sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{direct_sum}
+  \Phi_{{lmn}} = -\sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{direct_sum}
 \end{equation}
@@
 \begin{equation}
-  \Phi(\mathbf{x}) = \sum_{l=0}^{\infty}\sum_{m=-l}^{l} \frac{4\pi}{2l + 1} q_{lm} \frac{Y_{lm}(\theta,\phi)}{r^{l+1}}, \label{spherical_harmonic_expansion}
+  \Phi(\mathbf{x}) = -\sum_{l=0}^{\infty}\sum_{m=-l}^{l} \frac{4\pi}{2l + 1} q_{lm} \frac{Y_{lm}(\theta,\phi)}{r^{l+1}}, \label{spherical_harmonic_expansion}
 \end{equation}
@@
 \begin{align}
-  \Phi(\mathbf{x}) &= \sum_{l=0}^{\infty} \left\{Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} \right. \notag \\
+  \Phi(\mathbf{x}) &= -\sum_{l=0}^{\infty} \left\{Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} \right. \notag \\
     &+ \left. \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right\}.\label{multipole_potential}
@@
 \begin{equation}
-  \Phi_{\text{sphere}}(r) = \frac{GM}{r} \times \begin{cases} (3R^2 - r^2)/(2 r^2) & r \leq R \\ 1 & r > R \end{cases},\label{eq:sphere-analytical}
+  \Phi_{\text{sphere}}(r) = -\frac{GM}{r} \times \begin{cases} (3R^2 - r^2)/(2 r^2) & r \leq R \\ 1 & r > R \end{cases},\label{eq:sphere-analytical}
 \end{equation}
@@
 \begin{align}
-  \Phi_{\text{cube}}(x,y,z) &= G\rho\sum_{i,j,k=0}^1\left[x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right)\right. \notag \\
+  \Phi_{\text{cube}}(x,y,z) &= -G\rho\sum_{i,j,k=0}^1\left[x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right)\right. \notag \\
   &+ \left. y_j z_k\, \text{tanh}^{-1}\left(\frac{x_i}{r_{ijk}}\right) + z_k x_i\, \text{tanh}^{-1}\left(\frac{y_j}{r_{ijk}}\right) \right.\notag \\

### 2015-02-10 papers/wdmerger_I/paper.tex
Rewrite of the free-fall section so that it is in dimensionless units.

@@
 the specified \textit{primary} mass and $M_S$ is the specified
-\textit{secondary} mass. The primary WD will always be on the left
+\textit{secondary} mass. The primary WD will always start on the left
 side of the computational domain for our simulations, and is more
@@
 
-A simple test to verify the Poisson solver implemented by CASTRO is
-the case of gravitational free fall. In this setup, two stars, each
-individually in an equilibrium state, are placed on the computational
-grid, separated by an initial distance $r_0$ along the $x$ axis with
-zero initial velocity. We choose stars of masses $0.6$ and $0.8\,
-M_\odot$, with the lower mass star on the left side (e.g. $x < 0$) of
-the domain and the higher mass star on the right, such that their
-center of mass coincides with the center of the
-domain. Gravitationally, the stars may be treated as point masses
-until the point of contact, so the equation of motion governing the
-distance $r$ between their centers of mass is that of simple free
-fall:
-\begin{equation}
-  \ddot{r}(t) = - \frac{GM}{r},
-\end{equation}
-where $G$ is the gravitational constant and $M$ is the total mass of
-the system. This differential equation has a closed-form solution for
-the evolution time as a function of separation:
-\begin{equation}
-  t(r) = \sqrt{\frac{r_0^3}{2GM}} \left[ \text{arccos}\left(\sqrt{\frac{r}{r_0}}\,\right) + \sqrt{\frac{r}{r_0} \left(1 - \frac{r}{r_0}\right)}\ \right]. \label{analyticalFreeFall}
-\end{equation}
-This result can be derived by recognizing that
-\begin{equation}
-  t(r) = \int_{r_0}^r \frac{dr}{v(r)}
-\end{equation}
-and inserting the velocity as a function of radial separation,
-\begin{equation}
-  v(r) = \sqrt{\frac{2GM}{r_0}\left(\frac{r_0}{r} - 1\right)}.
-\end{equation}
-We determine the initial separation to be consistent with the
-simulation performed in Section \ref{sec:kepler}; that is, we select
-an initial orbital period and use Kepler's third law to calculate the
-radius of a circular orbit with that period. We select an initial
-orbital period of $T = 100$ s for this simulation, so that the separation is 
-$r_0 = 3.61 \times 10^{9}\ \text{cm}$.
-We chose a relatively low resolution simulation to demonstrate the
-capabilities of CASTRO even while using modest resources. The
-computational grid is covered by a coarse grid of $48^3$ zones, with
-two levels of refinement above the coarse grid. Each refined grid
-carries an increase in resolution by a factor of 4 relative to the
-coarser grid below it. CASTRO initially assigns $93\%$ of the domain
-to be covered by the intermediate resolution grids, and $0.05\%$ of
-the domain to be covered by the finest resolution grids.
+A simple dynamical test to verify the gravitation physics in CASTRO is
+the case of gravitational free fall. We place two stars on the grid 
+in the manner of Section \ref{sec:initial_models}. The distance $a$ between 
+them corresponds to a chosen orbital period $T$, consistent with the total
+system mass $M$, but we disable the rotational source terms so that 
+the stars start at rest in an inertial reference frame. 
+Thus the stars will simply begin moving toward each other.
+As long as the stars remain approximately spherical, the stars can be 
+treated as point masses (this only seriously breaks down after the stars
+have come into contact). In dimensionless units where $r \to r / a$ and 
+$t \to 2\sqrt{2}\pi t / T$, the simple free fall equation of motion governing the
+distance $r$ between their centers of mass takes the form:
+\begin{equation}
+  \ddot{r}(t) = - \frac{1}{2r^2}.
+\end{equation}
+It is possible to derive a closed-form solution for the evolution time
+as a function of separation by starting with the integral formulation,
+\begin{equation}
+  t(r) = \int_{1}^{r} \frac{dr}{v(r)}.
+\end{equation}
+The velocity $v$ (in dimensionless units) can be found by noting that 
+$\ddot{r} = v\, dv / dr$ and then separating and integrating the equation 
+of motion. This yields 
+\begin{equation}
+  v(r) = \sqrt{\left(\frac{1}{r} - 1\right)}.
+\end{equation}
+For our problem $0 \leq r \leq 1$, so this is always valid. Integrating, we find
+\begin{equation}
+  t(r) = \text{arccos}\left(\sqrt{r}\right) + \sqrt{r \left(1 - r\right)}. \label{analyticalFreeFall}
+\end{equation}
+so that the point of contact would occur at $t = 1$. We actually stop the simulation
+at $t = 0.9$, which is when the effects from the extended sizes of the stars
+starts to become important. The results of our simulation for a $128^3$ zone 
+uniform grid are shown in Figure \ref{Fig:Free Fall}. They show excellent agreement
+between the theory and the simulation results, even for this very modest resolution.
 
@@
     to the initial separation, and the vertical axis gives the elapsed
-    time of the simulation, scaled to the total elapsed time
-    necessarily. The solid curve shows the analytical result,
+    time of the simulation, scaled to the time it would take two point masses
+    to collide. The solid curve shows the analytical result,
     calculated from Newtonian mechanics, and the circles show the
-    samples from the time evolution with CASTRO. The agreement between
-    theory and simulation is excellent, even for a simulation with
-    modest resolution.}
+    samples from the time evolution with CASTRO.}
   \label{Fig:Free Fall}
 \end{figure}
-
-The analytical result in Equation \ref{analyticalFreeFall} determines
-the total elapsed free-fall time,
-\begin{equation}
-  t_{\text{ff}} = \frac{\pi}{2} \sqrt{\frac{r_0^3}{2GM}} = \frac{T}{4\sqrt{2}}.
-\end{equation}
-The physical radius of each white dwarf is roughly $10\%$ of the
-initial separation, so we consider the evolution terminated when the
-radial separation reaches that value (at that point, the stars will no
-longer be in free-fall due to physical contact). Since the elapsed
-time goes roughly as the square root of the distance for small $r$, we
-consider the evolution terminated when $t = 0.99\, t_{\text{ff}}$. The
-results of our simulation are shown in Figure \ref{Fig:Free Fall}. The
-positions of the two stars are determined by calculating the center of
-mass of the right $(x > 0)$ and left $(x < 0)$ sides of the domain at 
-the end of each time step, and treating the centers of mass as the 
-respective location of the two stars.
 

### 2015-02-13 job_scripts/run_utils.sh
Create a compile/ directory in each test. It will include the makefile, and the executable and inputs files will be built there.

@@ -7,20 +7,13 @@
 
 function get_wdmerger_make_var {
 
-    make print-$1 | tail -1 | awk '{ print $NF }'
+    make print-$1 -C $compile_dir | tail -2 | head -1 | awk '{ print $NF }'
 
 }
 
 function get_castro_make_var {
 
-    make print-$1 -f $local_makefile  | tail -1 | awk '{ print $NF }'
-
-}
-
-
-function get_machine {
-    
-    MACHINE=$(get_make_var WHICHLINUX)
+    make print-$1 -f $local_makefile -C $compile_dir  | tail -2 | head -1 | awk '{ print $NF }'
 
 }
 
@@ -43,13 +36,13 @@ function get_last_checkpoint {
 
 function copy_files {
 
-    cp $CASTRO $1
+    cp $compile_dir/$CASTRO $1
     if [ -e "helm_table.dat" ]; then
-	cp helm_table.dat $1
+	cp $compile_dir/helm_table.dat $1
     fi
-    cp $inputs $1
-    cp $probin $1
-    cp $job_script $1
+    cp $compile_dir/$inputs $1
+    cp $compile_dir/$probin $1
+    cp $compile_dir/$job_script $1
 
 }
 
@@ -97,11 +90,11 @@ function run {
     fi
 
     if [ $MACHINE == "GENERICLINUX" ] ; then 
-	echo "echo \"mpiexec -n $nprocs $CASTRO $inputs > info.out\" | batch" > $job_script
+	echo "echo \"mpiexec -n $nprocs $CASTRO $inputs > info.out\" | batch" > $compile_dir/$job_script
     elif [ $MACHINE == "BLUE_WATERS" ]; then
-	sed -i "/#PBS -l nodes/c #PBS -l nodes=$ntasks:ppn=$ppn:xe" $job_script
-	sed -i "/#PBS -l walltime/c #PBS -l walltime=$walltime" $job_script
-	sed -i "/aprun/c aprun -n $nprocs -N $ppn $CASTRO $inputs \$\{restartString\}" $job_script
+	sed -i "/#PBS -l nodes/c #PBS -l nodes=$ntasks:ppn=$ppn:xe" $compile_dir/$job_script
+	sed -i "/#PBS -l walltime/c #PBS -l walltime=$walltime" $compile_dir/$job_script
+	sed -i "/aprun/c aprun -n $nprocs -N $ppn $CASTRO $inputs \$\{restartString\}" $compile_dir/$job_script
     fi
 
[…]

### 2015-02-13 tests/gravity/boundary_condition_comparison/run_test.sh

@@ -5,13 +5,13 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 for l in {0..20}
 do
   dir=$results_dir/$l
-  sed -i "/gravity.max_multipole_order/c gravity.max_multipole_order = $l" $inputs
+  sed -i "/gravity.max_multipole_order/c gravity.max_multipole_order = $l" $compile_dir/$inputs
   run $dir
 done
 
 # Now do the 'exact' direct summation, for comparison purposes
 
 dir=$results_dir/true
-sed -i "/gravity.direct_sum_bcs/c gravity.direct_sum_bcs = 1" $inputs
+sed -i "/gravity.direct_sum_bcs/c gravity.direct_sum_bcs = 1" $compile_dir/$inputs
 run $dir
 

### 2015-02-13 tests/gravity/uniform_cube_sphere/run_test.sh

@@ -9,8 +9,8 @@ do
   do
 
     dir=$results_dir/problem$problem/$ncell
-    sed -i "/problem/c problem = $problem" $probin
-    sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $inputs
+    sed -i "/problem/c problem = $problem" $compile_dir/$probin
+    sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $compile_dir/$inputs
     run $dir
 
   done

### 2015-02-13 tests/hydro/circular_orbit/run_test.sh

@@ -5,6 +5,6 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 for ncell in 128
 do
   dir=$results_dir/$ncell
-  sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $inputs
+  sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $compile_dir/$inputs
   run $dir
 done

### 2015-02-13 tests/hydro/evrard_collapse/run_test.sh

@@ -8,8 +8,8 @@ do
   do
     dir=$results_dir/gs$gst/ncell$ncell
     nprocs=16
-    sed -i "/castro.grav_source_type/c castro.grav_source_type = $gst" $inputs
-    sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $inputs
+    sed -i "/castro.grav_source_type/c castro.grav_source_type = $gst" $compile_dir/$inputs
+    sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $compile_dir/$inputs
     run $dir $nprocs
   done
 done

### 2015-02-13 tests/hydro/freefall/run_test.sh

@@ -5,6 +5,6 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 for ncell in 32 64
 do
   dir=$results_dir/$ncell
-  sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $inputs
+  sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $compile_dir/$inputs
   run $dir
 done

### 2015-02-13 tests/hydro/kelvin_helmholtz/run_test.sh

@@ -12,7 +12,7 @@ do
 
     # Loop over the resolutions in question
 
-    for ncell in 64 128 256 1024 2048 4096 #512 1024 #2048 4096
+    for ncell in 64 128 256 1024 2048 4096
     do
 
       # We only want to do the high-resolution convergence test
@@ -24,10 +24,10 @@ do
 
       dir=$results_dir/problem$problem/velocity$vel/$ncell
 
-      sed -i "/problem/c problem = $problem" $probin
-      sed -i "/bulk_velocity/c bulk_velocity = $vel" $probin
+      sed -i "/problem/c problem = $problem" $compile_dir/$probin
+      sed -i "/bulk_velocity/c bulk_velocity = $vel" $compile_dir/$probin
 
-      sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell" $inputs
+      sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell" $compile_dir/$inputs
 
       # Determine stopping time based on problem of interest
 
@@ -42,9 +42,9 @@ do
 	  plot_per=0.1
       fi
 
-      sed -i "/amr.plot_per/c amr.plot_per = $plot_per" $inputs
-      sed -i "/amr.check_per/c amr.check_per = $plot_per" $inputs
-      sed -i "/stop_time/c stop_time = $stop_time" $inputs
+      sed -i "/amr.plot_per/c amr.plot_per = $plot_per" $compile_dir/$inputs
+      sed -i "/amr.check_per/c amr.check_per = $plot_per" $compile_dir/$inputs
+      sed -i "/stop_time/c stop_time = $stop_time" $compile_dir/$inputs
 
       # Set number of processors based on amount of work
 

### 2015-02-13 tests/hydro/single_star_hse/run_test.sh

@@ -5,6 +5,6 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 for ncell in 32 64 128
 do
   dir=$results_dir/$ncell
-  sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $inputs
+  sed -i "/amr.n_cell/c amr.n_cell = $ncell $ncell $ncell" $compile_dir/$inputs
   run $dir
 done

### 2015-02-14 papers/wdmerger_I/paper.tex
Mention the diagnostics that we track in the paper.

@@
 \end{align}
-As with gravity, in three dimensions we also include the mass fluxes through all six cell faces.
+As with gravity, in three dimensions we include the mass fluxes through all six cell faces.
+
+In all simulations performed in a rotating reference frame, we will transform all relevant
+quantities back to the inertial reference frame when reporting them in this paper. In particular,
+for every zone we add to the displayed kinetic energy the rotational energy associated with motion 
+at the circular velocity corresponding to the distance of that zone from the rotation axis.
+We do the analogous calculation for angular momentum as well. When plotting the locations of 
+stars on the grid, we rotate them by an angle corresponding to the amount of time that has 
+passed since the simulation started.
 
@@
 the maximum timestep is set to be equal to one-half of the CFL limit.
+
+We track a number of diagnostic quantities at the end of every coarse grid timestep. 
+For all simulations, we record the total energy (including the breakdown into
+its components: kinetic, internal, gravitational potential, rotation), 
+the total angular momentum, and the center of mass of the system. 
+For stellar calculations, we separately record diagnostic 
+information about the stars, including their centers of mass, 
+and the volume contained within each star.
 

### 2015-02-14 papers/wdmerger_I/paper.tex
Add reference to the location of this software.

@@
 For all simulations, we record the total energy (including the breakdown into
-its components: kinetic, internal, gravitational potential, rotation), 
+its components: kinetic, internal, gravitational potential, rotation; we note
+that for the diagnostics we actually use $(\rho E)$ for calculation of the total energy,
+rather than explicitly calculating the sum of kinetic and internal, as this is
+the quantity that should be explicitly conserved), 
 the total angular momentum, and the center of mass of the system. 
@@
 and the volume contained within each star.
+
+The software used to generate the test problems in this paper
+(as well as the paper text itself),
+\texttt{wdmerger}\footnote{\url{https://github.com/BoxLib-Codes/wdmerger}},
+is freely available at an online repository hosting service.
+Version control in both the parent software (BoxLib, CASTRO) and in \texttt{wdmerger}
+permits us to reference the state of the code at the time a simulation
+was performed. To do so, in all plot files generated we store
+the complete set of parameters used in that simulation. We also store the 
+version commit hashes of the parent software in the plot files.
+We store the commit hashes of all three as comments in the metadata of 
+any EPS files generated as figures in this paper.
 

### 2015-02-14 papers/wdmerger_I/paper.tex
Revamp discussion of density resets to be consistent with recent CASTRO changes. Also, remove reference to the dt limiting, which we are no longer doing.

@@
 
-A related issue that required a code improvement was that in cases of
+A related issue that required a code improvement is that in cases of
 large density gradients such as the edge of a star, it is possible to
@@
 seriously affect the energy conservation properties of our
-simulation. Similarly, it is also possible in this physical situation
-for very large velocities to be generated in the material behind the
-star. The momentum transferred to the ambient zone is close to the
-momentum of the stellar material, but this is of much higher density
-and so the velocity of the ambient material must become very large to
-compensate. While this may be physically viable in certain
-circumstances, it poses a problem for our code, which uses subcycling
-in time (see below). The timestep is fixed at the beginning of the set
-of subcycled timesteps, and if in the middle of the cycle a velocity
-is generated that is so large that it violates the CFL criterion for
-this timestep, we cannot guarantee the stability of the resulting
-solution since we cannot alter the timestep. (Nor would we want to,
-because in some cases because the velocity becomes spuriously large
-and would drag down the global timestep.) Therefore we also insert a
-check at the end of every hydrodynamics timestep that resets the
-velocity and/or sound-speed (indirectly, using the temperature) of any
-zone that violates the CFL criterion given the timestep we chose at
-the beginning of the cycle. In practice, we find that this is mainly
-necessary only for simulations in the inertial reference frame of the
-white dwarf merger, which we do not normally do for the reasons stated
-in Section \ref{sec:kepler}.
+simulation. Our strategy for a reset is as follows: when the density of 
+a zone is below a pre-determined density floor (which is typically 
+$10^{-5}\ \text{g cm}^{-3}$ for our stellar simulations), we look
+at all adjacent zones and find the zone with the highest density.
+If it is above the density floor, then we set the field values 
+(density, momentum, energy, and temperature) of the
+resetted zone to be exactly equal to the field values of this 
+adjacent zone. If no adjacent zone reaches the density floor, then
+the zone is set to the density floor, and given a temperature equal 
+to the temperature floor for our simulations (which is typically 
+$10^{5}\ \text{K}$ for our stellar simulations). We then recompute 
+the thermodynamics to be consistent with these values. The 
+velocity of the zone is set to zero. This latter approach will only
+occur in very rare situations, and is mainly there as a safeguard.
 

### 2015-02-14 papers/wdmerger_I/paper.tex
Updated discussion of eos_input_is_constant in the paper

@@
 the iterative value approaches zero. In these cases we employ a
-``floor'' that limits how low the temperature can go (typically $10^4$
-or $10^5$ K). There is a choice here how to proceed: we can either
+``floor'' that limits how low the temperature can go (typically 
+$10^5$ K). There is a choice here how to proceed: we can either
 assign this floor value to the temperature and let that zone be
-thermodynamically inconsistent (the original behavior in Castro), or
+thermodynamically inconsistent (the original behavior in CASTRO), or
 we can adjust the internal energy to be thermodynamically consistent
 with the temperature, at the cost of violating energy conservation. We
-have found in test problems of one-dimensional shocks that the latter
-yields more accurate results, so we employ the latter method.
-
-Reactions?
+have found in some test problems of strong one-dimensional shocks that reach 
+the temperature floor that the latter yields more accurate results. 
+However, allowing the equation of state call to update the 
+internal energy can actually result in significant changes to the 
+total energy of the system over long periods of time, 
+due not just to resets in low-density zones but also to small 
+inconsistencies between the energy given to the EOS and the energy 
+that is consistent with the returned temperature. These inconsistencies
+are dependent on the threshold we set that determines how close the 
+returned energy should be to the original input energy. While this 
+threshold is typically very small in an individual zone (a relative 
+difference of $10^{-8}$ by default in CASTRO), over time and given 
+a large number of zones, this can result in a significant energy 
+drift. This is a serious enough problem that we opt for the energy 
+conserving approach for our simulations.
+
+CASTRO has the ability to do both nuclear reactions and radiative 
+transport (in the flux-limited diffusion approximation). For all simulations 
+in this paper we do not enable either, and will delay discussion of 
+these modules until a later paper in this series.
 
@@
 where $\rho$, $\mathbf{u}$, and $\mathbf{g}$ 
-are evaluated at the zone
-center.
+are evaluated at the zone center.
 
@@
 
-\section{Castro hydrodynamics changes}
+\section{CASTRO hydrodynamics changes}
 

### 2015-02-17 papers/wdmerger_I/paper.tex
Add acknowledgement of Noel and Platon.
Add acknowledgement to Adam Jacobs
Cite support from Blue Waters

@@
 
-This research was supported by NSF award AST-1211563.  This research
-used resources of the National Energy Research Scientific Computing
-Center, which is supported by the Office of Science of the
-U.S. Department of Energy under Contract No. DE-AC02-05CH11231.  An
+The authors thank Noel Scudder and Platon Karpov for their help with 
+this project, especially related to visualization of the results. 
+We also thank Adam Jacobs for his advice on and assistance with 
+running on supercomputing resources.
+
+This research was supported by NSF award AST-1211563. An
 award of computer time was provided by the Innovative and Novel
@@
 the Office of Science of the Department of Energy under Contract
-DE-AC05-00OR22725.
-Project AST006 supported use of the ORNL/Titan resource.
-This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number ACI-1053575.
+DE-AC05-00OR22725. Projects AST006 and AST106 supported use of the ORNL/Titan resource. 
+This research is part of the Blue Waters sustained-petascale computing project, 
+which is supported by the National Science Foundation (awards OCI-0725070 
+and ACI-1238993) and the state of Illinois. Blue Waters is a joint 
+effort of the University of Illinois at Urbana-Champaign and its 
+National Center for Supercomputing Applications.
+This research used resources of the National Energy Research Scientific Computing
+Center, which is supported by the Office of Science of the
+U.S. Department of Energy under Contract No. DE-AC02-05CH11231.  
+This work used the Extreme Science and Engineering Discovery Environment (XSEDE), 
+which is supported by National Science Foundation grant number ACI-1053575. 
 Project AST100037 supported use of the resources NICS/Kraken and NICS/Darter.
-
-%TODO: add blue waters citation
 

### 2015-02-18 papers/wdmerger_I/paper.tex
Add plot for the boundary condition comparison to the paper.

@@
 \begin{equation}
-  \Phi_{{lmn}} = -\sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{direct_sum}
+  \Phi_{{lmn}} = -\sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{eq:direct_sum}
 \end{equation}
@@
 
-% Insert multipole vs. exact calculation here
+As a demonstration of the method's accuracy, we consider the case of two 
+white dwarfs of mass ratio 2/3, loaded onto the grid by the method described 
+in Section \ref{sec:initial_models}. We terminated the simulation just after
+initialization, so that we perform only an initial Poisson solve for this 
+density distribution. We did this for values of $l_{\text{max}}$ ranging
+from 0 to 20, and we also did this using the numerically exact solution 
+provided by Equation \ref{eq:direct_sum}.  Defining the $L^2$
+norm of a field $f$ as
+\begin{equation}
+  \| f \|_2 = \left(\sum_{i,j,k} dx\, dy\, dz\, f_{ijk}^2\right)^{1/2},
+\end{equation}
+we computed the $L^2$ error of $\Phi$ on the entire domain for multipole 
+boundary conditions, which we will call $\Phi_l$, relative to $\Phi$ 
+obtained using the exact boundary conditions:
+\begin{equation}
+  \text{Error}_l = \frac{\|\Phi_l - \Phi_{\text{exact}}\|_2}{\|\Phi_{\text{exact}}\|_2}.
+\end{equation}
+The result is shown in Figure \ref{fig:bc_comparison}. At $l_{\text{max}} = 6$,
+the error is already well below $10^{-4}$ and we adopt this as our default 
+choice for all simulations with Poisson gravity. Our choice is also motivated by the 
+fact that at this value, the time for computing the boundary conditions on the
+coarse grid is comparable to the time for the multigrid solve itself; at higher values 
+the boundary condition calculation becomes much more expensive than the multigrid solve
+and so the overall returns are severely diminished. We note that in our implementation 
+there are no gains at all for going beyond $l_{\text{max}} \gtrsim 16$, as we 
+become limited by error due to the machine precision.
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.45]{plots/bc_comparison}
+  \caption{Error of $\Phi$ on the computational domain for a binary white dwarf simulation 
+    whose boundary conditions were computed using various values of the maximum multipole order,
+    relative to the exact solution determined by a brute force sum on the boundaries.
+    Circles represent the error at integer values, and they have been connected by a smooth 
+    line to guide the eye.\label{fig:bc_comparison}}
+\end{figure}
 
@@
 \end{equation}
-where $M = 4\pi R^3 / 3$ is the mass of the sphere. Defining the $L^2$
-norm of a field $f$ as
-\begin{equation}
-  \| f \|_2 = \left(\sum_{i,j,k} dx\, dy\, dz\, f_{ijk}^2\right)^{1/2},
-\end{equation}
-we measure the numerical error by calculating the $L^2$ norm of the error and normalizing it by the $L^2$ norm of the analytical solution:
+where $M = 4\pi R^3 / 3$ is the mass of the sphere. We measure the 
+numerical error by calculating the $L^2$ norm of the error and 
+normalizing it by the $L^2$ norm of the analytical solution:
 \begin{equation}
@@
   \centering
-  \includegraphics[scale=1.0]{reference}
+  \includegraphics[scale=1.0]{plots/reference}
   \caption{ \label{Fig:sod} Solution to Sod's problem with the original
@@
 %  \centering
-%  \includegraphics[width=3.1in]{nested2}\hspace{1em}
-%  \includegraphics[width=3.1in]{nested3}
+%  \includegraphics[width=3.1in]{plots/nested2}\hspace{1em}
+%  \includegraphics[width=3.1in]{plots/nested3}
 %  \caption{\label{fig:2levgrav} Schematic showing the solve for $\phi$ for a 2-level grid.}

### 2015-02-19 documentation/userguide.tex
Create a compile/ directory in each test. It will include the makefile, and the executable and inputs files will be built there.
Add documentation on how the software interfaces with various supercomputing resources.

@@
 Each test directory is set up with a uniform pattern. The \texttt{source/} subdirectory includes all of the source code
-needed to build the test. The \texttt{makefile} builds the executable and creates
+needed to build the test. The \texttt{compile/} directory contains the \texttt{makefile}, which builds the executable and creates
 the inputs files needed for the simulation. The \texttt{run\_test.sh} script executes the test by creating a subdirectory
@@
 \begin{itemize}
+  \item \texttt{cd compile}
   \item \texttt{make}
+  \item \texttt{cd ..}
   \item \texttt{bash run\_test.sh}
@@
 
+\chapter{System Configuration}
+
+Running the wdmerger software will require a little setup in terms of configuring the 
+scripts to work with the system you are currently using. The software is currently actively 
+maintained only for a small number of supercomputing resources (Blue Waters at NCSA 
+and Titan at OLCF, in particular). If you are using one of these machines, then 
+modifying the code for your purposes is straightforward. If you are not, you'll need to do a 
+small bit of ground work to get everything set up, but once you do the automated scripts 
+will handle everything else for you. Feel free to contact us for advice on how to do so 
+(see Section \ref{sec:contributors} for contact information), but be aware that it will be 
+hard for us to debug on systems that we do not have access to.
+
+It is important to note that although we have attempted to 
+make sure that this software builds and runs on a generic Linux workstation, most of the test 
+problems are simply too large to run on anything smaller than a medium-sized cluster. To get a flavor 
+of what is going on, though, you can comment out the larger runs from the various loops in the 
+\texttt{run\_test.sh} file and then run the smaller problem setup.
+
+\section{Account Details and System Options}
+
+All of the machine-dependent functionality exists in the file \texttt{job\_scripts/run\_utils.sh}.
+In particular, near the end of the file you can find a bash conditional that tests on the name of 
+certain machines. Each machine definition includes a number of variables that are needed to 
+properly configure all of the helper functions. If you are using a system already known to us,
+then the only thing you need to change is the \text{allocation} variable: this is the account 
+that jobs will be charged to on the machine.
+
+If you are establishing a new machine, then there are a few steps you will need to take. First, 
+add a section to this conditional that tests on what you want to call the name of the machine. 
+Then, make entries for the necessary variables:
+\begin{itemize}
+  \item \texttt{allocation}: the account jobs are charged to.
+  \item \texttt{exec}: the function used to submit batch scripts. For PBS-style submission 
+systems such as TORQUE, this will be \texttt{qsub}.
+  \item \texttt{ppn}: the number of processors per node on the system. Be careful with how 
+you define this. Many systems have nodes that have two integer cores for every floating point
+unit and will report the number of integer cores as the number of processors per node, 
+and for floating point-intensive applications like hydrodynamics adding as many threads/MPI tasks
+as there are integer cores may actually slow you down. You will have to test to find 
+the best setting for you.
+  \item \texttt{run\_ext}: the file extension of the standard output file when a job is in operation. 
+This is used by the run scripts to ensure that we don't submit a job in a directory where one is 
+already running. For many systems this will be ``.OU''.
+  \item \texttt{batch\_system}: the job submission system used on the machine. For many supercomputers 
+``PBS'' will be what you should insert here, which is a catchall for any derivative such as TORQUE 
+that uses normal PBS options for accepting jobs.
+  \item \texttt{archive\_method} (optional): the method used by the machine for saving files to 
+long-term storage. See Section \ref{sec:archiving}. This is optional, and if you don't set it then
+the scripts won't attempt to copy your data off the machine.
+\end{itemize}
+When you submit a job through a \texttt{run\_test.sh} file using the \texttt{run} command, 
+a batch script will be automatically created for you. For PBS systems you should be mostly 
+good to go as-is, but some systems vary particularly in the \texttt{-l nodes} option, so 
+you should make sure that the job script construction is in sync with how batch jobs 
+on that machine should appear. Locate the \texttt{create\_job\_script} function and adjust it 
+accordingly by changing the relevant \texttt{echo} statements. If you are using a different submission
+system, create a new section of the conditional in that function, and model it off what we do for PBS
+so that the relevant information is output to the newly-created job script (called \texttt{run\_script} 
+by default; you can change this in the definitions section as well).
+
+The last step is to tell the script how to know that you're using the machine you just defined. 
+Locate the \texttt{get\_machine} command and use the output of \texttt{uname -n} to uniquely determine 
+which system you are using.
+
+\section{Archiving}
+\label{sec:archiving}
+
+This software has some capability to automatically handle archiving of data to long-term storage,
+at least for systems we already know. Each time we run a job and we are continuing from an old checkpoint,
+we move all new plotfiles and checkpoints to the \texttt{output/} subdirectory of the job submission 
+location. We also make copies of the diagnostics files, and the standard output files. By default this is 
+all we do. But on systems where archiving is supported, we copy these files to the long-term storage.
+
+There are currently two methods we know about: \texttt{htar} for Titan, and Globus Online for Blue Waters.
+\texttt{htar} is a command that both creates a tar file and then copies it to the high-performance storage system 
+(HPSS) using the \texttt{hsi} command. It works on a file-by-file basis, so we loop over all of the files to 
+archive and run \texttt{htar} on them individually. If your system uses \texttt{htar}, then you should be all set.
+
+\href{https://www.globus.org/}{Globus Online} is an internet service that handles data transfer requests 
+between certain known endpoints. This is the choice used by Blue Waters for copying and retrieving data
+to their storage system. You will need to set up your account and authenticate through Globus before using it.
+Once you know your Globus username, you can adjust the \texttt{globus\_username} in \texttt{run\_utils.sh}. 
+We note that the Globus archive strategy is different. There is a cap on the number of simultaneous transfers 
+one can have, so submitting transfers for every plotfile and checkpoint will not work. Instead, when we 
+run the archiving script, we first move all the files to the output directory, and then we do a transfer 
+on the entire \texttt{output/} subdirectory. Globus will automatically sync the source and destination 
+directories, so no extra work is required of us. We choose the \texttt{sync\_level = 2} option, which 
+overwrites existing files if the source has a newer date than the destination. This allows us to 
+continually update the diagnostic files, which always have the same name.
+
+If your system uses a different archive strategy, you will need to update the \texttt{archive\_all} 
+and \texttt{archive} functions.
+
 \chapter{Contributors}
+\label{sec:contributors}
 

### 2015-02-19 papers/wdmerger_I/paper.tex
No longer using the graphics showing multi-level Poisson boundary conditions -- should be able to explain it easily enough in the text.
Slightly update description of how we store git commit hashes, consistent with the recent CASTRO changes.

@@
 The software used to generate the test problems in this paper
-(as well as the paper text itself),
+(as well as the manuscript itself),
 \texttt{wdmerger}\footnote{\url{https://github.com/BoxLib-Codes/wdmerger}},
@@
 permits us to reference the state of the code at the time a simulation
-was performed. To do so, in all plot files generated we store
-the complete set of parameters used in that simulation. We also store the 
-version commit hashes of the parent software in the plot files.
-We store the commit hashes of all three as comments in the metadata of 
-any EPS files generated as figures in this paper.
+was performed. In all plot files and diagnostic output generated by CASTRO, 
+and EPS figure files and plots generated by \texttt{wdmerger} for this work,
+we store the active \texttt{git} commit hashes of BoxLib, CASTRO and \texttt{wdmerger}.
 
@@
 
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=3.1in]{plots/nested2}\hspace{1em}
-%  \includegraphics[width=3.1in]{plots/nested3}
-%  \caption{\label{fig:2levgrav} Schematic showing the solve for $\phi$ for a 2-level grid.}
-%\end{figure}
-
-\clearpage
-
 \end{document}

### 2015-02-19 papers/wdmerger_I/paper.tex
Mention in the paper the ambiguity presented by whether we apply the rotation force before or after the gravitation force.

@@
 As with gravity, in three dimensions we include the mass fluxes through all six cell faces.
+
+We apply the rotation forces after the gravitation forces, but 
+there is some freedom in the order in which to apply the gravitational and rotational terms.
+This order may matter because the Coriolis force depends on the fluid velocity, and 
+in the predictor-corrector approach, we use the velocities both at 
+time-level $n$ and time-level $n+1$. If we update the latter with the gravitational force, 
+then the Coriolis force will see a different velocity than the one obtained through the 
+pure hydrodynamics step. (This does not matter for the energy equation in our new formulation,
+because the velocities used are always the time-level $n+1/2$ values coming from the Riemann solver.)
+In practice, this will not matter significantly for our simulations in this work 
+because the centrifugal force plays the dominant role in maintaining stability of non-contact 
+binary systems, and the centrifugal force does not depend on the fluid velocity.
+However, this issue may be worth exploring in future work in situations where the Coriolis 
+term is non-negligible in determining the system evolution.
 

### 2015-02-19 papers/wdmerger_I/paper.tex
In the paper directory, keep a notes directory with LaTeX source of any notes that are supplementary to the main paper.
Update the free fall figure.

@@
 at $t = 0.9$, which is when the effects from the extended sizes of the stars
-starts to become important. The results of our simulation for a $128^3$ zone 
+starts to become important. The results of our simulation for our default $256^3$ zone 
 uniform grid are shown in Figure \ref{Fig:Free Fall}. They show excellent agreement
-between the theory and the simulation results, even for this very modest resolution.
+between the theory and the simulation results, and this agreement holds even for 
+a more moderate $128^3$ zone simulation.
 
@@
   \centering
-  \includegraphics[scale=1.0]{freefall/plot_freefall}
+  \includegraphics[scale=0.45]{plots/freefall}
   \caption{Time evolution of two initially stationary white dwarfs,
@@
     calculated from Newtonian mechanics, and the circles show the
-    samples from the time evolution with CASTRO.}
+    samples from the time evolution with CASTRO. For visual clarity, we 
+    show only a small fraction of the timesteps.}
   \label{Fig:Free Fall}

### 2015-02-20 papers/wdmerger_I/paper.tex
Add uniform cube/sphere convergence test plot to paper.

@@
 origin. The problem domain for our simulations is $[-1.6, 1.6]^3$, and
-we take $R = 0.5$ and $\rho = 10^3$. The zones with $r > R$ are filled
-with an ambient material of very low density. We run this problem at
-multiple resolutions corresponding to jumps by a factor of two. For
+we take $R = 1.0 \text{cm}$ and $\rho = 10^3\ \text{g cm}^{-3}$. 
+The zones with $r > R$ are filled with an ambient material of very low density 
+$\rho = 10^{-8}\ \text{g cm}^{-3}$. We run this problem at multiple 
+resolutions corresponding to jumps by a factor of two. For
 comparison, at each grid point we evaluate the analytical potential of
@@
 \end{equation}
+We define the order of convergence $p$ between two simulations with a jump 
+in resolution of integer factor $m > 1$ as
+\begin{equation}
+  p = \text{log}_{m}\left(\frac{\text{Error}_{\text{low}}}{\text{Error}_{\text{high}}}\right).
+\end{equation}
+Here $\text{Error}_{\text{low}}$ is the $L^2$ error at the lower resolution 
+and $\text{Error}_{\text{high}}$ is the $L^2$ error at the higher resolution.
+We expect the error to converge at $p = 2$ given the discretization we choose. 
+For all simulations in this section and for all our main science simulations,
+we choose a relative error tolerance of $10^{-10}$ to be satisfied in the multigrid solve.
 
 \begin{figure}
-  \label{fig:gravity_convergence}
+  \centering
+  \includegraphics[scale=0.45]{plots/phi_comparison}
+  \caption{Comparison of the CASTRO gravitational potential to the analytical solution for: 
+    a sphere of uniform density; the same sphere, but with the potential normalized using the 
+    actual amount of mass on the grid instead of the mass of a perfect sphere; and, a 
+    cube of uniform density. Plotted also is a notional curve whose slope represents
+    perfect second order convergence.\label{fig:gravity_convergence}}
 \end{figure}
@@
 The results of this test are plotted in Figure
-\ref{fig:gravity_convergence}. We find that convergence is actually
-substantially better than second-order, and that the difference is
-largest at low resolutions. The explanation for this is that we are
-attempting to model a spherical object on a rectangular grid. At very
-low resolution, the object does not look very spherical. As the
-resolution is increased, the total amount of mass on the grid, as well
-as its location, will change as the sphere fills out. This means we
-are combining the true accuracy bonus from increased resolution with
-the artificial accuracy bonus from getting closer to solving the
-problem we are supposed to be solving.
+\ref{fig:gravity_convergence}. We find that at low resolution 
+convergence is actually substantially better than second-order. The 
+explanation for this is that we are attempting to model a spherical 
+object on a rectangular grid. At very low resolution, the object does 
+not look very spherical. As the resolution is increased, the total 
+amount of mass on the grid, as well as its location, will change 
+as the sphere fills out. This means we are combining the true accuracy 
+bonus from increased resolution with the artificial accuracy bonus 
+from getting closer to solving the problem we are supposed to be solving. 
+On the other hand, at the highest resolution shown the error slightly increases.
 
@@
 amount of mass actually on the grid, defined as the sum of the density
-multiplied by the volume of each cell for all cells that have greater
-than ambient density. The resolution study for this case is also
+multiplied by the volume of each cell for all cells whose density is 
+greater than ambient. The resolution study for this case is also
 plotted in Figure \ref{fig:gravity_convergence}. We still obtain
-convergence better than second-order, indicating that we still have
-the geometrical problem.
-
-The only way to fully eliminate this effect is to test a problem that
+convergence slightly better than second-order, indicating that we 
+still have the geometrical problem of the mass distribution changing.
+
+The only way to fully eliminate this effect is to use a test problem that
 does not change with resolution. The obvious companion is a cube of
@@
 the domain and thus has the same mass, so the only improvement comes
-from better sampling at higher resolution. The potential for this
+from better sampling at higher resolution. The gravitational potential for this
 object has been worked out analytically by \citet{waldvogel:1976} (see
@@
 R+z$, $z_1 = R-z$, and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
-\texttt{FORTRAN} and \texttt{C}, the inverse hyperbolic tangent is
+\texttt{Fortran} and \texttt{C}, the inverse hyperbolic tangent is
 \texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not}
 \texttt{atan2}). This formula is valid both inside and outside the
-cube. The normalized $L^2$ error norm for this problem is also shown
-in Figure \ref{fig:gravity_convergence}, and indeed we see
-near-perfect second-order scaling.
+cube. The normalized $L^2$ error for this problem is also shown
+in Figure \ref{fig:gravity_convergence}, and only for this problem 
+do we obtain perfect second-order scaling at all resolutions.
 
@@
 resolution. Since in the case of spherical objects on rectangular
-grids the effect is to artificially boost convergence with resolution,
+grids the effect may be to artificially boost convergence with resolution,
 in a simulation with spherical objects like stars one can envision a
@@
 factors.
-
-Another factor that may affect convergence testing is easily seen in
-Figure \ref{fig:gravity_convergence}, where at the highest resolution
-the error flattens out. This occurs because of a limitation in the
-accuracy of the multigrid solver. For problems of this size we find
-empirically that the multigrid solver does not converge for error
-tolerances smaller than about $10^{-11}$, which enforces a lower limit
-on the total error. This is also something to be aware of in assessing
-convergence for our full simulations.
 

### 2015-02-21 papers/wdmerger_I/paper.tex
Our old strategy for tracking the stellar centers of mass failed when the stars got significantly far away from their expected circular orbit locations. This has been fixed by storing the center of mass at every time step and updating it with the old COM velocity. We then refine our guess using the location-weighted sums over the domain. We also make the cutoff more physically meaningful by using the approximation to the Roche radius of Eggleton (1983). To ensure continuity at a restart, I restored Mike's old solution of writing out the mass and COM to a file in the checkpoint, and reading it

@@
 For stellar calculations, we separately record diagnostic 
-information about the stars, including their centers of mass, 
-and the volume contained within each star.
+information about the stars. Our strategy for tracking their 
+locations is as follows: at the beginning of the calculation, we store the 
+physical center of mass $\mathbf{x}_{\text{COM}}$ of the stars as determined 
+by Kepler's third law. We also store the space velocity $\mathbf{v}_{\text{COM}}$ 
+of the stars. Then, at each new time step we make a preliminary guess for their 
+location by updating the location using the old velocity, 
+$\mathbf{x}_{\text{COM}} \rightarrow \mathbf{x}_{\text{COM}} + \mathbf{v}_{\text{COM}} \Delta t$.
+We then refine our guess for the location and velocity of each star by computing a
+location-weighted sum of the mass and velocity over the computational domain. 
+To do this, we need a cutoff for determining what counts as part of the primary 
+and what counts as part of the secondary. For this we can use the formula provided by 
+\citet{eggleton:1983} for the effective Roche radius $r_L$ of a star in a binary,
+\begin{equation}
+  \frac{r_L}{a} = \frac{0.49 q^{2/3}}{0.6^{2/3} + \text{ln}(1 + q^{1/3})}.
+\end{equation}
+In this formula we can use $q = M_S / M_P$ for obtaining the Roche radius of the secondary,
+and use the inverse value of $q$ to obtain the Roche radius of the primary.
+To a reasonable approximation, all material within the effective Roche radius
+of each star can be counted as belonging to that star. From this we obtain the mass
+of each star as well as its location and velocity.
 

### 2015-02-21 source/Prob.cpp

@@ -121,8 +121,6 @@ Castro::wdCOM (Real time, Real& mass_p, Real& mass_s, Real* com_p, Real* com_s, 
 // at zones that are within twice the original radius of the white dwarf.
 
 void Castro::volInBoundary (Real               time,
-                	    Real*              com_p,
-			    Real*              com_s,
 			    Real&              vol_p,
 			    Real&              vol_s,
                             Real               rho_cutoff)
@@ -157,7 +155,7 @@ void Castro::volInBoundary (Real               time,
         const int* hi   = box.hiVect();
 
 	BL_FORT_PROC_CALL(CA_VOLUMEINDENSITYBOUNDARY,ca_volumeindensityboundary)
-	                  (BL_TO_FORTRAN(fab),lo,hi,dx,com_p,com_s,&sp,&ss,&rho_cutoff);
+	                  (BL_TO_FORTRAN(fab),lo,hi,dx,&sp,&ss,&rho_cutoff);
         vp += sp;
 	vs += ss;
     }

### 2015-02-21 source/modules/probdata.f90

@@ -21,6 +21,8 @@ module probdata_module
 
   integer :: npts_model
 
+  ! Initial binary orbit characteristics
+
   double precision :: mass_P_initial, mass_S_initial
   double precision :: central_density_P, central_density_S
   double precision :: radius_P_initial, radius_S_initial
@@ -67,6 +69,14 @@ module probdata_module
 
   ! Are we doing a single star simulation?
   logical :: single_star
+
+  ! Stores the center of mass location of the stars throughout the run
+  double precision :: com_P(3), com_S(3)
+  double precision :: vel_P(3), vel_S(3)
+  double precision :: mass_P,   mass_S
+
+  ! Stores the effective Roche radii
+  double precision :: roche_rad_P, roche_rad_S
 
 contains
 
@@ -260,7 +270,8 @@ contains
   subroutine binary_setup
 
     use bl_constants_module, only: ZERO, ONE
-    use meth_params_module, only: rot_period
+    use fundamental_constants_module, only: M_solar
+    use meth_params_module, only: do_rotation, rot_period
     use initial_model_module, only: init_1d
     use prob_params_module, only: center
     use meth_params_module, only: rot_axis
@@ -283,6 +294,15 @@ contains
     npts_model = 1024
 
     dx = 1.0d6
+
+    com_P = center
+    com_S = center
+
+    vel_P = ZERO
+    vel_S = ZERO
+
+    mass_P = mass_P_initial * M_solar
+    mass_S = mass_S_initial * M_solar
 
     ! Allocate arrays to hold the stellar models
 
@@ -326,6 +346,9 @@ contains
        center_P_initial(star_axis) = center_P_initial(star_axis) - a_P_initial
        center_S_initial(star_axis) = center_S_initial(star_axis) + a_S_initial
 
+       com_P = center_P_initial
+       com_S = center_S_initial
+
        ! Direction of initial motion -- essentially it's the position axis
        ! that is not star axis and that is also not the rotation axis.
 
@@ -340,6 +363,11 @@ contains
           if (star_axis .eq. 3) initial_motion_dir = 2
        else
           call bl_error("Error: probdata module: invalid choice for rot_axis.")
+       endif
+
+       if ( (do_rotation .ne. 1) .and. inertial ) then
+          vel_P(initial_motion_dir) = - orbital_speed_P
+          vel_S(initial_motion_dir) =   orbital_speed_S
        endif
 
     endif
[…]

### 2015-02-21 source/Problem.H
Our old strategy for tracking the stellar centers of mass failed when the stars got significantly far away from their expected circular orbit locations. This has been fixed by storing the center of mass at every time step and updating it with the old COM velocity. We then refine our guess using the location-weighted sums over the domain. We also make the cutoff more physically meaningful by using the approximation to the Roche radius of Eggleton (1983). To ensure continuity at a restart, I restored Mike's old solution of writing out the mass and COM to a file in the checkpoint, and reading it

@@ -6,8 +6,6 @@
 void wdCOM (Real time, Real& mass_p, Real& mass_s, Real* com_p, Real* com_s, Real* vel_p, Real* vel_s);
 
 void volInBoundary (Real               time,
-  	            Real*              com_p,
-		    Real*              com_s,
 		    Real&              vol_p,
 		    Real&              vol_s,
                     Real               rho_cutoff);

### 2015-02-21 source/Problem.f90

@@ -5,6 +5,7 @@ subroutine problem_checkpoint(int_dir_name, len)
   ! called by the IO processor during checkpoint
 
   use bl_IO_module
+  use probdata_module, only: com_P, com_S, vel_P, vel_S, mass_P, mass_S
 
   implicit none
 
@@ -18,6 +19,21 @@ subroutine problem_checkpoint(int_dir_name, len)
   do i = 1, len
      dir(i:i) = char(int_dir_name(i))
   enddo
+
+  un = unit_new()
+  open (unit=un, file=trim(dir)//"/COM", status="unknown")
+
+100 format(1x, g30.20, 1x, g30.20)
+
+  write (un,100) mass_P, mass_S
+  write (un,100) com_P(1), com_S(1)
+  write (un,100) com_P(2), com_S(2)
+  write (un,100) com_P(3), com_S(3)
+  write (un,100) vel_P(1), vel_S(1)
+  write (un,100) vel_P(2), vel_S(2)
+  write (un,100) vel_P(3), vel_S(3)
+
+  close (un)
 
 end subroutine problem_checkpoint
 
@@ -27,6 +43,7 @@ subroutine problem_restart(int_dir_name, len)
   ! called by ALL processors during restart 
 
   use bl_IO_module
+  use probdata_module, only: com_P, com_S, vel_P, vel_S, mass_P, mass_S
 
   implicit none
 
@@ -41,11 +58,27 @@ subroutine problem_restart(int_dir_name, len)
      dir(i:i) = char(int_dir_name(i))
   enddo
 
+  un = unit_new()
+  open (unit=un, file=trim(dir)//"/COM", status="old")
+
+100 format(1x, g30.20, 1x, g30.20)
+
+  ! here we read in, and this sets the values in the com module
+  read (un,100) mass_P, mass_S
+  read (un,100) com_P(1), com_S(1)
+  read (un,100) com_P(2), com_S(2)
+  read (un,100) com_P(3), com_S(3)
+  read (un,100) vel_P(1), vel_S(1)
+  read (un,100) vel_P(2), vel_S(2)
+  read (un,100) vel_P(3), vel_S(3)
+
+  close (un)
+
 end subroutine problem_restart
 
 
 
-! Returrn whether we're doing a single star simulation or not.
+! Return whether we're doing a single star simulation or not.
 
 subroutine get_single_star(flag)
 
@@ -66,6 +99,10 @@ end subroutine
[…]

### 2015-02-21 source/Problem_F.H

@@ -15,12 +15,21 @@ BL_FORT_PROC_DECL(WDCOM,wdcom)
      Real* com_s_x, Real* com_s_y, Real* com_s_z,
      Real* vel_p_x, Real* vel_p_y, Real* vel_p_z,
      Real* vel_s_x, Real* vel_s_y, Real* vel_s_z,
-     Real* mass_p, Real* mass_s);
+     Real* m_p, Real* m_s);
 
 BL_FORT_PROC_DECL(CA_VOLUMEINDENSITYBOUNDARY,ca_volumeindensityboundary)
-   (BL_FORT_FAB_ARG(rho), const int lo[], const int hi[],
-    const Real dx[], const Real com_p[], const Real com_s[], 
+   (BL_FORT_FAB_ARG(rho), const int lo[], const int hi[], const Real dx[],  
     Real* volp, Real* vols, Real* rho_cutoff);
 
 BL_FORT_PROC_DECL(GET_SINGLE_STAR,get_single_star)
    (int& flag);
+
+BL_FORT_PROC_DECL(SET_STAR_LOCATIONS,set_star_locations)
+   (const Real com_p[], const Real com_s[], 
+    const Real vel_p[], const Real vel_s[], 
+    const Real* mass_p, const Real* mass_s);
+
+BL_FORT_PROC_DECL(GET_STAR_LOCATIONS,get_star_locations)
+   (Real com_p[], Real com_s[], 
+    Real vel_p[], Real vel_s[], 
+    Real* mass_p, Real* mass_s);

### 2015-02-21 source/diagnostics/sum_integrated_quantities.cpp

@@ -251,10 +251,32 @@ Castro::sum_integrated_quantities ()
 
 #ifdef merger
 
+    // Compute the center of mass locations and velocities for the primary and secondary.
+    // We'll start by predicting the current locations of their centers by taking the 
+    // old locations and updating them using the old velocities and the time passed 
+    // since the last update. Send it back to Fortran, then clear it out so we can update
+    // using the full calculation.
+
+    BL_FORT_PROC_CALL(GET_STAR_LOCATIONS,get_star_locations)
+      (com_p, com_s, vel_p, vel_s, &mass_p, &mass_s);   
+
+    for ( int i = 0; i <= BL_SPACEDIM; i++ ) {
+      com_p[i] += vel_p[i] * dt * sum_interval;
+      com_s[i] += vel_s[i] * dt * sum_interval;
+    }
+
+    BL_FORT_PROC_CALL(SET_STAR_LOCATIONS,set_star_locations)
+      (com_p, com_s, vel_p, vel_s, &mass_p, &mass_s);   
+
+    for ( int i = 0; i <= BL_SPACEDIM; i++ ) {
+      com_p[i] = 0.0;
+      com_s[i] = 0.0;
+      mass_p   = 0.0;
+      mass_s   = 0.0;
+    }
+
     for (int lev = 0; lev <= finest_level; lev++)
     {
-
-      // Compute the center of mass locations and velocities for the primary and secondary.
 
       getLevel(lev).wdCOM(time, lev_mass_p, lev_mass_s, lev_com_p, lev_com_s, lev_vel_p, lev_vel_s);
 
@@ -307,11 +329,16 @@ Castro::sum_integrated_quantities ()
       }
     } 
 
+    // Send this information back to the Fortran probdata module
+
+    BL_FORT_PROC_CALL(SET_STAR_LOCATIONS,set_star_locations)
+      (com_p, com_s, vel_p, vel_s, &mass_p, &mass_s);
+
     // Compute effective radii of stars at various density cutoffs
 
     for (int lev = 0; lev <= finest_level; lev++)
         for (int i = 0; i <= 6; ++i) {
-	    getLevel(lev).volInBoundary(time, com_p, com_s, lev_vol_p[i], lev_vol_s[i], pow(10.0,i));
+	    getLevel(lev).volInBoundary(time, lev_vol_p[i], lev_vol_s[i], pow(10.0,i));
 	    vol_p[i] += lev_vol_p[i];
 	    vol_s[i] += lev_vol_s[i];
 	}

### 2015-02-22 documentation/userguide.tex
Add documentation on how to build your own problem.

@@
 \usepackage{natbib}
+
+\usepackage[usenames]{color}
+
+% For inserting code snippets;
+% use the lstlisting environment to access.
+% emacs doesn't parse $'s in code sections 
+% propertly, so that is why I have sprinkled 
+% some %$'s through the code. The coloring 
+% on the environment itself is still broken though.
+\usepackage{listings}
+
+\definecolor{gray}{rgb}       {0.8,0.8,0.8}
+\definecolor{light-blue}{rgb} {0.8,0.8,1.0}
+\definecolor{light-green}{rgb}{0.8,1.0,0.8}
+\definecolor{light-red}{rgb}  {1.0,0.9,0.9}
+
+\lstset{
+  basicstyle=\small\ttfamily,
+  frame=shadowbox, 
+  rulesepcolor=\color{gray},                                                             
+  backgroundcolor=\color{white}
+}
 
@@
 
-\chapter{Code Structure}
+A note on machine compatibility: this software is currently only developed and tested on
+machines that run Linux and come with the common GNU flavors of most pieces of software
+like \texttt{sed}. We do not guarantee that this software will work out of the box on OS X
+or other Unix derivatives; however, we generally try to avoid GNU-specific options, so that 
+trying to run this software on OS X will not be too painful. With that being said, this 
+software does heavily rely on \texttt{bash} and at present there are no plans to maintain
+compatibility with any other shell.
+
+\chapter{Directory Layout}
 
@@
 executable file. It is not normally built directly, but you can type \texttt{make} in that directory
-to build a vanilla executable built with the default settings.
+to build a vanilla executable with the default settings.
 
@@
 
-The \texttt{job\_scripts} directory contains utilities that are used for building and running
+The \texttt{job\_scripts} directory contains utilities that are used for running
 problem setups on various machine architectures.
@@
 that are being submitted as a result of this project.
+
+\chapter{How to Build a Problem}
+
+This chapter describes the necessary steps you need to take to build your own problem setup,
+and describes the underlying code that facilitates the job submission. If you are interested in
+doing white dwarf mergers and want to use the default settings, read Section \ref{sec:standard_problems}.
+If you are interested in setting up a different problem but using this software's job submission 
+system, you can do so; read Section \ref{sec:non-standard_problems} for more information.
+
+\section{Standard Problems}
+\label{sec:standard_problems}
+
+At a minimum, your working directory needs a \texttt{compile/} subdirectory with a \texttt{makefile}
+in it, and some \texttt{bash} script that you run to submit the jobs. If you want to build 
+with the default settings, then your makefile needs only to have the line:
+\begin{lstlisting}
+  include $(WDMERGER_HOME)/source/GNUmakefile
+\end{lstlisting}%$
+This includes the main makefile from the \texttt{source/} directory in \texttt{wdmerger}. 
+The main makefile sets all of our build options (compiler, number of space dimensions, 
+equation of state, etc.) and then includes the CASTRO makefile, which then runs as normal.
+If you want to adjust the build options, then simply add a line before this include. As an 
+example, if you want to build using a gamma-law equation of state, set:
+\begin{lstlisting}
+  EOS_dir = gamma_law_general/
+
+  include $(WDMERGER_HOME)/source/GNUmakefile 
+\end{lstlisting}%$
+
+The \texttt{make} variables in the main makefile are all set conditionally, so as long 
+as you set your variable before including the main makefile, it will be built as you expect.
+
+Once your makefile is setup, you can simply type \texttt{make} (or \texttt{make -j} 
+to use more processors and speed up the operation) at the command line. 
+If you're using Cray compilers on a supercomputer this will take several minutes;
+with GNU compilers it should be quick. When everything is compiled, if it was
+successful you should have an executable that looks like \texttt{Castro3d.Linux.Cray.Cray.MPI.ex}
+in the \texttt{compile} directory.
+
+Now \texttt{cd} up one level, and use your favorite text editor to create a simple 
+\texttt{bash} script, which for example purposes we will call \texttt{run.sh}.
+The first thing we should do is tell the script where to look to pull in 
+all the helper functions we'll need to execute a job:
+
+\begin{lstlisting}
+source $WDMERGER_HOME/job_scripts/run_utils.sh
+\end{lstlisting}%$
+
+Decide what directory you want to build the problem in. It is expected 
+to be a subdirectory of the \texttt{results/} directory, which will be
+automatically created for you:
+\begin{lstlisting}
+dir=results/test_problem/
+\end{lstlisting}
+Now, optionally decide how many processors you want to run with, and 
+how long you want the job to run:
+\begin{lstlisting}
+nprocs=128
+walltime=4:00:00
+\end{lstlisting}
+If you do not, the job submission script will default to running the job
+using one node for one hour. Finally, run your job:
+\begin{lstlisting}
+run $dir $nprocs $walltime
+\end{lstlisting}%$
+That is all you need to submit a job. Return to the terminal, and type \texttt{bash run.sh}
+and your job will execute. The helper function \texttt{run} does several things. First, 
+it creates the directory named \texttt{\$dir}. Then, it copies into \texttt{\$dir} the \texttt{inputs}
+and \texttt{probin} files from the root \texttt{wdmerger} source directory; these are the runtime
+files CASTRO needs to determine what your problem setup will be. They contain the default 
+parameters for merger runs. It also copies any auxiliary files one might need, like the 
+table for the equation of state, as well as the actual CASTRO executable you just built in the 
+\texttt{compile/} directory. Finally, it builds a script that you'll use for actually submitting 
+the job to the batch submission system on the machine you're on, and then submits that job. 
+Thus for a typical setup, your run directory will contain the following files when you submit:
+\begin{itemize}
+  \item \texttt{Castro3d*.ex}
+  \item \texttt{helm\_table.dat}
+  \item \texttt{inputs}
+  \item \texttt{probin}
+  \item \texttt{run\_script}
+\end{itemize}
+Typically you'll want to run multiple jobs at once, that vary things like the number of processors 
+or the runtime parameters. For example, suppose we want to run the same problem but we want to 
+explore what happens when we change the rotational period of the system. The main helper functions 
+will automatically handle this for you as long as your variable is part of a valid namespace in the inputs files.
+The orbital period is handled by the inputs parameter \texttt{castro.rotational\_period}. Therefore 
+we simply need to update \texttt{rotational\_period}. A sample script follows:
+
+\begin{lstlisting}
+source $WDMERGER_HOME/job_scripts/run_utils.sh
+
+nprocs=512
+walltime=12:00:00
+
+for period in 25.0 50.0 75.0 100.0
+do
+  dir=results/period$period/
+  rotational_period=$period
+  stop_time=$period
+  run $dir $nprocs $walltime
+done
+\end{lstlisting}%$
+This will create the subdirectories \texttt{results/period25.0} through \texttt{results/period100.0}
+and in each directory the inputs file will have the corresponding \texttt{castro.rotational\_period}.
+Each job will stop after one rotational period has been completed. When all have been completed, you should 
+type \texttt{bash run.sh} once more; this will move all output files (checkpoints, plot files, diagnostics)
+to the \texttt{output/} subdirectory of that problem location, and on supported HPC systems will also copy these 
+files to the long-term storage system (see Section \ref{sec:archiving} for more details). 
+Generally our analysis scripts expect all the output to be in the 
+\texttt{output/} subdirectory. Also, if the machine crashes before your run completes, running \texttt{bash run.sh}
+again will automatically restart your job from the most recent checkpoint.
+
+\section{Non-Standard Problems}
+\label{sec:non-standard_problems}
+
+If you want to build your own problem setup, you can do that too. The makefile and job submission 
+requirements are the same as in Section \ref{sec:standard_problems}, but you will need to provide 
+your own source files. They should all go in a \texttt{source/} subdirectory in your problem location.
+At minimum, you'll need a \texttt{Prob\_3d.f90} file (or the equivalent for lower dimensions) and a 
+probin file. The probin file can be empty, it just needs to be there because otherwise the helper 
+routines will make a copy of the default probin file. To get a sense of how this is done, look at 
+some of the existing test problems, in particular the \texttt{kelvin\_helmholtz} test. To tell the 
+code that you want to compile this source code instead of the default code, make a file called 
+\texttt{full\_source} and place it in the \texttt{source/} directory (you can just use \texttt{touch}
+to create an empty file). Then follow the rest of the instructions as normal, starting with the 
+compile step.
+
+The only source code that will be built from the default location is \texttt{sum\_integrated\_quantities.cpp},
+which controls the diagnostic output. Our version is a bit more extensive than what comes in CASTRO
+by default, so we include this in every build.
 
@@
 Each test directory is set up with a uniform pattern. The \texttt{source/} subdirectory includes all of the source code
-needed to build the test. The \texttt{compile/} directory contains the \texttt{makefile}, which builds the executable and creates
-the inputs files needed for the simulation. The \texttt{run\_test.sh} script executes the test by creating a subdirectory
+needed to build the test. The \texttt{compile/} directory contains the \texttt{GNUmakefile}, which builds the executable.
+The \texttt{run\_test.sh} script executes the test by creating a subdirectory
 for each parameter set and then submitting it with a batch submission script (for a vanilla Linux installation,
@@
 there will be recommended options and specialized submission scripts). The \texttt{analysis.py} Python script
-is used to analyze the output and generature figures. In summary, to run a test, perform the following series of commands:
+is used to analyze the output and generate figures. In summary, to run a test, perform the following series of commands:
 \begin{itemize}
@@
   \item \texttt{bash run\_test.sh}
+  \item \texttt{bash run\_test.sh} (after the execution has completed, to archive output)
   \item \texttt{python analysis.py} (after the execution has completed)

### 2015-02-23 documentation/userguide.tex
Added documentation about problem setup and parameters.

@@
 
-\clearpage
-
 \mainmatter
@@
 that are being submitted as a result of this project.
+
+\chapter{Merger Problem}
+
+This chapter describes the setup of the white dwarf merger problem. It includes descriptions of 
+the physics involved, the code options and runtime parameters, the diagnostic output 
+and other I/O information, and the analysis routines. A full description of the physics 
+and the associated numerical algorithms is included in the CASTRO code paper \cite{castro}.
+
+\section{Physical Description}
+
+We model two white dwarfs in a rotating binary system. The white dwarfs are composed of 
+a mixture of elements that is principally carbon and oxygen, and are individually supported 
+against gravity by degenerate electron material at non-zero temperature. This matter is described well 
+for arbitrary degeneracy and relativity levels by the Helmholtz equation of state \cite{timmes_swesty:2000}.
+It includes ideal gas contributions from the ions, a simple blackbody recipe for the contribution from 
+the radiation field, and optionally Coulomb corrections.
+
+CASTRO includes full self-gravity with the Poisson equation, as well as the ability to perform a
+simulation in a rotating reference frame with fixed rotation period. By default the $z$ axis will 
+be the rotation axis, with the stars initially aligned along the $x$ axis. We select the 
+orbital period of the system, and the masses of the \textit{primary} (more massive) and 
+\textit{secondary} (less massive) white dwarfs. We then use Kepler's third law to determine the 
+appropriate initial orbital distance. The center of mass of the system corresponds with the center 
+of the domain, which is the origin; the domain is $1.024 \times 10^{10}\ \text{cm}$ wide in each 
+coordinate direction, so that each dimension runs from $-5.12 \times 10^{9}\ \text{cm}$ to 
+$+5.12 \times 10^{9}\ \text{cm}$. In a rotating reference frame, we set the stars at rest. 
+We can instead run a simulation in the inertial reference frame. To make it equivalent, we
+start all zones in the domain with a velocity corresponding to the equivalent rotating 
+reference frame velocity. This means that the stars will be synchronously rotating (though we
+may relax this assumption in future work).
+
+At present nuclear reactions are not enabled, but they will be soon.
+
+\section{Runtime Parameters}
+
+There are a few CASTRO input parameters that you will typically want to control when running problems. 
+Among the most major is whether we are doing rotation (\texttt{castro.do\_rotation}) and if so, 
+what the rotational period is (\texttt{castro.rotational\_period}). You'll also want to change the 
+\texttt{stop\_time}, which is when the simulation will terminate. You should also avoid flooding 
+yourself with plotfiles and checkpoints by setting \texttt{amr.plot\_files\_output} and 
+\texttt{amr.checkpoint\_files\_output}, respectively; these are the simulation time differences
+between successive plotfiles and checkpoints. Note that CASTRO does not adjust the timestep 
+to meet this criterion, so it will generate these at the first timestep that passes this interval.
+
+The \texttt{wdmerger} parameters of interest are all controlled through the \texttt{probin} file. 
+These control the layout of the stars on the grid, the tagging options for refinement, and some other things. 
+A \texttt{probin} file is composed of three Fortran namelists; a namelist is enclosed in 
+\texttt{\&namelist ... /}. The \texttt{fortin} namelist controls the main system parameters. 
+In the following list we describe the parameters; you can see their current defaults by 
+examining \texttt{source/probin}.
+\begin{itemize}
+  \item \texttt{mass\_P}: mass of the primary white dwarf, in solar masses (real)
+  \item \texttt{mass\_S}: mass of the secondary white dwarf, in solar masses (real; if less than zero, 
+    then no secondary star will be generated, and the primary star will be loaded onto the grid at the 
+    center of the domain)
+  \item \texttt{nsub}: number of sub-intervals to divide a zone into when interpolating from the initial model (integer)
+  \item \texttt{orbital\_kick}: whether or not to kick the stars with an orbital velocity corresponding to their 
+    Keplerian rotation rate (logical; this only makes sense in the inertial reference frame with \texttt{castro.do\_rotation = 0})
+  \item \texttt{damping}: whether or not to exert a constant damping force on the orbit (logical)
+  \item \texttt{damping\_alpha}: strength of damping force (real)
+  \item \texttt{bulk\_velx}: uniform velocity to give to all gas in the domain in the $x$-direction (real)
+  \item \texttt{bulk\_vely}: uniform velocity to give to all gas in the domain in the $y$-direction (real)
+  \item \texttt{bulk\_velz}: uniform velocity to give to all gas in the domain in the $z$-direction (real)
+  \item \texttt{interp\_temp}: whether or not to take temperature as given when interpolating from our initial model (logical;
+    if false, we will interpolate based on pressure)
+  \item \texttt{maxTaggingRadius}: fractional size of the domain outside of which we won't perform any refinement (real)
+\end{itemize}
+
+The \texttt{tagging} namelist is described in the AMR chapter of the CASTRO user's guide. At present, we 
+only tag based on absolute values of the density and temperature.
+
+The \texttt{extern} namelist contains parameters that are used by the microphysics modules. In particular, 
+the equation of state usually relies on parameters passed in this way. For the Helmholtz EOS, the relevant 
+parameters are \texttt{eos\_input\_is\_constant}, which determines whether or not we update the internal 
+energy on an EOS call, and \texttt{use\_eos\_coulomb}, which determines whether we include Coulomb 
+interactions in the EOS evaluation.
 

### 2015-02-23 documentation/userguide.tex
To obtain automatic inputs replacement when using the run() script, you now need to include the appropriate namespace -- but with an underscore, not a period, after the namespace, because of bash variable naming rules. If you want to adjust the rotational period, use castro_rotational_period. as an example. The helper routines will detect these namespaces and perform the appropriate substitution before checking the inputs file.

@@
 The orbital period is handled by the inputs parameter \texttt{castro.rotational\_period}. Therefore 
-we simply need to update \texttt{rotational\_period}. A sample script follows:
+we simply need to update a variable named \texttt{castro\_rotational\_period}. 
+Note that we use an underscore (\texttt{castro_}) and not a period (\texttt{castro.}) after the namespace.
+This is simply because \texttt{bash} does not permit periods in variable names; but the helper routine will 
+automatically take care of this for you when looking for the appropriate variable in the inputs file.
+A sample script follows:
 
@@
   dir=results/period$period/
-  rotational_period=$period
+  castro_rotational_period=$period
   stop_time=$period

### 2015-02-23 job_scripts/run_utils.sh

@@ -360,13 +360,41 @@ function copy_files {
     for var in $input_vars
     do
 
+	# Get the namespace of the variable by 
+	# looking at what is before the first underscore.
+	# If it is not a valid namespace (or one of the few variables
+	# without a namespace), then skip this variable.
+
+	namespace=$(echo $var | cut -d _ -f 1)
+
+	is_inputs_var=0
+
+	if ([ $var == "stop_time" ] || [ $var == "max_step" ])
+	then
+
+	    is_inputs_var=1
+	    inputs_var_name=$var	    
+
+	elif ( [ $namespace == "amr" ] || [ $namespace == "castro" ] || 
+	       [ $namespace == "geometry" ] || [ $namespace == "gravity" ] || 
+               [ $namespace == "mg" ] )
+	then
+
+	    # Remove the namespace from the variable, then
+	    # replace it with the correct period.
+	    is_inputs_var=1
+	    inputs_var_end=$(echo ${var#$namespace\_})
+	    inputs_var_name="$namespace.$inputs_var_end"
+
+	fi
+
 	# The -q option to grep means be quiet and only 
 	# return a flag indicating whether $var is in inputs.
 	# The [[:space:]]* tells grep to ignore any whitespace between the 
 	# variable and the equals sign. See:
         # http://www.linuxquestions.org/questions/programming-9/grep-ignoring-spaces-or-tabs-817034/
 
-	if grep -q "$var[[:space:]]*=" $dir/inputs 
+	if ([ $is_inputs_var -eq 1 ] && (grep -q "$inputs_var_name[[:space:]]*=" $dir/inputs)) 
 	then
 
 	    # OK, so the parameter name does exist in the inputs file;
@@ -375,14 +403,21 @@ function copy_files {
 	    # ${!var} is the value held by that variable. See:
 	    # http://www.tldp.org/LDP/abs/html/bashver2.html#EX78
 	    # http://stackoverflow.com/questions/10955479/name-of-variable-passed-to-function-in-bash
-            sed -i "s/$var.*=.*/$var = ${!var}/g" $dir/inputs
+            sed -i "s/$inputs_var_name.*=.*/$inputs_var_name = ${!var}/g" $dir/inputs
 	fi
 
 	# Do the same thing for the probin file.
+	# Since we don't have the namespace to rely on 
+	# to guard against false positives, we only replace
+	# in cases that there's a space both before and 
+	# after the variable, signifying a full word match.
[…]

### 2015-02-23 tests/gravity/boundary_condition_comparison/run_test.sh

@@ -5,8 +5,8 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 mass_P=0.90
 mass_S=0.60
 
-do_hydro=0
-do_rotation=0
+castro_do_hydro=0
+castro_do_rotation=0
 max_step=0
 
 # Now set up the runs
@@ -24,8 +24,8 @@ fi
 for l in $(seq $minell $maxell)
 do
   dir=$results_dir/$l
-  direct_sum_bcs=0
-  max_multipole_order=$l
+  gravity_direct_sum_bcs=0
+  gravity_max_multipole_order=$l
   run $dir $nprocs $walltime
 done
 
@@ -37,5 +37,5 @@ fi
 # Now do the 'exact' direct summation, for comparison purposes
 
 dir=$results_dir/true
-direct_sum_bcs=1
+gravity_direct_sum_bcs=1
 run $dir $nprocs $walltime

### 2015-02-23 tests/gravity/uniform_cube_sphere/run_test.sh

@@ -2,17 +2,17 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 
 # Problem inputs choices
 
-prob_lo="-1.6 -1.6 -1.6"
-prob_hi=" 1.6  1.6  1.6"
+geometry_prob_lo="-1.6 -1.6 -1.6"
+geometry_prob_hi=" 1.6  1.6  1.6"
 
-do_hydro=0
-do_rotation=0
+castro_do_hydro=0
+castro_do_rotation=0
 
-small_dens="1.e-10"
+castro_small_dens="1.e-10"
 
 max_step=0
 
-direct_sum_bcs=1
+gravity_direct_sum_bcs=1
 
 # Loop over problem choices
 
@@ -26,7 +26,7 @@ do
 
     dir=$results_dir/p$problem/n$ncell
     problem=$p
-    n_cell="$ncell $ncell $ncell"
+    amr_n_cell="$ncell $ncell $ncell"
 
     if [ $MACHINE == "BLUE_WATERS" ]; then
 	if   [ $ncell -eq 16  ]; then

### 2015-02-23 tests/hydro/circular_orbit/run_test.sh

@@ -5,12 +5,12 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 mass_P=0.90
 mass_S=0.90
 
-rotational_period=100.0
-stop_time=$rotational_period
+castro_rotational_period=100.0
+stop_time=$castro_rotational_period
 
 num_periods=50
 
-final_stop_time=$(echo "$(num_periods) * $(rotational_period)" | bc -l)
+final_stop_time=$(echo "$(num_periods) * $(castro_rotational_period)" | bc -l)
 
 # Loop over the resolutions in question
 
@@ -24,15 +24,15 @@ do
 
       # First do the relevant updates in the inputs file.
       
-      n_cell="$ncell $ncell $ncell"
-      grav_source_type=$gs
+      amr_n_cell="$ncell $ncell $ncell"
+      castro_grav_source_type=$gs
 
       if [ $rs == 0 ]; then
-	  do_rotation=0
+	  castro_do_rotation=0
 	  orbital_kick=T
       else
-	  do_rotation=1
-	  rot_source_type=$rs
+	  castro_do_rotation=1
+	  castro_rot_source_type=$rs
 	  orbital_kick=F
       fi
 
@@ -56,17 +56,17 @@ do
 	    # the stopping time equal to the final stopping time.
 	    # Otherwise, add a full orbital period.
 
-	    if [ $(echo "$time > ($final_stop_time - $rotational_period)" | bc -l) -eq 1 ]; then
+	    if [ $(echo "$time > ($final_stop_time - $castro_rotational_period)" | bc -l) -eq 1 ]; then
 		new_time=$final_stop_time
 	    else
-		new_time=$(echo "$time + $rotational_period" | bc -l)
+		new_time=$(echo "$time + $castro_rotational_period" | bc -l)
 	    fi
 
 	    # While we are here, we don't want to do too many orbits for the options
 	    # we do not prefer. We will stop at a single orbit for everytning but our
 	    # preferred options in both the rotating and inertial frames.
 
-	    new_time_flag=$(echo "$new_time >= $rotational_period" | bc -l)
+	    new_time_flag=$(echo "$new_time >= $castro_rotational_period" | bc -l)
 
 	    if [ $new_time_flag -eq 1 ]; then
 		if [[ $gs -ne 4 ]]; then

### 2015-02-23 tests/hydro/evrard_collapse/run_test.sh

@@ -4,10 +4,10 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 
 stop_time=15.0
 
-prob_lo=" -2.56e9 -2.56e9 -2.56e9"
-prob_hi="  2.56e9  2.56e9  2.56e9"
+geometry_prob_lo=" -2.56e9 -2.56e9 -2.56e9"
+geometry_prob_hi="  2.56e9  2.56e9  2.56e9"
 
-do_rotation=0
+castro_do_rotation=0
 
 # Loop over the resolutions in question
 
@@ -17,8 +17,8 @@ do
   do
     dir=$results_dir/gs$gs/n$ncell
 
-    grav_source_type=$gs
-    n_cell="$ncell $ncell $ncell"
+    castro_grav_source_type=$gs
+    amr_n_cell="$ncell $ncell $ncell"
 
     if [ $MACHINE == "BLUE_WATERS" ]; then
 

### 2015-02-23 tests/hydro/freefall/run_test.sh

@@ -5,11 +5,11 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 mass_P=0.90
 mass_S=0.81
 
-do_rotation=0
-rotational_period=100.0
+castro_do_rotation=0
+castro_rotational_period=100.0
 
-plot_per=1.0
-check_per=1.0
+amr_plot_per=1.0
+amr_check_per=1.0
 
 # We can work out the stopping time using the formula
 # t_freefall = rotational_period / (4 * sqrt(2))
@@ -29,7 +29,7 @@ stop_time=$(echo "0.90 * $(rotational_period) / 4.0 / e(0.5*l(2))" | bc -l)
 for ncell in 64 128 256
 do
   dir=$results_dir/n$ncell
-  n_cell="$ncell $ncell $ncell"
+  amr_n_cell="$ncell $ncell $ncell"
 
   if [ $MACHINE == "BLUE_WATERS" ]; then
 

### 2015-02-23 tests/hydro/kelvin_helmholtz/run_test.sh

@@ -2,13 +2,13 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 
 # Problem-specific variables
 
-is_periodic="1 1"
+geometry_is_periodic="1 1"
 
-prob_lo="0.0 0.0"
-prob_hi="1.0 1.0"
+geometry_prob_lo="0.0 0.0"
+geometry_prob_hi="1.0 1.0"
 
-lo_bc="0 0"
-hi_bc="0 0"
+castro_lo_bc="0 0"
+castro_hi_bc="0 0"
 
 # Loop over problem choices
 
@@ -36,22 +36,22 @@ do
 
       problem=$p
       bulk_velocity=$vel
-      n_cell="$ncell $ncell $ncell"
+      amr_n_cell="$ncell $ncell $ncell"
       
       # Determine stopping time based on problem of interest
 
       if [ $p -eq 1 ]; then
 	  stop_time=2.0
-	  plot_per=0.05
+	  amr_plot_per=0.05
       elif [ $p -eq 2 ]; then
 	  stop_time=2.0
-	  plot_per=0.05
+	  amr_plot_per=0.05
       elif [ $p -eq 3 ]; then
 	  stop_time=10.0
-	  plot_per=0.1
+	  amr_plot_per=0.1
       fi
 
-      check_per=$plot_per
+      amr_check_per=$amr_plot_per
 
       # Set number of processors based on amount of work
 

### 2015-02-23 tests/hydro/single_star_hse/run_test.sh

@@ -10,7 +10,7 @@ mass_S="-1.00"
 for ncell in 64 128 256
 do
   dir=$results_dir/n$ncell
-  n_cell="$ncell $ncell $ncell"
+  amr_n_cell="$ncell $ncell $ncell"
 
   if [ $MACHINE == "BLUE_WATERS" ]; then
 

### 2015-02-23 tests/performance/strong_scaling/run_test.sh

@@ -4,8 +4,8 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 
 max_step=10
 
-plot_files_output=0
-checkpoint_files_output=0
+amr_plot_files_output=0
+amr_checkpoint_files_output=0
 
 # For strong scaling, we fix the problem size and increase the number of processors.
 

### 2015-02-23 tests/performance/weak_scaling/run_test.sh

@@ -4,15 +4,15 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 
 max_step=10
 
-plot_files_output=0
-checkpoint_files_output=0
+amr_plot_files_output=0
+amr_checkpoint_files_output=0
 
 # For weak scaling, we increase the problem size in parallel with the number of processors.
 
 for ncell in 64 128 256 512
 do
   dir=$results_dir/$nprocs
-  n_cell="$ncell $ncell $ncell"
+  amr_n_cell="$ncell $ncell $ncell"
 
   if [ $ncell -eq "64" ]; then
       nprocs="16"

### 2015-02-23 papers/wdmerger_I/paper.tex
Our old strategy for tracking the stellar centers of mass failed when the stars got significantly far away from their expected circular orbit locations. This has been fixed by storing the center of mass at every time step and updating it with the old COM velocity. We then refine our guess using the location-weighted sums over the domain. We also make the cutoff more physically meaningful by using the approximation to the Roche radius of Eggleton (1983). To ensure continuity at a restart, I restored Mike's old solution of writing out the mass and COM to a file in the checkpoint, and reading it

@@
 \usepackage{bm}
+
+% allows us to force the location of a figure
+\usepackage{float}
 
@@
 of each star can be counted as belonging to that star. From this we obtain the mass
-of each star as well as its location and velocity.
+of each star as well as its location and velocity. Once we have the new centers of mass,
+we compute the effective radius of each star at various density cutoffs. This involves 
+computing the volume $V$ of all zones within the Roche radius of the star's center 
+whose density is greater than the cutoff. We then compute $r_{\text{eff}} = (3V/4\pi)^{1/3}.$
 
@@
 moving at a uniform bulk velocity will appear different than an
-equivalent stationary simulation. If true, this means that codes such
-as CASTRO may be fundamentally inappropriate for merger problems
-(though any errors should be greatly diminished in a reference frame
-that rotates with the initial binary orbit). Recently, this has come
-up in two ways which are of note for us in the present study. We will
-explain these situations and run tests to determine whether this
-actually is a significant concern.
-
-\citet{arepo} performed a Kelvin-Helmholtz instability test and showed
+equivalent stationary simulation. If true, we need to understand 
+the importance of this effect when deciding whether to trust the 
+output of a code like CASTRO when applied for merger problems.
+Recently, this has come up in two ways which are of note for us 
+in the present study. We will explain these situations and display 
+the results of tests we have run to determine whether this 
+actually is a significant concern for our study.
+
+\citet{arepo} (hereafter, S10) performed a Kelvin-Helmholtz instability test and showed
 that (at low resolution) a fixed-grid code failed to develop the
 expected fluid instability when the whole fluid was moving at a
-strongly supersonic uniform velocity. This contrasted with the results
+strongly supersonic uniform velocity. (See also \citet{wadsley:2008}, 
+who used the FLASH code to simulate a hot bubble subject to mixing 
+by the Kelvin-Helmholtz instability, and also found that the mixing was affected by a 
+uniform bulk velocity.) This contrasted with the results
 of the moving-mesh code AREPO being presented in that study, which
-demonstrated Galilean invariance even at large bulk velocities. If
-verified, this claim would have important consequences on how much we
-can trust the ability of CASTRO to test the violent merger progenitor
-model. Shearing between the material flowing out of the secondary and
-material near the surface of the primary may trigger fluid
+demonstrated Galilean invariance even at large bulk velocities. 
+Inability to correctly model the Kelvin-Helmholtz instability would 
+ave important consequences for how much we can trust the ability of 
+CASTRO to test the violent merger progenitor model. Shearing between 
+the material flowing out of the secondary and material near the 
+surface of the primary may trigger fluid
 instabilities that play an important role in the evolution of that
@@
 that a detonation (or lack thereof) is not numerically
-seeded. However, \citet{robertson:2010} observe that Galilean
+seeded.
+
+\begin{figure*}[ht]
+  % Note that since we have periods in the filename, we need to double-
+  % bracket the filename so that includegraphics doesn't think it's 
+  % looking for something with the wrong extension. See:
+  % http://tex.stackexchange.com/a/10575
+  \centering
+  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p2_low_res_collage}}}
+  \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
+    conditions given by Equations \ref{eq:kh_ic_b_ramp} and \ref{eq:kh_ic_b}. 
+    The rows each represent a different bulk fluid velocity $v$ and 
+    the columns each represent a grid resolution $n$ (the number of 
+    zones per spatial dimension). The highest velocity simulation, 
+    $v=100$, corresponds to approximately Mach 70. Compare to 
+    \citet{robertson:2010}, Figure 7. \label{fig:kh_p2}}
+\end{figure*}
+
+\citet{robertson:2010} (hereafter, R10) observe that Galilean
 invariance of simulation results for the Euler equations occurs only
@@
 velocity, which substantially reduces the numerical
-diffusion. \citeauthor{robertson:2010} argue that the differences seen
+diffusion. R10 argue that the differences seen
 between the moving-mesh and fixed-grid code are caused by the
@@
 fundamentally alter the large-scale modes. Small-scale instabilities
-are seeded in this problem by the choice of a sharp initial
-discontinuity between the fluids. Crucially though,
-\citeauthor{robertson:2010} point out that this problem does not
+are seeded by the choice of a sharp initial discontinuity between the 
+fluids in the problem posed by S10. Crucially though,
+R10 point out that this problem does not
 converge with resolution and so it is not possible to know the correct
 behavior of this problem. As such, we do not know whether the
-small-scale modes found in AREPO are real, so the problem is not
+small-scale modes found in AREPO are real, and the problem is not
 useful in formally discriminating between methodologies. They instead
@@
 resolution and the uniform flow speed).  We will see whether we can
-reproduce this result, as well as a result by \cite{wadsley:2008}, who
-used the FLASH code to simulate a hot bubble subject to mixing by the
-Kelvin-Helmholtz instability, and also found that the mixing was
-affected by a uniform bulk velocity.
+reproduce this result.
 
@@
 the following way. The problem domain runs from 0 to 1 in both the $x$
-and $y$ directions. Though this is a two-dimensional test, we run
-CASTRO in 3D and replicate the problem along 8 cells in the $z$
-direction; the physical size of these zones is determined so that $dz
-= dx = dy$, a requirement at present in CASTRO. The problem involves a
-fluid slab of density $\rho_1 = 2.0$ traveling rightward in the
-$x$-direction at velocity $v_1 = 0.5$, sandwiched by a fluid of
-density $\rho_2 = 1.0$ traveling leftward at velocity $v_2 =
+and $y$ directions. This is a two-dimensional test, so we run
+CASTRO in 2D mainly to avoid extra computational expense; in 3D, it 
+would merely involve replicating the problem in the $z$ direction.
+The problem involves a fluid slab of density $\rho_2 = 2.0$ traveling rightward in the
+$x$-direction at velocity $v_2 = 0.5$, sandwiched by a fluid of
+density $\rho_1 = 1.0$ traveling leftward at velocity $v_1 =
 -0.5$. The density gradient is in the $y$ direction, so this creates a
@@
 sets of initial conditions (ICs), which we follow
-\citeauthor{robertson:2010} in calling ICs A and B. They differ in
+R10 in calling ICs A and B. They differ in
 their ramp function ($R_A$ and $R_B$ respectively), as well as the
@@
 \begin{align}
-  R_A &= \begin{cases} 0 & |y - 0.5| > 0.25 \\ 1 & |y - 0.5| < 0.25 \end{cases} \\
-  R_B &= \Big\{\left[1 + e^{-2(y-0.25)/\Delta_y}\right]\left[1 + e^{2(y-0.75)/\Delta_y}\right]\Big\}^{-1} \\
-  v^\prime_A &= w_0\, \text{sin}\left(n_A\, \pi\, x\right) \left\{e^{-(y-0.25)^2 / 2\sigma^2} + e^{-(y-0.75)^2/2\sigma^2}\right\} \\
-  v^\prime_B &= w_0\, \text{sin}\left(n_B\, \pi\, x\right).
+  R_A &= \begin{cases} 0 & |y - 0.5| > 0.25 \\ 1 & |y - 0.5| < 0.25 \end{cases} \label{eq:kh_ic_a_ramp}\\
+  R_B &= \Big\{\left[1 + e^{-2(y-0.25)/\Delta_y}\right]\left[1 + e^{2(y-0.75)/\Delta_y}\right]\Big\}^{-1} \label{eq:kh_ic_b_ramp}\\
+  v^\prime_A &= w_0\, \text{sin}\left(n_A\, \pi\, x\right) \left\{e^{-(y-0.25)^2 / 2\sigma^2} + e^{-(y-0.75)^2/2\sigma^2}\right\} \label{eq:kh_ic_a}\\
+  v^\prime_B &= w_0\, \text{sin}\left(n_B\, \pi\, x\right). \label{eq:kh_ic_b}
 \end{align}
+
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p3_low_res_collage}}}
+  \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
+    conditions given by Equations \ref{eq:kh_p3_rho} through \ref{eq:kh_p3_vy}. 
+    The meaning of the rows and columns is the same as in Figure \ref{fig:kh_p2}.
+    \label{fig:kh_p3}}
+\end{figure*}
+
 Here $w_0 = 0.1$ is the scale of the velocity perturbation, $\sigma =
 0.05/\sqrt{2}$ controls the width of the Gaussian for IC A, and
-$\Delta_y$ is the transition distance scale for the smooth ramp of IC
+$\Delta y = 0.05$ is the transition distance scale for the smooth ramp of IC
 B. The pressure everywhere is set to $p = 2.5$, and we run this with a
 gamma-law equation of state set to $\gamma = 5/3$. Plotfiles are
-generated every 0.01 seconds, and the problem is run until $t = 2$.
+generated every 0.05 seconds, and the problem is run until $t = 2$.
 
@@
 for each set of initial conditions run the problem at resolutions of
-$64^2$, $128^2$, $256^2$, $512^2$. In addition, for each initial
-condition we run a simulation at the very high resolution of $2560^2$
-for the stationary problem only. This serves as a reference solution
-to gauge the extent to which the bulk flow affects the development of
-the fluid instability.
-
-\subsubsection{Hot Bubble}\label{sec:hot_bubble}
+$64^2$, $128^2$, $256^2$, $512^2$. For context, in these units the 
+sound speed is $c\approx 0.7$. In addition, for each initial
+condition we run simulations at the higher resolutions of $1024^2$,
+$2048^2$, and $4096^2$ for the stationary problem only. These serve 
+as a reference solution to gauge the extent to which the bulk flow 
+affects the development of the fluid instability, and to determine 
+if the problem is numerically converged.
+
+We find the same result as R10 for IC A, which is equivalent to the 
+test proposed by S10: at low resolutions and high bulk velocity, the 
+Kelvin-Helmholtz instability completely fails to develop. However, 
+the problem does not converge even qualitatively at the highest 
+resolutions we used. Our results are very similar to Figure 3 of 
+R10 so we do not show them here. For IC B, our results can be seen 
+for the normal resolutions and all velocities in Figure \ref{fig:kh_p2}.
+At low resolutions and very large bulk velocities, the fluid 
+does get significantly disrupted by numerical error. However, this 
+effect quickly converges away with resolution and qualitatively 
+at $512^2$ resolution the solution is nearly identical to the stationary 
+$v=0$ problem. We agree with R10 that this problem does converge 
+with resolution and is not subject to numerically-seeded secondary 
+instabilities at the stopping time. This is evident even at low resolutions
+by examining the first row of Figure \ref{fig:kh_p2}.
+
+\begin{figure*}[ht]
+  \includegraphics[trim = 0in 1.25in 0.25in 1.25in, clip, scale=0.65]{{{plots/kh_p3_high_res_collage}}}
+  \caption{Time series of the Kelvin-Helmholtz problem proposed by \citet{mcnally:2012}
+    as the simulation is just starting to go non-linear. The rows represent resolution, 
+    where $n$ is the number of grid cells per spatial dimension, and the columns are 
+    different snapshots in time.\label{fig:kh_p3_high_res}}
+\end{figure*}
+
+\citet{mcnally:2012} published another Kelvin-Helmholtz problem that 
+is well-posed in the sense that it converges with resolution and 
+is not subject to uncontrollable numerical instabilities. Though they 
+were not explicitly interested in the question of Galilean invariance, 
+we visit that issue here to see what can be learned. The initial 
+conditions for this problem are:
+\begin{align}
+  \rho &= \begin{cases} \rho_1 - \rho_m e^{\frac{y-0.25}{\Delta y}} & 0.25 > y \geq 0 \\ 
+                        \rho_2 + \rho_m e^{\frac{0.25-y}{\Delta y}} & 0.5 > y \geq 0.25 \\
+                        \rho_2 + \rho_m e^{\frac{y-0.75}{\Delta y}} & 0.75 > y \geq 0.5 \\
+                        \rho_1 - \rho_m e^{\frac{0.75-y}{\Delta y}} & 1 > y \geq 0.75 \end{cases} \label{eq:kh_p3_rho}\\
+  v_x &= \begin{cases} v_1 - v_m e^{\frac{y-0.25}{\Delta y}} & 0.25 > y \geq 0 \\
+                       v_2 + v_m e^{\frac{0.25-y}{\Delta y}} & 0.5 > y \geq 0.25 \\
+                       v_2 + v_m e^{\frac{y-0.75}{\Delta y}} & 0.75 > y \geq 0.5 \\
+                       v_1 - v_m e^{\frac{0.75-y}{\Delta y}} & 1 > y \geq 0.75 \end{cases} \label{eq:kh_p3_vx} \\
+  v_y &= w_0\, \text{sin}\left(4\pi x\right). \label{eq:kh_p3_vy}
+\end{align}
+Here $\Delta y = 0.025$, $w_0 = 0.01$, $v_m = (v_1 - v_2) / 2$, $\rho_m = (\rho_1 - \rho_2) / 2$, 
+and the other symbols have the same meaning as above (this means the flow direction
+is reversed compared to the original paper, so as to achieve consistency with the 
+other simulations presented here). We run this problem at all the same resolutions 
+and bulk velocities as the previous two problems. The results for the normal resolutions 
+at $t = 2.0$ are displayed in Figure \ref{fig:kh_p3}. We see a similar pattern as for 
+the test proposed by R10: as we get to higher flow speeds we need to have higher 
+spatial resolution to compensate for the increased numerical diffusion. However, 
+the accuracy is much lower for the highest bulk velocities. This is probably because 
+the initial conditions feature a transition that is twice as sharp (compare the $\Delta y$ 
+parameters in each case). This means that there are smaller scale numerical 
+perturbations on the grid, which can be amplified by numerical noise in the time 
+evolution of the system.
+
+\citet{hopkins:2014} performed this test as part of the testing of their code GIZMO. 
+They showed the late-time evolution of this system, when non-linear effects have 
+taken over and significantly disrupted the initial flow. However, at low resolution the 
+tested grid algorithm had failed to disrupt both for $v = 0$ and $v = 10$. We too ran 
+this problem until $t = 10$, and confirm that the Kelvin-Helmholtz instability damps out 
+at low resolution but goes strongly non-linear and disrupts the flow at high resolution.
+However, we strongly emphasize the point that this does not objectively 
+demonstrate a defiency in fixed-grid codes for this problem. We can only determine 
+the validity of a method when we have a trustworthy, converged solution to compare to, and 
+this is lacking for this problem at late times. As observed by \citeauthor{mcnally:2012}, 
+this is because the secondary instabilities form for this problem when the whorls of 
+the Kelvin-Helmholtz tendrils stretch out and create gradients that approach the 
+grid resolution. This is prime breeding ground for numerical noise. But because 
+the nature of this noise will depend on the resolution, it will be very different for 
+simulations at different resolutions. If these instabilities are seeded because of 
+this resolution-dependent noise and are not seeded instead in a controlled manner 
+such that they appear at the same time and location, then we simply cannot draw 
+any conclusions that bear on the question of verification from this test at late times. 
+Figure \ref{fig:kh_p3_high_res} provides a sense of this by examining the crucial 
+time at which the transition from the linear to the non-linear regime is occurring. 
+At all of these very high resolutions the secondary instabilities develop, but they 
+occur at different times and have different spatial scales for each resolution.
+
+We conclude that large bulk motions of fluid can have very significant effects 
+on numerical calculations of shear mixing in fixed-grid codes, but that this effect 
+diminishes with increasing resolution. As a result, we need only be confident that we are 
+sufficiently resolving the major mixing regions on the white dwarf surfaces. 
+If we find that this mixing occurs near the grid resolution scale, this will imply 
+that we need to ramp up the resolution in these regions using AMR. If this becomes 
+too expensive, we would need to be skeptical of any conclusions that could be drawn 
+about the effect of the mixing on the nuclear burning.
 

### 2015-02-26 papers/wdmerger_I/paper.tex
Change default initial model resolution to 25 km and make it a probin parameter. Add discussion of this choice to the paper.

@@
 iteration over the entire integration procedure to find the central
-density needed to yield the desired total mass.  We note that the grid
-spacing $\Delta x$ is chosen such that it is equal to the zone size on
-the finest level of our problem, which is known in advance since we
-specify the domain size and the maximum number of possible levels of
-refinement.
+density needed to yield the desired total mass.  The grid spacing is
+$\Delta x = 25\ \text{km}$. We chose this value because no simulation
+we perform is likely to exceed this grid resolution inside the stars 
+themselves; for our normal domain size (see below), this corresponds to 
+three jumps in refinement by a factor of four. We find that for low 
+resolution runs, this is a better choice than selecting the 1D grid 
+spacing to be equal to the 3D grid spacing.
 
@@
 interpolating these to the cell centers, and then calling the equation
-of state to initialize the remaining terms.  The interpolation process
-divides each zone into $n_{\text{sub}}$ sub-zones of equal volume for
+of state to initialize the remaining terms. It is possible to interpolate
+instead by using pressure instead of temperature, as pressure is more 
+closely related to hydrostatic balance, but the EOS we use is so 
+insensitive to temperature that this mapping can result in large 
+deviations from the isothermal assumption we started with.  The 
+interpolation process divides each zone into $n_{\text{sub}}$ 
+sub-zones of equal volume for
 the purpose of sampling the 1D model, and the sub-zones are added
 together to obtain the full zone's characteristics. This
-sub-grid-scale interpolation is useful near the edge of the star,
-where the density falls off rapidly with radius.
+sub-grid-scale interpolation is useful especially near the edge of the star,
+where the density falls off rapidly with radius. Typically we take 
+$n_{\text{sub}} = 4$.
 
@@
 frame that rotates with the WDs, and if we are in the inertial frame
-the velocity is set equal to the rigid rotation rate corresponding to
-the specified period $T$.
+the velocity in every zone is set equal to the rigid rotation rate 
+corresponding to the distance of that zone from the rotation axis, given
+the specified period $T$. Thus the inertial frame and rotating frame 
+simulations are starting off with the same initial conditions: two white 
+dwarfs locked in synchronous rotation. This is the simplest assumption to 
+make, but in the future we may explore relaxing this requirement.
 
@@
 have commented on the importance of equilibrium initial conditions in
-determining the evolution of the system. In particular, we plan to
-compare the self-consistent field method (which finds an equilibrium
-solution \textit{a priori}) and the method of \cite{rosswog:2004},
-which starts with our method and then attempts to guide the evolution
-toward the equilibrium state using suitably defined damping
-terms. However, these issues are not considered for the present study,
-as we are not yet attempting to study the merger process.
-
-As a consequence of starting in a non-equilibrium setting, there are 
+determining the evolution of the system. As a consequence of starting 
+in a non-equilibrium setting, there are 
 large density and pressure gradients near the white dwarf surfaces
@@
 whose density is greater than the cutoff. We then compute $r_{\text{eff}} = (3V/4\pi)^{1/3}.$
+
+The computational domain has a total size of $1.024 \times 10^{10}\ \text{cm}$ in each 
+spatial dimension, and is centered at the origin. Our coarse grid has $256^3$ zones,
+corresponding to a spatial resolution of 400 km. For the present study, we choose a simple 
+refinement strategy: all zones within the Roche radius of each star are tagged for maximal 
+refinement. We also ensure that the outer part of the domain is never tagged for refinement.
+In future work we will add criteria that tag for refinement the gas between the stars, which is
+expected to feature nuclear burning.
 

### 2015-02-26 source/probdata.f90

@@ -69,6 +69,10 @@ module probdata_module
   ! Are we doing a single star simulation?
   logical :: single_star
 
+  ! Resolution of 1D initial model
+  double precision :: initial_model_dx
+  integer          :: initial_model_npts
+
   ! Tagging criteria
   double precision :: maxTaggingRadius
 
@@ -150,7 +154,9 @@ contains
          stellar_temp, stellar_C12, stellar_O16, &
          star_axis, &
          maxTaggingRadius, &
-         bulk_velx, bulk_vely, bulk_velz
+         bulk_velx, bulk_vely, bulk_velz, &
+         initial_model_dx, &
+         initial_model_npts
 
     maxTaggingRadius = 0.75d0
 
@@ -180,6 +186,17 @@ contains
     bulk_velx = ZERO
     bulk_vely = ZERO
     bulk_velz = ZERO
+
+    ! For the grid spacing for our model, we'll use 
+    ! 25 km. No simulation we do is likely to have a resolution
+    ! higher than that inside the stars (it represents
+    ! three jumps by a factor of four compared to our 
+    ! normal coarse grid resolution). By using 1024 
+    ! grid points, the size of the 1D domain will be 2.56e9 cm,
+    ! which is larger than any reasonable mass white dwarf.
+
+    initial_model_dx = 2.5d6
+    initial_model_npts = 1024
 
     ! Read namelist to override the defaults
     untin = 9 
@@ -293,10 +310,6 @@ contains
     radius_P_initial = ZERO
     radius_S_initial = ZERO
 
-    npts_model = 1024
-
-    dx = 1.0d6
-
     com_P = center
     com_S = center
 
@@ -319,8 +332,9 @@ contains
 
     ! Generate primary and secondary WD
 
-    call init_1d(model_P_r, model_P_state, npts_model, dx, radius_P_initial, mass_P_initial, &
-                 central_density_P, stellar_temp, stellar_comp, ambient_state)
+    call init_1d(model_P_r, model_P_state, initial_model_npts, initial_model_dx, &
+                 radius_P_initial, mass_P_initial, central_density_P, &
+                 stellar_temp, stellar_comp, ambient_state)
 
     if (ioproc == 1 .and. init == 1) then
         print *, "Generated initial model for primary WD of mass", mass_P_initial, &
@@ -331,8 +345,9 @@ contains
 
     if (.not. single_star) then
 
-       call init_1d(model_S_r, model_S_state, npts_model, dx, radius_S_initial, mass_S_initial, &
[…]

### 2015-02-27 source/probdata.f90
Correction to previous commit: three levels of refinement == 6.25 km.

@@ -190,22 +190,22 @@ contains
     bulk_velz = ZERO
 
     ! For the grid spacing for our model, we'll use 
-    ! 25 km. No simulation we do is likely to have a resolution
+    ! 6.25 km. No simulation we do is likely to have a resolution
     ! higher than that inside the stars (it represents
     ! three jumps by a factor of four compared to our 
-    ! normal coarse grid resolution). By using 1024 
+    ! normal coarse grid resolution). By using 2048
     ! grid points, the size of the 1D domain will be 2.56e9 cm,
     ! which is larger than any reasonable mass white dwarf.
 
-    initial_model_dx = 2.5d6
-    initial_model_npts = 1024
+    initial_model_dx = 6.25d5
+    initial_model_npts = 4096
 
     ! initial_model_mass_tol is tolerance used for getting the total WD mass 
     ! equal to the desired mass. It can be reasonably small, since there
     ! will always be a central density value that can give the desired
     ! WD mass on the grid we use.
 
-    initial_model_mass_tol = 1.d-4
+    initial_model_mass_tol = 1.d-6
 
     ! hse_tol is the tolerance used when iterating over a zone to force
     ! it into HSE by adjusting the current density (and possibly

### 2015-03-04 papers/wdmerger_I/paper.tex
Correction to previous commit: three levels of refinement == 6.25 km.
First pass at a performance section

@@
 density needed to yield the desired total mass.  The grid spacing is
-$\Delta x = 25\ \text{km}$. We chose this value because no simulation
+$\Delta x = 6.25\ \text{km}$. We chose this value because no simulation
 we perform is likely to exceed this grid resolution inside the stars 
@@
 %==========================================================================
-\section{Performance}\label{sec:Performance}
-
-strong scaling plot
-
-space filling curve?
-
-OMP vs. MPI?
-
-breakdown by component?
+\section{Parallel Strategy and Performance}\label{sec:Performance}
+
+CASTRO is designed to be deployed on high-performance computing systems using 
+many thousands of processors simultaneously. It is worth briefly examining 
+our strategy for parallelizing the problem over many computational nodes 
+and our performance in situations similar to production science simulations. 
+This is especially true because some aspects of our approach to parallelism 
+have changed since the first CASTRO paper, and we have obtained improved 
+performance in certain settings.
+
+The BoxLib framework that CASTRO is based on splits each AMR level into a number 
+of boxes that collectively span the level. These boxes are distributed to processors 
+through MPI parallelism; each MPI task in general holds multiple boxes and 
+an update includes a loop over all the boxes an MPI task owns. The distribution 
+obeys a load-balancing algorithm that attempts to equalize the amount of work 
+done by each processor. BoxLib contains a number of strategies for distributing 
+work in this way, and by default uses a space-filling curve approach with a 
+Morton ordering. By experiment we have found that the most efficient load-balancing 
+strategy for our problem is actually a simple knapsack algorithm. In this approach, 
+the amount of work owned by a processor is proportional to the number of grid cells 
+associated with that processor, and the algorithm attempts to ensure that all 
+processors have a similar number of total grid cells. We demand an efficiency of 0.9,
+meaning that the average workload per processor should be no smaller than 90\% of the 
+maximum workload found on any processor. However, we find that in practice the 
+performance is largely insensitive to this choice.
+
+The size and shape of grid boxes is an important consideration for efficiency. 
+Boxes that are very small suffer from a host of problems, including the larger 
+amount of communication required between hydrodynamics solves. Additionally, 
+the multigrid solver will be less efficient if the boxes are small because there 
+will be fewer available levels for coarsening and performing V-cycles. Conversely, 
+boxes that are too large mean that there isn't enough work to go around when we 
+have a large number of processors. Good performance is the result of a careful 
+balance between these two effects. On the lower end, we require that all boxes 
+be a multiple of 16 zones in each dimension; multigrid efficiency sharply decreases 
+if this factor is any lower. On the upper end, we select the maximum grid size 
+based on the number of processors we use and the total number of cells in the 
+simulation. This size will therefore in general vary on different AMR levels. 
+Generally we select a value in between 32 and 64 zones per dimension.
+
+We use OpenMP to accelerate the work associated with the boxes owned by each 
+MPI task. Originally CASTRO used OpenMP to accelerate individual loops in the 
+hydrodynamics routines, such as the piecewise-parabolic edge state reconstruction 
+and the conservative flux update. However, there is a significant amount of 
+overhead associated with generating a new OpenMP region at each of the many 
+different loops in a hydrodynamics algorithm. This makes such a strategy 
+sub-optimal for use on many-core processors and GPUs. We have recently switched 
+to a tiling approach, where an OpenMP region is generated at the start of 
+the hydrodynamics routine and the individual threads separately work on 
+different partitions of each box. This results in much less overhead for 
+the threading. In general we obtain more efficient simulations than 
+could be obtained using MPI only, because there are fewer boxes and thus 
+less communication for a given number of processor cores. We are currently 
+developing an approach to evaluating the equation of state and nuclear reactions 
+(and possibly the entire hydrodynamics algorithm) on GPUs, which will allow us
+to take advantage of the significant computational resources embedded in GPUs on 
+certain systems.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.4]{plots/weak_scaling}
+  \caption{CASTRO weak scaling test, performed on the Blue Waters machine at 
+    NCSA. Each processor had a fixed amount of work, and we increased the 
+    number of simulation zones in concert with the number of processors. The 
+    solid curve represents perfect weak scaling, while the blue circles show 
+    CASTRO's performance at each processor count. The vertical axis measures 
+    the median time per timestep, normalized to this value for the smallest 
+    processor count.\label{fig:weak_scaling}}
+\end{figure}
+
+To examine the parallel performance of CASTRO, we performed both strong scaling 
+and weak scaling tests on the Blue Waters machine at the National Center for 
+Supercomputing Applications. For the weak scaling test, whose results are 
+shown in Figure \ref{fig:weak_scaling}, we ran a uniform grid binary white 
+dwarf simulation for resolutions of $128^3$ zones through $2048^3$ zones. The 
+number of processors was scaled with the number of zones so that each processor 
+had the same amount of work; the smallest test used 8 processors and the largest 
+used 32,768 (note that the number of processor cores on a Blue Waters node is twice 
+the number of floating point units on that node). The test was run for 10 timesteps,
+with each timestep including two Poisson solves and a hydrodynamics update (though 
+for a uniform grid calculation we generally do not need to perform any multigrid 
+iterations for the first Poisson solve in a timestep, since the density distribution 
+has not changed since the end of the last timestep). We disabled plotfile and 
+checkpoint writing, as well as calculation of diagnostic information (the latter 
+can contribute to a significant fraction of the run time at large processor 
+counts if computed every timestep). We computed the median wall time required per 
+time step for each simulation, and then normalized this to the median time per 
+timestep for the smallest simulation. We find excellent weak scaling through 
+4,096 processors. At the largest run, the simulation time required is 
+more than 1.5 times the amount required for the smallest simulation. 
+This is due entirely to the increased cost of the multigrid Poisson solve 
+in each timestep and this cannot be mitigated except by improving 
+communication or computation efficiency in the multigrid solver. We observe 
+that this weak scaling behavior with Poisson gravity is a significant 
+improvement over the results presented in the first CASTRO paper.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.4]{plots/strong_scaling}
+  \caption{CASTRO strong scaling test performed on the Blue Waters machine at
+    NCSA. The vertical axis measures the median time per timestep, and the 
+    horizontal axis measures the number of processors in the simulation. Data 
+    points are normalized to the time per timestep for the smallest number 
+    of processors. The blue circles show the data for a simulation with two AMR
+    levels (one coarse and one fine), while the red circles show the data 
+    for a test with three AMR levels (one coarse and two fine). The fine levels 
+    increase the resolution only in the regions around the stars. For each case 
+    we draw a solid curve representing perfect strong scaling.\label{fig:strong_scaling}}
+\end{figure}
+
+The strong scaling test we perform uses a grid setup similar to what we 
+use for well-resolved binary simulations. With a single refined level, 
+we have approximately $5 \times 10^7$ zones spread over $\sim 1000$ grids.
+On a second refined level, there are $2 \times 10^9$ zones spread over 
+$\sim 16000$ grids. We run a scaling test for both cases, with the 
+highest processor count in each case chosen so that the number of 
+MPI tasks is similar to the number of grids. There are no relative gains 
+to be achieved from further parallelism. The results are found in 
+\ref{fig:strong_scaling}. We find excellent scaling for low numbers 
+of processors, though the long wall time per timestep required for the 
+smallest processor counts makes such a simulation infeasible in practice.
+The best tradeoff between reasonable wall time and parallel efficiency 
+seems to be when there are approximately 2--4 grids per processor. The 
+scaling behavior worsens at the highest processor counts, but this is 
+an expected consequence of processors becoming work-starved. In general 
+we find very good strong scaling behavior in the regime we are presently 
+interested in, simulating the early phases of a simulation at moderate 
+resolution. The strong scaling behavior is acceptable, though not perfect, 
+at very large processor counts when self-gravity is considered.
 

### 2015-03-04 analysis/wdmerger.py
Guard against the possibility that we didn't have gravity verbosity on.

@@ -207,10 +207,10 @@ def timing(output_filename):
 
     grav_per_timestep = (nlevs - 1) * 2 * ref_ratio + 4
 
-    for n in range(len(coarseSteps)):
-
-        for i in range(grav_per_timestep):
-            coarseSteps[n] -= float(grav_time[n+i])
+    if (len(grav_time) > 0):
+        for n in range(len(coarseSteps)):
+            for i in range(grav_per_timestep):
+                coarseSteps[n] -= float(grav_time[n+i])
 
     med_timestep_no_grav = np.median(coarseSteps)
 

### 2015-03-11 papers/wdmerger_I/paper.tex
use IOP + set the paper size to letter so pdflatex makes things right
Insert description of the coupled rotation update. This used to be in the appendix and the content is now in the main rotation section. The method is implemented for rot_source_type == 4 in CASTRO.

@@
 in the edge state prediction in a way that is analogous to the gravity source.
-We evaluate all quantities at cell centers and use the same predictor-corrector 
-approach that we use for the gravity source terms to the momentum equations.
+We evaluate all quantities at cell centers. We use the same predictor-corrector 
+approach that we use for the gravity source terms to the momentum equations. However, a slight 
+difference is that the Coriolis force for each velocity component is coupled to other velocity 
+components. If the rotation is about the $z$-axis, then the discrete update to 
+$v_x^{n+1}$ depends on the value of $v_y^{n+1}$, and vice versa. If we fix the value of 
+the time-level $n+1$ quantities after coming out of the hydrodynamics update, there 
+would be a slight inconsistency between the $x$ and $y$ components of the velocity.
+
+We propose a more accurate coupling that directly solves this implicit system of coupled 
+equations. We denote by $(\widetilde{\rho \mathbf{u}})$ the value of the momentum after 
+updating it with the centrifugal force, and the time-level $n$ Coriolis force. The remaining 
+update for the time-level $n+1$ Coriolis force then appears as:
+\begin{equation}
+  (\rho \mathbf{u})^{n+1} = (\widetilde{\rho\mathbf{u}}) + \frac{\Delta t}{2} \left(-2\, {\bm\omega} \times (\rho\mathbf{u})^{n+1}\right)
+\end{equation}
+To proceed further, we assume that the rotation is about the $z$ axis with frequency $\omega$. 
+Then there is no update to the $z$-momentum, and the other equations are:
+\begin{align}
+  (\rho u)^{n+1} &= (\widetilde{\rho u}) + \omega \Delta t (\rho v)^{n+1}) \\
+  (\rho v)^{n+1} &= (\widetilde{\rho v}) - \omega \Delta t (\rho u)^{n+1})
+\end{align}
+We can directly solve this coupled system:
+\begin{align}
+  (\rho u)^{n+1} &= \frac{ (\widetilde{\rho u}) + \omega \Delta t (\widetilde{\rho v})}{1 + \omega^2 \Delta t^2} \\
+  (\rho v)^{n+1} &= \frac{ (\widetilde{\rho v}) - \omega \Delta t (\widetilde{\rho u})}{1 + \omega^2 \Delta t^2}
+\end{align}
+We use this form of the momentum update in CASTRO. This improvement is small
+but increases the accuracy of our rotating white dwarf systems over long time-scales.
 
 The update to the energy equation can be determined by taking the dot product of the velocity
-with the momentum source terms. The Coriolis term vanishes identically, as it must
-because the Coriolis term does no work on the fluid. The update from the centrifugal force becomes
+with the momentum source terms. The Coriolis term vanishes identically, and so
+ the Coriolis term does no work on the fluid. The update from the centrifugal force becomes
 \begin{equation}

### 2015-03-22 documentation/userguide.tex

@@
 and in each directory the inputs file will have the corresponding \texttt{castro.rotational\_period}.
-Each job will stop after one rotational period has been completed. When all have been completed, you should 
-type \texttt{bash run.sh} once more; this will move all output files (checkpoints, plot files, diagnostics)
+Each job will stop after one rotational period has been completed. At the end, the script will automatically move all 
+output files (checkpoints, plot files, diagnostics)
 to the \texttt{output/} subdirectory of that problem location, and on supported HPC systems will also copy these 

### 2015-03-22 papers/wdmerger_I/paper.tex

@@
 %\documentclass[12pt,preprint]{aastex}
-\documentclass{emulateapj}
+\documentclass[iop]{emulateapj}
 % for \sout
@@
 \normalem
+
+% these lines seem necessary for pdflatex to get the paper size right
+\pdfpagewidth 8.5in
+\pdfpageheight 11.0in
 
@@
 corresponding to a spatial resolution of 400 km. For the present study, we choose a simple 
-refinement strategy: all zones within the Roche radius of each star are tagged for maximal 
-refinement. We also ensure that the outer part of the domain is never tagged for refinement.
-In future work we will add criteria that tag for refinement the gas between the stars, which is
-expected to feature nuclear burning.
+refinement strategy: all zones within twice the Roche radius of each star are tagged for maximal 
+refinement. The extra buffer from doubling the size of this region ensures that the sharp density
+gradients near the edge of the star are within the zone of refinement. We also ensure that the 
+outer part of the domain is never tagged for refinement. In future work we will add criteria 
+that tag for refinement the gas between the stars, which is expected to feature nuclear burning.
 

### 2015-03-23 job_scripts/run_utils.sh
Make our strategy for archiving inputs and probin files the same as the other archived files.

@@ -492,11 +492,14 @@ function archive_all {
 
   for file in $inputs_list
   do
-      f=$(basename file)
-      if ([ ! -e $directory/output/$f ] || [ $directory/output/$f -ot $directory/$f ]); then	  
-	  cp $file $directory/output/
-	  archivelist=$archivelist" "$f
+      f=$(basename $file)
+      if [ -e $directory/output/$f ]; then
+	  if [ $directory/output/$f -nt $directory/$f ]; then	  
+	      continue
+	  fi
       fi
+      cp $file $directory/output/
+      archivelist=$archivelist" "$f
   done
 
   probin_list=$(find $directory -maxdepth 1 -name "*probin*")
@@ -504,10 +507,13 @@ function archive_all {
   for file in $probin_list
   do
       f=$(basename $file)
-      if ([ ! -e $directory/output/$f ] || [ $directory/output/$f -ot $directory/$f ]); then
-	  cp $file $directory/output/
-	  archivelist=$archivelist" "$f
-      fi
+      if [ -e $directory/output/$f ]; then
+	  if [ $directory/output/$f -nt $directory/$f ]; then
+	      continue
+	  fi
+     fi
+     cp $file $directory/output/
+     archivelist=$archivelist" "$f
   done
 
   # If there is nothing to archive,

### 2015-03-23 papers/wdmerger_I/paper.tex

@@
 
-The history of double degenerate theory can broadly be described as a
-series of three major paradigm shifts. The first attempts to model the
-results of the merger process came in the
+The first attempts to model the results of the merger process came in the
 1980s. \cite{nomotoiben:1985} demonstrated that off-center carbon
@@
 
-The first three-dimensional simulations of double degenerate systems
-were performed by \citet{benz:1990}, who used the smoothed particle
-hydrodynamics (SPH) method to simulate the merger process. They found
-that if the lower-mass star (generally called the ``secondary'') was
+Three-dimensional simulations of merging double degenerate systems were 
+first performed by \citet{benz:1990}, who used the smoothed particle
+hydrodynamics (SPH) method to simulate the merger process. This was 
+followed later by a number of authors 
+\citep{rasio_shapiro:1995,segretain:1997,guerrero:2004,yoon:2007,loren-aguilar:2009,raskin:2012}.
+The main finding of these early 3D SPH simulations was that if the 
+lower-mass star (generally called the ``secondary'') was
 close enough to the more massive star (the ``primary'') to begin mass
@@
 centrifugally-supported accretion disk surrounding the core and
-envelope. They noted that the envelope temperature was hot enough to
-generate carbon fusion but neglected nuclear reactions in their
-simulation. Later simulations of similar setups
-\citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009,raskin:2012}
-corroborated this result. This validation of the prediction of
-\cite{tutukov_yungelson:1979} did not instill confidence in the
-community. \cite{mochkovitch_livio:1990} and \cite{livio:2000}
-observed that turbulent viscosity would be sufficiently large for
-angular momentum to be removed from the disk at a rate high enough to
-generate the troublesome accretion timescales. Based on all of this
+envelope. Carbon fusion might commence in the disk, but not at a 
+high enough rate to generate a nuclear detonation. \cite{mochkovitch_livio:1990} 
+and \cite{livio:2000}  also observed that turbulent viscosity in this disk 
+would be sufficiently large for angular momentum to be removed from the 
+disk at a rate high enough to generate the troublesome accretion 
+timescales discussed by \cite{tutukov_yungelson:1979}. Based on this
 evidence, the review of \cite{hillebrandtniemeyer2000} argued that the
@@
 a more detailed treatment of the viscous transport in the outer
-regions of the remnant and found that while the centrifugally
-supported envelope would be converted into hot envelope material on a
-viscous timescale, their simulations still led to off-center carbon
+regions of the remnant and found that viscous dissipation in the centrifugally
+supported envelope would substantially heat up the envelope on a  
+viscous timescale, but their simulations still led to off-center carbon
 burning. \cite{vankerkwijk:2010} argued that equal-mass mergers would
@@
 
-The most recent shift in perspective on this problem started with two
-series of papers appearing at roughly the same time. One began with
-\cite{pakmor:2010} and \cite{pakmor:2011}. This group used the SPH
-method to study the merger of equal-mass ($0.9\ M_\odot$)
-carbon-oxygen white dwarfs and found that a hotspot was generated near
-the surface of the primary white dwarf. They argued that this region
-had a temperature and density sufficient to trigger a thermonuclear
-detonation. They propagated this detonation throughout the system and
-found that it would observationally appear as a subluminous Type Ia
-supernova. This was the first time a DD simulation successfully
-reproduced at least some characteristics of a SN
+A recent shift in perspective on this problem started around 2010.
+\cite{pakmor:2010} used the SPH method to study the merger of 
+equal-mass ($0.9\ M_\odot$) carbon-oxygen white dwarfs and found 
+that a hotspot was generated near the surface of the primary 
+white dwarf. They argued that this region had a temperature 
+and density sufficient to trigger a thermonuclear
+detonation. They inserted a detonation which propagated throughout 
+the system. They found that the result would observationally 
+appear as a subluminous Type Ia supernova. This was the first time 
+a DD simulation successfully reproduced at least some characteristics of a SN
 Ia. \cite{pakmor:2011} tried a few different mass combinations and
@@
 Around the same time, however, \cite{guillochon:2010} and
-\cite{dan:2011} came to significantly different conclusions. They
-pointed out that the previously mentioned simulations generally all
-shared a significant drawback, which was that their initial conditions
+\cite{dan:2011} pointed out that the previously mentioned simulations 
+generally shared a significant drawback, which was that their initial conditions
 were not carefully constructed. \cite{motl:2002}, \cite{dsouza:2006},
@@
 white dwarf systems after constructing exact equilibrium initial
-conditions. In contrast to the other studies, they found that paying
-close attention to the accuracy of the initial conditions of the
-simulation has important effects. In particular, earlier work placed
-the stars too close together and ignored the effects of tidal forces
-on changing the shape of the secondary, leading to the merger
+conditions. Earlier work placed the stars too close together 
+and ignored the effects of tidal forces that change the shape of the 
+secondary, leading to the merger
 happening artificially too quickly. When the initial conditions are
@@
 
-The very latest developments give some areas of agreement and some of
-remaining uncertainty. \cite{pakmor:2012} performed a merger scenario
+The latest developments have resulted in some possible areas of convergence.
+\cite{pakmor:2012} performed a merger scenario
 with a $1.1\ M_\odot$ and $0.9\ M_\odot$ setup, with better treatment
-of the initial conditions, and also found that the merger process
-happened over more than ten orbits. Nevertheless, they still found
+of the initial conditions, and indeed found that the merger process
+happened over more than ten orbits. Nevertheless, they still determined
 that a carbon-oxygen detonation would occur, in line with their
 earlier results. \cite{moll:2014} was also able to find a detonation
-in a similarly massive system. \cite{dan:2012} and \cite{dan:2014}
-performed a large sweep of the parameter space for merger pairs and
+in a similarly massive system. Very recently, \cite{kashyap:2015} 
+found that a detonation occurred in a similarly massive system. Notably,
+the detonation occurred self-consistently and did not need to be inserted 
+by hand. \cite{dan:2012} and \cite{dan:2014} performed a large sweep 
+of the parameter space for merger pairs and
 found that pure carbon-oxygen systems would generally not lead to
 detonations and violent mergers except for the most massive
-systems. They did find that for systems containing helium, many of
-them would detonate and potentially lead to SNe Ia, either through the
+systems. They did find that for systems containing helium, many
+would detonate and potentially lead to SNe Ia, either through the
 aforementioned instabilities in the accretion stream, or during the
@@
 white dwarf, and found that this robustly led to a detonation of the
-white dwarf. Thus at least a preliminary agreement may be that systems
-containing helium could robustly lead to events resembling SNe Ia, as
-well as very massive carbon-oxygen binaries.
+white dwarf. For now there is preliminary support for the hypothesis
+that systems with helium shells (or helium WDs), and very massive carbon-oxygen binaries,
+could robustly lead to events resembling SNe Ia.
 
@@
 across simulation codes and algorithms is important for gauging
-confidence in this result. With only a couple of exceptions, most of
-the existing results that study the viability of the DD progenitor as
-an explanation for Type Ia supernovae (that is, including a
-temperature-dependent equation of state and nuclear reactions) have
+confidence in this result. Most of the existing results that study 
+the viability of double degenerate systems as progenitors for
+Type Ia supernovae (that is, including a realistic 
+equation of state and nuclear reactions) have
 used the SPH method. SPH codes have a number of features which do aid
@@
 none of these improvements have yet been incorporated into an SPH
-model of a WD merger. Another reason to be skeptical is that the
-detonation itself was inserted manually in \cite{pakmor:2010} based on
-small-scale detonation simulations. A self-consistent evolution of a
-thermonuclear detonation, if it existed, would significantly build
-confidence in the progenitor model.
+model of a WD merger. Another reason for caution is that other than the
+most recent results of \cite{kashyap:2015}, no simulation has self-consistently
+resulted in a thermonuclear detonation. Reproducibility of the detonation 
+through direct numerical simulation is critical for building 
+confidence in this progenitor model.
 

### 2015-03-27 documentation/userguide.tex
The sponge parameters can now be set via the probin file.

@@
   \item \texttt{maxTaggingRadius}: fractional size of the domain outside of which we won't perform any refinement (real)
+  \item \texttt{sponge\_dist}: we can apply a sponge that damps the velocities starting at this fraction multiplied by the 
+    distance from the center to the domain boundary (real)
+  \item \texttt{sponge\_width}: smoothing width of the sponge; the sponge is fully applied at {\tt sponge\_dist} + {\tt sponge\_width},
+    in the same coordinates (real)
+  \item \texttt{sponge\_timescale}: the sponge decreases the velocity by a factor $1 / (1 + (\Delta t / t_s)$, where 
+    $t_s$ is the {\tt sponge\_timescale} in seconds (real)    
 \end{itemize}

### 2015-03-27 source/probdata.f90

@@ -106,6 +106,11 @@ module probdata_module
   double precision :: c(0:1,0:1,0:1,3)  ! Interpolation coefficients for points
   integer          :: rloc(3,3)         ! Indices of zones nearby to these points
 
+  ! Sponge
+  double precision :: sponge_dist      ! Fraction of the domain size interior to which we should not sponge
+  double precision :: sponge_width     ! Interval over which we should smooth out the sponging (in units of the domain size)
+  double precision :: sponge_timescale ! Typical timescale to use in determining sponging strength
+
 contains
 
   ! This routine calls all of the other subroutines at the beginning
@@ -186,7 +191,8 @@ contains
          initial_model_dx, &
          initial_model_npts, &
          initial_model_mass_tol, &
-         initial_model_hse_tol
+         initial_model_hse_tol, &
+         sponge_dist, sponge_width, sponge_timescale
 
     maxTaggingRadius = 0.75d0
 
@@ -256,6 +262,10 @@ contains
     ! temperature).  hse_tol should be very small (~ 1.e-10).
 
     initial_model_hse_tol = 1.d-10
+
+    sponge_dist      = 0.75
+    sponge_width     = 0.1
+    sponge_timescale = 0.01
 
     ! Read namelist to override the defaults
     untin = 9 

### 2015-03-27 source/sponge_3d.f90

@@ -11,6 +11,7 @@ contains
     use bl_constants_module, only: M_PI, HALF, ZERO, ONE
     use meth_params_module , only: NVAR, URHO, UMX, UMY, UMZ, UEDEN
     use prob_params_module, only: problo, probhi, center
+    use probdata_module, only: sponge_dist, sponge_width, sponge_timescale
 
     implicit none
     integer          :: lo(3),hi(3),domlo(3),domhi(3)
@@ -21,17 +22,14 @@ contains
     double precision :: time, dt
     double precision :: dx, dy, dz
 
-    double precision :: xx, yy, zz
-    double precision :: xctr, yctr, zctr
-
-    double precision :: ke_old, ke_new, sponge_mult, delta_sponge, &
-                        radius, r_sponge, fac
-
+    double precision :: xx, yy, zz, radius
+    double precision :: ke_old, ke_new
+    double precision :: sponge_radius, sponge_delta, sponge_mult, fac, alpha
     integer :: i,j,k
 
-
-    r_sponge = 0.25*(probhi(1)-problo(1))
-    delta_sponge = 0.025*(probhi(1)-problo(1))
+    sponge_radius = sponge_dist * max( maxval(abs(problo-center)), maxval(abs(probhi-center)) )
+    sponge_delta  = sponge_width * max( maxval(abs(problo-center)), maxval(abs(probhi-center)) )
+    alpha = dt / sponge_timescale
 
     do k = lo(3),hi(3)
        zz = problo(3) + dble(k + HALF)*dz
@@ -42,33 +40,32 @@ contains
           do i = lo(1),hi(1)
              xx = problo(1) + dble(i + HALF)*dx
 
+             ke_old = HALF * ( uout(i,j,k,UMX)**2 + &
+                               uout(i,j,k,UMY)**2 + &
+                               uout(i,j,k,UMZ)**2 ) / uout(i,j,k,URHO)
+
              radius = sqrt((xx-center(1))**2 + (yy-center(2))**2 + (zz-center(3))**2)
 
              sponge_mult = ONE
 
-             if (radius > r_sponge) then
-                if (radius < r_sponge+delta_sponge) then
-                   fac = HALF*(ONE - cos(M_PI*(radius-r_sponge)/delta_sponge))
+             if (radius >= sponge_radius) then
[…]

### 2015-04-03 papers/wdmerger_I/paper.tex
Add discussion to the paper of the sponge and the hydro boundary conditions.

@@
 
+For ghost cells at the physical edge of the domain, we always use quantities 
+corresponding to the initial ambient material, to prevent material from raining down
+on to the grid in response to the disequilibrium outside the stars. This is an artificial 
+choice, but the dynamics outside of the stellar orbit is arbitrary anyway, due to the 
+lack of knowledge about the density and velocity field inside this gas. In addition, we employ 
+a ``sponge'' similar to that described by \citet{maestro3}. After the hydrodynamics update (but before 
+the gravity and rotation correction terms), we apply a damping force to the momentum 
+equation as follows:
+\begin{equation}
+  (\rho \mathbf{u})^{n+1} \to \frac{(\rho \mathbf{u}^{n+1})}{1 + (\Delta t / \Delta t_S) f_S},
+\end{equation} 
+where $\Delta t_S$ is a timescale for the sponge to operate on, and $f_S$ is the damping factor.
+We choose it so that that the sponge is non-operational inside a radius $r_S$ from the origin, and 
+fully applied at a radius $r_S^\prime \equiv r_S + \Delta r_S$. We then smooth the sponge 
+out between $r_S$ and $r_S^\prime$:
+\begin{equation}
+  f_{S} = \begin{dcases} 0 & r < r_S \\ \frac{1}{2}\left(1 - \text{cos}\left[\pi\left(\dfrac{r - r_S}{\Delta r_S}\right)\right]\right) & r_S \leq r < r_S^\prime \\ 1 & r \geq r_S^\prime. \end{dcases}
+\end{equation}
+For the simulations in this paper we set $r_S$ to be 75\% of the distance from the origin to the domain boundaries,
+and $\Delta r_S$ so that the sponge extends another 10\% of that distance. We set 
+$\Delta t_S = 0.01$ s, which is of the same order as the as the CFL timestep 
+for typical problem setups. We find that this sponge is necessary especially in the 
+rotating reference frame, as standing instabilities can create very large velocities 
+in the ambient fluid that drag down the global timestep by up to an order of magnitude.
+However, this provides another reason to distrust the dynamics of the low-density fluid
+outside of the stellar orbits.
+
 The software used to generate the test problems in this paper

### 2015-04-03 analysis/wdmerger.py
Add a python function that returns the value of an input variable from an inputs file.

@@ -52,7 +52,48 @@ def get_inputs_filename(dir):
         inputs_list.sort(key=lambda s: int(s[7:]))
 
         return inputs_list[-1]
-        
+ 
+       
+
+# Get a CASTRO variable value from an inputs file.
+
+def get_inputs_var(inputs, var):
+
+    # Read in all the inputs file lines and search for the one
+    # that starts with the desired variable name.
+
+    inputs_file = open(inputs, 'r')
+    lines = inputs_file.readlines()
+
+    lines = filter(lambda s: s.split() != [], lines)
+    line = filter(lambda s: s.split()[0] == var, lines)
+
+    # The variable is the last item in a line before the comment.
+    # This should work correctly even if there is no comment.
+
+    var = (line[0].split('=')[1]).split('#')[0]
+    var = var.strip()
+
+    # Now, convert it into a list if it has multiple entries.
+
+    if (var.split() != []):
+        var = var.split()
+
+    # Convert this to a floating point array, if possible.
+    # If this fails, we'll just leave it as a string.
+
+    try:
+        var = np.array(var,dtype='float')    
+
+        # Now convert this to an integer array, if possible.
+
+        if (var[0].is_integer()):
+            var = np.array(var,dtype='int')
+    except:
+        pass
+
+    return var
+
 
 
 #

### 2015-04-03 papers/wdmerger_I/paper.tex
Add sections to the paper corresponding to the single-star hydrostatic equilibrium tests.

@@
 \cite{pakmor:2010} used the SPH method to study the merger of 
-equal-mass ($0.9\ M_\odot$) carbon-oxygen white dwarfs and found 
+equal-mass ($0.9\ \msolar$) carbon-oxygen white dwarfs and found 
 that a hotspot was generated near the surface of the primary 
@@
 \cite{pakmor:2012} performed a merger scenario
-with a $1.1\ M_\odot$ and $0.9\ M_\odot$ setup, with better treatment
+with a $1.1\ \msolar$ and $0.9\ \msolar$ setup, with better treatment
 of the initial conditions, and indeed found that the merger process
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/single_star_static_1e3_radius}
+  \caption{Time evolution of the effective radius of a $0.9 \msolar$ 
+    white dwarf, seeded onto the grid using a one-dimensional hydrostatic
+    model and evolved without further relaxation. The lines represent 
+    different number of zones per spatial dimension; when this number is 
+    greater than 256, it represents an effective resolution obtained 
+    using AMR levels that cover the star. The radius is determined 
+    using the volume of the grid that has a density greater than $10^3\ \text{g cm}^{-3}.$
+    \label{fig:single_star_static_radius}}
+\end{figure}
+
 This problem was studied in the first CASTRO paper, but is worth
@@
 deviation from equilibrium is affected by grid resolution.
+
+We loaded a single star of mass $0.9\ \msolar$ onto the grid at the origin, 
+and evolved it for 200 seconds. Our diagnostic of choice is the effective 
+radius of the star, determined by the volume of the grid that has a density 
+greater than $10^3\ \text{g cm}^{-3}$ (see Section \ref{sec:initial_models} 
+for details on this measure). This choice of density is intended to 
+mark a reasonable outer edge to the star that is not immediately susceptible 
+to the numerical errors prevalent near the physical edge of the star. Figure 
+\ref{fig:single_star_static_radius} shows our results at various resolutions. 
+As expected, the star quickly approaches an equilibrium size that is different 
+(and in this case larger) than the one-dimensional model. However, the magnitude 
+of this change becomes smaller with resolution. The star is only approximately 
+in equilibrium Only when the coarse grid of $256^3$ zones has a level of 
+refinement that jumps by a factor of four. However, even then there is a slight 
+uptick in the size toward the event, implying that the numerical stability 
+is not guaranteed for arbitrarily long timescales. This result suggests that when 
+doing the equilibrium initial models that will former the basis of later calculations,
+we should carefully monitor the evolution of the stars before applying any 
+artificial damping to cause the merger, to ensure that the merger is due to 
+this applied force and not the intrinsic numerical instability of the stars.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/single_star_static_ke}
+  \caption{Time evolution of the kinetic energy of a $0.9 \msolar$ 
+    white dwarf. The lines have the same meaning as in Figure \ref{fig:single_star_static_radius}.
+    \label{fig:single_star_static_ke}}
+\end{figure}
 
@@
 velocity distribution on the computational domain are given by:
+
 \begin{align}
@@
 \end{align}
+
 Here $R(y)$ is a ramp function that describes the transition between
@@
 \begin{align}
-  \rho &= \begin{cases} \rho_1 - \rho_m e^{\frac{y-0.25}{\Delta y}} & 0.25 > y \geq 0 \\ 
-                        \rho_2 + \rho_m e^{\frac{0.25-y}{\Delta y}} & 0.5 > y \geq 0.25 \\
-                        \rho_2 + \rho_m e^{\frac{y-0.75}{\Delta y}} & 0.75 > y \geq 0.5 \\
-                        \rho_1 - \rho_m e^{\frac{0.75-y}{\Delta y}} & 1 > y \geq 0.75 \end{cases} \label{eq:kh_p3_rho}\\
-  v_x &= \begin{cases} v_1 - v_m e^{\frac{y-0.25}{\Delta y}} & 0.25 > y \geq 0 \\
-                       v_2 + v_m e^{\frac{0.25-y}{\Delta y}} & 0.5 > y \geq 0.25 \\
-                       v_2 + v_m e^{\frac{y-0.75}{\Delta y}} & 0.75 > y \geq 0.5 \\
-                       v_1 - v_m e^{\frac{0.75-y}{\Delta y}} & 1 > y \geq 0.75 \end{cases} \label{eq:kh_p3_vx} \\
+  \rho &= \begin{dcases} \rho_1 - \rho_m e^{(y-0.25)/\Delta y} & 0.25 > y \geq 0 \\ 
+                         \rho_2 + \rho_m e^{(0.25-y)/\Delta y} & 0.5 > y \geq 0.25 \\
+                         \rho_2 + \rho_m e^{(y-0.75)/\Delta y} & 0.75 > y \geq 0.5 \\
+                         \rho_1 - \rho_m e^{(0.75-y)/\Delta y} & 1 > y \geq 0.75 \end{dcases} \label{eq:kh_p3_rho}\\
+  v_x &= \begin{dcases} v_1 - v_m e^{(y-0.25)/\Delta y} & 0.25 > y \geq 0 \\
+                        v_2 + v_m e^{(0.25-y)/\Delta y} & 0.5 > y \geq 0.25 \\
+                        v_2 + v_m e^{(y-0.75)/\Delta y} & 0.75 > y \geq 0.5 \\
+                        v_1 - v_m e^{(0.75-y)/\Delta y} & 1 > y \geq 0.75 \end{dcases} \label{eq:kh_p3_vx} \\
   v_y &= w_0\, \text{sin}\left(4\pi x\right). \label{eq:kh_p3_vy}
@@
 
-\subsubsection{Moving Star}\label{sed:moving_star}
+\subsubsection{Moving Star}\label{sec:moving_star}
+
+To see this effect in action for a stellar simulation, we repeated the test of Section 
+\ref{sec:HSE} with a bulk velocity on the grid. We chose a velocity of $2.56 \times 10^{8}\ \text{cm s}^{-1}$.
+For context, this is comparable to the orbital velocities of the stars in Section \ref{sec:kepler}.
+This test was inspired by \citet{tasker:2008}, who considered a moving galaxy cluster 
+and who obtained a long timescale evolution by using periodic boundary conditions, 
+so that the cluster would cross the domain multiple times throughout the evolution.
+We believe that periodic boundary conditions are unrealistic for our type of simulation,
+so we prefer to do one continuous simulation where the star does not cross the boundaries.
+Since our normal grid was not large enough to allow the motion to continue for very long, we expanded 
+the domain size by a factor of four, and then included an extra refined level around the star to 
+keep the effective resolution the same. We started the star off in the lower left corner of the domain, and 
+pointed its velocity towards the upper right corner. This allowed us to evolve the star for the 
+same length of time as for the original test. We note that getting the gravity boundary conditions right 
+required us to move the origin of the problem at the bulk velocity, so that the multipole moments 
+were computed with respect to the stellar center.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/single_star_compare_1e3_radius}
+  \caption{A variation on Figure \ref{fig:single_star_static_radius} where
+    we now compare the ``static'' case to ``motion'' simulations where the 
+    star moves across the grid at a fixed linear speed. The lines represent 
+    the effective number of zones per dimension inside the stellar material;
+    due to the expanded size of the grid in the ``motion'' case, the 
+    physical resolution is the same in each column in the legend.
+    \label{fig:single_star_compare_radius}}
+\end{figure}
+
+In Figure \ref{fig:single_star_compare_radius}, we take the results of Section 
+\ref{sec:HSE} (the ``static'' case), and plot on top of it the results of this 
+new simulation (the ``motion'' case). We see immediately that this bulk velocity
+causes the star to be much worse at maintaining hydrostatic equilibrium. Not only 
+is the absolute size of the star significantly larger (nearly a factor of two 
+at the lowest feasible resolution we consider), but also there is a clear upward 
+trend in the size that has not terminated at any resolution by the end of the simulation.
+This again emphasizes the results mentioned earlier, that we must be careful not 
+to trust any simulation with significant mass transfer if we are not confident that the 
+mass transfer is seeded in a controllable manner and free from numerical noise.
 

### 2015-04-03 tests/hydro/single_star_hse/analyze.py

@@ -9,9 +9,12 @@ import os
 ncell_arr = [ 256, 512, 1024 ]
 vel_arr   = [ 'motion', 'static' ]
 
-linestyles = ["-", "--","-.",":"]
+linestyles = ["-", "--","-"]
+markers = ["", "", "o"]
 
 nCols = 7
+
+# Effective radii of the star
 
 for vel in vel_arr:
 
@@ -22,7 +25,7 @@ for vel in vel_arr:
         eps_filename = "plots/single_star_" + str(vel) + "_1e" + str(c) + "_radius.eps"
 
         if (os.path.isfile(eps_filename)):
-            print "Plot with filename " + eps_filename + " already exists; skipping."
+            print "Plot with filename " + eps_filename + " already exists; skipping."            
             continue        
 
         print "Generating plot with filename " + eps_filename
@@ -34,9 +37,11 @@ for vel in vel_arr:
             # Account for the fact that we run at higher resolution when we're in the moving frame.
 
             if ( vel == 'motion' ):
-                n = n * 4
+                nc = n * 4
+            elif ( vel == 'static' ):
+                nc = n
 
-            diag_filename = "results/" + str(vel) + "/n" + str(n) + "/output/star_diag.out"
+            diag_filename = "results/" + str(vel) + "/n" + str(nc) + "/output/star_diag.out"
 
             timeCol = wdmerger.get_column('TIME', diag_filename)
 
@@ -49,13 +54,15 @@ for vel in vel_arr:
 
             radCol /= radCol[0]
 
-            plt.plot(timeCol, radCol, linestyle=linestyles[idx], linewidth=4.0, label=r'n = ' + str(n))
+            plt.plot(timeCol, radCol, linestyle=linestyles[idx], linewidth=4.0, marker = markers[idx], markevery=500, markersize=8.0, label=r'n = ' + str(nc))
 
             idx += 1
 
         plt.legend(loc='upper center',ncol=len(ncell_arr))
         plt.xlabel("Time (s)", fontsize=20)
         plt.ylabel("Radius / Initial Radius", fontsize=20)
+
+        plt.xkcd()
 
         if ( vel == 'motion' ):
             ylim_l = 0.95
@@ -71,6 +78,56 @@ for vel in vel_arr:
         wdmerger.insert_commits_into_eps(eps_filename, diag_filename, 'diag')
 
         plt.clf()
+
+
+
[…]

### 2015-04-05 papers/wdmerger_I/paper.tex
Change title

@@
 %==========================================================================
-\title{Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code 
-       Verification}
-
-\shorttitle{DWD Mergers. I. Methodology}
+\title{Double White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification}
+
+\shorttitle{WD Mergers I. Methodology}
 \shortauthors{Katz et al. (2015)}
@@
 relative to other methods. This is the first paper in a series designed to 
-describe simulations of these systems with the adaptive-mesh refinement 
-hydrodynamics code CASTRO. In this paper we describe our numerical 
-methodology. CASTRO solves the Euler equations of hydrodynamics 
+describe simulations of these systems using a hydrodynamics code with 
+using the adaptive-mesh refinement approach. In this paper we describe our numerical 
+methodology and discuss our implementation in the code CASTRO. CASTRO 
+solves the Euler equations of hydrodynamics 
 and the Poisson equation for self-gravity. Standard techniques for 

### 2015-04-05 papers/wdmerger_I/paper.tex
Add a conclusion

@@
 
+In this paper we have described the major components of a framework for 
+simulating mergers of white dwarfs. While there is much evidence for the 
+hypothesis that mergers (or collisions) of white dwarfs are significant 
+contributors to the rate of Type Ia supernovae and related astronomical 
+transients, the theoretical view on these systems is far from complete. 
+Studying these systems over the long timescales relevant to dynamical 
+mass transfer requires careful attention to the numerical methods used, 
+to ensure that numerical instabilities or other errors do not unduly 
+influence the system. Here we have described a number of common problems
+that may creep up, including violation of the conservation of energy, 
+a lack of hydrostatic equilibrium (at low resolution) of the stars 
+even when not acted on by external gravitational forces, and large 
+velocities that can be generated near the edges of stars due to the 
+numerically sharp gradients. Some of the issues are simply unresolvable
+at the resolutions achievable on modern supercomputing systems; 
+for example, it is very difficult to adequately resolve the stellar 
+surface of a white dwarf on a three-dimensional grid, and there is 
+justifiable room for suspicion regarding what happens there. But others
+are avoidable with care: energy non-conservation can be substantially 
+mollified by using a form of the gravitational work that is explicitly 
+conservative, and we have observed that this can be done for rotation 
+source terms too.
+
+We presented a set of numerical tests that show where we can 
+and cannot trust these techniques. We spent much time considering 
+the role of bulk motions on the grid and we conclude that there 
+are real issues with substantial bulk velocities on static grids, 
+that can diminish the quality of the resulting solutions to the 
+fluid equations. However, these effects diminish with increasing 
+resolution. We therefore make no explicit claims about the usefulness 
+of Langrangian versus Eulerian methods and instead simply observe 
+that whenever a simulation is performed, it is important to have 
+a measure of numerical accuracy and a sense of whether we are 
+witnessing properties that converge with increasing resolution.
+These problems do suggest that, where possible, we should seek to 
+minimize bulk motions on static grids. A comparison of orbit 
+simulations in both rotating and inertial reference frames 
+demonstrates that in practice this is not so simple, and that 
+a rotating reference frame has its own numerical issues for 
+the type of simulation we desire to perform here. Perhaps contrary 
+to expectations, the stars appear to be more stable in an inertial 
+frame where they are constantly in motion on the grid. It 
+is not easy to predict the correct behavior of such systems, 
+so the best recourse for assessing confidence in a model from a 
+verification standpoint is still to see whether the observed 
+behavior converges with resolution.
+
+Future work on this project will focus on how to build 
+reliable equilibrium initial models of the stars and 
+an examination of how the mass transfer episode depends on 
+these initial conditions, and then to enable the nuclear 
+reaction network and determine whether self-consistent 
+thermonuclear detonations are ignited and result in 
+events that appear similar to Type Ia supernovae. Other 
+areas ripe for future study include: the effect of radiation 
+on the merger process; the extent to which the result 
+depends on the initial composition of the stars (for example, 
+by including surface layers of helium and/or by using 
+white dwarf models generated by modern stellar evolution 
+codes); the gravitational wave signals generated by 
+white dwarf mergers; and, collisions of white dwarfs. All of 
+these are possible under the framework we have established.
+
 \acknowledgments

### 2015-05-05 papers/wdmerger_II/paper.tex
Add a directory for paper II

--- 2015-04-03 papers/wdmerger_I/paper.tex
@@
-%\documentclass[12pt,preprint]{aastex}
-\documentclass[iop]{emulateapj}
+%\documentclass[12pt,preprint]{../aastex}
+\documentclass[iop]{../emulateapj}
 % for \sout
@@
 \normalem
-
-% these lines seem necessary for pdflatex to get the paper size right
-\pdfpagewidth 8.5in
-\pdfpageheight 11.0in
 
@@
 
-% allows Greek symbols to be bold
-\usepackage{bm}
-
-% allows us to force the location of a figure
-\usepackage{float}
-
-% allows comment sections
-\usepackage{verbatim}
-
-\newcommand{\rhocutoff}{\rho_\mathrm{cutoff}}
-\newcommand{\rhoanelastic}{\rho_\mathrm{anelastic}}
-
 \newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-\newcommand{\Tcutoff}{T_\mathrm{cutoff}}
 
@@
 %==========================================================================
-\title{Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code 
-       Verification}
+\title{Double White Dwarf Mergers with CASTRO\\ II. Equilibrium Initial Conditions}
 
-\shorttitle{DWD Mergers. I. Methodology}
-\shortauthors{Katz et al. (2015)}
+\shorttitle{DWD Mergers. II. Equilibrium ICs}
+\shortauthors{Max}
 
-\author{Max P. Katz, Michael Zingale, Alan C. Calder, F. Douglas Swesty, Ann S. Almgren, Weiqun Zhang}
+\author{TBD}
 %==========================================================================
@@
 \begin{abstract}
-The Type Ia supernova progenitor problem is one of the most perplexing and 
-exciting problems in astrophysics, requiring detailed numerical modelling to 
-complement the observations of these explosions. One possibility that has 
-merited recent theoretical attention is the white dwarf merger scenario.
-This can naturally explain many of the observed characteristics of 
-Type Ia supernovae, as well as their distribution in time and space.
-However, to date there have been few fully self-consistent simulations 
-of binary white dwarf systems using mesh-based hydrodynamics, 
-relative to other methods. This is the first paper in a series designed to 
-describe simulations of these systems with the adaptive-mesh refinement 
-hydrodynamics code CASTRO. In this paper we describe our numerical 
-methodology. CASTRO solves the Euler equations of hydrodynamics 
-and the Poisson equation for self-gravity. Standard techniques for 
-coupling gravitation and rotation forces to the hydrodynamics do 
-not do an adequate job conserving the total energy of the system, 
-but recent advances in the literature have made advances on this 
-problem and we discuss our implementation here. We also run an 
-extensive set of test problems to verify that our software can sufficiently
-model a system where large amounts of mass are advected on the computational 
-domain over very long timescales. Future papers in this series will describe
-our treatment of the initial conditions of these systems, and will 
-examine the early phases of the merger to determine its viability
-for triggering a thermonuclear detonation.
+We describe a method for constructing equilibrium initial conditions for a binary star system,
+for a general equation of state. We extend this method to include the effects of adaptive
+mesh refinement in the algorithm.
 
@@
 
-Type Ia supernovae (SNe Ia) are currently some of the most exciting
-events to study in astrophysics. These bright, brief pulses of light
-in the distant universe have led to a number of important discoveries
-in recent years, including the discovery of the accelerated expansion
-of the universe \citep{perlmutter1999,riess1998}. However, their
-origin is shrouded in mystery. It has long been expected that these
-events arise from the thermonuclear explosions of white dwarfs
-\citep{hoyle_fowler:1960}, but the cause of these explosions is
-uncertain. In particular, it is not clear what process causes the
-temperatures in these white dwarfs to become hot enough for explosive
-burning of its constituent nuclei. The model favored initially by the
-community was the so-called single-degenerate (SD) model
-\citep{whelan_iben:1973}. Accretion of material from a companion star
-such as a red giant would cause the star to approach the Chandrasekhar
-mass, and in doing so the temperature and density in the center would
-become sufficient for thermonuclear fusion to proceed. However, in
-recent years many alternative progenitor models have been discussed. A
-leading candidate for explaining some or most of these explosions is
-the double-degenerate (DD) model, in which two white dwarfs merge and
-the merged object reaches the conditions necessary for a thermonuclear
-ignition \citep{ibentutukov:1984,webbink:1984}. Another is the double
-detonation scenario, where accretion of material onto a
-sub-Chandrasekhar white dwarf would lead to a detonation inside the
-accreted envelope, and this would send a compressional wave into the
-core of the star that would trigger a secondary detonation. A recent
-review of the progenitor models can be found in
-\citet{hillebrandt:2013}.
-
-There are several observational reasons why double-degenerate systems
-are a promising progenitor system for at least a substantial fraction
-of normal SNe Ia. No conclusive evidence exists for a surviving
-companion star of a SN Ia; this is naturally explained by the DD model
-because both WDs are likely to be destroyed in the merger
-process. Similarly, pre-explosion images of the SN Ia systems have
-never clearly turned up a companion star, and in some cases a large
-fraction of the parameter space for the nature of the companion star
-is excluded. Additionally, not enough progenitor systems are seen for
-the SD case to match the observed local SN Ia rate, whereas the number
-of white dwarf binaries may be sufficient to account for this
-rate. Finally, the DD model can naturally explain the fact that many
-SNe Ia are observed to occur at very long delay times after the stars
-were formed, since the progenitor systems only become active once both
-stars have evolved off the main sequence. A thorough review of the
-observational evidence about SNe Ia and further discussion of these
-ideas can be found in \cite{maoz:2014}.
-
-The first attempts to model the results of the merger process came in the
-1980s. \cite{nomotoiben:1985} demonstrated that off-center carbon
-ignition would occur in the more massive white dwarf as it accreted
-mass near the Eddington rate from the less massive white dwarf
-overflowing its Roche lobe. \cite{saionomoto:1985} tracked the
-evolution of the flame and found that it propagated quiescently into
-the center, converting the carbon-oxygen white dwarf into an
-oxygen-neon-magnesium white dwarf. This would then be followed by
-collapse into a neutron star -- a result with significantly different
-observational properties compared to a SN Ia. This scenario, termed
-accretion-induced collapse, would be avoided only if the accretion
-rate were well below the Eddington rate. \cite{tutukov_yungelson:1979}
-observed that this could happen if the mass loss from the secondary
-was higher than the Eddington rate and thus the accreted material
-formed an accretion disc, which might rain down on the primary more
-slowly. The main finding was that double degenerate systems would not
-obviously lead to Type Ia supernovae.
-
-Three-dimensional simulations of merging double degenerate systems were 
-first performed by \citet{benz:1990}, who used the smoothed particle
-hydrodynamics (SPH) method to simulate the merger process. This was 
-followed later by a number of authors 
-\citep{rasio_shapiro:1995,segretain:1997,guerrero:2004,yoon:2007,loren-aguilar:2009,raskin:2012}.
-The main finding of these early 3D SPH simulations was that if the 
-lower-mass star (generally called the ``secondary'') was
-close enough to the more massive star (the ``primary'') to begin mass
-transfer on a dynamical time scale, the secondary completely disrupted
-and formed a hot envelope around the primary, with a
-centrifugally-supported accretion disk surrounding the core and
-envelope. Carbon fusion might commence in the disk, but not at a 
-high enough rate to generate a nuclear detonation. \cite{mochkovitch_livio:1990} 
-and \cite{livio:2000}  also observed that turbulent viscosity in this disk 
-would be sufficiently large for angular momentum to be removed from the 
-disk at a rate high enough to generate the troublesome accretion 
-timescales discussed by \cite{tutukov_yungelson:1979}. Based on this
-evidence, the review of \cite{hillebrandtniemeyer2000} argued that the
-model was only viable if the accretion-induced collapse problem could
-be avoided. Later work by \cite{shen:2012} and \cite{schwab:2012} used
-a more detailed treatment of the viscous transport in the outer
-regions of the remnant and found that viscous dissipation in the centrifugally
-supported envelope would substantially heat up the envelope on a  
-viscous timescale, but their simulations still led to off-center carbon
-burning. \cite{vankerkwijk:2010} argued that equal-mass mergers would
-lead to the conditions necessary for carbon detonation in the center
-of the merged object, but \cite{shen:2012} questioned this for similar
-reasons related to how viscous transport would convert rotational
-motion into pressure support. \cite{zhu:2013} followed this with an
-expanded parameter space study and argued that many of their
-carbon-oxygen systems had the potential to detonate. The study of the
-long-term evolution of the remnants is thus still an open subject of
-research.
-
-A recent shift in perspective on this problem started around 2010.
-\cite{pakmor:2010} used the SPH method to study the merger of 
-equal-mass ($0.9\ \msolar$) carbon-oxygen white dwarfs and found 
-that a hotspot was generated near the surface of the primary 
-white dwarf. They argued that this region had a temperature 
-and density sufficient to trigger a thermonuclear
-detonation. They inserted a detonation which propagated throughout 
-the system. They found that the result would observationally 
-appear as a subluminous Type Ia supernova. This was the first time 
-a DD simulation successfully reproduced at least some characteristics of a SN
-Ia. \cite{pakmor:2011} tried a few different mass combinations and
-found empirically that this would hold as long as the secondary was at
-least 80\% as massive as the primary. These events, where the merger
-process resulted in the detonation of the system during the merger
-coalescence -- avoiding the much longer time-scale evolution -- were
-termed ``violent'' mergers.
-
-Around the same time, however, \cite{guillochon:2010} and
-\cite{dan:2011} pointed out that the previously mentioned simulations 
-generally shared a significant drawback, which was that their initial conditions
-were not carefully constructed. \cite{motl:2002}, \cite{dsouza:2006},
-and \cite{motl:2007} (the first three-dimensional grid-based
-simulations of mass transfer in binary white dwarf systems) pioneered
-the study of looking at the long-term dynamical evolution of binary
-white dwarf systems after constructing exact equilibrium initial
-conditions. Earlier work placed the stars too close together 
-and ignored the effects of tidal forces that change the shape of the 
-secondary, leading to the merger
-happening artificially too quickly. When the initial conditions are
-constructed in exact equilibrium, the system can be stable for tens of
-orbital periods, substantially changing the character of the mass
-transfer phase. However, one limitation of this series of studies is
-that the authors used a polytropic equation of state and thus could
-not consider nuclear reactions. \cite{guillochon:2010} and
-\cite{dan:2011} improved on this using a realistic equation of state,
-a nuclear reaction network, and a similar approach to the equilibrium
-initial conditions, and found substantial agreement with the idea that
-mass transfer occurs in a stable manner over tens of orbital
-periods. They also found that, assuming the material accreted onto the
-surface of the primary was primarily helium, explosive surface
-detonations would occur as a result of accretion stream instabilities
-during the mass transfer phase prior to the full merger. This could
-trigger a double-detonation explosion and thus perhaps a SN Ia.
-
-The latest developments have resulted in some possible areas of convergence.
-\cite{pakmor:2012} performed a merger scenario
-with a $1.1\ \msolar$ and $0.9\ \msolar$ setup, with better treatment
-of the initial conditions, and indeed found that the merger process
-happened over more than ten orbits. Nevertheless, they still determined
-that a carbon-oxygen detonation would occur, in line with their
-earlier results. \cite{moll:2014} was also able to find a detonation
-in a similarly massive system. Very recently, \cite{kashyap:2015} 
-found that a detonation occurred in a similarly massive system. Notably,
-the detonation occurred self-consistently and did not need to be inserted 
-by hand. \cite{dan:2012} and \cite{dan:2014} performed a large sweep 
-of the parameter space for merger pairs and
-found that pure carbon-oxygen systems would generally not lead to
-detonations and violent mergers except for the most massive
-systems. They did find that for systems containing helium, many
-would detonate and potentially lead to SNe Ia, either through the
-aforementioned instabilities in the accretion stream, or during the
-contact phase, similar to the violent carbon-oxygen
-mergers. \cite{pakmor:2013} added a thin helium shell on their primary
-white dwarf, and found that this robustly led to a detonation of the
-white dwarf. For now there is preliminary support for the hypothesis
-that systems with helium shells (or helium WDs), and very massive carbon-oxygen binaries,
-could robustly lead to events resembling SNe Ia.
-
-Given the above, why is another approach using a different simulation
-code warranted? First and foremost, reproducibility of the results
-across simulation codes and algorithms is important for gauging
-confidence in this result. Most of the existing results that study 
-the viability of double degenerate systems as progenitors for
-Type Ia supernovae (that is, including a realistic 
-equation of state and nuclear reactions) have
-used the SPH method. SPH codes have a number of features which do aid
-them in the study of these systems, such as excellent conservation of
-angular momentum. However, the question of whether a prompt detonation
-in a merger actually happens depends in detail on the nature of the
-gas at the interface between the two stars, which is at much lower
-density than the rest of the stellar material. The SPH codes use
-uniform mass particles, so their effective resolution is
-\textit{lowest} at the stellar surface. In contrast, a grid-based code
-with adaptive mesh refinement can zoom in on the regions where
-hotspots will develop, while also maintaining high enough resolution
-in the high-density regions to adequately capture the large-scale mass
-transfer dynamics. However, there are also outstanding questions of
-convergence in SPH (e.g.\ \citealt{zhu_SPH:2014}) and whether the method
-correctly captures fluid instabilities. This is an important question
-for white dwarf mergers because of the likely importance small-scale
-instabilities will have on the evolution of the low-density gas at the
-primary's surface. The pioneering work of \cite{agertz:2007} compared
-grid and SPH codes and found some important differences. Most relevant
-for this discussion is that the SPH codes could not adequately handle
-mixing from the Kelvin-Helmholtz instability in the test they
-propose. As pointed out by \cite{price:2008}, this is not a result of
-SPH being inherently unable to model this instability, but instead it
-is attributed to the fact that the standard SPH evolution equations do
-not have a mechanism for capturing discontinuities in internal
-energy. \citeauthor{price:2008} showed that the addition of an
-artificial thermal conductivity can dramatically improve the ability
-of the SPH codes to exhibit this instability. There have since been a
-number of other papers discussing this issue, but to our knowledge
-none of these improvements have yet been incorporated into an SPH
-model of a WD merger. Another reason for caution is that other than the
-most recent results of \cite{kashyap:2015}, no simulation has self-consistently
-resulted in a thermonuclear detonation. Reproducibility of the detonation 
-through direct numerical simulation is critical for building 
-confidence in this progenitor model.
-
-This is the first in a series of papers designed to address these
-outstanding theoretical issues for white dwarf mergers. This work will
-discuss the verification of our hydrodynamics code for simulating
-these events. Later efforts will look at the initial conditions of the
-system, the robustness with which a hotspot is found from which a
-detonation could occur, and the importance of the initial white dwarf
-models, which should be more sophisticated than simple carbon-oxygen
-mixtures and in principle should use results from modern stellar
-evolution calculations. Section \ref{sec:Numerical Methodology}
-describes our code and why it can provide useful results compared to
-other methodologies used for this problem. Section
-\ref{sec:initial_models} describes the method we use for setting up a
-binary white dwarf simulation. Section \ref{sec:Tests} discusses a few
-test problems that we use to demonstrate that our code accurately
-solves the equations of fluid dynamics. Section \ref{sec:Performance}
-demonstrates that the software scales well for supercomputer
-applications. Finally, Section \ref{sec:Conclusions and Discussion}
-recaps what we have shown and highlights some of the future work we
-plan to do.
-
-%==========================================================================
-% Numerical Methodology
-%==========================================================================
-\section{Numerical Methodology}\label{sec:Numerical Methodology}
-
-To study the white dwarf merger problem, we use the grid-based
-hydrodynamics code CASTRO \citep{castro}. CASTRO solves the Euler
-equations, along with the inclusion of optional modules for gravity,
-nuclear reactions and thermodynamics. CASTRO is based on the BoxLib
-adaptive-mesh refinement (AMR) framework \citep{rendleman:2000}, which
-represents fluid data on a mesh where regions of interest have higher
-spatial resolution. CASTRO is highly parallel and is designed for
-large-scale use on modern supercomputers; see Section
-\ref{sec:Performance} for information on how CASTRO performs for our
-problem. The next few subsections describe our approach to each of the
-physics components used in this work. We direct the reader to the
-original code paper for a full description of CASTRO's approach to
-solving the equations of hydrodynamics. In this work, we report mainly
-on the changes we have made to the code since its original release,
-for the purpose of approaching this problem.
-
-\subsection{Hydrodynamics}
-
-The Euler equations for hydrodynamics (in the absence of source terms) in conservative form are:
-\begin{align}
-  \frac{\partial \rho}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}) \label{eq:euler_density}\\
-  \frac{\partial \rho \mathbf{u}}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}\mathbf{u}) - \bm{\nabla}p \label{eq:euler_momentum}\\
-  \frac{\partial \rho E}{\partial t} &= -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u}). \label{eq:euler_energy}
-\end{align}
-Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity
-vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the
-total specific energy, where $e$ is the internal (thermal) specific
-energy (energy per unit mass).
-
-We use the unsplit piecewise-parabolic method (PPM) solver in CASTRO
-to advance the hydrodynamics system in time \citep{ppmunsplit}.  A
-number of changes were made to the solver, which are detailed in the
-Appendix.  These changes bring the algorithm more in line with that of
-\cite{ppm}. CASTRO as originally released featured a slightly modified
-version of the higher resolution limiters of
-\cite{colella_sekora:2008}, which can be accessed in the code using
-\texttt{castro.ppm\_type = 2}. The advantage of this limiter is that
-it preserves physical extrema rather than clipping them off as in the
-original approach of \cite{ppm}. However, we found these limiters to
-be unsatisfactory for our problem. There are many regions in our
-problem with large density gradients (such as the interface between
-the star's atmosphere and the ambient gas outside of it) and in these
-regions the algorithm can yield negative densities. This often results
-from the limiters interpreting these gradients as being true
-minima. As a result, we use the original limiter, which is strictly
-monotonicity preserving in the parabolic profiles it generates; this
-is activated with \texttt{castro.ppm\_type = 1}.
-
-A related issue that required a code improvement is that in cases of
-large density gradients such as the edge of a star, it is possible to
-generate negative densities in zones even with the more strongly
-limited PPM. This can occur if a region of large density is moving
-away from an ambient zone at relatively large speeds; then the net
-density flux in the ambient zones can be large enough to unphysically
-drag the density below zero. In practice, this will occur at the
-trailing edge of a star that is moving across a grid. In such a
-situation, there are two main approaches one could take: either
-explicitly introduce a positivity-guaranteeing diffusive flux, or
-reset the characteristics of the affected zone. We choose the latter
-approach. Even though it is non-conservative, it preserves a
-characteristic we value, which is to keep the edge of the stars
-relatively sharp, as they physically should be. Since the mass of the
-affected zones is typically already fairly low, this should not
-seriously affect the energy conservation properties of our
-simulation. Our strategy for a reset is as follows: when the density of 
-a zone is below a pre-determined density floor (which is typically 
-$10^{-5}\ \text{g cm}^{-3}$ for our stellar simulations), we look
-at all adjacent zones and find the zone with the highest density.
-If it is above the density floor, then we set the field values 
-(density, momentum, energy, and temperature) of the
-resetted zone to be exactly equal to the field values of this 
-adjacent zone. If no adjacent zone reaches the density floor, then
-the zone is set to the density floor, and given a temperature equal 
-to the temperature floor for our simulations (which is typically 
-$10^{5}\ \text{K}$ for our stellar simulations). We then recompute 
-the thermodynamics to be consistent with these values. The 
-velocity of the zone is set to zero. This latter approach will only
-occur in very rare situations, and is mainly there as a safeguard.
-
-CASTRO's approach to adaptive mesh refinement, based on its underlying
-BoxLib framework, is to refine zones based on certain user-specified
-criteria that tag regions of interest for higher spatial
-resolution. Data is represented on one of a number of AMR levels,
-where each level corresponds to a set of zones at the same resolution,
-which covers a subset of the domain covered by the level immediately
-below it. We typically call the level 0 grid the \textit{coarse} grid,
-which has the lowest spatial resolution. Each finer, higher-level grid
-has a higher resolution than the grid below it by some integer factor
-$N$, which is restricted to be $N = 2\ \text{or}\ 4$ in the code. The
-zones are strictly contained within the rectangular extent of the
-underlying coarser zones (the code is not restricted to representing
-only Cartesian geometries, but we use a Cartesian mesh with uniform
-spacing in each dimension for the present study). For the time
-evolution of the AMR system we use subcycling, where each AMR level is
-advanced at a different timestep and a correction step is applied at
-the end to synchronize the various levels. Generally the number of
-subcycled timesteps is equal to the jump in refinement between levels,
-so for example on a grid with three levels and two jumps of four in
-refinement, the level 2 zones will have 16 times higher spatial
-resolution than the coarse grid and there will be 16 level 2 timesteps
-per level 0 timestep.
-
-The boundary conditions on the hyperbolic system are simply
-zero-gradient zones that allow material to flow directly out of the
-domain. Using AMR, we make the coarse grid large enough that the
-boundaries are relatively far from the region of interest. This
-ensures that any boundary effects do not pollute the inner region
-where the stars will eventually make contact.  We further make the
-restriction that refined grids cannot reach the domain boundary.
-
-\subsection{Microphysics}
-
-The equation of state (EOS) for our simulations is the Helmholtz EOS
-\citep{timmes_swesty:2000}. This models an electron-positron gas of
-arbitrary relativity and degeneracy over a wide range of temperatures
-and densities. Thermodynamic quantities are calculated as derivatives
-of the Helmholtz free energy, and the values are interpolated from a
-table. The natural variables of the Helmholtz free energy are
-temperature and density, and calling the EOS is simplest in this
-form. However, in hydrodynamics we often have the density, and
-internal energy as independent variables, and we want to obtain the
-temperature, pressure, and other quantities. To do this, we employ a
-Newton-Raphson iteration over the temperature (given some sufficient
-starting guess) until we find the temperature that corresponds to the
-desired internal energy. Sometimes this process fails to converge and
-the iterative value approaches zero. In these cases we employ a
-``floor'' that limits how low the temperature can go (typically 
-$10^5$ K). There is a choice here how to proceed: we can either
-assign this floor value to the temperature and let that zone be
-thermodynamically inconsistent (the original behavior in CASTRO), or
-we can adjust the internal energy to be thermodynamically consistent
-with the temperature, at the cost of violating energy conservation. We
-have found in some test problems of strong one-dimensional shocks that reach 
-the temperature floor that the latter yields more accurate results. 
-However, allowing the equation of state call to update the 
-internal energy can actually result in significant changes to the 
-total energy of the system over long periods of time, 
-due not just to resets in low-density zones but also to small 
-inconsistencies between the energy given to the EOS and the energy 
-that is consistent with the returned temperature. These inconsistencies
-are dependent on the threshold we set that determines how close the 
-returned energy should be to the original input energy. While this 
-threshold is typically very small in an individual zone (a relative 
-difference of $10^{-8}$ by default in CASTRO), over time and given 
-a large number of zones, this can result in a significant energy 
-drift. This is a serious enough problem that we opt for the energy 
-conserving approach for our simulations.
-
-CASTRO has the ability to do both nuclear reactions and radiative 
-transport (in the flux-limited diffusion approximation). For all simulations 
-in this paper we do not enable either, and will delay discussion of 
-these modules until a later paper in this series.
-
-\subsection{Gravity}
-\label{sec:gravity}
-
-We solve the Poisson equation for self-gravity for our problem,
-\begin{equation}
-  \nabla^2 \Phi(\mathbf{x}) = 4\pi G\, \rho(\mathbf{x}),
-\end{equation}
-where $\Phi$ is the gravitational potential, $G$ is the gravitational
-constant, and $\rho$ is the mass density.\footnote{In the CASTRO code, the 
-right-hand side is negated and therefore $\Phi$ is positive. We use the 
-sign convention that is typical for astrophysics in this paper. 
-When $\Phi$ appears in the code it is negated to compensate for this.} 
-The solution of this equation in CASTRO is described in \cite{castro}, and
-consists of both level and composite solves, and a final
-synchronization at the end.
-
-\subsubsection{Coupling to Hydrodynamics}\label{sec:gravity_hydro_coupling}
-
-The effect of gravity on the hydrodynamical evolution is typically
-incorporated by the use of a source term for the momentum and energy
-equations. In a finite volume methodology, the momentum source term appears in 
-integral form as
-\begin{equation}
-  \left.\frac{\partial (\rho \mathbf{u})}{\partial t}\right|_{\text{grav}} = \frac{1}{\Delta V} \int \rho \mathbf{g}\, dV
-\end{equation}
-and for the energy source term it is
-\begin{equation}
-  \left.\frac{\partial (\rho E)}{\partial t}\right|_{\text{grav}} = \frac{1}{\Delta V} \int \rho \mathbf{u}\cdot\mathbf{g}\, dV \label{eq:cell_center_gravity_source}.
-\end{equation}
-Here $\Delta V$ is the cell's volume.
-In most hydrodynamics codes these are discretized as $\rho\,
-\mathbf{g}$ and $\rho\, \mathbf{u}\,\cdot\mathbf{g}$, respectively, 
-where $\rho$, $\mathbf{u}$, and $\mathbf{g}$ 
-are evaluated at the zone center.
-
-There are two ways that this source term enters the hydrodynamics update.
-First, during the hydrodynamics update, we alter the edge states that enter
-into the determination of the fluxes. To second order in space and time, 
-this can be done using the cell-centered
-source term evaluated at time-level $n$. The details of this are described in 
-the appendix. Second, after the hydrodynamics step, we add the time-centered source term
-to the state. First we will describe how we do this for the momentum,
-and then we will describe our approach for the energy. This discussion will be somewhat detailed.
-However, we believe that the attention is necessary because of the importance of accuracy
-in the gravitational source terms for our problem. The stability of the white dwarf binary
-system is dependent in large part upon accurate coupling of the hydrodynamics and gravity;
-an error in this approach could lead to, for example, a spurious mass transfer episode
-that might lead us to very different conclusions about the long term stability of such a system.
-Such considerations are generally unimportant for spherically-symmetric single star calculations,
-but are of the utmost importance in a simulation where the global gravitational field can change 
-quite significantly over the course of the simulation.
-
-In a system with self-gravity, total momentum is conserved if the spatial domain
-includes all of the mass of the system. This must be the 
-case because each mass element exerts an equal and opposite gravitational force 
-on every other mass element. However, the standard approach does not necessarily
-guarantee that momentum is conserved numerically. We cannot represent a vacuum state 
-in our code, so there is a small but non-zero density on the edge of the grid. 
-This allows momentum to leak out of the domain even if the gravitational source term 
-is written in an explicitly conservative manner. To see this, one can write the 
-density in terms of the potential and then consider its spatial discretization. For simplicity,
-we will consider one spatial dimension.
-\begin{align}
-  -\rho_{i}  \frac{d\Phi_{i}}{dx} &= -\frac{1}{4\pi G} \frac{d^2\Phi_i}{dx^2} \frac{d \Phi}{dx} \notag \\
-  &= -\frac{1}{4\pi G} \left[\frac{\Phi_{i-1} - 2 \Phi_{i} + \Phi_{i+1}}{\Delta x^2}\right] \left[ \frac{\Phi_{i+1} - \Phi_{i-1}}{2\Delta x} \right] \notag \\
-  &= -\frac{1}{8\pi G \Delta x^3} \left[ \Phi_{i+1}^2 - \Phi_{i-1}^2 - 2\Phi_i\left(\Phi_{i+1} - \Phi_{i-1}\right) \right] \label{eq:momentum_discretization}
-\end{align}
-It is easy to verify that adding the source terms two cells to the left
-and right of this zone will result in complete cancellation of the source terms.
-However, if the potential if non-zero outside of the domain, then there will be
-momentum lost or gained from the grid, which will be encapsulated in the ghost cells
-just outside the domain. In addition, when we replace the Laplacian above by the full
-three-dimensional stencil including the $y$ and $z$ derivatives, and depending on the
-discretization these will not be cancelled at all. This latter problem can be resolved by
-writing the momentum update in an explicitly conservative way.
-
-\citet[Chapter 4]{shu:1992} observes that it is possible to describe the source term 
-for the momentum equation by taking the divergence of a gravitational stress tensor,
-\begin{equation}
-  G_{ij} = -\frac{1}{4\pi G}\left(g_i g_j - \frac{1}{2}|\mathbf{g}|^2\delta_{ij}\right).
-\end{equation}
-The momentum equations are then written explicitly in conservative form.
-The flux at any zone boundary is added to one cell and
-subtracted from another, so that the total momentum in the domain interior stays constant to
-within numerical roundoff error. This result can be derived by analytically recasting 
-Equation \ref{eq:momentum_discretization}. In the continuum limit, the two momentum
-formulations are identical. Thus the latter has been advocated by, for example, 
-\cite{jiang:2013} for the ATHENA code. However, in a finite discretization 
-the divergence of the gravitational acceleration is no longer guaranteed to equal
-the zone density. In particular, we find that the mixing of the gravitational components
-means that the truncation error in the gravitational field can lead to large errors
-that imply a density much different than the zone's actual density. This is especially
-problematic in a simulation with a low-density ambient medium, where even a small error 
-in the momentum update can lead to large changes in a zone's momentum. By continuing to explicitly
-use the cell density in the momentum update, we can forgo this possibility: the size of the update
-will always be suitably small if the zone's density is small. Thus for our simulations
-we continue to use the standard source term for the momentum.
-
-Time centering of this source term is done in CASTRO using a predictor-corrector approach.
-At the start of a coarse grid timestep, we solve the gravitational potential for the density $\rho^n$.
-We then add to the momenta a prediction of the source term that is first-order accurate in time, 
-$\Delta t\, \rho^n\, \mathbf{g}^n$. After the hydrodynamics update, we recalculate
-the gravitational potential based on the new density, $\rho^n$, and then add 
-$-(\Delta t/2) \rho^n \mathbf{g}^n + (\Delta t/2) \rho^{n+1} \mathbf{g}^{n+1}$ to the momenta.
-
-Dealing with the energy equation is more difficult. The central
-challenge is to write down a form of the discretized energy equation
-that explicitly conserves total energy when coupled to gravity. When
-gravity is included, the total energy is
-\begin{equation}
-  \rho E_{\text{tot}} = \rho E + \frac{1}{2}\rho\Phi = \frac{1}{2}\rho \mathbf{u}^2 + \rho e + \frac{1}{2}\rho\Phi, \label{eq:total_energy_gravity}
-\end{equation}
-where $\rho E$ is the total gas energy from the pure hydrodynamics equation,
-and the factor of $1/2$ in the gravitational energy term is necessary
-for simulations with self-gravity to prevent double-counting of
-interactions. The standard approach has been to note that the total energy update
-can be written down by taking the dot product of the velocity and the momentum source term,
-so that the time-centered energy update is
-\begin{equation}
-  \left.\Delta (\rho E)\right|_\text{grav} = \frac{\Delta t}{2} \left[ \rho^n \mathbf{u}^n \cdot \mathbf{g}^n + \rho^{n+1} \mathbf{u}^{n+1} \cdot \mathbf{g}^{n+1}\right].
-\end{equation}
-However, this does not conserve the total energy of Equation \ref{eq:total_energy_gravity}.
-In particular, we notice in our simulations a long-term secular change in the gravitational potential energy
-that is not accounted for in the gas energy. One solution to this problem has been offered by 
-\cite{jiang:2013} and \cite{marcello:2012}. They propose to evolve the total energy instead of the gas energy 
-in the hydrodynamics equations. When this is done, one can write down an explicitly conservative 
-formulation:
-\begin{equation}
-  \frac{\partial \rho E_{\text{tot}}}{\partial t} = -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u} + \mathbf{F_g}).
-\end{equation}
-Here $\mathbf{F_g}$ is a flux term which explicitly accounts for the change in the total energy due to self-gravity. 
-By using Equation \ref{eq:cell_center_gravity_source} and the continuity equation, it is straightforward to show that
-\begin{equation}
-  {\bm\nabla} \cdot \mathbf{F_g} = -\rho \mathbf{u} \cdot \mathbf{g} - \frac{\partial}{\partial t}\left(\frac{1}{2}\rho\Phi\right). \label{eq:gradient_F_g}
-\end{equation}
-From here one can use the continuity equation to explicitly calculate the flux $\mathbf{F_g}$, as in \cite{jiang:2013}:
-\begin{equation}
-  \mathbf{F_g} = \frac{1}{8\pi G}\left(\Phi\bm{\nabla}\dot{\Phi} - \dot{\Phi}\bm{\nabla}\Phi\right) + \rho\mathbf{u}\Phi \label{eq:jiang_flux}.
-\end{equation}
-Here the dots refer to time derivatives. This has some attractive features. 
-In particular, for a simulation with AMR, a flux-based approach
-can quite naturally be accommodated in a way that conserves energy across levels, when fluxes
-at interfaces between coarse and fine grids are matched. We note in passing that although
-\citeauthor{jiang:2013} claim that this approach requires a Poisson solve for the time derivative of the potential,
-$\dot{\Phi}$, this is not needed as in practice we will already have the potential at time-levels
-$n$ and $n+1$, and to second order in time, $\dot{\Phi}^{n+1/2} = (\Phi^{n+1} - \Phi^{n}) / \Delta t.$
-
-Now, the first term on the right-hand side of Equation \ref{eq:gradient_F_g} is the old source term. The second term simply 
-accounts for the time rate of change of the gravitational potential energy. But since
-\begin{equation}
-  \frac{\partial(\rho E_{\text{tot}})}{\partial t} = \frac{\partial (\rho E)}{\partial t} + \frac{\partial (1/2) \rho \Phi}{\partial t},
-\end{equation}
-this does not actually add any information, since we can subtract off the time derivative 
-of the potential energy and be back where we started. This demonstrates that the problem
-with the original approach lies in the numerical discretization; if this is done correctly,
-we can still properly conserve total energy.
-
-Before pursuing that, we wish to point out the reason why we prefer not to use this 
-approach of solving a total energy equation. The expression for the flux $\mathbf{F_g}$
-given in \cite{jiang:2013} includes terms that are proportional to $\Phi$ and ${\bm{\nabla}} \Phi$,
-but that are not explicitly proportional to the zone density $\rho$. Thus the approach suffers
-from the same problems as the conservative momentum formulation discussed above. Truncation errors
-in the gravitational field will result in an approximation to the density that can be quite 
-different from the actual density, especially in low-density regions. In practice,
-we have found this to be such a serious problem that this approach is not feasible for our problem.
-The formulation of \cite{marcello:2012} does not suffer from the same problem. However, a 
-remaining problem is that in this formalism, the internal energy
-is calculated by subtracting the potential energy and the kinetic energy from the 
-total energy. Due to numerical discretization errors, however, this can result in a serious 
-error in the internal energy. This problem can be eased by the dual energy formalism,
-but not completely. This means that we will potentially have very large errors in a 
-zone's temperature, which can have serious repercussions in the reacting flows we will eventually consider,
-as many reaction rates are very sensitive to temperature. Therefore our goal is instead to consider an approach
-that is explicitly \textit{non}-conservative for the gas energy $(\rho E)$, but which
-instead conserves the total energy $(\rho E_{\text{tot}})$ over the computational domain 
-(subject, of course, to small errors due to energy leaking from the domain boundary).
-
-The question of how to do this was resolved by \cite{arepo}. 
-The essential problem with the standard approach is that 
-fluid movements in a grid code are calculated using
-fluxes at zone boundaries, while the standard gravity source term is
-an estimate of the flux using the zone center quantities. This can
-lead to a striking mismatch when the gradient across a zone is
-large. We have found that this is especially important at the edge of
-a star, where the large density gradient means that the momentum and
-energy fluxes can transition sharply even across half of a zone
-width. Another problem is that in a merger calculation in which two
-stars are moving on a grid, large quantities of material are being
-advected, so even a slight error in the calculation for the
-high-density material can manifest itself as a long-term error in
-energy conservation. This becomes apparent in Section
-\ref{sec:kepler}.
-
-We start with the integral formulation of the energy source term (Equation
-\ref{eq:cell_center_gravity_source}), but do not discretize it at
-the cell center.  Instead, we recognize that in the limit that the velocity is 
-constant, $\mathbf{u} = {\bm\nabla}\left[\mathbf{r}\cdot\mathbf{u}\right]$,
-where $\mathbf{r}$ is the position vector originating from the zone center.
-Then, since in the cell-centered approach we assume that the density, velocity,
-and gravitational acceleration are constant across the zone, we can include all terms
-inside the gradient operator. This allows us to use the divergence theorem to evaluate
-the quantities on the zone faces. To be second-order accurate in time, the fields must 
-all be evaluated at time-level $n+1/2$, but for clarity we will neglect the superscript.
-\begin{align}
-  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{\Delta t}{\Delta V} \int \rho \mathbf{u}\cdot\mathbf{g}\, dV_i \notag \\
-     &= \frac{\Delta t}{\Delta V} \int \rho\, {\bm \nabla} \left[\mathbf{r} \cdot \mathbf{u}\right]\cdot\mathbf{g}\, dV_i \notag \\
-     &= \frac{\Delta t}{\Delta V} \int {\bm \nabla} \cdot \left[ \rho\, (\mathbf{r} \cdot \mathbf{u})\, \mathbf{g}\right]\, dV_i \notag \\
-     &= \frac{\Delta t}{\Delta V} \int \rho\, (\mathbf{r} \cdot \mathbf{u})\, \mathbf{g} \cdot \mathbf{dS}_i.
-\end{align}
-Here $\mathbf{dS}$ is a surface element whose vector points normal to
-and away from the surface. In one dimension, this expression evaluates to
-\begin{align}
-  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{\Delta t}{\Delta V} \rho_{i-1/2}\, \frac{\Delta x}{2} u_{x_{i-1/2}} g_{x_{i-1/2}} \Delta A_{i-1/2} \notag \\
-             &+ \frac{\Delta t}{\Delta V} \rho_{i+1/2}\, \frac{\Delta x}{2} u_{x_{i+1/2}} g_{x_{i+1/2}} \Delta A_{i+1/2} \notag \\
-             &= \frac{1}{2} \rho_{i-1/2} u_{x_{i-1/2}} g_{x_{i-1/2}} \notag \\
-             &+ \frac{1}{2} \rho_{i+1/2} u_{x_{i+1/2}} g_{x_{i+1/2}}.\label{eq:edge_centered_gravity_source}
-\end{align}
-(We note that this differs slightly from Equation 94 in \cite{arepo} in that
-we evaluate the gravitational acceleration at the zone face rather than the zone center.
-We are free to do this since we assumed a uniform gravitational field to begin with.)
-In the limit of constant fields, this expression simply averages out to the standard 
-source term at the zone center. However, when the fields vary this can give very different results.
-
-As pointed out by \citeauthor{arepo}, $(\rho u)$ evaluated at the faces at time level $n+1/2$ is the mass flux
-that is obtained from the Riemann solver in the hydrodynamics step,
-so $(\rho u)_{i-1/2}\, \Delta t\, \Delta A = \Delta m_{i-1/2}$ is the mass exchanged between zones $i$ and $i-1$.
-Additionally, to second order $g_{x_{i-1/2}} \Delta x = \Phi_{i} - \Phi_{i-1}$. Therefore the energy gained by the zone is
-\begin{align}
-  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{1}{2} \Delta m_{i-1/2} \left(\Phi_{i} - \Phi_{i-1}\right) \notag \\
-     &+ \frac{1}{2} \Delta m_{i+1/2} \left(\Phi_{i+1} - \Phi_{i}\right)
-\end{align}
-This expression says that the energy gained by a zone is due to a mass element moving through a potential gradient
-between two zones. Now, the key observation is that this does not have the telescoping property of a conservative flux;
-instead, the \textit{same} amount of gravitational potential energy is gained by zones $i$ and $i-1$ by the mass flux 
-between them. When this source term is summed up over all zones, it will exactly compensate for the change in 
-gravitational potential energy, except for any energy lost through the boundaries. The extension of this energy change 
-to multiple dimensions simply means that we include the mass fluxes and potential gradients across the $y$ and $z$ faces.
-We have found that in practice this method for conserving the total energy is the least susceptible to deleterious effects
-from numerical discretization errors. However, it does mean that there can be sharp changes in the total energy of a zone
-in regions near a white dwarf's surface, because the mass fluxes into the ambient medium can be quite large compared to
-the mass of the zone. This effect diminishes at higher resolutions as the gradients become better resolved.
-
-Our strategy for coupling the gravity to the hydrodynamics on a uniform grid can thus be summarized as follows. 
-At the beginning of a timestep, we solve the Poisson equation given the density $\rho^n$. We then solve the 
-Euler equations and advance the system in time by $\Delta t$ using zone-centered fluxes at time-level $n+1/2$. 
-We add $\Delta t \rho^n \mathbf{g}^n$ to the momentum equations. We then solve the Poisson equation again, given
-the density $\rho^{n+1}$. We then add $-(\Delta t/2) \rho^{n} \mathbf{g}^{n} + (\Delta t / 2) \rho^{n+1} \mathbf{g}^{n+1}$
-to the momentum equations, time-centering the source term. We simultaneously add $\sum_{j} -(1/2) \Delta m_{j} (\Phi - \Phi_j)$
-to the zone's energy, where the index $j$ runs over all adjacent zones. This conserves the total energy on the grid
-to roundoff error, modulo any losses from an open domain boundary.
-
-\subsubsection{Boundary Conditions}\label{sec:gravity_boundary_conditions}
-
-Analytic solutions to the Poisson equation customarily assume that the
-potential vanishes at large distances from the region of non-zero
-density. However, on a finite computational domain it is usually not
-possible to have the edges of the domain be far enough away that the
-potential can be taken to be zero there. Solving the Poisson equation
-therefore requires knowledge of the values of the potential on the
-edges of the computational domain. In principle, they can be computed
-by doing a direct sum over the mass distribution inside the domain,
-where the mass in each zone is treated as a point source:
-\begin{equation}
-  \Phi_{{lmn}} = -\sum_{{i, j, k}} \frac{G \rho_{{ijk}}}{|\mathbf{x}_{{lmn}} - \mathbf{x}_{{ijk}}|}\, \Delta V_{{ijk}}.\label{eq:direct_sum}
-\end{equation}
-Here $(i, j, k)$ are the indices of cells inside the domain, and $(l,m, n)$ 
-are the indices of boundary locations. $\Delta V$ is the volume of the
-zone. We have implemented this as an option\footnote{It is controlled
-  with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in
-CASTRO. If there are $N$ zones per spatial dimension, then there are
-$6 N^2$ boundary zones, and each boundary zone requires a sum over
-$N^3$ zones, so the direct computation of the boundary conditions
-scales as $\mathcal{O}(N^5)$.  This method is expensive enough that it is not used
-for hydrodynamics simulations (though it is useful for comparison to
-approximate solutions).
-
-In a typical simulation we place the boundaries of the domain far
-enough away from the region containing most of the mass that some
-method of approximation to this direct summation is justified. Many
-approaches exist in the literature. The original release of CASTRO
-featured the crudest possible approximation: a monopole prescription,
-where the boundary values were computed by summing up all the mass on
-the domain and treating it as a point source at the domain
-center. This is exactly correct only for a spherically symmetric mass
-distribution, and therefore is best suited for problems like
-single-star Type Ia supernova simulations that employ
-self-gravity. However, for a problem like that of a binary star system
-with significant departures from spherical symmetry, this assumption
-fails to produce accurate boundary values. This results in a
-significant drift of the center of the mass of the system over time,
-which will be discussed in Section \ref{sec:kepler}.
-
-The most natural extension of the monopole prescription is to include
-higher-order multipole moments. If the entire mass distribution is
-enclosed, then the potential can be expanded in a series of spherical
-harmonics $Y_{lm}$:
-\begin{equation}
-  \Phi(\mathbf{x}) = -\sum_{l=0}^{\infty}\sum_{m=-l}^{l} \frac{4\pi}{2l + 1} q_{lm} \frac{Y_{lm}(\theta,\phi)}{r^{l+1}}, \label{spherical_harmonic_expansion}
-\end{equation}
-where $q_{lm}$ are the so-called multipole moments. The origin of the
-coordinate system is taken to be the center of the computational
-domain, and $r$ is the distance to the origin. The multipole moments
-can be calculated by expanding the Green's function for the Poisson
-equation as a series of spherical harmonics.
-\begin{comment}
-, which yields
-\begin{equation}
-  q_{lm} = \int Y^*_{lm}(\theta^\prime, \phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3x^\prime. \label{multipole_moments_original}
-\end{equation}
-\end{comment}
-After some algebraic simplification of Equation
-\ref{spherical_harmonic_expansion} using the addition theorem for
-spherical harmonics,
-\begin{comment}
-\begin{align}
-  &\frac{4\pi}{2l+1} \sum_{m=-l}^{l} Y^*_{lm}(\theta^\prime,\phi^\prime)\, Y_{lm}(\theta, \phi) = P_l(\text{cos}\, \theta) P_l(\text{cos}\, \theta^\prime) \notag \\
-  &\ \ + 2 \sum_{m=1}^{l} \frac{(l-m)!}{(l+m)!} P_{l}^{m}(\text{cos}\, \theta)\, P_{l}^{m}(\text{cos}\, \theta^\prime)\, \left[\text{cos}(m\phi)\, \text{cos}(m\phi^\prime) + \text{sin}(m\phi)\, \text{sin}(m\phi^\prime)\right].
-\end{align}
-\end{comment}
-the potential outside of the mass distribution can be written as:
-\begin{align}
-  \Phi(\mathbf{x}) &= -\sum_{l=0}^{\infty} \left\{Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} \right. \notag \\
-    &+ \left. \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right\}.\label{multipole_potential}
-\end{align}
-$P_l(x)$ are the Legendre polynomials and $P_{l}^{m}(x)$ are the associated Legendre polynomials.
-$Q_l^{(0)}$ and $Q_{lm}^{(C,S)}$ are variants of the multipole moments that involve integrals of $P_l$ and $P_l^m$, respectively, over the computational domain.
-\begin{comment}
-\begin{align}
-  Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
-  Q_{lm}^{(C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
-  Q_{lm}^{(S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime.
-\end{align}
-\end{comment}
-This approach becomes computationally feasible when we cut off the
-outer summation in Equation \ref{multipole_potential} at some finite
-value of $l_{\text{max}}$. If it is of sufficiently high order, we
-will accurately capture the distribution of mass on the grid. In
-practice we first evaluate the discretized analog of the modified
-multipole moments for $0 \leq l \leq l_{\text{max}}$ and $1 \leq m
-\leq l$, an operation that scales as $N^3$. We then directly compute
-the value of the potential on all of the $6N^2$ boundary zones. Since
-the multipole moments only need to be calculated once per Poisson
-solve, the full operation scales only as $N^3$. The amount of time
-required to calculate the boundary conditions will be directly related
-to the chosen value of $l_{\text{max}}$, so there is a trade-off
-between computational expense and accuracy of the result.
-
-As a demonstration of the method's accuracy, we consider the case of two 
-white dwarfs of mass ratio 2/3, loaded onto the grid by the method described 
-in Section \ref{sec:initial_models}. We terminated the simulation just after
-initialization, so that we perform only an initial Poisson solve for this 
-density distribution. We did this for values of $l_{\text{max}}$ ranging
-from 0 to 20, and we also did this using the numerically exact solution 
-provided by Equation \ref{eq:direct_sum}.  Defining the $L^2$
-norm of a field $f$ as
-\begin{equation}
-  \| f \|_2 = \left(\sum_{i,j,k} dx\, dy\, dz\, f_{ijk}^2\right)^{1/2},
-\end{equation}
-we computed the $L^2$ error of $\Phi$ on the entire domain for multipole 
-boundary conditions, which we will call $\Phi_l$, relative to $\Phi$ 
-obtained using the exact boundary conditions:
-\begin{equation}
-  \text{Error}_l = \frac{\|\Phi_l - \Phi_{\text{exact}}\|_2}{\|\Phi_{\text{exact}}\|_2}.
-\end{equation}
-The result is shown in Figure \ref{fig:bc_comparison}. At $l_{\text{max}} = 6$,
-the error is already well below $10^{-4}$ and we adopt this as our default 
-choice for all simulations with Poisson gravity. Our choice is also motivated by the 
-fact that at this value, the time for computing the boundary conditions on the
-coarse grid is comparable to the time for the multigrid solve itself; at higher values 
-the boundary condition calculation becomes much more expensive than the multigrid solve
-and so the overall returns are severely diminished. We note that in our implementation 
-there are no gains at all for going beyond $l_{\text{max}} \gtrsim 16$, as we 
-become limited by error due to the machine precision.
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.45]{plots/bc_comparison}
-  \caption{Error of $\Phi$ on the computational domain for a binary white dwarf simulation 
-    whose boundary conditions were computed using various values of the maximum multipole order,
-    relative to the exact solution determined by a brute force sum on the boundaries.
-    Circles represent the error at integer values, and they have been connected by a smooth 
-    line to guide the eye.\label{fig:bc_comparison}}
-\end{figure}
-
-\subsubsection{Convergence Testing}\label{sec:gravity_convergence_testing}
-
-Since the results of a merger simulation depend strongly on gravity,
-it is important to check whether proper numerical convergence is
-achieved for the Poisson solver. To do so, we created a simple test
-that inserts a sphere of radius $R$ and uniform mass density $\rho$
-onto our grid, and used CASTRO to calculate the gravitational
-potential $\Phi$ of this setup. We ensure that $R$ is an integer
-multiple of the grid spacing, and the center of the sphere is at the
-origin. The problem domain for our simulations is $[-1.6, 1.6]^3$, and
-we take $R = 1.0 \text{cm}$ and $\rho = 10^3\ \text{g cm}^{-3}$. 
-The zones with $r > R$ are filled with an ambient material of very low density 
-$\rho = 10^{-8}\ \text{g cm}^{-3}$. We run this problem at multiple 
-resolutions corresponding to jumps by a factor of two. For
-comparison, at each grid point we evaluate the analytical potential of
-a uniform sphere, which can be easily determined using Gauss' law:
-\begin{equation}
-  \Phi_{\text{sphere}}(r) = -\frac{GM}{r} \times \begin{cases} (3R^2 - r^2)/(2 r^2) & r \leq R \\ 1 & r > R \end{cases},\label{eq:sphere-analytical}
-\end{equation}
-where $M = 4\pi R^3 / 3$ is the mass of the sphere. We measure the 
-numerical error by calculating the $L^2$ norm of the error and 
-normalizing it by the $L^2$ norm of the analytical solution:
-\begin{equation}
-  \text{Error} = \frac{\|\Phi - \Phi_{\text{sphere}}\|_2}{\|\Phi_{\text{sphere}}\|_2}.
-\end{equation}
-We define the order of convergence $p$ between two simulations with a jump 
-in resolution of integer factor $m > 1$ as
-\begin{equation}
-  p = \text{log}_{m}\left(\frac{\text{Error}_{\text{low}}}{\text{Error}_{\text{high}}}\right).
-\end{equation}
-Here $\text{Error}_{\text{low}}$ is the $L^2$ error at the lower resolution 
-and $\text{Error}_{\text{high}}$ is the $L^2$ error at the higher resolution.
-We expect the error to converge at $p = 2$ given the discretization we choose. 
-For all simulations in this section and for all our main science simulations,
-we choose a relative error tolerance of $10^{-10}$ to be satisfied in the multigrid solve.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.45]{plots/phi_comparison}
-  \caption{Comparison of the CASTRO gravitational potential to the analytical solution for: 
-    a sphere of uniform density; the same sphere, but with the potential normalized using the 
-    actual amount of mass on the grid instead of the mass of a perfect sphere; and, a 
-    cube of uniform density. Plotted also is a notional curve whose slope represents
-    perfect second order convergence.\label{fig:gravity_convergence}}
-\end{figure}
-
-The results of this test are plotted in Figure
-\ref{fig:gravity_convergence}. We find that at low resolution 
-convergence is actually substantially better than second-order. The 
-explanation for this is that we are attempting to model a spherical 
-object on a rectangular grid. At very low resolution, the object does 
-not look very spherical. As the resolution is increased, the total 
-amount of mass on the grid, as well as its location, will change 
-as the sphere fills out. This means we are combining the true accuracy 
-bonus from increased resolution with the artificial accuracy bonus 
-from getting closer to solving the problem we are supposed to be solving. 
-On the other hand, at the highest resolution shown the error slightly increases.
-
-We can solve one of these two sources of error by evaluating Equation
-\ref{eq:sphere-analytical} with a mass $M$ that corresponds to the
-amount of mass actually on the grid, defined as the sum of the density
-multiplied by the volume of each cell for all cells whose density is 
-greater than ambient. The resolution study for this case is also
-plotted in Figure \ref{fig:gravity_convergence}. We still obtain
-convergence slightly better than second-order, indicating that we 
-still have the geometrical problem of the mass distribution changing.
-
-The only way to fully eliminate this effect is to use a test problem that
-does not change with resolution. The obvious companion is a cube of
-uniform density $\rho$, where now $R$ is half of the side length of
-the cube. At each resolution we use the same $R$ as for the sphere,
-which ensures that the cube always fills exactly the same fraction of
-the domain and thus has the same mass, so the only improvement comes
-from better sampling at higher resolution. The gravitational potential for this
-object has been worked out analytically by \citet{waldvogel:1976} (see
-also a similar result in \citealt{hummer:1996}). The potential is in
-Equation 15 of that paper, though the last term is missing a factor of
-$1/2$, which destroys the symmetry. Inserting this missing factor and
-performing a simple coordinate transformation so that the center of
-the cube is at the origin, the potential is
-\begin{align}
-  \Phi_{\text{cube}}(x,y,z) &= -G\rho\sum_{i,j,k=0}^1\left[x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right)\right. \notag \\
-  &+ \left. y_j z_k\, \text{tanh}^{-1}\left(\frac{x_i}{r_{ijk}}\right) + z_k x_i\, \text{tanh}^{-1}\left(\frac{y_j}{r_{ijk}}\right) \right.\notag \\
-  &\left. - \frac{x_i^2}{2}\,\text{tan}^{-1}\left(\frac{y_j z_k}{x_i r_{ijk}}\right) - \frac{y_j^2}{2}\,\text{tan}^{-1}\left(\frac{z_k x_i}{y_j r_{ijk}}\right) \right. \notag \\
-  &- \left. \frac{z_k^2}{2}\,\text{tan}^{-1}\left(\frac{x_i y_j}{z_k r_{ijk}}\right)\right]
-\end{align}
-where $x_0 = R + x$, $x_1 = R - x$, $y_0 = R + y$, $y_1 = R-y$, $z_0 =
-R+z$, $z_1 = R-z$, and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
-\texttt{Fortran} and \texttt{C}, the inverse hyperbolic tangent is
-\texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not}
-\texttt{atan2}). This formula is valid both inside and outside the
-cube. The normalized $L^2$ error for this problem is also shown
-in Figure \ref{fig:gravity_convergence}, and only for this problem 
-do we obtain perfect second-order scaling at all resolutions.
-
-The main lesson here is that in a convergence study, it is important
-to ensure that the physical problem does not change with
-resolution. Since in the case of spherical objects on rectangular
-grids the effect may be to artificially boost convergence with resolution,
-in a simulation with spherical objects like stars one can envision a
-scenario of being fooled into believing apparently good convergence
-results that are simply a convolution of artificially high
-gravitational convergence and poor convergence in the hydrodynamics. A
-convergence study in this case is only fully valid if there is reason
-to be confident that this effect is negligible compared to other
-factors.
-
-\subsection{Rotation}\label{sec:rotation}
-
-For the evolution of binary systems, it is most natural to evolve the
-two stars in a frame that is co-rotating at the same period as the
-orbital period. Since the publication of the original code paper, CASTRO 
-now has the ability to evolve systems in a rotating reference frame. 
-Source terms corresponding to the Coriolis and centrifugal 
-force terms are added to the momentum and energy equations. In this frame, 
-the stars essentially remain stationary in their original positions due to the
-centrifugal force supporting against the gravitational attraction, and
-will remain this way as long as significant mass transfer does not
-occur. \cite{swc:2000} demonstrated (in the context of neutron star
-mergers) that conservation of angular momentum is much easier to
-obtain in the rotating reference frame than in an inertial frame in
-which stars advect large amounts of material around the domain. We
-wish to emphasize that although it is commonly stated in the
-literature that grid-based codes poorly conserve angular momentum,
-it is only generally true that grid-based codes do not exactly conserve 
-angular momentum when the equations are written in conservative form
-for linear momentum. (See also \cite{motl:2002} for an example of how 
-to evolve the hydrodynamics equations for angular momentum.) 
-The extent to which angular momentum conservation is violated
-will be a function of the resolution. When this is sufficiently high
-and a rotating reference frame is employed, excellent conservation
-properties can result, as demonstrated in Section \ref{sec:kepler}. 
-In particular, at reasonable resolution for a binary orbit our code 
-conserves angular momentum well enough that this no longer becomes 
-a reason to use a Lagrangian method such as smoothed-particle hydrodynamics.
-We note that as the stars begin to coalesce, the rotating reference frame
-will no longer provide a good approximation to the spatial motion of
-the stars and then they will begin to significantly move around the
-domain. This is not necessarily problematic because the most important
-feature of the rotating frame is that it helps ensure that the initial
-coalescence is not the result of spurious numerical loss of angular
-momentum. When significant mass transfer sets in and evolution
-proceeds on a dynamical timescale, the conservation properties may be
-slightly worse but angular momentum conservation is also less
-important.
-
-In a rotating reference frame with angular frequency vector $\bm{\omega}$, the momentum equation becomes:
-\begin{equation}
-  \left.\frac{\partial(\rho \mathbf{u})}{\partial t}\right|_{\text{rot}} = -2\, {\bm\omega} \times (\rho\mathbf{u}) - \rho {\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right).
-\end{equation}
-Here $\mathbf{r}$ is the position vector with respect to the origin. Typically we choose $\bm{\omega} = (0, 0, 2\pi / T)^T$,
-with the rotation axis coincident with the $z$ axis at $x = y = 0$.
-$T$ is the rotation period, which is the most natural quantity to specify
-for a rotating stellar system. As described in the appendix, we include this source term
-in the edge state prediction in a way that is analogous to the gravity source.
-We evaluate all quantities at cell centers. We use the same predictor-corrector 
-approach that we use for the gravity source terms to the momentum equations. However, a slight 
-difference is that the Coriolis force for each velocity component is coupled to other velocity 
-components. If the rotation is about the $z$-axis, then the discrete update to 
-$v_x^{n+1}$ depends on the value of $v_y^{n+1}$, and vice versa. If we fix the value of 
-the time-level $n+1$ quantities after coming out of the hydrodynamics update, there 
-would be a slight inconsistency between the $x$ and $y$ components of the velocity.
-
-We propose a more accurate coupling that directly solves this implicit system of coupled 
-equations. We denote by $(\widetilde{\rho \mathbf{u}})$ the value of the momentum after 
-updating it with the centrifugal force, and the time-level $n$ Coriolis force. The remaining 
-update for the time-level $n+1$ Coriolis force then appears as:
-\begin{equation}
-  (\rho \mathbf{u})^{n+1} = (\widetilde{\rho\mathbf{u}}) + \frac{\Delta t}{2} \left(-2\, {\bm\omega} \times (\rho\mathbf{u})^{n+1}\right)
-\end{equation}
-To proceed further, we assume that the rotation is about the $z$ axis with frequency $\omega$. 
-Then there is no update to the $z$-momentum, and the other equations are:
-\begin{align}
-  (\rho u)^{n+1} &= (\widetilde{\rho u}) + \omega \Delta t (\rho v)^{n+1}) \\
-  (\rho v)^{n+1} &= (\widetilde{\rho v}) - \omega \Delta t (\rho u)^{n+1})
-\end{align}
-We can directly solve this coupled system:
-\begin{align}
-  (\rho u)^{n+1} &= \frac{ (\widetilde{\rho u}) + \omega \Delta t (\widetilde{\rho v})}{1 + \omega^2 \Delta t^2} \\
-  (\rho v)^{n+1} &= \frac{ (\widetilde{\rho v}) - \omega \Delta t (\widetilde{\rho u})}{1 + \omega^2 \Delta t^2}
-\end{align}
-We use this form of the momentum update in CASTRO. This improvement is small
-but increases the accuracy of our rotating white dwarf systems over long time-scales.
-
-The update to the energy equation can be determined by taking the dot product of the velocity
-with the momentum source terms. The Coriolis term vanishes identically, and so
- the Coriolis term does no work on the fluid. The update from the centrifugal force becomes
-\begin{equation}
-  \left.\frac{\partial(\rho E)}{\partial t}\right|_{\text{rot}} = \frac{1}{\Delta V}\int \rho \mathbf{u} \cdot \mathbf{f}_R\, dV,
-\end{equation}
-with $\mathbf{f}_R \equiv  -{\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right)$. 
-This expression is identical to the gravity source under the interchange of $\mathbf{g}$ with $\mathbf{f}_R$.
-As observed by \cite{marcello:2012}, we can similarly write down a rotational potential,
-\begin{equation}
-  \Phi_R = -\frac{1}{2} \left| {\bm\omega} \times \mathbf{r} \right|^2.
-\end{equation}
-Therefore we can retrace the steps of Section \ref{sec:gravity_hydro_coupling} to obtain
-an update to $(\rho E)$ that conserves the total energy in the presence of rotation,
-\begin{equation}
-  (\rho E_{\text{tot}}) = (\rho E) + \frac{1}{2} \rho \Phi + \rho \Phi_R.
-\end{equation}
-In one dimension, this would be
-\begin{align}
-  \left.\Delta(\rho E_i)\right|_{\text{rot}} &= \frac{1}{2} \Delta m_{i-1/2} (\Phi_{R_{i}} - \Phi_{R_{i-1}}) \notag \\
-  &+ \frac{1}{2} \Delta m_{i+1/2} (\Phi_{R_{i+1}} - \Phi_{R_{i}}).
-\end{align}
-As with gravity, in three dimensions we include the mass fluxes through all six cell faces.
-
-We apply the rotation forces after the gravitation forces, but 
-there is some freedom in the order in which to apply the gravitational and rotational terms.
-This order may matter because the Coriolis force depends on the fluid velocity, and 
-in the predictor-corrector approach, we use the velocities both at 
-time-level $n$ and time-level $n+1$. If we update the latter with the gravitational force, 
-then the Coriolis force will see a different velocity than the one obtained through the 
-pure hydrodynamics step. (This does not matter for the energy equation in our new formulation,
-because the velocities used are always the time-level $n+1/2$ values coming from the Riemann solver.)
-In practice, this will not matter significantly for our simulations in this work 
-because the centrifugal force plays the dominant role in maintaining stability of non-contact 
-binary systems, and the centrifugal force does not depend on the fluid velocity.
-However, this issue may be worth exploring in future work in situations where the Coriolis 
-term is non-negligible in determining the system evolution.
-
-In all simulations performed in a rotating reference frame, we will transform all relevant
-quantities back to the inertial reference frame when reporting them in this paper. In particular,
-for every zone we add to the displayed kinetic energy the rotational energy associated with motion 
-at the circular velocity corresponding to the distance of that zone from the rotation axis.
-We do the analogous calculation for angular momentum as well. When plotting the locations of 
-stars on the grid, we rotate them by an angle corresponding to the amount of time that has 
-passed since the simulation started.
-
-%==========================================================================
-% Initial Models and Problem Setup
-%==========================================================================
-
-\section{Initial Models and Problem Setup}
-\label{sec:initial_models}
-
-At the start of any full simulation, we generate initial model white
-dwarfs by integrating the equation of hydrostatic equilibrium, taking
-the temperature and composition to be constant, and using the general
-stellar equation of state.  This results in a single non-linear
-equation to find the density in a zone given the conditions in the
-zone beneath it:
-\begin{equation}
-\frac{p_{i+1} - p_i}{\Delta x} = \frac{1}{2} (\rho_i + \rho_{i+1}) g_{i+1/2}.
-\end{equation}
-This equation is a function of $\rho_{i+1}$ only since the pressure is
-uniquely determined by the density in this case. Here, $\rho_i$ and $p_i$
-are known, $g_{i+1/2}$ is the gravitational acceleration at the
-interface between zones $i$ and $i+1$, found by simply adding up all
-the mass from zones $1, \ldots, i$ to get the enclosed mass,
-$M_{i+1/2}$, and then setting $g_{i+1/2} =
--GM_{i+1/2}/r_{i+1/2}^2$. We solve this equation for $\rho_{i+1}$
-using a Newton-Raphson iteration.
-
-We desire to specify the mass of the white dwarf, as well as its
-temperature and composition. To start the integration off, we
-therefore need to guess at a central density.  We then do a secant
-iteration over the entire integration procedure to find the central
-density needed to yield the desired total mass.  The grid spacing is
-$\Delta x = 6.25\ \text{km}$. We chose this value because no simulation
-we perform is likely to exceed this grid resolution inside the stars 
-themselves; for our normal domain size (see below), this corresponds to 
-three jumps in refinement by a factor of four. We find that for low 
-resolution runs, this is a better choice than selecting the 1D grid 
-spacing to be equal to the 3D grid spacing.
-
-We map the 1D model onto the 3D Cartesian grid by taking density,
-temperature, and composition as the independent variables,
-interpolating these to the cell centers, and then calling the equation
-of state to initialize the remaining terms. It is possible to interpolate
-instead by using pressure instead of temperature, as pressure is more 
-closely related to hydrostatic balance, but the EOS we use is so 
-insensitive to temperature that this mapping can result in large 
-deviations from the isothermal assumption we started with.  The 
-interpolation process divides each zone into $n_{\text{sub}}$ 
-sub-zones of equal volume for
-the purpose of sampling the 1D model, and the sub-zones are added
-together to obtain the full zone's characteristics. This
-sub-grid-scale interpolation is useful especially near the edge of the star,
-where the density falls off rapidly with radius. Typically we take 
-$n_{\text{sub}} = 4$.
-
-For a single star simulation, the star is simply placed at the center
-of the computational domain, which we take to be the origin. For a
-binary star simulation, we take as parameters the mass of the two
-white dwarfs and the initial orbital period $T$. Using Kepler's third
-law and assuming a circular orbit, we can then work out the orbital
-separation $a$:
-\begin{equation}
-  a = \left(\frac{GM T^2}{4\pi^2}\right)^{1/3}.
-\end{equation}
-Here $M = M_P + M_S$ is the total mass of the system, where $M_P$ is
-the specified \textit{primary} mass and $M_S$ is the specified
-\textit{secondary} mass. The primary WD will always start on the left
-side of the computational domain for our simulations, and is more
-massive than the secondary. This reflects the usual terminology in the
-literature where the primary WD is the accretor and the secondary is
-the donor. The center of mass is located at the center of the
-computational domain, and the stars lie along the $x$ axis, so that
-the primary's center of mass is located at $x = -(M_S / M)\, a$ and
-the secondary's center of mass is located at $x = (M_P / M)\, a$.
-
-The initial velocity is taken to be zero in if we are in the reference
-frame that rotates with the WDs, and if we are in the inertial frame
-the velocity in every zone is set equal to the rigid rotation rate 
-corresponding to the distance of that zone from the rotation axis, given
-the specified period $T$. Thus the inertial frame and rotating frame 
-simulations are starting off with the same initial conditions: two white 
-dwarfs locked in synchronous rotation. This is the simplest assumption to 
-make, but in the future we may explore relaxing this requirement.
-
-We do not attempt to enforce equilibrium with an additional relaxation
-step. This will be an important part of future work in this series, as
-numerous groups working on binary evolution
-\citep{swc:2000,motl:2002,rosswog:2004,dan:2011,pakmor:2012:gadget}
-have commented on the importance of equilibrium initial conditions in
-determining the evolution of the system. As a consequence of starting 
-in a non-equilibrium setting, there are 
-large density and pressure gradients near the white dwarf surfaces
-that result in large amounts of mass flowing into the ambient
-medium. This can result in spurious non-physical consequences such as 
-the total density or energy going negative in a zone. To compensate 
-for this, we start the simulation with a timestep that is a few orders 
-of magnitude smaller than that required by the CFL criterion, and allow
-the timestep to increase by 1\% each timestep so that the timestep reaches 
-its maximum allowed by the velocities on the grid over a span of approximately 
-1000 timesteps. This allows the ambient medium just outside the white dwarf
-to come closer to equilibrium with the surface without having 
-discontinuous jumps in the density or energy. For all simulations, 
-the maximum timestep is set to be equal to one-half of the CFL limit.
-
-We track a number of diagnostic quantities at the end of every coarse grid timestep. 
-For all simulations, we record the total energy (including the breakdown into
-its components: kinetic, internal, gravitational potential, rotation; we note
-that for the diagnostics we actually use $(\rho E)$ for calculation of the total energy,
-rather than explicitly calculating the sum of kinetic and internal, as this is
-the quantity that should be explicitly conserved), 
-the total angular momentum, and the center of mass of the system. 
-For stellar calculations, we separately record diagnostic 
-information about the stars. Our strategy for tracking their 
-locations is as follows: at the beginning of the calculation, we store the 
-physical center of mass $\mathbf{x}_{\text{COM}}$ of the stars as determined 
-by Kepler's third law. We also store the space velocity $\mathbf{v}_{\text{COM}}$ 
-of the stars. Then, at each new time step we make a preliminary guess for their 
-location by updating the location using the old velocity, 
-$\mathbf{x}_{\text{COM}} \rightarrow \mathbf{x}_{\text{COM}} + \mathbf{v}_{\text{COM}} \Delta t$.
-We then refine our guess for the location and velocity of each star by computing a
-location-weighted sum of the mass and velocity over the computational domain. 
-To do this, we need a cutoff for determining what counts as part of the primary 
-and what counts as part of the secondary. For this we can use the formula provided by 
-\citet{eggleton:1983} for the effective Roche radius $r_L$ of a star in a binary,
-\begin{equation}
-  \frac{r_L}{a} = \frac{0.49 q^{2/3}}{0.6^{2/3} + \text{ln}(1 + q^{1/3})}.
-\end{equation}
-In this formula we can use $q = M_S / M_P$ for obtaining the Roche radius of the secondary,
-and use the inverse value of $q$ to obtain the Roche radius of the primary.
-To a reasonable approximation, all material within the effective Roche radius
-of each star can be counted as belonging to that star. From this we obtain the mass
-of each star as well as its location and velocity. Once we have the new centers of mass,
-we compute the effective radius of each star at various density cutoffs. This involves 
-computing the volume $V$ of all zones within the Roche radius of the star's center 
-whose density is greater than the cutoff. We then compute $r_{\text{eff}} = (3V/4\pi)^{1/3}.$
-
-The computational domain has a total size of $1.024 \times 10^{10}\ \text{cm}$ in each 
-spatial dimension, and is centered at the origin. Our coarse grid has $256^3$ zones,
-corresponding to a spatial resolution of 400 km. For the present study, we choose a simple 
-refinement strategy: all zones within twice the Roche radius of each star are tagged for maximal 
-refinement. The extra buffer from doubling the size of this region ensures that the sharp density
-gradients near the edge of the star are within the zone of refinement. We also ensure that the 
-outer part of the domain is never tagged for refinement. In future work we will add criteria 
-that tag for refinement the gas between the stars, which is expected to feature nuclear burning.
-
-For ghost cells at the physical edge of the domain, we always use quantities 
-corresponding to the initial ambient material, to prevent material from raining down
-on to the grid in response to the disequilibrium outside the stars. This is an artificial 
-choice, but the dynamics outside of the stellar orbit is arbitrary anyway, due to the 
-lack of knowledge about the density and velocity field inside this gas. In addition, we employ 
-a ``sponge'' similar to that described by \citet{maestro3}. After the hydrodynamics update (but before 
-the gravity and rotation correction terms), we apply a damping force to the momentum 
-equation as follows:
-\begin{equation}
-  (\rho \mathbf{u})^{n+1} \to \frac{(\rho \mathbf{u}^{n+1})}{1 + (\Delta t / \Delta t_S) f_S},
-\end{equation} 
-where $\Delta t_S$ is a timescale for the sponge to operate on, and $f_S$ is the damping factor.
-We choose it so that that the sponge is non-operational inside a radius $r_S$ from the origin, and 
-fully applied at a radius $r_S^\prime \equiv r_S + \Delta r_S$. We then smooth the sponge 
-out between $r_S$ and $r_S^\prime$:
-\begin{equation}
-  f_{S} = \begin{dcases} 0 & r < r_S \\ \frac{1}{2}\left(1 - \text{cos}\left[\pi\left(\dfrac{r - r_S}{\Delta r_S}\right)\right]\right) & r_S \leq r < r_S^\prime \\ 1 & r \geq r_S^\prime. \end{dcases}
-\end{equation}
-For the simulations in this paper we set $r_S$ to be 75\% of the distance from the origin to the domain boundaries,
-and $\Delta r_S$ so that the sponge extends another 10\% of that distance. We set 
-$\Delta t_S = 0.01$ s, which is of the same order as the as the CFL timestep 
-for typical problem setups. We find that this sponge is necessary especially in the 
-rotating reference frame, as standing instabilities can create very large velocities 
-in the ambient fluid that drag down the global timestep by up to an order of magnitude.
-However, this provides another reason to distrust the dynamics of the low-density fluid
-outside of the stellar orbits.
-
-The software used to generate the test problems in this paper
-(as well as the manuscript itself),
-\texttt{wdmerger}\footnote{\url{https://github.com/BoxLib-Codes/wdmerger}},
-is freely available at an online repository hosting service.
-Version control in both the parent software (BoxLib, CASTRO) and in \texttt{wdmerger}
-permits us to reference the state of the code at the time a simulation
-was performed. In all plot files and diagnostic output generated by CASTRO, 
-and EPS figure files and plots generated by \texttt{wdmerger} for this work,
-we store the active \texttt{git} commit hashes of BoxLib, CASTRO and \texttt{wdmerger}.
-
-%==========================================================================
-% Numerical Test Problems
-%==========================================================================
-\section{Numerical Test Problems}\label{sec:Tests}
-
-Merger simulations face a number of numerical difficulties that are
-not present in single-degenerate Type Ia and core-collapse supernova
-simulations. In Section \ref{sec:gravity}, we discussed how the lack
-of spherical symmetry necessitates a careful look at the gravity
-solver. There are also hydrodynamical issues: the merger process will
-result in substantial motion of stellar material across the grid. This
-bulk motion presents an opportunity for advection errors to build up,
-and is only partially mitigated by evolving the white dwarfs in a
-co-rotating frame. It is therefore important to be be aware of the
-behavior of the code in such circumstances. The behavior of CASTRO for
-many standard hydrodynamics test problems was detailed in the original
-code paper \citep{castro}, and in the interest of brevity we will not
-repeat them here. Instead, we focus on a subset of problems that
-highlight the special difficulties introduced in merger
-simulations. These problems couple the hydrodynamics, gravity and
-equation of state modules. We observe that while in most non-trivial
-three-dimensional problems this creates a complexity that makes it
-impossible to determine exact analytical solutions, it is
-straightforward to devise problems for which certain global properties
-should obey simple, expected behaviors. Where possible, these should
-be quantified and a convergence study performed. We recommend that all
-simulation codes should check their performance on such problems to
-ensure their reliability when coupling the various physics modules
-necessary to build a complete and realistic simulation.
-
-\subsection{Maintaining Hydrostatic Equilibrium}\label{sec:HSE}
-
-In Section \ref{sec:initial_models} we describe the process by which
-we generate initial stellar models. While the 1D models are in
-hydrostatic equilibrium to within a small error, interpolation onto
-the 3D Cartesian grid will introduce perturbations into the solution
-\citep{zingale:2002}. Although we ensure that the initial models are
-generated with the same equation of state and are as well resolved as
-our finest grid, there will still be a hydrodynamical error associated
-with the fact that the rectangular grid cannot faithfully represent a
-spherical star. Additionally, the gravitational potential obtained by
-the multigrid solver will differ slightly from the one assumed by the
-initial model, and the operator splitting between the gravity and
-hydrodynamics should also result in small errors. As a result, we
-expect that the star will oscillate slightly about an equilibrium
-point, but that the amplitude of this oscillation should decrease with
-increasing resolution.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.45]{plots/single_star_static_1e3_radius}
-  \caption{Time evolution of the effective radius of a $0.9 \msolar$ 
-    white dwarf, seeded onto the grid using a one-dimensional hydrostatic
-    model and evolved without further relaxation. The lines represent 
-    different number of zones per spatial dimension; when this number is 
-    greater than 256, it represents an effective resolution obtained 
-    using AMR levels that cover the star. The radius is determined 
-    using the volume of the grid that has a density greater than $10^3\ \text{g cm}^{-3}.$
-    \label{fig:single_star_static_radius}}
-\end{figure}
-
-This problem was studied in the first CASTRO paper, but is worth
-revisiting here. A single star explosion simulation may only last a
-couple of seconds, and the CASTRO paper studied the behavior of the
-star after one second of evolution. However, the dynamical timescale
-of a typical carbon-oxygen white dwarf is on the order of 1--10
-seconds. Additionally, a binary orbit is typically on the order of
-10--100 seconds when a merger simulation starts, and with equilibrium
-initial conditions the system may survive for tens of orbits before
-the secondary is disrupted. When this does happen, we want to be
-confident that it was because of the dynamics of the merger process
-and not because of an instability in an individual star. Our goal here
-is thus to populate a single star onto our three-dimensional
-coordinate grid and evolve it for a period of time long enough to
-assess whether the star is truly stable, and to probe how the size of
-deviation from equilibrium is affected by grid resolution.
-
-We loaded a single star of mass $0.9\ \msolar$ onto the grid at the origin, 
-and evolved it for 200 seconds. Our diagnostic of choice is the effective 
-radius of the star, determined by the volume of the grid that has a density 
-greater than $10^3\ \text{g cm}^{-3}$ (see Section \ref{sec:initial_models} 
-for details on this measure). This choice of density is intended to 
-mark a reasonable outer edge to the star that is not immediately susceptible 
-to the numerical errors prevalent near the physical edge of the star. Figure 
-\ref{fig:single_star_static_radius} shows our results at various resolutions. 
-As expected, the star quickly approaches an equilibrium size that is different 
-(and in this case larger) than the one-dimensional model. However, the magnitude 
-of this change becomes smaller with resolution. The star is only approximately 
-in equilibrium Only when the coarse grid of $256^3$ zones has a level of 
-refinement that jumps by a factor of four. However, even then there is a slight 
-uptick in the size toward the event, implying that the numerical stability 
-is not guaranteed for arbitrarily long timescales. This result suggests that when 
-doing the equilibrium initial models that will former the basis of later calculations,
-we should carefully monitor the evolution of the stars before applying any 
-artificial damping to cause the merger, to ensure that the merger is due to 
-this applied force and not the intrinsic numerical instability of the stars.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.45]{plots/single_star_static_ke}
-  \caption{Time evolution of the kinetic energy of a $0.9 \msolar$ 
-    white dwarf. The lines have the same meaning as in Figure \ref{fig:single_star_static_radius}.
-    \label{fig:single_star_static_ke}}
-\end{figure}
-
-\subsection{Gravitational Free Fall}\label{sec:Gravitational Free Fall}
-
-A simple dynamical test to verify the gravitation physics in CASTRO is
-the case of gravitational free fall. We place two stars on the grid 
-in the manner of Section \ref{sec:initial_models}. The distance $a$ between 
-them corresponds to a chosen orbital period $T$, consistent with the total
-system mass $M$, but we disable the rotational source terms so that 
-the stars start at rest in an inertial reference frame. 
-Thus the stars will simply begin moving toward each other.
-As long as the stars remain approximately spherical, the stars can be 
-treated as point masses (this only seriously breaks down after the stars
-have come into contact). In dimensionless units where $r \to r / a$ and 
-$t \to 2\sqrt{2}\pi t / T$, the simple free fall equation of motion governing the
-distance $r$ between their centers of mass takes the form:
-\begin{equation}
-  \ddot{r}(t) = - \frac{1}{2r^2}.
-\end{equation}
-It is possible to derive a closed-form solution for the evolution time
-as a function of separation by starting with the integral formulation,
-\begin{equation}
-  t(r) = \int_{1}^{r} \frac{dr}{v(r)}.
-\end{equation}
-The velocity $v$ (in dimensionless units) can be found by noting that 
-$\ddot{r} = v\, dv / dr$ and then separating and integrating the equation 
-of motion. This yields 
-\begin{equation}
-  v(r) = \sqrt{\left(\frac{1}{r} - 1\right)}.
-\end{equation}
-For our problem $0 \leq r \leq 1$, so this is always valid. Integrating, we find
-\begin{equation}
-  t(r) = \text{arccos}\left(\sqrt{r}\right) + \sqrt{r \left(1 - r\right)}. \label{analyticalFreeFall}
-\end{equation}
-so that the point of contact would occur at $t = 1$. We actually stop the simulation
-at $t = 0.9$, which is when the effects from the extended sizes of the stars
-starts to become important. The results of our simulation for our default $256^3$ zone 
-uniform grid are shown in Figure \ref{Fig:Free Fall}. They show excellent agreement
-between the theory and the simulation results, and this agreement holds even for 
-a more moderate $128^3$ zone simulation.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.45]{plots/freefall}
-  \caption{Time evolution of two initially stationary white dwarfs,
-    mutually attracted to each other by the gravitational force. The
-    horizontal axis gives the separation of the white dwarfs, scaled
-    to the initial separation, and the vertical axis gives the elapsed
-    time of the simulation, scaled to the time it would take two point masses
-    to collide. The solid curve shows the analytical result,
-    calculated from Newtonian mechanics, and the circles show the
-    samples from the time evolution with CASTRO. For visual clarity, we 
-    show only a small fraction of the timesteps.}
-  \label{Fig:Free Fall}
-\end{figure}
-
-\subsection{Galilean Invariance}\label{sec:galileo}
-
-It is often stated in the literature that Eulerian methods for
-hydrodynamics with grids fixed in space do not obey the Galilean
-invariance of the underlying Euler equations, so that simulations
-moving at a uniform bulk velocity will appear different than an
-equivalent stationary simulation. If true, we need to understand 
-the importance of this effect when deciding whether to trust the 
-output of a code like CASTRO when applied for merger problems.
-Recently, this has come up in two ways which are of note for us 
-in the present study. We will explain these situations and display 
-the results of tests we have run to determine whether this 
-actually is a significant concern for our study.
-
-\citet{arepo} (hereafter, S10) performed a Kelvin-Helmholtz instability test and showed
-that (at low resolution) a fixed-grid code failed to develop the
-expected fluid instability when the whole fluid was moving at a
-strongly supersonic uniform velocity. (See also \citet{wadsley:2008}, 
-who used the FLASH code to simulate a hot bubble subject to mixing 
-by the Kelvin-Helmholtz instability, and also found that the mixing was affected by a 
-uniform bulk velocity.) This contrasted with the results
-of the moving-mesh code AREPO being presented in that study, which
-demonstrated Galilean invariance even at large bulk velocities. 
-Inability to correctly model the Kelvin-Helmholtz instability would 
-ave important consequences for how much we can trust the ability of 
-CASTRO to test the violent merger progenitor model. Shearing between 
-the material flowing out of the secondary and material near the 
-surface of the primary may trigger fluid
-instabilities that play an important role in the evolution of that
-gas, which is the site of the initial detonation in the prompt
-explosion model. \citet{guillochon:2010} showed for their simulation
-that Kelvin-Helmholtz instabilities produced this way may raise the
-temperature of the accreting material enough to ignite a
-detonation. Therefore if we are not correctly reproducing the
-characteristics of the Kelvin-Helmholtz instability in the case where
-there is significant mass motion on the grid, we cannot be confident
-that a detonation (or lack thereof) is not numerically
-seeded.
-
-\begin{figure*}[ht]
-  % Note that since we have periods in the filename, we need to double-
-  % bracket the filename so that includegraphics doesn't think it's 
-  % looking for something with the wrong extension. See:
-  % http://tex.stackexchange.com/a/10575
-  \centering
-  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p2_low_res_collage}}}
-  \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
-    conditions given by Equations \ref{eq:kh_ic_b_ramp} and \ref{eq:kh_ic_b}. 
-    The rows each represent a different bulk fluid velocity $v$ and 
-    the columns each represent a grid resolution $n$ (the number of 
-    zones per spatial dimension). The highest velocity simulation, 
-    $v=100$, corresponds to approximately Mach 70. Compare to 
-    \citet{robertson:2010}, Figure 7. \label{fig:kh_p2}}
-\end{figure*}
-
-\citet{robertson:2010} (hereafter, R10) observe that Galilean
-invariance of simulation results for the Euler equations occurs only
-because of truncation error in the discretization of the fluid
-equations. This takes the form of a numerical diffusion term which is
-dependent on velocity (and also resolution). The advantage of a
-moving-mesh code is that the mesh everywhere moves with the local flow
-velocity, which substantially reduces the numerical
-diffusion. R10 argue that the differences seen
-between the moving-mesh and fixed-grid code are caused by the
-interaction of this numerical diffusion with small-scale instabilities
-(which may be physical or numerical) which couple with and
-fundamentally alter the large-scale modes. Small-scale instabilities
-are seeded by the choice of a sharp initial discontinuity between the 
-fluids in the problem posed by S10. Crucially though,
-R10 point out that this problem does not
-converge with resolution and so it is not possible to know the correct
-behavior of this problem. As such, we do not know whether the
-small-scale modes found in AREPO are real, and the problem is not
-useful in formally discriminating between methodologies. They instead
-propose an alternate test with a smoother initial contact. This
-converges to the same solution qualitatively in both the stationary
-and bulk velocity cases, indicating that the code does generally
-maintain Galilean invariance (to some specified error that depends on
-resolution and the uniform flow speed).  We will see whether we can
-reproduce this result.
-
-A related question is whether our code reliably simulates the bulk
-motion of the stars across the grid, and whether such bulk motion
-affects the stability of the star. This concern is prompted by the
-study of \cite{tasker:2008}, who studied the effect of uniform
-translation on the stability of a spherically symmetric model for a
-galaxy cluster. They compared the radial profile of the cluster at
-initialization and after a period of time evolution. Using FLASH and
-ENZO, they found that a static cluster retains its shape at high
-enough resolution, while uniform translation of the cluster causes
-mixing of the core material due to numerical diffusion which results
-in an underestimation of the core's true density. The SPH codes they
-used did a better job maintaining the core density. We will perform a
-variant of this test using white dwarf models.
-
-\subsubsection{Kelvin-Helmholtz Instability}\label{sec:khi}
-
-Following \cite{robertson:2010}, we set up a Kelvin-Helmholtz test in
-the following way. The problem domain runs from 0 to 1 in both the $x$
-and $y$ directions. This is a two-dimensional test, so we run
-CASTRO in 2D mainly to avoid extra computational expense; in 3D, it 
-would merely involve replicating the problem in the $z$ direction.
-The problem involves a fluid slab of density $\rho_2 = 2.0$ traveling rightward in the
-$x$-direction at velocity $v_2 = 0.5$, sandwiched by a fluid of
-density $\rho_1 = 1.0$ traveling leftward at velocity $v_1 =
--0.5$. The density gradient is in the $y$ direction, so this creates a
-velocity shear along the interface between the fluids. The density and
-velocity distribution on the computational domain are given by:
-
-\begin{align}
-  \rho &= \rho_1 + R(y)\left[\rho_2 - \rho_1\right] \\
-  v_x  &= v_1 + R(y)\left[v_2 - v_1\right] \\
-  v_y  &= v_{\text{bulk}} + v^\prime
-\end{align}
-
-Here $R(y)$ is a ramp function that describes the transition between
-the two fluids, while $v_{\text{bulk}}$ is the bulk motion of the
-fluid in the $y$ direction and $v^\prime$ is the velocity perturbation
-that seeds the instability. The problem will be established for two
-sets of initial conditions (ICs), which we follow
-R10 in calling ICs A and B. They differ in
-their ramp function ($R_A$ and $R_B$ respectively), as well as the
-initial perturbation ($v^\prime_A$ and $v^\prime_B$ respectively), and
-the frequency of the perturbation ($n_A = 4$ and $n_B = 2$):
-\begin{align}
-  R_A &= \begin{cases} 0 & |y - 0.5| > 0.25 \\ 1 & |y - 0.5| < 0.25 \end{cases} \label{eq:kh_ic_a_ramp}\\
-  R_B &= \Big\{\left[1 + e^{-2(y-0.25)/\Delta_y}\right]\left[1 + e^{2(y-0.75)/\Delta_y}\right]\Big\}^{-1} \label{eq:kh_ic_b_ramp}\\
-  v^\prime_A &= w_0\, \text{sin}\left(n_A\, \pi\, x\right) \left\{e^{-(y-0.25)^2 / 2\sigma^2} + e^{-(y-0.75)^2/2\sigma^2}\right\} \label{eq:kh_ic_a}\\
-  v^\prime_B &= w_0\, \text{sin}\left(n_B\, \pi\, x\right). \label{eq:kh_ic_b}
-\end{align}
-
-\begin{figure*}[ht]
-  \centering
-  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p3_low_res_collage}}}
-  \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
-    conditions given by Equations \ref{eq:kh_p3_rho} through \ref{eq:kh_p3_vy}. 
-    The meaning of the rows and columns is the same as in Figure \ref{fig:kh_p2}.
-    \label{fig:kh_p3}}
-\end{figure*}
-
-Here $w_0 = 0.1$ is the scale of the velocity perturbation, $\sigma =
-0.05/\sqrt{2}$ controls the width of the Gaussian for IC A, and
-$\Delta y = 0.05$ is the transition distance scale for the smooth ramp of IC
-B. The pressure everywhere is set to $p = 2.5$, and we run this with a
-gamma-law equation of state set to $\gamma = 5/3$. Plotfiles are
-generated every 0.05 seconds, and the problem is run until $t = 2$.
-
-We run the problem for $v_\text{bulk} = [0, 1, 3, 10, 30, 100]$, and
-for each set of initial conditions run the problem at resolutions of
-$64^2$, $128^2$, $256^2$, $512^2$. For context, in these units the 
-sound speed is $c\approx 0.7$. In addition, for each initial
-condition we run simulations at the higher resolutions of $1024^2$,
-$2048^2$, and $4096^2$ for the stationary problem only. These serve 
-as a reference solution to gauge the extent to which the bulk flow 
-affects the development of the fluid instability, and to determine 
-if the problem is numerically converged.
-
-We find the same result as R10 for IC A, which is equivalent to the 
-test proposed by S10: at low resolutions and high bulk velocity, the 
-Kelvin-Helmholtz instability completely fails to develop. However, 
-the problem does not converge even qualitatively at the highest 
-resolutions we used. Our results are very similar to Figure 3 of 
-R10 so we do not show them here. For IC B, our results can be seen 
-for the normal resolutions and all velocities in Figure \ref{fig:kh_p2}.
-At low resolutions and very large bulk velocities, the fluid 
-does get significantly disrupted by numerical error. However, this 
-effect quickly converges away with resolution and qualitatively 
-at $512^2$ resolution the solution is nearly identical to the stationary 
-$v=0$ problem. We agree with R10 that this problem does converge 
-with resolution and is not subject to numerically-seeded secondary 
-instabilities at the stopping time. This is evident even at low resolutions
-by examining the first row of Figure \ref{fig:kh_p2}.
-
-\begin{figure*}[ht]
-  \includegraphics[trim = 0in 1.25in 0.25in 1.25in, clip, scale=0.65]{{{plots/kh_p3_high_res_collage}}}
-  \caption{Time series of the Kelvin-Helmholtz problem proposed by \citet{mcnally:2012}
-    as the simulation is just starting to go non-linear. The rows represent resolution, 
-    where $n$ is the number of grid cells per spatial dimension, and the columns are 
-    different snapshots in time.\label{fig:kh_p3_high_res}}
-\end{figure*}
-
-\citet{mcnally:2012} published another Kelvin-Helmholtz problem that 
-is well-posed in the sense that it converges with resolution and 
-is not subject to uncontrollable numerical instabilities. Though they 
-were not explicitly interested in the question of Galilean invariance, 
-we visit that issue here to see what can be learned. The initial 
-conditions for this problem are:
-\begin{align}
-  \rho &= \begin{dcases} \rho_1 - \rho_m e^{(y-0.25)/\Delta y} & 0.25 > y \geq 0 \\ 
-                         \rho_2 + \rho_m e^{(0.25-y)/\Delta y} & 0.5 > y \geq 0.25 \\
-                         \rho_2 + \rho_m e^{(y-0.75)/\Delta y} & 0.75 > y \geq 0.5 \\
-                         \rho_1 - \rho_m e^{(0.75-y)/\Delta y} & 1 > y \geq 0.75 \end{dcases} \label{eq:kh_p3_rho}\\
-  v_x &= \begin{dcases} v_1 - v_m e^{(y-0.25)/\Delta y} & 0.25 > y \geq 0 \\
-                        v_2 + v_m e^{(0.25-y)/\Delta y} & 0.5 > y \geq 0.25 \\
-                        v_2 + v_m e^{(y-0.75)/\Delta y} & 0.75 > y \geq 0.5 \\
-                        v_1 - v_m e^{(0.75-y)/\Delta y} & 1 > y \geq 0.75 \end{dcases} \label{eq:kh_p3_vx} \\
-  v_y &= w_0\, \text{sin}\left(4\pi x\right). \label{eq:kh_p3_vy}
-\end{align}
-Here $\Delta y = 0.025$, $w_0 = 0.01$, $v_m = (v_1 - v_2) / 2$, $\rho_m = (\rho_1 - \rho_2) / 2$, 
-and the other symbols have the same meaning as above (this means the flow direction
-is reversed compared to the original paper, so as to achieve consistency with the 
-other simulations presented here). We run this problem at all the same resolutions 
-and bulk velocities as the previous two problems. The results for the normal resolutions 
-at $t = 2.0$ are displayed in Figure \ref{fig:kh_p3}. We see a similar pattern as for 
-the test proposed by R10: as we get to higher flow speeds we need to have higher 
-spatial resolution to compensate for the increased numerical diffusion. However, 
-the accuracy is much lower for the highest bulk velocities. This is probably because 
-the initial conditions feature a transition that is twice as sharp (compare the $\Delta y$ 
-parameters in each case). This means that there are smaller scale numerical 
-perturbations on the grid, which can be amplified by numerical noise in the time 
-evolution of the system.
-
-\citet{hopkins:2014} performed this test as part of the testing of their code GIZMO. 
-They showed the late-time evolution of this system, when non-linear effects have 
-taken over and significantly disrupted the initial flow. However, at low resolution the 
-tested grid algorithm had failed to disrupt both for $v = 0$ and $v = 10$. We too ran 
-this problem until $t = 10$, and confirm that the Kelvin-Helmholtz instability damps out 
-at low resolution but goes strongly non-linear and disrupts the flow at high resolution.
-However, we strongly emphasize the point that this does not objectively 
-demonstrate a defiency in fixed-grid codes for this problem. We can only determine 
-the validity of a method when we have a trustworthy, converged solution to compare to, and 
-this is lacking for this problem at late times. As observed by \citeauthor{mcnally:2012}, 
-this is because the secondary instabilities form for this problem when the whorls of 
-the Kelvin-Helmholtz tendrils stretch out and create gradients that approach the 
-grid resolution. This is prime breeding ground for numerical noise. But because 
-the nature of this noise will depend on the resolution, it will be very different for 
-simulations at different resolutions. If these instabilities are seeded because of 
-this resolution-dependent noise and are not seeded instead in a controlled manner 
-such that they appear at the same time and location, then we simply cannot draw 
-any conclusions that bear on the question of verification from this test at late times. 
-Figure \ref{fig:kh_p3_high_res} provides a sense of this by examining the crucial 
-time at which the transition from the linear to the non-linear regime is occurring. 
-At all of these very high resolutions the secondary instabilities develop, but they 
-occur at different times and have different spatial scales for each resolution.
-
-We conclude that large bulk motions of fluid can have very significant effects 
-on numerical calculations of shear mixing in fixed-grid codes, but that this effect 
-diminishes with increasing resolution. As a result, we need only be confident that we are 
-sufficiently resolving the major mixing regions on the white dwarf surfaces. 
-If we find that this mixing occurs near the grid resolution scale, this will imply 
-that we need to ramp up the resolution in these regions using AMR. If this becomes 
-too expensive, we would need to be skeptical of any conclusions that could be drawn 
-about the effect of the mixing on the nuclear burning.
-
-\subsubsection{Moving Star}\label{sec:moving_star}
-
-To see this effect in action for a stellar simulation, we repeated the test of Section 
-\ref{sec:HSE} with a bulk velocity on the grid. We chose a velocity of $2.56 \times 10^{8}\ \text{cm s}^{-1}$.
-For context, this is comparable to the orbital velocities of the stars in Section \ref{sec:kepler}.
-This test was inspired by \citet{tasker:2008}, who considered a moving galaxy cluster 
-and who obtained a long timescale evolution by using periodic boundary conditions, 
-so that the cluster would cross the domain multiple times throughout the evolution.
-We believe that periodic boundary conditions are unrealistic for our type of simulation,
-so we prefer to do one continuous simulation where the star does not cross the boundaries.
-Since our normal grid was not large enough to allow the motion to continue for very long, we expanded 
-the domain size by a factor of four, and then included an extra refined level around the star to 
-keep the effective resolution the same. We started the star off in the lower left corner of the domain, and 
-pointed its velocity towards the upper right corner. This allowed us to evolve the star for the 
-same length of time as for the original test. We note that getting the gravity boundary conditions right 
-required us to move the origin of the problem at the bulk velocity, so that the multipole moments 
-were computed with respect to the stellar center.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.45]{plots/single_star_compare_1e3_radius}
-  \caption{A variation on Figure \ref{fig:single_star_static_radius} where
-    we now compare the ``static'' case to ``motion'' simulations where the 
-    star moves across the grid at a fixed linear speed. The lines represent 
-    the effective number of zones per dimension inside the stellar material;
-    due to the expanded size of the grid in the ``motion'' case, the 
-    physical resolution is the same in each column in the legend.
-    \label{fig:single_star_compare_radius}}
-\end{figure}
-
-In Figure \ref{fig:single_star_compare_radius}, we take the results of Section 
-\ref{sec:HSE} (the ``static'' case), and plot on top of it the results of this 
-new simulation (the ``motion'' case). We see immediately that this bulk velocity
-causes the star to be much worse at maintaining hydrostatic equilibrium. Not only 
-is the absolute size of the star significantly larger (nearly a factor of two 
-at the lowest feasible resolution we consider), but also there is a clear upward 
-trend in the size that has not terminated at any resolution by the end of the simulation.
-This again emphasizes the results mentioned earlier, that we must be careful not 
-to trust any simulation with significant mass transfer if we are not confident that the 
-mass transfer is seeded in a controllable manner and free from numerical noise.
-
-\subsection{Keplerian Orbit}\label{sec:kepler}
-
-should show:
-
-effect of resolution
-
-inertial vs. rotating frame
-
-effect of ppm\_reference and gravity update type
-
-effect of boundary conditions (size of domain and possibly multipole dependence)
-
-%==========================================================================
-% Performance
-%==========================================================================
-\section{Parallel Strategy and Performance}\label{sec:Performance}
-
-CASTRO is designed to be deployed on high-performance computing systems using 
-many thousands of processors simultaneously. It is worth briefly examining 
-our strategy for parallelizing the problem over many computational nodes 
-and our performance in situations similar to production science simulations. 
-This is especially true because some aspects of our approach to parallelism 
-have changed since the first CASTRO paper, and we have obtained improved 
-performance in certain settings.
-
-The BoxLib framework that CASTRO is based on splits each AMR level into a number 
-of boxes that collectively span the level. These boxes are distributed to processors 
-through MPI parallelism; each MPI task in general holds multiple boxes and 
-an update includes a loop over all the boxes an MPI task owns. The distribution 
-obeys a load-balancing algorithm that attempts to equalize the amount of work 
-done by each processor. BoxLib contains a number of strategies for distributing 
-work in this way, and by default uses a space-filling curve approach with a 
-Morton ordering. By experiment we have found that the most efficient load-balancing 
-strategy for our problem is actually a simple knapsack algorithm. In this approach, 
-the amount of work owned by a processor is proportional to the number of grid cells 
-associated with that processor, and the algorithm attempts to ensure that all 
-processors have a similar number of total grid cells. We demand an efficiency of 0.9,
-meaning that the average workload per processor should be no smaller than 90\% of the 
-maximum workload found on any processor. However, we find that in practice the 
-performance is largely insensitive to this choice.
-
-The size and shape of grid boxes is an important consideration for efficiency. 
-Boxes that are very small suffer from a host of problems, including the larger 
-amount of communication required between hydrodynamics solves. Additionally, 
-the multigrid solver will be less efficient if the boxes are small because there 
-will be fewer available levels for coarsening and performing V-cycles. Conversely, 
-boxes that are too large mean that there isn't enough work to go around when we 
-have a large number of processors. Good performance is the result of a careful 
-balance between these two effects. On the lower end, we require that all boxes 
-be a multiple of 16 zones in each dimension; multigrid efficiency sharply decreases 
-if this factor is any lower. On the upper end, we select the maximum grid size 
-based on the number of processors we use and the total number of cells in the 
-simulation. This size will therefore in general vary on different AMR levels. 
-Generally we select a value in between 32 and 64 zones per dimension.
-
-We use OpenMP to accelerate the work associated with the boxes owned by each 
-MPI task. Originally CASTRO used OpenMP to accelerate individual loops in the 
-hydrodynamics routines, such as the piecewise-parabolic edge state reconstruction 
-and the conservative flux update. However, there is a significant amount of 
-overhead associated with generating a new OpenMP region at each of the many 
-different loops in a hydrodynamics algorithm. This makes such a strategy 
-sub-optimal for use on many-core processors and GPUs. We have recently switched 
-to a tiling approach, where an OpenMP region is generated at the start of 
-the hydrodynamics routine and the individual threads separately work on 
-different partitions of each box. This results in much less overhead for 
-the threading. In general we obtain more efficient simulations than 
-could be obtained using MPI only, because there are fewer boxes and thus 
-less communication for a given number of processor cores. We are currently 
-developing an approach to evaluating the equation of state and nuclear reactions 
-(and possibly the entire hydrodynamics algorithm) on GPUs, which will allow us
-to take advantage of the significant computational resources embedded in GPUs on 
-certain systems.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.4]{plots/weak_scaling}
-  \caption{CASTRO weak scaling test, performed on the Blue Waters machine at 
-    NCSA. Each processor had a fixed amount of work, and we increased the 
-    number of simulation zones in concert with the number of processors. The 
-    solid curve represents perfect weak scaling, while the blue circles show 
-    CASTRO's performance at each processor count. The vertical axis measures 
-    the median time per timestep, normalized to this value for the smallest 
-    processor count.\label{fig:weak_scaling}}
-\end{figure}
-
-To examine the parallel performance of CASTRO, we performed both strong scaling 
-and weak scaling tests on the Blue Waters machine at the National Center for 
-Supercomputing Applications. For the weak scaling test, whose results are 
-shown in Figure \ref{fig:weak_scaling}, we ran a uniform grid binary white 
-dwarf simulation for resolutions of $128^3$ zones through $2048^3$ zones. The 
-number of processors was scaled with the number of zones so that each processor 
-had the same amount of work; the smallest test used 8 processors and the largest 
-used 32,768 (note that the number of processor cores on a Blue Waters node is twice 
-the number of floating point units on that node). The test was run for 10 timesteps,
-with each timestep including two Poisson solves and a hydrodynamics update (though 
-for a uniform grid calculation we generally do not need to perform any multigrid 
-iterations for the first Poisson solve in a timestep, since the density distribution 
-has not changed since the end of the last timestep). We disabled plotfile and 
-checkpoint writing, as well as calculation of diagnostic information (the latter 
-can contribute to a significant fraction of the run time at large processor 
-counts if computed every timestep). We computed the median wall time required per 
-time step for each simulation, and then normalized this to the median time per 
-timestep for the smallest simulation. We find excellent weak scaling through 
-4,096 processors. At the largest run, the simulation time required is 
-more than 1.5 times the amount required for the smallest simulation. 
-This is due entirely to the increased cost of the multigrid Poisson solve 
-in each timestep and this cannot be mitigated except by improving 
-communication or computation efficiency in the multigrid solver. We observe 
-that this weak scaling behavior with Poisson gravity is a significant 
-improvement over the results presented in the first CASTRO paper.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.4]{plots/strong_scaling}
-  \caption{CASTRO strong scaling test performed on the Blue Waters machine at
-    NCSA. The vertical axis measures the median time per timestep, and the 
-    horizontal axis measures the number of processors in the simulation. Data 
-    points are normalized to the time per timestep for the smallest number 
-    of processors. The blue circles show the data for a simulation with two AMR
-    levels (one coarse and one fine), while the red circles show the data 
-    for a test with three AMR levels (one coarse and two fine). The fine levels 
-    increase the resolution only in the regions around the stars. For each case 
-    we draw a solid curve representing perfect strong scaling.\label{fig:strong_scaling}}
-\end{figure}
-
-The strong scaling test we perform uses a grid setup similar to what we 
-use for well-resolved binary simulations. With a single refined level, 
-we have approximately $5 \times 10^7$ zones spread over $\sim 1000$ grids.
-On a second refined level, there are $2 \times 10^9$ zones spread over 
-$\sim 16000$ grids. We run a scaling test for both cases, with the 
-highest processor count in each case chosen so that the number of 
-MPI tasks is similar to the number of grids. There are no relative gains 
-to be achieved from further parallelism. The results are found in 
-\ref{fig:strong_scaling}. We find excellent scaling for low numbers 
-of processors, though the long wall time per timestep required for the 
-smallest processor counts makes such a simulation infeasible in practice.
-The best tradeoff between reasonable wall time and parallel efficiency 
-seems to be when there are approximately 2--4 grids per processor. The 
-scaling behavior worsens at the highest processor counts, but this is 
-an expected consequence of processors becoming work-starved. In general 
-we find very good strong scaling behavior in the regime we are presently 
-interested in, simulating the early phases of a simulation at moderate 
-resolution. The strong scaling behavior is acceptable, though not perfect, 
-at very large processor counts when self-gravity is considered.
+\cite{castro}
 
@@
 %==========================================================================
-\section{Conclusions and Discussion}\label{sec:Conclusions and Discussion}
+\section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}
 
@@
 
-The authors thank Noel Scudder and Platon Karpov for their help with 
-this project, especially related to visualization of the results. 
-We also thank Adam Jacobs for his advice on and assistance with 
-running on supercomputing resources.
-
-This research was supported by NSF award AST-1211563. An
+This research was supported by NSF award AST-1211563.  This research
+used resources of the National Energy Research Scientific Computing
+Center, which is supported by the Office of Science of the
+U.S. Department of Energy under Contract No. DE-AC02-05CH11231.  An
 award of computer time was provided by the Innovative and Novel
@@
 the Office of Science of the Department of Energy under Contract
-DE-AC05-00OR22725. Projects AST006 and AST106 supported use of the ORNL/Titan resource. 
-This research is part of the Blue Waters sustained-petascale computing project, 
-which is supported by the National Science Foundation (awards OCI-0725070 
-and ACI-1238993) and the state of Illinois. Blue Waters is a joint 
-effort of the University of Illinois at Urbana-Champaign and its 
-National Center for Supercomputing Applications.
-This research used resources of the National Energy Research Scientific Computing
-Center, which is supported by the Office of Science of the
-U.S. Department of Energy under Contract No. DE-AC02-05CH11231.  
-This work used the Extreme Science and Engineering Discovery Environment (XSEDE), 
-which is supported by National Science Foundation grant number ACI-1053575. 
-Project AST100037 supported use of the resources NICS/Kraken and NICS/Darter.
+DE-AC05-00OR22725.
 
@@
 
-\section{CASTRO hydrodynamics changes}
-
-The basic PPM algorithm in CASTRO has undergone a number of changes
-since the original code paper \citep{castro}.  A discussion of the
-pure hydrodynamics changes along with verification of CASTRO when
-using the stellar equation of state was given in
-\citet{zingalekatz:2015}.  Here we discuss the changes that affect
-multispecies flow and source terms.
-
-Summary of changes:
-\begin{itemize}
-\item New gravity update type: fixes some energy conservation
-
-\item changing I's to use limit of the parabola instead of the
-  cell-center: not sure if this does much
-
-\item Colella \& Glaz Riemann solver: does a much better job with
-  temperature in strong shocks
-
-\item Tracing of gravity: seems to help with the material at the edge
-  of the star
-
-\item implicit update of the rotation terms?  (I coded this up once, but
-  it is not currently in the code)
-\end{itemize}
-
-\subsection{Reference States}
-
-For all the runs, the PPM reconstruction is done using the original limiters for
-the parabolic profiles \citep{ppm}.  The prediction of the interface
-states appears as:
-\begin{equation}
-\label{eq:ppmstatel}
-q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
-   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
-        \tilde{q}_L  - \mathcal{I}^{(\nu)}_+(q_i)
-       \right ] r_i^{(\nu)}
-\end{equation}
-where $q$ is the vector of primitive variables, $l_i^{(\nu)}$ and
-$r_i^{(\nu)}$ are the left and right eigenvectors with eigenvalue
-$\lambda_i^{(\nu)}$, with $\nu$ the index of the characteristic wave of
-the system.  The sum is over all the waves that result from the
-characteristic structure of the problem, but designed such that only
-waves moving toward the interface contribute to the interface value,
-$q_{i+1/2,L}^{n+1/2}$.  The reference state, $\tilde{q}_L$ is
-chosen to minimize the work of the characteristic projection.
-Finally, $\mathcal{I}_+^{(\nu)}(q)$ is the
-average under the parabolic profile of quantity $q$ of all the
-information that can reach the right interface of the zone $i$ as
-carried by the wave $\nu$.  The reader is referred to
-\citet{ppmunsplit} for further defaults.
-
-Since the original CASTRO paper, the reference state implementation
-has been switched to:
-\begin{equation}
-\label{eq:refchoice}
-\tilde{q}_L = \left \{ \begin{array}{cc}
-       \mathcal{I}_+^{(+)}(q_i) & \mathrm{if~} u + c > 0 \\
-       q_i                    & \mathrm{otherwise}
-\end{array}
-\right .
-\end{equation}
-where the $(+)$ superscript here means the fastest wave moving to the right
-(the $u+c$ eigenvalue).   This is simply the average under the largest
-portion of the parabolic profile that could possible reach the interface 
-over the timestep.  This is
-in agreement with \citet{ppmunsplit} (eq. 90).  The flattening
-in the original CASTRO paper has also been updated as discussed
-in \citet{zingalekatz:2015}.
-
-We comment on the choice of reference state for passively-advected
-quantities (like $X_k$ or the transverse velocity).
-First, consider one of the variables present in one-dimensional flow
-(density, velocity in the normal direction, and pressure).  Here we
-our reference state is as in Eq.~\ref{eq:refchoice} and we have.
-Ignoring flattening, if there are no waves moving toward our
-interface, then Eq.~\ref{eq:ppmstatel} reduces to:
-\begin{equation}
-q_{i+1/2,L}^{n+1/2} = \tilde{q}_L = q_i
-\end{equation}
-If instead only the fastest wave is moving toward the interface, then
-only the term corresponding to the fastest wave in the sum will be
-added in Eq.~\ref{eq:ppmstatel}, but our choice of reference state makes
-that term 0 by design, and our interface state is:
-\begin{equation}
-q_{i+1/2,L}^{n+1/2} = \tilde{q}_L = \mathcal{I}_+^{+}(q_i)
-\end{equation}
-This is the desired behavior for each of these cases.
-
-The details of the reference state for passively advected quantities 
-(which includes the transverse velocities in the 1-d tracing) is not
-typically discussed.  If we use the same idea of the reference state as
-in Eq.~\ref{eq:refchoice}, and consider a quantity $\xi$ which should only
-jump across the contact, then our interface state becomes:
-\begin{equation}
-\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L -
-  \underbrace{l_i^\evz \cdot \left [
-        \tilde{\xi}_L  - \mathcal{I}^\evz_+(\xi_i)
-       \right ] r_i^\evz}_{\text{only if~$u \ge 0$}}
-\end{equation}
-Again, ignoring flattening, if $u \ge 0$, then we have
-\begin{equation}
-\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L -
-  \left (\tilde{\xi}_L  - \mathcal{I}^\evz_+(\xi_i) \right ) = \mathcal{I}^\evz_+(\xi_i)
-\end{equation}
-(where we used the fact that the eigenvectors are normalized to 1 and
-don't mix in any other states when dealing with passive terms).  This
-is the expected behavior---we see a state that is traced only by the
-contact wave.  However, if $u < 0$ but $u + c \ge 0$, then we instead get:
-\begin{equation}
-\xi_{i+1/2,L}^{n+1/2} = \tilde{\xi}_L = \mathcal{I}^\evp_+(\xi_i)
-\end{equation}
-Here we used the same definition of the reference state and see that our
-interface state sees the profile traced under the fastest wave, not the
-contact.  This is not the correct behavior for a passively-advected
-quantity.
-
-The fix for passively-advected quantities is to simply ignore the 
-idea of a reference state and just test on the speed of the contact
-itself, setting:
-\begin{equation}
-\xi_{i+1/2,L}^{n+1/2} = \left \{ \begin{array}{cc}
-       \mathcal{I}_+^\evz(\xi_i) & \mathrm{if~} u  > 0 \\
-       \xi_i                    & \mathrm{otherwise}
-\end{array}
-\right .
-\end{equation}
-
-\subsection{Gravity and Rotation Tracing}
-
-We note a few additional differences between the original PPM
-implementation of \citet{ppm} and CASTRO.  In the original PPM
-implementation, the gravitational acceleration was reconstructed as a parabola, and
-this was traced under to find the forcing that affects the interface
-for each wave.  CASTRO followed \citet{ppmunsplit} which instead adds
-$(\Delta t/2)g$ to the interface states at the end of the
-reconstruction.  In the current implementation, we do the original
-parabolic reconstruction and characteristic tracing of the gravitation
-source.  Our system with the source appears as:
-\begin{equation}
-q_t + A(q) q_x = G
-\end{equation}
-where $G = (0, g, 0)^T$---i.e. the gravitational source only affects
-$u$, not $\rho$ or $p$.  Note that in the PPM paper, they put $G$ on 
-the lefthand side of the primitive variable equation, so our signs are
-opposite.  Our projections are now:
-\begin{equation}
-\sum_{\nu; \lambda^\enu \ge 0}l^\enu \cdot (\tilde{q} - \mathcal{I}^\enu_+(q) - \tfrac{\Delta t}{2} G) r^\enu
-\end{equation}
-for the left state, and
-\begin{equation}
-\sum_{\nu; \lambda^\enu \le 0} l^\enu \cdot (\tilde{q} - \mathcal{I}^\enu_-(q) - \tfrac{\Delta t}{2} G) r^\enu 
-\end{equation}
-for the right state.  Since $G$ is only non-zero for velocity, only
-the velocity changes.  Writing out the sum (and performing the vector products), we
-get:
-\begin{eqnarray}
-u_{i+1/2,L}^{n+1/2} =
-   \tilde{u}_+ 
-  &-& \frac{1}{2} \left [
-      \left (\tilde{u}_+ - \mathcal{I}_+^\evm(u) - \frac{\Delta t}{2} \mathcal{I}^\evm_+(g) \right ) - 
-       \frac{\tilde{p}_+ - \mathcal{I}_+^\evm(p)}{C} \right ] \nonumber \\
-  &-& \frac{1}{2} \left [
-      \left (\tilde{u}_+ - \mathcal{I}_+^\evp(u) - \frac{\Delta t}{2} \mathcal{I}^\evp_+(g) \right ) +
-       \frac{\tilde{p}_+ - \mathcal{I}_+^\evp(p)}{C} \right ]
-\end{eqnarray}
-These differ from the expression in the PPM paper, where $\Delta t G$,
-not $\Delta t/2 G$ is used in the projection, however we believe that
-the factor of $1/2$ is correct.  To see this, notice that if both
-waves are moving toward the interface, then the source term that is
-added to the interface state is $(\Delta t/4) (\mathcal{I}_+^\evm(g) +
-\mathcal{I}_+^\evp(g))$ for the left state, which reduces to $(\Delta
-t/2) g$ for constant g---this matches the result from Taylor
-expanding to the interface at the half-time (as in \citealt{ppmunsplit}).
-
-There is one additional effect of this change---now the gravitational
-source is seen by all Riemann solves (including the transverse solves)
-whereas previously it was only added to the final unsplit interface
-states.  Both methods are second-order accurate.
-
-The tracing of gravity can be controlled in the public version of CASTRO
-with the parameter {\tt castro.ppm\_trace\_grav}.
-
-Similarly, we also trace under the centrifugal and Coriolis forces to capture
-the effects of rotation on the edge states. For the primitive variables, the rotation
-source term is $F_R = (0, f_R, 0)^T$, where $f_R = -2\, {\bm{\omega}} \times \mathbf{u} - {\bm\omega}\times \left({\bm\omega} \times \mathbf{r}\right).$
-Just like the gravity source, the rotation source terms only update the velocities. Therefore 
-the rotation source terms can be included in a manner identical to that of the gravity source
-terms, with the replacement of $G$ with $F_R$ and $g$ with $f_R$ in the above. This is controlled in CASTRO
-with the parameter {\tt castro.ppm\_trace\_rot}.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=1.0]{plots/reference}
-  \caption{ \label{Fig:sod} Solution to Sod's problem with the original
-    reference state and the new reference state, as compared to the
-    exact solution.  We note that the new reference state shows a
-    slightly sharper shock and contact, but also has a dip at the tail
-    of the rarefaction.}
-\end{figure}
-
-\clearpage
-
 \end{document}

### 2015-05-26 papers/wdmerger_I/paper.tex
remove extra )
Add a directory for paper II
We were writing down the wrong formalism for the rotation--energy coupling; it is a static force and so it does not make sense to use the same approach we use for gravity, which dynamically responds to the movement of fluid on the grid.

@@
 \begin{align}
-  (\rho u)^{n+1} &= (\widetilde{\rho u}) + \omega \Delta t (\rho v)^{n+1}) \\
-  (\rho v)^{n+1} &= (\widetilde{\rho v}) - \omega \Delta t (\rho u)^{n+1})
+  (\rho u)^{n+1} &= (\widetilde{\rho u}) + \omega \Delta t (\rho v)^{n+1} \\
+  (\rho v)^{n+1} &= (\widetilde{\rho v}) - \omega \Delta t (\rho u)^{n+1}
 \end{align} 
@@
 with $\mathbf{f}_R \equiv  -{\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right)$. 
-This expression is identical to the gravity source under the interchange of $\mathbf{g}$ with $\mathbf{f}_R$.
+This expression is identical in form to the gravity source under the interchange of $\mathbf{g}$ with $\mathbf{f}_R$.
 As observed by \cite{marcello:2012}, we can similarly write down a rotational potential,
@@
 \end{equation}
-Therefore we can retrace the steps of Section \ref{sec:gravity_hydro_coupling} to obtain
-an update to $(\rho E)$ that conserves the total energy in the presence of rotation,
+So in the presence of rotation the total energy takes the form:
 \begin{equation}
@@
 \end{equation}
-In one dimension, this would be
+Since the rotation field $\Phi_R$ is static in time and does not change in response 
+to moving fluid, we do not need the formalism of Section \ref{sec:gravity_hydro_coupling}. 
+Instead, the time rate of change in the energy can be written as the flux of the quantity 
+$\rho \Phi_R$. In one dimension, this would be
 \begin{align}
-  \left.\Delta(\rho E_i)\right|_{\text{rot}} &= \frac{1}{2} \Delta m_{i-1/2} (\Phi_{R_{i}} - \Phi_{R_{i-1}}) \notag \\
-  &+ \frac{1}{2} \Delta m_{i+1/2} (\Phi_{R_{i+1}} - \Phi_{R_{i}}).
+  \left.\Delta(\rho E_i)\right|_{\text{rot}} &= \frac{1}{2} \Delta m_{i-1/2} (\Phi_{R_{i}} + \Phi_{R_{i-1}}) \notag \\
+  &+ \frac{1}{2} \Delta m_{i+1/2} (\Phi_{R_{i+1}} + \Phi_{R_{i}}).
 \end{align}
 As with gravity, in three dimensions we include the mass fluxes through all six cell faces.
+Note that whereas the gravity force explicitly changes the gas energy $E_g$, the rotation force 
+is manifestly conservative with respect to $E_g$ (except at domain boundaries), and only 
+redistributes existing quantities of $E_g$ between zones. This asymmetry between the gravity 
+and rotation forces may not be aesthetically pleasing but is necessary to avoid the problems 
+mentioned earlier that come from explicitly evolving the equation for $E_\text{tot}$.
 

### 2015-06-10 papers/wdmerger_I/paper.tex
Another rewrite of the energy description to match what we're doing in CASTRO now.

--- 2015-04-05 papers/wdmerger_I/paper.tex
@@
 incorporated by the use of a source term for the momentum and energy
-equations. In a finite volume methodology, the momentum source term appears in 
-integral form as
+equations. In a finite volume methodology, the momentum source term 
+as traditionally written appears in integral form as
 \begin{equation}
@@
 \end{equation}
-and for the energy source term it is
+and for the energy source term the analogue it is
 \begin{equation}
@@
 
-There are two ways that this source term enters the hydrodynamics update.
+There are two ways that these source terms enter the hydrodynamics update.
 First, during the hydrodynamics update, we alter the edge states that enter
-into the determination of the fluxes. To second order in space and time, 
+into the determination of the fluxes. (This only applies for the momentum source term;
+the gravitational force does not directly do work on the internal energy, which is used 
+to infer the pressure.) To second order in space and time, 
 this can be done using the cell-centered
 source term evaluated at time-level $n$. The details of this are described in 
-the appendix. Second, after the hydrodynamics step, we add the time-centered source term
+the appendix. Second, after the hydrodynamics step, we add the time-centered source terms
 to the state. First we will describe how we do this for the momentum,
@@
 in the momentum update can lead to large changes in a zone's momentum. By continuing to explicitly
-use the cell density in the momentum update, we can forgo this possibility: the size of the update
+use the cell density in the momentum update, we can avoid this possibility: the size of the update
 will always be suitably small if the zone's density is small. Thus for our simulations
@@
 
-Dealing with the energy equation is more difficult. The central
-challenge is to write down a form of the discretized energy equation
-that explicitly conserves total energy when coupled to gravity. When
-gravity is included, the total energy is
+For the energy equation, the central challenge is to write down a form of the 
+discretized energy equation that explicitly conserves total energy when 
+coupled to gravity. When gravity is included, the total energy is
 \begin{equation}
@@
 for simulations with self-gravity to prevent double-counting of
-interactions. The standard approach has been to note that the total energy update
-can be written down by taking the dot product of the velocity and the momentum source term,
-so that the time-centered energy update is
-\begin{equation}
-  \left.\Delta (\rho E)\right|_\text{grav} = \frac{\Delta t}{2} \left[ \rho^n \mathbf{u}^n \cdot \mathbf{g}^n + \rho^{n+1} \mathbf{u}^{n+1} \cdot \mathbf{g}^{n+1}\right].
-\end{equation}
-However, this does not conserve the total energy of Equation \ref{eq:total_energy_gravity}.
-In particular, we notice in our simulations a long-term secular change in the gravitational potential energy
-that is not accounted for in the gas energy. One solution to this problem has been offered by 
-\cite{jiang:2013} and \cite{marcello:2012}. They propose to evolve the total energy instead of the gas energy 
-in the hydrodynamics equations. When this is done, one can write down an explicitly conservative 
-formulation:
-\begin{equation}
-  \frac{\partial \rho E_{\text{tot}}}{\partial t} = -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u} + \mathbf{F_g}).
-\end{equation}
-Here $\mathbf{F_g}$ is a flux term which explicitly accounts for the change in the total energy due to self-gravity. 
-By using Equation \ref{eq:cell_center_gravity_source} and the continuity equation, it is straightforward to show that
-\begin{equation}
-  {\bm\nabla} \cdot \mathbf{F_g} = -\rho \mathbf{u} \cdot \mathbf{g} - \frac{\partial}{\partial t}\left(\frac{1}{2}\rho\Phi\right). \label{eq:gradient_F_g}
-\end{equation}
-From here one can use the continuity equation to explicitly calculate the flux $\mathbf{F_g}$, as in \cite{jiang:2013}:
-\begin{equation}
-  \mathbf{F_g} = \frac{1}{8\pi G}\left(\Phi\bm{\nabla}\dot{\Phi} - \dot{\Phi}\bm{\nabla}\Phi\right) + \rho\mathbf{u}\Phi \label{eq:jiang_flux}.
-\end{equation}
-Here the dots refer to time derivatives. This has some attractive features. 
-In particular, for a simulation with AMR, a flux-based approach
-can quite naturally be accommodated in a way that conserves energy across levels, when fluxes
-at interfaces between coarse and fine grids are matched. We note in passing that although
-\citeauthor{jiang:2013} claim that this approach requires a Poisson solve for the time derivative of the potential,
-$\dot{\Phi}$, this is not needed as in practice we will already have the potential at time-levels
-$n$ and $n+1$, and to second order in time, $\dot{\Phi}^{n+1/2} = (\Phi^{n+1} - \Phi^{n}) / \Delta t.$
-
-Now, the first term on the right-hand side of Equation \ref{eq:gradient_F_g} is the old source term. The second term simply 
-accounts for the time rate of change of the gravitational potential energy. But since
-\begin{equation}
-  \frac{\partial(\rho E_{\text{tot}})}{\partial t} = \frac{\partial (\rho E)}{\partial t} + \frac{\partial (1/2) \rho \Phi}{\partial t},
-\end{equation}
-this does not actually add any information, since we can subtract off the time derivative 
-of the potential energy and be back where we started. This demonstrates that the problem
-with the original approach lies in the numerical discretization; if this is done correctly,
-we can still properly conserve total energy.
-
-Before pursuing that, we wish to point out the reason why we prefer not to use this 
-approach of solving a total energy equation. The expression for the flux $\mathbf{F_g}$
-given in \cite{jiang:2013} includes terms that are proportional to $\Phi$ and ${\bm{\nabla}} \Phi$,
-but that are not explicitly proportional to the zone density $\rho$. Thus the approach suffers
-from the same problems as the conservative momentum formulation discussed above. Truncation errors
-in the gravitational field will result in an approximation to the density that can be quite 
-different from the actual density, especially in low-density regions. In practice,
-we have found this to be such a serious problem that this approach is not feasible for our problem.
-The formulation of \cite{marcello:2012} does not suffer from the same problem. However, a 
-remaining problem is that in this formalism, the internal energy
-is calculated by subtracting the potential energy and the kinetic energy from the 
-total energy. Due to numerical discretization errors, however, this can result in a serious 
-error in the internal energy. This problem can be eased by the dual energy formalism,
-but not completely. This means that we will potentially have very large errors in a 
-zone's temperature, which can have serious repercussions in the reacting flows we will eventually consider,
-as many reaction rates are very sensitive to temperature. Therefore our goal is instead to consider an approach
-that is explicitly \textit{non}-conservative for the gas energy $(\rho E)$, but which
-instead conserves the total energy $(\rho E_{\text{tot}})$ over the computational domain 
-(subject, of course, to small errors due to energy leaking from the domain boundary).
-
-The question of how to do this was resolved by \cite{arepo}. 
-The essential problem with the standard approach is that 
-fluid movements in a grid code are calculated using
-fluxes at zone boundaries, while the standard gravity source term is
-an estimate of the flux using the zone center quantities. This can
-lead to a striking mismatch when the gradient across a zone is
-large. We have found that this is especially important at the edge of
-a star, where the large density gradient means that the momentum and
-energy fluxes can transition sharply even across half of a zone
-width. Another problem is that in a merger calculation in which two
-stars are moving on a grid, large quantities of material are being
-advected, so even a slight error in the calculation for the
-high-density material can manifest itself as a long-term error in
-energy conservation. This becomes apparent in Section
-\ref{sec:kepler}.
-
-We start with the integral formulation of the energy source term (Equation
-\ref{eq:cell_center_gravity_source}), but do not discretize it at
-the cell center.  Instead, we recognize that in the limit that the velocity is 
-constant, $\mathbf{u} = {\bm\nabla}\left[\mathbf{r}\cdot\mathbf{u}\right]$,
-where $\mathbf{r}$ is the position vector originating from the zone center.
-Then, since in the cell-centered approach we assume that the density, velocity,
-and gravitational acceleration are constant across the zone, we can include all terms
-inside the gradient operator. This allows us to use the divergence theorem to evaluate
-the quantities on the zone faces. To be second-order accurate in time, the fields must 
-all be evaluated at time-level $n+1/2$, but for clarity we will neglect the superscript.
-\begin{align}
-  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{\Delta t}{\Delta V} \int \rho \mathbf{u}\cdot\mathbf{g}\, dV_i \notag \\
-     &= \frac{\Delta t}{\Delta V} \int \rho\, {\bm \nabla} \left[\mathbf{r} \cdot \mathbf{u}\right]\cdot\mathbf{g}\, dV_i \notag \\
-     &= \frac{\Delta t}{\Delta V} \int {\bm \nabla} \cdot \left[ \rho\, (\mathbf{r} \cdot \mathbf{u})\, \mathbf{g}\right]\, dV_i \notag \\
-     &= \frac{\Delta t}{\Delta V} \int \rho\, (\mathbf{r} \cdot \mathbf{u})\, \mathbf{g} \cdot \mathbf{dS}_i.
-\end{align}
-Here $\mathbf{dS}$ is a surface element whose vector points normal to
-and away from the surface. In one dimension, this expression evaluates to
-\begin{align}
-  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{\Delta t}{\Delta V} \rho_{i-1/2}\, \frac{\Delta x}{2} u_{x_{i-1/2}} g_{x_{i-1/2}} \Delta A_{i-1/2} \notag \\
-             &+ \frac{\Delta t}{\Delta V} \rho_{i+1/2}\, \frac{\Delta x}{2} u_{x_{i+1/2}} g_{x_{i+1/2}} \Delta A_{i+1/2} \notag \\
-             &= \frac{1}{2} \rho_{i-1/2} u_{x_{i-1/2}} g_{x_{i-1/2}} \notag \\
-             &+ \frac{1}{2} \rho_{i+1/2} u_{x_{i+1/2}} g_{x_{i+1/2}}.\label{eq:edge_centered_gravity_source}
-\end{align}
-(We note that this differs slightly from Equation 94 in \cite{arepo} in that
-we evaluate the gravitational acceleration at the zone face rather than the zone center.
-We are free to do this since we assumed a uniform gravitational field to begin with.)
-In the limit of constant fields, this expression simply averages out to the standard 
-source term at the zone center. However, when the fields vary this can give very different results.
-
-As pointed out by \citeauthor{arepo}, $(\rho u)$ evaluated at the faces at time level $n+1/2$ is the mass flux
-that is obtained from the Riemann solver in the hydrodynamics step,
-so $(\rho u)_{i-1/2}\, \Delta t\, \Delta A = \Delta m_{i-1/2}$ is the mass exchanged between zones $i$ and $i-1$.
-Additionally, to second order $g_{x_{i-1/2}} \Delta x = \Phi_{i} - \Phi_{i-1}$. Therefore the energy gained by the zone is
-\begin{align}
-  \left.\Delta\left(\rho E\right)_i\right|_{\text{grav}} &= \frac{1}{2} \Delta m_{i-1/2} \left(\Phi_{i} - \Phi_{i-1}\right) \notag \\
-     &+ \frac{1}{2} \Delta m_{i+1/2} \left(\Phi_{i+1} - \Phi_{i}\right)
-\end{align}
-This expression says that the energy gained by a zone is due to a mass element moving through a potential gradient
-between two zones. Now, the key observation is that this does not have the telescoping property of a conservative flux;
-instead, the \textit{same} amount of gravitational potential energy is gained by zones $i$ and $i-1$ by the mass flux 
-between them. When this source term is summed up over all zones, it will exactly compensate for the change in 
-gravitational potential energy, except for any energy lost through the boundaries. The extension of this energy change 
-to multiple dimensions simply means that we include the mass fluxes and potential gradients across the $y$ and $z$ faces.
-We have found that in practice this method for conserving the total energy is the least susceptible to deleterious effects
-from numerical discretization errors. However, it does mean that there can be sharp changes in the total energy of a zone
-in regions near a white dwarf's surface, because the mass fluxes into the ambient medium can be quite large compared to
-the mass of the zone. This effect diminishes at higher resolutions as the gradients become better resolved.
-
-Our strategy for coupling the gravity to the hydrodynamics on a uniform grid can thus be summarized as follows. 
-At the beginning of a timestep, we solve the Poisson equation given the density $\rho^n$. We then solve the 
-Euler equations and advance the system in time by $\Delta t$ using zone-centered fluxes at time-level $n+1/2$. 
-We add $\Delta t \rho^n \mathbf{g}^n$ to the momentum equations. We then solve the Poisson equation again, given
-the density $\rho^{n+1}$. We then add $-(\Delta t/2) \rho^{n} \mathbf{g}^{n} + (\Delta t / 2) \rho^{n+1} \mathbf{g}^{n+1}$
-to the momentum equations, time-centering the source term. We simultaneously add $\sum_{j} -(1/2) \Delta m_{j} (\Phi - \Phi_j)$
-to the zone's energy, where the index $j$ runs over all adjacent zones. This conserves the total energy on the grid
-to roundoff error, modulo any losses from an open domain boundary.
+interactions.
+
+Historically many simulation codes with gravity have not used a conservative formulation 
+of the energy equation, but it is straightforward to do so. In particular, conservation of 
+total energy requires only that a change in gravitational energy is compensated for by a 
+change in gas energy, and that energy changes due to mass transfer are explicitly and 
+exactly tracked. At each zone face in a given timestep, there is some transfer of mass $\Delta \rho$ between 
+the zones adjacent to that face, which is governed by the continuity equation. Suppose the 
+transfer of mass is positive to the right, and the face is located at the (one-dimensional)
+index $i-1/2$. Energy is transferred because an amount of mass $\Delta \rho_{i-1/2}$ moves from 
+potential $\Phi_{i-1}$ to potential $\Phi_{i}$; if these potentials are different, it requires some 
+work exerted by or on the zone to push the mass to this different potential energy. Therefore 
+the net change of gravitational energy is $\left[(1/2)\Delta \rho_{i-1/2} (\Phi_{i} - \Phi_{i-1})\right]$,
+and if total energy is conserved, then the zone must have a change in gas energy 
+$(\rho E)$ that is equal in magnitude and opposite in sign to the change in potential energy. 
+Thus the update in a zone for the total energy that conserves $(\rho E_{\text{tot}})$ is:
+\begin{equation}
+  \Delta (\rho E) = -\frac{1}{2}\sum_{f} \Delta \rho_{f} (\Phi_{f+1/2} - \Phi_{f-1/2}),\label{eq:grav_energy_conservation_update}
+\end{equation}
+where the sum is over the cell faces $f$ and the indices $f+1/2$ and $f-1/2$ refer to 
+the zone centers immediately to the left and right in the direction perpendicular to the face.
+Note that while total energy is conserved, the gas energy is not: this does not have the 
+telescoping property of a flux formalism where energy gained by one zone is lost by another. 
+The energy is being exchanged with a potential field that responds indirectly when the mass moves.
+
+Now the change in the mass $\Delta \rho_{f}$ is proportional to the mass flux through the surface with area $\Delta A$:
+\begin{equation}
+  \Delta \rho_f = \frac{\Delta t}{\Delta V} (\Delta A\, \rho v_{f}),\label{eq:mass_transfer}
+\end{equation}
+So as long as we record these fluxes through the zone faces after coming out of the hydrodynamics step, 
+we are able to conserve the total energy completely (except for any energy loss or gain through 
+physical domain boundaries). In order for the method to be second-order accurate in time, 
+no change is needed to the fluxes since those are already time-centered 
+in the hydrodynamics update, so we need only to use a time-centered $\Phi$ (which can be computed 
+by averaging the time-level $n$ and $n+1$ potentials; we already have the latter because CASTRO 
+re-computes the potential at the new time after the hydrodynamics step, and we can apply this energy 
+at the end of the timestep).
+
+As observed by \cite{arepo}, when examining Equation \ref{eq:mass_transfer} one can see the relationship 
+between this method and the more common (non-conservative) approach of evaluating the change in 
+gas energy using the work done $(\mathbf{v} \cdot \rho \mathbf{g})$ by the gravitational force at the 
+cell center. Analytically this form expresses the same core idea via the work-energy theorem, and its 
+major flaw is that it evaluates the energy change at the cell center when in fact the mass transfer is 
+happening at the cell edges, and the energy change is actually partially compensated by the potential 
+in the adjacent zone that the mass is moving to. This is important enough that for simulations where the 
+potential can change substantially over the course of the evolution, massive amounts of energy can be lost. 
+This standard approach is therefore unacceptable in the case of a problem like white dwarf mergers, and 
+the fix to this energy leaking -- evaluating the energy transfer at the six zone faces instead of 
+the single zone center -- has a very minor cost in terms of code complexity and computational time.
+
+It is worth pointing out that while Equation \ref{eq:grav_energy_conservation_update} was correctly obtained 
+by \cite{arepo} (and indeed that is the inspiration for this approach), there is a subtle error in that derivation. 
+Their argument was that if we integrate the work done over the volume of the zone, this volume integral can be 
+converted into a surface integral by use of the divergence theorem (see their Equation 93). Then we can simply 
+evaluate the work done at each zone face. However, this application of the divergence theorem appears to be 
+mathematically incorrect, despite the fact that this is very similar in spirit to the approach discussed here. 
+The author fortuitously obtains the correct factor of $1/2$ corresponding to the change in gravitational self-energy.
+However, some reflection reveals that there should be no way to derive this double-counting factor simply by integrating 
+the work done: that factor is derived from the unique nature of the gravitational potential, which is generated 
+by the same distribution that it exerts a force on for a simulation with self-gravity, and it has to be added in 
+intentionally to ensure conservation of total energy for these particular types of potentials. We can apply 
+the approach of considering energy transfer between the gas energy and the potential energy to any potential field, 
+whether or not it is static, changing in time, or generated by the fluid distribution itself. 
+Indeed, we will do so for the rotational potential in Section \ref{sec:rotation}, where the 
+factor of $1/2$ does not appear. However, it is important also to note that this is not related to a 
+global flow of energy or the global nature of the gravitational potential, as claimed by \cite{marcello:2012}. 
+This double-counting factor would apply even if there were only two point masses on the grid, since 
+work would only be required to bring in the second mass from infinity; the first one can be 
+placed at no energy cost. This transfer of energy is explicitly \textit{local} because the 
+mass transfer is local. It is true that the mass transfer changes the potential globally, 
+but that is implicitly accounted for when using the time-centered potential, and no 
+additional steps need to be taken.
+
+Another approach recently taken in the literature for conserving total energy has been to evolve an 
+equation for the total energy $(\rho E_{\text{tot}})$; see \cite{marcello:2012} and \cite{jiang:2013}.
+This can be done by adding a term corresponding to the ``flux'' of gravitational potential energy. We 
+avoid this approach in general because there will be regions on the computational domain where the total 
+energy is dominated by potential energy (especially the low-density regions near the edge of the white dwarfs),
+and the gas energy can only be retrieved by first subtracting $-\rho \Phi/2$ from the total energy. Like 
+\cite{arepo}, we find that this can result in some serious errors due to numerical discretization, yielding 
+unphysical energies or temperatures. We observe also that the specific implementation of \cite{jiang:2013} 
+has terms in their gravitational flux that are not proportional to $\rho$, and so it can lead to the 
+same troubles that plague the tensor-based formalism for the momentum equation, where small errors 
+in the discretization of the gravitational potential can lead to very large changes in the energy of the gas.
+(We note in passing too that although \citeauthor{jiang:2013} claim their approach requires a Poisson 
+solve for the time derivative of the potential, $\dot{\Phi}$, this is not needed as in practice 
+we will already have the potential at time-levels $n$ and $n+1$, and to second order in time, 
+$\dot{\Phi}^{n+1/2} = (\Phi^{n+1} - \Phi^{n}) / \Delta t.$)
 
@@
 \begin{equation}
-  \left.\frac{\partial(\rho E)}{\partial t}\right|_{\text{rot}} = \frac{1}{\Delta V}\int \rho \mathbf{u} \cdot \mathbf{f}_R\, dV,
-\end{equation}
-with $\mathbf{f}_R \equiv  -{\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right)$. 
-This expression is identical to the gravity source under the interchange of $\mathbf{g}$ with $\mathbf{f}_R$.
+  \left.\frac{\partial(\rho E)}{\partial t}\right|_{\text{rot}} = \frac{1}{\Delta V}\int \rho \mathbf{u} \cdot \mathbf{f}^R\, dV,
+\end{equation}
+with $\mathbf{f}^R \equiv  -{\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right)$. 
+This expression is identical in form to the gravity source under the interchange of $\mathbf{g}$ with $\mathbf{f}^R$.
 As observed by \cite{marcello:2012}, we can similarly write down a rotational potential,
 \begin{equation}
-  \Phi_R = -\frac{1}{2} \left| {\bm\omega} \times \mathbf{r} \right|^2.
-\end{equation}
-Therefore we can retrace the steps of Section \ref{sec:gravity_hydro_coupling} to obtain
-an update to $(\rho E)$ that conserves the total energy in the presence of rotation,
-\begin{equation}
-  (\rho E_{\text{tot}}) = (\rho E) + \frac{1}{2} \rho \Phi + \rho \Phi_R.
-\end{equation}
-In one dimension, this would be
+  \Phi^R = -\frac{1}{2} \left| {\bm\omega} \times \mathbf{r} \right|^2.
+\end{equation}
+So in the presence of rotation the total energy takes the form:
+\begin{equation}
+  (\rho E_{\text{tot}}) = (\rho E) + \frac{1}{2} \rho \Phi + \rho \Phi^R.
+\end{equation}
+Given that we can write down a potential energy for the rotation field, then we can use the machinery of 
+Section \ref{sec:gravity_hydro_coupling}. We again continue to evolve explicitly an equation for 
+the gas energy, and allow it to change in response to work done by or on the rotational potential.
 \begin{align}
-  \left.\Delta(\rho E_i)\right|_{\text{rot}} &= \frac{1}{2} \Delta m_{i-1/2} (\Phi_{R_{i}} - \Phi_{R_{i-1}}) \notag \\
-  &+ \frac{1}{2} \Delta m_{i+1/2} (\Phi_{R_{i+1}} - \Phi_{R_{i}}).
+  \left.\Delta(\rho E)\right|_{\text{rot}} &= - \sum_{f} \Delta \rho_{f} (\Phi^R_{f+1/2} - \Phi^R_{f-1/2})
 \end{align}
-As with gravity, in three dimensions we include the mass fluxes through all six cell faces.
+Again we observe that this results a change in the global integral of $(\rho E)$, but keeps 
+the global integral of $(\rho E_{\text{tot}})$ conserved (except at physical domain boundaries).
 

### 2015-07-09 papers/wdmerger_I/paper.tex
Add a description of our gravitational wave signal diagnostics to Paper I.

@@
 
+A final diagnostic quantity we consider is the gravitational wave emission by the 
+binary system. White dwarfs are not strongly affected by general relativistic effects;
+the orbital motions are much slower than the speed of light, and the relativity parameter 
+$GM / c^2 R$, which measures the ratio of the Schwarzschild radius of a mass $M$ to the actual 
+radius $R$ of the object, is much less than unity for a white dwarf. It is justified to 
+neglect any relativistic effects in computing the dynamical evolution of the system. However,
+the system emits gravitational waves during its evolution; this energy loss is what drives 
+the initial inspiral over very long timescales, and occurs at a greater rate during 
+the coalescence phase, though then being inconsequential on a dynamical timescale. The 
+frequency of the gravitational waves will be similar in size to the frequency of the 
+orbital motion, which is in the range 10-100 mHz for our problem. This is well outside 
+the range of currently existing gravitational wave detectors but is very well suited 
+for proposed space-based detectors such as eLISA.
+
+We follow the prescription of \citet{blanchet:1990}. At distances far from the 
+gravitational wave source, we can consider the leading term in the gravitational 
+wave signal:
+\begin{equation}
+  h^{TT}_{ij}(t,\mathbf{x}) = \frac{2G}{c^4 r}P_{ijkl}(\mathbf{n}) \ddot{Q}_{kl}(t - r/c).
+\end{equation}
+$h$ is the perturbation to the spacetime metric and is commonly called the \textit{strain}; 
+for laser interferometers, it measures the relative change in the distance between mirrors. 
+The ``TT'' superscript indicates that we work in the commonly used tranverse-traceless gauge.
+This strain is measured at time $t$ and position $\mathbf{x}$ relative to the binary system.
+$r\equiv |\mathbf{x}|$ is the distance from the observer to the binary system. The unit vector 
+$\mathbf{n} \equiv \mathbf{x} / r$ then measures the direction of the outgoing wave with 
+respect to the observer, and $P_{ikjl}(\mathbf{n})$ is an operator in the transverse-traceless 
+gauge that projects a tensor onto the direction orthogonal to $\mathbf{n}$. $Q_{kl}$ is the 
+quadrupole moment tensor:
+\begin{equation}
+  Q_{kl} = \int dV \rho \left(x^k x^l - \frac{1}{3}\delta^{kl} \mathbf{x}^2\right).
+\end{equation}
+The argument $(t - r/c)$ indicates that to get the strain at time $t$ we evaluate the second derivative of the 
+quadrupole moment at the retarded time $t - r/c$. In practice we ignore this in calculating this strain,
+instead using the current simulation time, and so this merely yields a time delay of order $r/c$ between 
+when the waves are emitted and when we observe them.
+
+Therefore the primary component of the calculation is the evaluation of the second time derivative of $Q_{kl}$.
+Explicitly constructing a discretized form of this derivative, using the current state and the state at 
+previous times, is undesirable because of the inherent imprecision (its accuracy depends on the size of the timestep),
+in addition to the logistical challenges that may be implied by saving and using previous simulation states. 
+\citet{blanchet:1990} provide a prescription for this time derivative purely in terms of the state at a given time:
+\begin{equation}
+  \ddot{Q}_{kl} = \text{STF}\left\{2\int dV \rho (v^k v^l + x^k g^l)\right\}.
+\end{equation}
+The symmetric trace-free (STF) operator is defined as:
+\begin{equation}
+  \text{STF}\left\{A^{ij}\right\} = \frac{1}{2}A^{ij} + \frac{1}{2}A^{ji} - \frac{1}{3} \delta^{ij} \sum_{k}A^{kk}.
+\end{equation}
+
+The strategy is then as follows. At the end of the coarse timestep, we first calculate $\ddot{Q}_{kl}$
+using an integral over the domain (using refined grids as usual by masking out the contributions from 
+coarse grids underlying more accurate fine grids). This quantity is independent of the observer. Then, 
+we pick an observing location $\mathbf{x}$ relative to the domain, evaluate the projection operator, 
+and then perform the relevant tensor contraction to determine the strain tensor. We can 
+repeat this process for any number of observing locations at minimal cost, since the quadruple tensor 
+only needs to be calculated once. Now, gravitational waves only excite modes orthogonal to their 
+direction of travel. These are the ``plus'' and ``cross'' modes, $h_+$ and $h_\times$, named after 
+the types of spatial distortions they exhibit. We calculate the signal at a distance $r$ along 
+the $x$, $y$ and $z$ axes. For the latter, as an example, $h_{+} = h_{11} = -h_{22}$ and 
+$h_{\times} = h_{12} = h_{21}$. All other entries vanish. By default we take $r = 10$ kpc; 
+as shown by \citet{loren-aguilar:2005}, this is a typical distance scale over which an 
+experiment such as LISA could detect a coalescing binary white dwarf system. 
+However, the strain at any other distance is easily calculated and goes as the inverse of the distance.
+
 The computational domain has a total size of $1.024 \times 10^{10}\ \text{cm}$ in each 
@@
 We also thank Adam Jacobs for his advice on and assistance with 
-running on supercomputing resources.
+running on supercomputing resources. Special thanks are given 
+to the organizers of the 2015 Caltech Gravitational Wave 
+Astrophysics School, and their supporters: the NSF 
+under CAREER award PHY-1151197, the Sherman Fairchild 
+Foundation, the LIGO Laboratory at Caltech, and Caltech.
 

### 2015-07-12 documentation/userguide.tex
Fill in some more probin parameters in the user's guide.

--- 2015-03-22 documentation/userguide.tex
@@
 
-Welcome to the user's guide for the wdmerger software package. This software is designed to
+Welcome to the user's guide for the \texttt{wdmerger} software package. This software is designed to
 simulate binary white dwarf systems, and is intended to provide useful information on the 
 viability of mergers of white dwarfs as a progenitor for Type Ia supernovae. It is a set of 
-source files designed to use the CASTRO code \citep{castro}.
+source files designed for use with the \texttt{CASTRO} code \citep{castro}.
 
 This software requires a few other software packages to be installed before it can be used.
-The science software includes \href{https://ccse.lbl.gov/BoxLib/}{BoxLib} and 
-\href{https://ccse.lbl.gov/Downloads/downloadCASTRO.html}{CASTRO}. This software is written 
+The science software includes \href{http://boxlib-codes.github.io/}{BoxLib} and 
+\href{http://boxlib-codes.github.io/Castro/}{CASTRO}. You should also 
+download the \href{https://github.com/BoxLib-Codes/Microphysics}{Microphysics} repository, 
+which stores the equation of state and nuclear network we will use. This software is written 
 in C++ and Fortran, and compilers for each are needed. The gcc compilers are sufficient.
@@
 If python and yt are not installed on your system, you will still be able to compile and 
-run this code, but you will need some other mechanism for analyzing the BoxLib output data.
+run this code, but you will need some other mechanism for analyzing the \texttt{BoxLib} output data.
 We have some routines for this purpose available in Fortran and may be able to provide them
@@
 \begin{itemize}
-  \item \texttt{BOXLIB\_HOME}, the location of the BoxLib code's top-level directory
-  \item \texttt{CASTRO\_DIR}, the location of the CASTRO code's top-level directory
+  \item \texttt{BOXLIB\_HOME}, the location of the \texttt{BoxLib} code's top-level directory
+  \item \texttt{CASTRO\_DIR}, the location of the \texttt{CASTRO} code's top-level directory
   \item \texttt{WDMERGER\_HOME}, the location of this software's top-level directory
+  \item \texttt{MICROPHYSICS\_DIR}, the location of the \texttt{Microphysics} code's top-level directory
 \end{itemize}
@@
 
-The \texttt{source} directory contains all of the Fortran routines needed to build a CASTRO 
-executable file. It is not normally built directly, but you can type \texttt{make} in that directory
-to build a vanilla executable with the default settings.
+The \texttt{source} directory contains all of the Fortran routines needed to build a \texttt{CASTRO} 
+executable file. You can type \texttt{make} in this directory to build a vanilla executable 
+with the default settings, though usually we will build the software from a different location.
 
 The \texttt{analysis} directory contains the Python analysis scripts used for analyzing output
-data (both in BoxLib and plaintext format).
+data (both in \texttt{BoxLib} and plain-text format).
 
 The \texttt{tests} directory includes problem setups that are used for testing 
-the hydrodynamics code CASTRO and exercises a wide range of its capabilities.
+the hydrodynamics code \texttt{CASTRO} and exercises a wide range of its capabilities.
 
@@
 and other I/O information, and the analysis routines. A full description of the physics 
-and the associated numerical algorithms is included in the CASTRO code paper \cite{castro}.
+and the associated numerical algorithms is included in the \texttt{CASTRO} code paper \cite{castro}.
 
@@
 
-CASTRO includes full self-gravity with the Poisson equation, as well as the ability to perform a
+\texttt{CASTRO} includes full self-gravity with the Poisson equation, as well as the ability to perform a
 simulation in a rotating reference frame with fixed rotation period. By default the $z$ axis will 
@@
 
-At present nuclear reactions are not enabled, but they will be soon.
+For the nuclear network we use the \texttt{aprox13} network provided by Frank Timmes, which 
+we have ported for use in the \texttt{Microphysics} repository. It is an alpha-chain network 
+with isotopes ranging from ${}^4$He to ${}^{56}$Ni.
 
@@
 
-There are a few CASTRO input parameters that you will typically want to control when running problems. 
-Among the most major is whether we are doing rotation (\texttt{castro.do\_rotation}) and if so, 
-what the rotational period is (\texttt{castro.rotational\_period}). You'll also want to change the 
-\texttt{stop\_time}, which is when the simulation will terminate. You should also avoid flooding 
-yourself with plotfiles and checkpoints by setting \texttt{amr.plot\_files\_output} and 
+There are a few \texttt{CASTRO} input parameters that you will typically want to control when running problems. 
+Among the most important is whether we are doing rotation (\texttt{castro.do\_rotation}) and if so, 
+what the rotational period is (\texttt{castro.rotational\_period}). (Note that the rotational period 
+is used to set the orbital properties even if we're not rotating, so you should always set this appropriately.)
+You'll also want to change the \texttt{stop\_time}, which is when the simulation will terminate. 
+You should also avoid flooding yourself with plotfiles and checkpoints by setting \texttt{amr.plot\_files\_output} and 
 \texttt{amr.checkpoint\_files\_output}, respectively; these are the simulation time differences
-between successive plotfiles and checkpoints. Note that CASTRO does not adjust the timestep 
+between successive plotfiles and checkpoints. Note that \texttt{CASTRO} does not adjust the timestep 
 to meet this criterion, so it will generate these at the first timestep that passes this interval.
@@
 In the following list we describe the parameters; you can see their current defaults by 
-examining \texttt{source/probin}.
+examining \texttt{source/probin}. (Some parameters in there are currently beta quality and 
+not recommended for use, and for a couple others there will be no practical reason to change. 
+Stick to the ones in this list for now.)
 \begin{itemize}
@@
     center of the domain)
+  \item \texttt{central\_density\_P}: alternatively, we can specify the central density of the primary white dwarf instead of the mass (real)
+  \item \texttt{central\_density\_S}: alternatively, we can specify the central density of the secondary white dwarf instead of the mass (real)
   \item \texttt{nsub}: number of sub-intervals to divide a zone into when interpolating from the initial model (integer)
@@
     Keplerian rotation rate (logical; this only makes sense in the inertial reference frame with \texttt{castro.do\_rotation = 0})
+  \item \texttt{collision}: whether to set the stars moving towards each other at free-fall speed for a collision (logical)
+  \item \texttt{collision\_separation}: if we're doing a collision, number of WD radii to set the stars apart from each other initially (real)
   \item \texttt{damping}: whether or not to exert a constant damping force on the orbit (logical)
@@
     if false, we will interpolate based on pressure)
+  \item \texttt{ambient\_density}: density of the rarefied ambient medium outside the stars (real)
+  \item \texttt{stellar\_temp}: initial temperature of the stars (and also ambient medium) (real)
+  \item \texttt{stellar\_C12, stellar\_O16}: mass fractions of ${}^{12}$C and ${}^{16}$O in the stars (real)
+  \item \texttt{star\_axis}: what coordinate axis the stars will be lined up along initially (integer; $x$ = 1, $y$ = 2, $z$ = 3)
   \item \texttt{maxTaggingRadius}: fractional size of the domain outside of which we won't perform any refinement (real)
+  \item \texttt{sponge\_dist}: we can apply a sponge that damps the velocities starting at this fraction multiplied by the 
+    distance from the center to the domain boundary (real)
+  \item \texttt{sponge\_width}: smoothing width of the sponge; the sponge is fully applied at {\tt sponge\_dist} + {\tt sponge\_width},
+    in the same coordinates (real)
+  \item \texttt{sponge\_timescale}: the sponge decreases the velocity by a factor $1 / (1 + (\Delta t / t_s)$, where 
+    $t_s$ is the {\tt sponge\_timescale} in seconds (real)
+  \item \texttt{gw\_dist}: for the gravitational wave signals we calculate, this is the distance in kiloparsecs from the observer to the binary system (real)
+  \item \texttt{fill\_ambient\_bc}: whether we should set ghost cells on the domain boundary to have the ambient state properties (logical)
 \end{itemize}
 
-The \texttt{tagging} namelist is described in the AMR chapter of the CASTRO user's guide. At present, we 
+The \texttt{tagging} namelist is described in the AMR chapter of the \texttt{CASTRO} user's guide. At present, we 
 only tag based on absolute values of the density and temperature.
@@
 The main makefile sets all of our build options (compiler, number of space dimensions, 
-equation of state, etc.) and then includes the CASTRO makefile, which then runs as normal.
+equation of state, etc.) and then includes the \texttt{CASTRO} makefile, which then runs as normal.
 If you want to adjust the build options, then simply add a line before this include. As an 
@@
 and \texttt{probin} files from the root \texttt{wdmerger} source directory; these are the runtime
-files CASTRO needs to determine what your problem setup will be. They contain the default 
+files \texttt{CASTRO} needs to determine what your problem setup will be. They contain the default 
 parameters for merger runs. It also copies any auxiliary files one might need, like the 
-table for the equation of state, as well as the actual CASTRO executable you just built in the 
+table for the equation of state, as well as the actual \texttt{CASTRO} executable you just built in the 
 \texttt{compile/} directory. Finally, it builds a script that you'll use for actually submitting 
@@
 we simply need to update a variable named \texttt{castro\_rotational\_period}. 
-Note that we use an underscore (\texttt{castro_}) and not a period (\texttt{castro.}) after the namespace.
+Note that we use an underscore (\texttt{castro\_}) and not a period (\texttt{castro.}) after the namespace.
 This is simply because \texttt{bash} does not permit periods in variable names; but the helper routine will 
@@
 The \texttt{tests} directory contains a number of example problems that we use
-to assess the reliability of CASTRO in performing hydrodynamics simulations 
+to assess the reliability of \texttt{CASTRO} in performing hydrodynamics simulations 
 with self-gravity. This section includes a list of the existing problems 
@@
 and michael.zingale@stonybrook.edu, respectively. Alan Calder and Doug Swesty at Stony Brook are 
-also co-investigators on this project. The CASTRO code was developed primarily at the 
+also co-investigators on this project. The \texttt{CASTRO} code was developed primarily at the 
 Lawrence Berkeley National Laboratory. Staff there that have contributed directly and 

### 2015-07-13 tests/hydro/evrard_collapse/run_test.sh
We have moved the Evrard collapse problem over to Castro, so from here we only compile that code instead of retaining a duplicate of the source code.

@@ -1,37 +1,29 @@
+TEST_DIR=$CASTRO_DIR/Exec/evrard_collapse
+
+cp $TEST_DIR/inputs source/
+cp $TEST_DIR/probin source/
+
 source $WDMERGER_HOME/job_scripts/run_utils.sh
 
-# Problem-specific variables
-
-stop_time=15.0
-
-geometry_prob_lo=" -2.56e9 -2.56e9 -2.56e9"
-geometry_prob_hi="  2.56e9  2.56e9  2.56e9"
-
-castro_do_rotation=0
-
-# Loop over the resolutions in question
+# Loop over the gravity source options
 
 for gs in 1 2 3 4
 do
-  for ncell in 64 128
-  do
-    dir=$results_dir/gs$gs/n$ncell
+  dir=$results_dir/gs$gs
 
-    castro_grav_source_type=$gs
-    amr_n_cell="$ncell $ncell $ncell"
+  castro_grav_source_type=$gs
 
-    if [ $MACHINE == "BLUE_WATERS" ]; then
+  if [ $MACHINE == "BLUE_WATERS" ]; then
 
-      if [ $ncell -eq 64 ]; then
-	  nprocs=32
-	  walltime=1:00:00
-      elif [ $ncell -eq 128 ]; then
-	  nprocs=128
-	  walltime=2:00:00
-      fi
-
+    if [ $ncell -eq 64 ]; then
+      nprocs=32
+      walltime=1:00:00
+    elif [ $ncell -eq 128 ]; then
+      nprocs=128
+      walltime=2:00:00
     fi
 
-    run $dir $nprocs $walltime
-  done
+  fi
+
+  run $dir $nprocs $walltime
 done

### 2015-07-13 tests/hydro/evrard_collapse/source/Prob_3d.f90

@@ -1,266 +1,0 @@
-   subroutine PROBINIT (init,name,namlen,problo,probhi)
-     
-     use probdata_module, only: initialize
-
-     implicit none
-
-     integer :: init, namlen
-     integer :: name(namlen)
-     double precision :: problo(3), probhi(3)
-
-     call initialize(name, namlen)
-
-   end subroutine PROBINIT
-
-
-   ! ::: -----------------------------------------------------------
-   ! ::: This routine is called at problem setup time and is used
-   ! ::: to initialize data on each grid.  
-   ! ::: 
-   ! ::: NOTE:  all arrays have one cell of ghost zones surrounding
-   ! :::        the grid interior.  Values in these cells need not
-   ! :::        be set here.
-   ! ::: 
-   ! ::: INPUTS/OUTPUTS:
-   ! ::: 
-   ! ::: level     => amr level of grid
-   ! ::: time      => time at which to init data             
-   ! ::: lo,hi     => index limits of grid interior (cell centered)
-   ! ::: nstate    => number of state components.  You should know
-   ! :::		   this already!
-   ! ::: state     <=  Scalar array
-   ! ::: delta     => cell size
-   ! ::: xlo,xhi   => physical locations of lower left and upper
-   ! :::              right hand corner of grid.  (does not include
-   ! :::		   ghost region).
-   ! ::: -----------------------------------------------------------
-   subroutine ca_initdata(level,time,lo,hi,nscal, &
-                          state,state_l1,state_l2,state_l3,state_h1,state_h2,state_h3, &
-                          delta,xlo,xhi)
-
-     use probdata_module
-     use eos_module, only: eos_input_re, eos
-     use eos_type_module, only: eos_t
-     use meth_params_module, only : NVAR, URHO, UMX, UMY, UMZ, UTEMP, &
-          UEDEN, UEINT, UFS, rot_period
-     use network, only : nspec
-     use bl_constants_module
-     use model_parser_module, only: idens_model, itemp_model, ipres_model, ispec_model
-     use fundamental_constants_module, only: Gconst, M_solar
-     use prob_params_module, only: center
-
[…]

### 2015-07-13 tests/hydro/evrard_collapse/source/probdata.f90

@@ -1,161 +1,0 @@
-module probdata_module
-
-  use network, only: nspec, network_species_index
-  use eos_type_module, only: eos_t
-  use eos_module, only: eos_input_rt, eos
-
-  ! Probin file
-  character (len=:), allocatable :: probin
-
-  ! Determine if we are the I/O processor
-  integer :: ioproc
-  
-  double precision, allocatable :: model_r(:), model_state(:,:)
-
-  integer :: npts_model
-
-  double precision :: mass, radius
-  double precision :: stellar_comp(nspec)
-
-  ! Smallest allowed mass fraction
-  double precision :: smallx
-
-  ! Smallest allowed velocity on the grid
-  double precision :: smallu
-  
-  ! Density of ambient gas around the star
-  double precision :: ambient_density
-  
-  ! Controls interpolation from 1D model to 3D model
-  integer :: nsub
-
-contains
-
-  ! This routine calls all of the other subroutines at the beginning
-  ! of a simulation to fill in the basic problem data.
-
-  subroutine initialize(name, namlen)
-
-    use bl_constants_module, only: ZERO
-    use bl_error_module, only: bl_error
-
-    implicit none
-
-    integer :: namlen, i
-    integer :: name(namlen)
-
-    ! Build "probin" filename -- the name of the file containing the fortin namelist.
-    allocate(character(len=namlen) :: probin)
-    do i = 1, namlen
-       probin(i:i) = char(name(i))
-    enddo
-
-    ! Determine if we are the I/O processor, and save it to the ioproc variable
-
-    call get_ioproc
-
-    ! Read in the namelist to set problem parameters
-
-    call read_namelist
-
-    ! Finalize ambient state, and get small_pres
-
-    call set_small
-
-  end subroutine
-
-
-
-  ! This routine reads in the namelist
-
-  subroutine read_namelist
-
-    use bl_constants_module, only: ZERO, HALF
-    use meth_params_module
-
-    implicit none
-
-    integer :: untin, i
-
-    integer :: iC12, iO16
-    double precision :: stellar_C12, stellar_O16
-
-    namelist /fortin/ &
-         mass, radius, &
-         ambient_density, &
-         nsub, &
-         stellar_C12, stellar_O16
-
-    ! Set namelist defaults
-
-    nsub = 1
-
-    mass   = 1.0d0
-    radius = 1.0d9
-
[…]

### 2015-07-13 documentation/userguide.tex

@@
 \usepackage{amssymb}
+\usepackage{bm}
 
@@
 \begin{document}
+
+% For interpreting the journal names in the bibilography; this is copied and 
+% amended from emulateapj.cls.
+
+\newcommand\aj{AJ}%        % Astronomical Journal 
+\newcommand\araa{ARA\&A}%  % Annual Review of Astron and Astrophys 
+\newcommand\apj{ApJ}%    % Astrophysical Journal 
+\newcommand\apjl{ApJ}%     % Astrophysical Journal, Letters 
+\newcommand\apjs{ApJS}%    % Astrophysical Journal, Supplement 
+\newcommand\ao{Appl.~Opt.}%   % Applied Optics 
+\newcommand\apss{Ap\&SS}%  % Astrophysics and Space Science 
+\newcommand\aap{A\&A}%     % Astronomy and Astrophysics 
+\newcommand\aapr{A\&A~Rev.}%  % Astronomy and Astrophysics Reviews 
+\newcommand\aaps{A\&AS}%    % Astronomy and Astrophysics, Supplement 
+\newcommand\azh{AZh}%       % Astronomicheskii Zhurnal 
+\newcommand\baas{BAAS}%     % Bulletin of the AAS 
+\newcommand\icarus{Icarus}% % Icarus
+\newcommand\jrasc{JRASC}%   % Journal of the RAS of Canada 
+\newcommand\memras{MmRAS}%  % Memoirs of the RAS 
+\newcommand\mnras{MNRAS}%   % Monthly Notices of the RAS 
+\newcommand\pra{Phys.~Rev.~A}% % Physical Review A: General Physics 
+\newcommand\prb{Phys.~Rev.~B}% % Physical Review B: Solid State 
+\newcommand\prc{Phys.~Rev.~C}% % Physical Review C 
+\newcommand\prd{Phys.~Rev.~D}% % Physical Review D 
+\newcommand\pre{Phys.~Rev.~E}% % Physical Review E 
+\newcommand\prl{Phys.~Rev.~Lett.}% % Physical Review Letters 
+\newcommand\pasp{PASP}%     % Publications of the ASP 
+\newcommand\pasj{PASJ}%     % Publications of the ASJ 
+\newcommand\qjras{QJRAS}%   % Quarterly Journal of the RAS 
+\newcommand\skytel{S\&T}%   % Sky and Telescope 
+\newcommand\solphys{Sol.~Phys.}% % Solar Physics 
+\newcommand\sovast{Soviet~Ast.}% % Soviet Astronomy 
+\newcommand\ssr{Space~Sci.~Rev.}% % Space Science Reviews 
+\newcommand\zap{ZAp}%       % Zeitschrift fuer Astrophysik 
+\newcommand\nat{Nature}%  % Nature 
+\newcommand\iaucirc{IAU~Circ.}% % IAU Cirulars 
+\newcommand\aplett{Astrophys.~Lett.}%  % Astrophysics Letters 
+\newcommand\apspr{Astrophys.~Space~Phys.~Res.}% % Astrophysics Space Physics Research 
+\newcommand\bain{Bull.~Astron.~Inst.~Netherlands}% % Bulletin Astronomical Institute of the Netherlands 
+\newcommand\fcp{Fund.~Cosmic~Phys.}%   % Fundamental Cosmic Physics 
+\newcommand\gca{Geochim.~Cosmochim.~Acta}% % Geochimica Cosmochimica Acta 
+\newcommand\grl{Geophys.~Res.~Lett.}%  % Geophysics Research Letters 
+\newcommand\jcp{J.~Chem.~Phys.}%     % Journal of Chemical Physics 
+\newcommand\jgr{J.~Geophys.~Res.}%     % Journal of Geophysics Research 
+\newcommand\jqsrt{J.~Quant.~Spec.~Radiat.~Transf.}%   % Journal of Quantitiative Spectroscopy and Radiative Trasfer 
+\newcommand\memsai{Mem.~Soc.~Astron.~Italiana}% % Mem. Societa Astronomica Italiana 
+\newcommand\nphysa{Nucl.~Phys.~A}%     % Nuclear Physics A 
+\newcommand\physrep{Phys.~Rep.}%       % Physics Reports 
+\newcommand\physscr{Phys.~Scr}%        % Physica Scripta 
+\newcommand\planss{Planet.~Space~Sci.}%  % Planetary Space Science 
+\newcommand\procspie{Proc.~SPIE}%      % Proceedings of the SPIE
+
+\newcommand\actaa{Acta Astron.}%  % Acta Astronomica
+\newcommand\caa{Chinese Astron. Astrophys.}%  % Chinese Astronomy and Astrophysics
+\newcommand\cjaa{Chinese J. Astron. Astrophys.}%  % Chinese Journal of Astronomy and Astrophysics
+\newcommand\jcap{J. Cosmology Astropart. Phys.}%  % Journal of Cosmology and Astroparticle Physics
+\newcommand\na{New A}%  % New Astronomy
+\newcommand\nar{New A Rev.}%  % New Astronomy Review
+\newcommand\pasa{PASA}%  % Publications of the Astron. Soc. of Australia
+\newcommand\rmxaa{Rev. Mexicana Astron. Astrofis.}%  % Revista Mexicana de Astronomia y Astrofisica
 
@@
 interactions in the EOS evaluation.
+
+\section{Diagnostic Output}
+
+\texttt{wdmerger} provides two kinds of diagnostic output at runtime. One set keeps track of quantities affecting 
+the entire grid (such as the total amount of mass and energy on the grid); the other tabulates properties 
+of the individual stars. The former is, by default, printed to \texttt{grid\_diag.out} and the latter to 
+\texttt{star\_diag.out}. This output can be used to keep track of how the run is going, but more often is 
+saved for the purposes of plotting later on (since it may be too expensive to write out enough plotfiles to 
+generate these types of curves at the end of the run).
+
+The routines that generate both of these are found in \\\texttt{source/diagnostics/sum\_integrated\_quantities.cpp}. 
+This is a replacement of the \texttt{CASTRO} function that does something similar. We use the same 
+conventions; for example, if summing over the whole grid and we encounter a region that is covered by 
+a fine grid, we zero out the contribution from the coarse grid and only count the fine grid quantities 
+in that location.
+
+\subsection{Grid Quantities}
+
+The following are the quantities that involve operations over the entire grid, as well as the continuous 
+integral representation of them. Discretization is typically done by setting $\int dV \rightarrow \Delta V$ and 
+assuming cell-centered values for the integrands.
+\begin{itemize}
+  \item \texttt{TOTAL ENERGY}: Total amount of energy stored in the system.
+  \begin{equation}
+    \int dV\, \left(-\frac{1}{2}\rho \phi + \rho E + \rho \Phi_{\text{rot}}\right)
+  \end{equation}
+
+  \item \texttt{TOTAL E GRID}: Total amount of energy represented on the computational grid; differs from
+    \texttt{total\_energy} only when rotation is enabled.
+  \begin{equation}
+    \int dV\, \left(\frac{1}{2}\rho \phi + \rho E\right)
+  \end{equation}
+
+  \item \texttt{GAS ENERGY}: Kinetic plus internal energy.
+  \begin{equation}
+    \int dV\, \rho E
+  \end{equation}
+
+  \item \texttt{KIN. ENERGY}: Kinetic energy (calculated using velocities; in general will differ from $\rho(E - e)$).
+  \begin{equation}
+    \int dV\, \frac{1}{2} \rho \mathbf{v}^2
+  \end{equation}
+
+  \item \texttt{ROT. ENERGY}: Energy associated with rotating reference frame.
+  \begin{equation}
+    \int dV\, \rho\, {\bm \omega} \cdot (\mathbf{r} \times \mathbf{p}) + \int dV\, \frac{1}{2} {\bm \omega}^T \cdot \mathbf{I} \cdot {\bm \omega}
+  \end{equation}
+
+  \item \texttt{GRAV. ENERGY}: Gravitational potential energy ($\phi < 0$).
+  \begin{equation}
+    \int dV\, \frac{1}{2}\rho\phi
+  \end{equation}
+
+  \item \texttt{INT. ENERGY}: Internal/thermal energy of the gas.
+  \begin{equation}
+    \int dV\, \rho e
+  \end{equation}
+
+  \item \texttt{XMOM, YMOM, ZMOM}: Total momentum.
+  \begin{equation}
+    \int dV\, \rho (\mathbf{u} + {\bm \omega} \times \mathbf{r})
+  \end{equation}
+
+  \item \texttt{XMOM GRID, YMOM GRID, ZMOM GRID}: Momentum on grid only; no rotation.
+  \begin{equation}
+    \int dV\, \rho \mathbf{u}
+  \end{equation}
+
+  \item \texttt{XMOM ROT., YMOM ROT., ZMOM ROT.}: Momentum due only to rotation.
+  \begin{equation}
+    \int dV\, \rho {\bm \omega} \times \mathbf{r}
+  \end{equation}
+
+  \item \texttt{ANG. MOM. X, ANG. MOM. Y, ANG. MOM. Z}: Total angular momentum.
+  \begin{equation}
+    \int dV\, \rho(\mathbf{r} \times \mathbf{u} + \int dV\, \mathbf{I} \cdot {\bm \omega}
+  \end{equation}
+
+  \item \texttt{ANG. MOM. X GRID, ANG. MOM. Y GRID, ANG. MOM. Z GRID}: Angular momentum on grid only; no rotation.
+  \begin{equation}
+    \int dV\, \rho\mathbf{r} \times \mathbf{u}
+  \end{equation}
+
+  \item \texttt{ANG. MOM. X ROT., ANG. MOM. Y ROT., ANG. MOM. Z ROT.}: Angular momentum due only to rotation.
+  \begin{equation}
+    \int dV\, \mathbf{I} \cdot {\bm \omega}
+  \end{equation}
+
+  \item \texttt{MASS}: Mass.
+  \begin{equation}
+    \int dV\, \rho
+  \end{equation}
+
+  \item \texttt{X COM, Y COM, Z COM}: Center of mass.
+  \begin{equation}
+    \int dV\, \rho \mathbf{x}
+  \end{equation}
+
+  \item \texttt{X COM VEL, Y COM VEL, Z COM VEL}: Velocity of center of mass.
+  \begin{equation}
+    \int dV\, \rho \mathbf{v}
+  \end{equation}
+
+  \item \texttt{h\_+, h\_x (rotation axis, star axis, motion axis)}: Gravitational wave strain in ``plus'' and ``cross'' polarization modes, observed along the rotation axis, the axis the stars were originally aligned along, and the axis the stars originally moved on.
+(By default, there are $z$, $x$, and $y$ respectively).
+See \texttt{wdmerger} paper I for details on how these are calculated.
+
+  \subsection{Star Quantities}
+
+  Now we consider the quantities describing the stars.
+First we must describe our strategy for determining what parts of the domain correspond to the primary and what parts correspond to the secondary.
+At the beginning of the simulation, we establish the binary so that each star's center of mass lies on the \texttt{star\_axis} and is zero in the other directions.
+At each subsequent timestep, we make a guess at the location of the new center of mass: $\mathbf{x}_{\text{COM}} \to \mathbf{x}_{\text{COM}} + \mathbf{v}_{\text{COM}} + \Delta t$.
+Here $\mathbf{v}_{\text{COM}}$ is the velocity of the center of mass at the last timestep (which at the beginning of the simulation is simply the Keplerian orbital speed).
+Then we refine this estimate by doing an integral on the domain.
+We define $\mathbf{x}^\prime_P$ and $\mathbf{x}^\prime_S$ as the tentative locations, and then do a center-of-mass calculation on all zones within the Roche radius $r_L$ of that star, which encompasses the effective gravitational domain of that star.
+We use the commonly employed definition of the Roche radius of \citet{eggleton:1983},
+  \begin{equation}
+    \frac{r_L}{a} = \frac{0.49 q^{2/3}}{0.6^{2/3} + \text{ln}(1 + q^{1/3})}.
+  \end{equation}
+  Here $a$ is the orbital separation (i.e.
+$\left|\mathbf{x}^\prime_P - \mathbf{x}^\prime_S\right|$ ), and if $q = M_S / M_P$ is the mass ratio then this yields the Roche radius of the secondary.
+$q$ is inverted to get the Roche radius of the primary.
+Once we do this calculation, we have a new $\mathbf{x}_P$ and $\mathbf{x}_S$, as well as $\mathbf{v}_P$ and $\mathbf{v}_S$ and $M_P$ and $M_S$.
+We also then define the distance between the stars as $|\mathbf{x}_P - \mathbf{x}_S|$.
+We also calculate the angle that the line joining them makes with respect to the original axis; this is a measure of how much they have rotated.
+With the center of masses located, we can add any diagnostics of interest we want related to the stars.
+At present we only have one.
+The \texttt{PRIMARY 1E$n$ RADIUS} and \texttt{SECONDARY 1E$n$ RADIUS} are the radii from the center of the stars at which the average density is $10^n\ \text{g}\ \text{cm}^{-3}$.
+
+\end{itemize}
 

### 2015-07-13 documentation/userguide.tex
New arrangement of the tests directory, to more closely reflect the fact that our tests are now either tests of our problem or problems in the CASTRO test suite.

@@
 \begin{itemize}
-  \item The \texttt{gravity} subdirectory includes problems that test the Poisson 
-self-gravity module in CASTRO.
+  \item The \texttt{binary} subdirectory includes problems that test the binary star setup.
   \begin{itemize}
@@
       and compares the resulting gravitational potential to the exact result.
+    \item The \texttt{circular\_orbit} test loads two white dwarfs on a grid and then orbits
+      them for a large number of periods, to test conservation of linear momentum,
+      angular momentum and energy convergence.
+    \item The \texttt{freefall} test loads two white dwarfs onto a grid and allows them to fall towards
+      each other under their mutual gravitational influence. We check to see whether the timescale for them
+      colliding matches the analytical expectation for two point masses.
+  \end{itemize}
+
+  \item The \texttt{single} subdirectory is for tests that only use a single star.
+  \begin{itemize}
+    \item The \texttt{single\_star\_hse} test loads a single white dwarf on a grid and checks how well 
+      hydrostatic equilibrium is maintained.
+  \end{itemize}
+
+  \item The \texttt{castro} subdirectory is for running tests in the CASTRO test suite that are relevant
+  to issues that the merger problem will face.
+  \begin{itemize}
     \item The \texttt{uniform\_cube\_sphere} test is designed to check the convergence
@@
         onto the computational domain and then calculates the potential at several different
-spatial resolutions to check if the potential converges.
-  \end{itemize}
-  \item The \texttt{hydro} subdirectory tests includes tests that use the hydrodynamics 
-module, including many with self-gravity.
-  \begin{itemize}
-    \item The \texttt{circular\_orbit} test loads two white dwarfs on a grid and then orbits
-them for a large number of periods, to test conservation of linear momentum,
-angular momentum and energy convergence.
+        spatial resolutions to check if the potential converges
     \item The \texttt{evrard\_collapse} test runs the test problem described originally by 
@@
       It is a good test of the energy conservation properties of the code.
-    \item The \texttt{freefall} test loads two white dwarfs onto a grid and allows them to fall towards
-each other under their mutual gravitational influence. We check to see whether the timescale for them
-colliding matches the analytical expectation for two point masses.
     \item The \texttt{kelvin\_helmoltz} test runs several varieties of a 2D shearing fluid that exhibits the 
       Kelvin-Helmholtz instability. The relevant papers are \cite{robertson:2010}, \cite{arepo}, and \cite{mcnally:2012}.
-    \item The \texttt{single\_star\_hse} test loads a single white dwarf on a grid and checks how well 
-hydrostatic equilibrium is maintained.
   \end{itemize}
+
   \item The \texttt{performance} subdirectory examines the scaling and performance properties of CASTRO.
@@
       checks the scaling properties.
+    \item The \texttt{threading} test is useful for determining the optimal number of OpenMP threads for your machine.
+    \item The \texttt{load\_balancing} test finds which load balancing strategy makes the most sense for our problem.
   \end{itemize}

### 2015-07-13 documentation/userguide.tex
Phase out the ability to have different source code. The user could still compile their own code and then use our job submission system, but they need to provide their own source code and makefile to do this.

@@
 and describes the underlying code that facilitates the job submission. If you are interested in
-doing white dwarf mergers and want to use the default settings, read Section \ref{sec:standard_problems}.
+doing white dwarf mergers and want to use the default source code, read Section \ref{sec:standard_problems}.
 If you are interested in setting up a different problem but using this software's job submission 
-system, you can do so; read Section \ref{sec:non-standard_problems} for more information.
+system, you can do so; Section \ref{sec:non-standard_problems} gives some advice on that.
 
@@
 
-If you want to build your own problem setup, you can do that too. The makefile and job submission 
+If you want to build your own problem setup, you can do that too. The job submission 
 requirements are the same as in Section \ref{sec:standard_problems}, but you will need to provide 
-your own source files. They should all go in a \texttt{source/} subdirectory in your problem location.
-At minimum, you'll need a \texttt{Prob\_3d.f90} file (or the equivalent for lower dimensions) and a 
-probin file. The probin file can be empty, it just needs to be there because otherwise the helper 
-routines will make a copy of the default probin file. To get a sense of how this is done, look at 
-some of the existing test problems, in particular the \texttt{kelvin\_helmholtz} test. To tell the 
-code that you want to compile this source code instead of the default code, make a file called 
-\texttt{full\_source} and place it in the \texttt{source/} directory (you can just use \texttt{touch}
-to create an empty file). Then follow the rest of the instructions as normal, starting with the 
-compile step.
-
-The only source code that will be built from the default location is \texttt{sum\_integrated\_quantities.cpp},
-which controls the diagnostic output. Our version is a bit more extensive than what comes in CASTRO
-by default, so we include this in every build.
+your own source files and makefile. They can go in the \texttt{compile/} subdirectory in your problem location.
+At minimum, you'll need a \texttt{Prob\_3d.f90} file (or the equivalent for lower dimensions) and
+corresponding inputs and probin files. The problems in \texttt{tests/castro} are similar to this;
+they are ultimately building using the source code in the CASTRO test suite, but using a
+\texttt{run\_test.sh} that is very similar to the one used for merger problems.
 

### 2015-07-13 job_scripts/run_utils.sh

@@ -635,16 +635,16 @@ function copy_files {
     fi
 
     if [ ! -e "$dir/$inputs" ]; then
-        if [ -e "$source_dir/$inputs" ]; then
-            cp $source_dir/inputs $dir/$inputs
+        if [ -e "$compile_dir/$inputs" ]; then
+            cp $compile_dir/inputs $dir/$inputs
         else
             cp $WDMERGER_HOME/source/inputs $dir/$inputs
 	fi
     fi
 
     if [ ! -e "$dir/$probin" ]; then
-	if [ -e "$source_dir/$probin" ]; then
-	    cp $source_dir/probin $dir/$probin
+	if [ -e "$compile_dir/$probin" ]; then
+	    cp $compile_dir/probin $dir/$probin
 	else
 	    cp $WDMERGER_HOME/source/probin $dir/$probin
 	fi
@@ -1202,7 +1202,6 @@ fi
 # Directory to compile the executable in
 
 compile_dir="compile"
-source_dir="source"
 
 # Some variables we need for storing job information.
 
@@ -1240,14 +1239,6 @@ if [ -d $compile_dir ]; then
 
     if [ ! -d $plots_dir ]; then
 	mkdir $plots_dir
-    fi
-
-    # Source code directory
-
-    source_dir="source"
-    
-    if [ ! -d $source_dir ]; then
-	mkdir $source_dir
     fi
 
 fi

### 2015-07-13 tests/castro/kelvin_helmholtz/run_test.sh

@@ -2,8 +2,8 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 
 TEST_DIR=$CASTRO_DIR/Exec/KH
 
-cp $TEST_DIR/inputs.2d source/inputs
-cp $TEST_DIR/probin source/probin
+cp $TEST_DIR/inputs.2d $compile_dir/inputs
+cp $TEST_DIR/probin $compile_dir/probin
 
 # Loop over problem choices
 

### 2015-07-13 tests/castro/uniform_cube_sphere/run_test.sh

@@ -2,8 +2,8 @@ source $WDMERGER_HOME/job_scripts/run_utils.sh
 
 TEST_DIR=$CASTRO_DIR/Exec/uniform_cube_sphere
 
-cp $TEST_DIR/inputs source/
-cp $TEST_DIR/probin source/
+cp $TEST_DIR/inputs $compile_dir/
+cp $TEST_DIR/probin $compile_dir/
 
 # Loop over problem choices
 

### 2015-07-19 papers/wdmerger_I/paper.tex
Switch to solely using the journal commands provided by the emulateapj class, so that we are fully consistent.
Add a note on refluxing to the paper.

@@
-%\documentclass[12pt,preprint]{../aastex}
 \documentclass[iop]{../emulateapj}
@@
 
+A final issue worth noting relates to complications posed by mesh refinement. Like 
+many AMR codes, CASTRO implements a refluxing algorithm at the end of a timestep. 
+At any locations on a particular grid level that are adjacent to, but not covered by,
+zones from a finer grid, there will be a disagreement about the fluxes crossing 
+this coarse-fine interface. If left uncorrected, this results in a violation of 
+the conservation condition. Therefore the value of the coarse flux on this interface 
+is replaced by the sum of the fine fluxes at that interface, at the end of the 
+coarse timestep. In general, this change in mass distribution on the grid 
+results in a change in the gravitational potential, meaning that an additional 
+corrective source term ought to be added. \cite{miniati:2007} and \cite{castro} 
+describe an algorithm for doing so. We consider the size of this effect to be 
+negligible and therefore we avoid the expense of solving for the potential change caused by 
+this mass change. Our grid boundaries are always located outside of the dense 
+stellar interiors that generate almost all of the gravitational potential. 
+Consequently the refluxing operation only changes the solution in regions of 
+ambient density, and even there only to the extent that there is disagreement 
+between the coarse and fine fluxes -- this clearly has a 
+vanishingly small effect on the gravitational potential.
+
 \subsubsection{Boundary Conditions}\label{sec:gravity_boundary_conditions}

### 2015-07-19 papers/wdmerger_I/paper.tex
Clean up the appendix.

@@
 
-\subsection{Hydrodynamics}
+\subsection{Hydrodynamics}\label{sec:Hydrodynamics}
 
@@
 
-Summary of changes:
-\begin{itemize}
-\item New gravity update type: fixes some energy conservation
-
-\item changing I's to use limit of the parabola instead of the
-  cell-center: not sure if this does much
-
-\item Colella \& Glaz Riemann solver: does a much better job with
-  temperature in strong shocks
-
-\item Tracing of gravity: seems to help with the material at the edge
-  of the star
-
-\item implicit update of the rotation terms?  (I coded this up once, but
-  it is not currently in the code)
-\end{itemize}
-
 \subsection{Reference States}
@@
 For all the runs, the PPM reconstruction is done using the original limiters for
-the parabolic profiles \citep{ppm}.  The prediction of the interface
+the parabolic profiles \citep{ppm}; see Section \ref{sec:Hydrodynamics} for a brief 
+discussion about the limiters.  The prediction of the interface
 states appears as:
@@
 carried by the wave $\nu$.  The reader is referred to
-\citet{ppmunsplit} for further defaults.
+\citet{ppmunsplit} for further details.
 
@@
 We comment on the choice of reference state for passively-advected
-quantities (like $X_k$ or the transverse velocity).
-First, consider one of the variables present in one-dimensional flow
-(density, velocity in the normal direction, and pressure).  Here we
-our reference state is as in Eq.~\ref{eq:refchoice} and we have.
+quantities (like $X_k$ or the transverse velocity), which is not typically 
+discussed. First, consider one of the variables present in one-dimensional flow
+(density, velocity in the normal direction, and pressure), and let our 
+reference state be as in Eq.~\ref{eq:refchoice}.
 Ignoring flattening, if there are no waves moving toward our
@@
 added in Eq.~\ref{eq:ppmstatel}, but our choice of reference state makes
-that term 0 by design, and our interface state is:
+that term zero by design, and our interface state is:
 \begin{equation}
@@
 
-The details of the reference state for passively advected quantities 
-(which includes the transverse velocities in the 1-d tracing) is not
-typically discussed.  If we use the same idea of the reference state as
+However, now consider the same approach applied to passively advected quantities.
+If we use the same idea of the reference state as
 in Eq.~\ref{eq:refchoice}, and consider a quantity $\xi$ which should only
@@
 \end{equation}
-(where we used the fact that the eigenvectors are normalized to 1 and
+(where we used the fact that the eigenvectors are normalized to unity and
 don't mix in any other states when dealing with passive terms).  This
@@
 this was traced under to find the forcing that affects the interface
-for each wave.  CASTRO followed \citet{ppmunsplit} which instead adds
-$(\Delta t/2)g$ to the interface states at the end of the
+for each wave.  CASTRO originally followed \citet{ppmunsplit} which instead adds
+$(\Delta t/2)g$ to the interface states for velocity at the end of the
 reconstruction.  In the current implementation, we do the original
 parabolic reconstruction and characteristic tracing of the gravitation
-source.  Our system with the source appears as:
+source. This can be controlled in CASTRO with the parameter 
+{\tt castro.ppm\_trace\_grav}. Our system with the source appears as:
 \begin{equation}
@@
 $u$, not $\rho$ or $p$.  Note that in the PPM paper, they put $G$ on 
-the lefthand side of the primitive variable equation, so our signs are
+the left-hand side of the primitive variable equation, so our signs are
 opposite.  Our projections are now:
@@
 \end{eqnarray}
-These differ from the expression in the PPM paper, where $\Delta t G$,
-not $\Delta t/2 G$ is used in the projection, however we believe that
-the factor of $1/2$ is correct.  To see this, notice that if both
+(The expression in the PPM paper contains $\Delta t G$, not $(\Delta t/2) G$,
+but we believe that the factor of $1/2$ is correct.  To see this, notice that if both
 waves are moving toward the interface, then the source term that is
@@
 t/2) g$ for constant g---this matches the result from Taylor
-expanding to the interface at the half-time (as in \citealt{ppmunsplit}).
+expanding to the interface at the half-time (as in \citealt{ppmunsplit}).)
 
@@
 states.  Both methods are second-order accurate.
-
-The tracing of gravity can be controlled in the public version of CASTRO
-with the parameter {\tt castro.ppm\_trace\_grav}.
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[scale=1.0]{plots/reference}
-  \caption{ \label{Fig:sod} Solution to Sod's problem with the original
-    reference state and the new reference state, as compared to the
-    exact solution.  We note that the new reference state shows a
-    slightly sharper shock and contact, but also has a dip at the tail
-    of the rarefaction.}
-\end{figure}
-
 \clearpage

### 2015-07-20 papers/wdmerger_I/paper.tex
Add a first cut at the section on the circular orbit test.

--- 2015-07-09 papers/wdmerger_I/paper.tex
@@
-%\documentclass[12pt,preprint]{../aastex}
 \documentclass[iop]{../emulateapj}
-% for \sout
-\usepackage{ulem}
-% makes sure \em{} is italic rather than underlined (corrects ulem from line above)
-\normalem
 
@@
 \usepackage{verbatim}
-
-\newcommand{\rhocutoff}{\rho_\mathrm{cutoff}}
-\newcommand{\rhoanelastic}{\rho_\mathrm{anelastic}}
-
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-\newcommand{\Tcutoff}{T_\mathrm{cutoff}}
 
@@
 
-should show:
-
-effect of resolution
-
-inertial vs. rotating frame
-
-effect of ppm\_reference and gravity update type
-
-effect of boundary conditions (size of domain and possibly multipole dependence)
+We now consider the phase of the binary system where the stars are orbiting each other 
+at distances great enough that the initial orbits should be approximately Keplerian. 
+There are a number of effects worth looking into here. For simplicity, we choose two 
+cases to demonstrate the simulation behavior: an equal mass case of two $0.9\msolar$ 
+white dwarfs, and an unequal mass case of $0.9\msolar$ and $0.6\msolar$ white dwarfs.
+In each case, the initial orbital period is 100 seconds.
+
+For some of the algorithms described earlier in this work, a single orbit of these 
+systems is enough to examine their effects. In Section \ref{sec:gravity_boundary_conditions},
+we discussed the replacement of a monopole boundary condition solver for the gravitational 
+potential with a more general multipole solver for the boundaries. To test the relevance 
+of this effect, we considered a single orbit of the unequal mass system and measured 
+the distance between the two white dwarfs at the beginning of the simulation and after 
+the full orbital period. This distance should not change significantly over that timescale.
+We performed this test for maximum multipole moments ranging from 0 (the monopole term) to 16.
+The results are shown in Figure \ref{fig:gravity_bcs}. Terms in the boundary potential 
+that vary faster than $r^{-5}$ are effectively negligible in determining the outcome of the orbit, 
+justifying our typical choice of maintaining terms up to $r^{-7}$.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/gravity_bcs}
+  \caption{Absolute magnitude of the relative change in the distance of two unequal mass white dwarfs after one orbital period. 
+           The stars were evolved in an inertial reference frame. The horizontal axis is the number of terms or multipole moments 
+           captured in the series expansion for the potential at the domain boundary.\label{fig:gravity_bcs}}
+\end{figure}
+
+Another that we consider is the energy conservation of the system. Recalling Section \ref{sec:gravity_hydro_coupling}, 
+there are several different methods of applying the gravitational source term to the hydrodynamics equations. In CASTRO we presently 
+have four options, controlled by the parameter {\tt castro.grav\_source\_type}, which we will shortern to {\tt gs} for the present 
+discussion. ${\tt gs} = 1$ is the standard cell-centered source term for gravity. ${\tt gs} = 2$ is a slight variation on this scheme 
+that determines the value of the energy source term after the momentum source term has been applied; the former approach uses 
+the uncorrected momenta in calculating $\rho \mathbf{u} g$, which we have found to be more accurate. ${\tt gs} = 3$ is entirely 
+different: after calculating the new momenta, we reset the total energy to be equal to the internal energy plus the 
+kinetic energy. This has the virtue of ensuring that there is no conflict due to discretization between the momentum and 
+energy equations, and also correctly ensuring that the gravitational force does not directly change the internal energy -- 
+and thus the temperature -- of the fluid. However, it explicitly sacrifices total energy conservation. ${\tt gs} = 4$ is 
+the new conservative method of evaluating the energy source terms at cell faces. The results for the change in energy 
+after a single orbit are seen in the first column of Table \ref{table:sources}. The first two versions give reasonable and 
+similar levels of energy conservation. The third has total energy changes on the order of 100\%, but this itself 
+does not have a severe effect on the dynamics because in this scheme the total energy variable is effectively a placeholder 
+value of the kinetic energy plus internal energy, rather than being evolved directly. The last scheme is nearly two orders 
+of magnitude better in energy conservation, justifying the effort in varying the scheme.
+
+In Table \ref{table:sources} we show also the effects on energy conservation of using the inertial reference frame. 
+We use ${\tt rs}$ for the CASTRO parameter {\tt castro.rot\_source\_type}, and we have the same four options for updating
+the energy source term as for gravity. The ${\tt rs} = 0$ column means that rotation is turned off and we are 
+in the inertial frame. We see that the choice of rotational coupling is much less important than the choice of gravity coupling. 
+Interestingly, the ``conservative'' ${\tt rs} = 4$ is slightly worse in energy conservation than the non-conservative, 
+cell-centered ${\tt rs} = 2$ algorithm. This may be because the rotational force grows with distance, and so losses at the 
+domain boundaries may be enhanced by explicitly evaluating the flux at cell edges.
+
+\input{plots/sources.table}
+
+We are most interested in the stability of these systems over long timescales. To this end, we consider the same systems as above, 
+but evolve them for 25 orbital periods. In Figure \ref{fig:circular_orbit_comparison} we illustrate the evolution of these systems 
+by plotting the center of mass locations of the white dwarfs on the orbital ($xy$) plane. For the equal mass case in the inertial 
+reference frame, the curves essentially fall on top of each other, indicating that the stars are indeed orbiting at the initial 
+distance without substantial deviation. In the unequal mass case, the stars now orbit at different distances but do not much 
+deviate from their initial distance. We do observe a secular decline in the center of mass of these stars, indicating that 
+eventually the stellar orbits will degrade due to numerical error even without intervention. The co-rotating frame is different.
+For clarity of visualization, we rotate these results back into the inertial frame before displaying their orbits.
+The stellar orbit in the equal mass case does not show the same secular decline, and the system appears to be stable. However, 
+the cost is that the oscillations about this equilibrium are slightly larger than in the inertial frame. The unequal mass case displays 
+this effect starkly: while the average distance is not declining, there are huge oscillations about this average. In addition, 
+the stars move substantially within the rotating frame itself, yielding the chaotic-looking orbits in the figure. The oscillations 
+grow with time and will eventually cause the stars to merge; however, the stars would eventually merge even in the inertial frame 
+case, and at a faster rate than the equal mass case. This suggests that this mass ratio is too large to ignore tidal effects, and 
+that the tidal effects play out much differently in the rotating frame.
+
+\begin{figure*}
+  \centering
+  \includegraphics[scale=0.7]{plots/circular_orbit_comparison}
+  \caption{Positions of the white dwarfs in the orbital plane for four cases evolved over 25 orbital periods. 
+           The $x$ and $y$ axes are normalized to the size of the domain, so that $x = -0.5$ is the left edge 
+           and $x = 0.5$ is the right edge. The dashed blue curve is the position of the primary white dwarf, 
+           and the solid red curve is the position of the secondary. In plot (a) we have the equal mass system 
+           evolved in the inertial reference frame, and in plot (c) we have the same system evolved in a rotating frame, 
+           where the positions have been transformed back to the inertial frame for comparison. 
+           Plots (b) and (d) are analogous but for the unequal mass system.\label{fig:circular_orbit_comparison}}
+\end{figure*}
+
+Turning to the conservation properties of the system, we examine as fairly typical cases the equal mass system 
+in the inertial frame for energy conservation (Figure \ref{fig:energy_conservation_equal}), and the unequal mass 
+system in the rotating frame for angular momentum conservation (Figure \ref {fig:angular_momentum_conservation_unequal}). 
+Angular momentum is conserved to within a few percent over the 25 orbits, while energy conservation is about 
+an order of magnitude better. We note that while this is already a fairly good level of energy conservation, it is not 
+nearly as good as the results of \citet{marcello:2012}. This is because we reset the internal energy to a level 
+corresponding to our temperature floor when it goes negative, while \citeauthor{marcello:2012} do not reset and 
+instead ignore the internal energy if it is negative. The resets impose an artificial floor on our ability to 
+conserve energy, but they only happen in low-density regions and do not much affect the large-scale dynamics. 
+Meanwhile, angular momentum conservation is not quite as good. This is linked to the decline (or increase) in 
+the size of the orbit. This implies that we ought to be careful in concluding that at these moderate resolutions 
+we can safely evolve systems for many dozens of orbits; this needs to be verified to ensure that an observed 
+inspiral and merger is physically (not numerically) motivated.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/equal_energy_rot_0.eps}
+  \caption{Absolute magnitude of the relative change in energy of two equal mass white dwarfs through 25 orbital periods,
+           evolved in an inertial reference frame. The decline and recovery is a change in sign of the energy difference.
+           \label{fig:energy_conservation_equal}}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot_1.eps}
+  \caption{Absolute magnitude of the relative change in angular momentum of two equal mass white dwarfs after 25 orbital periods, 
+           evolved in a co-rotating reference frame. We consider only the component of the angular moment along the rotational 
+           axis. The decline and recovery is a change in sign of the angular momentum difference.
+           \label{fig:angular_momentum_conservation_unequal}}
+\end{figure}
 
@@
 a rotating reference frame has its own numerical issues for 
-the type of simulation we desire to perform here. Perhaps contrary 
-to expectations, the stars appear to be more stable in an inertial 
-frame where they are constantly in motion on the grid. It 
+the type of simulation we desire to perform here. Based on 
+the tests performed here, in some cases the stars may be 
+\textit{more} stable in an inertial frame despite 
+constantly being in motion on the grid. It 
 is not easy to predict the correct behavior of such systems, 
-so the best recourse for assessing confidence in a model from a 
-verification standpoint is still to see whether the observed 
-behavior converges with resolution.
+though one recourse for assessing confidence in a model from a 
+verification standpoint is to see whether the observed 
+behavior converges with resolution. However, it is not clear 
+whether at practical resolutions the results in the 
+rotating and inertial frames will converge to each other.
 

### 2015-07-20 papers/wdmerger_I/paper.tex

@@
 $\dot{\Phi}^{n+1/2} = (\Phi^{n+1} - \Phi^{n}) / \Delta t.$)
+
+A final issue worth noting relates to complications posed by mesh refinement. Like 
+many AMR codes, CASTRO implements a refluxing algorithm at the end of a timestep. 
+At any locations on a particular grid level that are adjacent to, but not covered by,
+zones from a finer grid, there will be a disagreement about the fluxes crossing 
+this coarse-fine interface. If left uncorrected, this results in a violation of 
+the conservation condition. Therefore the value of the coarse flux on this interface 
+is replaced by the sum of the fine fluxes at that interface, at the end of the 
+coarse timestep. In general, this change in mass distribution on the grid 
+results in a change in the gravitational potential, meaning that an additional 
+corrective source term ought to be added. \cite{miniati:2007} and \cite{castro} 
+describe an algorithm for doing so. We consider the size of this effect to be 
+negligible and therefore we avoid the expense of solving for the potential change caused by 
+this mass change. Our grid boundaries are always located outside of the dense 
+stellar interiors that generate almost all of the gravitational potential. 
+Consequently the refluxing operation only changes the solution in regions of 
+ambient density, and even there only to the extent that there is disagreement 
+between the coarse and fine fluxes -- this clearly has a 
+vanishingly small effect on the gravitational potential.
 

### 2015-07-28 papers/wdmerger_I/paper.tex
Fix incorrect plot filename
Add reference to MacMillan's original calculation of the gravitational potential of a cube.

@@
 object has been worked out analytically by \citet{waldvogel:1976} (see
-also a similar result in \citealt{hummer:1996}). The potential is in
+also a similar result by \citet{hummer:1996}, and an earlier calculation 
+by \citet{macmillan:1958}). The potential is in
 Equation 15 of that paper, though the last term is missing a factor of
@@
   \centering
-  \includegraphics[scale=0.45]{plots/equal_energy_rot_0.eps}
+  \includegraphics[scale=0.45]{plots/equal_energy_rot0.eps}
   \caption{Absolute magnitude of the relative change in energy of two equal mass white dwarfs through 25 orbital periods,
@@
   \centering
-  \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot_1.eps}
+  \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot1.eps}
   \caption{Absolute magnitude of the relative change in angular momentum of two equal mass white dwarfs after 25 orbital periods, 

### 2015-07-28 papers/wdmerger_I/paper.tex
Replace references using \autoref from the hyperref package, as advocated by AstroBetter in http://www.astrobetter.com/blog/2014/09/29/latex-hyperref-and-emulateapj/.

@@
 \usepackage{verbatim}
+
+% allows in-document links; see http://www.astrobetter.com/blog/2014/09/29/latex-hyperref-and-emulateapj/
+\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref} 
+\usepackage[all]{hypcap} %Links go to figures;breaks on deluxetables       
+\renewcommand*{\backref}[1]{[#1]}
+
+% Override choices in \autoref
+\def\sectionautorefname{Section}
+\def\subsectionautorefname{Section}
+\def\subsubsectionautorefname{Section}
 
@@
 mixtures and in principle should use results from modern stellar
-evolution calculations. Section \ref{sec:Numerical Methodology}
+evolution calculations. \autoref{sec:Numerical Methodology}
+%\hyperref[sec:Numerical Methodology]{Section \ref*{sec:Numerical Methodology}}
 describes our code and why it can provide useful results compared to
-other methodologies used for this problem. Section
-\ref{sec:initial_models} describes the method we use for setting up a
-binary white dwarf simulation. Section \ref{sec:Tests} discusses a few
+other methodologies used for this problem. 
+\autoref{sec:initial_models} describes the method we use for setting up a
+binary white dwarf simulation. \autoref{sec:Tests} discusses a few
 test problems that we use to demonstrate that our code accurately
-solves the equations of fluid dynamics. Section \ref{sec:Performance}
+solves the equations of fluid dynamics. \autoref{sec:Performance}
 demonstrates that the software scales well for supercomputer
-applications. Finally, Section \ref{sec:Conclusions and Discussion}
+applications. Finally, \autoref{sec:Conclusions and Discussion}
 recaps what we have shown and highlights some of the future work we
@@
 spatial resolution. CASTRO is highly parallel and is designed for
-large-scale use on modern supercomputers; see Section
-\ref{sec:Performance} for information on how CASTRO performs for our
+large-scale use on modern supercomputers; see 
+\autoref{sec:Performance} for information on how CASTRO performs for our
 problem. The next few subsections describe our approach to each of the
@@
 within numerical roundoff error. This result can be derived by analytically recasting 
-Equation \ref{eq:momentum_discretization}. In the continuum limit, the two momentum
+\autoref{eq:momentum_discretization}. In the continuum limit, the two momentum
 formulations are identical. Thus the latter has been advocated by, for example, 
@@
 
-As observed by \cite{arepo}, when examining Equation \ref{eq:mass_transfer} one can see the relationship 
+As observed by \cite{arepo}, when examining \autoref{eq:mass_transfer} one can see the relationship 
 between this method and the more common (non-conservative) approach of evaluating the change in 
@@
 
-It is worth pointing out that while Equation \ref{eq:grav_energy_conservation_update} was correctly obtained 
+It is worth pointing out that while \autoref{eq:grav_energy_conservation_update} was correctly obtained 
 by \cite{arepo} (and indeed that is the inspiration for this approach), there is a subtle error in that derivation. 
@@
 whether or not it is static, changing in time, or generated by the fluid distribution itself. 
-Indeed, we will do so for the rotational potential in Section \ref{sec:rotation}, where the 
+Indeed, we will do so for the rotational potential in \autoref{sec:rotation}, where the 
 factor of $1/2$ does not appear. However, it is important also to note that this is not related to a 
@@
 in the discretization of the gravitational potential can lead to very large changes in the energy of the gas.
-(We note in passing too that although \citeauthor{jiang:2013} claim their approach requires a Poisson 
+(We note in passing too that although \cite{jiang:2013} claim their approach requires a Poisson 
 solve for the time derivative of the potential, $\dot{\Phi}$, this is not needed as in practice 
@@
 significant drift of the center of the mass of the system over time,
-which will be discussed in Section \ref{sec:kepler}.
+which will be discussed in \autoref{sec:kepler}.
 
@@
 equation as a series of spherical harmonics.
-\begin{comment}
-, which yields
-\begin{equation}
-  q_{lm} = \int Y^*_{lm}(\theta^\prime, \phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3x^\prime. \label{multipole_moments_original}
-\end{equation}
-\end{comment}
-After some algebraic simplification of Equation
-\ref{spherical_harmonic_expansion} using the addition theorem for
-spherical harmonics,
-\begin{comment}
-\begin{align}
-  &\frac{4\pi}{2l+1} \sum_{m=-l}^{l} Y^*_{lm}(\theta^\prime,\phi^\prime)\, Y_{lm}(\theta, \phi) = P_l(\text{cos}\, \theta) P_l(\text{cos}\, \theta^\prime) \notag \\
-  &\ \ + 2 \sum_{m=1}^{l} \frac{(l-m)!}{(l+m)!} P_{l}^{m}(\text{cos}\, \theta)\, P_{l}^{m}(\text{cos}\, \theta^\prime)\, \left[\text{cos}(m\phi)\, \text{cos}(m\phi^\prime) + \text{sin}(m\phi)\, \text{sin}(m\phi^\prime)\right].
-\end{align}
-\end{comment}
+%% \begin{equation}
+%%   q_{lm} = \int Y^*_{lm}(\theta^\prime, \phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3x^\prime. \label{multipole_moments_original}
+%% \end{equation}
+After some algebraic simplification of \autoref{spherical_harmonic_expansion} 
+using the addition theorem for spherical harmonics,
+%% \begin{align}
+%%   &\frac{4\pi}{2l+1} \sum_{m=-l}^{l} Y^*_{lm}(\theta^\prime,\phi^\prime)\, Y_{lm}(\theta, \phi) = P_l(\text{cos}\, \theta) P_l(\text{cos}\, \theta^\prime) \notag \\
+%%   &\ \ + 2 \sum_{m=1}^{l} \frac{(l-m)!}{(l+m)!} P_{l}^{m}(\text{cos}\, \theta)\, P_{l}^{m}(\text{cos}\, \theta^\prime)\, \left[\text{cos}(m\phi)\, \text{cos}(m\phi^\prime) + \text{sin}(m\phi)\, \text{sin}(m\phi^\prime)\right].
+%% \end{align}
 the potential outside of the mass distribution can be written as:
@@
 $Q_l^{(0)}$ and $Q_{lm}^{(C,S)}$ are variants of the multipole moments that involve integrals of $P_l$ and $P_l^m$, respectively, over the computational domain.
-\begin{comment}
-\begin{align}
-  Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
-  Q_{lm}^{(C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
-  Q_{lm}^{(S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime.
-\end{align}
-\end{comment}
+%% \begin{align}
+%%   Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
+%%   Q_{lm}^{(C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
+%%   Q_{lm}^{(S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime.
+%% \end{align}
 This approach becomes computationally feasible when we cut off the
-outer summation in Equation \ref{multipole_potential} at some finite
+outer summation in \autoref{multipole_potential} at some finite
 value of $l_{\text{max}}$. If it is of sufficiently high order, we
@@
 white dwarfs of mass ratio 2/3, loaded onto the grid by the method described 
-in Section \ref{sec:initial_models}. We terminated the simulation just after
+in \autoref{sec:initial_models}. We terminated the simulation just after
 initialization, so that we perform only an initial Poisson solve for this 
@@
 from 0 to 20, and we also did this using the numerically exact solution 
-provided by Equation \ref{eq:direct_sum}.  Defining the $L^2$
+provided by \autoref{eq:direct_sum}.  Defining the $L^2$
 norm of a field $f$ as
@@
 \end{equation}
-The result is shown in Figure \ref{fig:bc_comparison}. At $l_{\text{max}} = 6$,
+The result is shown in \autoref{fig:bc_comparison}. At $l_{\text{max}} = 6$,
 the error is already well below $10^{-4}$ and we adopt this as our default 
@@
 
-The results of this test are plotted in Figure
-\ref{fig:gravity_convergence}. We find that at low resolution 
+The results of this test are plotted in \autoref{fig:gravity_convergence}. 
+We find that at low resolution 
 convergence is actually substantially better than second-order. The 
@@
 
-We can solve one of these two sources of error by evaluating Equation
-\ref{eq:sphere-analytical} with a mass $M$ that corresponds to the
+We can solve one of these two sources of error by evaluating 
+\autoref{eq:sphere-analytical} with a mass $M$ that corresponds to the
 amount of mass actually on the grid, defined as the sum of the density
@@
 greater than ambient. The resolution study for this case is also
-plotted in Figure \ref{fig:gravity_convergence}. We still obtain
+plotted in \autoref{fig:gravity_convergence}. We still obtain
 convergence slightly better than second-order, indicating that we 
@@
 \end{align}
-where $x_0 = R + x$, $x_1 = R - x$, $y_0 = R + y$, $y_1 = R-y$, $z_0 =
-R+z$, $z_1 = R-z$, and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
+where $x_0 = R + x$, $x_1 = R - x$, $y_0 = R + y$, 
+$y_1 = R - y$, $z_0 = R + z$, $z_1 = R - z$, 
+and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
 \texttt{Fortran} and \texttt{C}, the inverse hyperbolic tangent is
@@
 cube. The normalized $L^2$ error for this problem is also shown
-in Figure \ref{fig:gravity_convergence}, and only for this problem 
+in \autoref{fig:gravity_convergence}, and only for this problem 
 do we obtain perfect second-order scaling at all resolutions.
@@
 and a rotating reference frame is employed, excellent conservation
-properties can result, as demonstrated in Section \ref{sec:kepler}. 
+properties can result, as demonstrated in \autoref{sec:kepler}. 
 In particular, at reasonable resolution for a binary orbit our code 
@@
 Given that we can write down a potential energy for the rotation field, then we can use the machinery of 
-Section \ref{sec:gravity_hydro_coupling}. We again continue to evolve explicitly an equation for 
+\autoref{sec:gravity_hydro_coupling}. We again continue to evolve explicitly an equation for 
 the gas energy, and allow it to change in response to work done by or on the rotational potential.
@@
 not present in single-degenerate Type Ia and core-collapse supernova
-simulations. In Section \ref{sec:gravity}, we discussed how the lack
+simulations. In \autoref{sec:gravity}, we discussed how the lack
 of spherical symmetry necessitates a careful look at the gravity
@@
 
-In Section \ref{sec:initial_models} we describe the process by which
+In \autoref{sec:initial_models} we describe the process by which
 we generate initial stellar models. While the 1D models are in
@@
 radius of the star, determined by the volume of the grid that has a density 
-greater than $10^3\ \text{g cm}^{-3}$ (see Section \ref{sec:initial_models} 
+greater than $10^3\ \text{g cm}^{-3}$ (see \autoref{sec:initial_models} 
 for details on this measure). This choice of density is intended to 
 mark a reasonable outer edge to the star that is not immediately susceptible 
-to the numerical errors prevalent near the physical edge of the star. Figure 
-\ref{fig:single_star_static_radius} shows our results at various resolutions. 
+to the numerical errors prevalent near the physical edge of the star. 
+\autoref{fig:single_star_static_radius} shows our results at various resolutions. 
 As expected, the star quickly approaches an equilibrium size that is different 
@@
   \caption{Time evolution of the kinetic energy of a $0.9 \msolar$ 
-    white dwarf. The lines have the same meaning as in Figure \ref{fig:single_star_static_radius}.
+    white dwarf. The lines have the same meaning as in \autoref{fig:single_star_static_radius}.
     \label{fig:single_star_static_ke}}
@@
 the case of gravitational free fall. We place two stars on the grid 
-in the manner of Section \ref{sec:initial_models}. The distance $a$ between 
+in the manner of \autoref{sec:initial_models}. The distance $a$ between 
 them corresponds to a chosen orbital period $T$, consistent with the total
@@
 starts to become important. The results of our simulation for our default $256^3$ zone 
-uniform grid are shown in Figure \ref{Fig:Free Fall}. They show excellent agreement
+uniform grid are shown in \autoref{fig:freefall}. They show excellent agreement
 between the theory and the simulation results, and this agreement holds even for 
@@
     show only a small fraction of the timesteps.}
-  \label{Fig:Free Fall}
+  \label{fig:freefall}
 \end{figure}
@@
   \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
-    conditions given by Equations \ref{eq:kh_ic_b_ramp} and \ref{eq:kh_ic_b}. 
+    conditions given by \autoref{eq:kh_ic_b_ramp} and \autoref{eq:kh_ic_b}. 
     The rows each represent a different bulk fluid velocity $v$ and 
@@
   \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
-    conditions given by Equations \ref{eq:kh_p3_rho} through \ref{eq:kh_p3_vy}. 
-    The meaning of the rows and columns is the same as in Figure \ref{fig:kh_p2}.
+    conditions given by \autoref{eq:kh_p3_rho} through \autoref{eq:kh_p3_vy}. 
+    The meaning of the rows and columns is the same as in \autoref{fig:kh_p2}.
     \label{fig:kh_p3}}
@@
 R10 so we do not show them here. For IC B, our results can be seen 
-for the normal resolutions and all velocities in Figure \ref{fig:kh_p2}.
+for the normal resolutions and all velocities in \autoref{fig:kh_p2}.
 At low resolutions and very large bulk velocities, the fluid 
@@
 instabilities at the stopping time. This is evident even at low resolutions
-by examining the first row of Figure \ref{fig:kh_p2}.
+by examining the first row of \autoref{fig:kh_p2}.
 
@@
 and bulk velocities as the previous two problems. The results for the normal resolutions 
-at $t = 2.0$ are displayed in Figure \ref{fig:kh_p3}. We see a similar pattern as for 
+at $t = 2.0$ are displayed in \autoref{fig:kh_p3}. We see a similar pattern as for 
 the test proposed by R10: as we get to higher flow speeds we need to have higher 
@@
 any conclusions that bear on the question of verification from this test at late times. 
-Figure \ref{fig:kh_p3_high_res} provides a sense of this by examining the crucial 
+\autoref{fig:kh_p3_high_res} provides a sense of this by examining the crucial 
 time at which the transition from the linear to the non-linear regime is occurring. 
@@
 
-To see this effect in action for a stellar simulation, we repeated the test of Section 
-\ref{sec:HSE} with a bulk velocity on the grid. We chose a velocity of $2.56 \times 10^{8}\ \text{cm s}^{-1}$.
-For context, this is comparable to the orbital velocities of the stars in Section \ref{sec:kepler}.
+To see this effect in action for a stellar simulation, we repeated the test of  
+\autoref{sec:HSE} with a bulk velocity on the grid. We chose a velocity of $2.56 \times 10^{8}\ \text{cm s}^{-1}$.
+For context, this is comparable to the orbital velocities of the stars in \autoref{sec:kepler}.
 This test was inspired by \citet{tasker:2008}, who considered a moving galaxy cluster 
@@
   \includegraphics[scale=0.45]{plots/single_star_compare_1e3_radius}
-  \caption{A variation on Figure \ref{fig:single_star_static_radius} where
+  \caption{A variation on \autoref{fig:single_star_static_radius} where
     we now compare the ``static'' case to ``motion'' simulations where the 
@@
 
-In Figure \ref{fig:single_star_compare_radius}, we take the results of Section 
-\ref{sec:HSE} (the ``static'' case), and plot on top of it the results of this 
+In \autoref{fig:single_star_compare_radius}, we take the results of 
+\autoref{sec:HSE} (the ``static'' case), and plot on top of it the results of this 
 new simulation (the ``motion'' case). We see immediately that this bulk velocity
@@
 For some of the algorithms described earlier in this work, a single orbit of these 
-systems is enough to examine their effects. In Section \ref{sec:gravity_boundary_conditions},
+systems is enough to examine their effects. In \autoref{sec:gravity_boundary_conditions},
 we discussed the replacement of a monopole boundary condition solver for the gravitational 
@@
 We performed this test for maximum multipole moments ranging from 0 (the monopole term) to 16.
-The results are shown in Figure \ref{fig:gravity_bcs}. Terms in the boundary potential 
+The results are shown in \autoref{fig:gravity_bcs}. Terms in the boundary potential 
 that vary faster than $r^{-5}$ are effectively negligible in determining the outcome of the orbit, 
@@
 
-Another that we consider is the energy conservation of the system. Recalling Section \ref{sec:gravity_hydro_coupling}, 
+Another that we consider is the energy conservation of the system. Recalling \autoref{sec:gravity_hydro_coupling}, 
 there are several different methods of applying the gravitational source term to the hydrodynamics equations. In CASTRO we presently 
@@
 the new conservative method of evaluating the energy source terms at cell faces. The results for the change in energy 
-after a single orbit are seen in the first column of Table \ref{table:sources}. The first two versions give reasonable and 
+after a single orbit are seen in the first column of \autoref{table:sources}. The first two versions give reasonable and 
 similar levels of energy conservation. The third has total energy changes on the order of 100\%, but this itself 
@@
 
-In Table \ref{table:sources} we show also the effects on energy conservation of using the inertial reference frame. 
+In \autoref{table:sources} we show also the effects on energy conservation of using the inertial reference frame. 
 We use ${\tt rs}$ for the CASTRO parameter {\tt castro.rot\_source\_type}, and we have the same four options for updating
@@
 
+\capstartfalse
 \input{plots/sources.table}
+\capstarttrue
 
 We are most interested in the stability of these systems over long timescales. To this end, we consider the same systems as above, 
-but evolve them for 25 orbital periods. In Figure \ref{fig:circular_orbit_comparison} we illustrate the evolution of these systems 
+but evolve them for 25 orbital periods. In \autoref{fig:circular_orbit_comparison} we illustrate the evolution of these systems 
 by plotting the center of mass locations of the white dwarfs on the orbital ($xy$) plane. For the equal mass case in the inertial 
@@
 Turning to the conservation properties of the system, we examine as fairly typical cases the equal mass system 
-in the inertial frame for energy conservation (Figure \ref{fig:energy_conservation_equal}), and the unequal mass 
-system in the rotating frame for angular momentum conservation (Figure \ref {fig:angular_momentum_conservation_unequal}). 
+in the inertial frame for energy conservation (\autoref{fig:energy_conservation_equal}), and the unequal mass 
+system in the rotating frame for angular momentum conservation (\autoref{fig:angular_momentum_conservation_unequal}). 
 Angular momentum is conserved to within a few percent over the 25 orbits, while energy conservation is about 
@@
 Supercomputing Applications. For the weak scaling test, whose results are 
-shown in Figure \ref{fig:weak_scaling}, we ran a uniform grid binary white 
+shown in \autoref{fig:weak_scaling}, we ran a uniform grid binary white 
 dwarf simulation for resolutions of $128^3$ zones through $2048^3$ zones. The 
@@
 to be achieved from further parallelism. The results are found in 
-\ref{fig:strong_scaling}. We find excellent scaling for low numbers 
+\autoref{fig:strong_scaling}. We find excellent scaling for low numbers 
 of processors, though the long wall time per timestep required for the 
@@
 For all the runs, the PPM reconstruction is done using the original limiters for
-the parabolic profiles \citep{ppm}; see Section \ref{sec:Hydrodynamics} for a brief 
+the parabolic profiles \citep{ppm}; see \autoref{sec:Hydrodynamics} for a brief 
 discussion about the limiters.  The prediction of the interface
@@
 (density, velocity in the normal direction, and pressure), and let our 
-reference state be as in Eq.~\ref{eq:refchoice}.
+reference state be as in Eq.~\autoref{eq:refchoice}.
 Ignoring flattening, if there are no waves moving toward our
-interface, then Eq.~\ref{eq:ppmstatel} reduces to:
+interface, then Eq.~\autoref{eq:ppmstatel} reduces to:
 \begin{equation}
@@
 only the term corresponding to the fastest wave in the sum will be
-added in Eq.~\ref{eq:ppmstatel}, but our choice of reference state makes
+added in Eq.~\autoref{eq:ppmstatel}, but our choice of reference state makes
 that term zero by design, and our interface state is:
@@
 If we use the same idea of the reference state as
-in Eq.~\ref{eq:refchoice}, and consider a quantity $\xi$ which should only
+in Eq.~\autoref{eq:refchoice}, and consider a quantity $\xi$ which should only
 jump across the contact, then our interface state becomes:

### 2015-07-29 papers/wdmerger_I/paper.tex
Add references and acknowledgements for a few more community resources.

@@
 we store the active \texttt{git} commit hashes of BoxLib, CASTRO and \texttt{wdmerger}.
+Line plots are generated using the \texttt{matplotlib} library for \texttt{Python} 
+\cite{matplotlib}, while slice plots and other multi-dimensional visualizations are 
+generated using the \cite{yt} code.
 
@@
 
+This research has made use of the AstroBetter blog and wiki. 
+This research has made use of NASA's Astrophysics Data System 
+Bibliographic Services.
+
 This research was supported by NSF award AST-1211563. An
@@
 (density, velocity in the normal direction, and pressure), and let our 
-reference state be as in Eq.~\autoref{eq:refchoice}.
+reference state be as in \autoref{eq:refchoice}.
 Ignoring flattening, if there are no waves moving toward our
-interface, then Eq.~\autoref{eq:ppmstatel} reduces to:
+interface, then \autoref{eq:ppmstatel} reduces to:
 \begin{equation}
@@
 only the term corresponding to the fastest wave in the sum will be
-added in Eq.~\autoref{eq:ppmstatel}, but our choice of reference state makes
+added in \autoref{eq:ppmstatel}, but our choice of reference state makes
 that term zero by design, and our interface state is:
@@
 If we use the same idea of the reference state as
-in Eq.~\autoref{eq:refchoice}, and consider a quantity $\xi$ which should only
+in \autoref{eq:refchoice}, and consider a quantity $\xi$ which should only
 jump across the contact, then our interface state becomes:

### 2015-07-30 papers/wdmerger_I/paper.tex
Cleanup and standardization of the references

@@
 in the discretization of the gravitational potential can lead to very large changes in the energy of the gas.
-(We note in passing too that although \cite{jiang:2013} claim their approach requires a Poisson 
+(We note in passing too that although \citeauthor{jiang:2013}\ claim their approach requires a Poisson 
 solve for the time derivative of the potential, $\dot{\Phi}$, this is not needed as in practice 
@@
     $v=100$, corresponds to approximately Mach 70. Compare to 
-    \citet{robertson:2010}, Figure 7. \label{fig:kh_p2}}
+    Robertson et al. (2010), Figure 7. \label{fig:kh_p2}}
 \end{figure*}
@@
   \includegraphics[trim = 0in 1.25in 0.25in 1.25in, clip, scale=0.65]{{{plots/kh_p3_high_res_collage}}}
-  \caption{Time series of the Kelvin-Helmholtz problem proposed by \citet{mcnally:2012}
+  \caption{Time series of the Kelvin-Helmholtz problem proposed by McNally et al. (2012)
     as the simulation is just starting to go non-linear. The rows represent resolution, 
@@
 
-\citet{hopkins:2014} performed this test as part of the testing of their code GIZMO. 
+\citet{hopkins:2015} performed this test as part of the testing of their code GIZMO. 
 They showed the late-time evolution of this system, when non-linear effects have 
@@
 
-\bibliographystyle{../apj}
-\bibliography{../refs}
-
-\clearpage
 \appendix
@@
 
+\bibliographystyle{../apj}
+\bibliography{../refs}
+
+\clearpage
+
 \end{document}

### 2015-07-31 papers/wdmerger_I/paper.tex
Fix a weird space after a figure
Rotating the strain tensor itself doesn't work. Instead, what we need to do is rotate positions and velocities first, and only then evaluate the quadrupole moment tensor.

@@
 we choose a relative error tolerance of $10^{-10}$ to be satisfied in the multigrid solve.
-
-\begin{figure}
+\begin{figure}[h]
   \centering
@@
 \end{figure}
-
 The results of this test are plotted in \autoref{fig:gravity_convergence}.
+
 We find that at low resolution 
@@
 using an integral over the domain (using refined grids as usual by masking out the contributions from 
-coarse grids underlying more accurate fine grids). This quantity is independent of the observer. Then, 
+coarse grids underlying more accurate fine grids). This quantity is independent of the observer. If we 
+are using a rotating reference frame, we first convert velocities and positions back to the inertial 
+frame before evaluating the integral. Then, 
 we pick an observing location $\mathbf{x}$ relative to the domain, evaluate the projection operator, 

### 2015-07-31 source/Problem.f90

@@ -377,11 +377,12 @@ subroutine quadrupole_tensor_double_dot(rho,  r_l1, r_l2, r_l3, r_h1, r_h2, r_h3
   double precision, intent(in   ) :: dx(3), time
   double precision, intent(inout) :: Qtt(3,3)
 
-  integer          :: i, j, k, l, m
+  integer          :: i, j, k, l, m, dir
   double precision :: dV
-  double precision :: r(3), vel(3), grav(3), rhoInv
+  double precision :: r(3), pos(3), vel(3), grav(3), rhoInv
   double precision :: dQtt(3,3)
   double precision :: omega(3)
+  double precision :: theta(3), rot_matrix(3,3)
 
   dV = dx(1) * dx(2) * dx(3)
   dQtt(:,:) = ZERO
@@ -391,7 +392,35 @@ subroutine quadrupole_tensor_double_dot(rho,  r_l1, r_l2, r_l3, r_h1, r_h2, r_h3
   else
      omega = ZERO
   endif
-  
+
+  ! If we are in a rotating reference frame, then we want
+  ! to rotate this by an amount corresponding to the time
+  ! that has passed since the beginning of the simulation.
+  ! This way it will correspond to an inertial observer
+  ! at large distances. There is some clipping due to material
+  ! that leaves or enters the domain boundaries, but this should
+  ! average out over time, and will be negligible in magnitude.
+        
+  theta = omega * time
+
+  ! This is the 3D rotation matrix for converting between reference frames.
+  ! It is the composition of rotations along the x, y, and z axes. Therefore 
+  ! it allows for the case where we are rotating about multiple axes. Normally 
+  ! we use the right-hand convention for constructing the usual rotation matrix, 
+  ! but this is the transpose of that rotation matrix to account for the fact 
+  ! that we are rotating *back* to the inertial frame, rather than from the 
+  ! inertial frame to the rotating frame.
+
+  rot_matrix(1,1) =  cos(theta(2)) * cos(theta(3))
+  rot_matrix(1,2) = -cos(theta(2)) * sin(theta(3))
+  rot_matrix(1,3) =  sin(theta(2))
+  rot_matrix(2,1) =  cos(theta(1)) * sin(theta(3)) + sin(theta(1)) * sin(theta(2)) * cos(theta(3))
+  rot_matrix(2,2) =  cos(theta(1)) * cos(theta(3)) - sin(theta(1)) * sin(theta(2)) * sin(theta(3))
[…]

### 2015-07-31 papers/wdmerger_I/paper.tex
Define the projection operator in the strain tensor definition, from BDS.

@@
 respect to the observer, and $P_{ikjl}(\mathbf{n})$ is an operator in the transverse-traceless 
-gauge that projects a tensor onto the direction orthogonal to $\mathbf{n}$. $Q_{kl}$ is the 
-quadrupole moment tensor:
+gauge that projects a tensor onto the direction orthogonal to $\mathbf{n}$:
+\begin{align}
+  P_{ijkl}(\mathbf{n}) &= \left(\delta_{ik} - n_i n_k\right)\left(\delta_{jl} - n_j n_l\right) \notag \\
+                      &- \frac{1}{2}\left(\delta_{ij} - n_i n_j\right)\left( \delta_{kl} - n_k n_l\right).
+\end{align}
+$Q_{kl}$ is the quadrupole moment tensor:
 \begin{equation}
@@
 the types of spatial distortions they exhibit. We calculate the signal at a distance $r$ along 
-the $x$, $y$ and $z$ axes. For the latter, as an example, $h_{+} = h_{11} = -h_{22}$ and 
-$h_{\times} = h_{12} = h_{21}$. All other entries vanish. By default we take $r = 10$ kpc; 
+the $x$, $y$ and $z$ axes. For the latter, as an example, $h_{+} = h_{11} = -h_{22} \propto (\ddot{Q}_{11} - \ddot{Q}_{22})/2$ and 
+$h_{\times} = h_{12} = h_{21} \propto \ddot{Q}_{12}$. All other entries vanish. By default we take $r = 10$ kpc; 
 as shown by \citet{loren-aguilar:2005}, this is a typical distance scale over which an 
@@
 Line plots are generated using the \texttt{matplotlib} library for \texttt{Python} 
-\cite{matplotlib}, while slice plots and other multi-dimensional visualizations are 
-generated using the \cite{yt} code.
+\citep{matplotlib}, while slice plots and other multi-dimensional visualizations are 
+generated using the \texttt{yt} code \citep{yt}.
 

### 2015-08-08 papers/wdmerger_I/paper.tex
Add GitHub links for BoxLib and Castro.

@@
 To study the white dwarf merger problem, we use the grid-based
-hydrodynamics code CASTRO \citep{castro}. CASTRO solves the Euler
+hydrodynamics code CASTRO\footnote{CASTRO can be obtained at \url{https://github.com/BoxLib-Codes/Castro}.} \citep{castro}.
+CASTRO solves the Euler
 equations, along with the inclusion of optional modules for gravity,
 nuclear reactions and thermodynamics. CASTRO is based on the BoxLib
+\footnote{BoxLib can be obtained at \url{https://github.com/BoxLib-Codes/BoxLib}.}
 adaptive-mesh refinement (AMR) framework \citep{rendleman:2000}, which
@@
 (as well as the manuscript itself),
-\texttt{wdmerger}\footnote{\url{https://github.com/BoxLib-Codes/wdmerger}},
+\texttt{wdmerger}\footnote{\texttt{wdmerger} can be obtained at \url{https://github.com/BoxLib-Codes/wdmerger}.},
 is freely available at an online repository hosting service.

### 2015-08-08 papers/wdmerger_I/paper.tex

--- 2015-07-31 papers/wdmerger_I/paper.tex
@@
 \newcommand{\msolar}{\mathrm{M}_\odot}
+
+% Software names
+\newcommand{\boxlib}{\texttt{BoxLib}}
+\newcommand{\castro}{\texttt{CASTRO}}
+\newcommand{\wdmerger}{\texttt{wdmerger}}
+\newcommand{\python}{\texttt{Python}}
+\newcommand{\matplotlib}{\texttt{matplotlib}}
+\newcommand{\yt}{\texttt{yt}}
 
@@
 using the adaptive-mesh refinement approach. In this paper we describe our numerical 
-methodology and discuss our implementation in the code CASTRO. CASTRO 
+methodology and discuss our implementation in the code \castro. \castro\ 
 solves the Euler equations of hydrodynamics 
@@
 To study the white dwarf merger problem, we use the grid-based
-hydrodynamics code CASTRO \citep{castro}. CASTRO solves the Euler
+hydrodynamics code \castro\footnote{\castro\ can be obtained at \url{https://github.com/BoxLib-Codes/Castro}.} \citep{castro}.
+\castro\ solves the Euler
 equations, along with the inclusion of optional modules for gravity,
-nuclear reactions and thermodynamics. CASTRO is based on the BoxLib
+nuclear reactions and thermodynamics. \castro\ is based on the \boxlib
+\footnote{\boxlib\ can be obtained at \url{https://github.com/BoxLib-Codes/BoxLib}.}
 adaptive-mesh refinement (AMR) framework \citep{rendleman:2000}, which
 represents fluid data on a mesh where regions of interest have higher
-spatial resolution. CASTRO is highly parallel and is designed for
+spatial resolution. \castro\ is highly parallel and is designed for
 large-scale use on modern supercomputers; see 
-\autoref{sec:Performance} for information on how CASTRO performs for our
+\autoref{sec:Performance} for information on how \castro\ performs for our
 problem. The next few subsections describe our approach to each of the
 physics components used in this work. We direct the reader to the
-original code paper for a full description of CASTRO's approach to
+original code paper for a full description of \castro's approach to
 solving the equations of hydrodynamics. In this work, we report mainly
@@
 
-We use the unsplit piecewise-parabolic method (PPM) solver in CASTRO
+We use the unsplit piecewise-parabolic method (PPM) solver in \castro\
 to advance the hydrodynamics system in time \citep{ppmunsplit}.  A
@@
 Appendix.  These changes bring the algorithm more in line with that of
-\cite{ppm}. CASTRO as originally released featured a slightly modified
+\cite{ppm}. \castro\ as originally released featured a slightly modified
 version of the higher resolution limiters of
@@
 
-CASTRO's approach to adaptive mesh refinement, based on its underlying
-BoxLib framework, is to refine zones based on certain user-specified
+\castro's approach to adaptive mesh refinement, based on its underlying
+\boxlib\ framework, is to refine zones based on certain user-specified
 criteria that tag regions of interest for higher spatial
@@
 assign this floor value to the temperature and let that zone be
-thermodynamically inconsistent (the original behavior in CASTRO), or
+thermodynamically inconsistent (the original behavior in \castro), or
 we can adjust the internal energy to be thermodynamically consistent
@@
 threshold is typically very small in an individual zone (a relative 
-difference of $10^{-8}$ by default in CASTRO), over time and given 
+difference of $10^{-8}$ by default in \castro), over time and given 
 a large number of zones, this can result in a significant energy 
@@
 
-CASTRO has the ability to do both nuclear reactions and radiative 
+\castro\ has the ability to do both nuclear reactions and radiative 
 transport (in the flux-limited diffusion approximation). For all simulations 
@@
 where $\Phi$ is the gravitational potential, $G$ is the gravitational
-constant, and $\rho$ is the mass density.\footnote{In the CASTRO code, the 
+constant, and $\rho$ is the mass density.\footnote{In the \castro\ code, the 
 right-hand side is negated and therefore $\Phi$ is positive. We use the 
@@
 When $\Phi$ appears in the code it is negated to compensate for this.} 
-The solution of this equation in CASTRO is described in \cite{castro}, and
+The solution of this equation in \castro\ is described in \cite{castro}, and
 consists of both level and composite solves, and a final
@@
 
-Time centering of this source term is done in CASTRO using a predictor-corrector approach.
+Time centering of this source term is done in \castro\ using a predictor-corrector approach.
 At the start of a coarse grid timestep, we solve the gravitational potential for the density $\rho^n$.
@@
 in the hydrodynamics update, so we need only to use a time-centered $\Phi$ (which can be computed 
-by averaging the time-level $n$ and $n+1$ potentials; we already have the latter because CASTRO 
+by averaging the time-level $n$ and $n+1$ potentials; we already have the latter because \castro\ 
 re-computes the potential at the new time after the hydrodynamics step, and we can apply this energy 
@@
 A final issue worth noting relates to complications posed by mesh refinement. Like 
-many AMR codes, CASTRO implements a refluxing algorithm at the end of a timestep. 
+many AMR codes, \castro\ implements a refluxing algorithm at the end of a timestep. 
 At any locations on a particular grid level that are adjacent to, but not covered by,
@@
   with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in
-CASTRO. If there are $N$ zones per spatial dimension, then there are
+\castro. If there are $N$ zones per spatial dimension, then there are
 $6 N^2$ boundary zones, and each boundary zone requires a sum over
@@
 method of approximation to this direct summation is justified. Many
-approaches exist in the literature. The original release of CASTRO
+approaches exist in the literature. The original release of \castro\
 featured the crudest possible approximation: a monopole prescription,
@@
 that inserts a sphere of radius $R$ and uniform mass density $\rho$
-onto our grid, and used CASTRO to calculate the gravitational
+onto our grid, and used \castro\ to calculate the gravitational
 potential $\Phi$ of this setup. We ensure that $R$ is an integer
@@
   \includegraphics[scale=0.45]{plots/phi_comparison}
-  \caption{Comparison of the CASTRO gravitational potential to the analytical solution for: 
+  \caption{Comparison of the \castro\ gravitational potential to the analytical solution for: 
     a sphere of uniform density; the same sphere, but with the potential normalized using the 
@@
 two stars in a frame that is co-rotating at the same period as the
-orbital period. Since the publication of the original code paper, CASTRO 
+orbital period. Since the publication of the original code paper, \castro\ 
 now has the ability to evolve systems in a rotating reference frame. 
@@
 \end{align}
-We use this form of the momentum update in CASTRO. This improvement is small
+We use this form of the momentum update in \castro. This improvement is small
 but increases the accuracy of our rotating white dwarf systems over long time-scales.
@@
 (as well as the manuscript itself),
-\texttt{wdmerger}\footnote{\url{https://github.com/BoxLib-Codes/wdmerger}},
+\wdmerger\footnote{\wdmerger\ can be obtained at \url{https://github.com/BoxLib-Codes/wdmerger}.},
 is freely available at an online repository hosting service.
-Version control in both the parent software (BoxLib, CASTRO) and in \texttt{wdmerger}
+Version control in both the parent software (\boxlib, \castro) and in \wdmerger\
 permits us to reference the state of the code at the time a simulation
-was performed. In all plot files and diagnostic output generated by CASTRO, 
-and EPS figure files and plots generated by \texttt{wdmerger} for this work,
-we store the active \texttt{git} commit hashes of BoxLib, CASTRO and \texttt{wdmerger}.
-Line plots are generated using the \texttt{matplotlib} library for \texttt{Python} 
+was performed. In all plot files and diagnostic output generated by \castro, 
+and EPS figure files and plots generated by \wdmerger\ for this work,
+we store the active \texttt{git} commit hashes of \boxlib, \castro, and \wdmerger.
+Line plots are generated using the \matplotlib\ library for \python\ 
 \citep{matplotlib}, while slice plots and other multi-dimensional visualizations are 
-generated using the \texttt{yt} code \citep{yt}.
+generated using the \yt\ code \citep{yt}.
 
@@
 co-rotating frame. It is therefore important to be be aware of the
-behavior of the code in such circumstances. The behavior of CASTRO for
+behavior of the code in such circumstances. The behavior of \castro\ for
 many standard hydrodynamics test problems was detailed in the original
@@
 
-This problem was studied in the first CASTRO paper, but is worth
+This problem was studied in the first \castro\ paper, but is worth
 revisiting here. A single star explosion simulation may only last a
-couple of seconds, and the CASTRO paper studied the behavior of the
+couple of seconds, and the \castro\ paper studied the behavior of the
 star after one second of evolution. However, the dynamical timescale
@@
 
-A simple dynamical test to verify the gravitation physics in CASTRO is
+A simple dynamical test to verify the gravitation physics in \castro\ is
 the case of gravitational free fall. We place two stars on the grid 
@@
     calculated from Newtonian mechanics, and the circles show the
-    samples from the time evolution with CASTRO. For visual clarity, we 
+    samples from the time evolution with \castro. For visual clarity, we 
     show only a small fraction of the timesteps.}
@@
 the importance of this effect when deciding whether to trust the 
-output of a code like CASTRO when applied for merger problems.
+output of a code like \castro\ when applied for merger problems.
 Recently, this has come up in two ways which are of note for us 
@@
 ave important consequences for how much we can trust the ability of 
-CASTRO to test the violent merger progenitor model. Shearing between 
+\castro\ to test the violent merger progenitor model. Shearing between 
 the material flowing out of the secondary and material near the 
@@
 and $y$ directions. This is a two-dimensional test, so we run
-CASTRO in 2D mainly to avoid extra computational expense; in 3D, it 
+\castro\ in 2D mainly to avoid extra computational expense; in 3D, it 
 would merely involve replicating the problem in the $z$ direction.
@@
 Another that we consider is the energy conservation of the system. Recalling \autoref{sec:gravity_hydro_coupling}, 
-there are several different methods of applying the gravitational source term to the hydrodynamics equations. In CASTRO we presently 
+there are several different methods of applying the gravitational source term to the hydrodynamics equations. In \castro\ we presently 
 have four options, controlled by the parameter {\tt castro.grav\_source\_type}, which we will shortern to {\tt gs} for the present 
@@
 In \autoref{table:sources} we show also the effects on energy conservation of using the inertial reference frame. 
-We use ${\tt rs}$ for the CASTRO parameter {\tt castro.rot\_source\_type}, and we have the same four options for updating
+We use ${\tt rs}$ for the \castro\ parameter {\tt castro.rot\_source\_type}, and we have the same four options for updating
 the energy source term as for gravity. The ${\tt rs} = 0$ column means that rotation is turned off and we are 
@@
 
-CASTRO is designed to be deployed on high-performance computing systems using 
+\castro\ is designed to be deployed on high-performance computing systems using 
 many thousands of processors simultaneously. It is worth briefly examining 
@@
 This is especially true because some aspects of our approach to parallelism 
-have changed since the first CASTRO paper, and we have obtained improved 
+have changed since the first \castro\ paper, and we have obtained improved 
 performance in certain settings.
 
-The BoxLib framework that CASTRO is based on splits each AMR level into a number 
+The \boxlib\ framework that \castro\ is based on splits each AMR level into a number 
 of boxes that collectively span the level. These boxes are distributed to processors 
@@
 obeys a load-balancing algorithm that attempts to equalize the amount of work 
-done by each processor. BoxLib contains a number of strategies for distributing 
+done by each processor. \boxlib\ contains a number of strategies for distributing 
 work in this way, and by default uses a space-filling curve approach with a 
@@
 We use OpenMP to accelerate the work associated with the boxes owned by each 
-MPI task. Originally CASTRO used OpenMP to accelerate individual loops in the 
+MPI task. Originally \castro\ used OpenMP to accelerate individual loops in the 
 hydrodynamics routines, such as the piecewise-parabolic edge state reconstruction 
@@
   \includegraphics[scale=0.4]{plots/weak_scaling}
-  \caption{CASTRO weak scaling test, performed on the Blue Waters machine at 
+  \caption{\castro\ weak scaling test, performed on the Blue Waters machine at 
     NCSA. Each processor had a fixed amount of work, and we increased the 
@@
     solid curve represents perfect weak scaling, while the blue circles show 
-    CASTRO's performance at each processor count. The vertical axis measures 
+    \castro's performance at each processor count. The vertical axis measures 
     the median time per timestep, normalized to this value for the smallest 
@@
 
-To examine the parallel performance of CASTRO, we performed both strong scaling 
+To examine the parallel performance of \castro, we performed both strong scaling 
 and weak scaling tests on the Blue Waters machine at the National Center for 
@@
 that this weak scaling behavior with Poisson gravity is a significant 
-improvement over the results presented in the first CASTRO paper.
+improvement over the results presented in the first \castro\ paper.
 
@@
   \includegraphics[scale=0.4]{plots/strong_scaling}
-  \caption{CASTRO strong scaling test performed on the Blue Waters machine at
+  \caption{\castro\ strong scaling test performed on the Blue Waters machine at
     NCSA. The vertical axis measures the median time per timestep, and the 
@@
 
-The basic PPM algorithm in CASTRO has undergone a number of changes
+The basic PPM algorithm in \castro\ has undergone a number of changes
 since the original code paper \citep{castro}.  A discussion of the
-pure hydrodynamics changes along with verification of CASTRO when
+pure hydrodynamics changes along with verification of \castro\ when
 using the stellar equation of state was given in
@@
 
-Since the original CASTRO paper, the reference state implementation
+Since the original \castro\ paper, the reference state implementation
 has been switched to:
@@
 in agreement with \citet{ppmunsplit} (eq. 90).  The flattening
-in the original CASTRO paper has also been updated as discussed
+in the original \castro\ paper has also been updated as discussed
 in \citet{zingalekatz:2015}.
@@
 We note a few additional differences between the original PPM
-implementation of \citet{ppm} and CASTRO.  In the original PPM
+implementation of \citet{ppm} and \castro\.  In the original PPM
 implementation, the gravitational acceleration was reconstructed as a parabola, and
 this was traced under to find the forcing that affects the interface
-for each wave.  CASTRO originally followed \citet{ppmunsplit} which instead adds
+for each wave.  \castro\ originally followed \citet{ppmunsplit} which instead adds
 $(\Delta t/2)g$ to the interface states for velocity at the end of the
@@
 parabolic reconstruction and characteristic tracing of the gravitation
-source. This can be controlled in CASTRO with the parameter 
+source. This can be controlled in \castro\ with the parameter 
 {\tt castro.ppm\_trace\_grav}. Our system with the source appears as:
@@
 the rotation source terms can be included in a manner identical to that of the gravity source
-terms, with the replacement of $G$ with $F_R$ and $g$ with $f_R$ in the above. This is controlled in CASTRO
+terms, with the replacement of $G$ with $F_R$ and $g$ with $f_R$ in the above. This is controlled in \castro\
 with the parameter {\tt castro.ppm\_trace\_rot}.

### 2015-08-10 papers/wdmerger_I/paper.tex
Add gravitational wave verification to paper.

@@
 
+As a simple verification test to ensure our gravitational wave calculations are correct, we plot the 
+gravitational wave strain along the rotation axis for the first two periods of the unequal mass system. 
+At this early time the orbit is circular and so to a good approximation we expect that the gravitational 
+wave signal should be that of two point masses, whose positions are:
+\begin{align}
+  \mathbf{r}_P(t) &= -a_P\, \text{cos}(\omega t) \hat{x} - a_P\, \text{sin}(\omega t) \hat{y} \\
+  \mathbf{r}_S(t) &= a_S\, \text{cos}(\omega t) \hat{x} + a_S\, \text{sin}(\omega t) \hat{y}.
+\end{align}
+Then the mass distribution is $\rho(\mathbf{r}) = M_P\, \delta^3(\mathbf{r} - \mathbf{r}_P) + M_S\, \delta^3(\mathbf{r} - \mathbf{r}_S)$.
+From this it is straightforward to calculate the quadruopole tensor, take its second time derivative, and then apply the 
+projection operator to get the gravitational wave polarizations along the rotation axis:
+\begin{align}
+  h_+ &= -4\frac{G\mu}{c^4 r}\left[G M_{\text{tot}} \omega \right]^{2/3}\, \text{cos}(2\omega t) \\
+  h_\times &= -4\frac{G\mu}{c^4 r}\left[G M_{\text{tot}} \omega \right]^{2/3}\, \text{sin}(2\omega t).
+\end{align}
+$\mu$ is the reduced mass, while $M_{\text{tot}}$ is the total mass. From this we see that the 
+gravitational wave frequency is twice the orbital frequency, and that the two polarizations 
+are out of phase by $90^\circ$ in time. We compare this analytical expectation to the 
+numerical results in \autoref{fig:gw_strain}. We find very good agreement in this case, and this 
+level of agreement holds in the rotating frame as well.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/unequal_gw_rot0.eps}
+  \caption{Gravitational wave strain polarizations for the first two orbital periods of the 
+           unequal mass system. The curves with markers are the numerical data, while the 
+           curves without markers are the analytical results for two point masses.\label{fig:gw_strain}}
+\end{figure}
+
 %==========================================================================

### 2015-08-10 papers/wdmerger_I/paper.tex
Add the spatial convergence test to the paper.

@@
 
+Finally we consider whether the dynamical behavior of the system converges with resolution. 
+In \autoref{fig:unequal_spatial_convergence} we plot the first 10 seconds of the orbit for 
+the unequal mass system, at three different resolutions in the inertial frame: our default 
+resolution of $256^3$ zones, as well as a single level of refinement with a jump by a factor of 
+two (effective resolution $512^3$) or a jump by a factor of four (effective resolution $1024^3$). 
+It is clear that at the latter resolution (corresponding to physical resolution of 100 km), 
+we have achieved convergent behavior. In the rotating frame (not shown here), the results are 
+not quite the same but for these resolutions the results do tend toward convergence.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/spatial_convergence_rot0.eps}
+  \caption{Distance between the two white dwarfs in the unequal mass system, for the first 
+           10 seconds of their orbit. The distance is scaled by the initial orbital distance. 
+           We plot at three different resolutions, corresponding to the number of 
+           effective zones per dimension in the refined regions. Note that the $n = 1024$ curve
+           falls nearly exactly on top of the $n = 512$ curve.\label{fig:unequal_spatial_convergence}}
+\end{figure}
+
 %==========================================================================

### 2015-08-10 papers/wdmerger_I/paper.tex
A pass through the paper to clean some things up.

@@
 The Type Ia supernova progenitor problem is one of the most perplexing and 
-exciting problems in astrophysics, requiring detailed numerical modelling to 
+exciting problems in astrophysics, requiring detailed numerical modeling to 
 complement the observations of these explosions. One possibility that has 
@@
 problem and we discuss our implementation here. We also run an 
-extensive set of test problems to verify that our software can sufficiently
+extensive set of test problems to determine whether our software can sufficiently
 model a system where large amounts of mass are advected on the computational 
@@
 temperatures in these white dwarfs to become hot enough for explosive
-burning of its constituent nuclei. The model favored initially by the
-community was the so-called single-degenerate (SD) model
+burning of their constituent nuclei. The model favored initially by the
+community was the single-degenerate (SD) model
 \citep{whelan_iben:1973}. Accretion of material from a companion star
@@
 detonation scenario, where accretion of material onto a
-sub-Chandrasekhar white dwarf would lead to a detonation inside the
-accreted envelope, and this would send a compressional wave into the
-core of the star that would trigger a secondary detonation. A recent
+sub-Chandrasekhar mass white dwarf leads to a detonation inside the
+accreted envelope, sending a compressional wave into the
+core of the star that triggers a secondary detonation. A recent
 review of the progenitor models can be found in
@@
 There are several observational reasons why double-degenerate systems
-are a promising progenitor system for at least a substantial fraction
+are a promising progenitor model for at least a substantial fraction
 of normal SNe Ia. No conclusive evidence exists for a surviving
@@
 in this paper we do not enable either, and will delay discussion of 
-these modules until a later paper in this series.
+these modules until later papers in this series.
 
@@
 The solution of this equation in \castro\ is described in \cite{castro}, and
-consists of both level and composite solves, and a final
+consists of both level and composite solves, and (optionally) a final
 synchronization at the end.
@@
 \end{align}
-It is easy to verify that adding the source terms two cells to the left
-and right of this zone will result in complete cancellation of the source terms.
+It is easy to verify that adding the source terms for the current zone and the two zones 
+to the left and right will result in complete cancellation of the source terms.
 However, if the potential if non-zero outside of the domain, then there will be
@@
 just outside the domain. In addition, when we replace the Laplacian above by the full
-three-dimensional stencil including the $y$ and $z$ derivatives, and depending on the
+three-dimensional stencil including the $y$ and $z$ derivatives, depending on the
 discretization these will not be cancelled at all. This latter problem can be resolved by
@@
 the divergence of the gravitational acceleration is no longer guaranteed to equal
-the zone density. In particular, we find that the mixing of the gravitational components
+the zone density. In particular, we find that the mixing of the gravitational accleration components
 means that the truncation error in the gravitational field can lead to large errors
@@
 $\Delta t\, \rho^n\, \mathbf{g}^n$. After the hydrodynamics update, we recalculate
-the gravitational potential based on the new density, $\rho^n$, and then add 
+the gravitational potential based on the new density, $\rho^{n+1}$, and then add 
 $-(\Delta t/2) \rho^n \mathbf{g}^n + (\Delta t/2) \rho^{n+1} \mathbf{g}^{n+1}$ to the momenta.
@@
 
-Now the change in the mass $\Delta \rho_{f}$ is proportional to the mass flux through the surface with area $\Delta A$:
+The change in the mass $\Delta \rho_{f}$ is proportional to the mass flux through the surface with area $\Delta A$:
 \begin{equation}
@@
 It is worth pointing out that while \autoref{eq:grav_energy_conservation_update} was correctly obtained 
-by \cite{arepo} (and indeed that is the inspiration for this approach), there is a subtle error in that derivation. 
+by \cite{arepo} (whose work was the primary inspiration for this approach), there is a subtle error in that derivation. 
 Their argument was that if we integrate the work done over the volume of the zone, this volume integral can be 
@@
 in the discretization of the gravitational potential can lead to very large changes in the energy of the gas.
-(We note in passing too that although \citeauthor{jiang:2013}\ claim their approach requires a Poisson 
-solve for the time derivative of the potential, $\dot{\Phi}$, this is not needed as in practice 
-we will already have the potential at time-levels $n$ and $n+1$, and to second order in time, 
-$\dot{\Phi}^{n+1/2} = (\Phi^{n+1} - \Phi^{n}) / \Delta t.$)
-
-A final issue worth noting relates to complications posed by mesh refinement. Like 
-many AMR codes, \castro\ implements a refluxing algorithm at the end of a timestep. 
-At any locations on a particular grid level that are adjacent to, but not covered by,
-zones from a finer grid, there will be a disagreement about the fluxes crossing 
-this coarse-fine interface. If left uncorrected, this results in a violation of 
-the conservation condition. Therefore the value of the coarse flux on this interface 
-is replaced by the sum of the fine fluxes at that interface, at the end of the 
-coarse timestep. In general, this change in mass distribution on the grid 
-results in a change in the gravitational potential, meaning that an additional 
-corrective source term ought to be added. \cite{miniati:2007} and \cite{castro} 
-describe an algorithm for doing so. We consider the size of this effect to be 
-negligible and therefore we avoid the expense of solving for the potential change caused by 
-this mass change. Our grid boundaries are always located outside of the dense 
-stellar interiors that generate almost all of the gravitational potential. 
-Consequently the refluxing operation only changes the solution in regions of 
-ambient density, and even there only to the extent that there is disagreement 
-between the coarse and fine fluxes -- this clearly has a 
-vanishingly small effect on the gravitational potential.
 
@@
 
-Analytic solutions to the Poisson equation customarily assume that the
+Analytical solutions to the Poisson equation customarily assume that the
 potential vanishes at large distances from the region of non-zero
@@
 are the indices of boundary locations. $\Delta V$ is the volume of the
-zone. We have implemented this as an option\footnote{It is controlled
-  with the \texttt{gravity.direct\_sum\_bcs} input parameter.} in
-\castro. If there are $N$ zones per spatial dimension, then there are
+zone. If there are $N$ zones per spatial dimension, then there are
 $6 N^2$ boundary zones, and each boundary zone requires a sum over
@@
 for hydrodynamics simulations (though it is useful for comparison to
-approximate solutions).
+approximate solutions, so we have implemented it as an option in \castro).
 
@@
 the error is already well below $10^{-4}$ and we adopt this as our default 
-choice for all simulations with Poisson gravity. Our choice is also motivated by the 
-fact that at this value, the time for computing the boundary conditions on the
-coarse grid is comparable to the time for the multigrid solve itself; at higher values 
-the boundary condition calculation becomes much more expensive than the multigrid solve
-and so the overall returns are severely diminished. We note that in our implementation 
-there are no gains at all for going beyond $l_{\text{max}} \gtrsim 16$, as we 
-become limited by error due to the machine precision.
+choice for all simulations with Poisson gravity. In \autoref{sec:kepler} we 
+see that there are no gains to be had by increasing the accuracy further. 
 \begin{figure}[h]
@@
 origin. The problem domain for our simulations is $[-1.6, 1.6]^3$, and
-we take $R = 1.0 \text{cm}$ and $\rho = 10^3\ \text{g cm}^{-3}$. 
+we take $R = 1.0\ \text{cm}$ and $\rho = 10^3\ \text{g cm}^{-3}$. 
 The zones with $r > R$ are filled with an ambient material of very low density 
-$\rho = 10^{-8}\ \text{g cm}^{-3}$. We run this problem at multiple 
+($10^{-8}\ \text{g cm}^{-3}$). We run this problem at multiple 
 resolutions corresponding to jumps by a factor of two. For
@@
 
-We can solve one of these two sources of error by evaluating 
-\autoref{eq:sphere-analytical} with a mass $M$ that corresponds to the
-amount of mass actually on the grid, defined as the sum of the density
-multiplied by the volume of each cell for all cells whose density is 
-greater than ambient. The resolution study for this case is also
+We can solve one of these two sources of error by changing the density on  
+the grid so that the total mass $M$ is actually what we intend it to be.
+The resolution study for this case is also
 plotted in \autoref{fig:gravity_convergence}. We still obtain
 convergence slightly better than second-order, indicating that we 
-still have the geometrical problem of the mass distribution changing.
+have not eliminated the geometrical problem of the mass distribution changing.
 
@@
 by \citet{macmillan:1958}). The potential is in
-Equation 15 of that paper, though the last term is missing a factor of
-$1/2$, which destroys the symmetry. Inserting this missing factor and
-performing a simple coordinate transformation so that the center of
-the cube is at the origin, the potential is
+Equation 15 of that paper\footnote{The last term in that equation is missing a factor of
+$1/2$, which destroys the symmetry. We have inserted this missing factor and
+performed a simple coordinate transformation so that the center of
+the cube is at the origin.}:
 \begin{align}
@@
 and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
-\texttt{Fortran} and \texttt{C}, the inverse hyperbolic tangent is
+\texttt{Fortran} and \texttt{C}, the inverse hyperbolic tangent used here is
 \texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not}
@@
 The extent to which angular momentum conservation is violated
-will be a function of the resolution. When this is sufficiently high
-and a rotating reference frame is employed, excellent conservation
-properties can result, as demonstrated in \autoref{sec:kepler}. 
+will be a function of the resolution. When this is sufficiently high, 
+excellent conservation properties can result, as demonstrated in \autoref{sec:kepler}. 
 In particular, at reasonable resolution for a binary orbit our code 
 conserves angular momentum well enough that this no longer becomes 
-a reason to use a Lagrangian method such as smoothed-particle hydrodynamics.
+a strong argument in favor of using a Lagrangian method.
 We note that as the stars begin to coalesce, the rotating reference frame
@@
 \end{equation}
-So in the presence of rotation the total energy takes the form:
+In the presence of rotation the total energy becomes:
 \begin{equation}
@@
 
-In all simulations performed in a rotating reference frame, we will transform all relevant
-quantities back to the inertial reference frame when reporting them in this paper. In particular,
-for every zone we add to the displayed kinetic energy the rotational energy associated with motion 
-at the circular velocity corresponding to the distance of that zone from the rotation axis.
-We do the analogous calculation for angular momentum as well. When plotting the locations of 
+In all simulations performed in a rotating reference frame, we transform all relevant
+quantities back to the inertial reference frame when reporting them. In particular,
+for every zone we adjust the momenta and kinetic energy to account for rotation at 
+frequency $\omega$. When plotting the locations of 
 stars on the grid, we rotate them by an angle corresponding to the amount of time that has 
@@
 dwarfs by integrating the equation of hydrostatic equilibrium, taking
-the temperature and composition to be constant, and using the general
+the temperature and composition to be constant, and using the
 stellar equation of state.  This results in a single non-linear
@@
 interface between zones $i$ and $i+1$, found by simply adding up all
-the mass from zones $1, \ldots, i$ to get the enclosed mass,
+the mass from zones $1$ to $i$ to get the enclosed mass,
 $M_{i+1/2}$, and then setting $g_{i+1/2} =
@@
 physical center of mass $\mathbf{x}_{\text{COM}}$ of the stars as determined 
-by Kepler's third law. We also store the space velocity $\mathbf{v}_{\text{COM}}$ 
+by Kepler's third law. We also store the velocity $\mathbf{v}_{\text{COM}}$ 
 of the stars. Then, at each new time step we make a preliminary guess for their 
@@
 $\mathbf{n} \equiv \mathbf{x} / r$ then measures the direction of the outgoing wave with 
-respect to the observer, and $P_{ikjl}(\mathbf{n})$ is an operator in the transverse-traceless 
-gauge that projects a tensor onto the direction orthogonal to $\mathbf{n}$:
+respect to the observer, and $P_{ijkl}(\mathbf{n})$ is an operator that projects a tensor 
+onto the direction orthogonal to $\mathbf{n}$:
 \begin{align}
@@
 \begin{equation}
-  \text{STF}\left\{A^{ij}\right\} = \frac{1}{2}A^{ij} + \frac{1}{2}A^{ji} - \frac{1}{3} \delta^{ij} \sum_{k}A^{kk}.
+  \text{STF}\left\{A_{ij}\right\} = \frac{1}{2}A_{ij} + \frac{1}{2}A_{ji} - \frac{1}{3} \delta_{ij} \sum_{k}A_{kk}.
 \end{equation}
@@
 
-For ghost cells at the physical edge of the domain, we always use quantities 
-corresponding to the initial ambient material, to prevent material from raining down
-on to the grid in response to the disequilibrium outside the stars. This is an artificial 
-choice, but the dynamics outside of the stellar orbit is arbitrary anyway, due to the 
-lack of knowledge about the density and velocity field inside this gas. In addition, we employ 
-a ``sponge'' similar to that described by \citet{maestro3}. After the hydrodynamics update (but before 
-the gravity and rotation correction terms), we apply a damping force to the momentum 
+We employ a ``sponge'' similar to that described by \citet{maestro3} for the outer regions of the 
+computational domain. After the hydrodynamics update, we apply a damping force to the momentum 
 equation as follows:
@@
 was performed. In all plot files and diagnostic output generated by \castro, 
-and EPS figure files and plots generated by \wdmerger\ for this work,
+and figure files generated by \wdmerger,
 we store the active \texttt{git} commit hashes of \boxlib, \castro, and \wdmerger.
@@
 code paper \citep{castro}, and in the interest of brevity we will not
-repeat them here. Instead, we focus on a subset of problems that
+repeat them all here. Instead, we focus on a subset of problems that
 highlight the special difficulties introduced in merger
@@
 should obey simple, expected behaviors. Where possible, these should
-be quantified and a convergence study performed. We recommend that all
-simulation codes should check their performance on such problems to
-ensure their reliability when coupling the various physics modules
-necessary to build a complete and realistic simulation.
+be quantified and a convergence study performed, and that will 
+be the focus of the current section.
 
@@
 and not because of an instability in an individual star. Our goal here
-is thus to populate a single star onto our three-dimensional
+is thus to install a single star onto our three-dimensional
 coordinate grid and evolve it for a period of time long enough to
@@
 of this change becomes smaller with resolution. The star is only approximately 
-in equilibrium Only when the coarse grid of $256^3$ zones has a level of 
+in equilibrium when the coarse grid of $256^3$ zones has a level of 
 refinement that jumps by a factor of four. However, even then there is a slight 
@@
 uniform grid are shown in \autoref{fig:freefall}. They show excellent agreement
-between the theory and the simulation results, and this agreement holds even for 
-a more moderate $128^3$ zone simulation.
+between the theory and the simulation results.
 
@@
 Inability to correctly model the Kelvin-Helmholtz instability would 
-ave important consequences for how much we can trust the ability of 
+have important consequences for how much we can trust the ability of 
 \castro\ to test the violent merger progenitor model. Shearing between 
@@
 
-\citet{robertson:2010} (hereafter, R10) observe that Galilean
+\citet{robertson:2010} (hereafter, R10) observe that violation of Galilean
 invariance of simulation results for the Euler equations occurs only

### 2015-08-13 papers/wdmerger_I/paper.tex
fix en dashes -> em dashes

@@
 oxygen-neon-magnesium white dwarf. This would then be followed by
-collapse into a neutron star -- a result with significantly different
+collapse into a neutron star---a result with significantly different
 observational properties compared to a SN Ia. This scenario, termed
@@
 process resulted in the detonation of the system during the merger
-coalescence -- avoiding the much longer time-scale evolution -- were
+coalescence---avoiding the much longer time-scale evolution---were
 termed ``violent'' mergers.
@@
 This standard approach is therefore unacceptable in the case of a problem like white dwarf mergers, and 
-the fix to this energy leaking -- evaluating the energy transfer at the six zone faces instead of 
-the single zone center -- has a very minor cost in terms of code complexity and computational time.
+the fix to this energy leaking---evaluating the energy transfer at the six zone faces instead of 
+the single zone center---has a very minor cost in terms of code complexity and computational time.
 
@@
 kinetic energy. This has the virtue of ensuring that there is no conflict due to discretization between the momentum and 
-energy equations, and also correctly ensuring that the gravitational force does not directly change the internal energy -- 
-and thus the temperature -- of the fluid. However, it explicitly sacrifices total energy conservation. ${\tt gs} = 4$ is 
+energy equations, and also correctly ensuring that the gravitational force does not directly change the internal energy---and thus the temperature---of the fluid.
+However, it explicitly sacrifices total energy conservation.
+${\tt gs} = 4$ is
 the new conservative method of evaluating the energy source terms at cell faces. The results for the change in energy 

### 2015-08-13 papers/wdmerger_I/paper.tex
Mike's first round of commments

@@
 mass near the Eddington rate from the less massive white dwarf
-overflowing its Roche lobe. \cite{saionomoto:1985} tracked the
+overflowing its Roche lobe.\MarginPar{ref?} \cite{saionomoto:1985} tracked the
 evolution of the flame and found that it propagated quiescently into
@@
 in a similarly massive system. Very recently, \cite{kashyap:2015} 
-found that a detonation occurred in a similarly massive system. Notably,
+found that a detonation occurred in a similarly massive system.\MarginPar{``similarly massive system'' repeated twice in succession} Notably,
 the detonation occurred self-consistently and did not need to be inserted 
@@
 
-The Euler equations for hydrodynamics (in the absence of source terms) in conservative form are:
+The Euler equations for hydrodynamics (in the absence of source terms) in conservative form are: \MarginPar{we usually use $U$ for the velocity vector, with $U = (u, v, w)$}
 \begin{align}
@@
 explicitly introduce a positivity-guaranteeing diffusive flux, or
-reset the characteristics of the affected zone. We choose the latter
+reset the characteristics \MarginPar{``characteristics'' has a very precise meaning for hyperbolic systems, use a different word} of the affected zone. We choose the latter
 approach. Even though it is non-conservative, it preserves a
@@
 (density, momentum, energy, and temperature) of the
-resetted zone to be exactly equal to the field values of this 
+reset zone to be exactly equal to the field values of this 
 adjacent zone. If no adjacent zone reaches the density floor, then
@@
 to the temperature floor for our simulations (which is typically 
-$10^{5}\ \text{K}$ for our stellar simulations). We then recompute 
+$10^{5}\ \text{K}$ \MarginPar{do we really need it this cold?  I thought we were doing $10^6$~K} for our stellar simulations). We then recompute 
 the thermodynamics to be consistent with these values. The 
@@
 has a higher resolution than the grid below it by some integer factor
-$N$, which is restricted to be $N = 2\ \text{or}\ 4$ in the code. The
+$N$, which is restricted to be $N = 2\ \text{or}\ 4$ in \castro. The
 zones are strictly contained within the rectangular extent of the
@@
 advanced at a different timestep and a correction step is applied at
-the end to synchronize the various levels. Generally the number of
+the end to synchronize the various levels. Generally \MarginPar{why ``Generally''?  Is there ever as case where the subcycled step isn't the jump in refinement?} the number of
 subcycled timesteps is equal to the jump in refinement between levels,
@@
 
-There are two ways that these source terms enter the hydrodynamics update.
+There are two ways that these source terms enter the hydrodynamics update. \MarginPar{``hydrodynamics update'' appears twice in succession}
 First, during the hydrodynamics update, we alter the edge states that enter
@@
 this can be done using the cell-centered
-source term evaluated at time-level $n$. The details of this are described in 
+source term evaluated at time-level $n$. \MarginPar{we trace, so it's not just the cell-centered source term} The details of this are described in 
 the appendix. Second, after the hydrodynamics step, we add the time-centered source terms
@@
 single-star Type Ia supernova simulations that employ
-self-gravity. However, for a problem like that of a binary star system
+self-gravity. \MarginPar{also try for Castro core-collapse sims---maybe reference a paper on each subject from castro} However, for a problem like that of a binary star system
 with significant departures from spherical symmetry, this assumption
@@
 multiple of the grid spacing, and the center of the sphere is at the
-origin. The problem domain for our simulations is $[-1.6, 1.6]^3$, and
+origin. The problem domain for our simulations is $[-1.6 \text{cm}, 1.6 \text{cm}]^3$, and
 we take $R = 1.0\ \text{cm}$ and $\rho = 10^3\ \text{g cm}^{-3}$. 
@@
 The resolution study for this case is also
-plotted in \autoref{fig:gravity_convergence}. We still obtain
+plotted in \autoref{fig:gravity_convergence}. \MarginPar{explain that this is the ``normalized sphere'' in the figure} We still obtain
 convergence slightly better than second-order, indicating that we 
@@
 components. If the rotation is about the $z$-axis, then the discrete update to 
-$v_x^{n+1}$ depends on the value of $v_y^{n+1}$, and vice versa. If we fix the value of 
+$v_x^{n+1}$ \MarginPar{we are using $u, v, w$ below, and is more consistent with other papers / User's Guide} depends on the value of $v_y^{n+1}$, and vice versa. If we fix the value of 
 the time-level $n+1$ quantities after coming out of the hydrodynamics update, there 
@@
 stars on the grid, we rotate them by an angle corresponding to the amount of time that has 
-passed since the simulation started.
+passed since the simulation started. \MarginPar{show the transforms.  Also, be more explicit---is this just done for the plotfiles?}
 
@@
 \label{sec:initial_models}
+
+\MarginPar{use subsections to break the distinct topics apart in this section---it all runs together}
 
@@
 the purpose of sampling the 1D model, and the sub-zones are added
-together to obtain the full zone's characteristics. This
+together to obtain the full zone's characteristics \MarginPar{use ``state'' instead of ``characteristics''}. This
 sub-grid-scale interpolation is useful especially near the edge of the star,

### 2015-08-13 papers/wdmerger_I/paper.tex
Updates in response to Mike's first pass

--- 2015-08-13 papers/wdmerger_I/paper.tex
@@
 that a carbon-oxygen detonation would occur, in line with their
-earlier results. \cite{moll:2014} was also able to find a detonation
-in a similarly massive system. Very recently, \cite{kashyap:2015} 
-found that a detonation occurred in a similarly massive system. Notably,
+earlier results. \cite{moll:2014} and \cite{kashyap:2015} were also 
+able to find a detonation in similarly massive systems. Notably,
 the detonation occurred self-consistently and did not need to be inserted 
@@
 evolution calculations. \autoref{sec:Numerical Methodology}
-%\hyperref[sec:Numerical Methodology]{Section \ref*{sec:Numerical Methodology}}
 describes our code and why it can provide useful results compared to
 other methodologies used for this problem. 
-\autoref{sec:initial_models} describes the method we use for setting up a
+\autoref{sec:implementation} describes the method we use for setting up a
 binary white dwarf simulation. \autoref{sec:Tests} discusses a few
@@
 \end{align}
-Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity
+Here $\rho$ is the mass density, $\mathbf{u} = (u, v, w)$ is the fluid velocity
 vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the
@@
 explicitly introduce a positivity-guaranteeing diffusive flux, or
-reset the characteristics of the affected zone. We choose the latter
+reset the properties of the affected zone. We choose the latter
 approach. Even though it is non-conservative, it preserves a
@@
 advanced at a different timestep and a correction step is applied at
-the end to synchronize the various levels. Generally the number of
+the end to synchronize the various levels. The number of
 subcycled timesteps is equal to the jump in refinement between levels,
@@
 
-There are two ways that these source terms enter the hydrodynamics update.
+There are two ways that these source terms enter the system evolution. 
 First, during the hydrodynamics update, we alter the edge states that enter
@@
 this can be done using the cell-centered
-source term evaluated at time-level $n$. The details of this are described in 
+source term evaluated at time-level $n$. We choose a more accurate approach, 
+which is also second order, of tracing
+under the source term; the details of this are described in 
 the appendix. Second, after the hydrodynamics step, we add the time-centered source terms
@@
 distribution, and therefore is best suited for problems like
-single-star Type Ia supernova simulations that employ
-self-gravity. However, for a problem like that of a binary star system
+single-star Type Ia supernova simulations (e.g. \cite{malone:2014})
+that employ self-gravity. However, for a problem like that of a binary star system
 with significant departures from spherical symmetry, this assumption
@@
 white dwarfs of mass ratio 2/3, loaded onto the grid by the method described 
-in \autoref{sec:initial_models}. We terminated the simulation just after
+in \autoref{sec:implementation}. We terminated the simulation just after
 initialization, so that we perform only an initial Poisson solve for this 
@@
 the grid so that the total mass $M$ is actually what we intend it to be.
-The resolution study for this case is also
+The resolution study for this case (the ``normalized sphere'') is also
 plotted in \autoref{fig:gravity_convergence}. We still obtain
@@
 components. If the rotation is about the $z$-axis, then the discrete update to 
-$v_x^{n+1}$ depends on the value of $v_y^{n+1}$, and vice versa. If we fix the value of 
+$u^{n+1}$ depends on the value of $v^{n+1}$, and vice versa. If we fix the value of 
 the time-level $n+1$ quantities after coming out of the hydrodynamics update, there 
@@
 In all simulations performed in a rotating reference frame, we transform all relevant
-quantities back to the inertial reference frame when reporting them. In particular,
-for every zone we adjust the momenta and kinetic energy to account for rotation at 
-frequency $\omega$. When plotting the locations of 
-stars on the grid, we rotate them by an angle corresponding to the amount of time that has 
-passed since the simulation started.
+quantities back to the inertial reference frame when reporting them in analysis routines 
+and visualization (though the data is saved to plotfiles while still in the rotating frame). In particular,
+for every zone we adjust the position, momentum, and energy to account for rotation.
+If the position is $\mathbf{x}$ in the inertial frame and $\mathbf{x}^\prime$ in 
+the rotating frame, and the rotation vector is $\bm{\omega}$, the transformation rules are:
+\begin{align}  
+  \mathbf{x}(t) &= \mathbf{R}\mathbf{x}^\prime(t) \\
+  \mathbf{v}(t) &= \mathbf{v}^\prime(t) + \bm{\omega} \times \left(\mathbf{R} \mathbf{x}^\prime(t)\right)
+\end{align}
+The rotation matrix $\mathbf{R}$ is:
+\begin{equation}
+  \mathbf{R} = \mathbf{R}_z({\bm{\theta}}_3) \mathbf{R}_y({\bm{\theta}}_2) \mathbf{R}_x({\bm{\theta}}_1) 
+             %% = \left( \begin{array}{ccc} 
+             %%     \text{cos}(\theta_2)\, \text{cos}(\theta_3) & 
+             %%     -\text{cos}(\theta_2)\, \text{sin}(\theta_3) & 
+             %%     \text{sin}(\theta_2) \\
+             %%     \text{cos}(\theta_1)\, \text{sin}(\theta_3) + \text{sin}(\theta_1)\, \text{sin}(\theta_2)\, \text{cos}(\theta_3) &
+             %%     \text{cos}\theta_1)\, \text{cos}(\theta_3) - \text{sin}(\theta_1)\, \text{sin}(\theta_2)\, \text{sin}(\theta_3) &
+             %%     -\text{sin}(\theta_1)\, \text{cos}(\theta_2) \\
+             %%     \text{sin}(\theta_1)\, \text{sin}(\theta_3) - \text{cos}(\theta_1)\, \text{sin}(\theta_2)\, \text{cos}(\theta_3) &
+             %%     \text{sin}(\theta_1)\, \text{cos}(\theta_3) + \text{cos}(\theta_1)\, \text{sin}(\theta_2)\, \text{sin}(\theta_3) &
+             %%     \text{cos}(\theta_1)\, \text{cos}(\theta_2) 
+             %%   \end{array} \right)
+\end{equation}
+where $\mathbf{R}_x$, $\mathbf{R}_y$, and $\mathbf{R}_z$ are the standard rotation matrices about 
+the $x$, $y$, and $z$ axes, and $\bm{\theta} = \bm{\omega} t$.
 
 %==========================================================================
-% Initial Models and Problem Setup
+% Problem Description and Software Implementation
 %==========================================================================
 
-\section{Initial Models and Problem Setup}
+\section{Problem Description and Software Implementation}
+\label{sec:implementation}
+
+In this section we describe our white dwarf merger software, and focus in 
+particular on the initial white dwarf models (\autoref{sec:initial_models}), 
+the initial problem setup (\autoref{sec:initial_state}), and analysis 
+(\autoref{sec:analysis}) components.
+
+The software used to generate the test problems in this paper
+(as well as the manuscript itself),
+\wdmerger\footnote{\wdmerger\ can be obtained at \url{https://github.com/BoxLib-Codes/wdmerger}.},
+is freely available at an online repository hosting service.
+Version control in both the parent software (\boxlib, \castro) and in \wdmerger\
+permits us to reference the state of the code at the time a simulation
+was performed. In all plot files and diagnostic output generated by \castro, 
+and figure files generated by \wdmerger,
+we store the active \texttt{git} commit hashes of \boxlib, \castro, and \wdmerger.
+Line plots are generated using the \matplotlib\ library for \python\ 
+\citep{matplotlib}, while slice plots and other multi-dimensional visualizations are 
+generated using the \yt\ code \citep{yt}.
+
+\subsection{White Dwarf Models}
 \label{sec:initial_models}
@@
 the purpose of sampling the 1D model, and the sub-zones are added
-together to obtain the full zone's characteristics. This
+together to obtain the full zone's state. This
 sub-grid-scale interpolation is useful especially near the edge of the star,
@@
 $n_{\text{sub}} = 4$.
+
+\subsection{Initial State}
+\label{sec:initial_state}
 
@@
 the maximum timestep is set to be equal to one-half of the CFL limit.
+
+The computational domain has a total size of $1.024 \times 10^{10}\ \text{cm}$ in each 
+spatial dimension, and is centered at the origin. Our coarse grid has $256^3$ zones,
+corresponding to a spatial resolution of 400 km. For the present study, we choose a simple 
+refinement strategy: all zones within twice the Roche radius of each star are tagged for maximal 
+refinement. The extra buffer from doubling the size of this region ensures that the sharp density
+gradients near the edge of the star are within the zone of refinement. We also ensure that the 
+outer part of the domain is never tagged for refinement. In future work we will add criteria 
+that tag for refinement the gas between the stars, which is expected to feature nuclear burning.
+
+We employ a ``sponge'' similar to that described by \citet{maestro3} for the outer regions of the 
+computational domain. After the hydrodynamics update, we apply a damping force to the momentum 
+equation as follows:
+\begin{equation}
+  (\rho \mathbf{u})^{n+1} \to \frac{(\rho \mathbf{u}^{n+1})}{1 + (\Delta t / \Delta t_S) f_S},
+\end{equation} 
+where $\Delta t_S$ is a timescale for the sponge to operate on, and $f_S$ is the damping factor.
+We choose it so that that the sponge is non-operational inside a radius $r_S$ from the origin, and 
+fully applied at a radius $r_S^\prime \equiv r_S + \Delta r_S$. We then smooth the sponge 
+out between $r_S$ and $r_S^\prime$:
+\begin{equation}
+  f_{S} = \begin{dcases} 0 & r < r_S \\ \frac{1}{2}\left(1 - \text{cos}\left[\pi\left(\dfrac{r - r_S}{\Delta r_S}\right)\right]\right) & r_S \leq r < r_S^\prime \\ 1 & r \geq r_S^\prime. \end{dcases}
+\end{equation}
+For the simulations in this paper we set $r_S$ to be 75\% of the distance from the origin to the domain boundaries,
+and $\Delta r_S$ so that the sponge extends another 10\% of that distance. We set 
+$\Delta t_S = 0.01$ s, which is of the same order as the as the CFL timestep 
+for typical problem setups. We find that this sponge is necessary especially in the 
+rotating reference frame, as standing instabilities can create very large velocities 
+in the ambient fluid that drag down the global timestep by up to an order of magnitude.
+However, this provides another reason to distrust the dynamics of the low-density fluid
+outside of the stellar orbits.
+
+\subsection{Analysis}
+\label{sec:analysis}
 
@@
 
+\subsubsection{Gravitational Waves}
+\label{sec:grvatational_waves}
+
 A final diagnostic quantity we consider is the gravitational wave emission by the 
@@
 However, the strain at any other distance is easily calculated and goes as the inverse of the distance.
-
-The computational domain has a total size of $1.024 \times 10^{10}\ \text{cm}$ in each 
-spatial dimension, and is centered at the origin. Our coarse grid has $256^3$ zones,
-corresponding to a spatial resolution of 400 km. For the present study, we choose a simple 
-refinement strategy: all zones within twice the Roche radius of each star are tagged for maximal 
-refinement. The extra buffer from doubling the size of this region ensures that the sharp density
-gradients near the edge of the star are within the zone of refinement. We also ensure that the 
-outer part of the domain is never tagged for refinement. In future work we will add criteria 
-that tag for refinement the gas between the stars, which is expected to feature nuclear burning.
-
-We employ a ``sponge'' similar to that described by \citet{maestro3} for the outer regions of the 
-computational domain. After the hydrodynamics update, we apply a damping force to the momentum 
-equation as follows:
-\begin{equation}
-  (\rho \mathbf{u})^{n+1} \to \frac{(\rho \mathbf{u}^{n+1})}{1 + (\Delta t / \Delta t_S) f_S},
-\end{equation} 
-where $\Delta t_S$ is a timescale for the sponge to operate on, and $f_S$ is the damping factor.
-We choose it so that that the sponge is non-operational inside a radius $r_S$ from the origin, and 
-fully applied at a radius $r_S^\prime \equiv r_S + \Delta r_S$. We then smooth the sponge 
-out between $r_S$ and $r_S^\prime$:
-\begin{equation}
-  f_{S} = \begin{dcases} 0 & r < r_S \\ \frac{1}{2}\left(1 - \text{cos}\left[\pi\left(\dfrac{r - r_S}{\Delta r_S}\right)\right]\right) & r_S \leq r < r_S^\prime \\ 1 & r \geq r_S^\prime. \end{dcases}
-\end{equation}
-For the simulations in this paper we set $r_S$ to be 75\% of the distance from the origin to the domain boundaries,
-and $\Delta r_S$ so that the sponge extends another 10\% of that distance. We set 
-$\Delta t_S = 0.01$ s, which is of the same order as the as the CFL timestep 
-for typical problem setups. We find that this sponge is necessary especially in the 
-rotating reference frame, as standing instabilities can create very large velocities 
-in the ambient fluid that drag down the global timestep by up to an order of magnitude.
-However, this provides another reason to distrust the dynamics of the low-density fluid
-outside of the stellar orbits.
-
-The software used to generate the test problems in this paper
-(as well as the manuscript itself),
-\wdmerger\footnote{\wdmerger\ can be obtained at \url{https://github.com/BoxLib-Codes/wdmerger}.},
-is freely available at an online repository hosting service.
-Version control in both the parent software (\boxlib, \castro) and in \wdmerger\
-permits us to reference the state of the code at the time a simulation
-was performed. In all plot files and diagnostic output generated by \castro, 
-and figure files generated by \wdmerger,
-we store the active \texttt{git} commit hashes of \boxlib, \castro, and \wdmerger.
-Line plots are generated using the \matplotlib\ library for \python\ 
-\citep{matplotlib}, while slice plots and other multi-dimensional visualizations are 
-generated using the \yt\ code \citep{yt}.
 
@@
 radius of the star, determined by the volume of the grid that has a density 
-greater than $10^3\ \text{g cm}^{-3}$ (see \autoref{sec:initial_models} 
+greater than $10^3\ \text{g cm}^{-3}$ (see \autoref{sec:implementation} 
 for details on this measure). This choice of density is intended to 
@@
 the case of gravitational free fall. We place two stars on the grid 
-in the manner of \autoref{sec:initial_models}. The distance $a$ between 
+in the manner of \autoref{sec:implementation}. The distance $a$ between 
 them corresponds to a chosen orbital period $T$, consistent with the total

### 2015-08-14 papers/wdmerger_I/paper.tex

@@
 
-The computational domain has a total size of $1.024 \times 10^{10}\ \text{cm}$ in each 
-spatial dimension, and is centered at the origin. Our coarse grid has $256^3$ zones,
-corresponding to a spatial resolution of 400 km. For the present study, we choose a simple 
-refinement strategy: all zones within twice the Roche radius of each star are tagged for maximal 
-refinement. The extra buffer from doubling the size of this region ensures that the sharp density
-gradients near the edge of the star are within the zone of refinement. We also ensure that the 
-outer part of the domain is never tagged for refinement. In future work we will add criteria 
-that tag for refinement the gas between the stars, which is expected to feature nuclear burning.
-
-We employ a ``sponge'' similar to that described by \citet{maestro3} for the outer regions of the 
-computational domain. After the hydrodynamics update, we apply a damping force to the momentum 
+The computational domain has a total size of $1.024 \times
+10^{10}\ \text{cm}$ in each spatial dimension, and is centered at the
+origin. Our coarse grid has $256^3$ zones, corresponding to a spatial
+resolution of 400 km. For the present study, we choose a simple
+refinement strategy: all zones within twice the Roche radius of each
+star are tagged for maximal refinement. The extra buffer from doubling
+the size of this region ensures that the sharp density gradients near
+the edge of the star are within the zone of refinement. We also ensure
+that the outer part of the domain is never tagged for refinement. In
+future work we will add criteria that tag for refinement the gas
+between the stars, which is expected to feature nuclear burning.
+
+We employ a ``sponge'' similar to that described by \citet{maestro3}
+for the outer regions of the computational domain. After the
+hydrodynamics update, we apply a damping force to the momentum
 equation as follows:
@@
 \end{equation} 
-where $\Delta t_S$ is a timescale for the sponge to operate on, and $f_S$ is the damping factor.
-We choose it so that that the sponge is non-operational inside a radius $r_S$ from the origin, and 
-fully applied at a radius $r_S^\prime \equiv r_S + \Delta r_S$. We then smooth the sponge 
-out between $r_S$ and $r_S^\prime$:
+where $\Delta t_S$ is a timescale for the sponge to operate on, and
+$f_S$ is the damping factor.  We choose it so that that the sponge is
+non-operational inside a radius $r_S$ from the origin, and fully
+applied at a radius $r_S^\prime \equiv r_S + \Delta r_S$. We then
+smooth the sponge out between $r_S$ and $r_S^\prime$:
 \begin{equation}
@@
 \end{equation}
-For the simulations in this paper we set $r_S$ to be 75\% of the distance from the origin to the domain boundaries,
-and $\Delta r_S$ so that the sponge extends another 10\% of that distance. We set 
-$\Delta t_S = 0.01$ s, which is of the same order as the as the CFL timestep 
-for typical problem setups. We find that this sponge is necessary especially in the 
-rotating reference frame, as standing instabilities can create very large velocities 
-in the ambient fluid that drag down the global timestep by up to an order of magnitude.
-However, this provides another reason to distrust the dynamics of the low-density fluid
-outside of the stellar orbits.
+For the simulations in this paper we set $r_S$ to be 75\% of the \MarginPar{maybe a plot of the sponge profile?}
+distance from the origin to the domain boundaries, and $\Delta r_S$ so
+that the sponge extends another 10\% of that distance. We set $\Delta
+t_S = 0.01$ s, which is of the same order as the as the CFL timestep
+for typical problem setups. We find that this sponge is necessary
+especially in the rotating reference frame, as standing instabilities
+can create very large velocities in the ambient fluid that drag down
+the global timestep by up to an order of magnitude.  However, this
+provides another reason to distrust the dynamics of the low-density
+fluid outside of the stellar orbits. \MarginPar{I am not sure I understand 
+this statement}
 

### 2015-08-14 papers/wdmerger_I/paper.tex
more comments

@@
 
-We loaded a single star of mass $0.9\ \msolar$ onto the grid at the origin, 
-and evolved it for 200 seconds. Our diagnostic of choice is the effective 
-radius of the star, determined by the volume of the grid that has a density 
-greater than $10^3\ \text{g cm}^{-3}$ (see \autoref{sec:implementation} 
-for details on this measure). This choice of density is intended to 
-mark a reasonable outer edge to the star that is not immediately susceptible 
-to the numerical errors prevalent near the physical edge of the star. 
-\autoref{fig:single_star_static_radius} shows our results at various resolutions. 
-As expected, the star quickly approaches an equilibrium size that is different 
-(and in this case larger) than the one-dimensional model. However, the magnitude 
-of this change becomes smaller with resolution. The star is only approximately 
-in equilibrium when the coarse grid of $256^3$ zones has a level of 
-refinement that jumps by a factor of four. However, even then there is a slight 
-uptick in the size toward the event, implying that the numerical stability 
-is not guaranteed for arbitrarily long timescales. This result suggests that when 
-doing the equilibrium initial models that will former the basis of later calculations,
-we should carefully monitor the evolution of the stars before applying any 
-artificial damping to cause the merger, to ensure that the merger is due to 
-this applied force and not the intrinsic numerical instability of the stars.
+We loaded a single star of mass $0.9\ \msolar$ onto the grid at the
+origin, and evolved it for 200 seconds. Our diagnostic of choice is
+the effective radius of the star, determined by the volume of the grid
+that has a density greater than $10^3\ \text{g cm}^{-3}$ (see
+\autoref{sec:implementation} for details on this measure). This choice
+of density is intended to mark a reasonable outer edge to the star
+that is not immediately susceptible to the numerical errors prevalent
+near the physical edge of the star.
+\autoref{fig:single_star_static_radius} shows our results at various
+resolutions.  As expected, the star quickly approaches an equilibrium
+size that is different (and in this case larger) than the
+one-dimensional model. However, the magnitude of this change becomes
+smaller with resolution. The star is only approximately in equilibrium
+when the coarse grid of $256^3$ zones has a level of refinement that
+jumps by a factor of four. However, even then there is a slight uptick
+in the size toward the event, implying that the numerical stability is
+not guaranteed for arbitrarily long timescales. This result suggests
+that when doing the equilibrium initial models that will former the
+basis of later calculations, we should carefully monitor the evolution
+of the stars before applying any artificial damping to cause the
+merger, to ensure that the merger is due to this applied force and not
+the intrinsic numerical instability of the stars. \MarginPar{you never referenced figure 4}
 
@@
     The meaning of the rows and columns is the same as in \autoref{fig:kh_p2}.
-    \label{fig:kh_p3}}
+    \label{fig:kh_p3}  {\color{red} You should note this is $R_B$}}
 \end{figure*}
@@
 
-\citet{hopkins:2015} performed this test as part of the testing of their code GIZMO. 
-They showed the late-time evolution of this system, when non-linear effects have 
-taken over and significantly disrupted the initial flow. However, at low resolution the 
-tested grid algorithm had failed to disrupt both for $v = 0$ and $v = 10$. We too ran 
-this problem until $t = 10$, and confirm that the Kelvin-Helmholtz instability damps out 
-at low resolution but goes strongly non-linear and disrupts the flow at high resolution.
-However, we strongly emphasize the point that this does not objectively 
-demonstrate a defiency in fixed-grid codes for this problem. We can only determine 
-the validity of a method when we have a trustworthy, converged solution to compare to, and 
-this is lacking for this problem at late times. As observed by \citeauthor{mcnally:2012}, 
-this is because the secondary instabilities form for this problem when the whorls of 
-the Kelvin-Helmholtz tendrils stretch out and create gradients that approach the 
-grid resolution. This is prime breeding ground for numerical noise. But because 
-the nature of this noise will depend on the resolution, it will be very different for 
-simulations at different resolutions. If these instabilities are seeded because of 
-this resolution-dependent noise and are not seeded instead in a controlled manner 
-such that they appear at the same time and location, then we simply cannot draw 
-any conclusions that bear on the question of verification from this test at late times. 
-\autoref{fig:kh_p3_high_res} provides a sense of this by examining the crucial 
-time at which the transition from the linear to the non-linear regime is occurring. 
-At all of these very high resolutions the secondary instabilities develop, but they 
-occur at different times and have different spatial scales for each resolution.
+\citet{hopkins:2015} performed this test as part of the testing of
+their code GIZMO.  They showed the late-time evolution of this system,
+when non-linear effects have taken over and significantly disrupted
+the initial flow. However, at low resolution the tested grid algorithm
+had failed to disrupt both for $v = 0$ and $v = 10$. We too ran this
+problem until $t = 10$, and confirm that the Kelvin-Helmholtz
+instability damps out at low resolution but goes strongly non-linear
+and disrupts the flow at high resolution.  However, we strongly
+emphasize the point that this does not objectively demonstrate a
+defiency in fixed-grid codes for this problem. We can only determine
+the validity of a method when we have a trustworthy, converged
+solution to compare to, and this is lacking for this problem at late
+times. As observed by \citeauthor{mcnally:2012}, this is because the
+secondary instabilities form for this problem when the whirls of the
+Kelvin-Helmholtz tendrils stretch out and create gradients that
+approach the grid resolution. This is prime breeding ground for
+numerical noise. But because the nature of this noise will depend on
+the resolution, it will be very different for simulations at different
+resolutions. If these instabilities are seeded because of this
+resolution-dependent noise and are not seeded instead in a controlled
+manner such that they appear at the same time and location, then we
+simply cannot draw any conclusions that bear on the question of
+verification from this test at late times.
+\autoref{fig:kh_p3_high_res} provides a sense of this by examining the
+crucial time at which the transition from the linear to the non-linear
+regime is occurring.  At all of these very high resolutions the
+secondary instabilities develop, but they occur at different times and
+have different spatial scales for each resolution.
 
@@
 
-To see this effect in action for a stellar simulation, we repeated the test of  
-\autoref{sec:HSE} with a bulk velocity on the grid. We chose a velocity of $2.56 \times 10^{8}\ \text{cm s}^{-1}$.
-For context, this is comparable to the orbital velocities of the stars in \autoref{sec:kepler}.
-This test was inspired by \citet{tasker:2008}, who considered a moving galaxy cluster 
-and who obtained a long timescale evolution by using periodic boundary conditions, 
-so that the cluster would cross the domain multiple times throughout the evolution.
-We believe that periodic boundary conditions are unrealistic for our type of simulation,
-so we prefer to do one continuous simulation where the star does not cross the boundaries.
-Since our normal grid was not large enough to allow the motion to continue for very long, we expanded 
-the domain size by a factor of four, and then included an extra refined level around the star to 
-keep the effective resolution the same. We started the star off in the lower left corner of the domain, and 
-pointed its velocity towards the upper right corner. This allowed us to evolve the star for the 
-same length of time as for the original test. We note that getting the gravity boundary conditions right 
-required us to move the origin of the problem at the bulk velocity, so that the multipole moments 
-were computed with respect to the stellar center.
+To see this effect in action for a stellar simulation, we repeated the
+test of \autoref{sec:HSE} with a bulk velocity on the grid. We chose a
+velocity of $2.56 \times 10^{8}\ \text{cm s}^{-1}$.  \MarginPar{Approx
+  what Mach number is this?}  {For context, this is comparable to the
+  orbital velocities of the stars in \autoref{sec:kepler}.  This test
+  was inspired by \citet{tasker:2008}, who considered a moving galaxy
+  cluster and who obtained a long timescale evolution by using
+  periodic boundary conditions, so that the cluster would cross the
+  domain multiple times throughout the evolution.  We believe that
+  periodic boundary conditions are unrealistic for our type of
+  simulation, so we prefer to do one continuous simulation where the
+  star does not cross the boundaries.  Since our normal grid was not
+  large enough to allow the motion to continue for very long, we
+  expanded the domain size by a factor of four, and then included an
+  extra refined level around the star to keep the effective resolution
+  the same. We started the star off in the lower left corner of the
+  domain, and pointed its velocity towards the upper right
+  corner. This allowed us to evolve the star for the same length of
+  time as for the original test. We note that getting the gravity
+  boundary conditions right required us to move the origin of the
+  problem at the bulk velocity, so that the multipole moments were
+  computed with respect to the stellar center.  \MarginPar{are you
+    sure that the single WD left quiet in the {\em corner} of the
+    bigger domain is as stable as it was in the earlier test?}
 
@@
 In \autoref{fig:single_star_compare_radius}, we take the results of
-\autoref{sec:HSE} (the ``static'' case), and plot on top of it the results of this 
-new simulation (the ``motion'' case). We see immediately that this bulk velocity
-causes the star to be much worse at maintaining hydrostatic equilibrium. Not only 
-is the absolute size of the star significantly larger (nearly a factor of two 
-at the lowest feasible resolution we consider), but also there is a clear upward 
-trend in the size that has not terminated at any resolution by the end of the simulation.
-This again emphasizes the results mentioned earlier, that we must be careful not 
-to trust any simulation with significant mass transfer if we are not confident that the 
-mass transfer is seeded in a controllable manner and free from numerical noise.
+\autoref{sec:HSE} (the ``static'' case), and plot on top of it the
+results of this new simulation (the ``motion'' case). We see
+immediately that this bulk velocity causes the star to be much worse
+at maintaining hydrostatic equilibrium. Not only is the absolute size
+of the star significantly larger (nearly a factor of two at the lowest
+feasible resolution we consider), but also there is a clear upward
+trend in the size that has not terminated at any resolution by the end
+of the simulation.  This again emphasizes the results mentioned
+earlier, that we must be careful not to trust any simulation with
+significant mass transfer if we are not confident that the mass
+transfer is seeded in a controllable manner and free from numerical
+noise.  \MarginPar{I really think that we should be able to do better here---this is where I'd like to see ppm\_type = 2, or better hydro-gravity coupling (eventually)}
 

### 2015-08-14 papers/wdmerger_I/paper.tex
finish comments on the main paper

@@
 
-Another that we consider is the energy conservation of the system. Recalling \autoref{sec:gravity_hydro_coupling}, 
-there are several different methods of applying the gravitational source term to the hydrodynamics equations. In \castro\ we presently 
-have four options, controlled by the parameter {\tt castro.grav\_source\_type}, which we will shortern to {\tt gs} for the present 
-discussion. ${\tt gs} = 1$ is the standard cell-centered source term for gravity. ${\tt gs} = 2$ is a slight variation on this scheme 
-that determines the value of the energy source term after the momentum source term has been applied; the former approach uses 
-the uncorrected momenta in calculating $\rho \mathbf{u} g$, which we have found to be more accurate. ${\tt gs} = 3$ is entirely 
-different: after calculating the new momenta, we reset the total energy to be equal to the internal energy plus the 
-kinetic energy. This has the virtue of ensuring that there is no conflict due to discretization between the momentum and 
-energy equations, and also correctly ensuring that the gravitational force does not directly change the internal energy---and thus the temperature---of the fluid.
-However, it explicitly sacrifices total energy conservation.
-${\tt gs} = 4$ is
-the new conservative method of evaluating the energy source terms at cell faces. The results for the change in energy 
-after a single orbit are seen in the first column of \autoref{table:sources}. The first two versions give reasonable and 
-similar levels of energy conservation. The third has total energy changes on the order of 100\%, but this itself 
-does not have a severe effect on the dynamics because in this scheme the total energy variable is effectively a placeholder 
-value of the kinetic energy plus internal energy, rather than being evolved directly. The last scheme is nearly two orders 
-of magnitude better in energy conservation, justifying the effort in varying the scheme.
+Another that we consider is the energy conservation of the
+system. Recalling \autoref{sec:gravity_hydro_coupling}, there are
+several different methods of applying the gravitational source term to
+the hydrodynamics equations. In \castro\ we presently have four
+options, controlled by the parameter {\tt castro.grav\_source\_type},
+which we will shortern to {\tt gs} for the present discussion. ${\tt
+  gs} = 1$ is the standard cell-centered source term for
+gravity. ${\tt gs} = 2$ is a slight variation on this scheme that
+determines the value of the energy source term after the momentum
+source term has been applied; the former approach uses the uncorrected
+momenta in calculating $\rho \mathbf{u} g$, which we have found to be
+more accurate.
+\MarginPar{which is more accurate?
+Also, the ``standard approach'' is {\tt gs = 2} --- this is what is written, for exmaple, in the PPM paper}${\tt gs} = 3$ is entirely different: after calculating
+the new momenta, we reset the total energy to be equal to the internal
+energy plus the kinetic energy. This has the virtue of ensuring that
+there is no conflict due to discretization between the momentum and
+energy equations, and also correctly ensuring that the gravitational
+force does not directly change the internal energy---and thus the
+temperature---of the fluid. However, it explicitly sacrifices total
+energy conservation. ${\tt gs} = 4$ is the new conservative method of
+evaluating the energy source terms at cell faces. The results for the
+change in energy after a single orbit are seen in the first column of
+\autoref{table:sources}. The first two versions give reasonable and
+similar levels of energy conservation. The third has total energy
+changes on the order of 100\%, but this itself does not have a severe
+effect on the dynamics because in this scheme the total energy
+variable is effectively a placeholder value of the kinetic energy plus
+internal energy, rather than being evolved directly. The last scheme
+is nearly two orders of magnitude better in energy conservation,
+justifying the effort in varying the scheme.
 
@@
 
-We are most interested in the stability of these systems over long timescales. To this end, we consider the same systems as above, 
-but evolve them for 25 orbital periods. In \autoref{fig:circular_orbit_comparison} we illustrate the evolution of these systems 
-by plotting the center of mass locations of the white dwarfs on the orbital ($xy$) plane. For the equal mass case in the inertial 
-reference frame, the curves essentially fall on top of each other, indicating that the stars are indeed orbiting at the initial 
-distance without substantial deviation. In the unequal mass case, the stars now orbit at different distances but do not much 
-deviate from their initial distance. We do observe a secular decline in the center of mass of these stars, indicating that 
-eventually the stellar orbits will degrade due to numerical error even without intervention. The co-rotating frame is different.
-For clarity of visualization, we rotate these results back into the inertial frame before displaying their orbits.
-The stellar orbit in the equal mass case does not show the same secular decline, and the system appears to be stable. However, 
-the cost is that the oscillations about this equilibrium are slightly larger than in the inertial frame. The unequal mass case displays 
-this effect starkly: while the average distance is not declining, there are huge oscillations about this average. In addition, 
-the stars move substantially within the rotating frame itself, yielding the chaotic-looking orbits in the figure. The oscillations 
-grow with time and will eventually cause the stars to merge; however, the stars would eventually merge even in the inertial frame 
-case, and at a faster rate than the equal mass case. This suggests that this mass ratio is too large to ignore tidal effects, and 
-that the tidal effects play out much differently in the rotating frame.
+We are most interested in the stability of these systems over long
+timescales. To this end, we consider the same systems as above, but
+evolve them for 25 orbital periods. In
+\autoref{fig:circular_orbit_comparison} we illustrate the evolution of
+these systems by plotting the center of mass locations of the white
+dwarfs on the orbital ($xy$) plane. For the equal mass case in the
+inertial reference frame, the curves essentially fall on top of each
+other, indicating that the stars are indeed orbiting at the initial
+distance without substantial deviation. In the unequal mass case, the
+stars now orbit at different distances but do not much deviate from
+their initial distance. We do observe a secular decline in the center
+of mass of these stars, indicating that eventually the stellar orbits
+will degrade due to numerical error even without intervention. The
+co-rotating frame is different.  For clarity of visualization, we
+rotate these results back into the inertial frame before displaying
+their orbits.  The stellar orbit in the equal mass case does not show
+the same secular decline, and the system appears to be
+stable. However, the cost is that the oscillations about this
+equilibrium are slightly larger than in the inertial frame. The
+unequal mass case displays this effect starkly: while the average
+distance is not declining, there are huge oscillations about this
+average. In addition, the stars move substantially within the rotating
+frame itself, yielding the chaotic-looking orbits in the figure. The
+oscillations grow with time and will eventually cause the stars to
+merge; however, the stars would eventually merge even in the inertial
+frame case, and at a faster rate than the equal mass case. This
+suggests that this mass ratio is too large to ignore tidal effects,
+and that the tidal effects play out much differently in the rotating
+frame. \MarginPar{this is interesting---has anyone else seen this?}
 
@@
   \includegraphics[scale=0.7]{plots/circular_orbit_comparison}
-  \caption{Positions of the white dwarfs in the orbital plane for four cases evolved over 25 orbital periods. 
-           The $x$ and $y$ axes are normalized to the size of the domain, so that $x = -0.5$ is the left edge 
-           and $x = 0.5$ is the right edge. The dashed blue curve is the position of the primary white dwarf, 
-           and the solid red curve is the position of the secondary. In plot (a) we have the equal mass system 
-           evolved in the inertial reference frame, and in plot (c) we have the same system evolved in a rotating frame, 
-           where the positions have been transformed back to the inertial frame for comparison. 
-           Plots (b) and (d) are analogous but for the unequal mass system.\label{fig:circular_orbit_comparison}}
+  \caption{Positions of the white dwarfs in the orbital plane for four
+    cases evolved over 25 orbital periods.  The $x$ and $y$ axes are
+    normalized to the size of the domain, so that $x = -0.5$ is the
+    left edge and $x = 0.5$ is the right edge. The dashed blue curve
+    is the position of the primary white dwarf, and the solid red
+    curve is the position of the secondary. In plot (a) we have the
+    equal mass system evolved in the inertial reference frame, and in
+    plot (c) we have the same system evolved in a rotating frame,
+    where the positions have been transformed back to the inertial
+    frame for comparison.  Plots (b) and (d) are analogous but for the
+    unequal mass system.\label{fig:circular_orbit_comparison}}
 \end{figure*}
 
-Turning to the conservation properties of the system, we examine as fairly typical cases the equal mass system 
-in the inertial frame for energy conservation (\autoref{fig:energy_conservation_equal}), and the unequal mass 
-system in the rotating frame for angular momentum conservation (\autoref{fig:angular_momentum_conservation_unequal}). 
-Angular momentum is conserved to within a few percent over the 25 orbits, while energy conservation is about 
-an order of magnitude better. We note that while this is already a fairly good level of energy conservation, it is not 
-nearly as good as the results of \citet{marcello:2012}. This is because we reset the internal energy to a level 
-corresponding to our temperature floor when it goes negative, while \citeauthor{marcello:2012} do not reset and 
-instead ignore the internal energy if it is negative. The resets impose an artificial floor on our ability to 
-conserve energy, but they only happen in low-density regions and do not much affect the large-scale dynamics. 
-Meanwhile, angular momentum conservation is not quite as good. This is linked to the decline (or increase) in 
-the size of the orbit. This implies that we ought to be careful in concluding that at these moderate resolutions 
-we can safely evolve systems for many dozens of orbits; this needs to be verified to ensure that an observed 
-inspiral and merger is physically (not numerically) motivated.
+Turning to the conservation properties of the system, we examine as
+fairly typical cases the equal mass system in the inertial frame for
+energy conservation (\autoref{fig:energy_conservation_equal}), and the
+unequal mass system in the rotating frame for angular momentum
+conservation (\autoref{fig:angular_momentum_conservation_unequal}).
+Angular momentum is conserved to within a few percent over the 25
+orbits, while energy conservation is about an order of magnitude
+better. We note that while this is already a fairly good level of
+energy conservation, it is not nearly as good as the results of
+\citet{marcello:2012}. This is because we reset the internal energy to
+a level corresponding to our temperature floor when it goes negative,
+while \citeauthor{marcello:2012} do not reset and instead ignore the
+internal energy if it is negative. The resets impose an artificial
+floor on our ability to conserve energy, but they only happen in
+low-density regions and do not much affect the large-scale dynamics.
+Meanwhile, angular momentum conservation is not quite as good.\MarginPar{``as good'' as what?}  This is
+linked to the decline (or increase) in the size of the orbit. This
+implies that we ought to be careful in concluding that at these
+moderate resolutions we can safely evolve systems for many dozens of
+orbits; this needs to be verified to ensure that an observed inspiral
+and merger is physically (not numerically) motivated.
 
@@
 
-We use OpenMP to accelerate the work associated with the boxes owned by each 
-MPI task. Originally \castro\ used OpenMP to accelerate individual loops in the 
-hydrodynamics routines, such as the piecewise-parabolic edge state reconstruction 
-and the conservative flux update. However, there is a significant amount of 
-overhead associated with generating a new OpenMP region at each of the many 
-different loops in a hydrodynamics algorithm. This makes such a strategy 
-sub-optimal for use on many-core processors and GPUs. We have recently switched 
-to a tiling approach, where an OpenMP region is generated at the start of 
-the hydrodynamics routine and the individual threads separately work on 
-different partitions of each box. This results in much less overhead for 
-the threading. In general we obtain more efficient simulations than 
-could be obtained using MPI only, because there are fewer boxes and thus 
-less communication for a given number of processor cores. We are currently 
-developing an approach to evaluating the equation of state and nuclear reactions 
-(and possibly the entire hydrodynamics algorithm) on GPUs, which will allow us
-to take advantage of the significant computational resources embedded in GPUs on 
-certain systems.
+We use OpenMP to accelerate the work associated with the boxes owned
+by each MPI task. Originally \castro\ used OpenMP to accelerate
+individual loops in the hydrodynamics routines, such as the
+piecewise-parabolic edge state reconstruction and the conservative
+flux update. However, there is a significant amount of overhead
+associated with generating a new OpenMP region at each of the many
+different loops in a hydrodynamics algorithm. This makes such a
+strategy sub-optimal for use on many-core processors and GPUs. We have
+recently switched to a tiling approach \MarginPar{ref Weiqun's paper}, where an OpenMP region is
+generated at the start of the hydrodynamics routine and the individual
+threads separately work on different partitions of each box. This
+results in much less overhead for the threading. In general we obtain
+more efficient simulations than could be obtained using MPI only,
+because there are fewer boxes and thus less communication for a given
+number of processor cores. We are currently developing an approach to
+evaluating the equation of state and nuclear reactions (and possibly
+the entire hydrodynamics algorithm) on GPUs, which will allow us to
+take advantage of the significant computational resources embedded in
+GPUs on certain systems.
 
@@
 
-To examine the parallel performance of \castro, we performed both strong scaling 
-and weak scaling tests on the Blue Waters machine at the National Center for 
-Supercomputing Applications. For the weak scaling test, whose results are 
-shown in \autoref{fig:weak_scaling}, we ran a uniform grid binary white 
-dwarf simulation for resolutions of $128^3$ zones through $2048^3$ zones. The 
-number of processors was scaled with the number of zones so that each processor 
-had the same amount of work; the smallest test used 8 processors and the largest 
-used 32,768 (note that the number of processor cores on a Blue Waters node is twice 
-the number of floating point units on that node). The test was run for 10 timesteps,
-with each timestep including two Poisson solves and a hydrodynamics update (though 
-for a uniform grid calculation we generally do not need to perform any multigrid 
-iterations for the first Poisson solve in a timestep, since the density distribution 
-has not changed since the end of the last timestep). We disabled plotfile and 
-checkpoint writing, as well as calculation of diagnostic information (the latter 
-can contribute to a significant fraction of the run time at large processor 
-counts if computed every timestep). We computed the median wall time required per 
-time step for each simulation, and then normalized this to the median time per 
-timestep for the smallest simulation. We find excellent weak scaling through 
-4,096 processors. At the largest run, the simulation time required is 
-more than 1.5 times the amount required for the smallest simulation. 
-This is due entirely to the increased cost of the multigrid Poisson solve 
-in each timestep and this cannot be mitigated except by improving 
-communication or computation efficiency in the multigrid solver. We observe 
-that this weak scaling behavior with Poisson gravity is a significant 
+To examine the parallel performance of \castro, we performed both
+strong scaling and weak scaling tests on the Blue Waters machine at
+the National Center for Supercomputing Applications. For the weak
+scaling test, whose results are shown in \autoref{fig:weak_scaling},
+we ran a uniform grid binary white dwarf simulation for resolutions of
+$128^3$ zones through $2048^3$ zones. \MarginPar{is this really weak scaling?} The number of processors was
+scaled with the number of zones so that each processor had the same
+amount of work; the smallest test used 8 processors and the largest
+used 32,768 (note that the number of processor cores on a Blue Waters
+node is twice the number of floating point units on that node). The
+test was run for 10 timesteps, with each timestep including two
+Poisson solves and a hydrodynamics update (though for a uniform grid
+calculation we generally do not need to perform any multigrid
+iterations for the first Poisson solve in a timestep, since the
+density distribution has not changed since the end of the last
+timestep). We disabled plotfile and checkpoint writing, as well as
+calculation of diagnostic information (the latter can contribute to a
+significant fraction of the run time at large processor counts if
+computed every timestep). We computed the median wall time required
+per time step for each simulation, and then normalized this to the
+median time per timestep for the smallest simulation. We find
+excellent weak scaling through 4,096 processors. At the largest run,
+the simulation time required is more than 1.5 times the amount
+required for the smallest simulation.  This is due entirely to the
+increased cost of the multigrid Poisson solve in each timestep and
+this cannot be mitigated except by improving communication or
+computation efficiency in the multigrid solver. We observe that this
+weak scaling behavior with Poisson gravity is a significant
 improvement over the results presented in the first \castro\ paper.
@@
 resolution. The strong scaling behavior is acceptable, though not perfect, 
-at very large processor counts when self-gravity is considered.
+at very large processor counts when self-gravity is considered. \MarginPar{does this use the new MG initialization of $\Phi$ from the old step?}
 

### 2015-08-14 papers/wdmerger_I/paper.tex
Deal with a number of recent comments

--- 2015-08-14 papers/wdmerger_I/paper.tex
@@
 \begin{equation}
-  f_{S} = \begin{dcases} 0 & r < r_S \\ \frac{1}{2}\left(1 - \text{cos}\left[\pi\left(\dfrac{r - r_S}{\Delta r_S}\right)\right]\right) & r_S \leq r < r_S^\prime \\ 1 & r \geq r_S^\prime. \end{dcases}
-\end{equation}
-For the simulations in this paper we set $r_S$ to be 75\% of the \MarginPar{maybe a plot of the sponge profile?}
+  f_{S} = \begin{dcases} 0 & r < r_S \\ \frac{1}{2}\left(1 - \text{cos}\left[\pi\left(\dfrac{r - r_S}{\Delta r_S}\right)\right]\right) & r_S \leq r < r_S^\prime \\ 1 & r \geq r_S^\prime.
+\end{dcases}\label{eq:sponge_frac}
+\end{equation}
+For the simulations in this paper we set $r_S$ to be 75\% of the 
 distance from the origin to the domain boundaries, and $\Delta r_S$ so
-that the sponge extends another 10\% of that distance. We set $\Delta
+that the sponge extends another 10\% of that distance. The resulting profile 
+is displayed in Figure \autoref{fig:sponge}. We set $\Delta
 t_S = 0.01$ s, which is of the same order as the as the CFL timestep
@@
 can create very large velocities in the ambient fluid that drag down
-the global timestep by up to an order of magnitude.  However, this
-provides another reason to distrust the dynamics of the low-density
-fluid outside of the stellar orbits. \MarginPar{I am not sure I understand 
-this statement}
+the global timestep by up to an order of magnitude.  This means, however, 
+that while the sponge is applied we should be cautious of imputing any physical 
+meaning to what is happening in the low-density gas far from the stars.
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.4]{plots/sponge}
+  \caption{Radial profile of the hydrodynamical sponge we apply (\autoref{eq:sponge_frac}). 
+           We subtract $f_S$ from unity; the value of $1 - f_S$ indicates what happens 
+           to the sponged function after the sponge is applied. The sponge has no effect 
+           in the inner part of the domain, and is fully applied at the outer edge.\label{fig:sponge}}
+\end{figure}
 
@@
 smaller with resolution. The star is only approximately in equilibrium 
-when the coarse grid of $256^3$ zones has a level of refinement that
-jumps by a factor of four. However, even then there is a slight uptick
-in the size toward the event, implying that the numerical stability is
-not guaranteed for arbitrarily long timescales. This result suggests
+by this measure when the coarse grid of $256^3$ zones has a level of 
+refinement that jumps by a factor of four. However, even then there is a slight uptick
+in the size toward the end, implying that the numerical stability is
+not guaranteed for arbitrarily long timescales. For another view, we 
+consider the kinetic energy on the grid, in \autoref{fig:single_star_static_ke}. 
+This is a more holistic measure that weights the contribution by the density. 
+At the end of the simulation the kinetic energy is not lower at the highest 
+resolution than at the lower resolutions. This result suggests
 that when doing the equilibrium initial models that will former the
@@
 merger, to ensure that the merger is due to this applied force and not
-the intrinsic numerical instability of the stars. \MarginPar{you never referenced figure 4}
-
+the intrinsic numerical instability of the stars.
 \begin{figure}
@@
   \includegraphics[scale=0.45]{plots/single_star_static_ke}
-  \caption{Time evolution of the kinetic energy of a $0.9 \msolar$ 
+  \caption{Time evolution of the kinetic energy of a $0.9\, \msolar$ 
     white dwarf. The lines have the same meaning as in \autoref{fig:single_star_static_radius}.
@@
 test of \autoref{sec:HSE} with a bulk velocity on the grid. We chose a
-velocity of $2.56 \times 10^{8}\ \text{cm s}^{-1}$.  \MarginPar{Approx
-  what Mach number is this?}  {For context, this is comparable to the
-  orbital velocities of the stars in \autoref{sec:kepler}.  This test
+velocity of $2.56 \times 10^{8}\ \text{cm s}^{-1}$. For context, this is 
+comparable to the orbital velocities of the stars in \autoref{sec:kepler}, and the Mach
+number is of order unity in the stellar core at this speed.  This test
 was inspired by \citet{tasker:2008}, who considered a moving galaxy
@@
 
-Another that we consider is the energy conservation of the system. Recalling \autoref{sec:gravity_hydro_coupling}, 
-there are several different methods of applying the gravitational source term to the hydrodynamics equations. In \castro\ we presently 
-have four options, controlled by the parameter {\tt castro.grav\_source\_type}, which we will shortern to {\tt gs} for the present 
-discussion. ${\tt gs} = 1$ is the standard cell-centered source term for gravity. ${\tt gs} = 2$ is a slight variation on this scheme 
-that determines the value of the energy source term after the momentum source term has been applied; the former approach uses 
-the uncorrected momenta in calculating $\rho \mathbf{u} g$, which we have found to be more accurate. ${\tt gs} = 3$ is entirely 
-different: after calculating the new momenta, we reset the total energy to be equal to the internal energy plus the 
-kinetic energy. This has the virtue of ensuring that there is no conflict due to discretization between the momentum and 
-energy equations, and also correctly ensuring that the gravitational force does not directly change the internal energy---and thus the temperature---of the fluid.
-However, it explicitly sacrifices total energy conservation.
-${\tt gs} = 4$ is
-the new conservative method of evaluating the energy source terms at cell faces. The results for the change in energy 
-after a single orbit are seen in the first column of \autoref{table:sources}. The first two versions give reasonable and 
-similar levels of energy conservation. The third has total energy changes on the order of 100\%, but this itself 
-does not have a severe effect on the dynamics because in this scheme the total energy variable is effectively a placeholder 
-value of the kinetic energy plus internal energy, rather than being evolved directly. The last scheme is nearly two orders 
-of magnitude better in energy conservation, justifying the effort in varying the scheme.
+Another that we consider is the energy conservation of the
+system. Recalling \autoref{sec:gravity_hydro_coupling}, there are
+several different methods of applying the gravitational source term to
+the hydrodynamics equations. In \castro\ we presently have four
+options, controlled by the parameter {\tt castro.grav\_source\_type},
+which we will shortern to {\tt gs} for the present discussion. 
+${\tt gs} = 1$ and ${\tt gs} = 2$ are variations on the standard 
+cell-centered source term for gravity. The difference between them is that 
+${\tt gs} = 2$ determines the value of the energy source term after the momentum
+source term has been applied, while ${\tt gs} = 1$ uses the uncorrected
+momenta in calculating $\rho \mathbf{u} g$. We have found ${\tt gs} = 2$ to be
+more accurate. ${\tt gs} = 3$ is entirely different: after calculating
+the new momenta, we reset the total energy to be equal to the internal
+energy plus the kinetic energy. This has the virtue of ensuring that
+there is no conflict due to discretization between the momentum and
+energy equations, and also correctly ensuring that the gravitational
+force does not directly change the internal energy---and thus the
+temperature---of the fluid. However, it explicitly sacrifices total
+energy conservation. ${\tt gs} = 4$ is the new conservative method of
+evaluating the energy source terms at cell faces. The results for the
+change in energy after a single orbit are seen in the first column of
+\autoref{table:sources}. The first two versions give reasonable and
+similar levels of energy conservation. The third has total energy
+changes on the order of 100\%, but this itself does not have a severe
+effect on the dynamics because in this scheme the total energy
+variable is effectively a placeholder value of the kinetic energy plus
+internal energy, rather than being evolved directly. The last scheme
+is nearly two orders of magnitude better in energy conservation,
+justifying the effort in varying the scheme.
 
@@
 
-We are most interested in the stability of these systems over long timescales. To this end, we consider the same systems as above, 
-but evolve them for 25 orbital periods. In \autoref{fig:circular_orbit_comparison} we illustrate the evolution of these systems 
-by plotting the center of mass locations of the white dwarfs on the orbital ($xy$) plane. For the equal mass case in the inertial 
-reference frame, the curves essentially fall on top of each other, indicating that the stars are indeed orbiting at the initial 
-distance without substantial deviation. In the unequal mass case, the stars now orbit at different distances but do not much 
-deviate from their initial distance. We do observe a secular decline in the center of mass of these stars, indicating that 
-eventually the stellar orbits will degrade due to numerical error even without intervention. The co-rotating frame is different.
-For clarity of visualization, we rotate these results back into the inertial frame before displaying their orbits.
-The stellar orbit in the equal mass case does not show the same secular decline, and the system appears to be stable. However, 
-the cost is that the oscillations about this equilibrium are slightly larger than in the inertial frame. The unequal mass case displays 
-this effect starkly: while the average distance is not declining, there are huge oscillations about this average. In addition, 
-the stars move substantially within the rotating frame itself, yielding the chaotic-looking orbits in the figure. The oscillations 
-grow with time and will eventually cause the stars to merge; however, the stars would eventually merge even in the inertial frame 
-case, and at a faster rate than the equal mass case. This suggests that this mass ratio is too large to ignore tidal effects, and 
-that the tidal effects play out much differently in the rotating frame.
+We are most interested in the stability of these systems over long
+timescales. To this end, we consider the same systems as above, but
+evolve them for 25 orbital periods. In
+\autoref{fig:circular_orbit_comparison} we illustrate the evolution of
+these systems by plotting the center of mass locations of the white
+dwarfs on the orbital ($xy$) plane. For the equal mass case in the
+inertial reference frame, the curves essentially fall on top of each
+other, indicating that the stars are indeed orbiting at the initial
+distance without substantial deviation. In the unequal mass case, the
+stars now orbit at different distances but do not much deviate from
+their initial distance. We do observe a secular decline in the center
+of mass of these stars, indicating that eventually the stellar orbits
+will degrade due to numerical error even without intervention. The
+co-rotating frame is different.  For clarity of visualization, we
+rotate these results back into the inertial frame before displaying
+their orbits.  The stellar orbit in the equal mass case does not show
+the same secular decline, and the system appears to be
+stable. However, the cost is that the oscillations about this
+equilibrium are slightly larger than in the inertial frame. The
+unequal mass case displays this effect starkly: while the average
+distance is not declining, there are huge oscillations about this
+average. In addition, the stars move substantially within the rotating
+frame itself, yielding the chaotic-looking orbits in the figure. The
+oscillations grow with time and will eventually cause the stars to
+merge; however, the stars would eventually merge even in the inertial
+frame case, and at a faster rate than the equal mass case. This
+suggests that this mass ratio is too large to ignore tidal effects,
+and that the tidal effects play out much differently in the rotating
+frame.
 
@@
   \includegraphics[scale=0.7]{plots/circular_orbit_comparison}
-  \caption{Positions of the white dwarfs in the orbital plane for four cases evolved over 25 orbital periods. 
-           The $x$ and $y$ axes are normalized to the size of the domain, so that $x = -0.5$ is the left edge 
-           and $x = 0.5$ is the right edge. The dashed blue curve is the position of the primary white dwarf, 
-           and the solid red curve is the position of the secondary. In plot (a) we have the equal mass system 
-           evolved in the inertial reference frame, and in plot (c) we have the same system evolved in a rotating frame, 
-           where the positions have been transformed back to the inertial frame for comparison. 
-           Plots (b) and (d) are analogous but for the unequal mass system.\label{fig:circular_orbit_comparison}}
+  \caption{Positions of the white dwarfs in the orbital plane for four
+    cases evolved over 25 orbital periods.  The $x$ and $y$ axes are
+    normalized to the size of the domain, so that $x = -0.5$ is the
+    left edge and $x = 0.5$ is the right edge. The dashed blue curve
+    is the position of the primary white dwarf, and the solid red
+    curve is the position of the secondary. In plot (a) we have the
+    equal mass system evolved in the inertial reference frame, and in
+    plot (c) we have the same system evolved in a rotating frame,
+    where the positions have been transformed back to the inertial
+    frame for comparison.  Plots (b) and (d) are analogous but for the
+    unequal mass system.\label{fig:circular_orbit_comparison}}
 \end{figure*}
 
-Turning to the conservation properties of the system, we examine as fairly typical cases the equal mass system 
-in the inertial frame for energy conservation (\autoref{fig:energy_conservation_equal}), and the unequal mass 
-system in the rotating frame for angular momentum conservation (\autoref{fig:angular_momentum_conservation_unequal}). 
-Angular momentum is conserved to within a few percent over the 25 orbits, while energy conservation is about 
-an order of magnitude better. We note that while this is already a fairly good level of energy conservation, it is not 
-nearly as good as the results of \citet{marcello:2012}. This is because we reset the internal energy to a level 
-corresponding to our temperature floor when it goes negative, while \citeauthor{marcello:2012} do not reset and 
-instead ignore the internal energy if it is negative. The resets impose an artificial floor on our ability to 
-conserve energy, but they only happen in low-density regions and do not much affect the large-scale dynamics. 
-Meanwhile, angular momentum conservation is not quite as good. This is linked to the decline (or increase) in 
-the size of the orbit. This implies that we ought to be careful in concluding that at these moderate resolutions 
-we can safely evolve systems for many dozens of orbits; this needs to be verified to ensure that an observed 
-inspiral and merger is physically (not numerically) motivated.
+Turning to the conservation properties of the system, we examine as
+fairly typical cases the equal mass system in the inertial frame for
+energy conservation (\autoref{fig:energy_conservation_equal}), and the
+unequal mass system in the rotating frame for angular momentum
+conservation (\autoref{fig:angular_momentum_conservation_unequal}).
+Angular momentum is conserved to within a few percent over the 25
+orbits, while energy conservation is about an order of magnitude
+better. We note that while this is already a fairly good level of
+energy conservation, it is not nearly as good as the results of
+\citet{marcello:2012}. This is because we reset the internal energy to
+a level corresponding to our temperature floor when it goes negative,
+while \citeauthor{marcello:2012} do not reset and instead ignore the
+internal energy if it is negative. The resets impose an artificial
+floor on our ability to conserve energy, but they only happen in
+low-density regions and do not much affect the large-scale dynamics.
+Meanwhile, relative angular momentum conservation is not quite as good 
+as relative energy conservation.  This is
+linked to the decline (or increase) in the size of the orbit. This
+implies that we ought to be careful in concluding that at these
+moderate resolutions we can safely evolve systems for many dozens of
+orbits; this needs to be verified to ensure that an observed inspiral
+and merger is physically (not numerically) motivated.
 
@@
 
-We use OpenMP to accelerate the work associated with the boxes owned by each 
-MPI task. Originally \castro\ used OpenMP to accelerate individual loops in the 
-hydrodynamics routines, such as the piecewise-parabolic edge state reconstruction 
-and the conservative flux update. However, there is a significant amount of 
-overhead associated with generating a new OpenMP region at each of the many 
-different loops in a hydrodynamics algorithm. This makes such a strategy 
-sub-optimal for use on many-core processors and GPUs. We have recently switched 
-to a tiling approach, where an OpenMP region is generated at the start of 
-the hydrodynamics routine and the individual threads separately work on 
-different partitions of each box. This results in much less overhead for 
-the threading. In general we obtain more efficient simulations than 
-could be obtained using MPI only, because there are fewer boxes and thus 
-less communication for a given number of processor cores. We are currently 
-developing an approach to evaluating the equation of state and nuclear reactions 
-(and possibly the entire hydrodynamics algorithm) on GPUs, which will allow us
-to take advantage of the significant computational resources embedded in GPUs on 
-certain systems.
+We use OpenMP to accelerate the work associated with the boxes owned
+by each MPI task. Originally \castro\ used OpenMP to accelerate
+individual loops in the hydrodynamics routines, such as the
+piecewise-parabolic edge state reconstruction and the conservative
+flux update. However, there is a significant amount of overhead
+associated with generating a new OpenMP region at each of the many
+different loops in a hydrodynamics algorithm. This makes such a
+strategy sub-optimal for use on many-core processors and GPUs. We have
+recently switched to a tiling approach where an OpenMP region is
+generated at the start of the hydrodynamics routine and the individual
+threads separately work on different partitions of each box. This
+results in much less overhead for the threading. In general we obtain
+more efficient simulations than could be obtained using MPI only,
+because there are fewer boxes and thus less communication for a given
+number of processor cores. We are currently developing an approach to
+evaluating the equation of state and nuclear reactions (and possibly
+the entire hydrodynamics algorithm) on GPUs, which will allow us to
+take advantage of the significant computational resources embedded in
+GPUs on certain systems.
 
@@
 
-To examine the parallel performance of \castro, we performed both strong scaling 
-and weak scaling tests on the Blue Waters machine at the National Center for 
-Supercomputing Applications. For the weak scaling test, whose results are 
-shown in \autoref{fig:weak_scaling}, we ran a uniform grid binary white 
-dwarf simulation for resolutions of $128^3$ zones through $2048^3$ zones. The 
-number of processors was scaled with the number of zones so that each processor 
-had the same amount of work; the smallest test used 8 processors and the largest 
-used 32,768 (note that the number of processor cores on a Blue Waters node is twice 
-the number of floating point units on that node). The test was run for 10 timesteps,
-with each timestep including two Poisson solves and a hydrodynamics update (though 
-for a uniform grid calculation we generally do not need to perform any multigrid 
-iterations for the first Poisson solve in a timestep, since the density distribution 
-has not changed since the end of the last timestep). We disabled plotfile and 
-checkpoint writing, as well as calculation of diagnostic information (the latter 
-can contribute to a significant fraction of the run time at large processor 
-counts if computed every timestep). We computed the median wall time required per 
-time step for each simulation, and then normalized this to the median time per 
-timestep for the smallest simulation. We find excellent weak scaling through 
-4,096 processors. At the largest run, the simulation time required is 
-more than 1.5 times the amount required for the smallest simulation. 
-This is due entirely to the increased cost of the multigrid Poisson solve 
-in each timestep and this cannot be mitigated except by improving 
-communication or computation efficiency in the multigrid solver. We observe 
-that this weak scaling behavior with Poisson gravity is a significant 
+To examine the parallel performance of \castro, we performed both
+strong scaling and weak scaling tests on the Blue Waters machine at
+the National Center for Supercomputing Applications. For the weak
+scaling test, whose results are shown in \autoref{fig:weak_scaling},
+we ran a uniform grid binary white dwarf simulation for resolutions of
+$128^3$ zones through $2048^3$ zones. The number of processors was
+scaled with the number of zones so that each processor had the same
+amount of work; the smallest test used 8 processors and the largest
+used 32,768 (note that the number of processor cores on a Blue Waters
+node is twice the number of floating point units on that node). The
+test was run for 10 timesteps, with each timestep including two
+Poisson solves and a hydrodynamics update (though for a uniform grid
+calculation we generally do not need to perform any multigrid
+iterations for the first Poisson solve in a timestep, since the
+density distribution has not changed since the end of the last
+timestep). We disabled plotfile and checkpoint writing, as well as
+calculation of diagnostic information (the latter can contribute to a
+significant fraction of the run time at large processor counts if
+computed every timestep). We computed the median wall time required
+per time step for each simulation, and then normalized this to the
+median time per timestep for the smallest simulation. We find
+excellent weak scaling through 4,096 processors. At the largest run,
+the simulation time required is more than 1.5 times the amount
+required for the smallest simulation.  This is due entirely to the
+increased cost of the multigrid Poisson solve in each timestep and
+this cannot be mitigated except by improving communication or
+computation efficiency in the multigrid solver. We observe that this
+weak scaling behavior with Poisson gravity is a significant
 improvement over the results presented in the first \castro\ paper.

### 2015-08-14 papers/wdmerger_I/plots/sponge.py

@@ -1,0 +1,37 @@
+# This Python program is used to create a plot displaying the sponge 
+# function we use in the CASTRO hydrodynamics for the wdmerger problem.
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+def sponge(r):
+    sp
+
+rs = 0.75
+rt = 0.85
+
+r = np.linspace(0.0, 1.0, 1000)
+f = np.zeros(len(r))
+
+idx = np.where(r < rs)
+f[idx] = 0.0
+
+idx = np.where(r < rt)
+idx = np.where(r[idx] >= rs)
+f[idx] = 0.5 * (1.0 - np.cos(np.pi * (r[idx] - rs) / (rt - rs)))
+
+idx = np.where(r >= rt)
+f[idx] = 1.0
+
+plt.plot(r, 1.0 - f, linewidth=4.0)
+plt.xlabel('Radius', fontsize=20)
+plt.ylabel(r'$1 - f_S$', fontsize=20)
+
+plt.xlim([0.0, 1.0])
+plt.ylim([-0.05, 1.05])
+
+plt.tick_params(labelsize=16)
+
+plt.tight_layout()
+
+plt.savefig('sponge.eps')

### 2015-08-18 papers/wdmerger_I/paper.tex
Rewording of the section on why we cannot integrate the work done to get the change in gravitational energy.

@@
 converted into a surface integral by use of the divergence theorem (see their Equation 93). Then we can simply 
-evaluate the work done at each zone face. However, this application of the divergence theorem appears to be 
-mathematically incorrect, despite the fact that this is very similar in spirit to the approach discussed here. 
-The author fortuitously obtains the correct factor of $1/2$ corresponding to the change in gravitational self-energy.
-However, some reflection reveals that there should be no way to derive this double-counting factor simply by integrating 
-the work done: that factor is derived from the unique nature of the gravitational potential, which is generated 
-by the same distribution that it exerts a force on for a simulation with self-gravity, and it has to be added in 
-intentionally to ensure conservation of total energy for these particular types of potentials. We can apply 
-the approach of considering energy transfer between the gas energy and the potential energy to any potential field, 
-whether or not it is static, changing in time, or generated by the fluid distribution itself. 
-Indeed, we will do so for the rotational potential in \autoref{sec:rotation}, where the 
-factor of $1/2$ does not appear. However, it is important also to note that this is not related to a 
-global flow of energy or the global nature of the gravitational potential, as claimed by \cite{marcello:2012}. 
-This double-counting factor would apply even if there were only two point masses on the grid, since 
-work would only be required to bring in the second mass from infinity; the first one can be 
-placed at no energy cost. This transfer of energy is explicitly \textit{local} because the 
-mass transfer is local. It is true that the mass transfer changes the potential globally, 
-but that is implicitly accounted for when using the time-centered potential, and no 
-additional steps need to be taken.
+evaluate the work done at each zone face. However, this application of the divergence theorem is 
+mathematically incorrect, despite the fact that this is similar in spirit to the approach discussed here. 
+Moreover, using the divergence theorem to integrate the work done in a zone can never obtain 
+the factor of $1/2$ corresponding to the change in gravitational self-energy. That factor can be thought of
+as a double-counting factor that ensures that we correctly capture the energy of a pairwise interaction. 
+When one integrates the work done in the zone, there is a contribution to the energy that cannot be 
+expressed as a full time derivative or the divergence of a flux. This reflects the 
+global nature of the gravitational potential and means that we cannot get the correct energy expression 
+unless we integrate over the entire mass distribution; see \citet[Chapter 4]{shu:1992} and \citet{marcello:2012}. 
+However, this approach is still valid when considering energy transfer between the gas energy and a  
+potential field, if the change in energy is explicitly local. In the case of gravity, the mass flux
+that generates a change in energy is local, but the change in energy is stored in the entire gravitational 
+field, and is only captured by a global integral because of the long-range nature of the potential.
 

### 2015-08-20 papers/wdmerger_I/paper.tex
Fixed typos. Added a comment.  Added a reference to BoxLib tiling.

@@
 describe simulations of these systems using a hydrodynamics code with 
-using the adaptive-mesh refinement approach. In this paper we describe our numerical 
+the adaptive-mesh refinement approach. In this paper we describe our numerical 
 methodology and discuss our implementation in the code \castro. \castro\ 
@@
 through direct numerical simulation is critical for building 
-confidence in this progenitor model.
+confidence in this progenitor model. \MarginPar{Direct numerical
+  simulation has a specific meaning to CFD community.  That is the
+  simulation resolve the viscous scale.  So maybe we should remove the
+word 'direct'.}
 
@@
 is displayed in Figure \autoref{fig:sponge}. We set $\Delta
-t_S = 0.01$ s, which is of the same order as the as the CFL timestep
+t_S = 0.01$ s, which is of the same order as the CFL timestep
 for typical problem setups. We find that this sponge is necessary
@@
 and is only partially mitigated by evolving the white dwarfs in a
-co-rotating frame. It is therefore important to be be aware of the
+co-rotating frame. It is therefore important to be aware of the
 behavior of the code in such circumstances. The behavior of \castro\ for
@@
 generated at the start of the hydrodynamics routine and the individual
-threads separately work on different partitions of each box. This
+threads separately work on different partitions of each box \citep{boxlib-tiling}. This
 results in much less overhead for the threading. In general we obtain

### 2015-08-25 papers/wdmerger_I/paper.tex
Adopt the WD composition distribution of Dan et al. (2012), abandoning the old pure CO WD approach. This prompted a number of changes to the probdata module and the way initial model data is stored, including the addition of a new derived type for that data.

@@
 dwarfs by integrating the equation of hydrostatic equilibrium, taking
-the temperature and composition to be constant, and using the
+the temperature to be constant, and using the
 stellar equation of state.  This results in a single non-linear
@@
 spacing to be equal to the 3D grid spacing.
+
+The white dwarf composition is determined by the chosen mass. For 
+this paper we adopt the scheme of \cite{dan:2012}. Low-mass WDs 
+are pure helium; low-to-intermediate-mass WDs are an even carbon-oxygen 
+core with a relatively large helium envelope; intermediate-mass 
+WDs are a carbon-oxygen core with slightly more oxygen than carbon; 
+and, high-mass WDs are composed of oxygen, neon, and magnesium. 
+This choice of composition distribution broadly resembles the 
+results of stellar evolution calculations in the respective 
+mass ranges, though it does not match the calculations in detail.
 

### 2015-08-25 source/Prob_3d.f90

@@ -86,10 +86,10 @@
               dist_P = loc - center_P_initial
               dist_S = loc - center_S_initial
 
-              if (sum(dist_P**2) < radius_P_initial**2) then
-                 call interpolate_3d_from_1d(model_P_r, model_P_state, initial_model_npts, dist_P, delta, zone_state, nsub)
-              else if (sum(dist_S**2) < radius_S_initial**2) then
-                 call interpolate_3d_from_1d(model_S_r, model_S_state, initial_model_npts, dist_S, delta, zone_state, nsub)
+              if (sum(dist_P**2) < model_P % radius**2) then
+                 call interpolate_3d_from_1d(model_P, dist_P, delta, zone_state, nsub)
+              else if (sum(dist_S**2) < model_S % radius**2) then
+                 call interpolate_3d_from_1d(model_S, dist_S, delta, zone_state, nsub)
               else
                  zone_state = ambient_state
               endif
@@ -135,9 +135,9 @@
                 dist_P = loc - center_P_initial
                 dist_S = loc - center_S_initial
 
-                if (sum(dist_P**2) < radius_P_initial**2) then
+                if (sum(dist_P**2) < model_P % radius**2) then
                    state(i,j,k,UMX:UMZ) = state(i,j,k,UMX:UMZ) + vel_P(:) * state(i,j,k,URHO)
-                else if (sum(dist_S**2) < radius_S_initial**2) then
+                else if (sum(dist_S**2) < model_S % radius**2) then
                    state(i,j,k,UMX:UMZ) = state(i,j,k,UMX:UMZ) + vel_S(:) * state(i,j,k,URHO)
                 endif
 

### 2015-08-25 source/Problem.f90

@@ -642,7 +642,9 @@ subroutine setup_scf_relaxation(dx, problo, probhi)
     integer          :: ncell(3)
 
     double precision :: pos_l(3)
-    type (eos_t) :: eos_state
+    type (eos_t) :: eos_state, ambient_state
+
+    call get_ambient(ambient_state)
 
     ! Following Swesty, Wang and Calder (2000), and Motl et al. (2002), 
     ! we need to fix three points to uniquely determine an equilibrium 
@@ -705,17 +707,17 @@ subroutine setup_scf_relaxation(dx, problo, probhi)
 
     ! Convert the maximum densities into maximum enthalpies.
 
-    eos_state % T   = stellar_temp
-    eos_state % rho = central_density_P
-    eos_state % xn  = stellar_comp
+    eos_state % T   = model_P % central_temp
+    eos_state % rho = model_P % central_density
+    eos_state % xn  = model_P % core_comp
 
     call eos(eos_input_rt, eos_state)
 
     h_max_P = eos_state % h
 
-    eos_state % T   = stellar_temp
-    eos_state % rho = central_density_S
-    eos_state % xn  = stellar_comp
+    eos_state % T   = model_S % central_temp
+    eos_state % rho = model_S % central_density
+    eos_state % xn  = model_S % core_comp
 
     call eos(eos_input_rt, eos_state)
     
@@ -726,9 +728,7 @@ subroutine setup_scf_relaxation(dx, problo, probhi)
     ! against trying to compute a corresponding temperature
     ! in zones where the enthalpy is just too low for convergence.
 
-    eos_state % T   = stellar_temp
-    eos_state % rho = ambient_density
-    eos_state % xn  = stellar_comp
+    eos_state = ambient_state
 
     do while (eos_state % rho < 1.d11)
 

### 2015-08-25 source/initial_model.f90

@@ -4,332 +4,301 @@
 
 module initial_model_module
 
-use bl_types
-use bl_constants_module
-use bl_error_module, only: bl_error
-use eos_module, only: eos_input_rt, eos
-use eos_type_module, only: eos_t
-use network, only: nspec
-use model_parser_module, only: itemp_model, idens_model, ipres_model, ispec_model
-use fundamental_constants_module, only: Gconst, M_solar
-use interpolate_module, only: interpolate
+  use bl_types
+  use bl_constants_module
+  use bl_error_module, only: bl_error
+  use eos_module, only: eos_input_rt, eos
+  use eos_type_module, only: eos_t
+  use network, only: nspec
+  use model_parser_module, only: itemp_model, idens_model, ipres_model, ispec_model
+  use fundamental_constants_module, only: Gconst, M_solar
+  use interpolate_module, only: interpolate
 
+  type :: initial_model
+
+     ! Physical characteristics
+
+     double precision :: mass, envelope_mass
+     double precision :: central_density, central_temp
+     double precision :: max_density, min_density
+     double precision :: radius
+
+     ! Composition
+
+     double precision :: core_comp(nspec), envelope_comp(nspec)
+
+     ! Model storage
+
+     double precision :: dx
+     integer          :: npts
+     double precision :: mass_tol, hse_tol
+
+     double precision, allocatable :: r(:), rl(:), rr(:)
+     double precision, allocatable :: M_enclosed(:), g(:)
+     type (eos_t), allocatable :: state(:)
+
+  end type initial_model
+  
 contains
 
-  subroutine init_1d(model_r, model_hse, nx, dx, hse_tol, mass_tol, &
-                     radius, mass, central_density, temp_core, xn_core, ambient_state)
+  subroutine initialize_model(model, dx, npts, mass_tol, hse_tol)
+
+    type (initial_model) :: model
+    integer :: npts
+    double precision :: dx, mass_tol, hse_tol
+
+    integer :: i
+
+    model % dx = dx
+    model % npts = npts
+    model % mass_tol = mass_tol
+    model % hse_tol = hse_tol
+
+    allocate(model % r(npts))
+    allocate(model % rl(npts))
+    allocate(model % rr(npts))
+    allocate(model % M_enclosed(npts))
+    allocate(model % g(npts))
+    allocate(model % state(npts))
+
+    do i = 1, npts
+
[…]

### 2015-08-25 source/probdata.f90

@@ -3,26 +3,27 @@ module probdata_module
   use network, only: nspec, network_species_index
   use eos_type_module, only: eos_t
   use eos_module, only: eos_input_rt, eos
+  use bl_constants_module, only: ZERO, THIRD, HALF, ONE, TWO, THREE, M_PI, FOUR
+  use fundamental_constants_module, only: Gconst, M_solar, AU
+  use initial_model_module, only: initial_model
 
   ! Probin file
   character (len=:), allocatable :: probin
 
-  ! Determine if we are the I/O processor
+  ! For determining if we are the I/O processor.
   integer :: ioproc
   
-  ! Model info
-  character (len=80) :: model_P_name
-  character (len=80) :: model_S_name
-  
-  double precision, allocatable :: model_P_r(:), model_P_state(:,:)
-  double precision, allocatable :: model_S_r(:), model_S_state(:,:)
-
   ! Initial binary orbit characteristics
-  double precision :: mass_P_initial, mass_S_initial
+  ! Note that the envelope mass is included within the total mass of the star
+  double precision :: mass_P, mass_S
   double precision :: central_density_P, central_density_S
-  double precision :: radius_P_initial, radius_S_initial
   double precision :: orbital_speed_P, orbital_speed_S
-  double precision :: stellar_temp, stellar_comp(nspec)
+  double precision :: stellar_temp
+  double precision :: primary_envelope_mass, secondary_envelope_mass
+  double precision :: primary_envelope_comp(nspec), secondary_envelope_comp(nspec)
+
+  ! Ambient medium
+  double precision :: ambient_density, ambient_temp, ambient_comp(nspec)
 
   ! Smallest allowed mass fraction
   double precision :: smallx
@@ -51,18 +52,14 @@ module probdata_module
   double precision :: damping_alpha
 
   ! Binary properties
-  double precision :: a_P_initial, a_S_initial, a
-  
-  double precision, dimension(3) :: center_P_initial, center_S_initial
+  double precision :: a_P_initial, a_S_initial, a  
+  double precision :: center_P_initial(3), center_S_initial(3)
 
   integer :: star_axis
   integer :: initial_motion_dir
 
   ! Location of the physical center of the problem, as a fraction of domain size
   double precision :: center_fracx, center_fracy, center_fracz
-
-  ! Density of ambient medium
[…]

### 2015-08-31 source/Prob.cpp
Move SCF relaxation stuff to its own source files.

@@ -1,6 +1,5 @@
 #include "Castro.H"
 #include "Castro_F.H"
-#include "Problem_F.H"
 
 #include "Gravity.H"
 
@@ -332,267 +331,12 @@ void Castro::problem_post_init() {
 	BoxLib::Error();
       }
 
-      int finest_level = parent->finestLevel();
+      int relax_type = 0;
 
-      MultiFab& S_new = get_new_data(State_Type);
+      BL_FORT_PROC_CALL(GET_RELAX_TYPE,get_relax_type)(relax_type);
 
-      int j = 1;
-      int relax_max_iterations = 30;
-
-      const Real* dx  = parent->Geom(level).CellSize();
-
-      const Real* problo = parent->Geom(level).ProbLo();
-      const Real* probhi = parent->Geom(level).ProbHi();
-
-      const int*  domlo = geom.Domain().loVect();
-      const int*  domhi = geom.Domain().hiVect();
-
-      MultiFab gcoeff(grids,1,0,Fab_allocate);
-      gcoeff.setVal(0.0);
-
-      Box box_A( IntVect(), IntVect(2, 2, 2) );
-      Box box_B( IntVect(), IntVect(2, 2, 2) );
-      Box box_C( IntVect(), IntVect(2, 2, 2) );
-
-      FArrayBox cA(box_A);
-      FArrayBox cB(box_B);
-      FArrayBox cC(box_C);
-
-      cA.setVal(0.0);
-      cB.setVal(0.0);
-      cC.setVal(0.0);
-
-      int loc_A[3] = {0};
-      int loc_B[3] = {0};
-      int loc_C[3] = {0};
-
-      BL_FORT_PROC_CALL(SETUP_SCF_RELAXATION,setup_scf_relaxation)(dx, problo, probhi);
-
-      BL_FORT_PROC_CALL(GET_COEFF_INFO,get_coeff_info)(loc_A, loc_B, loc_C, cA.dataPtr(), cB.dataPtr(), cC.dataPtr());
-
-      // Get the phi MultiFab.
-
-      MultiFab& phi = get_new_data(PhiGrav_Type);
-
-      int ns          = NUM_STATE;
-      Real cur_time   = state[State_Type].curTime();
-
-      // Iterate until the system is relaxed by filling the level data 
[…]

### 2015-08-31 source/Problem.H

@@ -28,7 +28,10 @@ void gwstrain (Real time, Real& h_plus_rot, Real& h_cross_rot,
 
 void problem_post_init();
 
+// Self-consistent field relaxation routine.
+
+void scf_relaxation();
+
 // Parameter that controls whether we do a post-initialization relaxation.
 
 static int do_relax;
-

### 2015-08-31 source/Problem.f90

@@ -626,551 +626,6 @@ end subroutine get_roche_radii
 
 
 
-subroutine setup_scf_relaxation(dx, problo, probhi)
-
-    use bl_constants_module, only: HALF, ONE, TWO
-    use prob_params_module, only: center
-    use eos_module
-    use probdata_module
-
-    implicit none
-
-    double precision :: dx(3), problo(3), probhi(3)
-
-    double precision :: x, y, z
-    integer          :: n
-    integer          :: ncell(3)
-
-    double precision :: pos_l(3)
-    type (eos_t) :: eos_state, ambient_state
-
-    call get_ambient(ambient_state)
-
-    ! Following Swesty, Wang and Calder (2000), and Motl et al. (2002), 
-    ! we need to fix three points to uniquely determine an equilibrium 
-    ! configuration for two unequal mass stars. We can do this by 
-    ! specifying d_A, the distance from the center of mass to the 
-    ! inner point of the primary; d_B, the distance from the center 
-    ! of mass to the inner point of the secondary; and d_C, the 
-    ! distance from the center of mass to the outer point of the secondary. 
-    ! We give these widths in physical units, so in general these will 
-    ! be locations on the grid that are not coincident with a corner. 
-    ! If a location is at point (x, y, z), then we find the eight 
-    ! zone centers that surround this point, and do a tri-linear
-    ! reconstruction to estimate the relative weight of each of the eight 
-    ! zone centers in determining the value at that point.
-
-    ncell = NINT( (probhi - problo) / dx )
-
-    do n = 1, 3
-       d_vector(:,n) = center
-       if (n .eq. 1) then
-          d_vector(star_axis,n) = d_vector(star_axis,n) - d_A
-       else if (n .eq. 2) then
-          d_vector(star_axis,n) = d_vector(star_axis,n) + d_B
-       else if (n .eq. 3) then
-          d_vector(star_axis,n) = d_vector(star_axis,n) + d_C
-       endif
-    enddo
-
[…]

### 2015-08-31 source/Problem_F.H

@@ -52,53 +52,13 @@ BL_FORT_PROC_DECL(GW_STRAIN_TENSOR,gw_strain_tensor)
     Real* h_plus_motion, Real* h_cross_motion,
     Real* Qtt, const Real* time);
 
-BL_FORT_PROC_DECL(SETUP_SCF_RELAXATION,setup_scf_relaxation)
-   (const Real dx[], const Real problo[], const Real probhi[]);
-
-BL_FORT_PROC_DECL(GET_COEFF_INFO,get_coeff_info)
-(int* loc_A, int* loc_B, int* loc_C,
-    Real* c_A, Real* c_B, Real* c_C);
-
-BL_FORT_PROC_DECL(GET_OMEGASQ,get_omegasq)
-   (const int lo[], const int hi[], const int domlo[], const int domhi[],
-    BL_FORT_FAB_ARG(state),
-    BL_FORT_FAB_ARG(phi),
-    const Real dx[], const Real problo[], const Real probhi[], Real* omegasq);
-
 BL_FORT_PROC_DECL(SET_PERIOD,set_period)
    (const Real* period);
 
-BL_FORT_PROC_DECL(GET_BERNOULLI_CONST,get_bernoulli_const)
-   (const int lo[], const int hi[], const int domlo[], const int domhi[],
-    BL_FORT_FAB_ARG(state),
-    BL_FORT_FAB_ARG(phi),
-    const Real dx[], const Real problo[], const Real probhi[], 
-    Real* bernoulli_1, Real* bernoulli_2);
-
-BL_FORT_PROC_DECL(CONSTRUCT_ENTHALPY,construct_enthalpy)
-   (const int lo[], const int hi[], const int domlo[], const int domhi[],
-    BL_FORT_FAB_ARG(state),
-    BL_FORT_FAB_ARG(phi),
-    BL_FORT_FAB_ARG(enthalpy),
-    const Real dx[], const Real problo[], const Real probhi[], 
-    Real* bernoulli_1, Real* bernoulli_2, Real* h_max_1, Real* h_max_2);
-
-BL_FORT_PROC_DECL(UPDATE_DENSITY,update_density)
-   (const int lo[], const int hi[], const int domlo[], const int domhi[],
-    BL_FORT_FAB_ARG(state),
-    BL_FORT_FAB_ARG(phi),
-    BL_FORT_FAB_ARG(enthalpy),
-    const Real dx[], const Real problo[], const Real probhi[], 
[…]

### 2015-08-31 source/scf_relax.cpp

@@ -1,0 +1,276 @@
+#include "Castro.H"
+#include "Castro_F.H"
+
+#include "Gravity.H"
+
+#ifdef GRAVITY
+#ifdef ROTATION
+void Castro::scf_relaxation() {
+
+  int finest_level = parent->finestLevel();
+
+  MultiFab& S_new = get_new_data(State_Type);
+
+  int j = 1;
+  int relax_max_iterations = 30;
+
+  const Real* dx  = parent->Geom(level).CellSize();
+  
+  const Real* problo = parent->Geom(level).ProbLo();
+  const Real* probhi = parent->Geom(level).ProbHi();
+
+  const int*  domlo = geom.Domain().loVect();
+  const int*  domhi = geom.Domain().hiVect();
+
+  MultiFab gcoeff(grids,1,0,Fab_allocate);
+  gcoeff.setVal(0.0);
+
+  Box box_A( IntVect(), IntVect(2, 2, 2) );
+  Box box_B( IntVect(), IntVect(2, 2, 2) );
+  Box box_C( IntVect(), IntVect(2, 2, 2) );
+
+  FArrayBox cA(box_A);
+  FArrayBox cB(box_B);
+  FArrayBox cC(box_C);
+
+  cA.setVal(0.0);
+  cB.setVal(0.0);
+  cC.setVal(0.0);
+
+  int loc_A[3] = {0};
+  int loc_B[3] = {0};
+  int loc_C[3] = {0};
+
+  BL_FORT_PROC_CALL(SCF_SETUP_RELAXATION,scf_setup_relaxation)
+    (dx, problo, probhi);
+
+  BL_FORT_PROC_CALL(SCF_GET_COEFF_INFO,scf_get_coeff_info)
+    (loc_A, loc_B, loc_C, cA.dataPtr(), cB.dataPtr(), cC.dataPtr());
+
+  // Get the phi MultiFab.
+
+  MultiFab& phi = get_new_data(PhiGrav_Type);
+
+  int ns          = NUM_STATE;
+  Real cur_time   = state[State_Type].curTime();
+
+  // Iterate until the system is relaxed by filling the level data 
+  // and then doing a multilevel gravity solve.
+
+  int is_relaxed = 0;
+
+  while ( j <= relax_max_iterations ) {
+
+     // First step is to find the rotational frequency.
+
+     Real omegasq = 0.0;
+
+#ifdef _OPENMP
+#pragma omp parallel reduction(+:omegasq)
[…]

### 2015-08-31 source/scf_relax_F.H

@@ -1,0 +1,45 @@
+
+BL_FORT_PROC_DECL(SCF_SETUP_RELAXATION,scf_setup_relaxation)
+   (const Real dx[], const Real problo[], const Real probhi[]);
+
+BL_FORT_PROC_DECL(SCF_GET_COEFF_INFO,scf_get_coeff_info)
+(int* loc_A, int* loc_B, int* loc_C,
+    Real* c_A, Real* c_B, Real* c_C);
+
+BL_FORT_PROC_DECL(SCF_GET_OMEGASQ,scf_get_omegasq)
+   (const int lo[], const int hi[], const int domlo[], const int domhi[],
+    BL_FORT_FAB_ARG(state),
+    BL_FORT_FAB_ARG(phi),
+    const Real dx[], const Real problo[], const Real probhi[], Real* omegasq);
+
+BL_FORT_PROC_DECL(SCF_GET_BERNOULLI_CONST,scf_get_bernoulli_const)
+   (const int lo[], const int hi[], const int domlo[], const int domhi[],
+    BL_FORT_FAB_ARG(state),
+    BL_FORT_FAB_ARG(phi),
+    const Real dx[], const Real problo[], const Real probhi[], 
+    Real* bernoulli_1, Real* bernoulli_2);
+
+BL_FORT_PROC_DECL(SCF_CONSTRUCT_ENTHALPY,scf_construct_enthalpy)
+   (const int lo[], const int hi[], const int domlo[], const int domhi[],
+    BL_FORT_FAB_ARG(state),
+    BL_FORT_FAB_ARG(phi),
+    BL_FORT_FAB_ARG(enthalpy),
+    const Real dx[], const Real problo[], const Real probhi[], 
+    Real* bernoulli_1, Real* bernoulli_2, Real* h_max_1, Real* h_max_2);
+
+BL_FORT_PROC_DECL(SCF_UPDATE_DENSITY,scf_update_density)
+   (const int lo[], const int hi[], const int domlo[], const int domhi[],
+    BL_FORT_FAB_ARG(state),
+    BL_FORT_FAB_ARG(phi),
+    BL_FORT_FAB_ARG(enthalpy),
+    const Real dx[], const Real problo[], const Real probhi[], 
+    Real* h_max_1, Real* h_max_2,
+    Real* kin_eng, Real* pot_eng, Real* int_eng,
+    Real* left_mass, Real* right_mass,
[…]

### 2015-08-31 source/scf_relaxation.f90

@@ -1,0 +1,522 @@
+subroutine scf_setup_relaxation(dx, problo, probhi)
+
+    use bl_constants_module, only: HALF, ONE, TWO
+    use prob_params_module, only: center
+    use eos_module
+    use probdata_module
+
+    implicit none
+
+    double precision :: dx(3), problo(3), probhi(3)
+
+    double precision :: x, y, z
+    integer          :: n
+    integer          :: ncell(3)
+
+    double precision :: pos_l(3)
+    type (eos_t) :: eos_state, ambient_state
+
+    call get_ambient(ambient_state)
+
+    ! Following Swesty, Wang and Calder (2000), and Motl et al. (2002), 
+    ! we need to fix three points to uniquely determine an equilibrium 
+    ! configuration for two unequal mass stars. We can do this by 
+    ! specifying d_A, the distance from the center of mass to the 
+    ! inner point of the primary; d_B, the distance from the center 
+    ! of mass to the inner point of the secondary; and d_C, the 
+    ! distance from the center of mass to the outer point of the secondary. 
+    ! We give these widths in physical units, so in general these will 
+    ! be locations on the grid that are not coincident with a corner. 
+    ! If a location is at point (x, y, z), then we find the eight 
+    ! zone centers that surround this point, and do a tri-linear
+    ! reconstruction to estimate the relative weight of each of the eight 
+    ! zone centers in determining the value at that point.
+
+    ncell = NINT( (probhi - problo) / dx )
+
+    do n = 1, 3
+       d_vector(:,n) = center
+       if (n .eq. 1) then
+          d_vector(star_axis,n) = d_vector(star_axis,n) - d_A
+       else if (n .eq. 2) then
+          d_vector(star_axis,n) = d_vector(star_axis,n) + d_B
+       else if (n .eq. 3) then
+          d_vector(star_axis,n) = d_vector(star_axis,n) + d_C
+       endif
+    enddo
+
+    ! Locate the zone centers that bracket each point at the 
[…]

### 2015-08-31 papers/wdmerger_I/paper.tex

@@
 Type Ia supernovae, as well as their distribution in time and space.
-However, to date there have been few fully self-consistent simulations 
+To date there have been few fully self-consistent simulations 
 of binary white dwarf systems using mesh-based hydrodynamics, 
 relative to other methods. This is the first paper in a series designed to 
-describe simulations of these systems using a hydrodynamics code with 
+describe simulations of these systems using a mesh-based hydrodynamics code with 
 the adaptive-mesh refinement approach. In this paper we describe our numerical 
@@
 not do an adequate job conserving the total energy of the system, 
-but recent advances in the literature have made advances on this 
-problem and we discuss our implementation here. We also run an 
-extensive set of test problems to determine whether our software can sufficiently
+but recent advances in the literature allow progress on this 
+problem and we discuss our implementation here. We present an 
+extensive set of test problems demonstrating the extent to which our software can sufficiently
 model a system where large amounts of mass are advected on the computational 
 domain over very long timescales. Future papers in this series will describe
-our treatment of the initial conditions of these systems, and will 
+our treatment of the initial conditions of these systems and will 
 examine the early phases of the merger to determine its viability
@@
 
-Type Ia supernovae (SNe Ia) are currently some of the most exciting
+Type Ia supernovae (SNe Ia) are among the most exciting
 events to study in astrophysics. These bright, brief pulses of light
@@
 in recent years, including the discovery of the accelerated expansion
-of the universe \citep{perlmutter1999,riess1998}. However, their
-origin is shrouded in mystery. It has long been expected that these
+of the universe \citep{perlmutter1999,riess1998}. Their origin, though,
+is shrouded in mystery. It has long been expected that these
 events arise from the thermonuclear explosions of white dwarfs
@@
 mass, and in doing so the temperature and density in the center would
-become sufficient for thermonuclear fusion to proceed. However, in
-recent years many alternative progenitor models have been discussed. A
+become sufficient for thermonuclear fusion to proceed. In
+recent years the focus has shifted to a number of alternative progenitor models. A
 leading candidate for explaining some or most of these explosions is
@@
 accretion-induced collapse, would be avoided only if the accretion
-rate were well below the Eddington rate. \cite{tutukov_yungelson:1979}
-observed that this could happen if the mass loss from the secondary
+rate were well below the Eddington rate (see, e.g., \cite{fryer:1999}
+for a discussion of the possible implications of the accretion-induced 
+collapse scenario). \cite{tutukov_yungelson:1979}
+observed that the collapse could be avoided if the mass loss from the secondary
 was higher than the Eddington rate and thus the accreted material
@@
 the study of looking at the long-term dynamical evolution of binary
-white dwarf systems after constructing exact equilibrium initial
+white dwarf systems after constructing equilibrium initial
 conditions. Earlier work placed the stars too close together 
@@
 secondary, leading to the merger
-happening artificially too quickly. When the initial conditions are
+happening artificially too quickly \citep{fryer:2008}. When the initial conditions are
 constructed in exact equilibrium, the system can be stable for tens of
 orbital periods, substantially changing the character of the mass
-transfer phase. However, one limitation of this series of studies is
+transfer phase. One limitation of this series of studies is
 that the authors used a polytropic equation of state and thus could
@@
 
-Given the above, why is another approach using a different simulation
+Given the considerable research into the double degenerate problem 
+described above, why is another approach using a different simulation
 code warranted? First and foremost, reproducibility of the results
@@
 them in the study of these systems, such as excellent conservation of
-angular momentum. However, the question of whether a prompt detonation
-in a merger actually happens depends in detail on the nature of the
+angular momentum. A drawback relates to the fact that whether a prompt detonation
+in a merger happens depends in detail on the nature of the
 gas at the interface between the two stars, which is at much lower
@@
 resulted in a thermonuclear detonation. Reproducibility of the detonation 
-through direct numerical simulation is critical for building 
-confidence in this progenitor model. \MarginPar{Direct numerical
-  simulation has a specific meaning to CFD community.  That is the
-  simulation resolve the viscous scale.  So maybe we should remove the
-word 'direct'.}
-
-This is the first in a series of papers designed to address these
+through numerical simulation is critical for building 
+confidence in this progenitor model.
+
+This paper is the first in a series designed to address these
 outstanding theoretical issues for white dwarf mergers. This work will
@@
 binary white dwarf simulation. \autoref{sec:Tests} discusses a few
-test problems that we use to demonstrate that our code accurately
+test problems that we use to verify that our code accurately
 solves the equations of fluid dynamics. \autoref{sec:Performance}
@@
 it preserves physical extrema rather than clipping them off as in the
-original approach of \cite{ppm}. However, we found these limiters to
-be unsatisfactory for our problem. There are many regions in our
+original approach of \cite{ppm}. Despite the advantages of this limiter 
+we have found it to be unsatisfactory for our problem. There are many regions in our
 problem with large density gradients (such as the interface between
@@
 temperature and density, and calling the EOS is simplest in this
-form. However, in hydrodynamics we often have the density, and
+form. In hydrodynamics we often have the density and
 internal energy as independent variables, and we want to obtain the
@@
 \end{equation}
-and for the energy source term the analogue it is
+and for the energy source term it is
 \begin{equation}
@@
 and then we will describe our approach for the energy. This discussion will be somewhat detailed.
-However, we believe that the attention is necessary because of the importance of accuracy
+We believe that the attention is necessary because of the importance of accuracy
 in the gravitational source terms for our problem. The stability of the white dwarf binary
@@
 This allows momentum to leak out of the domain even if the gravitational source term 
-is written in an explicitly conservative manner. To see this, one can write the 
+is written in an explicitly conservative manner. To see this, one can use the Poisson equation to write the 
 density in terms of the potential and then consider its spatial discretization. For simplicity,
-we will consider one spatial dimension.
+we will consider one spatial dimension and a uniform discretization. Analogous results 
+may be readily obtained for the non-uniform case.
 \begin{align}
-  -\rho_{i}  \frac{d\Phi_{i}}{dx} &= -\frac{1}{4\pi G} \frac{d^2\Phi_i}{dx^2} \frac{d \Phi}{dx} \notag \\
+  -\rho_{i}  \frac{d\Phi_{i}}{dx} &= -\frac{1}{4\pi G} \frac{d^2\Phi_i}{dx^2} \frac{d \Phi_i}{dx} \notag \\
   &= -\frac{1}{4\pi G} \left[\frac{\Phi_{i-1} - 2 \Phi_{i} + \Phi_{i+1}}{\Delta x^2}\right] \left[ \frac{\Phi_{i+1} - \Phi_{i-1}}{2\Delta x} \right] \notag \\
@@
 to the left and right will result in complete cancellation of the source terms.
-However, if the potential if non-zero outside of the domain, then there will be
+The catch is that if the potential if non-zero outside of the domain, then there will be
 momentum lost or gained from the grid, which will be encapsulated in the ghost cells
@@
 three-dimensional stencil including the $y$ and $z$ derivatives, depending on the
-discretization these will not be cancelled at all. This latter problem can be resolved by
+discretization these may not be cancelled at all. This latter problem can be resolved by
 writing the momentum update in an explicitly conservative way.
@@
 formulations are identical. Thus the latter has been advocated by, for example, 
-\cite{jiang:2013} for the ATHENA code. However, in a finite discretization 
+\cite{jiang:2013} for the ATHENA code. A significant limitation to this approach is that in a finite discretization 
 the divergence of the gravitational acceleration is no longer guaranteed to equal
@@
 change in gas energy, and that energy changes due to mass transfer are explicitly and 
-exactly tracked. At each zone face in a given timestep, there is some transfer of mass $\Delta \rho$ between 
+exactly tracked. At each zone face in a given timestep, there is some transfer of mass $\Delta \rho \Delta V$ between 
 the zones adjacent to that face, which is governed by the continuity equation. Suppose the 
 transfer of mass is positive to the right, and the face is located at the (one-dimensional)
-index $i-1/2$. Energy is transferred because an amount of mass $\Delta \rho_{i-1/2}$ moves from 
+index $i-1/2$. Energy is transferred because an amount of mass $\Delta \rho_{i-1/2} \Delta V$ moves from 
 potential $\Phi_{i-1}$ to potential $\Phi_{i}$; if these potentials are different, it requires some 
@@
 the zone centers immediately to the left and right in the direction perpendicular to the face.
-Note that while total energy is conserved, the gas energy is not: this does not have the 
+Note that while total energy is conserved, the gas energy is not because \autoref{eq:grav_energy_conservation_update} does not have the 
 telescoping property of a flux formalism where energy gained by one zone is lost by another. 
@@
 
-The change in the mass $\Delta \rho_{f}$ is proportional to the mass flux through the surface with area $\Delta A$:
+The change in the density $\Delta \rho_{f}$ is proportional to the mass flux through the surface with area $\Delta A$:
 \begin{equation}
@@
 re-computes the potential at the new time after the hydrodynamics step, and we can apply this energy 
-at the end of the timestep).
+at the end of the timestep). Using this conservative energy formulation is very important for simulations where the 
+potential can change substantially over the course of the evolution, as massive amounts of energy can be lost. 
+The standard approach is therefore unacceptable in the case of a problem like white dwarf mergers, and 
+the fix to this energy leaking---evaluating the energy transfer at the six zone faces instead of 
+the single zone center---has a very minor cost in terms of code complexity and computational time.
 
@@
 gas energy using the work done $(\mathbf{v} \cdot \rho \mathbf{g})$ by the gravitational force at the 
-cell center. Analytically this form expresses the same core idea via the work-energy theorem, and its 
-major flaw is that it evaluates the energy change at the cell center when in fact the mass transfer is 
-happening at the cell edges, and the energy change is actually partially compensated by the potential 
-in the adjacent zone that the mass is moving to. This is important enough that for simulations where the 
-potential can change substantially over the course of the evolution, massive amounts of energy can be lost. 
-This standard approach is therefore unacceptable in the case of a problem like white dwarf mergers, and 
-the fix to this energy leaking---evaluating the energy transfer at the six zone faces instead of 
-the single zone center---has a very minor cost in terms of code complexity and computational time.
-
-It is worth pointing out that while \autoref{eq:grav_energy_conservation_update} was correctly obtained 
-by \cite{arepo} (whose work was the primary inspiration for this approach), there is a subtle error in that derivation. 
-Their argument was that if we integrate the work done over the volume of the zone, this volume integral can be 
-converted into a surface integral by use of the divergence theorem (see their Equation 93). Then we can simply 
-evaluate the work done at each zone face. However, this application of the divergence theorem is 
-mathematically incorrect, despite the fact that this is similar in spirit to the approach discussed here. 
-Moreover, using the divergence theorem to integrate the work done in a zone can never obtain 
-the factor of $1/2$ corresponding to the change in gravitational self-energy. That factor can be thought of
-as a double-counting factor that ensures that we correctly capture the energy of a pairwise interaction. 
-When one integrates the work done in the zone, there is a contribution to the energy that cannot be 
-expressed as a full time derivative or the divergence of a flux. This reflects the 
-global nature of the gravitational potential and means that we cannot get the correct energy expression 
-unless we integrate over the entire mass distribution; see \citet[Chapter 4]{shu:1992} and \citet{marcello:2012}. 
-However, this approach is still valid when considering energy transfer between the gas energy and a  
-potential field, if the change in energy is explicitly local. In the case of gravity, the mass flux
-that generates a change in energy is local, but the change in energy is stored in the entire gravitational 
-field, and is only captured by a global integral because of the long-range nature of the potential.
-
-Another approach recently taken in the literature for conserving total energy has been to evolve an 
-equation for the total energy $(\rho E_{\text{tot}})$; see \cite{marcello:2012} and \cite{jiang:2013}.
-This can be done by adding a term corresponding to the ``flux'' of gravitational potential energy. We 
-avoid this approach in general because there will be regions on the computational domain where the total 
+cell center. Analytically this form expresses the same core idea as \autoref{eq:grav_energy_conservation_update} 
+via the work-energy theorem, but a major flaw is that it evaluates the energy change at the cell center 
+when in fact the mass transfer is happening at the cell edges. However, there is a more fundamental issue. 
+As observed by \citet[Chapter 4]{shu:1992} and \citet{marcello:2012}, the energy source term $(-\rho \mathbf{u}\cdot\nabla\Phi)$ 
+includes a contribution that is not localizable and reflects the global nature of the gravitational potential. 
+It is only by integrating this source term over all space that the correct gravitational potential energy 
+for the system emerges. Integrating over only one zone cannot yield the correct description of this global 
+energy flow, and indeed therefore can never obtain the factor of 1/2 that arises from avoiding double-counting 
+of the pairwise gravitational interactions. So while \autoref{eq:grav_energy_conservation_update} was correctly obtained 
+by \cite{arepo} (whose work was the primary inspiration for this approach), their derivation that integrates the work 
+done over the volume of a zone does not correctly account\footnote{They argued that this volume integral 
+of the work can be converted into a surface integral by use of the divergence theorem (see their Equation 93). That 
+application of the divergence theorem is mathematically incorrect, because their formulation of the integrand cannot actually 
+be expressed only as the divergence of a vector field.} for this behavior. It is only by accounting for the global nature of the
+potential that we can obtain the correct factor of $1/2$ and have a correct accounting of the energy within a zone.
+The insight of applying the source term on the zone faces to conserve energy is still valid when 
+considering energy transfer between the gas energy and a potential field if the change in energy 
+is explicitly localizable. Indeed, we employ this formulation in \autoref{sec:rotation} for the case of the rotational potential.
+
+Another approach to conserving total energy recently taken in the literature is to evolve an 
+equation for the total energy $(\rho E_{\text{tot}})$; see \cite{marcello:2012} and \cite{jiang:2013} (see also 
+\cite{arepo}, Section 5.3). That is, one can replace the gas energy equation with a total energy equation, 
+and then the energy flux includes a term corresponding to the flux of gravitational potential energy. We 
+avoid this approach for our problem because there will be regions on the computational domain where the total 
 energy is dominated by potential energy (especially the low-density regions near the edge of the white dwarfs),
@@
 \cite{arepo}, we find that this can result in some serious errors due to numerical discretization, yielding 
-unphysical energies or temperatures. We observe also that the specific implementation of \cite{jiang:2013} 
-has terms in their gravitational flux that are not proportional to $\rho$, and so it can lead to the 
+unphysical energies or temperatures. We observe also that the implementation of \cite{jiang:2013} 
+has terms in the gravitational flux that are not proportional to $\rho$, and so can lead to the 
 same troubles that plague the tensor-based formalism for the momentum equation, where small errors 
@@
 potential vanishes at large distances from the region of non-zero
-density. However, on a finite computational domain it is usually not
+density. On a finite computational domain, however, it is usually not
 possible to have the edges of the domain be far enough away that the
@@
 therefore requires knowledge of the values of the potential on the
-edges of the computational domain. In principle, they can be computed
+edges of the computational domain. In principle, the boundary values can be computed
 by doing a direct sum over the mass distribution inside the domain,
@@
 the domain and treating it as a point source at the domain
-center. This is exactly correct only for a spherically symmetric mass
+center. This is correct only for a spherically symmetric mass
 distribution, and therefore is best suited for problems like
-single-star Type Ia supernova simulations (e.g. \cite{malone:2014})
-that employ self-gravity. However, for a problem like that of a binary star system
+single-star Type Ia supernova simulations (e.g.$\ $\cite{malone:2014})
+that employ self-gravity. For a problem like that of a binary star system
 with significant departures from spherical symmetry, this assumption
-fails to produce accurate boundary values. This results in a
-significant drift of the center of the mass of the system over time,
-which will be discussed in \autoref{sec:kepler}.
+fails to produce accurate boundary values, which we find in \autoref{sec:kepler}
+results in a significant drift of the center of the mass of the system over time.
 
@@
 As a demonstration of the method's accuracy, we consider the case of two 
-white dwarfs of mass ratio 2/3, loaded onto the grid by the method described 
+white dwarfs of mass ratio 2/3, loaded onto the grid by the method described below 
 in \autoref{sec:implementation}. We terminated the simulation just after
@@
 choice for all simulations with Poisson gravity. In \autoref{sec:kepler} we 
-see that there are no gains to be had by increasing the accuracy further. 
+show that there are no gains to be had by increasing the accuracy further. At 
+very high orders the approximation breaks down, as seen in \autoref{fig:bc_comparison}. 
+This is a result of the ambient material on the grid. At each boundary point we 
+assume that all of the mass on the grid is contained within a sphere whose radius is 
+the distance from that boundary point to the center of the domain. This does not hold 
+for boundary points in the centers of domain faces, because of the material in the 
+domain corners. This can be fixed by using multiple mass shells at diferent radii, but the error 
+is negligible in practice for the values of $l_{\text{max}}$ that we use.
 \begin{figure}[h]
@@
 multiple of the grid spacing, and the center of the sphere is at the
-origin. The problem domain for our simulations is $[-1.6 \text{cm}, 1.6 \text{cm}]^3$, and
+origin. The problem domain for our simulations is $[-1.6\ \text{cm}, 1.6\ \text{cm}]^3$, and
 we take $R = 1.0\ \text{cm}$ and $\rho = 10^3\ \text{g cm}^{-3}$. 
@@
 
-We find that at low resolution 
-convergence is actually substantially better than second-order. The 
-explanation for this is that we are attempting to model a spherical 
-object on a rectangular grid. At very low resolution, the object does 
-not look very spherical. As the resolution is increased, the total 
-amount of mass on the grid, as well as its location, will change 
-as the sphere fills out. This means we are combining the true accuracy 
-bonus from increased resolution with the artificial accuracy bonus 
-from getting closer to solving the problem we are supposed to be solving. 
-On the other hand, at the highest resolution shown the error slightly increases.
-
-We can solve one of these two sources of error by changing the density on  
+We find that at low resolution convergence is actually substantially better 
+than second-order. The explanation for this is that we are attempting to 
+model a spherical object on a rectangular grid. This results in two sources of error.
+First, at very low resolution, the object does not look very spherical due to the rectangular 
+grid representation, so the potential it produces is not quite that of a sphere. 
+As the resolution is increased, the distribution of the mass on the grid will change.
+Second, the total amount of mass on the grid will change as the sphere fills out. 
+So we are combining the true accuracy bonus from increased resolution 
+with the artificial accuracy bonus from getting closer to solving the problem 
+we are supposed to be solving. At high resolution this effect levels off, though, 
+as the representation of the sphere is not significantly different in 
+our two highest resolutions shown. For example, at $128$ zones per dimension 
+the amount of mass on the grid happens to be slightly closer to the true spherical 
+mass than at $256$ zones per dimension.
+
+We can solve the second source of error by changing the density on  
 the grid so that the total mass $M$ is actually what we intend it to be.
 The resolution study for this case (the ``normalized sphere'') is also
-plotted in \autoref{fig:gravity_convergence}. We still obtain
+plotted in \autoref{fig:gravity_convergence}. At low resolution We still obtain
 convergence slightly better than second-order, indicating that we 
@@
 The only way to fully eliminate this effect is to use a test problem that
-does not change with resolution. The obvious companion is a cube of
+does not change with resolution. The obvious companion problem is a cube of
 uniform density $\rho$, where now $R$ is half of the side length of
@@
 $y_1 = R - y$, $z_0 = R + z$, $z_1 = R - z$, 
-and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
-\texttt{Fortran} and \texttt{C}, the inverse hyperbolic tangent used here is
+and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. We note that if implemented in 
+\texttt{Fortran} or \texttt{C}, the inverse hyperbolic tangent used here is
 \texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not}
@@
 The extent to which angular momentum conservation is violated
-will be a function of the resolution. When this is sufficiently high, 
+will be a function of the resolution. When the resolution is sufficiently high, 
 excellent conservation properties can result, as demonstrated in \autoref{sec:kepler}. 
@@
 We evaluate all quantities at cell centers. We use the same predictor-corrector 
-approach that we use for the gravity source terms to the momentum equations. However, a slight 
+approach that we use for the gravity source terms to the momentum equations. A slight 
 difference is that the Coriolis force for each velocity component is coupled to other velocity 
@@
 then the Coriolis force will see a different velocity than the one obtained through the 
-pure hydrodynamics step. (This does not matter for the energy equation in our new formulation,
+pure hydrodynamics step. (The energy equation does not face the same issue in our new formulation,
 because the velocities used are always the time-level $n+1/2$ values coming from the Riemann solver.)
@@
 binary systems, and the centrifugal force does not depend on the fluid velocity.
-However, this issue may be worth exploring in future work in situations where the Coriolis 
+This issue may be worth exploring in future work in situations where the Coriolis 
 term is non-negligible in determining the system evolution.
@@
 large density and pressure gradients near the white dwarf surfaces
-that result in large amounts of mass flowing into the ambient
-medium. This can result in spurious non-physical consequences such as 
+that result in significant amounts of mass flowing out of the white dwarfs.
+This can result in spurious non-physical consequences such as 
 the total density or energy going negative in a zone. To compensate 
@@
 its maximum allowed by the velocities on the grid over a span of approximately 
-1000 timesteps. This allows the ambient medium just outside the white dwarf
-to come closer to equilibrium with the surface without having 
+1000 timesteps. This allows the gas at the surface of the white dwarf
+to come closer to equilibrium without having 
 discontinuous jumps in the density or energy. For all simulations, 
@@
 
-We employ a ``sponge'' similar to that described by \citet{maestro3}
-for the outer regions of the computational domain. After the
-hydrodynamics update, we apply a damping force to the momentum
+Outside of the stars we fill the rest of the domain with a very low density 
+ambient gas because our hydrodynamics representation requires the density to be 
+non-zero everywhere. This ambient material can create difficulties for the simulation.
+In addition to the negative densities or energies at the stellar surfaces mentioned earlier, 
+in the rotating reference frame we observe that standing instabilities can create very 
+large velocities in the ambient fluid that drag down the global timestep by 
+up to an order of magnitude.  To deal with this we employ a ``sponge'' similar 
+to that described by \citet{maestro3} for the outer regions of the computational domain. 
+After the hydrodynamics update, we apply a damping force to the momentum
 equation as follows:
@@
 that the sponge extends another 10\% of that distance. The resulting profile 
-is displayed in Figure \autoref{fig:sponge}. We set $\Delta
+is displayed in \autoref{fig:sponge}. We set $\Delta
 t_S = 0.01$ s, which is of the same order as the CFL timestep
-for typical problem setups. We find that this sponge is necessary
-especially in the rotating reference frame, as standing instabilities
-can create very large velocities in the ambient fluid that drag down
-the global timestep by up to an order of magnitude.  This means, however, 
-that while the sponge is applied we should be cautious of imputing any physical 
+for typical problem setups. While the sponge is applied we should be cautious of imputing any physical 
 meaning to what is happening in the low-density gas far from the stars.
@@
 locations is as follows: at the beginning of the calculation, we store the 
-physical center of mass $\mathbf{x}_{\text{COM}}$ of the stars as determined 
-by Kepler's third law. We also store the velocity $\mathbf{v}_{\text{COM}}$ 
+physical center of mass $\mathbf{x}_{c}$ of the stars as determined 
+by Kepler's third law. We also store the velocity $\mathbf{v}_{c}$ 
 of the stars. Then, at each new time step we make a preliminary guess for their 
 location by updating the location using the old velocity, 
-$\mathbf{x}_{\text{COM}} \rightarrow \mathbf{x}_{\text{COM}} + \mathbf{v}_{\text{COM}} \Delta t$.
+$\mathbf{x}_{c} \rightarrow \mathbf{x}_{c} + \mathbf{v}_{c} \Delta t$.
 We then refine our guess for the location and velocity of each star by computing a
@@
 
+When we do simulations with adaptive-mesh refinement, there are multiple levels of refinement 
+that contribute to a global integral. To deal with this we employ a ``mask'' which zeros out 
+the data in a zone on a given level if there is a refined region overlying that zone.
+
 \subsubsection{Gravitational Waves}
@@
 $GM / c^2 R$, which measures the ratio of the Schwarzschild radius of a mass $M$ to the actual 
-radius $R$ of the object, is much less than unity for a white dwarf. It is justified to 
-neglect any relativistic effects in computing the dynamical evolution of the system. However,
-the system emits gravitational waves during its evolution; this energy loss is what drives 
-the initial inspiral over very long timescales, and occurs at a greater rate during 
-the coalescence phase, though then being inconsequential on a dynamical timescale. The 
-frequency of the gravitational waves will be similar in size to the frequency of the 
-orbital motion, which is in the range 10-100 mHz for our problem. This is well outside 
-the range of currently existing gravitational wave detectors but is very well suited 
-for proposed space-based detectors such as eLISA.
-
-We follow the prescription of \citet{blanchet:1990}. At distances far from the 
+radius $R$ of the object, is much less than unity for a white dwarf. Thus at any given time the 
+relativistic effects are negligible compared to the Newtonian gravity and so we do not 
+directly include relativistic effects in computing the dynamical evolution of the system.
+A white dwarf binary system does emit gravitational waves during its evolution; this energy loss 
+is what drives the initial inspiral over very long timescales. Eventually it will drive the system 
+to become dynamically unstable due to the Newtonian tidal forces alone, though once that period begins, 
+the gravitational energy loss is inconsequential in affecting the dynamical evolution of the system. 
+The frequency of the gravitational waves emitted by the white dwarf binary 
+will be similar in size to the frequency of the orbital motion, which is in the range 
+10-100 mHz for our problem. This is well outside the range of currently existing 
+gravitational wave detectors but is very well suited for proposed space-based detectors such as eLISA.
+
+We follow the prescription of \citet{blanchet:1990} for computing a gravitational wave 
+signal for our simulation. At distances far from the 
 gravitational wave source, we can consider the leading term in the gravitational 
@@
 The argument $(t - r/c)$ indicates that to get the strain at time $t$ we evaluate the second derivative of the 
-quadrupole moment at the retarded time $t - r/c$. In practice we ignore this in calculating this strain,
-instead using the current simulation time, and so this merely yields a time delay of order $r/c$ between 
-when the waves are emitted and when we observe them.
+quadrupole moment at the retarded time $t - r/c$. In practice the retarded time is simply the simulation time
+and the observer would see the gravitational waves after a time delay of order $r/c$.
 
@@
 The strategy is then as follows. At the end of the coarse timestep, we first calculate $\ddot{Q}_{kl}$
-using an integral over the domain (using refined grids as usual by masking out the contributions from 
-coarse grids underlying more accurate fine grids). This quantity is independent of the observer. If we 
+using an integral over the domain. This quantity is independent of the observer. If we 
 are using a rotating reference frame, we first convert velocities and positions back to the inertial 
@@
 experiment such as LISA could detect a coalescing binary white dwarf system. 
-However, the strain at any other distance is easily calculated and goes as the inverse of the distance.
+The strain at any other distance is easily calculated and goes as the inverse of the distance.
 
@@
 size that is different (and in this case larger) than the
-one-dimensional model. However, the magnitude of this change becomes
+one-dimensional model, though the magnitude of this change becomes
 smaller with resolution. The star is only approximately in equilibrium 
 by this measure when the coarse grid of $256^3$ zones has a level of 
-refinement that jumps by a factor of four. However, even then there is a slight uptick
+refinement that jumps by a factor of four. Even then there is a slight uptick
 in the size toward the end, implying that the numerical stability is
@@
 resolution than at the lower resolutions. This result suggests
-that when doing the equilibrium initial models that will former the
+that when constructing the equilibrium initial models that will form the
 basis of later calculations, we should carefully monitor the evolution
@@
 As long as the stars remain approximately spherical, the stars can be 
-treated as point masses (this only seriously breaks down after the stars
+treated as point masses (this approximation only seriously breaks down after the stars
 have come into contact). In dimensionless units where $r \to r / a$ and 
@@
 uniform grid are shown in \autoref{fig:freefall}. They show excellent agreement
-between the theory and the simulation results.
+between the analytical solution and the simulation results.
 
@@
 moving at a uniform bulk velocity will appear different than an
-equivalent stationary simulation. If true, we need to understand 
+equivalent stationary simulation (e.g. \cite{arepo}). If true, we need to understand 
 the importance of this effect when deciding whether to trust the 
 output of a code like \castro\ when applied for merger problems.
-Recently, this has come up in two ways which are of note for us 
+Recently, concern for the issue of Galilean invariance has come up in two ways which are of note for us 
 in the present study. We will explain these situations and display 
@@
 have important consequences for how much we can trust the ability of 
-\castro\ to test the violent merger progenitor model. Shearing between 
+\castro\ to test the violent merger progenitor model, where a detonation 
+arises in the low-density material at the stellar surface. Shearing between 
 the material flowing out of the secondary and material near the 
@@
 interaction of this numerical diffusion with small-scale instabilities
-(which may be physical or numerical) which couple with and
+(that may be physical or numerical) which couple with and
 fundamentally alter the large-scale modes. Small-scale instabilities
@@
 R10 point out that this problem does not
-converge with resolution and so it is not possible to know the correct
+converge with resolution (because the initial perturbation is too sharp 
+and seeds numerical noise at the grid resolution level) 
+and so it is not possible to know the correct
 behavior of this problem. As such, we do not know whether the
@@
 test proposed by S10: at low resolutions and high bulk velocity, the 
-Kelvin-Helmholtz instability completely fails to develop. However, 
+Kelvin-Helmholtz instability completely fails to develop. Furthermore 
 the problem does not converge even qualitatively at the highest 
@@
 At low resolutions and very large bulk velocities, the fluid 
-does get significantly disrupted by numerical error. However, this 
+does get significantly disrupted by numerical error. This 
 effect quickly converges away with resolution and qualitatively 
@@
 the test proposed by R10: as we get to higher flow speeds we need to have higher 
-spatial resolution to compensate for the increased numerical diffusion. However, 
-the accuracy is much lower for the highest bulk velocities. This is probably because 
+spatial resolution to compensate for the increased numerical diffusion. The 
+accuracy is much lower for the highest bulk velocities. This is probably because 
 the initial conditions feature a transition that is twice as sharp (compare the $\Delta y$ 
@@
 when non-linear effects have taken over and significantly disrupted
-the initial flow. However, at low resolution the tested grid algorithm
+the initial flow. At low resolution the tested grid algorithm
 had failed to disrupt both for $v = 0$ and $v = 10$. We too ran this
@@
 instability damps out at low resolution but goes strongly non-linear
-and disrupts the flow at high resolution.  However, we strongly
+and disrupts the flow at high resolution.  We strongly
 emphasize the point that this does not objectively demonstrate a
@@
 solution to compare to, and this is lacking for this problem at late
-times. As observed by \citeauthor{mcnally:2012}, this is because the
+times. As observed by \citeauthor{mcnally:2012}, this lack of a solution is because the
 secondary instabilities form for this problem when the whirls of the
@@
 
-To see this effect in action for a stellar simulation, we repeated the
+To analyze the effects of velocity-dependent results for a stellar simulation, we repeated the
 test of \autoref{sec:HSE} with a bulk velocity on the grid. We chose a
@@
 problem at the bulk velocity, so that the multipole moments were
-computed with respect to the stellar center.  \MarginPar{are you
-  sure that the single WD left quiet in the {\em corner} of the
-  bigger domain is as stable as it was in the earlier test?}
+computed with respect to the stellar center.
 
@@
 transfer is seeded in a controllable manner and free from numerical
-noise.  \MarginPar{I really think that we should be able to do better here---this is where I'd like to see ppm\_type = 2, or better hydro-gravity coupling (eventually)}
+noise.
 
@@
 There are a number of effects worth looking into here. For simplicity, we choose two 
-cases to demonstrate the simulation behavior: an equal mass case of two $0.9\msolar$ 
-white dwarfs, and an unequal mass case of $0.9\msolar$ and $0.6\msolar$ white dwarfs.
+cases to demonstrate the simulation behavior: an equal mass case of two $0.9\ \msolar$ 
+white dwarfs, and an unequal mass case of $0.9\ \msolar$ and $0.6\ \msolar$ white dwarfs.
 In each case, the initial orbital period is 100 seconds.
@@
 
-Another that we consider is the energy conservation of the
+Another diagnostic that we consider is the energy conservation of the
 system. Recalling \autoref{sec:gravity_hydro_coupling}, there are
@@
 the new momenta, we reset the total energy to be equal to the internal
-energy plus the kinetic energy. This has the virtue of ensuring that
+energy plus the kinetic energy. This approach has the virtue of ensuring that
 there is no conflict due to discretization between the momentum and
@@
 the same secular decline, and the system appears to be
-stable. However, the cost is that the oscillations about this
+stable. The cost is that the oscillations about this
 equilibrium are slightly larger than in the inertial frame. The
-unequal mass case displays this effect starkly: while the average
-distance is not declining, there are huge oscillations about this
+unequal mass case displays this effect starkly, to the point where the system 
+becomes unstable: while the average distance is not substantially declining, 
+there are huge oscillations about this
 average. In addition, the stars move substantially within the rotating
@@
 This is especially true because some aspects of our approach to parallelism 
-have changed since the first \castro\ paper, and we have obtained improved 
+have changed since the first \castro\ paper \citep{castro}, and we have obtained improved 
 performance in certain settings.
@@
 meaning that the average workload per processor should be no smaller than 90\% of the 
-maximum workload found on any processor. However, we find that in practice the 
+maximum workload found on any processor. We find that in practice the 
 performance is largely insensitive to this choice.
@@
 that can diminish the quality of the resulting solutions to the 
-fluid equations. However, these effects diminish with increasing 
+fluid equations, but that these effects diminish with increasing 
 resolution. We therefore make no explicit claims about the usefulness 
@@
 verification standpoint is to see whether the observed 
-behavior converges with resolution. However, it is not clear 
+behavior converges with resolution. It is not clear 
 whether at practical resolutions the results in the 
@@
 
-The authors thank Noel Scudder and Platon Karpov for their help with 
-this project, especially related to visualization of the results. 
-We also thank Adam Jacobs for his advice on and assistance with 
+The authors thank John Bell for helpful discussions on gravity 
+and hydrodynamics issues. We also thank Noel Scudder and Platon Karpov 
+for their help with this project, especially related to visualization of the results, 
+and Adam Jacobs for his advice on and assistance with 
 running on supercomputing resources. Special thanks are given 
@@
 is the expected behavior---we see a state that is traced only by the
-contact wave.  However, if $u < 0$ but $u + c \ge 0$, then we instead get:
+contact wave.  If $u < 0$ but $u + c \ge 0$, then we instead get:
 \begin{equation}

### 2015-09-04 papers/wdmerger_I/paper.tex

@@
 happening artificially too quickly \citep{fryer:2008}. When the initial conditions are
-constructed in exact equilibrium, the system can be stable for tens of
+constructed in equilibrium, the system can be stable for tens of
 orbital periods, substantially changing the character of the mass
@@
 
-We do not attempt to enforce equilibrium with an additional relaxation
+In the current paper we do not attempt to enforce equilibrium with an additional relaxation
 step. This will be an important part of future work in this series, as
@@
 basis of later calculations, we should carefully monitor the evolution
-of the stars before applying any artificial damping to cause the
+of the stars when applying any artificial damping to cause the
 merger, to ensure that the merger is due to this applied force and not
@@
 depends on the initial composition of the stars (for example, 
-by including surface layers of helium and/or by using 
+by studying the dependence of the results on the size of helium 
+surface layers on carbon-oxygen white dwarfs, or by using 
 white dwarf models generated by modern stellar evolution 
-codes); the gravitational wave signals generated by 
-white dwarf mergers; and, collisions of white dwarfs. All of 
+codes); and, collisions of white dwarfs. All of 
 these are possible under the framework we have established.

### 2015-09-06 papers/wdmerger_I/paper.tex

@@
 %==========================================================================
-\title{Double White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification}
+\title{White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification}
 
@@
 discretized energy equation that explicitly conserves total energy when 
-coupled to gravity. When gravity is included, the total energy is
-\begin{equation}
-  \rho E_{\text{tot}} = \rho E + \frac{1}{2}\rho\Phi = \frac{1}{2}\rho \mathbf{u}^2 + \rho e + \frac{1}{2}\rho\Phi, \label{eq:total_energy_gravity}
-\end{equation}
-where $\rho E$ is the total gas energy from the pure hydrodynamics equation,
-and the factor of $1/2$ in the gravitational energy term is necessary
-for simulations with self-gravity to prevent double-counting of
-interactions.
-
-Historically many simulation codes with gravity have not used a conservative formulation 
-of the energy equation, but it is straightforward to do so. In particular, conservation of 
-total energy requires only that a change in gravitational energy is compensated for by a 
-change in gas energy, and that energy changes due to mass transfer are explicitly and 
-exactly tracked. At each zone face in a given timestep, there is some transfer of mass $\Delta \rho \Delta V$ between 
-the zones adjacent to that face, which is governed by the continuity equation. Suppose the 
-transfer of mass is positive to the right, and the face is located at the (one-dimensional)
-index $i-1/2$. Energy is transferred because an amount of mass $\Delta \rho_{i-1/2} \Delta V$ moves from 
-potential $\Phi_{i-1}$ to potential $\Phi_{i}$; if these potentials are different, it requires some 
-work exerted by or on the zone to push the mass to this different potential energy. Therefore 
-the net change of gravitational energy is $\left[(1/2)\Delta \rho_{i-1/2} (\Phi_{i} - \Phi_{i-1})\right]$,
-and if total energy is conserved, then the zone must have a change in gas energy 
-$(\rho E)$ that is equal in magnitude and opposite in sign to the change in potential energy. 
-Thus the update in a zone for the total energy that conserves $(\rho E_{\text{tot}})$ is:
+coupled to gravity. When gravity is included, the conserved total energy
+over the entire domain is
+\begin{equation}
+  \int \rho E_{\text{tot}}\, dV = \int dV \left(\rho E + \frac{1}{2}\rho\Phi\right), \label{eq:total_energy_gravity}
+\end{equation}
+where $\rho E$ is the total gas energy from the pure hydrodynamics equation. 
+The factor of 1/2 in the gravitational energy is necessary for simulations with
+self-gravity to prevent double-counting of interactions (since in dynamical evolution
+the relevant gravitational potential energy is $\rho \Phi$ and the gravitational force
+is $\rho \mathbf{g}$). Historically many simulation codes with gravity have not used
+a conservative formulation of the energy equation, but it is straightforward to do so.
+Our approach, and the discussion that follows, is based on that of \cite{arepo}.
+
+Conservation of total energy requires that a change in gravitational energy is compensated
+for by a change in gas energy, and that energy changes due to mass transfer are explicitly and 
+exactly tracked. Suppose that we have some fluid mass $\Delta M_{i+1/2} = \Delta \rho_{i+1/2} \Delta V$ leave some zone
+with index $i$ and enter some zone with index $i+1$. The subscript indicates that the mass change is
+occurring at the interface between the two zones, at index $i+1/2$. The work done by the gravitational
+force on the gas is $\Delta (\rho E) = W = \int F dx = (\Delta M_{i+1/2}\ g_{i+1/2}) (\Delta x / 2)$,
+where $g_{i+1/2}$ is the gravitational acceleration at the interface. The second term in parentheses
+is just the distance from the zone center to the zone edge: once the mass leaves the zone edge, it no longer
+needs to be tracked. To second order, $g_{i+1/2} = -(\Phi_{i+1} - \Phi_{i}) / \Delta x$, and also to second order the potential
+at the interface is given by $\Phi_{i+1/2} = (\Phi_{i+1} + \Phi_i) / 2$, so we can equivalently view the work done
+as $W = -\Delta M_{i+1/2} (\Phi_{i+1/2} - \Phi_i)$. Physically, this is just the negative of the gravitational
+potential energy change as the fluid is pushed from the cell center potential to the cell edge potential,
+exactly as the work-energy theorem implies.
+
+Now, in a hydrodynamics code mass changes correspond to hydrodynamic fluxes. In particular,
+the continuity equation tells us that the mass flux $F_\rho = \rho^{n+1/2}_{i+1/2} v^{n+1/2}_{i+1/2}$ yields
+an integrated mass motion through the interface $i+1/2$ over a timestep $\Delta t$ of:
+\begin{equation}
+  \Delta \rho_{i+1/2} = \frac{\Delta t}{\Delta V} \left(\rho^{n+1/2}_{i+1/2} v^{n+1/2}_{i+1/2} dA\right).
+\end{equation}
+Note that here $v_{i+1/2}$ is the component of the velocity perpendicular to the zone face, whose
+area is $dA$.
+
+Finally, then, we write the update in a zone for the total energy that conserves $(\rho E_{\text{tot}})$ as:
 \begin{equation}
@@
 \end{equation}
-where the sum is over the cell faces $f$ and the indices $f+1/2$ and $f-1/2$ refer to 
+where the sum is over the cell faces with indices $f$ and the indices $f+1/2$ and $f-1/2$ refer to 
 the zone centers immediately to the left and right in the direction perpendicular to the face.
-Note that while total energy is conserved, the gas energy is not because \autoref{eq:grav_energy_conservation_update} does not have the 
-telescoping property of a flux formalism where energy gained by one zone is lost by another. 
-The energy is being exchanged with a potential field that responds indirectly when the mass moves.
-
-The change in the density $\Delta \rho_{f}$ is proportional to the mass flux through the surface with area $\Delta A$:
-\begin{equation}
-  \Delta \rho_f = \frac{\Delta t}{\Delta V} (\Delta A\, \rho v_{f}),\label{eq:mass_transfer}
-\end{equation}
-So as long as we record these fluxes through the zone faces after coming out of the hydrodynamics step, 
+As long as we record the hydrodynamical fluxes through the zone faces after coming out of the hydrodynamics step, 
 we are able to conserve the total energy completely (except for any energy loss or gain through 
 physical domain boundaries). In order for the method to be second-order accurate in time, 
-no change is needed to the fluxes since those are already time-centered 
-in the hydrodynamics update, so we need only to use a time-centered $\Phi$ (which can be computed 
-by averaging the time-level $n$ and $n+1$ potentials; we already have the latter because \castro\ 
-re-computes the potential at the new time after the hydrodynamics step, and we can apply this energy 
-at the end of the timestep). Using this conservative energy formulation is very important for simulations where the 
-potential can change substantially over the course of the evolution, as massive amounts of energy can be lost. 
-The standard approach is therefore unacceptable in the case of a problem like white dwarf mergers, and 
-the fix to this energy leaking---evaluating the energy transfer at the six zone faces instead of 
-the single zone center---has a very minor cost in terms of code complexity and computational time.
-
-As observed by \cite{arepo}, when examining \autoref{eq:mass_transfer} one can see the relationship 
-between this method and the more common (non-conservative) approach of evaluating the change in 
-gas energy using the work done $(\mathbf{v} \cdot \rho \mathbf{g})$ by the gravitational force at the 
-cell center. Analytically this form expresses the same core idea as \autoref{eq:grav_energy_conservation_update} 
-via the work-energy theorem, but a major flaw is that it evaluates the energy change at the cell center 
-when in fact the mass transfer is happening at the cell edges. However, there is a more fundamental issue. 
-As observed by \citet[Chapter 4]{shu:1992} and \citet{marcello:2012}, the energy source term $(-\rho \mathbf{u}\cdot\nabla\Phi)$ 
-includes a contribution that is not localizable and reflects the global nature of the gravitational potential. 
-It is only by integrating this source term over all space that the correct gravitational potential energy 
-for the system emerges. Integrating over only one zone cannot yield the correct description of this global 
-energy flow, and indeed therefore can never obtain the factor of 1/2 that arises from avoiding double-counting 
-of the pairwise gravitational interactions. So while \autoref{eq:grav_energy_conservation_update} was correctly obtained 
-by \cite{arepo} (whose work was the primary inspiration for this approach), their derivation that integrates the work 
-done over the volume of a zone does not correctly account\footnote{They argued that this volume integral 
-of the work can be converted into a surface integral by use of the divergence theorem (see their Equation 93). That 
-application of the divergence theorem is mathematically incorrect, because their formulation of the integrand cannot actually 
-be expressed only as the divergence of a vector field.} for this behavior. It is only by accounting for the global nature of the
-potential that we can obtain the correct factor of $1/2$ and have a correct accounting of the energy within a zone.
-The insight of applying the source term on the zone faces to conserve energy is still valid when 
-considering energy transfer between the gas energy and a potential field if the change in energy 
-is explicitly localizable. Indeed, we employ this formulation in \autoref{sec:rotation} for the case of the rotational potential.
+we need to use a time-centered $\Phi$ (which can be computed by averaging the time-level $n$ and $n+1$ potentials;
+we already have the latter because \castro\ re-computes the potential at the new time after the hydrodynamics step,
+and we can apply this energy at the end of the timestep). Note that of course the hydrodynamical
+flux is already second-order accurate in time. In passing, we observe that the factor of $1/2$ that
+appears in \autoref{eq:grav_energy_conservation_update} has nothing to do with the factor of
+$1/2$ that appears in the statement of conservation of total energy, \autoref{eq:total_energy_gravity}.
+The former comes simply from the fact that the energy change is evaluated using the mass motion through
+a distance of half of the zone width. The latter is needed to ensure that these local changes in energy
+are not double-counted when doing a global integral, since the gravitational potential is self-generated.
+\autoref{eq:grav_energy_conservation_update} applies to any conservative potential $\Phi$, and we will
+use this to our advantage for the rotation forces in \autoref{sec:rotation}.
+
+As observed by \cite{arepo}, this is more accurate than the more common (non-conservative) approach
+of evaluating the change in gas energy using the work done $(\mathbf{v} \cdot \rho \mathbf{g})$
+by the gravitational force at the cell center. Analytically this form expresses the same core idea as
+\autoref{eq:grav_energy_conservation_update} via the work-energy theorem, but a major flaw is that
+it evaluates the energy change at the cell center when in fact the mass transfer is happening at
+the cell edges. This can result in a significant leaking of energy throughout the course of the
+evolution, dramatically affecting the course of the evolution. The standard approach is therefore
+unacceptable in the case of a problem like white dwarf mergers, and the fix to this energy
+leaking---evaluating the energy transfer at the six zone faces instead of the single zone
+center---has a very minor cost in terms of code complexity and computational time.
 
@@
 \begin{equation}
-  \Phi^R = -\frac{1}{2} \left| {\bm\omega} \times \mathbf{r} \right|^2.
-\end{equation}
-In the presence of rotation the total energy becomes:
-\begin{equation}
-  (\rho E_{\text{tot}}) = (\rho E) + \frac{1}{2} \rho \Phi + \rho \Phi^R.
+  \Phi^R = \frac{1}{2} \left| {\bm\omega} \times \mathbf{r} \right|^2.
+\end{equation}
+In the presence of rotation the conserved total energy becomes:
+\begin{equation}
+  \int dV (\rho E_{\text{tot}}) = \int dV \left( \rho E + \frac{1}{2} \rho \Phi + \rho \Phi^R \right).
 \end{equation}
@@
 \begin{align}
-  \left.\Delta(\rho E)\right|_{\text{rot}} &= - \sum_{f} \Delta \rho_{f} (\Phi^R_{f+1/2} - \Phi^R_{f-1/2})
+  \left.\Delta(\rho E)\right|_{\text{rot}} &= -\frac{1}{2}\sum_{f} \Delta \rho_{f} (\Phi^R_{f+1/2} - \Phi^R_{f-1/2})
 \end{align}
-Again we observe that this results a change in the global integral of $(\rho E)$, but keeps 
-the global integral of $(\rho E_{\text{tot}})$ conserved (except at physical domain boundaries).
 

### 2015-09-06 papers/wdmerger_II/paper.tex

@@
 %==========================================================================
-\title{Double White Dwarf Mergers with CASTRO\\ II. Equilibrium Initial Conditions}
+\title{White Dwarf Mergers on Adaptive Meshes\\ II. Collisions}
 
-\shorttitle{DWD Mergers. II. Equilibrium ICs}
-\shortauthors{Max}
+\shorttitle{WD Mergers. II. Collisions}
+\shortauthors{Katz et al. (2015)}
 
@@
 \begin{abstract}
-We describe a method for constructing equilibrium initial conditions for a binary star system,
-for a general equation of state. We extend this method to include the effects of adaptive
-mesh refinement in the algorithm.
+We consider the collisions of white dwarfs as possible progenitors of Type Ia 
+supernovae.
 
 \end{abstract}
-\keywords{hydrodynamics - methods: numerical - supernovae: general - white dwarfs}
+\keywords{supernovae: general - white dwarfs}
 

### 2015-09-08 papers/wdmerger_I/paper.tex
Point out that non-zero tolerances on gravity solve results in small deviations from conservation of energy.

@@
 As long as we record the hydrodynamical fluxes through the zone faces after coming out of the hydrodynamics step, 
-we are able to conserve the total energy completely (except for any energy loss or gain through 
+this algorithm is able to conserve the total energy completely (except for any energy loss or gain through 
 physical domain boundaries). In order for the method to be second-order accurate in time, 
@@
 and we can apply this energy at the end of the timestep). Note that of course the hydrodynamical
-flux is already second-order accurate in time. In passing, we observe that the factor of $1/2$ that
-appears in \autoref{eq:grav_energy_conservation_update} has nothing to do with the factor of
-$1/2$ that appears in the statement of conservation of total energy, \autoref{eq:total_energy_gravity}.
-The former comes simply from the fact that the energy change is evaluated using the mass motion through
-a distance of half of the zone width. The latter is needed to ensure that these local changes in energy
-are not double-counted when doing a global integral, since the gravitational potential is self-generated.
-\autoref{eq:grav_energy_conservation_update} applies to any conservative potential $\Phi$, and we will
-use this to our advantage for the rotation forces in \autoref{sec:rotation}.
+flux is already second-order accurate in time. We observe also that in practice we will not obtain 
+conservation of energy to machine precision even in the absence of open domain boundaries. The 
+method itself is conservative if it is time-centered and correctly evaluates the energy change 
+on cell faces \citep{marcello:2012}. However, in practice there is a non-zero numerical tolerance associated 
+with the Poisson gravity solver (in our case, the multigrid method) that results in a non-zero error 
+in the calculation of the gravitational potential. This results in a very small deviation from perfect 
+conservation. It is not usually larger than the other effects which result in energy non-conservation 
+for our simulations, such as resetting the state of zones that acquire a negative internal energy, and 
+in principle if desired it can be made smaller by using stricter tolerance levels on the gravity solve.
+
+In passing, the factor of $1/2$ that appears in \autoref{eq:grav_energy_conservation_update} 
+has nothing to do with the factor of $1/2$ that appears in the statement of conservation of total energy, 
+\autoref{eq:total_energy_gravity}. The former comes simply from the fact that the energy change is 
+evaluated using the mass motion through a distance of half of the zone width. The latter is needed 
+to ensure that these local changes in energy are not double-counted when doing a global integral, 
+since the gravitational potential is self-generated. \autoref{eq:grav_energy_conservation_update} 
+applies to any conservative potential $\Phi$, and we will use this to our advantage for the 
+rotation forces in \autoref{sec:rotation}.
 

### 2015-09-13 documentation/userguide.tex
The code is now configured to run in 2D axisymmetric coordinates. Plus I did a little general cleanup while I was working on that.

@@
 
+Though we are mainly interested in 3D simulations, it is sometimes useful to compare to 2D
+simulations which are much cheaper and can be run at higher resolutions. The only physically
+meaningful analogue to these mergers in lower dimensions occurs in 2D axisymmetric cylindrical
+coordinates. CASTRO represents these as $(r,z,\theta)$. In this coordinate system we enforce
+that the stars lie along the $z$ axis and that the center of the stars lies at $r = 0$. The
+$\theta$ dimension doesn't get directly evolved hydrodynamically but that is the dimension
+along which the rotational motion occurs, so in some circumstances the velocity in that
+direction will change with time because we still allow the $\theta$-velocity to participate in
+the Coriolis force. When rotation is enabled, the rotation is along the $z = 0$ axis in the $r$
+coordinate dimension. To try this out you should use \texttt{DIM = 2} in the makefile and
+use \texttt{inputs\_2d} in the \texttt{source/} directory.
+
 For the nuclear network we use the \texttt{aprox13} network provided by Frank Timmes, which 
@@
   \item \texttt{nsub}: number of sub-intervals to divide a zone into when interpolating from the initial model (integer)
-  \item \texttt{orbital\_kick}: whether or not to kick the stars with an orbital velocity corresponding to their 
+  \item \texttt{no\_orbital\_kick}: whether or not to kick the stars with an orbital velocity corresponding to their 
     Keplerian rotation rate (logical; this only makes sense in the inertial reference frame with \texttt{castro.do\_rotation = 0})

### 2015-09-13 source/Prob.cpp

@@ -25,7 +25,7 @@ Castro::wdCOM (Real time, Real& mass_p, Real& mass_s, Real* com_p, Real* com_s, 
     MultiFab*   mfxmom   = derive("xmom",time,0);
     MultiFab*   mfymom   = derive("ymom",time,0);
     MultiFab*   mfzmom   = derive("zmom",time,0);
-
+    
     BL_ASSERT(mfrho  != 0);
     BL_ASSERT(mfxmom != 0);
     BL_ASSERT(mfymom != 0);
@@ -67,17 +67,18 @@ Castro::wdCOM (Real time, Real& mass_p, Real& mass_s, Real* com_p, Real* com_s, 
 	FArrayBox& fabxmom = (*mfxmom)[mfi];
 	FArrayBox& fabymom = (*mfymom)[mfi];
 	FArrayBox& fabzmom = (*mfzmom)[mfi];
-    
+	
         const Box& box  = mfi.tilebox();
         const int* lo   = box.loVect();
         const int* hi   = box.hiVect();
 
 	BL_FORT_PROC_CALL(WDCOM,wdcom)
-            (BL_TO_FORTRAN(fabrho),
-	     BL_TO_FORTRAN(fabxmom),
-	     BL_TO_FORTRAN(fabymom),
-	     BL_TO_FORTRAN(fabzmom),
-	     lo,hi,dx,&time,
+            (BL_TO_FORTRAN_3D(fabrho),
+	     BL_TO_FORTRAN_3D(fabxmom),
+	     BL_TO_FORTRAN_3D(fabymom),
+	     BL_TO_FORTRAN_3D(fabzmom),
+	     BL_TO_FORTRAN_3D(volume[mfi]),
+	     ARLIM_3D(lo),ARLIM_3D(hi),ZFILL(dx),&time,
 	     &com_p_x, &com_p_y, &com_p_z,
 	     &com_s_x, &com_s_y, &com_s_z,
 	     &vel_p_x, &vel_p_y, &vel_p_z,
@@ -158,7 +159,7 @@ void Castro::volInBoundary (Real               time,
         const int* hi   = box.hiVect();
 
 	BL_FORT_PROC_CALL(CA_VOLUMEINDENSITYBOUNDARY,ca_volumeindensityboundary)
[…]

### 2015-09-13 source/Problem.H

@@ -1,11 +1,17 @@
 // Define a preprocessor directive that determines if we're doing a full merger simulation.
 // This is useful for sharing C++ files like the diagnostics between multiple test problems.
 
+#ifndef merger
 #define merger
+#endif
 
 // Preprocessor directive for allowing us to do a post-initialization step.
 
+#if (BL_SPACEDIM==3)
+#ifndef do_problem_post_init
 #define do_problem_post_init
+#endif
+#endif
 
 // Routine for determining white dwarf masses, centers of mass and velocities.
 

### 2015-09-13 source/Problem.f90

@@ -110,10 +110,11 @@ end subroutine
 ! going back to C++ for the full MPI reduce. We'll then update the module locations accordingly.
 ! The same is true for mass_S and mass_P versus m_s and m_p.
 
-subroutine wdcom(rho,  r_l1, r_l2, r_l3, r_h1, r_h2, r_h3, &
-                 xmom, x_l1, x_l2, x_l3, x_h1, x_h2, x_h3, &
-                 ymom, y_l1, y_l2, y_l3, y_h1, y_h2, y_h3, &
-                 zmom, z_l1, z_l2, z_l3, z_h1, z_h2, z_h3, &
+subroutine wdcom(rho,  r_lo, r_hi, &
+                 xmom, px_lo, px_hi, &
+                 ymom, py_lo, py_hi, &
+                 zmom, pz_lo, pz_hi, &
+                 vol,  vo_lo, vo_hi, &
                  lo, hi, dx, time, &
                  com_p_x, com_p_y, com_p_z, & 
                  com_s_x, com_s_y, com_s_z, &
@@ -127,15 +128,18 @@ subroutine wdcom(rho,  r_l1, r_l2, r_l3, r_h1, r_h2, r_h3, &
 
   implicit none
 
-  integer         , intent(in   ) :: r_l1, r_l2, r_l3, r_h1, r_h2, r_h3
-  integer         , intent(in   ) :: x_l1, x_l2, x_l3, x_h1, x_h2, x_h3
-  integer         , intent(in   ) :: y_l1, y_l2, y_l3, y_h1, y_h2, y_h3
-  integer         , intent(in   ) :: z_l1, z_l2, z_l3, z_h1, z_h2, z_h3
-
-  double precision, intent(in   ) :: rho(r_l1:r_h1,r_l2:r_h2,r_l3:r_h3)
-  double precision, intent(in   ) :: xmom(x_l1:x_h1,x_l2:x_h2,x_l3:x_h3)
-  double precision, intent(in   ) :: ymom(y_l1:y_h1,y_l2:y_h2,y_l3:y_h3)
-  double precision, intent(in   ) :: zmom(z_l1:z_h1,z_l2:z_h2,z_l3:z_h3)
+  integer         , intent(in   ) :: r_lo(3), r_hi(3)
+  integer         , intent(in   ) :: px_lo(3), px_hi(3)
+  integer         , intent(in   ) :: py_lo(3), py_hi(3)
[…]

### 2015-09-13 source/Problem_F.H

@@ -5,12 +5,13 @@ BL_FORT_PROC_DECL(PROBLEM_CHECKPOINT,problem_checkpoint)(int * int_dir_name, int
 BL_FORT_PROC_DECL(PROBLEM_RESTART,problem_restart)(int * int_dir_name, int * len);
 
 BL_FORT_PROC_DECL(WDCOM,wdcom)
-    (BL_FORT_FAB_ARG(rho), 
-     BL_FORT_FAB_ARG(xmom), 
-     BL_FORT_FAB_ARG(ymom), 
-     BL_FORT_FAB_ARG(zmom),
-     const int lo[], const int hi[],
-     const Real dx[], const Real* time,
+    (BL_FORT_FAB_ARG_3D(rho), 
+     BL_FORT_FAB_ARG_3D(xmom), 
+     BL_FORT_FAB_ARG_3D(ymom), 
+     BL_FORT_FAB_ARG_3D(zmom),
+     BL_FORT_FAB_ARG_3D(vol),
+     const int* lo, const int* hi,
+     const Real* dx, const Real* time,
      Real* com_p_x, Real* com_p_y, Real* com_p_z,
      Real* com_s_x, Real* com_s_y, Real* com_s_z,
      Real* vel_p_x, Real* vel_p_y, Real* vel_p_z,
@@ -18,7 +19,8 @@ BL_FORT_PROC_DECL(WDCOM,wdcom)
      Real* m_p, Real* m_s);
 
 BL_FORT_PROC_DECL(CA_VOLUMEINDENSITYBOUNDARY,ca_volumeindensityboundary)
-   (BL_FORT_FAB_ARG(rho), const int lo[], const int hi[], const Real dx[],  
+   (BL_FORT_FAB_ARG_3D(rho), BL_FORT_FAB_ARG_3D(vol),
+    const int* lo, const int* hi, const Real* dx,  
     Real* volp, Real* vols, Real* rho_cutoff);
 
 BL_FORT_PROC_DECL(GET_SINGLE_STAR,get_single_star)
@@ -37,15 +39,16 @@ BL_FORT_PROC_DECL(GET_STAR_DATA,get_star_data)
     Real* t_ff_p, Real* t_ff_s);
 
 BL_FORT_PROC_DECL(QUADRUPOLE_TENSOR_DOUBLE_DOT,quadrupole_tensor_double_dot)
[…]

### 2015-09-13 source/bc_fill_nd.f90

@@ -3,37 +3,36 @@
 
    subroutine ca_hypfill(adv,adv_lo,adv_hi,domlo,domhi,delta,xlo,time,bc)
 
-     use meth_params_module, only: NVAR, URHO, UMX, UMY, UMZ, UTEMP, &
-          UEDEN, UEINT, UFS, rot_period, do_rotation
-     use prob_params_module, only: center
-     use bl_constants_module
-     use probdata_module
-     use eos_module
-     use rotation_module, only: cross_product, get_omega
+     use bl_constants_module, only: HALF
+     use meth_params_module, only: NVAR
+     use prob_params_module, only: center, dim
+     use probdata_module, only: fill_ambient_bc, fill_ambient
 
      implicit none
 
      include 'bc_types.fi'
 
      integer          :: adv_lo(3), adv_hi(3)
-     integer          :: bc(3,2,*)
+     integer          :: bc(dim,2,NVAR)
      integer          :: domlo(3), domhi(3)
      double precision :: delta(3), xlo(3), time
      double precision :: adv(adv_lo(1):adv_hi(1),adv_lo(2):adv_hi(2),adv_lo(3):adv_hi(3),NVAR)
 
      integer          :: i, j, k, n
-     double precision :: loc(3), omega(3)
+     double precision :: loc(3)
 
-     type (eos_t)     :: ambient_state
-
+     ! First, use the generic filling routines to make sure we have valid data everywhere
+     ! on physical domain ghost cells.
+     
+     do n = 1,NVAR
+        call filcc_nd(adv(:,:,:,n),adv_lo,adv_hi,domlo,domhi,delta,xlo,bc(:,:,n))
+     enddo
+     
      ! Override the generic routine at the physical boundaries by
-     ! setting the material to the ambient state.
+     ! setting the material to the ambient state. Note that we
+     ! don't want to do this for interior/periodic boundaries, which have bc == 0.
 
      if (fill_ambient_bc) then
-
-        call get_ambient(ambient_state)
-
-        omega = get_omega(time)
 
         do k = adv_lo(3), adv_hi(3)
            loc(3) = xlo(3) + dble(k - adv_lo(3) + HALF)*delta(3) - center(3)
[…]

### 2015-09-13 source/diagnostics/sum_integrated_quantities.cpp

@@ -30,6 +30,7 @@ Castro::sum_integrated_quantities ()
     Real rho_e        = 0.0;
     Real rho_K        = 0.0;
     Real rho_phi      = 0.0;
+    Real rho_phirot   = 0.0;
 
     Real gravitational_energy = 0.0; 
     Real kinetic_energy       = 0.0; 
@@ -43,7 +44,6 @@ Castro::sum_integrated_quantities ()
     Real m_r_squared[3]          = { 0.0 };
 
     Real omega[3]       = { 0.0 };
-    Real rot_kin_eng    = 0.0;
     Real rot_mom[3]     = { 0.0 };
     Real rot_ang_mom[3] = { 0.0 };
 
@@ -145,6 +145,16 @@ Castro::sum_integrated_quantities ()
     BL_FORT_PROC_CALL(UPDATE_CENTER,update_center)(&time);
 #endif
 
+    // Determine the names of the species in the simulation.    
+    
+#ifdef merger
+    for (int i = 0; i < NumSpec; i++) {
+      species_names[i] = desc_lst[State_Type].name(FirstSpec+i);
+      species_names[i] = species_names[i].substr(4,std::string::npos);
+      species_mass[i]  = 0.0;	
+    }
+#endif    
+    
     for (int lev = 0; lev <= finest_level; lev++)
     {
 
@@ -152,38 +162,34 @@ Castro::sum_integrated_quantities ()
 
       Castro& ca_lev = getLevel(lev);
 
-      for ( int i = 0; i < BL_SPACEDIM; i++ ) {
+      for ( int i = 0; i < 3; i++ ) {
         lev_com[i]    = ca_lev.locWgtSum("density", time, i);
         com[i]       += lev_com[i];
       }
 
       // Calculate total mass, momentum and energy of system.
 
-      mass     += ca_lev.volWgtSum("density", time);
+      mass += ca_lev.volWgtSum("density", time);
 
-      mom_grid[0]     += ca_lev.volWgtSum("xmom", time);
-#if (BL_SPACEDIM >= 2)      
-      mom_grid[1]     += ca_lev.volWgtSum("ymom", time);
-#endif
-#if (BL_SPACEDIM == 3)
-      mom_grid[2]     += ca_lev.volWgtSum("zmom", time);
-#endif
+      mom_grid[0] += ca_lev.volWgtSum("xmom", time);
+      mom_grid[1] += ca_lev.volWgtSum("ymom", time);
+      mom_grid[2] += ca_lev.volWgtSum("zmom", time);
 
[…]

### 2015-09-13 source/probdata.f90

@@ -124,12 +124,18 @@ contains
   subroutine initialize(name, namlen, init_in)
 
     use bl_error_module, only: bl_error
-
+    use prob_params_module, only: dim
+    
     implicit none
 
     integer :: namlen, i, init_in
     integer :: name(namlen)
-
+ 
+    ! Safety check: we can't run this problem in one dimension.       
+    if (dim .eq. 1) then
+       call bl_error("Cannot run wdmerger problem in one dimension. Exiting.")
+    endif
+    
     init = init_in
 
     ! Build "probin" filename -- the name of the file containing the fortin namelist.
@@ -163,6 +169,7 @@ contains
   subroutine read_namelist
 
     use meth_params_module
+    use prob_params_module, only: dim, coord_type
 
     implicit none
 
@@ -309,6 +316,19 @@ contains
 
     call ensure_primary_mass_larger
 
+    ! We enforce star_axis = 2 (the z-axis) for two-dimensional problems.
+
+    if (dim .eq. 2) then
+
+       if (coord_type .ne. 1) then
+          call bl_error("We only support cylindrical coordinates in two dimensions. Set coord_type == 1.")
+       endif
+       
+       star_axis = 2
+       rot_axis = 1
+       
+    endif
+    
   end subroutine read_namelist
 
 
@@ -340,7 +360,7 @@ contains
     eos_state % rho = small_dens
     eos_state % T   = small_temp
     eos_state % xn  = ambient_comp
- 
+
     call eos(eos_input_rt, eos_state)
 
     small_pres = eos_state % p
@@ -378,23 +398,33 @@ contains
 
     use meth_params_module, only: do_rotation, rot_period
     use initial_model_module, only: initialize_model, establish_hse
-    use prob_params_module, only: center, problo, probhi
+    use prob_params_module, only: center, problo, probhi, dim
     use meth_params_module, only: rot_axis
 
     implicit none
 
-    type (eos_t) :: ambient_state
     double precision :: v_ff
-    double precision :: mass_temp
 
     ! Set up the center variable. We want it to be at 
     ! problo + center_frac * domain_width in each direction.
     ! center_frac is 1/2 by default, so the problem
     ! would be set up exactly in the center of the domain.
+    ! Note that we override this for 2D axisymmetric, as the
[…]

### 2015-09-13 source/problem_tagging_nd.f90

@@ -5,7 +5,7 @@ subroutine set_problem_tags(tag,tag_lo,tag_hi, &
                             dx,problo,time,level)
 
   use bl_constants_module, only: ZERO, HALF, TWO
-  use meth_params_module, only: URHO, UMX, UMY, UMZ, UEDEN, NVAR
+  use meth_params_module, only: NVAR
   use prob_params_module, only: center, probhi
   use probdata_module, only: maxTaggingRadius, com_P, com_S, roche_rad_P, roche_rad_S  
   

### 2015-09-13 source/scf_relax.cpp

@@ -5,6 +5,7 @@
 
 #ifdef GRAVITY
 #ifdef ROTATION
+#ifdef do_problem_post_init
 void Castro::scf_relaxation() {
 
   int finest_level = parent->finestLevel();
@@ -51,7 +52,6 @@ void Castro::scf_relaxation() {
 
   MultiFab& phi = get_new_data(PhiGrav_Type);
 
-  int ns    = NUM_STATE;
   Real time = state[State_Type].curTime();
 
   // Iterate until the system is relaxed by filling the level data 
@@ -272,5 +272,6 @@ void Castro::scf_relaxation() {
   }
 
 }
+#endif
 #endif
 #endif

### 2015-09-17 papers/wdmerger_I/paper.tex
Add an appendix demonstrating that the time-centered energy update conserves total energy in simulations with self-gravity.

@@
 to advance the hydrodynamics system in time \citep{ppmunsplit}.  A
-number of changes were made to the solver, which are detailed in the
-Appendix.  These changes bring the algorithm more in line with that of
+number of changes were made to the solver, which are detailed in \autoref{app:hydro}.
+These changes bring the algorithm more in line with that of
 \cite{ppm}. \castro\ as originally released featured a slightly modified
@@
 under the source term; the details of this are described in 
-the appendix. Second, after the hydrodynamics step, we add the time-centered source terms
+\autoref{app:hydro}. Second, after the hydrodynamics step, we add the time-centered source terms
 to the state. First we will describe how we do this for the momentum,
@@
 method itself is conservative if it is time-centered and correctly evaluates the energy change 
-on cell faces \citep{marcello:2012}. However, in practice there is a non-zero numerical tolerance associated 
+on cell faces. This was demonstrated empirically by \cite{jiang:2013} and is obvious in the case of a
+fixed external potential; it is not as obvious in the case of the gravitational potential, which
+changes in response to the mass distribution, so we give a short proof of this in \autoref{app:gravity}.
+However, in practice there is a non-zero numerical tolerance associated 
 with the Poisson gravity solver (in our case, the multigrid method) that results in a non-zero error 
@@
 $T$ is the rotation period, which is the most natural quantity to specify
-for a rotating stellar system. As described in the appendix, we include this source term
+for a rotating stellar system. As described in \autoref{app:hydro}, we include this source term
 in the edge state prediction in a way that is analogous to the gravity source.
@@
 \section{CASTRO hydrodynamics changes}
+\label{app:hydro}
 
@@
 
+\section{Proof of Energy Conservation in Simulations using Self-Gravity}
+\label{app:gravity}
+
+In \autoref{sec:gravity_hydro_coupling}, we described our approach to updating the gas energy
+in response to motions of fluid through the self-generated gravitational potential using 
+\autoref{eq:grav_energy_conservation_update}. While it is straightforward to observe that this approach
+should be conservative for an arbitrary fixed external potential $\Phi$, it is not as obvious that this
+should be so for a self-generated potential which changes in response to mass motions on the domain. To
+see that this still holds for the self-generated gravitational potential $\Phi$, let us start with 
+\autoref{eq:grav_energy_conservation_update} in a slightly revised form:
+\begin{equation}
+  \Delta(\rho E)_i = -\frac{1}{2}\sum_{j} \Delta\rho_{ij}(\Phi_i - \Phi_{j}) \label{eq:grav_energy_conservation_update_revised}
+\end{equation}
+where by $\Delta \rho_{ij}$ we mean the density transferred from zone $j$ to zone $i$, so that
+$\Delta \rho_{ij} = - \Delta \rho_{ji}$, and the sum is over all zone indices $j$ that are adjacent
+to zone $i$. Let us define $\Phi_{ij} = \Phi_{ji} = (\Phi_{i} + \Phi_{j}) / 2$ as the potential on the
+zone interface between zones $i$ and $j$. Then we have:
+\begin{equation}
+  \Delta(\rho E)_i = -\sum_{j} \Delta\rho_{ij}(\Phi_i - \Phi_{ij}).
+\end{equation}
+We can evalute the sum for all of the terms proportional to $\Phi_i$ by observing that the change in
+density from time-level $n$ to time-level $n+1$ is the sum of the density fluxes from all adjacent zones.
+\begin{equation*}
+  \Delta(\rho E)_i = - (\rho_i^{n+1} - \rho_i^{n}) \Phi_i + \sum_{j}\Delta \rho_{ij} \ \Phi_{ij}
+\end{equation*}
+Now let us sum this over all zones $i$ in the domain, and ignore the domain boundaries, or assume that they are
+far enough away from the region of compact support for $\rho$ that $\Phi$ is negligible there. As the second
+term on the right-hand side is antisymmetric in $i$ and $j$, it cancels when summing adjacent zones, and we have:
+\begin{equation*}
+  \sum_{i} (\rho E)_i^{n+1} - \sum_{i} (\rho E)_i^{n} = -\frac{1}{2}\sum_{i} (\Phi_{i}^{n+1} + \Phi_{i}^{n})(\rho_i^{n+1} - \rho_i^{n})
+\end{equation*}
+Note that, as explained the text, we are using a time-centered $\Phi$ to correspond to the mass fluxes
+at time-level $n+1/2$. Finally we re-write this in a form where the difference in total energy between time-levels
+$n$ and $n+1$ is on the left-hand side and any sources causing this to be non-zero are on the right-hand side:
+\begin{align}
+  \sum_{i} \left(\rho E + \frac{1}{2}\rho\Phi\right)_i^{n+1} - \sum_{i} \left(\rho E + \frac{1}{2}\rho\Phi\right)_i^{n} &= \frac{1}{2}\sum_{i} \left(\Phi_i^{n+1}\rho_i^{n} - \Phi_i^{n}\rho_i^{n+1}\right) \notag \\
+       &= \frac{1}{8\pi G} \sum_{i}\left(\Phi_{i}^{n+1}\nabla^2 \Phi_{i}^{n} - \Phi_i^{n}\nabla^2 \Phi_{i}^{n+1}\right) \label{eq:total_energy_difference}
+\end{align}
+\autoref{eq:total_energy_difference} expresses total energy conversation if and only if the right-hand side vanishes.
+We observe that the right-hand side has the form of a variant of the divergence theorem often called Green's second identity:
+\begin{equation}
+  \int (\Phi^{n}\nabla^2 \Phi^{n+1} - \Phi^{n+1}\nabla^2 \Phi^{n}) dV = \int \left(\Phi^{n} \nabla \Phi^{n+1} - \Phi^{n+1} \nabla \Phi^{n}\right) \cdot d\mathbf{S}, \label{eq:green_second_identity}
+\end{equation}
+where $d\mathbf{S}$ is the area element with vector component parallel to the outward normal. The analogous result holds for the
+discretized form in \autoref{eq:total_energy_difference}. With the assumptions used above, the right-hand side of
+\autoref{eq:green_second_identity} will vanish as the surface integral is evaluated at infinity, where the potential
+tends to zero. This concludes the proof that the method is conservative when the potential used at the zone interfaces
+is time-centered, even in light of the change of the potential over the timestep due to the mass motion that is causing the change in the energy.
+
+From the above discussion it is straightforward to see exactly why the method will not be fully conservative to machine
+precision in practice. First, we cannot simulate the domain out to infinity, so Green's second identity does not hold exactly
+and there will be some loss or addition of energy at domain boundaries. Second, \autoref{eq:total_energy_difference} holds in the
+continuum limit by using the Poisson equation, but in practice it is not exactly true that $\rho_i = 4\pi G \nabla^2 \Phi_{i}$ due
+to small errors in the potential at the level of the tolerances used in the Poisson solver.
+
 \clearpage

### 2015-09-19 papers/wdmerger_I/paper.tex

@@
 exciting problems in astrophysics, requiring detailed numerical modeling to 
-complement the observations of these explosions. One possibility that has 
+complement the observations of these explosions. One possible progenitor that has 
 merited recent theoretical attention is the white dwarf merger scenario.
@@
 To date there have been few fully self-consistent simulations 
-of binary white dwarf systems using mesh-based hydrodynamics, 
+of merging binary white dwarf systems using mesh-based hydrodynamics, 
 relative to other methods. This is the first paper in a series designed to 
 describe simulations of these systems using a mesh-based hydrodynamics code with 
-the adaptive-mesh refinement approach. In this paper we describe our numerical 
+adaptive-mesh refinement. In this paper we describe our numerical 
 methodology and discuss our implementation in the code \castro. \castro\ 
-solves the Euler equations of hydrodynamics 
+solves the Euler equations 
 and the Poisson equation for self-gravity. Standard techniques for 
@@
 but recent advances in the literature allow progress on this 
-problem and we discuss our implementation here. We present an 
-extensive set of test problems demonstrating the extent to which our software can sufficiently
+problem and we discuss our implementation here. We present a
+set of test problems demonstrating the extent to which our software can sufficiently
 model a system where large amounts of mass are advected on the computational 
@@
 \end{abstract}
+
 \keywords{hydrodynamics - methods: numerical - supernovae: general - white dwarfs}
@@
 model of a WD merger. Another reason for caution is that other than the
-most recent results of \cite{kashyap:2015}, no simulation has self-consistently
+most recent results of \cite{kashyap:2015}, no white dwarf merger simulation has self-consistently
 resulted in a thermonuclear detonation. Reproducibility of the detonation 
@@
 version of the higher resolution limiters of
-\cite{colella_sekora:2008}, which can be accessed in the code using
-\texttt{castro.ppm\_type = 2}. The advantage of this limiter is that
+\cite{colella_sekora:2008}, which can be used in the code by setting 
+\texttt{castro.ppm\_type = 2} in the inputs file, which is a set of code parameters
+accessed at runtime to determine the algorithms used in the simulation.
+The advantage of this limiter is that
 it preserves physical extrema rather than clipping them off as in the
@@
 monotonicity preserving in the parabolic profiles it generates; this
-is activated with \texttt{castro.ppm\_type = 1}.
+is activated with \texttt{castro.ppm\_type = 1} in the inputs file.
 
@@
 enclosed, then the potential can be expanded in a series of spherical
-harmonics $Y_{lm}$:
-\begin{equation}
-  \Phi(\mathbf{x}) = -\sum_{l=0}^{\infty}\sum_{m=-l}^{l} \frac{4\pi}{2l + 1} q_{lm} \frac{Y_{lm}(\theta,\phi)}{r^{l+1}}, \label{spherical_harmonic_expansion}
+harmonics $Y_{lm}(\theta,\phi)$ (where $\theta \in [0, \pi]$ is the usual polar angle
+with respect to the $z$ axis and $\phi \in [0, 2\pi)$ is the usual azimuthal angle with
+respect to the positive $x$ axis):
+\begin{equation}
+  \Phi(\mathbf{x}) = -\sum_{l=0}^{\infty}\sum_{m=-l}^{l} \frac{4\pi}{2l + 1} q_{lm} \frac{Y_{lm}(\theta,\phi)}{r^{l+1}}, \label{eq:spherical_harmonic_expansion}
 \end{equation}
@@
 can be calculated by expanding the Green's function for the Poisson
-equation as a series of spherical harmonics.
-%% \begin{equation}
-%%   q_{lm} = \int Y^*_{lm}(\theta^\prime, \phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3x^\prime. \label{multipole_moments_original}
-%% \end{equation}
-After some algebraic simplification of \autoref{spherical_harmonic_expansion} 
-using the addition theorem for spherical harmonics,
-%% \begin{align}
-%%   &\frac{4\pi}{2l+1} \sum_{m=-l}^{l} Y^*_{lm}(\theta^\prime,\phi^\prime)\, Y_{lm}(\theta, \phi) = P_l(\text{cos}\, \theta) P_l(\text{cos}\, \theta^\prime) \notag \\
-%%   &\ \ + 2 \sum_{m=1}^{l} \frac{(l-m)!}{(l+m)!} P_{l}^{m}(\text{cos}\, \theta)\, P_{l}^{m}(\text{cos}\, \theta^\prime)\, \left[\text{cos}(m\phi)\, \text{cos}(m\phi^\prime) + \text{sin}(m\phi)\, \text{sin}(m\phi^\prime)\right].
-%% \end{align}
+equation as a series of spherical harmonics. After some algebraic
+simplification of \autoref{eq:spherical_harmonic_expansion}, 
 the potential outside of the mass distribution can be written as:
@@
   \Phi(\mathbf{x}) &= -\sum_{l=0}^{\infty} \left\{Q_l^{(0)} \frac{P_l(\text{cos}\, \theta)}{r^{l+1}} \right. \notag \\
-    &+ \left. \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right\}.\label{multipole_potential}
+    &+ \left. \sum_{m = 1}^{l}\left[ Q_{lm}^{(C)}\, \text{cos}(m\phi) + Q_{lm}^{(S)}\, \text{sin}(m\phi)\right] \frac{P_{l}^{m}(\text{cos}\, \theta)}{r^{l+1}} \right\}.\label{eq:multipole_potential}
 \end{align}
 $P_l(x)$ are the Legendre polynomials and $P_{l}^{m}(x)$ are the associated Legendre polynomials.
-$Q_l^{(0)}$ and $Q_{lm}^{(C,S)}$ are variants of the multipole moments that involve integrals of $P_l$ and $P_l^m$, respectively, over the computational domain.
-%% \begin{align}
-%%   Q_l^{(0)}   &= \int P_l(\text{cos}\, \theta^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
-%%   Q_{lm}^{(C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
-%%   Q_{lm}^{(S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime.
-%% \end{align}
+$Q_l^{(0)}$ and $Q_{lm}^{(C,S)}$ are variants of the multipole moments that involve integrals of
+$P_l$ and $P_l^m$, respectively, over the computational domain; their definition is given in \autoref{app:multipole}.
+
 This approach becomes computationally feasible when we cut off the
-outer summation in \autoref{multipole_potential} at some finite
+outer summation in \autoref{eq:multipole_potential} at some finite
 value of $l_{\text{max}}$. If it is of sufficiently high order, we
@@
 show that there are no gains to be had by increasing the accuracy further. At 
-very high orders the approximation breaks down, as seen in \autoref{fig:bc_comparison}. 
+very high orders ($l \gtrsim 18$) the approximation breaks down, as seen in \autoref{fig:bc_comparison}. 
 This is a result of the ambient material on the grid. At each boundary point we 
@@
 
-In a rotating reference frame with angular frequency vector $\bm{\omega}$, the momentum equation becomes:
+In a rotating reference frame with angular frequency vector $\bm{\omega}$,
+non-inertial contribution to the momentum equation is:
 \begin{equation}
@@
 The frequency of the gravitational waves emitted by the white dwarf binary 
-will be similar in size to the frequency of the orbital motion, which is in the range 
+will be similar to the frequency of the orbital motion, which is in the range 
 10-100 mHz for our problem. This is well outside the range of currently existing 
-gravitational wave detectors but is very well suited for proposed space-based detectors such as eLISA.
+gravitational wave detectors but is very well suited for proposed space-based detectors such as eLISA \citep{eLISA}.
 
@@
 repeat this process for any number of observing locations at minimal cost, since the quadruple tensor 
-only needs to be calculated once. Now, gravitational waves only excite modes orthogonal to their 
+only needs to be calculated once. Gravitational waves only excite modes orthogonal to their 
 direction of travel. These are the ``plus'' and ``cross'' modes, $h_+$ and $h_\times$, named after 
@@
 
-Merger simulations face a number of numerical difficulties that are
+White dwarf merger simulations face a number of numerical difficulties that are
 not present in single-degenerate Type Ia and core-collapse supernova
@@
 of this effect, we considered a single orbit of the unequal mass system and measured 
-the distance between the two white dwarfs at the beginning of the simulation and after 
+the distance between the two white dwarf centers of mass at the beginning of the simulation and after 
 the full orbital period. This distance should not change significantly over that timescale.
@@
 In \autoref{table:sources} we show also the effects on energy conservation of using the inertial reference frame. 
-We use ${\tt rs}$ for the \castro\ parameter {\tt castro.rot\_source\_type}, and we have the same four options for updating
-the energy source term as for gravity. The ${\tt rs} = 0$ column means that rotation is turned off and we are 
+We use ${\tt rs}$ for the \castro\ parameter {\tt castro.rot\_source\_type}.
+Each option for ${\tt rs}$ is implemented in the same way as for
+the gravitational source term, simply swapping out the gravitational acceleration
+for the rotational acceleration (except for the improvement to the momentum update
+for ${\tt rs} = 4$ described in \autoref{sec:rotation}). 
+The ${\tt rs} = 0$ column means that rotation is turned off and we are 
 in the inertial frame. We see that the choice of rotational coupling is much less important than the choice of gravity coupling. 
@@
 
-The \boxlib\ framework that \castro\ is based on splits each AMR level into a number 
+The \boxlib\ framework that \castro\ is based on domain decomposes each AMR level into a number 
 of boxes that collectively span the level. These boxes are distributed to processors 
@@
 work in this way, and by default uses a space-filling curve approach with a 
-Morton ordering. By experiment we have found that the most efficient load-balancing 
+Morton ordering (e.g. \cite{sasidharan:2015,beichl:1998}). By experiment we have found that the most efficient load-balancing 
 strategy for our problem is actually a simple knapsack algorithm. In this approach, 
@@
 We note a few additional differences between the original PPM
-implementation of \citet{ppm} and \castro\.  In the original PPM
+implementation of \citet{ppm} and \castro.  In the original PPM
 implementation, the gravitational acceleration was reconstructed as a parabola, and
@@
 
+\section{Formulation of the Multipole Expansion for the Gravitational Potential}
+\label{app:multipole}
+
+The integral formulation of the gravitational potential, using a series expansion in spherical harmonics, is:
+\begin{equation}
+  \Phi(\mathbf{x}) = -G\sum_{l=0}^{\infty} \sum_{m=-l}^{l} \frac{4\pi}{2l+1} \int \rho(\mathbf{x}^\prime)\, Y_{lm}(\theta,\phi)\, Y_{lm}^*(\theta^\prime,\phi^\prime)\, \frac{r_{<}^{l}}{r_{>}^{l+1}}\, dV^\prime,
+\end{equation}
+where $\theta$ is the polar angle and $\phi$ is the azimuthal angle, $r \equiv |\mathbf{x}|$ is the radial distance, and at any point in the domain $r_{<}$ is the smaller of $r$ and $r^\prime$, and $r_{>}$ is the larger of the two.
+This immediately suggests writing the potential at any location as the sum of
+two series:
+\begin{equation*}
+  \Phi(\mathbf{x}) = -G\sum_{l=0}^{\infty} \sum_{m=-l}^{l} \frac{4\pi}{2l+1}\left[ q^{L}_{lm}(\mathbf{x})\, r^{-l-1} + q^{U}_{lm}(\mathbf{x})\, r^{-l-1} \right] Y_{lm}(\theta,\phi),
+\end{equation*}
+where we have defined two multipole moments as integrals over the domain:
+\begin{align}
+  q^{L}_{lm}(\mathbf{x}) &= \int dV^\prime\, \rho(\mathbf{x}^\prime)\, Y^*(\theta^\prime,\phi^\prime)\, \Theta(r - r^\prime)\, {r^\prime}^{l} \\
+  q^{U}_{lm}(\mathbf{x}) &= \int dV^\prime\, \rho(\mathbf{x}^\prime)\, Y^*(\theta^\prime,\phi^\prime)\, \Theta(r^\prime - r)\, {r^\prime}^{-l-1}.
+\end{align}
+$\Theta(r)$ is the standard step function, equal to one if the argument is positive and zero if the argument is negative. Geometrically,
+$q^{L}(\mathbf{x})$ is an integral containing only mass interior to $|\mathbf{x}|$, and $q^{U}(\mathbf{x})$ is an integral containing only
+mass exterior to $|\mathbf{x}|$. Provided that one has computed these two integrals for a point $\mathbf{x}$, one can use the series expansion
+to calculate the potential at that point in principle to arbitrary accuracy by including higher order terms.
+
+We prefer to work with solely real-valued quantities, and so we make use of the addition theorem for spherical harmonics \citep[Section 3.6]{jackson}:
+\begin{align}
+  \frac{4\pi}{2l+1} \sum_{m=-l}^{l} Y^*_{lm}(\theta^\prime,\phi^\prime)\, Y_{lm}(\theta, \phi) &= P_l(\text{cos}\, \theta) P_l(\text{cos}\, \theta^\prime) \notag \\
+   &+ 2 \sum_{m=1}^{l} \frac{(l-m)!}{(l+m)!} P_{l}^{m}(\text{cos}\, \theta)\, P_{l}^{m}(\text{cos}\, \theta^\prime)\, \left[\text{cos}(m\phi)\, \text{cos}(m\phi^\prime) + \text{sin}(m\phi)\, \text{sin}(m\phi^\prime)\right].
+ \end{align}
+The $P_l(x)$ are the Legendre polynomials and the $P_l^m(x)$ are the associated Legendre polynomials.
+We construct them using a stable recurrence relation given known values for $l = 0$ and $l = 1$.
+We can then formulate the expansion in a different way:
+\begin{align}
+  \Phi(\mathbf{x}) &= -G\sum_{l=0}^{\infty} \left\{ Q_{l}^{(L,0)}(\mathbf{x})\, P_l(\text{cos}\, \theta)\, r^{-l-1} + Q_{l}^{(U,0)}(\mathbf{x})\, P_l(\text{cos}\, \theta)\, r^{l} \right. \notag \\
+  &\hspace{0.7in} + \sum_{m=1}^{l} \left[ Q_{lm}^{(L,C)}(\mathbf{x})\, \text{cos}(m\phi) + Q_{lm}^{(L,S)}(\mathbf{x})\, \text{sin}(m\phi)\right] P_{l}^{m}(\text{cos}\, \theta)\, r^{-l-1} \notag \\
+  &\hspace{0.7in} + \left. \sum_{m=1}^{l} \left[ Q_{lm}^{(U,C)}(\mathbf{x})\, \text{cos}(m\phi) + Q_{lm}^{(U,S)}(\mathbf{x})\, \text{sin}(m\phi)\right] P_{l}^{m}(\text{cos}\, \theta)\, r^{l} \right\}
+\end{align}
+
+The multipole moments now take the form:
+\begin{align}
+  Q_l^{(L,0)}(\mathbf{x}) &= \int P_l(\text{cos}\, \theta^\prime)\, \Theta(r - r^\prime)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
+  Q_l^{(U,0)}(\mathbf{x}) &= \int P_l(\text{cos}\, \theta^\prime)\, \Theta(r^\prime - r)\, {r^{\prime}}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\  
+  Q_{lm}^{(L,C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, \Theta(r - r^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
+  Q_{lm}^{(U,C)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{cos}(m\phi^\prime)\, \Theta(r^\prime - r)\, {r^\prime}^{-l-1} \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
+  Q_{lm}^{(L,S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, \Theta(r - r^\prime)\, {r^\prime}^l \rho(\mathbf{x}^\prime)\, d^3 x^\prime \\
+  Q_{lm}^{(U,S)} &= 2\frac{(l-m)!}{(l+m)!} \int P_{l}^{m}(\text{cos}\, \theta^\prime)\, \text{sin}(m\phi^\prime)\, \Theta(r^\prime - r)\, {r^\prime}^{-l-1} \rho(\mathbf{x}^\prime)\, d^3 x^\prime.  
+\end{align}
+In practice, of course, we select some maximum value $l_{\text{max}}$ at which we terminate the summation, determined either by computational efficiency requirements or by the fact that there is little information at high orders for sufficiently smooth mass distributions.
+In \castro\ we have the capability to compute any of the above multipole moments, though in this paper we are only using the multipole expansion to calculate the boundary conditions on the potential, and so we neglect calculation of the moments with a $U$ subscript as we are assuming that all of the mass is interior to the boundary.
+\autoref{eq:multipole_potential} is directly recovered under these conditions.
+
 \clearpage

### 2015-09-21 papers/wdmerger_I/paper.tex
Add description of the source-term predictor to the appendix.

@@
 \subsection{Gravity and Rotation Tracing}
+\label{sec:gravity_rotation_tracing}
 
@@
 
+\subsection{Source Term Predictor for the Hydrodynamics}
+
+In the original release of \castro\ the value of the gravitational and rotation
+source terms used in constructing the edge-states for the hydrodynamics update
+was the time-level $n$ value. While this is formally second-order accurate, there
+is a better choice one can make. We have information about the trend of these
+source terms from previous timesteps, so we can use a predictor method to guess
+at a more accurate value of the gravitational and rotational fields at the
+$n+1/2$ time-level the hydro is evaluated at. Our method uses a lagged linear
+extrapolation. Going into the hydro update, we have both the time-level $n$ and
+time-level $n-1$ data for $g$ and $f_R$ (as defined in \autoref{sec:gravity_rotation_tracing}).
+From this one can construct a simple linear estimator of (say) the gravitational
+acceleration using a backward difference scheme at time-level $n$:
+\begin{equation}
+  \mathbf{g}(t_n + \Delta t) = \mathbf{g}(t_n) + \Delta t \left.\frac{d\mathbf{g}}{dt}\right|_{n} = \frac{3}{2}\mathbf{g}(t_n) - \frac{1}{2}\mathbf{g}(t_{n-1}).
+\end{equation}
+Here we have assumed a constant timestep for simplicity, and constructed the time derivative as
+$d\mathbf{g} / dt = (\mathbf{g}(t_n) - \mathbf{g}(t_{n-1})) / \Delta t$. We perform this
+predictor for the rotation field as well. While both this method and the original
+method have second-order convergence properties, we have seen from testing that
+this new method is slightly more accurate in absolute terms.
+
 \section{Proof of Energy Conservation in Simulations using Self-Gravity}

### 2015-09-21 papers/wdmerger_I/paper.tex
Point out that we don't do source-term predictor when we don't have good old-time data
Used numbered appendix option
More paper I updates in response to comments

@@
-\documentclass[iop]{../emulateapj}
+\documentclass[iop,numberedappendix]{../emulateapj}
 
@@
 Type Ia supernovae, as well as their distribution in time and space.
-To date there have been few fully self-consistent simulations 
-of merging binary white dwarf systems using mesh-based hydrodynamics, 
-relative to other methods. This is the first paper in a series designed to 
+To date there have been relatively few fully self-consistent simulations 
+of merging binary white dwarf systems using mesh-based hydrodynamics. 
+This is the first paper in a series designed to 
 describe simulations of these systems using a mesh-based hydrodynamics code with 
-adaptive-mesh refinement. In this paper we describe our numerical 
+adaptive mesh refinement. In this paper we describe our numerical 
 methodology and discuss our implementation in the code \castro. \castro\ 
-solves the Euler equations 
-and the Poisson equation for self-gravity. Standard techniques for 
-coupling gravitation and rotation forces to the hydrodynamics do 
-not do an adequate job conserving the total energy of the system, 
+solves the Euler equations, and for merging white dwarf simulations we use its capability to solve 
+the Poisson equation for self-gravity and couple the gravitational force to the hydrodynamics.
+Standard techniques for coupling gravitation and rotation forces to the hydrodynamics do 
+not do an adequate job conserving the total energy of the system for our problem, 
 but recent advances in the literature allow progress on this 
-problem and we discuss our implementation here. We present a
-set of test problems demonstrating the extent to which our software can sufficiently
-model a system where large amounts of mass are advected on the computational 
+issue and we discuss our implementation here. We present a
+set of test problems demonstrating the extent to which our software sufficiently
+models a system where large amounts of mass are advected on the computational 
 domain over very long timescales. Future papers in this series will describe
@@
 were not carefully constructed. \cite{motl:2002}, \cite{dsouza:2006},
-and \cite{motl:2007} (the first three-dimensional grid-based
+and \cite{motl:2007} (the first three-dimensional mesh-based
 simulations of mass transfer in binary white dwarf systems) pioneered
-the study of looking at the long-term dynamical evolution of binary
+the study of the long-term dynamical evolution of binary
 white dwarf systems after constructing equilibrium initial
@@
 able to find a detonation in similarly massive systems. Notably,
-the detonation occurred self-consistently and did not need to be inserted 
-by hand. \cite{dan:2012} and \cite{dan:2014} performed a large sweep 
+the detonation occurred self-consistently and did not need to be  
+intentionally triggered using an external source term.
+\cite{dan:2012} and \cite{dan:2014} performed a large sweep 
 of the parameter space for merger pairs and
@@
 used the SPH method. SPH codes have a number of features which do aid
-them in the study of these systems, such as excellent conservation of
-angular momentum. A drawback relates to the fact that whether a prompt detonation
+them in the study of these systems, such as conservation of
+angular momentum to machine precision when there are no source terms
+such as gravity (when gravity is used, conservation is limited by the
+tolerance of error in the gravity solver). A drawback relates to the fact that whether a prompt detonation
 in a merger happens depends in detail on the nature of the
@@
 uniform mass particles, so their effective resolution is
-\textit{lowest} at the stellar surface. In contrast, a grid-based code
+\textit{lowest} at the stellar surface. In contrast, a code
 with adaptive mesh refinement can zoom in on the regions where
@@
 in the high-density regions to adequately capture the large-scale mass
-transfer dynamics. However, there are also outstanding questions of
+transfer dynamics. There are also outstanding questions of
 convergence in SPH (e.g.\ \citealt{zhu_SPH:2014}) and whether the method
@@
 This paper is the first in a series designed to address these
-outstanding theoretical issues for white dwarf mergers. This work will
-discuss the verification of our hydrodynamics code for simulating
+outstanding theoretical issues for white dwarf mergers. This work 
+discusses the verification of our hydrodynamics code for simulating
 these events. Later efforts will look at the initial conditions of the
@@
 
-To study the white dwarf merger problem, we use the grid-based
+To study the white dwarf merger problem, we use the mesh-based
 hydrodynamics code \castro\footnote{\castro\ can be obtained at \url{https://github.com/BoxLib-Codes/Castro}.} \citep{castro}.
@@
 adaptive-mesh refinement (AMR) framework \citep{rendleman:2000}, which
-represents fluid data on a mesh where regions of interest have higher
+represents fluid data on a hierarchical mesh where regions of interest have higher
 spatial resolution. \castro\ is highly parallel and is designed for
@@
 density flux in the ambient zones can be large enough to unphysically
-drag the density below zero. In practice, this will occur at the
+drag the density below zero. In practice, this occurs at the
 trailing edge of a star that is moving across a grid. In such a
@@
 the thermodynamics to be consistent with these values. The 
-velocity of the zone is set to zero. This latter approach will only
-occur in very rare situations, and is mainly there as a safeguard.
+velocity of the zone is set to zero. This latter approach only
+occurs in very rare situations, and is mainly there as a safeguard.
 
@@
 so for example on a grid with three levels and two jumps of four in
-refinement, the level 2 zones will have 16 times higher spatial
-resolution than the coarse grid and there will be 16 level 2 timesteps
+refinement, the level 2 zones have 16 times higher spatial
+resolution than the coarse grid and there are 16 level 2 timesteps
 per level 0 timestep.
@@
 transport (in the flux-limited diffusion approximation). For all simulations 
-in this paper we do not enable either, and will delay discussion of 
+in this paper we do not enable either, and we delay discussion of 
 these modules until later papers in this series.
@@
 \autoref{app:hydro}. Second, after the hydrodynamics step, we add the time-centered source terms
-to the state. First we will describe how we do this for the momentum,
-and then we will describe our approach for the energy. This discussion will be somewhat detailed.
+to the state. First we describe how we do this for the momentum,
+and then we describe our approach for the energy. This discussion is somewhat detailed.
 We believe that the attention is necessary because of the importance of accuracy
@@
 density in terms of the potential and then consider its spatial discretization. For simplicity,
-we will consider one spatial dimension and a uniform discretization. Analogous results 
+we consider one spatial dimension and a uniform discretization. Analogous results 
 may be readily obtained for the non-uniform case.
@@
 It is easy to verify that adding the source terms for the current zone and the two zones 
-to the left and right will result in complete cancellation of the source terms.
+to the left and right results in complete cancellation of the source terms.
 The catch is that if the potential if non-zero outside of the domain, then there will be
@@
 for by a change in gas energy, and that energy changes due to mass transfer are explicitly and 
-exactly tracked. Suppose that we have some fluid mass $\Delta M_{i+1/2} = \Delta \rho_{i+1/2} \Delta V$ leave some zone
+exactly tracked. Suppose that we have some fluid mass $\Delta M_{i+1/2} = \Delta \rho_{i+1/2} \Delta V$ leave the zone
 with index $i$ and enter some zone with index $i+1$. The subscript indicates that the mass change is
@@
 since the gravitational potential is self-generated. \autoref{eq:grav_energy_conservation_update} 
-applies to any conservative potential $\Phi$, and we will use this to our advantage for the 
+applies to any conservative potential $\Phi$, and we use this to our advantage for the 
 rotation forces in \autoref{sec:rotation}.
@@
 and then the energy flux includes a term corresponding to the flux of gravitational potential energy. We 
-avoid this approach for our problem because there will be regions on the computational domain where the total 
+avoid this approach for our problem because there are regions on the computational domain where the total 
 energy is dominated by potential energy (especially the low-density regions near the edge of the white dwarfs),
@@
 Here $(i, j, k)$ are the indices of cells inside the domain, and $(l,m, n)$ 
-are the indices of boundary locations. $\Delta V$ is the volume of the
+are the indices of ghost zones outside the domain where the boundary values for the potential is specified\footnote{In \castro\ we actually
+specify the potential on cell edges, not on cell centers, but the idea is the same, and we use the location of
+the cell edge in computing the distance to each zone in the domain.}. $\Delta V$ is the volume of the
 zone. If there are $N$ zones per spatial dimension, then there are
@@
 solve, the full operation scales only as $N^3$. The amount of time
-required to calculate the boundary conditions will be directly related
+required to calculate the boundary conditions is directly related
 to the chosen value of $l_{\text{max}}$, so there is a trade-off
@@
 As a demonstration of the method's accuracy, we consider the case of two 
-white dwarfs of mass ratio 2/3, loaded onto the grid by the method described below 
+white dwarfs of mass ratio 2/3, using the initialization procedure described below 
 in \autoref{sec:implementation}. We terminated the simulation just after
@@
 we computed the $L^2$ error of $\Phi$ on the entire domain for multipole 
-boundary conditions, which we will call $\Phi_l$, relative to $\Phi$ 
+boundary conditions, which we call $\Phi_l$, relative to $\Phi$ 
 obtained using the exact boundary conditions:
@@
 achieved for the Poisson solver. To do so, we created a simple test
-that inserts a sphere of radius $R$ and uniform mass density $\rho$
+that initializes a sphere of radius $R$ and uniform mass density $\rho$
 onto our grid, and used \castro\ to calculate the gravitational
@@
 
-We can solve the second source of error by changing the density on  
+We can eliminate the second source of error by changing the density on  
 the grid so that the total mass $M$ is actually what we intend it to be.
 The resolution study for this case (the ``normalized sphere'') is also
-plotted in \autoref{fig:gravity_convergence}. At low resolution We still obtain
+plotted in \autoref{fig:gravity_convergence}. At low resolution we still obtain
 convergence slightly better than second-order, indicating that we 
@@
 wish to emphasize that although it is commonly stated in the
-literature that grid-based codes poorly conserve angular momentum,
-it is only generally true that grid-based codes do not exactly conserve 
+literature that fixed-mesh codes poorly conserve angular momentum,
+it is only generally true that mesh-based codes do not exactly conserve 
 angular momentum when the equations are written in conservative form
-for linear momentum. (See also \cite{motl:2002} for an example of how 
+for linear momentum. (See \cite{motl:2002} and \cite{byerly:2014} for examples of how 
 to evolve the hydrodynamics equations for angular momentum.) 
 The extent to which angular momentum conservation is violated
-will be a function of the resolution. When the resolution is sufficiently high, 
+is a function of the resolution. When the resolution is sufficiently high, 
 excellent conservation properties can result, as demonstrated in \autoref{sec:kepler}. 
@@
 
-We apply the rotation forces after the gravitation forces, but 
+We apply the rotational forces after the gravitational forces, but 
 there is some freedom in the order in which to apply the gravitational and rotational terms.
@@
 time-level $n$ and time-level $n+1$. If we update the latter with the gravitational force, 
-then the Coriolis force will see a different velocity than the one obtained through the 
+then the Coriolis force sees a different velocity than the one obtained through the 
 pure hydrodynamics step. (The energy equation does not face the same issue in our new formulation,
 because the velocities used are always the time-level $n+1/2$ values coming from the Riemann solver.)
-In practice, this will not matter significantly for our simulations in this work 
+In practice, this does not matter significantly for our simulations in this work 
 because the centrifugal force plays the dominant role in maintaining stability of non-contact 
@@
 the specified \textit{primary} mass and $M_S$ is the specified
-\textit{secondary} mass. The primary WD will always start on the left
+\textit{secondary} mass. The primary WD always starts on the left
 side of the computational domain for our simulations, and is more
@@
 The frequency of the gravitational waves emitted by the white dwarf binary 
-will be similar to the frequency of the orbital motion, which is in the range 
+is similar to the frequency of the orbital motion, which is in the range 
 10-100 mHz for our problem. This is well outside the range of currently existing 
@@
 many standard hydrodynamics test problems was detailed in the original
-code paper \citep{castro}, and in the interest of brevity we will not
+code paper \citep{castro}, and in the interest of brevity we do not
 repeat them all here. Instead, we focus on a subset of problems that
@@
 should obey simple, expected behaviors. Where possible, these should
-be quantified and a convergence study performed, and that will 
+be quantified and a convergence study performed, and that is
 be the focus of the current section.
@@
 generated with the same equation of state and are as well resolved as
-our finest grid, there will still be a hydrodynamical error associated
+our finest grid, there is still be a hydrodynamical error associated
 with the fact that the rectangular grid cannot faithfully represent a
@@
 
-A simple dynamical test to verify the gravitation physics in \castro\ is
+A simple dynamical test to verify the coupling between the gravity and hydrodynamics in \castro\ is
 the case of gravitational free fall. We place two stars on the grid 
@@
 invariance of the underlying Euler equations, so that simulations
-moving at a uniform bulk velocity will appear different than an
+moving at a uniform bulk velocity appear different than an
 equivalent stationary simulation (e.g. \cite{arepo}). If true, we need to understand 
@@
 Recently, concern for the issue of Galilean invariance has come up in two ways which are of note for us 
-in the present study. We will explain these situations and display 
+in the present study. We explain these situations and display 
 the results of tests we have run to determine whether this 
@@
 fluid in the $y$ direction and $v^\prime$ is the velocity perturbation
-that seeds the instability. The problem will be established for two
+that seeds the instability. The problem is established for two
 sets of initial conditions (ICs), which we follow
@@
 approach the grid resolution. This is prime breeding ground for
-numerical noise. But because the nature of this noise will depend on
-the resolution, it will be very different for simulations at different
+numerical noise. But because the nature of this noise depends on
+the resolution, it is very different for simulations at different
 resolutions. If these instabilities are seeded because of this
@@
 options, controlled by the parameter {\tt castro.grav\_source\_type},
-which we will shortern to {\tt gs} for the present discussion. 
+which we shorten to {\tt gs} for the present discussion. 
 ${\tt gs} = 1$ and ${\tt gs} = 2$ are variations on the standard 
@@
 amount of communication required between hydrodynamics solves. Additionally, 
-the multigrid solver will be less efficient if the boxes are small because there 
-will be fewer available levels for coarsening and performing V-cycles. Conversely, 
+the multigrid solver is less efficient if the boxes are small because there 
+are fewer available levels for coarsening and performing V-cycles. Conversely, 
 boxes that are too large mean that there isn't enough work to go around when we 
@@
 influence the system. Here we have described a number of common problems
-that may creep up, including violation of the conservation of energy, 
+that may occur, including violation of the conservation of energy, 
 a lack of hydrostatic equilibrium (at low resolution) of the stars 
@@
 $d\mathbf{g} / dt = (\mathbf{g}(t_n) - \mathbf{g}(t_{n-1})) / \Delta t$. We perform this
-predictor for the rotation field as well. While both this method and the original
-method have second-order convergence properties, we have seen from testing that
-this new method is slightly more accurate in absolute terms.
+predictor for the rotation field as well.
+
+While both this method and the original method have second-order convergence properties,
+we have seen from testing that this new method is slightly more accurate in absolute terms.
+Finally, we note that this predictor is not applied in cases where we do not have a suitable
+time-level $n-1$ value of the source term, such as in the first timestep of a simulation.
 
@@
 
-From the above discussion it is straightforward to see exactly why the method will not be fully conservative to machine
+From the above discussion it is straightforward to see exactly why the method is not fully conservative to machine
 precision in practice. First, we cannot simulate the domain out to infinity, so Green's second identity does not hold exactly
-and there will be some loss or addition of energy at domain boundaries. Second, \autoref{eq:total_energy_difference} holds in the
+and there is some loss or addition of energy at domain boundaries. Second, \autoref{eq:total_energy_difference} holds in the
 continuum limit by using the Poisson equation, but in practice it is not exactly true that $\rho_i = 4\pi G \nabla^2 \Phi_{i}$ due

### 2015-09-22 papers/wdmerger_I/paper.tex
Updated strategy for tagging

--- 2015-09-19 papers/wdmerger_I/paper.tex
@@
 resolution of 400 km. For the present study, we choose a simple
-refinement strategy: all zones within twice the Roche radius of each
-star are tagged for maximal refinement. The extra buffer from doubling
+refinement strategy: on the coarse grid, all zones within twice the Roche radius of each
+star are tagged for refinement. The extra buffer from doubling
 the size of this region ensures that the sharp density gradients near
-the edge of the star are within the zone of refinement. We also ensure
+the edge of the star are within the zone of refinement. On higher levels, 
+we tag all zones above a given density threshold (taken to be $1\ \text{g cm}^{-3}$ 
+in this paper) that corresponds to the stars themselves. We also ensure
 that the outer part of the domain is never tagged for refinement. In

### 2015-09-22 source/probdata.f90

@@ -90,7 +90,8 @@ module probdata_module
   double precision :: max_co_wd_mass, co_wd_he_shell_mass
 
   ! Tagging criteria
-  double precision :: maxTaggingRadius
+  double precision :: max_tagging_radius
+  double precision :: stellar_density_threshold
 
   ! Stores the center of mass location of the stars throughout the run
   double precision :: com_P(3), com_S(3)
@@ -199,7 +200,7 @@ contains
          max_hybrid_co_wd_mass, hybrid_co_wd_he_shell_mass, &
          max_co_wd_mass, &
          star_axis, initial_motion_dir, &
-         maxTaggingRadius, &
+         max_tagging_radius, stellar_density_threshold, &
          bulk_velx, bulk_vely, bulk_velz, &
          smallx, smallu, &
          center_fracx, center_fracy, center_fracz, &
@@ -250,7 +251,8 @@ contains
     star_axis = 1
     initial_motion_dir = 2
 
-    maxTaggingRadius = 0.75d0
+    max_tagging_radius = 0.75d0
+    stellar_density_threshold = 1.0d0
 
     do_initial_relaxation = .false.
     relaxation_timescale = 0.001

### 2015-09-22 papers/wdmerger_I/paper.tex
Updated scaling results

@@
 excellent weak scaling through 4,096 processors. At the largest run,
-the simulation time required is more than 1.5 times the amount
+the simulation time required is slightly than 1.5 times the amount
 required for the smallest simulation.  This is due entirely to the
@@
     points are normalized to the time per timestep for the smallest number 
-    of processors. The blue circles show the data for a simulation with two AMR
+    of processors. The green circles show the data for a simulation with one 
+    AMR level (a single uniform grid), the blue diamonds show the data for a simulation with two AMR
     levels (one coarse and one fine), while the red circles show the data 
@@
 The strong scaling test we perform uses a grid setup similar to what we 
-use for well-resolved binary simulations. With a single refined level, 
-we have approximately $5 \times 10^7$ zones spread over $\sim 1000$ grids.
-On a second refined level, there are $2 \times 10^9$ zones spread over 
-$\sim 16000$ grids. We run a scaling test for both cases, with the 
+use for well-resolved binary simulations. With only a uniform coarse grid, 
+there are approximately $2 \times 10^7$ zones. With a single refined level, 
+we have approximately $2 \times 10^8$ zones, typically spread over $\sim 2000$ grids.
+On a second refined level, there are a similar number of zones and grids 
+(the volume covered by this level is smaller, which offsets the greater resolution).
+We run a scaling test for all three cases, with the 
 highest processor count in each case chosen so that the number of 
-MPI tasks is similar to the number of grids. There are no relative gains 
+MPI tasks is similar to the number of grids. There are no gains 
 to be achieved from further parallelism. The results are found in 
-\autoref{fig:strong_scaling}. We find excellent scaling for low numbers 
-of processors, though the long wall time per timestep required for the 
-smallest processor counts makes such a simulation infeasible in practice.
-The best tradeoff between reasonable wall time and parallel efficiency 
-seems to be when there are approximately 2--4 grids per processor. The 
+\autoref{fig:strong_scaling}. We find excellent scaling for low to moderate numbers 
+of processors. Parallel efficiency is well maintained when there are 
+at least 2 grids per processor. The 
 scaling behavior worsens at the highest processor counts, but this is 
-an expected consequence of processors becoming work-starved. In general 
+an expected consequence of processors becoming work-starved. At the highest 
+processor count in this test, there is approximately only one grid per processor. In general 
 we find very good strong scaling behavior in the regime we are presently 

### 2015-09-22 science/wdmerger_I/performance/strong_scaling/analyze.py

@@ -4,15 +4,34 @@ from matplotlib import pyplot as plt
 import os
 import wdmerger
 
-small_nproc_arr = np.array([64, 128, 256, 512, 1024, 2048])
-small_time_arr = np.zeros(len(small_nproc_arr))
+one_level_nproc_arr = np.array([32, 64, 128, 256])
+one_level_time_arr = np.zeros(len(one_level_nproc_arr))
 
-large_nproc_arr = np.array([2048, 4096, 8192, 16384])
-large_time_arr = np.zeros(len(large_nproc_arr))
+two_level_nproc_arr = np.array([256, 512, 1024, 2048])
+two_level_time_arr = np.zeros(len(two_level_nproc_arr))
+
+three_level_nproc_arr = np.array([2048, 4096, 8192, 16384])
+three_level_time_arr = np.zeros(len(three_level_nproc_arr))
 
 idx = 0
 
-for nprocs in small_nproc_arr:
+for nprocs in one_level_nproc_arr:
+
+    # Open up the standard output for analysis.
+
+    dir = "results/one_level/n" + str(nprocs)
+
+    output_filename = wdmerger.get_last_output(dir)
+
+    # Get the timing information
+
+    [ one_level_time_arr[idx], _ ] = wdmerger.timing(output_filename)
+
+    idx += 1
+
+idx = 0
+
+for nprocs in two_level_nproc_arr:
 
     # Open up the standard output for analysis.
 
@@ -22,13 +41,13 @@ for nprocs in small_nproc_arr:
 
     # Get the timing information
 
-    [ small_time_arr[idx], _ ] = wdmerger.timing(output_filename)
+    [ two_level_time_arr[idx], _ ] = wdmerger.timing(output_filename)
 
     idx += 1
 
 idx = 0
 
-for nprocs in large_nproc_arr:
+for nprocs in three_level_nproc_arr:
 
     # Open up the standard output for analysis.
 
@@ -38,7 +57,7 @@ for nprocs in large_nproc_arr:
 
     # Get the timing information
 
-    [ large_time_arr[idx], _ ] = wdmerger.timing(output_filename)
+    [ three_level_time_arr[idx], _ ] = wdmerger.timing(output_filename)
 
     idx += 1
 
@@ -46,16 +65,20 @@ eps_filename = 'plots/strong_scaling.eps'
 
 # Divide time array by initial value to get speedup
 
[…]

### 2015-09-23 papers/wdmerger_I/paper.tex
Cite two recent papers

@@
 contact phase, similar to the violent carbon-oxygen
-mergers. \cite{pakmor:2013} added a thin helium shell on their primary
+mergers. \cite{sato:2015} also examined the parameter space and
+came to a similar conclusion for massive carbon-oxygen systems
+(and also looked at the possibility of detonations after the
+coalescence had completed), while \cite{tanikawa:2015} discussed
+the plausibility of helium detonations in the massive binary case.
+\cite{pakmor:2013} added a thin helium shell on their primary
 white dwarf, and found that this robustly led to a detonation of the

### 2015-10-02 documentation/userguide.tex

@@
 
-The \texttt{tests} directory includes problem setups that are used for testing 
-the hydrodynamics code \texttt{CASTRO} and exercises a wide range of its capabilities.
+The \texttt{science} directory includes problem setups that are or have been used in the papers
+that have been or will be submitted based on this software.
+
+The \texttt{tests} directory is designed to hold small versions of problems that can be run in a
+nightly test suite that regularly checks for code errors.
 
@@
 
-The routines that generate both of these are found in \\\texttt{source/diagnostics/sum\_integrated\_quantities.cpp}. 
+The routines that generate both of these are found in \\\texttt{source/sum\_integrated\_quantities.cpp}. 
 This is a replacement of the \texttt{CASTRO} function that does something similar. We use the same 
@@
 At minimum, you'll need a \texttt{Prob\_3d.f90} file (or the equivalent for lower dimensions) and
-corresponding inputs and probin files. The problems in \texttt{tests/castro} are similar to this;
+corresponding inputs and probin files. The problems in \texttt{science/wdmerger_I/castro} are similar to this;
 they are ultimately building using the source code in the CASTRO test suite, but using a
@@
 
-The \texttt{tests} directory contains a number of example problems that we use
+The \texttt{science/wdmerger_I/} directory contains a number of example problems that we use
 to assess the reliability of \texttt{CASTRO} in performing hydrodynamics simulations 

### 2015-12-05 papers/wdmerger_I/paper.tex
Fix inconsistent parentheses
Updated description of the source term predictor, and added mention of our new strategy to sum all the source terms into a single array.

@@
 \begin{equation}
-  (\rho \mathbf{u})^{n+1} \to \frac{(\rho \mathbf{u}^{n+1})}{1 + (\Delta t / \Delta t_S) f_S},
+  (\rho \mathbf{u})^{n+1} \to \frac{(\rho \mathbf{u})^{n+1}}{1 + (\Delta t / \Delta t_S) f_S},
 \end{equation} 
@@
 
-\subsection{Gravity and Rotation Tracing}
+\subsection{Source Term Predictor for the Hydrodynamics}
+\label{sec:source_term_predictor}
+
+In the original release of \castro\ we used the time-level $n$ value of the 
+gravitational and rotation source terms in constructing the edge-states for 
+the hydrodynamics update. While this is formally second-order accurate, there
+is a better choice one can make. We have information about the trend of these
+source terms from previous timesteps, so we can use a predictor method to guess
+at a more accurate value of the gravitational and rotational fields at the
+$n+1/2$ time-level the hydro is evaluated at. Our method uses a lagged linear
+extrapolation. Going into the hydro update, we have both the time-level $n$ and
+time-level $n-1$ data for $g$ (as well as the acceleration due to rotation).
+From this one can construct a simple linear estimator of (say) the gravitational
+acceleration using a backward difference scheme at time-level $n$:
+\begin{equation}
+  \mathbf{g}^{n+1/2} \approx \mathbf{g}^n + \frac{\Delta t_n}{2} \frac{d\mathbf{g}^n}{dt} \approx \mathbf{g}^n + \frac{\Delta t_n}{2}\left(\frac{\mathbf{g}^n - \mathbf{g}^{n-1}}{\Delta t_{n-1}}\right).
+\end{equation}
+
+While both this method and the original method have second-order convergence properties,
+we have seen from testing that this new method is slightly more accurate in absolute terms.
+Finally, we note that this predictor is not applied in cases where we do not have a suitable
+time-level $n-1$ value of the source term, such as in the first timestep of a simulation.
+
+We note also that we have changed slightly how we use source terms in \castro's 
+hydrodynamics. In the original release we would explicitly handle gravity, rotation, 
+and user-defined external source terms separately in constructing the edge states. 
+At present, we sum all of these source terms prior to starting the hydrodynamics 
+update, and use a single source array with data for all components of the state. 
+Consequently, we actually do the source-term predictor shown here on this source 
+term array, rather than individually on each component of the forcing.
+
+\subsection{Source Term Tracing}
 \label{sec:gravity_rotation_tracing}
@@
 $(\Delta t/2)g$ to the interface states for velocity at the end of the
-reconstruction.  In the current implementation, we do the original
-parabolic reconstruction and characteristic tracing of the gravitation
-source. This can be controlled in \castro\ with the parameter 
-{\tt castro.ppm\_trace\_grav}. Our system with the source appears as:
+reconstruction.  In the current implementation, we return to the original
+parabolic reconstruction and characteristic tracing. In fact, as described 
+in \autoref{sec:source_term_predictor}, since we send to the hydro a single 
+source term array that holds the sum of all the source terms (including gravity 
+and rotation), we do the parabolic reconstruction on the full source term data. 
+This can be controlled in \castro\ with the parameter {\tt castro.ppm\_trace\_sources}.
+
+For the following explanation of how the tracing works, we consider only gravity.
+Our system with the source appears as:
 \begin{equation}
@@
 
-Similarly, we also trace under the centrifugal and Coriolis forces to capture
-the effects of rotation on the edge states. For the primitive variables, the rotation
-source term is $F_R = (0, f_R, 0)^T$, where $f_R = -2\, {\bm{\omega}} \times \mathbf{u} - {\bm\omega}\times \left({\bm\omega} \times \mathbf{r}\right).$
-Just like the gravity source, the rotation source terms only update the velocities. Therefore 
-the rotation source terms can be included in a manner identical to that of the gravity source
-terms, with the replacement of $G$ with $F_R$ and $g$ with $f_R$ in the above. This is controlled in \castro\
-with the parameter {\tt castro.ppm\_trace\_rot}.
-
-\subsection{Source Term Predictor for the Hydrodynamics}
-
-In the original release of \castro\ the value of the gravitational and rotation
-source terms used in constructing the edge-states for the hydrodynamics update
-was the time-level $n$ value. While this is formally second-order accurate, there
-is a better choice one can make. We have information about the trend of these
-source terms from previous timesteps, so we can use a predictor method to guess
-at a more accurate value of the gravitational and rotational fields at the
-$n+1/2$ time-level the hydro is evaluated at. Our method uses a lagged linear
-extrapolation. Going into the hydro update, we have both the time-level $n$ and
-time-level $n-1$ data for $g$ and $f_R$ (as defined in \autoref{sec:gravity_rotation_tracing}).
-From this one can construct a simple linear estimator of (say) the gravitational
-acceleration using a backward difference scheme at time-level $n$:
-\begin{equation}
-  \mathbf{g}(t_n + \Delta t) = \mathbf{g}(t_n) + \Delta t \left.\frac{d\mathbf{g}}{dt}\right|_{n} = \frac{3}{2}\mathbf{g}(t_n) - \frac{1}{2}\mathbf{g}(t_{n-1}).
-\end{equation}
-Here we have assumed a constant timestep for simplicity, and constructed the time derivative as
-$d\mathbf{g} / dt = (\mathbf{g}(t_n) - \mathbf{g}(t_{n-1})) / \Delta t$. We perform this
-predictor for the rotation field as well.
-
-While both this method and the original method have second-order convergence properties,
-we have seen from testing that this new method is slightly more accurate in absolute terms.
-Finally, we note that this predictor is not applied in cases where we do not have a suitable
-time-level $n-1$ value of the source term, such as in the first timestep of a simulation.
-
 \section{Proof of Energy Conservation in Simulations using Self-Gravity}

### 2015-12-13 papers/wdmerger_I/paper.tex
Update of the paper with latest numerical results

@@
 cases to demonstrate the simulation behavior: an equal mass case of two $0.9\ \msolar$ 
-white dwarfs, and an unequal mass case of $0.9\ \msolar$ and $0.6\ \msolar$ white dwarfs.
+white dwarfs, and an unequal mass case of $0.9\ \msolar$ and $0.75\ \msolar$ white dwarfs.
+In both systems, the secondary should be stable against mass loss.
 In each case, the initial orbital period is 100 seconds.
@@
 dwarfs on the orbital ($xy$) plane. For the equal mass case in the
-inertial reference frame, the curves essentially fall on top of each
-other, indicating that the stars are indeed orbiting at the initial
-distance without substantial deviation. In the unequal mass case, the
-stars now orbit at different distances but do not much deviate from
-their initial distance. We do observe a secular decline in the center
-of mass of these stars, indicating that eventually the stellar orbits
-will degrade due to numerical error even without intervention. The
-co-rotating frame is different.  For clarity of visualization, we
+inertial reference frame, the curves fall nearly on top of each
+other for most of the run, indicating that the stars are indeed 
+orbiting at the initial distance, at least for a while. Towards the end 
+of the run, however, the orbit starts to decay significantly, and the center-of-mass
+distance of the two stars has decreased by about 10\% after 25 orbits.
+We attribute this to non-conservation of angular momentum, which occurs 
+because our code only explicitly conserves linear momentum. This orbital 
+decay resembles the effect seen by \citet{swc:2000} for the case of neutron stars. 
+In the unequal mass case, the magnitude of the orbital decay is smaller but 
+at the end of the run the secular decline in distance is also visible. In 
+both cases the stars would likely merge due to numerical error 
+after a long enough timescale.
+
+The co-rotating frame is different.  For clarity of visualization, we
 rotate these results back into the inertial frame before displaying
-their orbits.  The stellar orbit in the equal mass case does not show
-the same secular decline, and the system appears to be
-stable. The cost is that the oscillations about this
-equilibrium are slightly larger than in the inertial frame. The
-unequal mass case displays this effect starkly, to the point where the system 
-becomes unstable: while the average distance is not substantially declining, 
-there are huge oscillations about this
-average. In addition, the stars move substantially within the rotating
-frame itself, yielding the chaotic-looking orbits in the figure. The
-oscillations grow with time and will eventually cause the stars to
-merge; however, the stars would eventually merge even in the inertial
-frame case, and at a faster rate than the equal mass case. This
-suggests that this mass ratio is too large to ignore tidal effects,
-and that the tidal effects play out much differently in the rotating
-frame.
+their orbits.  In both the equal and unequal mass cases, the centrifugal force 
+pushes the stars outward toward a new equilibrium distance that is a few 
+percent larger than its initial distance. At the end of the run, the system is 
+relatively stable, with oscillations about the new equilibrium distance. In fact 
+these oscillations occur too in the inertial frame, but they are much more pronounced 
+here. In the unequal mass case this is coupled with severe precession of the orbit, 
+which results in chatoic-looking orbits when viewed from the rotating reference frame. 
+These result from the explicit numerical consideration of the Coriolis and centrifugal 
+terms, which do not appear in the inertial frame. So while the rotating frame 
+is clearly more stable than the inertial frame, the cost is that the specific 
+dynamics may be more suspect.
 
@@
 conservation (\autoref{fig:angular_momentum_conservation_unequal}).
-Angular momentum is conserved to within a few percent over the 25
+Angular momentum is conserved to within 10 percent over the 25
 orbits, while energy conservation is about an order of magnitude
@@
   \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot1.eps}
-  \caption{Absolute magnitude of the relative change in angular momentum of two equal mass white dwarfs after 25 orbital periods, 
+  \caption{Absolute magnitude of the relative change in angular momentum of two unequal mass white dwarfs after 25 orbital periods, 
            evolved in a co-rotating reference frame. We consider only the component of the angular moment along the rotational 
-           axis. The decline and recovery is a change in sign of the angular momentum difference.
+           axis.
            \label{fig:angular_momentum_conservation_unequal}}
@@
 As a simple verification test to ensure our gravitational wave calculations are correct, we plot the 
-gravitational wave strain along the rotation axis for the first two periods of the unequal mass system. 
+gravitational wave strain along the rotation axis for the first two periods of an unequal mass system. 
 At this early time the orbit is circular and so to a good approximation we expect that the gravitational 
@@
   \includegraphics[scale=0.45]{plots/unequal_gw_rot0.eps}
-  \caption{Gravitational wave strain polarizations for the first two orbital periods of the 
+  \caption{Gravitational wave strain polarizations for the first two orbital periods of an 
            unequal mass system. The curves with markers are the numerical data, while the 
@@
 we have achieved convergent behavior. In the rotating frame (not shown here), the results are 
-not quite the same but for these resolutions the results do tend toward convergence.
+not quite the same but for these resolutions the results do tend toward convergence. These 
+results suggest a minimum resolution of 200 km for getting the dynamics right.
 
@@
 a rotating reference frame has its own numerical issues for 
-the type of simulation we desire to perform here. Based on 
-the tests performed here, in some cases the stars may be 
-\textit{more} stable in an inertial frame despite 
-constantly being in motion on the grid. It 
+the type of simulation we desire to perform here. While decay 
+of the stellar orbit is a commonplace feature in the inertial frame 
+but easier to avoid in the rotating frame, the rotation forces 
+can result in likely unphysical oscillations of the stars. It 
 is not easy to predict the correct behavior of such systems,
+and therefore determine which frame is closer to being correct, 
 though one recourse for assessing confidence in a model from a 

### 2015-12-13 papers/wdmerger_I/paper.tex
More careful wording in the rotation section

@@
 angular momentum when the equations are written in conservative form
-for linear momentum. (See \cite{motl:2002} and \cite{byerly:2014} for examples of how 
-to evolve the hydrodynamics equations for angular momentum.) 
-The extent to which angular momentum conservation is violated
+for linear momentum. Indeed, \cite{motl:2002} and \cite{byerly:2014} 
+have evolved binary systems using the hydrodynamics equations written 
+in a form that explicitly conserves angular momentum, and it is 
+straightforward to convert an existing grid-based code to solve 
+the system of equations that \citeauthor{byerly:2014} present.
+Additionally, the extent to which angular momentum conservation is violated in our code
 is a function of the resolution. When the resolution is sufficiently high, 
-excellent conservation properties can result, as demonstrated in \autoref{sec:kepler}. 
-In particular, at reasonable resolution for a binary orbit our code 
-conserves angular momentum well enough that this no longer becomes 
-a strong argument in favor of using a Lagrangian method.
+excellent conservation properties can result. At reasonable resolution 
+for a binary orbit our code conserves angular momentum well enough 
+to keep the stars stable for a large number of orbits; however, at moderate 
+resolution in an inertial frame, there is a secular loss of angular 
+momentum that eventually will result in a spurious merger.
 We note that as the stars begin to coalesce, the rotating reference frame

### 2015-12-15 papers/wdmerger_I/paper.tex
Another pass through with minor corrections

@@
 merited recent theoretical attention is the white dwarf merger scenario.
-This can naturally explain many of the observed characteristics of 
+This has the potential to naturally explain many of the observed characteristics of 
 Type Ia supernovae, as well as their distribution in time and space.
@@
 adaptive mesh refinement. In this paper we describe our numerical 
-methodology and discuss our implementation in the code \castro. \castro\ 
-solves the Euler equations, and for merging white dwarf simulations we use its capability to solve 
-the Poisson equation for self-gravity and couple the gravitational force to the hydrodynamics.
+methodology and discuss our implementation in the compressible hydrodynamics code \castro.
+For merging white dwarf simulations we use \castro's capability to solve the Euler equations, and
+the Poisson equation for self-gravity, and couple the gravitational force to the hydrodynamics.
 Standard techniques for coupling gravitation and rotation forces to the hydrodynamics do 
@@
 uncertain. In particular, it is not clear what process causes the
-temperatures in these white dwarfs to become hot enough for explosive
+temperatures in these white dwarfs (WDs) to become hot enough for explosive
 burning of their constituent nuclei. The model favored initially by the
@@
 recent years the focus has shifted to a number of alternative progenitor models. A
-leading candidate for explaining some or most of these explosions is
+leading candidate for explaining at least some of these explosions is
 the double-degenerate (DD) model, in which two white dwarfs merge and
@@
 disk at a rate high enough to generate the troublesome accretion 
-timescales discussed by \cite{tutukov_yungelson:1979}. Based on this
+timescales discussed by \cite{tutukov_yungelson:1979} and mentioned above. Based on this
 evidence, the review of \cite{hillebrandtniemeyer2000} argued that the
@@
 lead to the conditions necessary for carbon detonation in the center
-of the merged object, but \cite{shen:2012} questioned this for similar
+of the merged object, but \cite{shen:2012} also questioned this for
 reasons related to how viscous transport would convert rotational
@@
 
-The latest developments have resulted in some possible areas of convergence.
+The latest violent merger developments have resulted in some possible areas of convergence.
 \cite{pakmor:2012} performed a merger scenario
@@
 found that pure carbon-oxygen systems would generally not lead to
-detonations and violent mergers except for the most massive
-systems. They did find that for systems containing helium, many
+detonations (and thus be violent mergers) except for the most massive
+systems. They did find that for systems with WDs containing helium, many
 would detonate and potentially lead to SNe Ia, either through the
 aforementioned instabilities in the accretion stream, or during the
-contact phase, similar to the violent carbon-oxygen
+contact phase, similar to the violent carbon-oxygen WD
 mergers. \cite{sato:2015} also examined the parameter space and
-came to a similar conclusion for massive carbon-oxygen systems
+came to a similar conclusion for massive carbon-oxygen WD systems
 (and also looked at the possibility of detonations after the
@@
 angular momentum to machine precision when there are no source terms
-such as gravity (when gravity is used, conservation is limited by the
-tolerance of error in the gravity solver). A drawback relates to the fact that whether a prompt detonation
+such as gravity (and conservation proportional to the level of
+tolerance of error in the gravity solver when gravity is used).
+A drawback relates to the fact that whether a prompt detonation
 in a merger happens depends in detail on the nature of the
 gas at the interface between the two stars, which is at much lower
-density than the rest of the stellar material. The SPH codes use
+density than the rest of the stellar material. The SPH codes for these
+simulations generally all use
 uniform mass particles, so their effective resolution is
@@
 \cite{colella_sekora:2008}, which can be used in the code by setting 
-\texttt{castro.ppm\_type = 2} in the inputs file, which is a set of code parameters
-accessed at runtime to determine the algorithms used in the simulation.
-The advantage of this limiter is that
+\texttt{castro.ppm\_type = 2} in the inputs file (the inputs file is
+a set of code parameters accessed at runtime to determine the algorithms
+used in the simulation). The advantage of this limiter is that
 it preserves physical extrema rather than clipping them off as in the
@@
 affected zones is typically already fairly low, this should not
-seriously affect the energy conservation properties of our
+seriously affect the dynamics or the energy conservation properties of our
 simulation. Our strategy for a reset is as follows: when the density of 
@@
 (density, momentum, energy, and temperature) of the
-reset zone to be exactly equal to the field values of this 
+reset zone to be equal to the field values of this 
 adjacent zone. If no adjacent zone reaches the density floor, then
@@
 velocity of the zone is set to zero. This latter approach only
-occurs in very rare situations, and is mainly there as a safeguard.
+occurs in very rare situations, and is there as a last resort.
 
@@
 zones are strictly contained within the rectangular extent of the
-underlying coarser zones (the code is not restricted to representing
-only Cartesian geometries, but we use a Cartesian mesh with uniform
-spacing in each dimension for the present study). For the time
+underlying coarser zones (at present, in 3D the code is restricted to representing
+only Cartesian geometries with uniform spacing in each dimension). For the time
 evolution of the AMR system we use subcycling, where each AMR level is
@@
 that is consistent with the returned temperature. These inconsistencies
-are dependent on the threshold we set that determines how close the 
-returned energy should be to the original input energy. While this 
-threshold is typically very small in an individual zone (a relative 
+are dependent on the tolerance of the Newton-Raphson iterative solve.
+While this error tolerance is typically very small in an individual zone (a relative 
 difference of $10^{-8}$ by default in \castro), over time and given 
@@
 
-\castro\ has the ability to do both nuclear reactions and radiative 
+\castro\ has the ability to model both nuclear reactions and radiative 
 transport (in the flux-limited diffusion approximation). For all simulations 
@@
 consists of both level and composite solves, and (optionally) a final
-synchronization at the end.
+synchronization at the end. We do not enable this final synchronization
+for the merger simulations, because the grid boundaries never lie in
+regions of high density, so the change in the potential due to the correction
+at coarse--fine interface is always negligible.
 
@@
 equations. In a finite volume methodology, the momentum source term 
-as traditionally written appears in integral form as
+often appears in integral form as
 \begin{equation}
@@
 source term evaluated at time-level $n$. We choose a more accurate approach, 
-which is also second order, of tracing
+which is also second order, of characteristic tracing
 under the source term; the details of this are described in 
@@
 exactly tracked. Suppose that we have some fluid mass $\Delta M_{i+1/2} = \Delta \rho_{i+1/2} \Delta V$ leave the zone
-with index $i$ and enter some zone with index $i+1$. The subscript indicates that the mass change is
+with index $i$ and enter the zone with index $i+1$. The subscript indicates that the mass change is
 occurring at the interface between the two zones, at index $i+1/2$. The work done by the gravitational
@@
 
-Now, in a hydrodynamics code mass changes correspond to hydrodynamic fluxes. In particular,
+Now, in a hydrodynamics code, mass changes correspond to hydrodynamic fluxes. In particular,
 the continuity equation tells us that the mass flux $F_\rho = \rho^{n+1/2}_{i+1/2} v^{n+1/2}_{i+1/2}$ yields
@@
 on cell faces. This was demonstrated empirically by \cite{jiang:2013} and is obvious in the case of a
-fixed external potential; it is not as obvious in the case of the gravitational potential, which
-changes in response to the mass distribution, so we give a short proof of this in \autoref{app:gravity}.
-However, in practice there is a non-zero numerical tolerance associated 
+fixed external potential; it is not as obvious in the case of the gravitational self-potential, which
+changes in response to changes in the mass distribution, so we give a short proof of this in
+\autoref{app:gravity}. However, in practice there is a non-zero numerical tolerance associated 
 with the Poisson gravity solver (in our case, the multigrid method) that results in a non-zero error 
@@
 
-In passing, the factor of $1/2$ that appears in \autoref{eq:grav_energy_conservation_update} 
+In passing, we hope to clear up a spot of potential confusion, that we feel is unclear in other papers
+on this subject: the factor of $1/2$ that appears in \autoref{eq:grav_energy_conservation_update} 
 has nothing to do with the factor of $1/2$ that appears in the statement of conservation of total energy, 
@@
 
-As observed by \cite{arepo}, this is more accurate than the more common (non-conservative) approach
+As observed by \cite{arepo}, this method is more accurate than the more common (non-conservative) approach
 of evaluating the change in gas energy using the work done $(\mathbf{v} \cdot \rho \mathbf{g})$
@@
 leaking---evaluating the energy transfer at the six zone faces instead of the single zone
-center---has a very minor cost in terms of code complexity and computational time.
+center---adds only a very minor cost in terms of code complexity and computational time.
 
@@
 also a similar result by \citet{hummer:1996}, and an earlier calculation 
-by \citet{macmillan:1958}). The potential is in
+by \citet{macmillan:1958}). The potential is given in
 Equation 15 of that paper\footnote{The last term in that equation is missing a factor of
@@
 and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. We note that if implemented in 
-\texttt{Fortran} or \texttt{C}, the inverse hyperbolic tangent used here is
+\texttt{Fortran} or \texttt{C}/\texttt{C++}, the inverse hyperbolic tangent used here is
 \texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not}
@@
 In a rotating reference frame with angular frequency vector $\bm{\omega}$,
-non-inertial contribution to the momentum equation is:
+the non-inertial contribution to the momentum equation is:
 \begin{equation}
@@
 uniquely determined by the density in this case. Here, $\rho_i$ and $p_i$
-are known, $g_{i+1/2}$ is the gravitational acceleration at the
+are known, and $g_{i+1/2}$ is the gravitational acceleration at the
 interface between zones $i$ and $i+1$, found by simply adding up all
@@
 resolution runs, this is a better choice than selecting the 1D grid 
-spacing to be equal to the 3D grid spacing.
+spacing to be comparable to the 3D grid spacing.
 
@@
 the donor. The center of mass is located at the center of the
-computational domain, and the stars lie along the $x$ axis, so that
+computational domain, and by default the stars lie along the $x$ axis, so that
 the primary's center of mass is located at $x = -(M_S / M)\, a$ and
 the secondary's center of mass is located at $x = (M_P / M)\, a$.
+The user may choose to initialize the stars along a different axis,
+and can also choose a non-zero orbital phase and/or eccentricity.
 
@@
 refinement strategy: on the coarse grid, all zones within twice the Roche radius of each
-star are tagged for refinement. The extra buffer from doubling
-the size of this region ensures that the sharp density gradients near
+star are tagged for refinement, using the formula provided by 
+\citet{eggleton:1983} for the effective Roche radius $r_L$ of a star in a binary,
+\begin{equation}
+  \frac{r_L}{a} = \frac{0.49 q^{2/3}}{0.6^{2/3} + \text{ln}(1 + q^{1/3})}.
+\end{equation}
+In this formula we can use $q = M_S / M_P$ for obtaining the Roche radius of the secondary,
+and use the inverse value of $q$ to obtain the Roche radius of the primary.
+The extra buffer from doubling the Roche radius ensures that the sharp density gradients near
 the edge of the star are within the zone of refinement. On higher levels, 
@@
 Outside of the stars we fill the rest of the domain with a very low density 
-ambient gas because our hydrodynamics representation requires the density to be 
+ambient gas because our hydrodynamics model requires the density to be 
 non-zero everywhere. This ambient material can create difficulties for the simulation.
@@
 distance from the origin to the domain boundaries, and $\Delta r_S$ so
-that the sponge extends another 10\% of that distance. The resulting profile 
-is displayed in \autoref{fig:sponge}. We set $\Delta
+that the sponge smoothing region extends another 10\% of that distance.
+The resulting profile is displayed in \autoref{fig:sponge}. We set $\Delta
 t_S = 0.01$ s, which is of the same order as the CFL timestep
-for typical problem setups. While the sponge is applied we should be cautious of imputing any physical 
+for typical problem setups. While the sponge is applied we should avoid imputing any physical 
 meaning to what is happening in the low-density gas far from the stars.
@@
 
-We track a number of diagnostic quantities at the end of every coarse grid timestep. 
+We track a number of diagnostic quantities at the end of coarse grid timesteps. 
 For all simulations, we record the total energy (including the breakdown into
-its components: kinetic, internal, gravitational potential, rotation; we note
+its components: kinetic, internal, gravitational potential, and rotation; we note
 that for the diagnostics we actually use $(\rho E)$ for calculation of the total energy,
@@
 the total angular momentum, and the center of mass of the system. 
-For stellar calculations, we separately record diagnostic 
+We also separately record diagnostic 
 information about the stars. Our strategy for tracking their 
@@
 To do this, we need a cutoff for determining what counts as part of the primary 
-and what counts as part of the secondary. For this we can use the formula provided by 
-\citet{eggleton:1983} for the effective Roche radius $r_L$ of a star in a binary,
-\begin{equation}
-  \frac{r_L}{a} = \frac{0.49 q^{2/3}}{0.6^{2/3} + \text{ln}(1 + q^{1/3})}.
-\end{equation}
-In this formula we can use $q = M_S / M_P$ for obtaining the Roche radius of the secondary,
-and use the inverse value of $q$ to obtain the Roche radius of the primary.
-To a reasonable approximation, all material within the effective Roche radius
-of each star can be counted as belonging to that star. From this we obtain the mass
+and what counts as part of the secondary. We use a simple criterion: the star
+that a zone ``belongs'' to is the one that exerts a larger magnitude
+gravitational force on that zone (as computed using the tentative data
+for that star's mass and radius). From this we obtain the corrected mass
 of each star as well as its location and velocity. Once we have the new centers of mass,
 we compute the effective radius of each star at various density cutoffs. This involves 
-computing the volume $V$ of all zones within the Roche radius of the star's center 
+computing the volume $V$ of all zones that belong to the star (in the sense described above) 
 whose density is greater than the cutoff. We then compute $r_{\text{eff}} = (3V/4\pi)^{1/3}.$
@@
 \begin{equation}
-  Q_{kl} = \int dV \rho \left(x^k x^l - \frac{1}{3}\delta^{kl} \mathbf{x}^2\right).
+  Q_{kl} = \int dV \rho \left(x_k x_l - \frac{1}{3}\delta_{kl} \mathbf{x}^2\right).
 \end{equation}
@@
 \begin{equation}
-  \ddot{Q}_{kl} = \text{STF}\left\{2\int dV \rho (v^k v^l + x^k g^l)\right\}.
+  \ddot{Q}_{kl} = \text{STF}\left\{2\int dV \rho (v_k v_l + x_k g_l)\right\}.
 \end{equation}
@@
 \citep{zingale:2002}. Although we ensure that the initial models are
-generated with the same equation of state and are as well resolved as
+generated with the same equation of state and are at least as well resolved as
 our finest grid, there is still be a hydrodynamical error associated
@@
 \end{equation}
-For our problem $0 \leq r \leq 1$, so this is always valid. Integrating, we find
+For our problem $0 < r \leq 1$, so this is always valid. Integrating, we find
 \begin{equation}
@@
 \citet{robertson:2010} (hereafter, R10) observe that violation of Galilean
-invariance of simulation results for the Euler equations occurs only
+invariance of simulation results for the Euler equations occurs 
 because of truncation error in the discretization of the fluid
@@
 emphasize the point that this does not objectively demonstrate a
-defiency in fixed-grid codes for this problem. We can only determine
+deficiency in fixed-grid codes for this problem. We can only determine
 the validity of a method when we have a trustworthy, converged
@@
 times. As observed by \citeauthor{mcnally:2012}, this lack of a solution is because the
-secondary instabilities form for this problem when the whirls of the
+secondary instabilities form for this problem when the whorls of the
 Kelvin-Helmholtz tendrils stretch out and create gradients that
@@
 on numerical calculations of shear mixing in fixed-grid codes, but that this effect 
-diminishes with increasing resolution. As a result, we need only be confident that we are 
-sufficiently resolving the major mixing regions on the white dwarf surfaces. 
-If we find that this mixing occurs near the grid resolution scale, this will imply 
-that we need to ramp up the resolution in these regions using AMR. If this becomes 
-too expensive, we would need to be skeptical of any conclusions that could be drawn 
+diminishes with increasing resolution. As a result, we must be confident that we are 
+sufficiently resolving the major mixing regions on the white dwarf surfaces,
+specifically that the density gradients occur over spatial scales much larger
+than the grid resolution. If we find instead that this mixing occurs near the
+grid resolution scale, this will imply that we need to ramp up the resolution
+in these regions using AMR. If this becomes too expensive, we would need to be
+skeptical of any conclusions that could be drawn 
 about the effect of the mixing on the nuclear burning.
@@
 problem at the bulk velocity, so that the multipole moments were
-computed with respect to the stellar center.
+always computed with respect to the current location of the stellar center.
 
@@
 in the inertial frame. We see that the choice of rotational coupling is much less important than the choice of gravity coupling. 
-Interestingly, the ``conservative'' ${\tt rs} = 4$ is slightly worse in energy conservation than the non-conservative, 
-cell-centered ${\tt rs} = 2$ algorithm. This may be because the rotational force grows with distance, and so losses at the 
-domain boundaries may be enhanced by explicitly evaluating the flux at cell edges.
+The ``conservative'' ${\tt rs} = 4$ is slightly better in energy conservation than the non-conservative, 
+cell-centered ${\tt rs} = 2$ algorithm, but it is a small effect.
 
@@
 terms, which do not appear in the inertial frame. So while the rotating frame 
-is clearly more stable than the inertial frame, the cost is that the specific 
-dynamics may be more suspect.
+is clearly more stable against mass transfer than the inertial frame,
+the cost is that the specific dynamics may be more suspect.
 
@@
 conservation (\autoref{fig:angular_momentum_conservation_unequal}).
-Angular momentum is conserved to within 10 percent over the 25
+For the former system Angular momentum is conserved to within 10 percent over the 25
 orbits, while energy conservation is about an order of magnitude
@@
 number of processor cores. We are currently developing an approach to
-evaluating the equation of state and nuclear reactions (and possibly
-the entire hydrodynamics algorithm) on GPUs, which will allow us to
+evaluating the hydrodynamics and microphysics modules on GPUs,
+which will allow us to
 take advantage of the significant computational resources embedded in
@@
   \includegraphics[scale=0.4]{plots/weak_scaling}
-  \caption{\castro\ weak scaling test, performed on the Blue Waters machine at 
+  \caption{\castro\ weak scaling test, performed on Blue Waters at 
     NCSA. Each processor had a fixed amount of work, and we increased the 
@@
 contributors to the rate of Type Ia supernovae and related astronomical 
-transients, the theoretical view on these systems is far from complete. 
+transients, the theoretical view of these systems is far from complete. 
 Studying these systems over the long timescales relevant to dynamical 

### 2015-12-15 papers/wdmerger_I/paper.tex
Acknowledge Stony Brook IACS for its support

@@
 
-This research has made use of the AstroBetter blog and wiki. 
-This research has made use of NASA's Astrophysics Data System 
-Bibliographic Services.
-
 This research was supported by NSF award AST-1211563. An
@@
 U.S. Department of Energy under Contract No. DE-AC02-05CH11231.
+Additional support and computer time was provided by the Stony Brook
+University Institute for Advanced Computational Science, in particular
+through approval to run on its LI-RED cluster.
 This work used the Extreme Science and Engineering Discovery Environment (XSEDE), 
@@
 Project AST100037 supported use of the resources NICS/Kraken and NICS/Darter.
+
+This research has made use of NASA's Astrophysics Data System 
+Bibliographic Services. In addition, this research has made use
+of the AstroBetter blog and wiki.
 

### 2015-12-15 papers/wdmerger_I/paper.tex
Fix IACS acknowledgement
Add affiliations

--- 2015-12-15 papers/wdmerger_I/paper.tex
@@
 
-\author{Max P. Katz, Michael Zingale, Alan C. Calder, F. Douglas Swesty, Ann S. Almgren, Weiqun Zhang}
+\author{Max P. Katz\altaffilmark{1}}
+\author{Michael Zingale\altaffilmark{1}}
+\author{Alan C. Calder\altaffilmark{1,2}}
+\author{F. Douglas Swesty\altaffilmark{1}}
+\author{Ann S. Almgren\altaffilmark{3}}
+\author{Weiqun Zhang\altaffilmark{3}}
+
+\altaffiltext{1}
+{
+  Department of Physics and Astronomy,
+  Stony Brook University, Stony Brook, NY, 11794-3800, USA
+}
+
+\altaffiltext{2}
+{
+  Institute for Advanced Computational Sciences,
+  Stony Brook University, Stony Brook, NY, 11794-5250, USA
+}
+
+\altaffiltext{3}
+{
+  Center for Computational Sciences and Engineering,
+  Lawrence Berkeley National Laboratory, Berkeley, CA 94720
+}
+
 %==========================================================================
@@
 U.S. Department of Energy under Contract No. DE-AC02-05CH11231.
+Results in this paper were obtained using the high-performance
+computing system at the Institute for Advanced Computational Science
+at Stony Brook University.
 This work used the Extreme Science and Engineering Discovery Environment (XSEDE), 

### 2015-12-16 papers/wdmerger_I/paper.tex
Fix IACS acknowledgement again

@@
 Results in this paper were obtained using the high-performance
-computing system at the Institute for Advanced Computational Science
-at Stony Brook University.
+LIred computing system at the Institute for Advanced Computational
+Science at Stony Brook University, which was obtained through
+the Empire State Development grant NYS \#28451. 
 This work used the Extreme Science and Engineering Discovery Environment (XSEDE), 

### 2015-12-17 papers/wdmerger_I/paper.tex
Remove underscores in reference names, and also remove .eps extensions from some figures

@@
 events arise from the thermonuclear explosions of white dwarfs
-\citep{hoyle_fowler:1960}, but the cause of these explosions is
+\citep{hoyle-fowler:1960}, but the cause of these explosions is
 uncertain. In particular, it is not clear what process causes the
@@
 community was the single-degenerate (SD) model
-\citep{whelan_iben:1973}. Accretion of material from a companion star
+\citep{whelan-iben:1973}. Accretion of material from a companion star
 such as a red giant would cause the star to approach the Chandrasekhar
@@
 for a discussion of the possible implications of the accretion-induced 
-collapse scenario). \cite{tutukov_yungelson:1979}
+collapse scenario). \cite{tutukov-yungelson:1979}
 observed that the collapse could be avoided if the mass loss from the secondary
@@
 followed later by a number of authors 
-\citep{rasio_shapiro:1995,segretain:1997,guerrero:2004,yoon:2007,loren-aguilar:2009,raskin:2012}.
+\citep{rasio-shapiro:1995,segretain:1997,guerrero:2004,yoon:2007,loren-aguilar:2009,raskin:2012}.
 The main finding of these early 3D SPH simulations was that if the 
@@
 envelope. Carbon fusion might commence in the disk, but not at a 
-high enough rate to generate a nuclear detonation. \cite{mochkovitch_livio:1990} 
+high enough rate to generate a nuclear detonation. \cite{mochkovitch-livio:1990} 
 and \cite{livio:2000}  also observed that turbulent viscosity in this disk 
@@
 disk at a rate high enough to generate the troublesome accretion 
-timescales discussed by \cite{tutukov_yungelson:1979} and mentioned above. Based on this
+timescales discussed by \cite{tutukov-yungelson:1979} and mentioned above. Based on this
 evidence, the review of \cite{hillebrandtniemeyer2000} argued that the
@@
 transfer dynamics. There are also outstanding questions of
-convergence in SPH (e.g.\ \citealt{zhu_SPH:2014}) and whether the method
+convergence in SPH (e.g.\ \citealt{zhu-SPH:2014}) and whether the method
 correctly captures fluid instabilities. This is an important question
@@
 version of the higher resolution limiters of
-\cite{colella_sekora:2008}, which can be used in the code by setting 
+\cite{colella-sekora:2008}, which can be used in the code by setting 
 \texttt{castro.ppm\_type = 2} in the inputs file (the inputs file is
@@
 The equation of state (EOS) for our simulations is the Helmholtz EOS
-\citep{timmes_swesty:2000}. This models an electron-positron gas of
+\citep{timmes-swesty:2000}. This models an electron-positron gas of
 arbitrary relativity and degeneracy over a wide range of temperatures
@@
   \centering
-  \includegraphics[scale=0.45]{plots/equal_energy_rot0.eps}
+  \includegraphics[scale=0.45]{plots/equal_energy_rot0}
   \caption{Absolute magnitude of the relative change in energy of two equal mass white dwarfs through 25 orbital periods,
@@
   \centering
-  \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot1.eps}
+  \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot1}
   \caption{Absolute magnitude of the relative change in angular momentum of two unequal mass white dwarfs after 25 orbital periods, 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/unequal_gw_rot0.eps}
+  \includegraphics[scale=0.45]{plots/unequal_gw_rot0}
   \caption{Gravitational wave strain polarizations for the first two orbital periods of an 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/spatial_convergence_rot0.eps}
+  \includegraphics[scale=0.45]{plots/spatial_convergence_rot0}
   \caption{Distance between the two white dwarfs in the unequal mass system, for the first 

### 2015-12-17 papers/wdmerger_I/paper.tex
dx -> delta x
Shorten abstract to fit within 250 word ApJ limit

@@
 \begin{abstract}
+
 The Type Ia supernova progenitor problem is one of the most perplexing and
 exciting problems in astrophysics, requiring detailed numerical modeling to
-complement the observations of these explosions. One possible progenitor that has 
-merited recent theoretical attention is the white dwarf merger scenario.
-This has the potential to naturally explain many of the observed characteristics of 
-Type Ia supernovae, as well as their distribution in time and space.
-To date there have been relatively few fully self-consistent simulations 
-of merging binary white dwarf systems using mesh-based hydrodynamics. 
-This is the first paper in a series designed to 
-describe simulations of these systems using a mesh-based hydrodynamics code with 
-adaptive mesh refinement. In this paper we describe our numerical 
-methodology and discuss our implementation in the compressible hydrodynamics code \castro.
-For merging white dwarf simulations we use \castro's capability to solve the Euler equations, and
-the Poisson equation for self-gravity, and couple the gravitational force to the hydrodynamics.
-Standard techniques for coupling gravitation and rotation forces to the hydrodynamics do 
-not do an adequate job conserving the total energy of the system for our problem, 
-but recent advances in the literature allow progress on this 
-issue and we discuss our implementation here. We present a
-set of test problems demonstrating the extent to which our software sufficiently
-models a system where large amounts of mass are advected on the computational 
-domain over very long timescales. Future papers in this series will describe
-our treatment of the initial conditions of these systems and will 
-examine the early phases of the merger to determine its viability
-for triggering a thermonuclear detonation.
+complement observations of these explosions. One possible progenitor that has
+merited recent theoretical attention is the white dwarf merger scenario, which
+has the potential to naturally explain many of the observed characteristics of
+Type Ia supernovae. To date there have been relatively few self-consistent simulations
+of merging white dwarf systems using mesh-based hydrodynamics. This is the first
+paper in a series describing simulations of these systems using a hydrodynamics
+code with adaptive mesh refinement. In this paper we describe our numerical
+methodology and discuss our implementation in the compressible hydrodynamics
+code CASTRO, which solves the Euler equations, and the Poisson equation for
+self-gravity, and couples the gravitational and rotation forces to the hydrodynamics.
+Standard techniques for coupling gravitation and rotation forces to the hydrodynamics
+do not adequately conserve the total energy of the system for our problem, but
+recent advances in the literature allow progress and we discuss our implementation
+here. We present a set of test problems demonstrating the extent to which our
+software sufficiently models a system where large amounts of mass are advected
+on the computational domain over long timescales. Future papers in this series
+will describe our treatment of the initial conditions of these systems and will
+examine the early phases of the merger to determine its viability for triggering
+a thermonuclear detonation.
 
@@
 \begin{equation}
-  \| f \|_2 = \left(\sum_{i,j,k} dx\, dy\, dz\, f_{ijk}^2\right)^{1/2},
+  \| f \|_2 = \left(\sum_{i,j,k} \Delta x\, \Delta y\, \Delta z\, f_{ijk}^2\right)^{1/2},
 \end{equation}

### 2015-12-17 papers/wdmerger_I/paper.tex
Rearrange figure placement

@@
 we choose a relative error tolerance of $10^{-10}$ to be satisfied in the multigrid solve.
+The results of this test are plotted in \autoref{fig:gravity_convergence}.
+
+We find that at low resolution convergence is actually substantially better 
+than second-order. The explanation for this is that we are attempting to 
+model a spherical object on a rectangular grid. This results in two sources of error.
+First, at very low resolution, the object does not look very spherical due to the rectangular 
+grid representation, so the potential it produces is not quite that of a sphere. 
+As the resolution is increased, the distribution of the mass on the grid will change.
 \begin{figure}[h]
@@
 \end{figure}
-The results of this test are plotted in \autoref{fig:gravity_convergence}.
-
-We find that at low resolution convergence is actually substantially better 
-than second-order. The explanation for this is that we are attempting to 
-model a spherical object on a rectangular grid. This results in two sources of error.
-First, at very low resolution, the object does not look very spherical due to the rectangular 
-grid representation, so the potential it produces is not quite that of a sphere. 
-As the resolution is increased, the distribution of the mass on the grid will change.
 Second, the total amount of mass on the grid will change as the sphere fills out. 
@@
 mass than at $256$ zones per dimension.
-
 We can eliminate the second source of error by changing the density on  

### 2015-12-17 papers/submit.sh
Script that will create a directory and corresponding tar file that is ready for ApJ submission

@@ -1,0 +1,31 @@
+submit_dir="submit"
+
+# Build the PDF so that we have the bibilography
+
+make
+
+# Move all the source files to the submit directory
+
+mkdir -p submit/
+
+cp -f paper.tex $submit_dir
+cp -f paper.bbl $submit_dir
+cp -f plots/*.eps $submit_dir
+cp -f plots/*.table $submit_dir
+cp -f ../apj.bst $submit_dir
+cp -f ../emulateapj.cls $submit_dir
+
+# Get rid of all references to the plots subdirectory
+
+sed -i "s/plots\///" $submit_dir/paper.tex
+
+# Get rid of all references to the parent directory
+
+sed -i "s/\.\.\///" $submit_dir/paper.tex
+
+# Replace the bibliography command with a direct input
+# of the bibliography file
+
+sed -i "s/bibliography{refs}/input{paper.bbl}/" $submit_dir/paper.tex
+
+tar -czvf submit.tar $submit_dir

### 2015-12-17 papers/wdmerger_I/submit/paper.tex
Add a copy of the files that were submitted to ApJ for paper 1

--- 2015-12-17 papers/wdmerger_I/paper.tex
@@
-\documentclass[iop,numberedappendix]{../emulateapj}
+\documentclass[iop,numberedappendix]{emulateapj}
 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/bc_comparison}
+  \includegraphics[scale=0.45]{bc_comparison}
   \caption{Error of $\Phi$ on the computational domain for a binary white dwarf simulation 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/phi_comparison}
+  \includegraphics[scale=0.45]{phi_comparison}
   \caption{Comparison of the \castro\ gravitational potential to the analytical solution for: 
@@
   \centering
-  \includegraphics[scale=0.4]{plots/sponge}
+  \includegraphics[scale=0.4]{sponge}
   \caption{Radial profile of the hydrodynamical sponge we apply (\autoref{eq:sponge_frac}). 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/single_star_static_1e3_radius}
+  \includegraphics[scale=0.45]{single_star_static_1e3_radius}
   \caption{Time evolution of the effective radius of a $0.9 \msolar$ 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/single_star_static_ke}
+  \includegraphics[scale=0.45]{single_star_static_ke}
   \caption{Time evolution of the kinetic energy of a $0.9\, \msolar$ 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/freefall}
+  \includegraphics[scale=0.45]{freefall}
   \caption{Time evolution of two initially stationary white dwarfs,
@@
   \centering
-  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p2_low_res_collage}}}
+  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{kh_t2.0_p2_low_res_collage}}}
   \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
@@
   \centering
-  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p3_low_res_collage}}}
+  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{kh_t2.0_p3_low_res_collage}}}
   \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
@@
 \begin{figure*}[ht]
-  \includegraphics[trim = 0in 1.25in 0.25in 1.25in, clip, scale=0.65]{{{plots/kh_p3_high_res_collage}}}
+  \includegraphics[trim = 0in 1.25in 0.25in 1.25in, clip, scale=0.65]{{{kh_p3_high_res_collage}}}
   \caption{Time series of the Kelvin-Helmholtz problem proposed by McNally et al. (2012)
@@
   \centering
-  \includegraphics[scale=0.45]{plots/single_star_compare_1e3_radius}
+  \includegraphics[scale=0.45]{single_star_compare_1e3_radius}
   \caption{A variation on \autoref{fig:single_star_static_radius} where
@@
   \centering
-  \includegraphics[scale=0.45]{plots/gravity_bcs}
+  \includegraphics[scale=0.45]{gravity_bcs}
   \caption{Absolute magnitude of the relative change in the distance of two unequal mass white dwarfs after one orbital period. 
@@
 \capstartfalse
-\input{plots/sources.table}
+\input{sources.table}
 \capstarttrue
@@
   \centering
-  \includegraphics[scale=0.7]{plots/circular_orbit_comparison}
+  \includegraphics[scale=0.7]{circular_orbit_comparison}
   \caption{Positions of the white dwarfs in the orbital plane for four
@@
   \centering
-  \includegraphics[scale=0.45]{plots/equal_energy_rot0}
+  \includegraphics[scale=0.45]{equal_energy_rot0}
   \caption{Absolute magnitude of the relative change in energy of two equal mass white dwarfs through 25 orbital periods,
@@
   \centering
-  \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot1}
+  \includegraphics[scale=0.45]{unequal_angular_momentum_rot1}
   \caption{Absolute magnitude of the relative change in angular momentum of two unequal mass white dwarfs after 25 orbital periods, 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/unequal_gw_rot0}
+  \includegraphics[scale=0.45]{unequal_gw_rot0}
   \caption{Gravitational wave strain polarizations for the first two orbital periods of an 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/spatial_convergence_rot0}
+  \includegraphics[scale=0.45]{spatial_convergence_rot0}
   \caption{Distance between the two white dwarfs in the unequal mass system, for the first 
@@
   \centering
-  \includegraphics[scale=0.4]{plots/weak_scaling}
+  \includegraphics[scale=0.4]{weak_scaling}
   \caption{\castro\ weak scaling test, performed on Blue Waters at 
@@
   \centering
-  \includegraphics[scale=0.4]{plots/strong_scaling}
+  \includegraphics[scale=0.4]{strong_scaling}
   \caption{\castro\ strong scaling test performed on the Blue Waters machine at
@@
 
-\bibliographystyle{../apj}
-\bibliography{../refs}
+\bibliographystyle{apj}
+\input{paper.bbl}
 

### 2015-12-22 papers/wdmerger_II/paper.tex
Sync up some things in paper II with the layout in paper I

@@
-\documentclass[iop]{../emulateapj}
-% for \sout
-\usepackage{ulem}
-% makes sure \em{} is italic rather than underlined (corrects ulem from line above)
-\normalem
+\documentclass[iop,numberedappendix]{../emulateapj}
+
+% these lines seem necessary for pdflatex to get the paper size right
+\pdfpagewidth 8.5in
+\pdfpageheight 11.0in
 
@@
 
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
+% allows Greek symbols to be bold
+\usepackage{bm}
+
+% allows us to force the location of a figure
+\usepackage{float}
+
+% allows comment sections
+\usepackage{verbatim}
+
+% allows in-document links; see http://www.astrobetter.com/blog/2014/09/29/latex-hyperref-and-emulateapj/
+\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref} 
+\usepackage[all]{hypcap} %Links go to figures;breaks on deluxetables       
+\renewcommand*{\backref}[1]{[#1]}
+
+% Override choices in \autoref
+\def\sectionautorefname{Section}
+\def\subsectionautorefname{Section}
+\def\subsubsectionautorefname{Section}
 
@@
 
-\newcommand{\evm}{{(-)}}
-\newcommand{\evz}{{(\circ)}}
-\newcommand{\evp}{{(+)}}
-\newcommand{\enu}{{(\nu)}}
+\newcommand{\msolar}{\mathrm{M}_\odot}
 
-\newcommand{\msolar}{\mathrm{M}_\odot}
+% Software names
+\newcommand{\boxlib}{\texttt{BoxLib}}
+\newcommand{\castro}{\texttt{CASTRO}}
+\newcommand{\wdmerger}{\texttt{wdmerger}}
+\newcommand{\python}{\texttt{Python}}
+\newcommand{\matplotlib}{\texttt{matplotlib}}
+\newcommand{\yt}{\texttt{yt}}
 
@@
 \shorttitle{WD Mergers. II. Collisions}
-\shortauthors{Katz et al. (2015)}
+\shortauthors{Katz et al. (2016)}
 
-\author{TBD}
+\author{Max P. Katz\altaffilmark{1}}
+\author{Michael Zingale\altaffilmark{1}}
+\author{Alan C. Calder\altaffilmark{1,2}}
+\author{F. Douglas Swesty\altaffilmark{1}}
+\author{Ann S. Almgren\altaffilmark{3}}
+\author{Weiqun Zhang\altaffilmark{3}}
+
+\altaffiltext{1}
+{
+  Department of Physics and Astronomy,
+  Stony Brook University, Stony Brook, NY, 11794-3800, USA
+}
+
+\altaffiltext{2}
+{
+  Institute for Advanced Computational Sciences,
+  Stony Brook University, Stony Brook, NY, 11794-5250, USA
+}
+
+\altaffiltext{3}
+{
+  Center for Computational Sciences and Engineering,
+  Lawrence Berkeley National Laboratory, Berkeley, CA 94720
+}
+
 %==========================================================================
@@
 
-\cite{castro}
+We consider the collisions of white dwarfs as possible progenitors of Type Ia 
+supernovae, using the code \castro\ \cite{castro}.
 
@@
 
-This research was supported by NSF award AST-1211563.  This research
-used resources of the National Energy Research Scientific Computing
-Center, which is supported by the Office of Science of the
-U.S. Department of Energy under Contract No. DE-AC02-05CH11231.  An
+This research was supported by NSF award AST-1211563. An
 award of computer time was provided by the Innovative and Novel
@@
 the Office of Science of the Department of Energy under Contract
-DE-AC05-00OR22725.
+DE-AC05-00OR22725. Project AST106 supported use of the ORNL/Titan resource. 
+This research used resources of the National Energy Research Scientific Computing
+Center, which is supported by the Office of Science of the
+U.S. Department of Energy under Contract No. DE-AC02-05CH11231.
+Results in this paper were obtained using the high-performance
+LIred computing system at the Institute for Advanced Computational
+Science at Stony Brook University, which was obtained through
+the Empire State Development grant NYS \#28451.
+
+This research has made use of NASA's Astrophysics Data System 
+Bibliographic Services. In addition, this research has made use
+of the AstroBetter blog and wiki.
 

### 2016-01-11 papers/wdmerger_II/paper.tex
Add an outline of the sections for paper II

@@
 \newcommand{\yt}{\texttt{yt}}
+\newcommand{\aproxthirteen}{\texttt{aprox13}}
+\newcommand{\aproxnineteen}{\texttt{aprox19}}
+\newcommand{\aproxtwentyone}{\texttt{aprox21}}
 
@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 %==========================================================================
+% Numerical Implementation
+%==========================================================================
+\section{Numerical Methods}
+\label{sec:numericalmethods}
+
+The hydrodynamical equations solved in this paper were presented in \citet{wdmergerI}.
+In this paper we focus our discussion on our implementation of the nuclear
+reaction network and integrator.
+
+%==========================================================================
+% Problem Setup
+%==========================================================================
+\section{Problem Setup}
+\label{sec:problemsetup}
+
+We implement the white dwarf collision problem in \castro\ using our \wdmerger\
+package. The white dwarf centers of mass are separated by a distance of four times
+the (secondary) white dwarf radius. Their initial velocity is that of
+two point-masses in free-fall towards each other, at the appropriate distance.
+
+%==========================================================================
+% 2D simulations
+%==========================================================================
+\section{Two-Dimensional Simulations}
+\label{sec:2D}
+
+Though the main focus of this work is fully three-dimensional simulations,
+there is much to be learned from simpler, and much computationally cheaper,
+two-dimensional axisymmetric collision simulations. These are implemented
+in a cylindrical ($R-z$) coordinate system. We use these for three purposes.
+First, in the regime where the 2D and 3D simulations can be usefully compared, 
+zero-impact parameter simulations, to understand to what extent the
+reduced-dimensionality approach can reproduce the 3D results and thus to
+conclude whether 2D simulations can be used for obtaining scientific data
+such as nucleosynthesis yields. Second, for testing the effect of various
+code choices such as the number of isotopes in the nuclear network. A third
+effect which comes for free with the others is that we can benchmark existing
+and new code tools such as our nuclear postprocessing analysis. Prepared
+with these choices and a better understanding of how reliable our approximations
+are, we can then proceed to performing 3D simulations with our code choices
+already made and with confidence in our analysis software.
+
+\subsection{Dynamics}
+\label{sec:2D:dynamics}
+
+First we briefly consider whether the dynamical evolution of the system
+is consistent with simple physics we understand. We ask two questions.
+First, is the time it takes for the white dwarfs to collide consistent with
+basic kinematics. Second, is the amount of energy released in the
+collision by the nuclear reactions comparable to the amount of kinetic
+energy we infer from observations of Type Ia supernovae?
+
+\subsection{Timestep Limiting}
+\label{sec:2D:timestep}
+
+Here we discuss the effect of limiting the hydrodynamic timestep based on
+changes in the internal energy in the nuclear reaction network.
+
+\subsection{Nuclear Network}
+\label{sec:2D:network}
+
+Now we move to a discussion of the effect of including various isotopes
+in the nuclear reaction network. In particular we compare the \aproxthirteen,
+\aproxnineteen, and \aproxtwentyone reaction networks.
+
+\subsection{Nucleosynthetic Postprocessing}
+\label{sec:2D:postprocessing}
+
+We consider also the results of our postprocessing step for nucleosynthesis
+yields, to understand whether the results for the isotopes not included in
+our networks make sense.
+
+%==========================================================================
+% 3D simulations
+%==========================================================================
+\section{Three-Dimensional Simulations}
+\label{sec:3D}
+
+Now that we have analyzed the various parts of our simulation software and
+discussed our code choices, we consider full 3D simulations to understand
+the potential for white dwarf collisions to be progenitors of Type Ia
+supernovae. We focus on two main quantities of interest: the total amount of
+nickel produced in the explosion, which directly powers the light curve
+observed in SNe Ia; and, the gravitational wave signature produced by
+these events.
+
+%==========================================================================
 % Conclusions
@@
 \section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}
+\label{sec:2D}
 

### 2016-01-16 papers/wdmerger_I/paper.tex
Updated paper I

@@
-\documentclass[iop,numberedappendix]{../emulateapj}
+\documentclass[twocolumn,numberedappendix]{../aastex6}
 
@@
 \usepackage{verbatim}
-
-% allows in-document links; see http://www.astrobetter.com/blog/2014/09/29/latex-hyperref-and-emulateapj/
-\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref} 
-\usepackage[all]{hypcap} %Links go to figures;breaks on deluxetables       
-\renewcommand*{\backref}[1]{[#1]}
 
@@
 \shorttitle{WD Mergers I. Methodology}
-\shortauthors{Katz et al. (2015)}
+\shortauthors{Katz et al. (2016)}
 
@@
 Another approach to conserving total energy recently taken in the literature is to evolve an 
-equation for the total energy $(\rho E_{\text{tot}})$; see \cite{marcello:2012} and \cite{jiang:2013} (see also 
+equation for the total energy $(\rho E_{\text{tot}})$; see \cite{jiang:2013} (see also 
 \cite{arepo}, Section 5.3). That is, one can replace the gas energy equation with a total energy equation, 
@@
   R_A &= \begin{cases} 0 & |y - 0.5| > 0.25 \\ 1 & |y - 0.5| < 0.25 \end{cases} \label{eq:kh_ic_a_ramp}\\
-  R_B &= \Big\{\left[1 + e^{-2(y-0.25)/\Delta_y}\right]\left[1 + e^{2(y-0.75)/\Delta_y}\right]\Big\}^{-1} \label{eq:kh_ic_b_ramp}\\
+  R_B &= \Big\{\left[1 + e^{-2(y-0.25)/\Delta y}\right]\left[1 + e^{2(y-0.75)/\Delta y}\right]\Big\}^{-1} \label{eq:kh_ic_b_ramp}
+\end{align}
+\begin{align}
   v^\prime_A &= w_0\, \text{sin}\left(n_A\, \pi\, x\right) \left\{e^{-(y-0.25)^2 / 2\sigma^2} + e^{-(y-0.75)^2/2\sigma^2}\right\} \label{eq:kh_ic_a}\\
@@
 \end{align}
-
-\begin{figure*}[ht]
-  \centering
-  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p3_low_res_collage}}}
-  \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
-    conditions given by \autoref{eq:kh_p3_rho} through \autoref{eq:kh_p3_vy}. 
-    The meaning of the rows and columns is the same as in \autoref{fig:kh_p2}.
-    \label{fig:kh_p3}  {\color{red} You should note this is $R_B$}}
-\end{figure*}
-
 Here $w_0 = 0.1$ is the scale of the velocity perturbation, $\sigma =
@@
 generated every 0.05 seconds, and the problem is run until $t = 2$.
+
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p3_low_res_collage}}}
+  \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
+    conditions given by \autoref{eq:kh_p3_rho} through \autoref{eq:kh_p3_vy},
+    which come from McNally et al. (2012).
+    The meaning of the rows and columns is the same as in \autoref{fig:kh_p2}.
+    \label{fig:kh_p3}}
+\end{figure*}
 
@@
                          \rho_2 + \rho_m e^{(y-0.75)/\Delta y} & 0.75 > y \geq 0.5 \\
-                         \rho_1 - \rho_m e^{(0.75-y)/\Delta y} & 1 > y \geq 0.75 \end{dcases} \label{eq:kh_p3_rho}\\
+                         \rho_1 - \rho_m e^{(0.75-y)/\Delta y} & 1 > y \geq 0.75 \end{dcases} \label{eq:kh_p3_rho}
+\end{align}
+\begin{align}
   v_x &= \begin{dcases} v_1 - v_m e^{(y-0.25)/\Delta y} & 0.25 > y \geq 0 \\
@@
 spatial resolution to compensate for the increased numerical diffusion. The 
-accuracy is much lower for the highest bulk velocities. This is probably because 
-the initial conditions feature a transition that is twice as sharp (compare the $\Delta y$ 
-parameters in each case). This means that there are smaller scale numerical 
-perturbations on the grid, which can be amplified by numerical noise in the time 
-evolution of the system.
+qualitative accuracy is much lower for the highest bulk velocities for this problem
+than for the previous problems. This is because the amplitude of the instability overall
+is smaller than for the previous problems, at least by $t = 2.0$, so it is easier
+for numerical diffusion at the shearing layer, caused by the high bulk velocities,
+to completely wipe out the instability. Like \citet{robertson:2010} found for
+their problem, we find for this problem that the convergence properties are not
+substantially affected by altering the perturbation frequency -- the results
+show the same qualitative pattern even if we halve this frequency.
 
@@
 
-\capstartfalse
 \input{plots/sources.table}
-\capstarttrue
 
@@
 conservation (\autoref{fig:angular_momentum_conservation_unequal}).
-For the former system Angular momentum is conserved to within 10 percent over the 25
+For the former system angular momentum is conserved to within 10 percent over the 25
 orbits, while energy conservation is about an order of magnitude
@@
 Finally we consider whether the dynamical behavior of the system converges with resolution. 
-In \autoref{fig:unequal_spatial_convergence} we plot the first 10 seconds of the orbit for 
+In \autoref{fig:unequal_spatial_convergence_inertial} we plot the first full orbit for 
 the unequal mass system, at three different resolutions in the inertial frame: our default 
@@
 It is clear that at the latter resolution (corresponding to physical resolution of 100 km), 
-we have achieved convergent behavior. In the rotating frame (not shown here), the results are 
-not quite the same but for these resolutions the results do tend toward convergence. These 
-results suggest a minimum resolution of 200 km for getting the dynamics right.
+we have achieved convergent behavior. In the rotating frame, the results also show
+convergent behavior but the convergence is not as fast with resolution as in the inertial frame; 
+see \autoref{fig:unequal_spatial_convergence_rotating}. At the two higher resolutions the white dwarf
+distance is qualitatively similar, and both are qualitatively different from the lower resolution. However,
+quantitatively the two higher resolution runs are not as similar to each other as the analogous runs in the
+inertial frame. Convergence with resolution is slightly slower in the rotating reference frame
+because in the rotating reference frame a stable, unchanging circular orbit requires balance between
+two forces with opposite sign (the gravitational and centrifugal forces), and slight perturbations from the
+circular orbit are amplified by the effect of the Corolis force. In the inertial frame, these numerical instabilities
+vanish, but the cost is that there is no centrifugal force to actively maintain the white dwarf distance,
+which is why it is much more likely for the orbit to prematurely decay. In either case, these results suggest
+at least a minimum resolution of 200 km for getting the dynamics qualitatively right. To put that into context,
+consider that the parameter study of \citet{dan:2014} used 40,000 SPH particles per simulation, or (for an equal mass
+binary) 20,000 particles per white dwarf. For, say, a $0.9\ \msolar + 0.9\ \msolar$ white dwarf binary on
+a $256^3$ zone simulation grid, there are 20,000 zones that fit within a white dwarf. We do not
+intend here to directly compare results between the two simulation methods. We limit ourselves to the
+observation that at least for grid-based codes, a parameter study such as the ones performed by
+\citet{dan:2012} and \citet{dan:2014} would likely not yield qualitatively convergent results if it were to use the
+same effective mass resolution. Instead the number of zones inside each star should at least be doubled.
 
@@
   \includegraphics[scale=0.45]{plots/spatial_convergence_rot0}
-  \caption{Distance between the two white dwarfs in the unequal mass system, for the first 
-           10 seconds of their orbit. The distance is scaled by the initial orbital distance. 
+  \caption{Distance between the two white dwarfs in the unequal mass system, for the first orbit.
+           The distance is scaled by the initial orbital distance. 
            We plot at three different resolutions, corresponding to the number of 
-           effective zones per dimension in the refined regions. Note that the $n = 1024$ curve
-           falls nearly exactly on top of the $n = 512$ curve.\label{fig:unequal_spatial_convergence}}
+           effective zones per dimension in the refined regions.
+           \label{fig:unequal_spatial_convergence_inertial}}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.45]{plots/spatial_convergence_rot1}
+  \caption{Distance between the two white dwarfs in the unequal mass system, for the first orbit.
+           The distance is scaled by the initial orbital distance. 
+           We plot at three different resolutions, corresponding to the number of 
+           effective zones per dimension in the refined regions.
+           \label{fig:unequal_spatial_convergence_rotating}}
 \end{figure}
@@
 excellent weak scaling through 4,096 processors. At the largest run,
-the simulation time required is slightly than 1.5 times the amount
+the simulation time required is slightly less than 1.5 times the amount
 required for the smallest simulation.  This is due entirely to the
@@
 
-The authors thank John Bell for helpful discussions on gravity 
+The authors thank John Bell, Volker Springel, and Dominic Marcello
+for helpful discussions on gravity 
 and hydrodynamics issues. We also thank Noel Scudder and Platon Karpov 
@@
 
+\software{BoxLib (\url{https://github.com/BoxLib-Codes/BoxLib}),
+          CASTRO (\citet{castro}, \url{https://github.com/BoxLib-Codes/Castro}),
+          wdmerger (\url{https://github.com/BoxLib-Codes/wdmerger}),
+          GCC (\url{https://gcc.gnu.org/}),
+          python (\url{https://www.python.org/}),
+          matplotlib (\citet{matplotlib}, \url{http://matplotlib.org/}),
+          yt (\citet{yt}, \url{http://yt-project.org/})}
+
 \clearpage
@@
 
-\bibliographystyle{../apj}
+\bibliographystyle{../aasjournal}
 \bibliography{../refs}

### 2016-01-16 papers/submit.sh

@@ -1,4 +1,4 @@
-submit_dir="submit"
+submit_dir="submit2"
 
 # Build the PDF so that we have the bibilography
 
@@ -6,14 +6,14 @@ make
 
 # Move all the source files to the submit directory
 
-mkdir -p submit/
+mkdir -p $submit_dir
 
 cp -f paper.tex $submit_dir
 cp -f paper.bbl $submit_dir
 cp -f plots/*.eps $submit_dir
 cp -f plots/*.table $submit_dir
-cp -f ../apj.bst $submit_dir
-cp -f ../emulateapj.cls $submit_dir
+cp -f ../aasjournal.bst $submit_dir
+cp -f ../aastex6.cls $submit_dir
 
 # Get rid of all references to the plots subdirectory
 
@@ -28,4 +28,4 @@ sed -i "s/\.\.\///" $submit_dir/paper.tex
 
 sed -i "s/bibliography{refs}/input{paper.bbl}/" $submit_dir/paper.tex
 
-tar -czvf submit.tar $submit_dir
+tar -czvf $submit_dir.tar $submit_dir

### 2016-01-16 documentation/userguide.tex
Updated note on networks

@@
 
-For the nuclear network we use the \texttt{aprox13} network provided by Frank Timmes, which 
+For the nuclear network we use by default the \texttt{aprox13} network provided by Frank Timmes, which 
 we have ported for use in the \texttt{Microphysics} repository. It is an alpha-chain network 
-with isotopes ranging from ${}^4$He to ${}^{56}$Ni.
+with isotopes ranging from ${}^4$He to ${}^{56}$Ni. We have also ported \texttt{aprox19} and
+\texttt{aprox21}, which include the alpha chain isotopes as well as other nuclides for hydrogen
+burning and more accurate iron-group calculations. It is also possible to use the \texttt{iso7}
+network, which has the alpha nuclides from helium to silicon as well as nickel. The energy generation
+rate will be less accurate but at least should give some qualitatively correct answers. You can also
+provide your own network following the instructions in the Microphysics repository, as long as it
+contains at least ${}^4$He, ${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, and ${}^{24}$Mg. This the minimum
+set of isotopes needed to construct the initial white dwarf models.
 
@@
 At minimum, you'll need a \texttt{Prob\_3d.f90} file (or the equivalent for lower dimensions) and
-corresponding inputs and probin files. The problems in \texttt{science/wdmerger_I/castro} are similar to this;
+corresponding inputs and probin files. The problems in \texttt{science/wdmerger\_I/castro} are similar to this;
 they are ultimately building using the source code in the CASTRO test suite, but using a
-\texttt{run\_test.sh} that is very similar to the one used for merger problems.
+\texttt{run\_test.sh} that is very similar to the one used for merger problems. Note that for problems
+in CASTRO/Exec, you'll need to append \texttt{CASTRO\_DIR=\$CASTRO\_DIR} to \texttt{make} at the
+command line, which will force an override of that variable so that you can compile
+outside of the root CASTRO directory.
 
@@
 
-The \texttt{science/wdmerger_I/} directory contains a number of example problems that we use
+The \texttt{science/wdmerger\_I/} directory contains a number of example problems that we use
 to assess the reliability of \texttt{CASTRO} in performing hydrodynamics simulations 

### 2016-01-16 papers/wdmerger_II/paper.tex
add the DOE NP grant
Switch over papers II and III to aastex6

@@
 
-% allows in-document links; see http://www.astrobetter.com/blog/2014/09/29/latex-hyperref-and-emulateapj/
-\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref} 
-\usepackage[all]{hypcap} %Links go to figures;breaks on deluxetables       
-\renewcommand*{\backref}[1]{[#1]}
-
 % Override choices in \autoref
@@
 \newcommand{\yt}{\texttt{yt}}
+\newcommand{\isoseven}{\texttt{iso7}}
 \newcommand{\aproxthirteen}{\texttt{aprox13}}
@@
 \section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}
-\label{sec:2D}
+\label{sec:conclusion}
 
@@
 
-This research was supported by NSF award AST-1211563. An
-award of computer time was provided by the Innovative and Novel
-Computational Impact on Theory and Experiment (INCITE) program.  This
-research used resources of the Oak Ridge Leadership Computing Facility
-located in the Oak Ridge National Laboratory, which is supported by
-the Office of Science of the Department of Energy under Contract
-DE-AC05-00OR22725. Project AST106 supported use of the ORNL/Titan resource. 
-This research used resources of the National Energy Research Scientific Computing
-Center, which is supported by the Office of Science of the
-U.S. Department of Energy under Contract No. DE-AC02-05CH11231.
-Results in this paper were obtained using the high-performance
-LIred computing system at the Institute for Advanced Computational
-Science at Stony Brook University, which was obtained through
-the Empire State Development grant NYS \#28451.
+This research was supported by NSF award AST-1211563 and DOE/Office of
+Nuclear Physics grant DE-FG02-87ER40317 to Stony Brook. An award of
+computer time was provided by the Innovative and Novel Computational
+Impact on Theory and Experiment (INCITE) program.  This research used
+resources of the Oak Ridge Leadership Computing Facility located in
+the Oak Ridge National Laboratory, which is supported by the Office of
+Science of the Department of Energy under Contract
+DE-AC05-00OR22725. Project AST106 supported use of the ORNL/Titan
+resource.  This research used resources of the National Energy
+Research Scientific Computing Center, which is supported by the Office
+of Science of the U.S. Department of Energy under Contract
+No. DE-AC02-05CH11231.  Results in this paper were obtained using the
+high-performance LIred computing system at the Institute for Advanced
+Computational Science at Stony Brook University, which was obtained
+through the Empire State Development grant NYS \#28451.
 

### 2016-01-16 papers/wdmerger_II/paper.tex
Add some content to the numerical methods section in paper II.

@@
 \newcommand{\castro}{\texttt{CASTRO}}
+\newcommand{\microphysics}{\texttt{Microphysics}}
 \newcommand{\wdmerger}{\texttt{wdmerger}}
@@
 \newcommand{\yt}{\texttt{yt}}
+\newcommand{\vode}{\texttt{VODE}}
 \newcommand{\isoseven}{\texttt{iso7}}
@@
 In this paper we focus our discussion on our implementation of the nuclear
-reaction network and integrator.
+reaction network and integrator. There are three main areas of concern: first,
+specifying what nuclides are in the network and what reaction rates link the
+various nuclides; second, once the system of ODEs governing the evolution of
+the nuclear species has been written down, how to append to them ODEs governing
+the evolution of the thermodynamics, in particular the temperature; third,
+coupling the nuclear reaction updates to the evolution of the hydrodynamical
+system. We will deal with each of these in turn in the following subsections.
+
+\subsection{Nuclear Network}
+\label{sec:network}
+
+White dwarfs are mainly composed of $\alpha$-chain particles, primarily ${}^4$He,
+${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, and ${}^{24}$ Mg. Therefore the core of
+any network appropriate for modeling nuclear burning in white dwarfs will be
+these alpha chain nuclides, with the idea being that links up the $\alpha$-chain
+will eventually get us to ${}^{56}$Ni, the nuclide responsible for the
+energy output of Type Ia supernovae. In this paper we consider four networks
+to do this, presented in order of increasing complexity. The most simple is
+\isoseven\ \citep{timmes:2000}, which includes all of the aforementioned isotopes and
+${}^{28}$Si (see also \citet{hix:1998}). ${}^{28}$Si effectively measures the
+equilibrium state of silicon-group elements, and ${}^{56}$Ni effectively measures
+the equilibrium state of iron-group elements, with the link between them governed
+by the effective loss or gain of seven $\alpha$-particles. This type of network
+was used by \citet{rosswog:2009} for their collision calculations in SPH.
+
+Next is \aproxthirteen\ \citep{timmes:1999,timmes:2000}. This includes
+all of the isotopes of \isoseven, and all of the $alpha$-chain particles between
+silicon and nickel (${}^{32}$S, ${}^{36}$Ar, ${}^{40}$Ca, ${}^{44}$Ti, ${}^{48}$Cr,
+and ${}^{52}$Fe). This network was used by \citet{hawley:2012} and \citet{raskin:2010}.
+\citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a very similar network
+that included additionally ${}^{60}$Zn. The \aproxnineteen\ network \citep{timmes:1999}
+builds on \aproxthirteen\ by including isotopes for hydrogen burning and explicit
+tracking of photodisintegration into ${}^{54}$Fe. This network was used by
+\citet{kushnir:2013}, \citet{kushnir:2014}, \citet{rosswog:2009} for their
+calculations with FLASH, and \citet{papish:2015}. Finally we will also
+consider \aproxtwentyone, which includes all of the above plus ${}^{56}$Cr
+and ${}^{56}$Fe and related reaction links. The primary virtue of using
+the latter two networks is that they allow us to track changes away from
+$Y_e = 0.5$.
+
+All four of these networks have been ported into a form that is consistent
+with the \boxlib\ codes, in the freely available \microphysics\ code
+repository\footnote{\microphysics\ can be obtained at \url{https://github.com/BoxLib-Codes/Microphysics}.},
+a collection of microphysical routines that are designed to be used in our
+hydrodynamics codes. These can be easily swapped at compile time by using the 
+appropriate makefile variable.
+
+\subsection{Nuclear Burning}
+\label{sec:burner}
+
+Given a set of nuclides and the reaction links between them, we now consider
+how a burning step is performed in our software. The integration of the stiff,
+implicit system is performed by \vode\ \citep{vode}, a copy of which is
+provided with \castro. To \vode\ we provide an integration state containing
+the molar abundances $Y_{n} = X_{n} / A_{n}$, where $X$ is the mass fraction
+of the abundance in the zone and $A$ is the number of nucleons. The integration
+of
+\begin{equation}
+  \frac{d\bm{Y}}{dt} = f(\mathbf{Y})
+\end{equation}
+is determined using a right-hand-side provided by the nuclear burning network.
+When it comes to the thermodynamical evolution, the internal energy $e$ of the zone
+will change when the nuclear abundances evolve. During the integration we track
+the energy release from the system using
+\begin{equation}
+  \frac{\partial e}{\partial t} = N_A \sum_{n} \frac{\partial Y_{n}}{\partial t} m_{n} c^2,
+\end{equation}
+where $c$ is the speed of light and $m_n$ is the mass of each nuclide.
+
+In a hydrostatic burn, we keep $\rho$ and $T$ fixed throughout, and use
+the energy released at the end to compute a final temperature that is
+thermodynamically consistent with the new internal energy. By contrast,
+in a self-heating burn, we allow the temperature to evolve in response
+to the burning (see \citet{maestro3}):
+\begin{equation}
+  \frac{dT}{dt} = \frac{1}{c_V}\left(\frac{\partial e}{\partial t} - \sum_n \frac{\partial e}{\partial X_n}\frac{\partial X_n}{\partial t}\right)
+\end{equation}
+Here $c_V$ is the specific heat at constant volume, and $(\partial e/\partial X_n)$
+is the derivative of the internal energy with respect to the mass fraction of a given
+species (with other thermodynamic variables held constant). Both quantities are provided by the
+equation of state. During this burn, we can keep these thermodynamic
+derivatives constant using their value at the beginning of the run, or at each step we
+can choose to re-evaluate the equation of state using the latest value of $(rho, T)$.
+The latter is more expensive but also more accurate, and we use it in this paper.
+A third option presented by \citet{raskin:2010} is a so-called ``hybrid'' mode.
+In this mode, by default we do a hydrostatic burn. If that burn fails, or if the net
+energy change is negative, we do the burn again in self-heating mode. All three options
+are implemented in our burner software.
+
+\subsection{Coupling to the Hydrodynamics}
+\label{sec:hydrocoupling}
+
+In \castro, the reactions are coupled to the hydrodynamics using Strang splitting.
+In a given timestep advance $\Delta t$, we first evolve the reactions alone through
+a time interval $\Delta t / 2$. Then, we evolve the hydrodynamics for $\Delta t$,
+and we evolve the reactions again for a further $\Delta t / 2$. The principal
+drawback of this approach is that the reactions and the hydrodynamics can become
+decoupled from each other. A common solution to this problem presented in
+the literature has been to limit the size of the timestep and thereby limit the
+extent of this decoupling \citep{raskin:2010,hawley:2012}, which we adopt here.
+(A related approach presented in \citet{kushnir:2013} is to limit the magnitude
+of the energy release in any given integration by scaling all the reaction rates.)
+At the start of each advance, we limit the size of the timestep to be the smaller
+of the minimum hydro timestep (limited by the CFL condition), and the minimum burning
+timestep. In a given zone, the burning timestep is defined as:
+\begin{equation}
+  \Delta t_{\text{burning}} = f_b\, \frac{e}{\dot{e}}.
+\end{equation}
+Here $\dot{e}$ is an estimate of the time rate of change of the internal energy
+from nuclear burning. We estimate this using the energy change in the last burning
+update, divided by the length of the burning timestep. The factor $f_b$ then
+determines by what fraction we would like to allow the internal energy to change
+in the current timestep (assuming that $\dot{e}$ does not change quickly from
+timestep to timestep, which is not always the case). By making $f_b$ smaller,
+we can control the magnitude of the decoupling between the reactions and the hydro.
+A typical choice in the literature is $f_b \sim 0.3$. This timestep limiting has
+now been implemented in \castro\ and the factor $f_b$ can be set at runtime.
 

### 2016-01-20 documentation/userguide.tex
Refined comment about analysis routines

@@
 If python and yt are not installed on your system, you will still be able to compile and 
-run this code, but you will need some other mechanism for analyzing the \texttt{BoxLib} output data.
-We have some routines for this purpose available in Fortran and may be able to provide them
-upon request.
+run this code, but we use yt for all analysis of \texttt{BoxLib} output data other than 
+the ASCII files that are generated at runtime. We have some routines for this purpose 
+available in Fortran and Python; some of the Fortran tools are provided in \texttt{BoxLib} 
+itself, in the \texttt{Tools/Postprocessing} subdirectory, and it is straightforward enough 
+to modify them for basic analysis. We may be able to provide others upon request.
 

### 2016-01-21 papers/wdmerger_II/paper.tex
Add a description to paper II of the new burning limiter

@@
 extent of this decoupling \citep{raskin:2010,hawley:2012}, which we adopt here.
-(A related approach presented in \citet{kushnir:2013} is to limit the magnitude
-of the energy release in any given integration by scaling all the reaction rates.)
-At the start of each advance, we limit the size of the timestep to be the smaller
-of the minimum hydro timestep (limited by the CFL condition), and the minimum burning
-timestep. In a given zone, the burning timestep is defined as:
+We define a burning-limited timestep $\Delta t_{b1}$ as:
 \begin{equation}
-  \Delta t_{\text{burning}} = f_b\, \frac{e}{\dot{e}}.
+  \Delta t_{b1} = f_{b1}\, \frac{e}{|\dot{e}|}.
 \end{equation}
@@
 from nuclear burning. We estimate this using the energy change in the last burning
-update, divided by the length of the burning timestep. The factor $f_b$ then
+update, divided by the length of the burning timestep. The factor $f_{b1}$ then
 determines by what fraction we would like to allow the internal energy to change
 in the current timestep (assuming that $\dot{e}$ does not change quickly from
-timestep to timestep, which is not always the case). By making $f_b$ smaller,
+timestep to timestep, which is not always the case). By making $f_{b1}$ smaller,
 we can control the magnitude of the decoupling between the reactions and the hydro.
-A typical choice in the literature is $f_b \sim 0.3$. This timestep limiting has
-now been implemented in \castro\ and the factor $f_b$ can be set at runtime.
+Typical choices in the literature are $f_{b1} = 0.2$ or $0.3$. This timestep limiting has
+now been implemented in \castro\ and the factor $f_{b1}$ can be set at runtime.
+
+\citet{kushnir:2013} point out this is not the only way for the burning to
+become numerically unstable; another failure mode is when the energy injection timescale
+is faster than the sound-crossing time in a zone. \citet{kushnir:2013} solved this by
+limiting the magnitude of the energy release after a burn; we take a slightly different
+approach and limit the timestep based on the previous burning, as with the other limiter.
+The goal is to ensure that the following condition holds:
+\begin{equation}
+  \frac{\partial e}{\partial t} = f_{b2} \frac{e}{t_s},
+\end{equation}
+The sound crossing time, $t_s$, is given by $\Delta x / c_s$, where $c_s$ is the sound speed
+and $\Delta x$ is the zone size. The parameter $f_{b2}$ then determines the ratio of the
+nuclear energy generation timescale to the sound-crossing time, in the limit of uniform
+timesteps. \citet{kushnir:2013} choose $f_{b2} = 0.1$ for their simulations.
+(As with the previous limiter, this parameter can be set at runtime in \castro.)
+We then re-write this as a limit on the size of the next timestep, using the same
+estimate as above for the nuclear energy generation rate $\dot{e}$ from the last burn:
+\begin{equation}
+  \Delta t_{b2} = \left[f_{b2}\, \frac{e}{|\dot{e}| \Delta t_{\rm{old}}}\right]^{-1}\, {t_s}
+\end{equation}
+Here $\Delta t_{\rm{old}}$ is the size of the timestep used in the burning advance that calculated
+$\dot{e}$. In the case of our Strang splitting approach it is half of the old advance
+timestep, $\Delta t^{n-1} / 2$. Thus the denominator inside the brackets is simply the net
+change in internal energy during the last burn.
+
+At the start of each advance, we limit the size of the timestep to be the smaller
+of the minimum hydro timestep (limited by the CFL condition), and the minimum of each
+burning timestep across all zones.
 

### 2016-01-27 papers/wdmerger_II/paper.tex
rho -> \rho

@@
 derivatives constant using their value at the beginning of the run, or at each step we
-can choose to re-evaluate the equation of state using the latest value of $(rho, T)$.
+can choose to re-evaluate the equation of state using the latest value of $(\rho, T)$.
 The latter is more expensive but also more accurate, and we use it in this paper.

### 2016-01-29 papers/wdmerger_II/paper.tex
Update discussion of the burning limiters to reflect new AMR approach

--- 2016-01-16 papers/wdmerger_II/paper.tex
@@
 the literature has been to limit the size of the timestep and thereby limit the
-extent of this decoupling \citep{raskin:2010,hawley:2012}, which we adopt here.
-(A related approach presented in \citet{kushnir:2013} is to limit the magnitude
-of the energy release in any given integration by scaling all the reaction rates.)
-At the start of each advance, we limit the size of the timestep to be the smaller
-of the minimum hydro timestep (limited by the CFL condition), and the minimum burning
-timestep. In a given zone, the burning timestep is defined as:
+extent of this decoupling \citep{raskin:2010,hawley:2012}, which we adopt here 
+and have implemented in \castro. Defining the nuclear energy injection timescale 
+$t_e$:
 \begin{equation}
-  \Delta t_{\text{burning}} = f_b\, \frac{e}{\dot{e}}.
+  t_e \equiv \frac{e}{|\dot{e}|},
 \end{equation}
-Here $\dot{e}$ is an estimate of the time rate of change of the internal energy
-from nuclear burning. We estimate this using the energy change in the last burning
-update, divided by the length of the burning timestep. The factor $f_b$ then
+where $\dot{e}$ is an estimate of the time rate of change of the internal energy
+from nuclear burning, we define a burning-limited timestep $\Delta t_{b}$:
+\begin{equation}
+  \Delta t_{b} = f_{\Delta t}\, t_e.
+\end{equation}
+We estimate $\dot{e}$ using the energy change in the last burning
+update, divided by the length of the burning timestep. The factor $f_{\Delta t}$ then
 determines by what fraction we would like to allow the internal energy to change
 in the current timestep (assuming that $\dot{e}$ does not change quickly from
-timestep to timestep, which is not always the case). By making $f_b$ smaller,
+timestep to timestep, which is not always the case). By making $f_{\Delta t}$ smaller,
 we can control the magnitude of the decoupling between the reactions and the hydro.
-A typical choice in the literature is $f_b \sim 0.3$. This timestep limiting has
-now been implemented in \castro\ and the factor $f_b$ can be set at runtime.
+We choose by default $f_{\Delta t} = 0.1$, a choice which is on par with others in 
+the literature. The sensitivity of results to this choice will be discussed in 
+\autoref{sec:2D:timestep}. The factor $f_{\Delta t}$ can be set at runtime in \castro.
+
+At the start of each advance, we limit the size of the timestep to be the smaller
+of the minimum hydro timestep (limited by the CFL condition), and the minimum of the
+burning timestep across all zones.
+
+\citet{kushnir:2013} point out this is not the only way for the burning to
+become numerically unstable; another failure mode is when the energy injection timescale
+$t_e$ is shorter than the sound-crossing time $t_s$ in a zone.
+The goal is to ensure that the following condition holds:
+\begin{equation}
+  t_s \leq f_{\Delta x} t_e \label{eq:burning_limiter_2}
+\end{equation}
+The sound crossing time, $t_s$, is given by $\Delta x / c_s$, 
+where $c_s$ is the sound speed and $\Delta x$ is the (minimum) 
+zone width. The parameter $f_{\Delta x}$ then determines the minimum
+ratio of the nuclear energy generation timescale to the 
+sound-crossing time. \citet{kushnir:2013} choose $f_{\Delta x} = 0.1$ 
+for their simulations, and we do too (this parameter can be set 
+at runtime in \castro).
+
+\citet{kushnir:2013} implemented this criterion by artificially 
+limiting the magnitude of the energy release after a burn. As 
+this is not physical, we take a different approach. Since we 
+cannot directly control the energy injection timescale, we 
+must find a way to alter the sound-crossing timescale. 
+We can achieve this by adding levels of refinement in 
+regions that do not satisfy \autoref{eq:burning_limiter_2},
+which effectively lowers the $\Delta x$ and thus the
+sound-crossing time. We keep tagging zones for refinement
+based on this criterion until the criterion is satisfied
+on the finest level. Since the concern is regions that 
+may detonate, we also tag nearby zones in a buffer region
+which do not themselves satisfy the criterion,
+so that a detonation in a single timestep cannot 
+escape into non-refined regions. The width of the buffer 
+region should thus be at least as large as the number of 
+timesteps before a regridding procedure is performed.
+We choose a value of two for both the number of zones in the 
+buffer region and the number of steps in between regrids,
+for all simulations in this paper.
+
+While this approach may add several AMR levels 
+to a simulation, we agree with \citet{kushnir:2013} that 
+solving this numerical instability is crucial to avoiding
+early, unphysical detonations, and so it is unavoidable
+if a correct evaluation of the burning phase is desired.
+A simulation that does not solve this problem in this way
+(or some analogous manner) will not obtain the correct amount 
+of burning, and will not converge properly with resolution. 
+Fortunately, the regions where this criterion is not satisfied 
+are fairly localized on the grid, so the amount of additional 
+computational work is not exceedingly large.
 

### 2016-02-14 papers/wdmerger_II/paper.tex
Update section on timestep limiting in paper II

@@
 and have implemented in \castro. Defining the nuclear energy injection timescale 
-$t_e$:
-\begin{equation}
-  t_e \equiv \frac{e}{|\dot{e}|},
-\end{equation}
+$\tau_e$, and the species evolution timescale $\tau_{X_k}$,
+\begin{align}
+  \tau_e &\equiv \frac{e}{|\dot{e}|} \\
+  \tau_{X_k} &\equiv \frac{X_k}{|\dot{X_k}|},
+\end{align}
 where $\dot{e}$ is an estimate of the time rate of change of the internal energy
-from nuclear burning, we define a burning-limited timestep $\Delta t_{b}$:
-\begin{equation}
-  \Delta t_{b} = f_{\Delta t}\, t_e.
-\end{equation}
-We estimate $\dot{e}$ using the energy change in the last burning
-update, divided by the length of the burning timestep. The factor $f_{\Delta t}$ then
-determines by what fraction we would like to allow the internal energy to change
-in the current timestep (assuming that $\dot{e}$ does not change quickly from
-timestep to timestep, which is not always the case). By making $f_{\Delta t}$ smaller,
-we can control the magnitude of the decoupling between the reactions and the hydro.
-We choose by default $f_{\Delta t} = 0.1$, a choice which is on par with others in 
-the literature. The sensitivity of results to this choice will be discussed in 
-\autoref{sec:2D:timestep}. The factor $f_{\Delta t}$ can be set at runtime in \castro.
+from nuclear burning, and $\dot{X_k}$ is an estimate of the time rate of change 
+of the mass fraction of the species with index $k$, we define burning-limited 
+timesteps $\Delta t_{be}$ and $\Delta t_{bX_k}$:
+\begin{align}
+  \Delta t_{be} &= f_{e}\, \tau_e \\
+  \Delta t_{bX_k} &= f_{X}\, \tau_{X_k}.
+\end{align}
+Given an estimate for $\dot{e}$, the factor $f_{e}$ determines by what 
+fraction we would like to allow the internal energy to change
+in the current timestep, under the assumption that $\dot{e}$ does not change from
+timestep to timestep. Similarly, given an estimate for $\dot{X_k}$, the factor $f_{X}$ 
+determines the maximum change in the mass fraction of any species. By making 
+$f_{e}$ and $f_{X}$ smaller, we can control the magnitude of the decoupling 
+between the reactions and the hydro. We choose by default $f_{X} = f_{e} = 0.1$, 
+a choice which is slightly smaller than others in the literature. (Note that previous 
+collision papers we are aware of only limited based on changes in internal energy, 
+not on cahnges in species.)  The sensitivity of results to this choice will be discussed in 
+\autoref{sec:2D:timestep}. The factors $f_{e}$ and $f_{X}$ can be set at runtime in \castro.
 
 At the start of each advance, we limit the size of the timestep to be the smaller
-of the minimum hydro timestep (limited by the CFL condition), and the minimum of the
-burning timestep across all zones.
+of the minimum hydro timestep (limited by the CFL condition), and the minimum of all the
+burning timesteps across all zones. To do this, we need a method for determining 
+$\dot{e}$ and $\dot{X_k}$. A typical choice in the literature has been to set, for example,
+\begin{equation}
+  \dot{e} = \frac{e^{n} - e^{n-1}}{\Delta t^{n-1}}, \label{eq:burning_mode_4}
+\end{equation}
+where is $e^n$ is the internal energy at the start of the current timestep and
+$e^{n-1}$ is the internal energy at the start of the previous timestep. 
+The obvious analogue is used for constructing the species rate of change.
+However, there are alternative methods of constructing this derivative estimate, 
+and we have found that these different methods have measurable consequences.
+We define four separate methods for calculating the time derivative, with 
+the above being method 4. Method 3 is similar to method 4 but replaces the
+denominator in Equation \ref{eq:burning_mode_4} with the change in 
+internal energy over the last timestep only from the nuclear reactions.
+Method 2 is the same as method 3 but we only use the change in internal 
+energy from the most recent nuclear burning step, that is, the second-half
+of the Strang-split burning from the last timestep (the denominator 
+then becomes $\Delta t / 2$). In method 1, the most accurate option and 
+the current default in \castro, we evaluate the right-hand-side of the 
+burning network given the current state to explicitly obtain the 
+instantaneous value of $\dot{e}$ and $\dot{X_k}$.
 
-\citet{kushnir:2013} point out this is not the only way for the burning to
-become numerically unstable; another failure mode is when the energy injection timescale
+To understand the consequences of this choice, and more broadly to 
+understand the limitations of Strang splitting, we consider the 
+basic outline of a single-level advance in an advection-reaction system:
+\begin{enumerate}
+  \item Evaluate timestep $\Delta t$ for the current advance
+  \item Advance the nuclear burning network by $\Delta t / 2$
+  \item Advance the hydrodynamics by $\Delta t$
+  \item Advance the nuclear burning network by $\Delta t / 2$
+  \item Return to Step 1
+\end{enumerate}
+Now, consider that during a head-on collision, initial nuclear burning 
+will occur at the contact point between the two stars. Because of 
+the staggered updates from splitting, the evolution effectively progresses 
+as a cycle between burning for $\Delta t$ and getting fresh material 
+advected into the contact point by the hydro update for $\Delta t$. 
+When the collision begins, $\Delta t$ is controlled by the hydrodynamic 
+stability criterion, and may be large enough that it is possible for 
+the burning advance in Step 4 to completely burn the freshly advected 
+material all the way to NSE. Consequently the evolution is no longer 
+a good approximation to smooth burning of the in-falling material but
+rather separate discrete burning and hydro steps, and the nature of 
+the burning evolution will be quite different. Furthermore, by the 
+time we return to Step 1 and estimate the next timestep size, all 
+of the burning rates will be small again, and the instantaneous 
+timestep limiter of method 1 may actually substantially overestimate 
+the needed timestep. The other methods will see that the energy/species  
+substantially changed over the last timestep, but will still 
+overestimate the needed timestep because the burning was quiescent
+for at least some portion of the last advance. In particular, our 
+experience has shown that using only the energy change criterion
+is particularly susceptible to this phenomenon; silicon-group 
+material can build up without changing the internal energy by a 
+large fraction, so the timestep limiter is never triggered, and 
+then in a single step a substantial amount of iron-group elements 
+can be be generated. This can have non-trivial effects on the 
+total amount of iron-group elements generated over the course of
+the simulation. We have found that the addition of the 
+limiter based on changes in species functionally precludes this.
+
+With the timestep limited the way we advocate in this paper, 
+the timesteps are generally short enough so that the errors 
+due to splitting are small. Other approaches to the coupling 
+between reactions and hydrodynamics have been proposed in the 
+broader literature, especially iterative methods such as 
+deferred corrections that allow each of these operators to 
+feel the lagged effects of the other operators. For example,
+in the context of low Mach number flows, \cite{nonaka:2012} have
+used the method of spectral deferred corrections \cite{SDC} to
+couple their advection-diffusion-reaction equation set. In our
+context this would involve using the advection as a source term
+for the burning network, and symmetrically including a source
+term to the hydrodynamics corresponding to the energy release
+from the burning. We are presently investigating such a method,
+and it may form the basis of further work on this subject.
+
+Finally, we return to a point we hinted at above. The timestep
+will only actually satisfy the energy criterion
+$\Delta t \leq f_e \tau_e$ and species criterion
+$\Delta t \leq f_{X_k} \tau_{X_k}$ when the estimates for
+$\dot{e}$ and $\dot{X_k}$ we generate are at least as large
+as the actual rate of change of energy and mass fractions
+over the timestep. However, this can assumption can fail
+quite radically during periods of runaway burning when
+the rate of change of these quantities is highly nonlinear.
+We do not want to neglect the errors caused by this approximation
+because they may build up over an extended period of nonlinear
+evolution and perhaps substantially change the final results.
+To this end, we have implemented a timestep retry option in
+\castro. The retry option works by first saving a copy of the
+full state, and then doing the advance. We then explicitly
+check whether the timestep criteria were violated during the advance,
+using the time-centered energy and mass fractions and a
+time-centered differencing of the old and new states to
+calculate the evolution timescale. If so, we reject this
+advance and restore the original state data. We then perform
+a subcycled update, with the maximum timestep chosen so that
+the criterion is satisfied in each subcycle. Note that this
+occurs entirely on a single level and is separate from the
+subcycling that is used by \boxlib\ to advance the different
+AMR levels in time; the full AMR algorithm does not ever
+know that the retry subcycling took place. This can help
+substantially to mollify the problem described above, where
+the time derivative estimation methods proposed were not
+sufficient to obtain the desired limiting. Thus we strongly
+recommend that hydrodynamics codes implement such an approach,
+which ensures that the coupling actually does stay at the
+level requested. The most significant drawback is that saving
+a copy of the old state effectively doubles the memory footprint
+of the code, so this option should be disabled when not needed.
+However, it may be crucial in determining the evolution of
+systems such as white dwarf collisions, where even a short
+period of incorrect advances can significantly pollute the
+final answer, and the additional memory expense is justified
+by obtaining a more accurate solution.
+
+\subsection{Numerically Unstable Burning}
+
+\citet{kushnir:2013} point out that an inappropriate timestep is 
+not the only way for the numerical discretization to cause 
+severe errors in the burning. Another failure mode is when
+the energy injection timescale
 $t_e$ is shorter than the sound-crossing time $t_s$ in a zone.
@@
 \begin{equation}
-  t_s \leq f_{\Delta x} t_e \label{eq:burning_limiter_2}
+  t_s \leq f_{s} t_e \label{eq:burning_limiter_2}
 \end{equation}
@@
 where $c_s$ is the sound speed and $\Delta x$ is the (minimum) 
-zone width. The parameter $f_{\Delta x}$ then determines the minimum
+zone width. The parameter $f_{s}$ then determines the minimum
 ratio of the nuclear energy generation timescale to the 
-sound-crossing time. \citet{kushnir:2013} choose $f_{\Delta x} = 0.1$ 
+sound-crossing time. \citet{kushnir:2013} choose $f_{s} = 0.1$ 
 for their simulations, and we do too (this parameter can be set 
@@
 are fairly localized on the grid, so the amount of additional 
-computational work is not exceedingly large.
+computational work is not insurmountable for a single simulation.
+However, this level of accuracy may make a parameter study of
+the problem unfeasible, given current computational resources.
 

### 2016-02-14 papers/wdmerger_II/paper.tex
Add discussion of the dual energy switch to paper II

@@
 
-Finally, we return to a point we hinted at above. The timestep
+Now we return to a point we hinted at above. The timestep
 will only actually satisfy the energy criterion
@@
 by obtaining a more accurate solution.
+
+One other point worth noting for the coupling of the reactions
+and the hydrodynamics relates to the dual-energy formalism used
+by \castro\ and many other hydrodynamics codes \cite{bryan:1995}.
+\castro\ evolves separately equations for the total energy $(\rho E)$ and
+for the internal energy $(\rho e)$. The former is conservative while the latter
+is not, so for accurate hydrodynamics we prefer to use the total
+energy when possible and to use the evolved internal energy variable
+only in situations where the kinetic energy dominates the contribution
+to the total energy. Indeed, for the cosmological purpose for which
+this was originally developed, \cite{ENZO} choose a value $\eta_2 = 0.1$
+so that $e$ is only updated to be equal to $E - \mathbf{v}^2/2$ if
+$e > \eta_2\, E$. In other cases the evolved value of the internal
+energy variable is preserved. However, this choice does not work
+for our application, because the ultimate cause of the nuclear
+burning in a white dwarf collision is the rapid conversion of
+kinetic energy from the white dwarf bulk motion to thermal energy
+as the white dwarfs slam into each other. Keeping $\eta_2$ large
+prevents this from happening, and consequently the temperature
+will never reach values high enough to generate significant
+amounts of nickel. For this paper we choose $\eta_2 = 10^{-4}$,
+which is low enough for the correct conversion of kinetic energy
+but not so low that we need to be concerned about roundoff issues
+caused by subtracting kinetic from total energy. This is also
+the value that was used by \cite{hawley:2012}.
 
@@
 
-\clearpage
-\appendix
-
 \end{document}

### 2016-02-23 papers/wdmerger_III/paper.tex
Switch over papers II and III to aastex6
Add a first cut at describing the hybrid advection scheme to paper III

--- 2015-05-05 papers/wdmerger_II/paper.tex
@@
-%\documentclass[12pt,preprint]{../aastex}
-\documentclass[iop]{../emulateapj}
-% for \sout
-\usepackage{ulem}
-% makes sure \em{} is italic rather than underlined (corrects ulem from line above)
-\normalem
+\documentclass[twocolumn,numberedappendix]{../aastex6}
+
+% these lines seem necessary for pdflatex to get the paper size right
+\pdfpagewidth 8.5in
+\pdfpageheight 11.0in
 
@@
 
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
+% allows Greek symbols to be bold
+\usepackage{bm}
+
+% allows us to force the location of a figure
+\usepackage{float}
+
+% allows comment sections
+\usepackage{verbatim}
+
+% Override choices in \autoref
+\def\sectionautorefname{Section}
+\def\subsectionautorefname{Section}
+\def\subsubsectionautorefname{Section}
 
@@
 %==========================================================================
-\title{Double White Dwarf Mergers with CASTRO\\ II. Equilibrium Initial Conditions}
+\title{White Dwarf Mergers on Adaptive Meshes\\ III. Mergers}
 
-\shorttitle{DWD Mergers. II. Equilibrium ICs}
-\shortauthors{Max}
+\shorttitle{WD Mergers. III. Mergers}
+\shortauthors{Katz et al. (2016)}
 
@@
 for a general equation of state. We extend this method to include the effects of adaptive
-mesh refinement in the algorithm.
+mesh refinement in the algorithm. Then we consider mergers of these binary stars.
 
@@
 
-\cite{castro}
+%==========================================================================
+% Numerical Methods
+%==========================================================================
+\section{Numerical Methods}
+
+\subsection{Hybrid Advection Scheme}
+
+It is well known that there is a trade-off between conservation of linear
+momentum and conservation of angular momentum in grid-based, Eulerian codes
+such as \castro. Although both are true statements analytically, we cannot
+numerically guarantee that both hold. Consequently the choice of which
+physical quantity to conserve must be motivated by analysis of the particular
+problem to be done. Ideally a simulation code will be capable of evolving the
+equations for both the linear momentum and the angular momentum, and be
+able to switch between them as appropriate. Prior to this work, \castro\
+already had the standard method implemented for evolving the linear momentum.
+To solve the need for an angular momentum evolution, we adopt the ``hybrid''
+advection scheme presented by \cite{byerly:2014} (see also \cite{motl:2002} for
+a similar equation set). Their approach is appropriate for physical situations
+where there is a dominant angular momentum axis and it is desirable to conserve
+that component as accurately as possible. This is certainly applicable for
+studies of the stability of binary stellar systems. This section contains a
+discussion of our implementation of that scheme in \castro. For the sake
+of simplicity, let us assume that the rotation axis of the binary system is the $z$
+axis and so the WDs orbit in the $xy$-plane. The core of the method is that instead
+of solving the Euler equations for the linear momenta $(\rho u)$ and $(\rho v)$,
+we instead solve the corresponding equations for the radial momentum
+$s_R \equiv \rho v_R$ (where $v_R$ is the radial velocity with respect to the
+rotation axis) and the angular momentum with respect to the $z$-axis,
+$\ell_z \equiv R\rho v_\phi$ (where $v_\phi$ is the azimuthal velocity).
+The equation for the linear $z$-momentum is unchanged. We want to be able
+to solve these equations while still operating in our preferred Cartesian
+coordinate system. \cite{byerly:2014} present these equations as the following;
+for the moment we neglect inclusion of external source terms such as gravity and rotation:
+\begin{align}
+  \frac{\partial(s_R)}{\partial t} + \nabla \cdot (s_R \mathbf{u}) &=
+    - \frac{1}{R}\left(x \frac{\partial}{\partial x} + y \frac{\partial}{\partial y}\right) p
+    + \frac{\ell_z^2}{\rho R^3} \label{eq:radial-momentum}\\
+    \frac{\partial(\ell_z)}{\partial t} + \nabla \cdot (\ell_z \mathbf{u}) &=
+  \left(y\frac{\partial}{\partial x} - x \frac{\partial}{\partial y}\right) p \label{eq:angular-momentum}
+\end{align}  
+Here $\mathbf{u}$ is the ordinary linear momentum used to advect any fluid quantity
+on the grid, and $R = \sqrt{x^2 + y^2}$, where the coordinates $x$ and $y$ are defined
+relative to the axis of rotation, and the origin is located on a zone corner
+so that singularities are avoided. Note that $v_R = (1 / R)(x u + y v)$, and
+$v_\phi = (1/R)(x v - y u)$. These equations can be straightforwardly derived from
+the Euler equations for linear momentum by appropriate multiplication of $x$, $y$,
+$x / R$, and $y/R$, and subsequent algebraic manipulation. We now rewrite this in the
+way that we actually discretize them for use in the code (by use of the product rule
+to rearrange the terms with derivatives in them):
+\begin{align}
+  \frac{\partial(s_R)}{\partial t} &= -\nabla \cdot (s_R \mathbf{u})
+    - \left(\frac{\partial(p (x/R))}{\partial x} + \frac{\partial(p (y/R))}{\partial y}\right) \notag \\
+    &\hspace{0.175in} + \frac{1}{\rho R^3}\left(\ell_z^2 + p\rho R^2\right) \label{eq:radial-momentum-rearranged}\\
+    \frac{\partial(\ell_z)}{\partial t} &= -\nabla \cdot (\ell_z \mathbf{u}) - \left(\frac{\partial (-py)}{\partial x} + \frac{\partial(px)}{\partial y}\right).\label{eq:angular-momentum-rearranged}
+\end{align}
+It is clear that the second term on the right-hand-side in each case can be swept up
+into the flux terms in the same way that the pressure term for the linear momenta
+is often added to the fluxes, to make the conservation form explicit. The remainder
+is the last term on the right-hand-side of the radial momentum equation, which
+exists as a cell-centered source term (which we deal with using a predictor-corrector
+method similar to how we implement external forcing like gravity and rotation; see
+\cite{castro} for details). It is important to observe that in this formulation
+the radial momentum equation is \textit{not} a conserved quantity: it should and
+will in general change with time over the course of the simulation. The above form,
+absent of source terms, makes this clear. The angular momentum term in 
+\autoref{eq:angular-momentum} can be thought of physically as expressing the fact
+that a particle with some initial angular momentum will have its radial momentum
+increase with time as it moves away from the origin, and can be thought of geometrically
+as expressing the curved nature of the underlying coordinates for the radial momentum
+\citep{motl:2002}. Note that while the version in \autoref{eq:radial-momentum-rearranged}
+is more convenient for our hydrodynamics methodology, the pressure-gradient terms
+in the first version are the ones appropriate for a radial field and so no physical meaning
+should be ascribed to the pressure term at the end of \autoref{eq:radial-momentum-rearranged}.
+
+In the absence of source terms, the angular momentum \textit{is} conserved
+to machine precision (ignoring the effects of physical domain boundaries). In the
+presence of source terms, the error in angular momentum conservation is of a
+similar order of magnitude to the error in linear momentum conservation under
+the influence of those source terms (see \autoref{eq:force-angular}). Since this
+error is quite small for gravitational and rotational forces in practice, angular
+momentum conservation is quite good, especially in comparison to the standard
+method of evolving only the linear momentum. Thus this method is most appropriate
+when the net radial momentum is small in comparison to the net angular momentum,
+that is, when the motion is primarily azimuthal in nature.
+
+To implement this method in \castro, we desired an approach that was flexible
+and would leave as little imprint on the code structure as possible.
+To this end, what we have done is to add three new ``hybrid'' momentum state
+variables, corresponding to the radial momentum, angular momentum, and the linear
+momentum component that is perpendicular to both. Whenever we update the normal
+momentum state variables with hydrodynamics fluxes, we add the analogous fluxes
+to the hybrid momenta -- that is, using the edge state values determined by the
+final multi-dimensional Riemann solve, we construct the value of the hybrid
+momenta on that zone edge, and then allow it to be transported as usual by the
+advective velocity $\mathbf{u}$. When we update the normal momenta with an
+external forcing $\mathbf{F}$ such as gravity or rotation, with $x$ and $y$
+components $F_x$ and $F_y$, we apply the update to the hybrid momenta as well:
+\begin{align}
+  \left.\frac{\partial(s_R)}{\partial t}\right|_{F} &= -F_x \frac{x}{R} - F_y\frac{y}{R} \label{eq:force-radial}\\
+  \left.\frac{\partial(\ell_z)}{\partial t}\right|_{F} &= F_x\, y - F_y\, x.\label{eq:force-angular}
+  \end{align}
+Then, when we are in a phase of the evolution when we want to conserve angular
+momentum, we make one additional change: at the end of a (single-level) advance,
+we recompute the normal momenta so that they are fully consistent with the hybrid
+momenta. This is all that is necessary to have a calculation that conserves angular
+momentum over the course of the simulation, without changing any of the
+core infrastructure for the hydrodynamics update. For example, when we compute
+the primitive variables from the conservative variables, we will still be
+getting them from the normal momenta, but these normal momenta are always consistent
+with the conserved angular momentum, so we get the same result as if we had done
+a much more involved update where we explicitly computed the primitive variables
+using the hybrid momenta. There is of course no unique choice for how to do this.
+For example, we could take the approach of explicitly re-calculating the normal
+momenta after any change has been made to the state. This would add significant code
+complexity without a clear benefit. The most significant difference would be
+that when evaluating the new-time value of source terms like rotation, the velocity
+field used is slightly different than it would otherwise be; but this is a high-order
+effect that does not change the core angular momentum conservation property. An
+analysis of the different mechanisms for implementing this update may be an interesting
+avenue for future research --- including various aspects like how to implement the
+source terms in the radial momentum equation. Of course, the answer will likely
+be problem-dependent. Indeed, we already faces this situation in general with the
+gravitational and rotational source terms, even independent of the hybrid momentum
+discussion --- the velocity field seen by the new-time rotation source terms depends
+on whether we apply the gravitational force before or after the rotation force.
 

### 2016-06-18 papers/wdmerger_II/paper.tex
Update paper II description of the integration

@@
 various nuclides; second, once the system of ODEs governing the evolution of
-the nuclear species has been written down, how to append to them ODEs governing
+the nuclear species has been written down, how to couple them to ODEs governing
 the evolution of the thermodynamics, in particular the temperature; third,
@@
 Given a set of nuclides and the reaction links between them, we now consider
-how a burning step is performed in our software. The integration of the stiff,
-implicit system is performed by \vode\ \citep{vode}, a copy of which is
+how a burning step is performed in our software. 
 provided with \castro. To \vode\ we provide an integration state containing
@@
 in a self-heating burn, we allow the temperature to evolve in response
-to the burning (see \citet{maestro3}):
+to the burning (see\footnote{In the cited paper, a term based on the
+thermodynamic chemical potential was included; we now believe
+that it is incorrect to include such a term in the burn, since it
+automatically sums to zero analytically.} \citet{maestro3}):
 \begin{equation}
-  \frac{dT}{dt} = \frac{1}{c_V}\left(\frac{\partial e}{\partial t} - \sum_n \frac{\partial e}{\partial X_n}\frac{\partial X_n}{\partial t}\right)
+  \frac{dT}{dt} = \frac{1}{c_V}\frac{\partial e}{\partial t}
 \end{equation}
-Here $c_V$ is the specific heat at constant volume, and $(\partial e/\partial X_n)$
-is the derivative of the internal energy with respect to the mass fraction of a given
-species (with other thermodynamic variables held constant). Both quantities are provided by the
-equation of state. During this burn, we can keep these thermodynamic
-derivatives constant using their value at the beginning of the run, or at each step we
+(Although $T$ evolves during the burn so that the integration is physically
+accurate, as for the hydrostatic method we discard the final value
+for $T$ at the end of the burn and recompute a temperature for the zone that is
+consistent with its new internal energy.) Here $c_V$ is the specific heat at
+constant volume, which is provided by the equation of state.  During this burn,
+we can keep $c_V$ constant using its initial value, or at each step we
 can choose to re-evaluate the equation of state using the latest value of $(\rho, T)$.
 The latter is more expensive but also more accurate, and we use it in this paper.
+In practice we find that the cost is negligible in comparison to the more expensive
+parts of the calculation, and it can significantly speed up convergence near NSE.
 A third option presented by \citet{raskin:2010} is a so-called ``hybrid'' mode.
@@
 energy change is negative, we do the burn again in self-heating mode. All three options
-are implemented in our burner software.
+are implemented in our burner software. The simulations shown in this work all
+use the self-heating mode unless otherwise specified.
+
+In our \microphysics\ repository we provide several software options for
+solving a set of coupled stiff ODEs. For this work we use an implementation of
+the well known variable-order Richardson extrapolation method presented by Stoer and
+Bulirsch \cite{stoer:1980}, that is similar to the integrator which ships with
+the original versions of the networks mentioned above. Previous work using the \boxlib\
+codes typically used the classic \vode\ integrator \cite{vode}. We do maintain a version
+of the software that is compatible with our software interfaces in the \microphysics\
+repository, but we have largely shifted to integrators which are written in modern
+Fortran as a consequence of our efforts to run our codes on GPUs. The Stoer and Bulirsch
+integrator we provide satisfies this criterion; we also provide a rewrite of the VODE
+BDF algorithm that uses modern Fortran features. The Stoer and Bulirsch algorithm
+relies on a uniform relative error tolerance for all of the ODEs in the system, which
+we set at $10^{-6}$ for all simulations described here.
 
@@
 the above being method 4. Method 3 is similar to method 4 but replaces the
-denominator in Equation \ref{eq:burning_mode_4} with the change in 
+denominator in \autoref{eq:burning_mode_4} with the change in 
 internal energy over the last timestep only from the nuclear reactions.
@@
 in the context of low Mach number flows, \cite{nonaka:2012} have
-used the method of spectral deferred corrections \cite{SDC} to
+used the method of spectral deferred corrections \citep{SDC} to
 couple their advection-diffusion-reaction equation set. In our
@@
 over the timestep. However, this can assumption can fail
-quite radically during periods of runaway burning when
-the rate of change of these quantities is highly nonlinear.
-We do not want to neglect the errors caused by this approximation
+during periods of runaway burning when the rate of change
+of these quantities is highly nonlinear. We may not want
+to neglect the errors caused by this approximation
 because they may build up over an extended period of nonlinear
@@
 To this end, we have implemented a timestep retry option in
-\castro. The retry option works by first saving a copy of the
-full state, and then doing the advance. We then explicitly
-check whether the timestep criteria were violated during the advance,
-using the time-centered energy and mass fractions and a
-time-centered differencing of the old and new states to
-calculate the evolution timescale. If so, we reject this
-advance and restore the original state data. We then perform
-a subcycled update, with the maximum timestep chosen so that
-the criterion is satisfied in each subcycle. Note that this
-occurs entirely on a single level and is separate from the
-subcycling that is used by \boxlib\ to advance the different
-AMR levels in time; the full AMR algorithm does not ever
-know that the retry subcycling took place. This can help
-substantially to mollify the problem described above, where
-the time derivative estimation methods proposed were not
-sufficient to obtain the desired limiting. Thus we strongly
-recommend that hydrodynamics codes implement such an approach,
-which ensures that the coupling actually does stay at the
-level requested. The most significant drawback is that saving
-a copy of the old state effectively doubles the memory footprint
-of the code, so this option should be disabled when not needed.
-However, it may be crucial in determining the evolution of
-systems such as white dwarf collisions, where even a short
-period of incorrect advances can significantly pollute the
-final answer, and the additional memory expense is justified
-by obtaining a more accurate solution.
+\castro, which re-computes an advance if it violated the
+stability criteria as judged from the end of the timestep.
+However we have found that the benefits for this problem are
+small and thus we do not use it for the simulations here.
 
@@
 However, this level of accuracy may make a parameter study of
-the problem unfeasible, given current computational resources.
+the problem challenging given current computational resources.
+When resources are limited, one can still use this method and set a
+limit on the number of levels of refinement possible in the simulation,
+so that the refinement will still go in the right places but may not
+refine all of the way to the regime of guaranteed stability.
 

### 2016-06-19 papers/wdmerger_II/paper.tex
Add notes on problem setup to Paper II

@@
 
-The hydrodynamical equations solved in this paper were presented in \citet{wdmergerI}.
+The hydrodynamical equations solved in this paper were presented in \citet{wdmergerI} (hereafter, Paper I).
 In this paper we focus our discussion on our implementation of the nuclear
@@
 Given a set of nuclides and the reaction links between them, we now consider
-how a burning step is performed in our software. 
-provided with \castro. To \vode\ we provide an integration state containing
-the molar abundances $Y_{n} = X_{n} / A_{n}$, where $X$ is the mass fraction
-of the abundance in the zone and $A$ is the number of nucleons. The integration
-of
+how a burning step is performed in our software. The goal is to integrate the
+vector ${\bm{Y}} = (Y_1, Y_2, \ldots, Y_n, e, T)$, where $Y_{n} = X_{n} / A_{n}$
+is the molar fraction of species $n$, with $X_n$ the mass fraction and $A_n$ the
+mass number of that species, $e$ is the energy released during the burn, and
+$T$ is the temperature. The equation describing its evolution is given by
 \begin{equation}
-  \frac{d\bm{Y}}{dt} = f(\mathbf{Y})
+  \frac{d\bm{Y}}{dt} = f(\mathbf{Y}),
 \end{equation}
-is determined using a right-hand-side provided by the nuclear burning network.
-When it comes to the thermodynamical evolution, the internal energy $e$ of the zone
-will change when the nuclear abundances evolve. During the integration we track
-the energy release from the system using
+where the components of the right-hand-side for the species come from the particular
+nuclear burning network we are using. The energy release $e$ of the zone's burn
+will change when the nuclear abundances evolve, according to
 \begin{equation}
@@
 \subsection{Numerically Unstable Burning}
+\label{sec:unstable_burning}
 
@@
 We implement the white dwarf collision problem in \castro\ using our \wdmerger\
-package. The white dwarf centers of mass are separated by a distance of four times
-the (secondary) white dwarf radius. Their initial velocity is that of
-two point-masses in free-fall towards each other, at the appropriate distance.
+package. We load the grid in the way most previous papers have done so:
+the white dwarf centers of mass are initially separated by a distance of four times
+the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
+consequently the initial distance, depends on the equation of state used, and this
+can vary somewhat depending on the relevant physics used, such as Coulomb corrections
+in the Helmholtz equation of state. However, the results do not really depend on the
+exact initial distance as long as there is enough time for the WDs to distort in
+response to tidal forces as they approach. Their initial velocity is that of
+two point-masses in free-fall towards each other coming in from infinity, such that
+the contact point is at the origin, and they approach each other on the $x$-axis.
+The composition of the WDs is as described in Paper I: low-mass WDs are helium-dominated,
+intermediate-mass WDs are carbon and oxygen dominated, and high-mass WDs contain
+oxygen, neon, and magnesium. By default the WDs approach each other head-on, with
+their respective centers lying on the $x$-axis, but we have also added an option
+to have the WDs approach each other at a non-zero impact parameter $b$, where the
+offset is perpendicular to the $x$-axis in the $y$ direction. We measure
+$b$ in units of the secondary WD's radius, so that the default head-on case
+corresponds to $b = 0$ and the case where the two WDs just graze past each other
+corresponds to $b = 1$ (in practice it happens slightly differently because
+of the WDs responding to each other's gravitational fields).
+
+For all the 3D simulations to follow, we use the same coarse grid as in Paper I:
+there are 256 zones per dimension, and the width of each axis is $1.024 \times 10^{10}$
+cm, yielding a base resolution of 400 km. When we use adaptive-mesh refinement, we will
+use the refinement scheme described in \autoref{sec:unstable_burning}. This is
+acceptable because the outcome of the collision problem is chiefly a function
+of the propagation of burning front through the WDs, so for this problem one
+can take some liberties in other parts of the algorithm for the purpose of
+computational efficiency, which can always be turned back on later if a more
+accurate answer is desired. So, for example, we do not need to add refinement
+everywhere in the WDs in the interest of getting the collision dynamics to be
+slightly more accurate. Additionally, we choose to solve the Poisson equation
+for the self-gravity of the system only on the coarse grid. The gravitational
+potential and acceleration on the finer levels are interpolated from the coarse
+grid.
+
+We have also added to \wdmerger\ the ability to use the 2D cylindrical ($R-z$)
+coordinate system evolution in \castro. In this coordinate system, we align
+the WDs along the $z$-axis (which is analogous to our $x$-axis in the Cartesian
+evolution) at the usual distance, with the center of the WDs at $R = 0$. The
+un-simulated $\phi$ dimension then would extend the WDs through a $2\pi$ revolution.
+The nature of the axisymmetry inherent to the cylindrical coordinate system
+means that we can only run head-on collisions with $b = 0$. The coarse grid
+is the same resolution as the 3D case: the width of the $z$ axis is $1.024 \times 10^{10}$
+cm and the width of the $R$ axis is $5.012 \times 10^{10}$ cm, with twice as
+many zones along the $z$ axis as the $R$ axis so that the equal 400 km
+resolution is maintained.
 
@@
 
-Though the main focus of this work is fully three-dimensional simulations,
-there is much to be learned from simpler, and much computationally cheaper,
-two-dimensional axisymmetric collision simulations. These are implemented
-in a cylindrical ($R-z$) coordinate system. We use these for three purposes.
-First, in the regime where the 2D and 3D simulations can be usefully compared, 
-zero-impact parameter simulations, to understand to what extent the
-reduced-dimensionality approach can reproduce the 3D results and thus to
-conclude whether 2D simulations can be used for obtaining scientific data
-such as nucleosynthesis yields. Second, for testing the effect of various
-code choices such as the number of isotopes in the nuclear network. A third
-effect which comes for free with the others is that we can benchmark existing
-and new code tools such as our nuclear postprocessing analysis. Prepared
-with these choices and a better understanding of how reliable our approximations
-are, we can then proceed to performing 3D simulations with our code choices
-already made and with confidence in our analysis software.
+While white dwarf collisions are three-dimensional events, there is much to
+be learned from simpler, and much computationally cheaper, two-dimensional
+axisymmetric collision simulations. These are implemented in the cylindrical
+coordinate system described in \autoref{sec:problemsetup}.
+We simulate collisions in 2D cylindrical coordinates for three purposes.
+First, we can quickly get up to speed on the nature of the problem and then
+benchmark existing and new code tools such as our nuclear postprocessing analysis.
+Second, in the regime where the 2D and 3D simulations can be usefully compared, 
+zero-impact parameter head-on simulations ($b = 0$), we can understand to
+what extent the reduced-dimensionality approach can reproduce the 3D results
+and thus determine whether 2D simulations can be used for obtaining scientific
+predictions, including nucleosynthesis yields and observable quantities.
+Third, we can cheaply test the effect of various code choices such as the number
+of isotopes in the nuclear network; we can afford to run the same problem many
+times while tweaking parameters one by one to learn their effects. 
+Prepared with these results and a better understanding of how reliable our
+approximations are, we can then proceed to performing 3D simulations with
+our code choices already made and with confidence that we understand the
+uncertainty resulting from these choices.
 
@@
 First, is the time it takes for the white dwarfs to collide consistent with
-basic kinematics. Second, is the amount of energy released in the
+basic kinematics? Second, is the amount of energy released in the
 collision by the nuclear reactions comparable to the amount of kinetic
-energy we infer from observations of Type Ia supernovae?
+energy we infer from observations of Type Ia supernovae? We must ensure
+that we are getting these basics right before we can advance to more
+complicated questions. This is not necessarily a trivial matter. We have
+observed, for example, that at low resolution, the WDs take longer to
+collide than at higher resolution, because the gravitational field is
+much less accurate.
 

### 2016-06-28 analysis/wdmerger.py
CASTRO_DIR -> CASTRO_HOME

@@ -26,8 +26,8 @@ def get_wdmerger_git_commit_hash():
 #
 
 def get_castro_dir():
-    CASTRO_DIR = os.getenv('CASTRO_DIR')
-    return CASTRO_DIR
+    CASTRO_HOME = os.getenv('CASTRO_HOME')
+    return CASTRO_HOME
 
 
 
@@ -596,9 +596,9 @@ def get_star_locs(plotfile):
 
 def get_castro_const(var_name):
 
-    CASTRO_DIR = get_castro_dir()
+    CASTRO_HOME = get_castro_dir()
 
-    file  = open(CASTRO_DIR + '/constants/constants_cgs.f90')
+    file  = open(CASTRO_HOME + '/constants/constants_cgs.f90')
     lines = file.readlines()
 
     const = None

### 2016-06-28 science/wdmerger_I/castro/evrard_collapse/run_test.sh

@@ -1,6 +1,6 @@
 source $WDMERGER_HOME/job_scripts/run_utils.sh
 
-TEST_DIR=$CASTRO_DIR/Exec/evrard_collapse
+TEST_DIR=$CASTRO_HOME/Exec/evrard_collapse
 
 cp $TEST_DIR/inputs $compile_dir/
 cp $TEST_DIR/probin $compile_dir/

### 2016-06-28 documentation/userguide.tex
Update bibliography file reference in the documentation

@@
 table for the equation of state, as well as the actual \texttt{CASTRO} executable you just built in the 
-\texttt{compile/} directory. Finally, it builds a script that you'll use for actually submitting 
-the job to the batch submission system on the machine you're on, and then submits that job. 
-Thus for a typical setup, your run directory will contain the following files when you submit:
+\texttt{compile/} directory, and a copy of the run utilities that manage the job itself (so that the 
+run doesn't change if you make local edits to the job management system). Finally, it builds a script 
+that you'll use for actually submitting the job to the batch submission system on the machine you're on, 
+and then submits that job. Thus for a typical setup, your run directory will contain at least the following 
+files when you submit:
 \begin{itemize}
@@
   \item \texttt{run\_script}
+  \item \texttt{job\_scripts/}
 \end{itemize}
+Note that if you want to set up this directory as above but \textit{not} actually submit the job, 
+set the \texttt{no\_submit} variable to any value in your run script. Similarly, if you want to 
+set up the directory to contain the inputs and probin file but none of the machine-dependent 
+scripts or the executable, set the \texttt{inputs\_only} variable to any value in your run script.
+
 Typically you'll want to run multiple jobs at once, that vary things like the number of processors 
@@
 
-\bibliographystyle{../papers/apj}
+\bibliographystyle{../papers/aasjournal}
 \bibliography{../papers/refs}

### 2016-06-28 papers/wdmerger_II/paper.tex

@@
 
+The simulation is terminated when the total energy is positive, indicating that
+the system has become unbound due to nuclear energy release, and has been
+constant or decreasing for the last five timesteps, indicating that no further
+meaningful nuclear energy generation is occurring and material is now beginning
+to stream off the simulation domain. As all simulations in this paper are of
+WD collisions that generate enough energy to unbind the system, this stopping
+criterion is applied throughout all simulations shown here.
+
 %==========================================================================

### 2016-07-08 papers/wdmerger_II/paper.tex
Add a discussion of the suppressed burning mode

@@
 In this mode, by default we do a hydrostatic burn. If that burn fails, or if the net
-energy change is negative, we do the burn again in self-heating mode. All three options
+energy change is negative, we do the burn again in self-heating mode. A final option
+is a burn that limits the changes due to a burn to avoid numerically unstable burning.
+This mode is discussed in \autoref{sec:unstable_burning} and which we will call a
+``suppressed'' burn for the remainder of this work. All four options
 are implemented in our burner software. The simulations shown in this work all
@@
 $t_e$ is shorter than the sound-crossing time $t_s$ in a zone.
+When the sound-crossing time is too long, energy is built up in
+a zone faster than it can be advected away by pressure waves.
+This is obviously a problem inherent to numerically discretized
+systems as the underlying fluid equations are continuous.
+This can lead to a numerically seeded detonation caused by the
+temperature building up too quickly in the zone; the detonation
+is spurious in this case and should be avoided if possible.
 The goal is to ensure that the following condition holds:
@@
 \citet{kushnir:2013} implemented this criterion by artificially 
-limiting the magnitude of the energy release after a burn. As 
-this is not physical, we take a different approach. Since we 
+limiting the magnitude of the energy release after a burn. We
+too have developed an option for our burner to do this,
+the ``suppressed'' burning mode. In a suppressed burn, we limit
+the changes to the state so that \autoref{eq:burning_limiter_2}
+is always satisfied. To achieve this we directly multiply the
+right-hand-side vector in the integration by a constant factor $F$
+for all variables, where $F$ is the multiplicative factor needed to
+be applied to $\dot{e}$ such that the equality in \autoref{eq:burning_limiter_2}
+holds. (If the inequality is already satisfied, then the integration
+vector is not modified.) We fix $t_s$ to be the value of the sound
+crossing time at the beginning of the burn (that is, we do not
+update it as the sound speed changes) and we fix the energy $e$
+that goes into the estimate for $t_e$ to be the value of the
+internal energy of the zone at the beginning of the burn. If
+instead one allowed $c_s$ and $e$ to evolve with the burn, one
+would obtain a less conservative limiter in the case of explosive
+burning, as $c_s$ and $e$ are both increasing in this case.
+As the point of the limiter is to ensure that the changes to the
+\textit{original} energy are small enough so that the following
+hydrodynamics update can advect away newly generated energy
+quickly enough to avoid a nuerically seeded detonation,
+we desire the most conservative version of the limiter. We discuss
+our results with the suppressed burning mode in \autoref{sec:2D}.
+
+Regardless of whether the suppressed burning mode works for this
+particular problem, it is not physical, so in general we have
+taken a different approach. If we insist that we cannot
 cannot directly control the energy injection timescale, we 

### 2016-07-09 papers/wdmerger_II/paper.tex
Begin the introduction of paper II

@@
 \label{sec:introduction}
+
+Binary white dwarf (WD) systems have long been thought to be potential progenitors
+of Type Ia supernovae (SNe Ia) \citep{ibentutukov:1984,webbink:1984}. Recently there
+has been a growing amount of observational evidence that double WD systems play
+an important role in explaining at least some SNe Ia \citep{maoz:2014}. These systems
+may be isolated or they may be in hierarchical triple systems, where a WD binary
+system in a tight inner orbit is gravitationally coupled to a third, outer star.
+In either case the binary WD pair can merge due to the emissions of gravitational waves,
+but in the hierarchical triple case, dynamical interactions between the WD binary and
+the outer star can prompt mergers as well \citep{thompson:2011,hamers:2013}. The mergers
+occur due to perturbations in the eccentricity of the binary orbit caused by the outer
+star, which significantly increase the gravitational wave emission of the system. In a
+subset of these mergers the eccentricity may be driven to such a high value that the
+merger resembles a head-on collision of the binary WDs. Whether the rate of these collisions
+may be high enough to significantly contribute to the observed SN Ia rate is uncertain,
+though \cite{katzdong:2012} suggest that this is plausible.
+
+Over the past few years a number of groups have performed simulations of WD collisions
+to understand whether they may yield astrophysical transients that look like SNe Ia
+\citep{rosswog:2009,raskin:2010,loren-aguilar:2010,hawley:2012,garcia-senz:2013,
+kushnir:2013,papish:2015,holcomb:2015}. Head-on collisions rapidly convert a significant
+amount of kinetic energy into thermal energy in a small region and thus set up the
+conditions for a thermonuclear detonation, and these previous simulations have indeed
+found that detonations occur and convert a large amount of carbon/oxygen material into
+iron-group elements.
 

### 2016-07-11 papers/wdmerger_II/paper.tex
Add some data to paper II

@@
 \author{Weiqun Zhang\altaffilmark{3}}
+\author{Frank X. Timmes\altaffilmark{4,5}}
 
@@
   Lawrence Berkeley National Laboratory, Berkeley, CA 94720
+}
+
+\altaffiltext{4}
+{
+  The Joint Institute for Nuclear Astrophysics, USA
+}
+
+\altaffiltext{5}
+{
+  School of Earth and Space Exploration, Arizona State University, Tempe, AZ, USA
 }
@@
 \begin{align}
-  \Delta t_{be} &= f_{e}\, \tau_e \\
-  \Delta t_{bX_k} &= f_{X}\, \tau_{X_k}.
+  \Delta t_{be} &= f_{e}\, \tau_e \label{eq:timestep_e}\\
+  \Delta t_{bX_k} &= f_{X}\, \tau_{X_k}. \label{eq:timestep_X}
 \end{align}
@@
 not on cahnges in species.)  The sensitivity of results to this choice will be discussed in 
-\autoref{sec:2D:timestep}. The factors $f_{e}$ and $f_{X}$ can be set at runtime in \castro.
+\autoref{sec:parameters:timestepping}. The factors $f_{e}$ and $f_{X}$ can be set at runtime in \castro.
 
@@
 we desire the most conservative version of the limiter. We discuss
-our results with the suppressed burning mode in \autoref{sec:2D}.
+our results with the suppressed burning mode in \autoref{sec:parameters}.
 
@@
 potential and acceleration on the finer levels are interpolated from the coarse
-grid.
+grid. This grants enhanced computational speed without a serious effect on the
+accuracy of the simulation (for the case where the fine grids cover the stars,
+we have checked that the more accurate gravitational forcing would only modify
+slightly the time to impact and the subsequent detonation process).
 
@@
 
+As in Paper I, the region outside the WDs is filled with a low-density ambient
+material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition is the
+same as that of the WDs (which is equal carbon and oxygen by mass, unless
+stated otherwise). Reactions in this region are unimportant, so for
+computational efficiency, we disable all nuclear burning for zones that have
+$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. The initial ambient
+temperature is $10^7\ \text{K}$, which is the same as the initial temperature
+everywhere throughout the WDs, and we set the temperature floor to the same
+value (any zone that reaches below this temperature will be reset).
+
+\castro\ uses the dual energy formalism of \cite{bryan:1995,ENZO}, where we evolve equations
+for both the total energy $E$ and the internal energy $e$, using one or the other to
+determine the pressure in the hydrodynamics step (if we are using the total energy, we
+compute the corresponding internal energy by subtracting the kinetic energy $K$). The switching
+parameter is $\eta_1 = 10^{-3}$, and if $(E - K) / E < \eta_1$, we use the internal energy variable $e$
+in the equation of state call that obtains the pressure. Otherwise we use $(E - K)$. The
+idea behind this is that in high Mach number flows where the kinetic energy dominates the
+total energy, we do not want to risk generating a negative estimate of the internal energy
+due to roundoff error. Similarly, at regular intervals when we enforce consistency of the
+internal energy variable (which occurs a few times per timestep), we sync the internal energy
+variable so that it is equal to $E - K$, skipping this step only if $(E - K) / E < \eta_2$.
+We adopt by default $\eta_2 = 10^{-4}$. This choice can matter much more for reacting
+hydrodynamics than non-reacting hydrodynamics. For example, a collision can only lead to a
+detonation if there is sufficient conversion of kinetic to thermal energy. Obviously
+this happens physically, but if the parameter $\eta_2$ is set high enough then it could
+in principle interfere by not allowing the simulation to efficiently convert kinetic energy
+to thermal energy. (This does not end up mattering much for our simulations because the
+kinetic energy does not strongly dominate the internal energy for colliding WDs.) And on
+a similar note, when using this formalism it is ambiguous which variable to use ($(E - K)$
+or $e$) when calculating the initial temperature that goes into the reaction step. We
+have introduced a new parameter $\eta_3$ that uses the same logic as the first two: if
+$(E - K) / E < \eta_3$, we use $(E - K)$ in the equation of state call that obtains the
+temperature; otherwise, we use $e$. However, unlike the other two, the \castro\ default
+is $\eta_3 = 1.0$, meaning that we normally use the internal energy variable.
+
 %==========================================================================
-% 2D simulations
+% Parameter Study
 %==========================================================================
-\section{Two-Dimensional Simulations}
-\label{sec:2D}
+\section{Parameter Study}
+\label{sec:parameters}
 
-While white dwarf collisions are three-dimensional events, there is much to
-be learned from simpler, and much computationally cheaper, two-dimensional
-axisymmetric collision simulations. These are implemented in the cylindrical
-coordinate system described in \autoref{sec:problemsetup}.
-We simulate collisions in 2D cylindrical coordinates for three purposes.
-First, we can quickly get up to speed on the nature of the problem and then
-benchmark existing and new code tools such as our nuclear postprocessing analysis.
-Second, in the regime where the 2D and 3D simulations can be usefully compared, 
-zero-impact parameter head-on simulations ($b = 0$), we can understand to
-what extent the reduced-dimensionality approach can reproduce the 3D results
-and thus determine whether 2D simulations can be used for obtaining scientific
-predictions, including nucleosynthesis yields and observable quantities.
-Third, we can cheaply test the effect of various code choices such as the number
-of isotopes in the nuclear network; we can afford to run the same problem many
-times while tweaking parameters one by one to learn their effects. 
-Prepared with these results and a better understanding of how reliable our
-approximations are, we can then proceed to performing 3D simulations with
-our code choices already made and with confidence that we understand the
-uncertainty resulting from these choices.
+Before turning to the effect that adaptive mesh refinement has on WD collisions,
+we will first examine the effect of a number of other code parameters. For most
+of the following we will use two-dimensional axisymmetric simulations which have
+impact parameter $b = 0$, unless otherwise specified, and we will stick to a
+moderate resolution (we use only the coarse grid described in \autoref{sec:problemsetup}).
+For all simulations in this section, we use a pair of $0.64\ \msolar$ WDs, a binary
+system that has been studied in most of the collision papers to date. This allows
+us to cheaply test how the collision responds to many effects. One limitation of
+this is that we will not find out exactly how a simulation differs in the case of a
+high-resolution 3D run when these parameters changed. Another is that we will not
+perform a full parameter study where we consider the full non-linear dependence of
+all parameters on all other parameters; here we are merely interested in qualitative
+trends that will help to understand the limitations of WD collision simulations.
 
-\subsection{Dynamics}
-\label{sec:2D:dynamics}
+With those caveats in mind, let us consider the tests and the results we obtained.
+Our primary metric for the following tests will be the amount of $^{56}$Ni generated
+in the collision, as this is the parameter most directly related to the observable
+quantities of interest for Type Ia supernovae. Specifically, we collect information
+at the end of every timestep about the total amount of nickel on the grid (in solar
+masses), and we will use the maximum value of this nickel mass. When the stopping
+criterion described in \autoref{sec:problemsetup} is used, this nickel mass is generally
+constant or decreasing by the end of the run. The typical amount of nickel mass generated
+for our runs at low resolution is $0.2\ \msolar$, which is significantly lower than
+that obtained by \cite{raskin:2010} and \cite{kushnir:2013} for the same problem.
+We will discuss this in \autoref{sec:resolution}.
 
-First we briefly consider whether the dynamical evolution of the system
-is consistent with simple physics we understand. We ask two questions.
-First, is the time it takes for the white dwarfs to collide consistent with
-basic kinematics? Second, is the amount of energy released in the
-collision by the nuclear reactions comparable to the amount of kinetic
-energy we infer from observations of Type Ia supernovae? We must ensure
-that we are getting these basics right before we can advance to more
-complicated questions. This is not necessarily a trivial matter. We have
-observed, for example, that at low resolution, the WDs take longer to
-collide than at higher resolution, because the gravitational field is
-much less accurate.
+\subsection{Timestepping}
+\label{sec:parameters:timestepping}
 
-\subsection{Timestep Limiting}
-\label{sec:2D:timestep}
+For the standard self-heating burns that we use, the most important parameters we
+examined come from the timestep limiting scheme described in \autoref{sec:hydrocoupling}.
+The parameters $f_{be}$ (from \autoref{eq:timestep_e}) and $f_{bX}$ (from
+\autoref{eq:timestep_X}) control the size of the timestep as a function of the
+burning rate. As the timestep gets smaller, the error due to the Strang splitting
+scheme coupling the reactions and the burning also gets smaller, and we expect
+that the results become more accurate.
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
+  \caption{Nickel production as a function of the timestep factors $f_{be}$
+           (solid blue, limiting the timestep on changes in internal energy) and
+           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
+           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
+           the timestep by at least an order of magnitude more.
+           \label{fig:timestep_factor}}
+\end{figure}
+\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
+production. As either $f$ becomes small, the restriction on the timestep becomes quite
+severe: for $f_{bX} = 0.01$, the minimum timestep in the simulation is smaller than
+$2 \times 10^{-9}\ \text{s}$ and the full simulation requires over 750,000 timesteps.
+Yet it is clear that even for this steep value of the limiter, which is much stricter
+than the limiter used in, e.g., \cite{hawley:2012} and \cite{raskin:2010}, the
+simulation has not fully converged in nickel production. These limiters are somewhat
+redundant in the sense that a given $f_{bX}$ corresponds to another (higher) $f_{be}$,
+as is evident by the fact that on the graph the curves look similar in shape but with
+a horizontal offset.
 
-Here we discuss the effect of limiting the hydrodynamic timestep based on
-changes in the internal energy in the nuclear reaction network.
+We also examined the effect of the method used to enforce the limiting, using the
+four limiting modes discussed in \autoref{sec:hydrocoupling}. The results are shown
+in \autoref{table:burninglimitermode} for the case $f_{be} = 0.3$ (we did not use
+limiting with $f_{bX}$ for this test). The two modes that use the full last timestep's
+worth of information to estimate $\dot{e}$ yield a slightly larger nickel mass than the
+two modes that only use more recent information, but the difference is negligible.
+
+\input{plots/burning_limiter_mode_m_P_0.64_m_S_0.64.tbl}
+
+%TODO: add a ``Nonaka plot''
+
+For all remaining tests in this section, unless otherwise stated, we take $f_{be} = 10^2$
+and disable limiting with $f_{bX}$ by setting it to a large number. This choice is a balance
+between accuracy and efficiency: we avoid the obviously wrong answers that occur for $f_{be} > 10^3$
+while ensuring that the timesteps are long enough that the following tests are computationally
+inexpensive.
 
 \subsection{Nuclear Network}
-\label{sec:2D:network}
+\label{sec:parameters:network}
 
 Now we move to a discussion of the effect of including various isotopes
-in the nuclear reaction network. In particular we compare the \aproxthirteen,
-\aproxnineteen, and \aproxtwentyone reaction networks.
+in the nuclear reaction network. In particular we compare the \isoseven, \aproxthirteen,
+\aproxnineteen, and \aproxtwentyone\ reaction networks. In all cases we run
+the same problem setup, and the unused isotopes (everything but carbon and oxygen)
+have their mass fractions in every zone set to a negligibly small number at initialization.
 
-\subsection{Nucleosynthetic Postprocessing}
-\label{sec:2D:postprocessing}
+\subsection{Burning Mode}
+\label{sec:parameters:burningmode}
 
-We consider also the results of our postprocessing step for nucleosynthesis
-yields, to understand whether the results for the isotopes not included in
-our networks make sense.
+\subsection{Other Parameters}
+
+We tested a few other parameters that ended up having no serious effect on the outcome.
+The dual energy parameter $\eta_2$ described in \autoref{sec:problemsetup} can have a
+minor effect on the nickel production, yielding variations on the order of a few percent,
+but there is no clear pattern. The dual energy parameter $\eta_3$ has no effect on the
+outcome: the answer is the same for $\eta_3 = 0$ and $\eta_3 = 1$, which makes sense
+because all of the reacting regions have a low enough kinetic energy relative to their
+total energy such that the internal energy variable is always kept in sync with $(E - K)$.
+(We did not look at $\eta_1$ because that has a direct effect only on the hydrodynamics,
+not the reactions.) We also tested whether our choice for the temperature floor mattered
+by lowering it to $10^6\ \text{K}$ and $10^5\ \text{K}$, and this had no effect on nickel
+production. It does allow parts of the WDs to become colder as they move, due to numerical
+errors in the advecting flow (the Helmholtz equation of state is very insensitive to
+temperature, so even small changes in the internal energy can create very large changes
+in the implied temperature), but the amount of energy injected by the collision is so
+large that the final result is insensitive to whatever the initial temperature of the
+WDs were. We also checked whether there was any consequence of our choice to disable the
+burning for small enough $\rho$ and $T$. Lowering the minimum $\rho$ for reactions to
+$10^0\ \text{g / cm}^{-3}$ from $10^6\ \text{g / cm}^{-3}$ had a negligible effect,
+and varying the minimum $T$ for reactions from $2 \times 10^{7}\ \text{K}$ to
+$4 \times 10^{8}\ \text{K}$ also had no meaningful effect.
+
+%TODO: maybe show an image of the temperature field in the WDs before colliding to demonstrate the point about numerical error?
+
+The choice of the WD composition to be equal carbon and oxygen by mass is a common
+choice in the literature and is broadly comparable to what stellar evolution codes
+yield for WDs of this mass. Still, it is an arbitrary approximation, especially the
+insistence that the composition is uniform throughout the WD. In \autoref{table:co}
+we list the nickel production as a function of the initial carbon mass fraction.
+This can have a few percent impact on the outcome for the plausible values we tested.
+In future work we may look at the effect of a more realistic progenitor structure,
+including the effects of a helium shell on the surface of the WDs \citep{holcomb:2015}
+and a non-uniform interior structure.
+
+\input{plots/co_frac_m_P_0.64_m_S_0.64.tbl}
 
 %==========================================================================
-% 3D simulations
+% Resolution Dependence
 %==========================================================================
-\section{Three-Dimensional Simulations}
-\label{sec:3D}
+\section{Resolution Dependence}
+\label{sec:resolution}
 
-Now that we have analyzed the various parts of our simulation software and
-discussed our code choices, we consider full 3D simulations to understand
-the potential for white dwarf collisions to be progenitors of Type Ia
-supernovae. We focus on two main quantities of interest: the total amount of
-nickel produced in the explosion, which directly powers the light curve
-observed in SNe Ia; and, the gravitational wave signature produced by
-these events.
+%==========================================================================
+% Gravitational Wave Signature
+%==========================================================================
+\section{Gravitational Wave Signature}
+\label{sec:gwsignature}
 

### 2016-07-12 papers/wdmerger_II/paper.tex
Add to the section on the burning modes

@@
 is a burn that limits the changes due to a burn to avoid numerically unstable burning.
-This mode is discussed in \autoref{sec:unstable_burning} and which we will call a
+This mode is discussed in \autoref{sec:unstable_burning} and we will call it a
 ``suppressed'' burn for the remainder of this work. All four options
@@
 \begin{equation}
-  \dot{e} = \frac{e^{n} - e^{n-1}}{\Delta t^{n-1}}, \label{eq:burning_mode_4}
+  \dot{e} = \frac{e^{n} - e^{n-1}}{\Delta t^{n-1}}, \label{eq:burning_limiter_mode_4}
 \end{equation}
@@
 We define four separate methods for calculating the time derivative, with 
-the above being method 4. Method 3 is similar to method 4 but replaces the
-denominator in \autoref{eq:burning_mode_4} with the change in 
+the above being mode 4. Mode 3 is similar to mode 4 but replaces the
+denominator in \autoref{eq:burning_limiter_mode_4} with the change in 
 internal energy over the last timestep only from the nuclear reactions.
-Method 2 is the same as method 3 but we only use the change in internal 
+Mode 2 is the same as mode 3 but we only use the change in internal 
 energy from the most recent nuclear burning step, that is, the second-half
 of the Strang-split burning from the last timestep (the denominator 
-then becomes $\Delta t / 2$). In method 1, the most accurate option and 
+then becomes $\Delta t / 2$). In mode 1, the most accurate option and 
 the current default in \castro, we evaluate the right-hand-side of the 
@@
 of the burning rates will be small again, and the instantaneous 
-timestep limiter of method 1 may actually substantially overestimate 
-the needed timestep. The other methods will see that the energy/species  
+timestep limiter of mode 1 may actually substantially overestimate 
+the needed timestep. The other modes will see that the energy/species  
 substantially changed over the last timestep, but will still 
@@
 we desire the most conservative version of the limiter. We discuss
-our results with the suppressed burning mode in \autoref{sec:parameters}.
+our results with the suppressed burning mode in \autoref{sec:parameters:burningmode}.
+Note that we have found that with this option enabled, it typically takes
+many more timesteps to complete a burn than in self-heating mode.
 
@@
 
+In \autoref{sec:burner} we observed that there are several alternatives to the traditional
+self-heating approach in a burn. These alternatives are relevant to a collision problem because,
+as pointed out by \cite{raskin:2010} and \cite{kushnir:2013}, it is possible for a typical
+zone in our simulation to release a very large amount of energy in a burn before it is cooled
+off by a subsequent hydrodynamic expansion step, leading to the numerically unstable burning
+problem described in \autoref{sec:unstable_burning}. \autoref{table:burningmode} lists nickel
+production for the four burning modes described earlier. The hydrostatic burn does produces
+slightly more nickel than the self-heating burn, consistent with the idea that by limiting the
+energy release from early carbon burning, we can forestall a detonation which is numerically
+spurious. Yet the difference is only about 5\%. The reason is that the hydrostatic burn controls
+the energy release during the burn, but \textit{not} what happens in the rest of the hydrodynamics
+step. And in the hydrodynamics step, the \castro\ algorithm performs multiple EOS calls to ensure
+that the temperature is synchronized with the internal energy at various points in the update.
+So in practice there is a temperature change due to the burn for our problem. This seems inevitable
+for any hydro code unless the hydrodynamics update includes an explicit equation for $T$ that is
+independent of what is happening for the internal energy. Similarly, the hybrid burn does not
+affect the total nickel production relative to the hydrostatic burn. The hybrid burn only changes
+the behavior for zones that have a net negative energy release from the hydrostatic burn, which
+occurs only after we have burned to NSE.
+
+\input{plots/burning_mode_m_P_0.64_m_S_0.64.tbl}
+
+In contrast, the suppressed burning mode (modelled after \citet{kushnir:2013}) produces only about
+$5 \times 10^{-4}\ \msolar$ of $^{56}$Ni. In a suppressed burn, we ensure that at every evaluation
+of the right-hand-side vector for the nuclear network integration, all quantities are scaled by the
+same factor, with this factor ensuring that the energy release is not large enough to permit a
+spurious detonation. This seems to be equivalent to the method used in \citet{kushnir:2013}. Yet
+we do not reproduce their claim that this fix is sufficient to ensure that the detonation occurs
+at the proper and time and place; indeed, the burning is quenched until very late in the collision,
+and the detonation barely forms at all. The suppression used therefore seems to be far too strong
+to permit an actual detonation when one should occur. We have also checked the same method at higher
+resolution and found that the nickel produced is significantly smaller than for a self-heating
+calculation at the same resolution.
+
+\subsection{Impact Parameter}
+\label{sec:parameters:impactparameter}
+
 \subsection{Other Parameters}

### 2016-07-12 papers/wdmerger_II/paper.tex
Correct some wording in an explanation

@@
 same factor, with this factor ensuring that the energy release is not large enough to permit a
-spurious detonation. This seems to be equivalent to the method used in \citet{kushnir:2013}. Yet
-we do not reproduce their claim that this fix is sufficient to ensure that the detonation occurs
-at the proper and time and place; indeed, the burning is quenched until very late in the collision,
-and the detonation barely forms at all. The suppression used therefore seems to be far too strong
-to permit an actual detonation when one should occur. We have also checked the same method at higher
-resolution and found that the nickel produced is significantly smaller than for a self-heating
-calculation at the same resolution.
+spurious detonation. This seems to be equivalent to the method used in \citet{kushnir:2013}. We
+reproduce their claim that this suppresses the detonation until later in the collision, after more
+material has approached the stalled shock. We also observe that, for a given resolution, the simultaneous
+detonations occur slightly outside the initial contact point, near the edges of the stalled shock. They
+establish as the main source of error in simulations that produce too little nickel. However, we find
+that the resulting detonations are not strong enough to convert a significant amount of material to NSE
+conditions; instead, only QSE conditions are reached near the center of the collision, leaving a significant
+amount of silicon and sulfur on the domain that is not further processed.
 

### 2016-07-14 papers/wdmerger_II/paper.tex

@@
 \begin{abstract}
-We consider the collisions of white dwarfs as possible progenitors of Type Ia 
-supernovae.
-
+Collisions of white dwarfs are a subset of white dwarf mergers that occur
+when the stars approach each other nearly head-on. These have the potential
+for explosive nucleosynthesis, leading to a thermonuclear detonation that
+disrupts both stars, possibly causing a Type Ia supernova. Simulation of
+these events is challenging due to the rapid evolution of the nuclear burning,
+posing numerical challenges related to both spatial and temporal resolution.
+In this study we explore a number of the algorithms that affect the outcome
+of reacting hydrodynamics simulations that lead to explosive burning and
+address the question of what needs to be done to ensure that these simulations
+are properly converged. We find that achieving a converged simulation in the
+code \castro\ is non-trivial and requires demanding computational resources.
+We also discuss the quantities obtained from our simulations that bear relevance
+to observations of these events, particularly production of nickel-group
+elements and gravitational wave emission during the collision.
 \end{abstract}
@@
 found that detonations occur and convert a large amount of carbon/oxygen material into
-iron-group elements.
+iron-group elements. These papers varied significantly in the hydrodynamic methods used
+(Lagrangian versus Eulerian methods), the methodology used for coupling nuclear
+reactions to the hydrodynamics (including variation in the number of isotopes in
+the nuclear network and in the evolution of the temperature), and the temporal and
+spatial resolution. Consequently they yielded estimates of production of nickel-group
+elements that varied significantly, leaving much uncertainty about how such an event
+would appear observationally and whether it bears any resemblance to a SN Ia.
 
-We consider the collisions of white dwarfs as possible progenitors of Type Ia 
-supernovae, using the code \castro\ \cite{castro}.
+In this paper we seek to add another view to the problem using the reactive hydrodynamics
+code \castro\ \citep{castro} and its underlying adaptive mesh refinement library \boxlib.
+This paper is a follow-up to \citet{wdmergerI} (hereafter, Paper I), which established
+the hydrodynamic methdology we are using to simulate mergers of white dwarfs.
+Paper I left out a discussion of the methodology we are using for nuclear reactions,
+and we fill out this information below in \autoref{sec:numericalmethods}. While it is
+our aim is to add information to the discussion about collisions of white dwarfs,
+we also view this work as a methodology study that asks to what extent we
+can trust the reliability of explosive thermonuclear reactions in Eulerian
+hydrodynamics simulations. As a corollary, to what extent does static and/or
+adaptive mesh refinement resolve the types of problems that occur in low resolution
+simulations? We discuss this in Section \autoref{sec:resolution}, though first
+we examine the response of the code to a number of user-specified parameters in
+\autoref{sec:parameters}. We then close with our thoughts on what our work implies
+for the broader astrophysical reacting hydrodynamics literature in \autoref{sec:conclusion}.
 
@@
 Next is \aproxthirteen\ \citep{timmes:1999,timmes:2000}. This includes
-all of the isotopes of \isoseven, and all of the $alpha$-chain particles between
+all of the isotopes of \isoseven, and all of the $\alpha$-chain particles between
 silicon and nickel (${}^{32}$S, ${}^{36}$Ar, ${}^{40}$Ca, ${}^{44}$Ti, ${}^{48}$Cr,
@@
 
+Paper I established our methodology for calculating the gravitational wave signature
+of a white dwarf merger using the post-Newtonian framework of \citet{blanchet:1990}.
+While a typical merger event has a characteristic profile including a chirp and ringdown,
+white dwarf collisions feature sharply different gravitational wave signatures, and
+hence may be interesting targets for observation for future low frequency gravitational
+wave detectors \citep{eLISA}. The main caveat for these events compared to events
+such as neutron star mergers is that due to the event being inherently less energetic,
+the typical distance scale on which we could observe a white dwarf merger or collision
+is the size of our own galaxy, with extragalactic events being unlikely to be
+detected, at least with the first generation of space-based gravitational wave
+detectors. \autoref{fig:gw_signal} is a plot of the gravitational wave strain for
+the representative case of \autoref{sec:parameters}. We use the convention that the
+absolute scale of the strain is normalized to a distance of 10 kpc (the strain scales
+inversely with the source distance). The shape of the profile agrees qualitatively
+with the gravitational wave signature predicted by \cite{loren-aguilar:2009:collisions},
+though because they used an unequal mass pair the results do not match exactly.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.425]{plots/gw_signal}
+  \caption{Gravitational wave strain for the head-on collision of two $0.64\ \msolar$
+           white dwarfs. The strain is normalized to a source distance of 10 kpc.
+           \label{fig:gw_signal}}
+\end{figure}
+
+%TODO: use a 3D collision for this?
+
 %==========================================================================

### 2016-07-15 papers/wdmerger_II/paper.tex
Update the gravitational wave signal

@@
 would appear observationally and whether it bears any resemblance to a SN Ia.
+See Table 4 of \cite{garcia-senz:2013} for a summary of the outcome of many of these studies.
 
@@
 While a typical merger event has a characteristic profile including a chirp and ringdown,
-white dwarf collisions feature sharply different gravitational wave signatures, and
+white dwarf collisions are noticeably distinct in their gravitational wave signatures, and
 hence may be interesting targets for observation for future low frequency gravitational
-wave detectors \citep{eLISA}. The main caveat for these events compared to events
+wave detectors \citep{eLISA}. The strain increases monotonically for a collision until impact,
+instead of oscillating back and forth at a characteristic frequency as for mergers.
+The main caveat for these events compared to events
 such as neutron star mergers is that due to the event being inherently less energetic,
@@
 detectors. \autoref{fig:gw_signal} is a plot of the gravitational wave strain for
-the representative case of \autoref{sec:parameters}. We use the convention that the
+the representative case of \autoref{sec:parameters}, where we disabled the normal stopping
+criterion and let the simulatin run until $t = 20$. We use the convention that the
 absolute scale of the strain is normalized to a distance of 10 kpc (the strain scales
 inversely with the source distance). The shape of the profile agrees qualitatively
-with the gravitational wave signature predicted by \cite{loren-aguilar:2009:collisions},
-though because they used an unequal mass pair the results do not match exactly.
+with the gravitational wave signatures predicted by \cite{loren-aguilar:2009:collisions}
+and \cite{garcia-senz:2013}.
 
@@
   \includegraphics[scale=0.425]{plots/gw_signal}
-  \caption{Gravitational wave strain for the head-on collision of two $0.64\ \msolar$
-           white dwarfs. The strain is normalized to a source distance of 10 kpc.
+  \caption{Gravitational wave strain $h_+$ for the head-on collision of two $0.64\ \msolar$
+           white dwarfs. The strain is normalized to a source distance of 10 kpc. We do not
+           show the other polarization, $h_\times$, as for an equal-mass head-on collision it vanishes
+           identically.
            \label{fig:gw_signal}}

### 2016-07-15 papers/wdmerger_II/paper.tex
Add impact parameter plot

@@
 
+As the impact parameter varies for a head-on collision, the cross-section of the contact
+zone between the two stars varies in size and geometry. In \autoref{fig:impact_parameter}, we
+see the dependence of the nickel production on the impact parameter $b$, measured in units
+of the WD radius. This test was performed using the 3D setup at our usual base resolution of
+400 km. For computational efficiency, timestep limiting based on burning was disabled.
+We sampled at $b = 0.0$, $0.01$, $0.02$, $0.03$, $0.04$, $0.05$, and then increasing
+in increments of $0.05$. There is no clear pattern in the data, with the nickel production varying
+significantly even for small changes in the impact parameter. The typical value is around $0.15\ \msolar$.
+It is worth nothing that for $b = 0.0$ the nickel production is a bit smaller for the 3D simulation
+than its equivalent 2D simulation, the default case described above.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.425]{plots/impact_parameter}
+  \caption[Nickel production dependence on the impact parameter]
+          {$^{56}$Ni generation as a function of the initial WD impact parameter for
+           a $0.64\ \msolar$ + $0.64\ \msolar$ WD collision.
+           \label{fig:impact_parameter}}
+\end{figure}
+
 \subsection{Other Parameters}
@@
 burning for small enough $\rho$ and $T$. Lowering the minimum $\rho$ for reactions to
-$10^0\ \text{g / cm}^{-3}$ from $10^6\ \text{g / cm}^{-3}$ had a negligible effect,
+$10^0\ \text{g cm}^{-3}$ from $10^6\ \text{g cm}^{-3}$ had a negligible effect,
 and varying the minimum $T$ for reactions from $2 \times 10^{7}\ \text{K}$ to

### 2016-07-17 papers/wdmerger_II/paper.tex
Update the impact parameter plot

@@
 We sampled at $b = 0.0$, $0.01$, $0.02$, $0.03$, $0.04$, $0.05$, and then increasing
-in increments of $0.05$. There is no clear pattern in the data, with the nickel production varying
-significantly even for small changes in the impact parameter. The typical value is around $0.15\ \msolar$.
+in increments of $0.05$. Relative to $b = 0$, there is a slight increase in nickel production for small
+but non-zero $b$, which then declines with increasing $b$ except for a bump at $b = 0.5$.
+For $b \ge 0.7$, no detonation forms at all, and the merged object remains a stable C/O WD
+with $M = 1.2\ \msolar$ at least until $t = 100$.
+
 It is worth nothing that for $b = 0.0$ the nickel production is a bit smaller for the 3D simulation
-than its equivalent 2D simulation, the default case described above.
+than its equivalent 2D simulation, the default case described above. There are differences both in
+the size of the timestep (which as we saw in \autoref{sec:parameters:timestepping} plays a role in
+the total nickel production) and in the discretized nature of the stars; the assumption of perfect
+azimuthal symmetry inherent to the 2D simulation breaks down for the 3D simulation, especially at
+low resolution, since the zones are rectangular.
 

### 2016-07-17 papers/wdmerger_II/paper.tex
Fix an exponent

@@
 is the same resolution as the 3D case: the width of the $z$ axis is $1.024 \times 10^{10}$
-cm and the width of the $R$ axis is $5.012 \times 10^{10}$ cm, with twice as
+cm and the width of the $R$ axis is $5.012 \times 10^{9}$ cm, with twice as
 many zones along the $z$ axis as the $R$ axis so that the equal 400 km

### 2016-07-18 papers/wdmerger_II/paper.tex
Write up the collision network dependence

@@
 have their mass fractions in every zone set to a negligibly small number at initialization.
+\autoref{table:networks} lists the $^{56}$Ni generation and total energy generation (determined
+by subtracting the initial energy from the final energy). The energy generated in the latter
+three cases agrees remarkably well, with a less than $0.5\%$ difference. \aproxthirteen\
+over-generates nickel at the $2\%$ level compared to the networks with more isotopes, a reasonably
+small difference. On the other hand, agreement with \isoseven\ is not very strong, as \isoseven\
+generates far too much nickel and far too little energy. Nevertheless while the quantitative
+agreement is not there, the collision looks qualitatively similar compared to the evolution with
+more complicated networks, so while \isoseven\ should not be used for predicting nucleosynthetic
+yields in production science runs, it may still be fair for use in proof-of-concept studies.
+
+\input{plots/networks_m_P_0.64_m_S_0.64.tbl}
+
+\isoseven\ also has the peculiar property that after the detonation had passed through the WDs,
+the region at the central contact point did not fully stop burning, yielding a quasistatic energy
+release that added a couple of million extra timesteps to the simulation (in addition to the few
+hundred it normally takes) before we manually terminated the run. The full effect of these additional
+timesteps only amounted to a $10^{-4}$ relative change in the nickel production. By examining the behavior
+of zones near the collision point, we discovered the reason for this. The zones in question have a composition that
+is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in
+comparison to, say, \aproxthirteen, reveals that $dX/dt$ for these isotopes is off by many orders of magnitude.
+This is a consequence of the choice inherent to the network to assume that the isotopes in between $^{28}$Si
+and $^{56}$Ni are in an effective equilibrium state. So an advective update that yields even a small change
+to the abundances in the zone will very strongly knock the zone out of equilibrium with respect to the burn
+step following the advective update. This then causes the burning timestep limiter to kick in and make the
+timestep very small so that the advective changes do not cause such strong effects on the burning equilibrium.
+This effect is an argument against the use of \isoseven\ for explosive, non-hydrostatic burning problems.
 

### 2016-07-18 papers/wdmerger_II/paper.tex
Modelled --> modeled, I guess
Add 3D gravitational wave signal

@@
 Regardless of whether the suppressed burning mode works for this
-particular problem, it is not physical, so in general we have
-taken a different approach. If we insist that we cannot
+particular problem, it is not physical, so we include for
+consideration a different approach. If we insist that we cannot
 cannot directly control the energy injection timescale, we 
@@
 
-While this approach may add several AMR levels 
-to a simulation, we agree with \citet{kushnir:2013} that 
-solving this numerical instability is crucial to avoiding
-early, unphysical detonations, and so it is unavoidable
+We agree with \citet{kushnir:2013} that solving this numerical
+instability is crucial to avoiding unphysical detonations.
+A simulation that does not solve this problem will not obtain
+the correct amount and will not converge properly with resolution.
+This may justify the addition of many AMR levels to the domain
 if a correct evaluation of the burning phase is desired.
-A simulation that does not solve this problem in this way
-(or some analogous manner) will not obtain the correct amount 
-of burning, and will not converge properly with resolution. 
-Fortunately, the regions where this criterion is not satisfied 
-are fairly localized on the grid, so the amount of additional 
-computational work is not insurmountable for a single simulation.
-However, this level of accuracy may make a parameter study of
-the problem challenging given current computational resources.
-When resources are limited, one can still use this method and set a
-limit on the number of levels of refinement possible in the simulation,
-so that the refinement will still go in the right places but may not
-refine all of the way to the regime of guaranteed stability.
+Even if one cannot afford the full resolution required by this
+AMR criterion, and chooses to limit the number of levels to some
+predetermined maximum based on a constraint of computing time,
+the added resolution will still go to the most-needed places.
+However, a significant drawback to this approach is that it
+does not turn on until after an unstable burning region has
+been generated, so this AMR criterion cannot help in the case
+where a spurious detonation begins in a single timestep on the coarse
+grid, though it will kick in immediately after that timestep
+to resolve the regions where the detonation has occurred.
+A possible remedy to be explored in future work is to add
+refinement pre-emptively in regions where the ratio $t_s / t_e$
+approaches $f_{s}$ but does not yet exceed it. Another
+possibility for this case would be to track how many levels
+would have been needed to prevent the numerical detonation,
+based on the stability criterion, and reject the timestep and
+retry it from the starting state using that many levels of refinement.
+The ability to reject a timestep and retry it is functionality that
+we have recently added to \castro, so we may try this in future work.
 
@@
 
-In contrast, the suppressed burning mode (modelled after \citet{kushnir:2013}) produces only about
+In contrast, the suppressed burning mode (modeled after \citet{kushnir:2013}) produces only about
 $5 \times 10^{-4}\ \msolar$ of $^{56}$Ni. In a suppressed burn, we ensure that at every evaluation
@@
 detected, at least with the first generation of space-based gravitational wave
-detectors. \autoref{fig:gw_signal} is a plot of the gravitational wave strain for
-the representative case of \autoref{sec:parameters}, where we disabled the normal stopping
-criterion and let the simulatin run until $t = 20$. We use the convention that the
-absolute scale of the strain is normalized to a distance of 10 kpc (the strain scales
-inversely with the source distance). The shape of the profile agrees qualitatively
-with the gravitational wave signatures predicted by \cite{loren-aguilar:2009:collisions}
-and \cite{garcia-senz:2013}.
+detectors.
 
@@
   \centering
-  \includegraphics[scale=0.425]{plots/gw_signal}
-  \caption{Gravitational wave strain $h_+$ for the head-on collision of two $0.64\ \msolar$
-           white dwarfs. The strain is normalized to a source distance of 10 kpc. We do not
-           show the other polarization, $h_\times$, as for an equal-mass head-on collision it vanishes
-           identically.
-           \label{fig:gw_signal}}
+  \includegraphics[scale=0.425]{plots/gw_signal_2D}
+  \caption{Gravitational wave strains for the head-on collision of two $0.64\ \msolar$
+           white dwarfs. The strain is normalized to a source distance of 10 kpc.
+           \label{fig:gw_signal_2D}}
 \end{figure}
 
-%TODO: use a 3D collision for this?
+\autoref{fig:gw_signal_2D} is a plot of the gravitational wave strain for
+the representative case of \autoref{sec:parameters} ($M_P = 0.64\ \msolar$, $M_S = 0.64\ \msolar$, $b = 0$),
+where we disabled the normal stopping criterion and let the simulation run until $t = 20$.
+We use the convention that the absolute scale of the strain is normalized to a distance
+of 10 kpc (the strain scales inversely with the source distance). We show the $+$ and $\times$
+polarizations for observers situated along the $x$, $y$, and $z$ axes (for the 2D case,
+this corresponds to the $z$, $R$, and $\phi$ axes, respectively, in the simulation).
+The shape of the profile agrees qualitatively with the gravitational wave signatures
+predicted by \cite{loren-aguilar:2009:collisions} and \cite{garcia-senz:2013}.
+\autoref{fig:gw_signal_3D} shows the strain prediction for a 3D collision event
+with $M_P = 0.8\ \msolar$, $M_S = 0.6\ \msolar$, and $b = 0.8$. This mass combination
+was chosen because \cite{loren-aguilar:2009:collisions} used the same combination for
+gravitational wave signal prediction, but we cannot replicate the large impact parameter
+they used on our simulation grid, so the detonation looks quite different than in their case.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.425]{plots/gw_signal_3D}
+  \caption{Gravitational wave strains for the off-center collision ($b = 0.8$) of
+           $0.8\ \msolar$ and $0.6\ \msolar$ WDs. The strain is normalized to a
+           source distance of 10 kpc.
+           \label{fig:gw_signal_3D}}
+\end{figure}
 

### 2016-07-20 papers/wdmerger_III/paper.tex
Clarify some wording for hybrid advection

@@
 $x / R$, and $y/R$, and subsequent algebraic manipulation. We now rewrite this in the
-way that we actually discretize them for use in the code (by use of the product rule
-to rearrange the terms with derivatives in them):
+way that illustrates how the update is done in the code:
 \begin{align}
   \frac{\partial(s_R)}{\partial t} &= -\nabla \cdot (s_R \mathbf{u})
-    - \left(\frac{\partial(p (x/R))}{\partial x} + \frac{\partial(p (y/R))}{\partial y}\right) \notag \\
-    &\hspace{0.175in} + \frac{1}{\rho R^3}\left(\ell_z^2 + p\rho R^2\right) \label{eq:radial-momentum-rearranged}\\
+    - \frac{1}{R}\left(x \frac{\partial p}{\partial x} + y \frac{\partial p}{\partial y}\right)
+    + \frac{\ell_z^2}{\rho R^3} \label{eq:radial-momentum-rearranged}\\
     \frac{\partial(\ell_z)}{\partial t} &= -\nabla \cdot (\ell_z \mathbf{u}) - \left(\frac{\partial (-py)}{\partial x} + \frac{\partial(px)}{\partial y}\right).\label{eq:angular-momentum-rearranged}
 \end{align}
-It is clear that the second term on the right-hand-side in each case can be swept up
+The first term is the standard advective flux term for any variable that advects with
+the flow, and the state quantity inside this advective term can be derived on cell edges
+using the primitive hydrodynamic variables after a Riemann solver has been performed.
+The second term on the right-hand-side of \autoref{eq:angular-momentum-rearranged} can be swept up
 into the flux terms in the same way that the pressure term for the linear momenta
 is often added to the fluxes, to make the conservation form explicit. The remainder
-is the last term on the right-hand-side of the radial momentum equation, which
-exists as a cell-centered source term (which we deal with using a predictor-corrector
-method similar to how we implement external forcing like gravity and rotation; see
-\cite{castro} for details). It is important to observe that in this formulation
+are the last two terms on the right-hand-side of \autoref{eq:radial-momentum-rearranged}.
+While maintaining second-order accuracy in time, the last term can be treated as a
+cell-centered source term that we deal with using a predictor-corrector
+method (similar to how we implement external forcing like gravity and rotation; see
+\cite{castro} for details). The pressure gradients in the second-to-last term can be
+constructed using the Riemann-solved edge states, and they are multiplied by zone
+coordinates that are cell-centered. In principle this is second-order accurate in time
+because the edge states are too; however, as a drawback, the pressure is unaware of the
+source terms that are applied in the corrector step after a hydrodynamics update. The
+benefit of this approach is that we do not need to construct a general cell-centered
+gradient of the pressure using cell-centered state data, nor do we need to perform the
+equation of state call that entails.
+
+It is important to observe that in this formulation
 the radial momentum equation is \textit{not} a conserved quantity: it should and
@@
 source terms in the radial momentum equation. Of course, the answer will likely
-be problem-dependent. Indeed, we already faces this situation in general with the
+be problem-dependent. Indeed, we already face this situation in general with the
 gravitational and rotational source terms, even independent of the hybrid momentum

### 2016-07-22 papers/wdmerger_II/paper.tex
Updates to papers

@@
 where the components of the right-hand-side for the species come from the particular
-nuclear burning network we are using. The energy release $e$ of the zone's burn
+nuclear burning network we are using. The energy $e$ of the zone
 will change when the nuclear abundances evolve, according to
@@
 
-In a hydrostatic burn, we keep $\rho$ and $T$ fixed throughout, and use
+We define several burning modes that determine how $T$ and $e$ are evolved
+during a nuclear burn. In a hydrostatic burn, which we call burning mode 0,
+we keep $\rho$ and $T$ fixed throughout, and use 
 the energy released at the end to compute a final temperature that is
 thermodynamically consistent with the new internal energy. By contrast,
-in a self-heating burn, we allow the temperature to evolve in response
+in a self-heating burn (mode 1), we allow the temperature to evolve in response
 to the burning (see\footnote{In the cited paper, a term based on the
@@
 (Although $T$ evolves during the burn so that the integration is physically
-accurate, as for the hydrostatic method we discard the final value
+accurate, as in the hydrostatic method we discard the final value
 for $T$ at the end of the burn and recompute a temperature for the zone that is
@@
 The latter is more expensive but also more accurate, and we use it in this paper.
-In practice we find that the cost is negligible in comparison to the more expensive
+In practice we find that the cost is small in comparison to the more expensive
 parts of the calculation, and it can significantly speed up convergence near NSE.
-A third option presented by \citet{raskin:2010} is a so-called ``hybrid'' mode.
+A third option (mode 2) presented by \citet{raskin:2010} is a so-called ``hybrid'' mode.
 In this mode, by default we do a hydrostatic burn. If that burn fails, or if the net
-energy change is negative, we do the burn again in self-heating mode. A final option
+energy change is negative, we do the burn again in self-heating mode. A final option (mode 3)
 is a burn that limits the changes due to a burn to avoid numerically unstable burning.
@@
 solving a set of coupled stiff ODEs. For this work we use an implementation of
-the well known variable-order Richardson extrapolation method presented by Stoer and
-Bulirsch \cite{stoer:1980}, that is similar to the integrator which ships with
+the well known variable-order Richardson extrapolation method presented by \citet{stoer:1980},
+that is similar to the integrator which ships with
 the original versions of the networks mentioned above. Previous work using the \boxlib\
-codes typically used the classic \vode\ integrator \cite{vode}. We do maintain a version
+codes typically used the classic \vode\ integrator \citep{vode}. We do maintain a version
 of the software that is compatible with our software interfaces in the \microphysics\
@@
 $f_{e}$ and $f_{X}$ smaller, we can control the magnitude of the decoupling 
-between the reactions and the hydro. We choose by default $f_{X} = f_{e} = 0.1$, 
-a choice which is slightly smaller than others in the literature. (Note that previous 
-collision papers we are aware of only limited based on changes in internal energy, 
-not on cahnges in species.)  The sensitivity of results to this choice will be discussed in 
+between the reactions and the hydro. A typical choice for $f_e$ parameter in the
+literature is in the range of 0.2 or 0.3, while to our knowledge a limiter based on
+$f_X$ has not been used by others performing these types of calculations. The sensitivity
+of results to the value of these timestep limiters will be discussed in 
 \autoref{sec:parameters:timestepping}. The factors $f_{e}$ and $f_{X}$ can be set at runtime in \castro.
@@
 overestimate the needed timestep because the burning was quiescent
-for at least some portion of the last advance. In particular, our 
-experience has shown that using only the energy change criterion
-is particularly susceptible to this phenomenon; silicon-group 
-material can build up without changing the internal energy by a 
-large fraction, so the timestep limiter is never triggered, and 
-then in a single step a substantial amount of iron-group elements 
-can be be generated. This can have non-trivial effects on the 
-total amount of iron-group elements generated over the course of
+for at least some portion of the last advance. Our experience has
+shown that none of these methods is flawless, and that limiting
+based on only changes in internal energy is particularly susceptible
+to this staggered burning phenomenon; silicon-group material can
+build up without changing the internal energy by a  large fraction,
+so the timestep limiter is never triggered, and then in a single step
+a substantial amount of iron-group elements can be be generated,
+perhaps forming a detonation. This can have non-trivial effects
+on the total amount of iron-group elements generated over the course of
 the simulation. We have found that the addition of the 
-limiter based on changes in species functionally precludes this.
+limiter based on changes in species functionally precludes this, and
+so the two methods can complement each other.
 
@@
 couple their advection-diffusion-reaction equation set. In our
-context this would involve using the advection as a source term
-for the burning network, and symmetrically including a source
-term to the hydrodynamics corresponding to the energy release
-from the burning. We are presently investigating such a method,
+context this would involve treating the full evolution equation
+for each of the state variables as an ODE with a directly coupled
+burning source term integrated at high order; the advective
+flux is evaluated at the standard second order accuracy
+and is included as an ODE source term.
+We are presently investigating such a method,
 and it may form the basis of further work on this subject.
@@
 $e > \eta_2\, E$. In other cases the evolved value of the internal
-energy variable is preserved. However, this choice does not work
+energy variable is preserved. However, this choice is somewhat unsafe
 for our application, because the ultimate cause of the nuclear
@@
 the energy injection timescale
-$t_e$ is shorter than the sound-crossing time $t_s$ in a zone.
+$\tau_e$ is shorter than the sound-crossing time $\tau_s$ in a zone.
 When the sound-crossing time is too long, energy is built up in
@@
 \begin{equation}
-  t_s \leq f_{s} t_e \label{eq:burning_limiter_2}
+  \tau_s \leq f_{s}\, \tau_e \label{eq:burning_limiter_2}
 \end{equation}
-The sound crossing time, $t_s$, is given by $\Delta x / c_s$, 
+The sound crossing time, $\tau_s$, is given by $\Delta x / c_s$, 
 where $c_s$ is the sound speed and $\Delta x$ is the (minimum) 
@@
 holds. (If the inequality is already satisfied, then the integration
-vector is not modified.) We fix $t_s$ to be the value of the sound
+vector is not modified.) We fix $\tau_s$ to be the value of the sound
 crossing time at the beginning of the burn (that is, we do not
 update it as the sound speed changes) and we fix the energy $e$
-that goes into the estimate for $t_e$ to be the value of the
+that goes into the estimate for $\tau_e$ to be the value of the
 internal energy of the zone at the beginning of the burn. If
@@
 hydrodynamics update can advect away newly generated energy
-quickly enough to avoid a nuerically seeded detonation,
+quickly enough to avoid a numerically seeded detonation,
 we desire the most conservative version of the limiter. We discuss
@@
 particular problem, it is not physical, so we include for
-consideration a different approach. If we insist that we cannot
+consideration a different approach. If we insist that we
 cannot directly control the energy injection timescale, we 
@@
 regions that do not satisfy \autoref{eq:burning_limiter_2},
-which effectively lowers the $\Delta x$ and thus the
+which effectively lowers $\Delta x$ and thus the
 sound-crossing time. We keep tagging zones for refinement
@@
 A possible remedy to be explored in future work is to add
-refinement pre-emptively in regions where the ratio $t_s / t_e$
+refinement pre-emptively in regions where the ratio $\tau_s / \tau_e$
 approaches $f_{s}$ but does not yet exceed it. Another
@@
 retry it from the starting state using that many levels of refinement.
-The ability to reject a timestep and retry it is functionality that
+As mentioned in \autoref{sec:burnerhydrocoupling}, the ability to reject
+a timestep and retry it is functionality that
 we have recently added to \castro, so we may try this in future work.
@@
 two point-masses in free-fall towards each other coming in from infinity, such that
-the contact point is at the origin, and they approach each other on the $x$-axis.
+the contact point is at the origin, and they approach each other along the $x$-axis.
 The composition of the WDs is as described in Paper I: low-mass WDs are helium-dominated,
@@
 hence may be interesting targets for observation for future low frequency gravitational
-wave detectors \citep{eLISA}. The strain increases monotonically for a collision until impact,
+wave detectors such as LISA \citep{eLISA}. The strain increases monotonically for a collision until impact,
 instead of oscillating back and forth at a characteristic frequency as for mergers.
@@
 
+\software{BoxLib\footnote{\url{https://github.com/BoxLib-Codes/BoxLib}},
+          CASTRO \citep{castro} (https://github.com/BoxLib-Codes/Castro),
+          wdmerger \citep{wdmergerI} (\url{https://github.com/BoxLib-Codes/wdmerger}),
+          GCC (\url{https://gcc.gnu.org/}),
+          python (\url{https://www.python.org/}),
+          matplotlib \citep{matplotlib}, \url{http://matplotlib.org/}),
+          yt \citep{yt}, \url{http://yt-project.org/})}
+
 \clearpage

### 2016-07-22 papers/wdmerger_III/paper.tex
Updates to papers

@@
 
-It is important to observe that in this formulation
-the radial momentum equation is \textit{not} a conserved quantity: it should and
-will in general change with time over the course of the simulation. The above form,
-absent of source terms, makes this clear. The angular momentum term in 
-\autoref{eq:angular-momentum} can be thought of physically as expressing the fact
+It is important to observe that in this formulation the radial momentum equation
+is \textit{not} a conserved quantity: it should and will in general change with
+time over the course of the simulation if the system is not in perfect rotational
+equilibrium. The above form, absent of source terms, makes this clear. The angular momentum term in
+\autoref{eq:radial-momentum} can be thought of physically as expressing the fact
 that a particle with some initial angular momentum will have its radial momentum
@@
 as expressing the curved nature of the underlying coordinates for the radial momentum
-\citep{motl:2002}. Note that while the version in \autoref{eq:radial-momentum-rearranged}
-is more convenient for our hydrodynamics methodology, the pressure-gradient terms
-in the first version are the ones appropriate for a radial field and so no physical meaning
-should be ascribed to the pressure term at the end of \autoref{eq:radial-momentum-rearranged}.
+\citep{motl:2002}. It is only the presence of a source term like gravity that can keep
+such a particle on an orbit at its original radius, and for such a case it may be helpful
+to think of the angular momentum source term as an outward centrifugal force that
+balances against the inward centripetal force provided by gravity.
 

### 2016-07-22 analysis/wdmerger.py

@@ -4,10 +4,14 @@
 #
 
 def get_wdmerger_git_commit_hash():
+    """Return current git commit hash of wdmerger."""
+
     import os
     import subprocess
-    cwd = os.getcwd()
+
     WDMERGER_HOME = os.getenv('WDMERGER_HOME')
+
+    cwd = os.getcwd()
     os.chdir(WDMERGER_HOME)
     hash = subprocess.check_output(['git', 'rev-parse', 'HEAD']).strip()
     os.chdir(cwd)
@@ -21,41 +25,40 @@ def get_wdmerger_git_commit_hash():
 #
 
 def get_castro_dir():
+    """Return the location of the CASTRO directory."""
+
     import os
+
     CASTRO_HOME = os.getenv('CASTRO_HOME')
+
     return CASTRO_HOME
 
 
 
-# Return the last inputs file in a directory.
+# Return the name of the current inputs file in a directory.
 
 def get_inputs_filename(directory):
+    """Return the name of the inputs file in a directory."""
 
     import os
 
-    # If there's a file named 'inputs', then this was a standard run.
+    # At present we have no reason to look for anything other than inputs.
 
-    if (os.path.isfile(directory + '/inputs')):
+    if os.path.isfile(directory + '/inputs'):
 
         return 'inputs'
 
-    # If there are files named inputs_*, corresponding to a chain, return the last one.
+    else:
 
-    elif (os.path.isfile(directory + '/inputs_1')):
-        
-        inputs_list = filter(lambda s: s[0:6] == "inputs", os.listdir(directory))
+        print "Error: no inputs file found in " + directory + "."
+        exit
 
-        # Now sort them numerically
-        
-        inputs_list.sort(key=lambda s: int(s[7:]))
 
-        return inputs_list[-1]
- 
-       
 
 # Get a CASTRO variable value from an inputs file.
 
 def get_inputs_var(inputs, var):
+    """Retrieve a CASTRO variable value from an inputs file."""
 
     import numpy as np
 
@@ -101,6 +104,7 @@ def get_inputs_var(inputs, var):
 # Get a variable value from a probin file.
 
 def get_probin_var(probin, var):
+    """Retrieve a variable value from a probin file."""
 
     import numpy as np
 
@@ -144,17 +148,21 @@ def get_probin_var(probin, var):
 
 
 #
-# Given a plotfile directory, return the CASTRO and BoxLib git commit hashes.
[…]

### 2016-09-12 papers/wdmerger_III/paper.tex
Update the description of the hybrid advection scheme

--- 2016-02-23 papers/wdmerger_III/paper.tex
@@
 \section{Numerical Methods}
+\label{sec:methodology}
+
+The numerical methodology for our problem was established in \cite{wdmergerI}.
+In this section we report only on the differences between our current methodology
+and the methodology in that work.
 
 \subsection{Hybrid Advection Scheme}
+\label{sec:hybrid}
 
@@
 momentum and conservation of angular momentum in grid-based, Eulerian codes
-such as \castro. Although both are true statements analytically, we cannot
+such as \castro. Although both quantities should be conserved analytically, we cannot
 numerically guarantee that both hold. Consequently the choice of which
 physical quantity to conserve must be motivated by analysis of the particular
-problem to be done. Ideally a simulation code will be capable of evolving the
+problem to be done. In \cite{wdmergerI}, we demonstrated that numerical conservation
+of linear momentum leads to white dwarf mergers that are spuriously caused by violation
+of numerical conservation of angular momentum. Ideally our simulation code should be capable of evolving the
 equations for both the linear momentum and the angular momentum, and be
@@
 already had the standard method implemented for evolving the linear momentum.
-To solve the need for an angular momentum evolution, we adopt the ``hybrid''
+To solve the need for an angular momentum evolution, which could end up being more
+appropriate for binary star systems than linear momentum evolution (a proposition we
+will test in \autoref{sec:results}), we adopt the ``hybrid''
 advection scheme presented by \cite{byerly:2014} (see also \cite{motl:2002} for
@@
 studies of the stability of binary stellar systems. This section contains a
-discussion of our implementation of that scheme in \castro. For the sake
-of simplicity, let us assume that the rotation axis of the binary system is the $z$
+discussion of our implementation of that scheme in \castro.
+
+For the sake of simplicity, let us assume that the rotation axis of the binary system is the $z$
 axis and so the WDs orbit in the $xy$-plane. The core of the method is that instead
@@
 rotation axis) and the angular momentum with respect to the $z$-axis,
-$\ell_z \equiv R\rho v_\phi$ (where $v_\phi$ is the azimuthal velocity).
+$\ell_z \equiv R\rho v_\phi$ (where $v_\phi$ is the azimuthal velocity
+and $R$ is the distance to the rotation axis).
 The equation for the linear $z$-momentum is unchanged. We want to be able
 to solve these equations while still operating in our preferred Cartesian
-coordinate system. \cite{byerly:2014} present these equations as the following;
-for the moment we neglect inclusion of external source terms such as gravity and rotation:
+coordinate system. \cite{byerly:2014} present these equations as the following
+(for the moment we neglect inclusion of external source terms such as gravity and rotation,
+and we note that we have reorganized terms slightly relative to their presentation):
 \begin{align}
-  \frac{\partial(s_R)}{\partial t} + \nabla \cdot (s_R \mathbf{u}) &=
-    - \frac{1}{R}\left(x \frac{\partial}{\partial x} + y \frac{\partial}{\partial y}\right) p
+  \frac{\partial(s_R)}{\partial t} &= -\nabla \cdot (s_R \mathbf{u})
+    - \frac{1}{R}\left(x \frac{\partial p}{\partial x} + y \frac{\partial p}{\partial y}\right)
     + \frac{\ell_z^2}{\rho R^3} \label{eq:radial-momentum}\\
-    \frac{\partial(\ell_z)}{\partial t} + \nabla \cdot (\ell_z \mathbf{u}) &=
-  \left(y\frac{\partial}{\partial x} - x \frac{\partial}{\partial y}\right) p \label{eq:angular-momentum}
+    \frac{\partial(\ell_z)}{\partial t} &= -\nabla \cdot (\ell_z \mathbf{u}) - \left(\frac{\partial (-py)}{\partial x} + \frac{\partial(px)}{\partial y}\right).\label{eq:angular-momentum}
 \end{align}
@@
 the Euler equations for linear momentum by appropriate multiplication of $x$, $y$,
-$x / R$, and $y/R$, and subsequent algebraic manipulation. We now rewrite this in the
-way that we actually discretize them for use in the code (by use of the product rule
-to rearrange the terms with derivatives in them):
-\begin{align}
-  \frac{\partial(s_R)}{\partial t} &= -\nabla \cdot (s_R \mathbf{u})
-    - \left(\frac{\partial(p (x/R))}{\partial x} + \frac{\partial(p (y/R))}{\partial y}\right) \notag \\
-    &\hspace{0.175in} + \frac{1}{\rho R^3}\left(\ell_z^2 + p\rho R^2\right) \label{eq:radial-momentum-rearranged}\\
-    \frac{\partial(\ell_z)}{\partial t} &= -\nabla \cdot (\ell_z \mathbf{u}) - \left(\frac{\partial (-py)}{\partial x} + \frac{\partial(px)}{\partial y}\right).\label{eq:angular-momentum-rearranged}
-\end{align}
-It is clear that the second term on the right-hand-side in each case can be swept up
+$x / R$, and $y/R$, and subsequent algebraic manipulation. The first term on the
+right-hand-side is the standard advective flux term for any variable that advects with
+the flow, and the state quantity inside this advective term can be derived on cell edges
+using the primitive hydrodynamic variables after a Riemann solve has been performed.
+The second term on the right-hand-side of \autoref{eq:angular-momentum} can be swept up
 into the flux terms in the same way that the pressure term for the linear momenta
 is often added to the fluxes, to make the conservation form explicit. The remainder
-is the last term on the right-hand-side of the radial momentum equation, which
-exists as a cell-centered source term (which we deal with using a predictor-corrector
-method similar to how we implement external forcing like gravity and rotation; see
-\cite{castro} for details). It is important to observe that in this formulation
-the radial momentum equation is \textit{not} a conserved quantity: it should and
-will in general change with time over the course of the simulation. The above form,
-absent of source terms, makes this clear. The angular momentum term in 
-\autoref{eq:angular-momentum} can be thought of physically as expressing the fact
-that a particle with some initial angular momentum will have its radial momentum
-increase with time as it moves away from the origin, and can be thought of geometrically
-as expressing the curved nature of the underlying coordinates for the radial momentum
-\citep{motl:2002}. Note that while the version in \autoref{eq:radial-momentum-rearranged}
-is more convenient for our hydrodynamics methodology, the pressure-gradient terms
-in the first version are the ones appropriate for a radial field and so no physical meaning
-should be ascribed to the pressure term at the end of \autoref{eq:radial-momentum-rearranged}.
+are the last two terms on the right-hand-side of \autoref{eq:radial-momentum}.
+While maintaining second-order accuracy in time, the last term can be treated as a
+cell-centered source term that we deal with using a predictor-corrector
+method (similar to how we implement external forcing like gravity and rotation; see
+\cite{castro} for details). The pressure gradients in the second-to-last term can be
+constructed using the Riemann-solved edge states, and they are multiplied by zone
+coordinates that are cell-centered. In principle this is second-order accurate in time
+because the edge states are too; however, as a drawback, the pressure is unaware of the
+source terms that are applied in the corrector step after a hydrodynamics update. The
+benefit of this approach is that we do not need to construct a general cell-centered
+gradient of the pressure using cell-centered state data, nor do we need to perform the
+equation of state call that entails.
+
+It is important to observe that in this formulation the radial momentum equation
+is \textit{not} a conserved quantity: it should and will in general change with
+time over the course of the simulation if the system is not in perfect rotational
+equilibrium. The above form, absent of source terms, makes this clear. The angular
+momentum term in \autoref{eq:radial-momentum} is effectively an outward centrifugal
+force that tends to increase the radial momentum with time. This is a reflection of
+the fact that mass on the domain with non-zero momentum tends to increase its radial
+momentum with time. For example, consider a star initially on the $x$ axis moving
+in the positive $y$ direction at constant speed: initially its radial momentum is
+zero, but as its $y$ coordinate increases, more of its velocity becomes radial rather
+than angular).  This term can thus also be thought of geometrically as expressing the
+curved nature of the underlying coordinates for the radial momentum
+\citep{motl:2002}. In the presence of a source term like gravity that can keep
+such a star on an orbit at its original radius, the gravitational force acts as a
+centripetal force that balances the centrifugal force, allowing the radial momentum
+to remain constant with time, as it should when the orbital radius is constant.
 
@@
 advective velocity $\mathbf{u}$. When we update the normal momenta with an
-external forcing $\mathbf{F}$ such as gravity or rotation, with $x$ and $y$
+external forcing $\mathbf{F}$ such as gravity, with $x$ and $y$
 components $F_x$ and $F_y$, we apply the update to the hybrid momenta as well:
@@
   \end{align}
-Then, when we are in a phase of the evolution when we want to conserve angular
-momentum, we make one additional change: at the end of a (single-level) advance,
-we recompute the normal momenta so that they are fully consistent with the hybrid
-momenta. This is all that is necessary to have a calculation that conserves angular
-momentum over the course of the simulation, without changing any of the
-core infrastructure for the hydrodynamics update. For example, when we compute
-the primitive variables from the conservative variables, we will still be
-getting them from the normal momenta, but these normal momenta are always consistent
-with the conserved angular momentum, so we get the same result as if we had done
-a much more involved update where we explicitly computed the primitive variables
-using the hybrid momenta. There is of course no unique choice for how to do this.
-For example, we could take the approach of explicitly re-calculating the normal
-momenta after any change has been made to the state. This would add significant code
-complexity without a clear benefit. The most significant difference would be
-that when evaluating the new-time value of source terms like rotation, the velocity
-field used is slightly different than it would otherwise be; but this is a high-order
-effect that does not change the core angular momentum conservation property. An
-analysis of the different mechanisms for implementing this update may be an interesting
-avenue for future research --- including various aspects like how to implement the
-source terms in the radial momentum equation. Of course, the answer will likely
-be problem-dependent. Indeed, we already faces this situation in general with the
-gravitational and rotational source terms, even independent of the hybrid momentum
-discussion --- the velocity field seen by the new-time rotation source terms depends
-on whether we apply the gravitational force before or after the rotation force.
+
+If in addition to evolving these variables we want to use them in place of the
+standard linear momenta for determining the hydrodynamic flow, only one change
+is needed for the standard timestep advance scheme: after source terms are
+applied to the state, we recompute the normal momenta so that they are fully
+consistent with the hybrid momenta. In practice for a \castro\ advance this
+synchronization of the linear and hybrid momenta occurs twice: once after the
+old-time source terms and the hydrodynamic update has been applied (these all
+use time-level $n$ data to construct the update, which is expected to already have
+synchronized linear and hybrid momenta), and once after new-time source
+terms have been applied. The use of adaptive mesh refinement implies sychronization
+in two other places: after a reflux that syncs up the fluxes between coarse and
+fine grids; and, after interpolation has been performed to obtain fine grid data
+from a coarser grid (since the spatially interpolated data may not respect the
+consistency relationship between the two momenta). In this way, the linear momenta
+are essentially placeholder variables that always reflect what the current radial
+and angular momenta are, but it is quite useful to express the scheme in this way
+because most of the \castro\ infrastructure is already set up to use the linear
+momentum state variables for calculating source terms, so the hybrid scheme requires
+minimal changes to the code architecture. For example, when we compute
+the primitive variables from the conservative variables in preparation for an advective
+hydrodynamic update, we will still be getting them from the linear momenta,
+but these linaer momenta are consistent with the conserved angular momentum,
+so we get the same result as if we had modified the code to explicitly compute
+the primitive variables as a function of the hybrid momenta. Consequently the
+scheme presented here adds the ability to explicitly conserve angular momentum
+in a simulation without a significant footprint on the simulation software.
+
+%==========================================================================
+% Results
+%==========================================================================
+\section{Simulation Results}\label{sec:results}
 
@@
 %==========================================================================
-\section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}
+\section{Conclusions and Discussion}\label{sec:Conclusions and Discussion}
 
 \acknowledgments
+
+The authors thank Dominic Marcello and Zachary Byerly for helpful comments
+regarding the hybrid advection scheme.
 

### 2016-09-12 papers/wdmerger_III/paper.tex
Describe the inertial frame variable scheme

@@
 momentum state variables with hydrodynamics fluxes, we add the analogous fluxes
-to the hybrid momenta -- that is, using the edge state values determined by the
+to the hybrid momenta --- that is, using the edge state values determined by the
 final multi-dimensional Riemann solve, we construct the value of the hybrid
@@
 hydrodynamic update, we will still be getting them from the linear momenta,
-but these linaer momenta are consistent with the conserved angular momentum,
+but these linear momenta are consistent with the conserved angular momentum,
 so we get the same result as if we had modified the code to explicitly compute
@@
 in a simulation without a significant footprint on the simulation software.
+
+\subsection{Rotation Revisited}\label{sec:rotation}
+
+Another significant change we have made in \castro\ since \citet{wdmergerI} deals
+with how rotation is applied as a source term. We noted in that paper that rotation
+can be used to reduce advection errors in a binary system due to the fact that
+material is moving less on the grid. However, we have also observed that rotation
+source terms introduce the opportunity for other numerical problems to arise. In
+particular, the most significant problem for us is that conservation of linear and
+angular momentum is harder to accomplish with the standard rotation source terms.
+Largely this is due to the fact that the rotation force is not expressed in a numerically
+conservative fashion; the forces are evaluated at cell centers and do not inherently
+have the telescoping property that allows for conservation of momentum in the case
+of hydrodynamic fluxes. While gravity has the property that in practice momentum
+conservation errors due to the gravitational force are typically quite small due to
+an underlying telescoping behavior in the gravitational force (see the discussion
+in Section 2.3.1 of \cite{wdmergerI}), this is harder to obtain for rotation (this is
+easy to see when considering the Coriolis force, as the standard Coriolis forcing
+involves cell-centered velocities that cannot trivially be evaluated in a way that
+leads to numerical cancellation --- though see \citet{audusse:2009} for a method that
+utilizes the edge-centered hydrodynamic fluxes to this end). A way around this issue
+is to evolve variables measured in the \textit{inertial} frame, even when the reference
+frame used for the advection is rotating. The benefits of this were observed by
+\citet{kley:1998} in the context of protostellar accretion disk simulations (and the
+method has since been used in a number of codes in that community, e.g. \citet{NIRVANA}
+and \citet{FARGO3D}), and also by \citet{call:2010}, whose method was adopted by
+\citet{byerly:2014} in the context of binary star simulations. The benefit of using
+inertial frame variables in this context is that the rotational terms vanish as an explicit
+source term to the angular momentum. \citet{kley:1998} points out that the rotational update
+to the $z$-component of the angular momentum measured in the rotating frame (which we
+temporarily call $\ell^\prime$) is:
+\begin{equation}
+  \left.\frac{\partial \ell}{\partial t}\right|^{\text{rotation}} = -2\rho\, v_r\, \omega\, R.\label{eq:rotating-angular-momentum}
+\end{equation}
+(Note that in \castro\ we construct all source terms, including rotation, for the linear
+momenta and then apply the source term to the hybrid momenta using \autoref{eq:force-radial} and
+\autoref{eq:force-angular}; \autoref{eq:rotating-angular-momentum} is automatically obtained this way.)
+The term on the right-hand-side vanishes if instead we evolve the inertial frame $\ell$, and there
+is no source term for the radial momentum either. If one is evolving the linear momenta measured
+in the inertial frame, \cite{byerly:2014} give the form of the update:
+\begin{equation}
+  \left.\frac{\partial (\rho \mathbf{u})}{\partial t}\right|^{\text{rotation}} = -{\bm{\omega}} \times (\rho \mathbf{u})
+\end{equation}
+This is equal to one-half of the Coriolis force (with the centrifugal force not used; see
+\cite{call:2010} for a further discussion of the meaning of this expression),
+and is needed because the linear momentum components change when advected over a
+rotating cylindrical grid. (Since the form of the update is similar to the Coriolis force,
+we can still use the implicit update technique from Section 2.4 of \cite{wdmergerI} when
+evaluating the new-time source term at time-level $n+1$.)
+
+The ability to evolve state variables measured in the inertial frame when doing a
+rotating frame simulation has been added as a runtime parameter in \castro. Note that
+other than evolving inertial frame variables, other aspects of the rotating frame are
+preserved --- for example, binary stars in a stable circular orbit will remain in the same
+location on the grid as where they started, just as in a standard rotating frame simulation,
+and the difference is mainly that the simulation velocity associated with those stationary
+stars will read as the equivalent velocity of the star in the inertial frame. The only
+meaningful wrinkle this introduces is that the hydrodynamic update needs to use the
+rotating frame velocities, not the inertial frame velocities. So in the conversion from
+conserved variables $\mathbf{U}$ to primitive variables $\mathbf{q}$, the latter being
+the ones used by the hydrodynamics module to construct an advective update, we subtract
+the rotating frame velocity from the conserved variables when constructing $\mathbf{q}$.
+Then the rest of the hydrodynamics update proceeds as normal with rotating frame velocities.
+As observed by \citet{byerly:2014}, this is acceptable because the divergence of the
+rotational velocity field is zero, so it can legitimately be analytically subtracted
+from the inertial velocity field. The payoff to the scheme is that we get the better
+numerical behavior associated with the rotating frame velocities, as they are smaller
+in magnitude with respect to the grid.
+
+\subsection{Miscellaneous Other Changes}
+\label{sec:hydro-other-changes}
+
+- Density flux limiter
 

### 2017-02-25 papers/wdmerger_I/submit2/paper.tex
Commit final submitted version of paper I

--- 2016-01-16 papers/wdmerger_I/paper.tex
@@
-\documentclass[twocolumn,numberedappendix]{../aastex6}
+\documentclass[twocolumn,numberedappendix]{aastex6}
 
@@
   \centering
-  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{plots/kh_t2.0_p3_low_res_collage}}}
+  \includegraphics[trim=0.75in 0.45in 0.85in 0.5in, clip, scale = 0.75]{{{kh_t2.0_p3_low_res_collage}}}
   \caption{2D Kelvin-Helmholtz instability test at $t = 2.0$ for the initial 
     conditions given by \autoref{eq:kh_p3_rho} through \autoref{eq:kh_p3_vy},
-    which come from McNally et al. (2012).
+    \textbf{which come from McNally et al. (2012)}.
     The meaning of the rows and columns is the same as in \autoref{fig:kh_p2}.
@@
 the test proposed by R10: as we get to higher flow speeds we need to have higher 
-spatial resolution to compensate for the increased numerical diffusion. The 
+spatial resolution to compensate for the increased numerical diffusion. \textbf{The 
 qualitative accuracy is much lower for the highest bulk velocities for this problem
@@
 substantially affected by altering the perturbation frequency -- the results
-show the same qualitative pattern even if we halve this frequency.
+show the same qualitative pattern even if we halve this frequency.}
 
@@
 
-\input{plots/sources.table}
+\input{sources.table}
 
@@
 It is clear that at the latter resolution (corresponding to physical resolution of 100 km), 
-we have achieved convergent behavior. In the rotating frame, the results also show
+we have achieved convergent behavior. \textbf{In the rotating frame, the results also show
 convergent behavior but the convergence is not as fast with resolution as in the inertial frame; 
@@
 \citet{dan:2012} and \citet{dan:2014} would likely not yield qualitatively convergent results if it were to use the
-same effective mass resolution. Instead the number of zones inside each star should at least be doubled.
+same effective mass resolution. Instead the number of zones inside each star should at least be doubled.}
 
@@
   \centering
-  \includegraphics[scale=0.45]{plots/spatial_convergence_rot0}
+  \includegraphics[scale=0.45]{spatial_convergence_rot0}
   \caption{Distance between the two white dwarfs in the unequal mass system, for the first orbit.
@@
   \centering
-  \includegraphics[scale=0.45]{plots/spatial_convergence_rot1}
+  \includegraphics[scale=0.45]{spatial_convergence_rot1}
   \caption{Distance between the two white dwarfs in the unequal mass system, for the first orbit.
@@
 excellent weak scaling through 4,096 processors. At the largest run,
-the simulation time required is slightly less than 1.5 times the amount
+the simulation time required is slightly \textbf{less} than 1.5 times the amount
 required for the smallest simulation.  This is due entirely to the
@@
 
-The authors thank John Bell, Volker Springel, and Dominic Marcello
+The authors thank John Bell, \textbf{Volker Springel, and Dominic Marcello}
 for helpful discussions on gravity 
@@
 
-\bibliographystyle{../aasjournal}
-\bibliography{../refs}
+\bibliographystyle{aasjournal}
+\input{paper.bbl}
 

### 2017-05-13 documentation/userguide.tex
Remove source/ and tests/, these are in Castro now

@@
 \newcommand{\castro}{{\sf Castro}}
-\newcommand{\boxlib}{{\sf BoxLib}}
+\newcommand{\amrex}{{\sf AMReX}}
 \newcommand{\yt}{{\sf yt}}
@@
 viability of mergers of white dwarfs as a progenitor for Type Ia supernovae. It is a set of 
-source files designed for use with the \texttt{CASTRO} code \citep{castro}.
+driver routines and analysis scripts designed for use with the \castro\ code \citep{castro}.
 
 This software requires a few other software packages to be installed before it can be used.
-The science software includes \href{http://boxlib-codes.github.io/}{BoxLib} and 
-\href{http://boxlib-codes.github.io/Castro/}{CASTRO}. You should also 
-download the \href{https://github.com/BoxLib-Codes/Microphysics}{Microphysics} repository, 
-which stores the equation of state and nuclear network we will use. This software is written 
-in C++ and Fortran, and compilers for each are needed. The gcc compilers are sufficient.
-In addition, it relies on several utilities that are 
+The science software includes \href{https://github.com/AMReX-Codes/amrex}{AMReX} and 
+\href{https://github.com/AMReX-Astro/Castro}{CASTRO}. \castro\ will also need
+the \href{https://github.com/BoxLib-Codes/Microphysics}{Microphysics} repository, 
+which stores the equation of state and nuclear network we will use. Instructions for
+building and installing \castro\ can be found in its documentation.
+This package relies on several utilities that are
 commonly distributed on Linux systems: \texttt{bash}, \texttt{GNU bc}, \texttt{GNU make}, 
@@
 If python and yt are not installed on your system, you will still be able to compile and 
-run this code, but we use yt for all analysis of \texttt{BoxLib} output data other than 
+run this code, but we use yt for all analysis of \amrex\ output data other than 
 the ASCII files that are generated at runtime. We have some routines for this purpose 
-available in Fortran and Python; some of the Fortran tools are provided in \texttt{BoxLib} 
+available in Fortran and Python; some of the Fortran tools are provided in \amrex\ 
 itself, in the \texttt{Tools/Postprocessing} subdirectory, and it is straightforward enough 
@@
 \begin{itemize}
-  \item \texttt{BOXLIB\_HOME}, the location of the \texttt{BoxLib} code's top-level directory
-  \item \texttt{CASTRO\_DIR}, the location of the \texttt{CASTRO} code's top-level directory
+  \item \texttt{AMREX\_HOME}, the location of the \amrex\ code's top-level directory
+  \item \texttt{CASTRO\_DIR}, the location of the \castro\ code's top-level directory
   \item \texttt{WDMERGER\_HOME}, the location of this software's top-level directory
@@
 
-The \texttt{source} directory contains all of the Fortran routines needed to build a \texttt{CASTRO} 
-executable file. You can type \texttt{make} in this directory to build a vanilla executable 
-with the default settings, though usually we will build the software from a different location.
-
 The \texttt{analysis} directory contains the Python analysis scripts used for analyzing output
-data (both in \texttt{BoxLib} and plain-text format).
+data (both in \amrex\ and plain-text format).
 
@@
 that have been or will be submitted based on this software.
-
-The \texttt{tests} directory is designed to hold small versions of problems that can be run in a
-nightly test suite that regularly checks for code errors.
 
@@
 and other I/O information, and the analysis routines. A full description of the physics 
-and the associated numerical algorithms is included in the \texttt{CASTRO} code paper \cite{castro}.
+and the associated numerical algorithms is included in the \castro\ code paper \cite{castro}.
 
@@
 
-\texttt{CASTRO} includes full self-gravity with the Poisson equation, as well as the ability to perform a
+\castro\ includes full self-gravity with the Poisson equation, as well as the ability to perform a
 simulation in a rotating reference frame with fixed rotation period. By default the $z$ axis will 
@@
 coordinate dimension. To try this out you should use \texttt{DIM = 2} in the makefile and
-use \texttt{inputs\_2d} in the \texttt{source/} directory.
+use \texttt{inputs\_2d} in the \texttt{wdmerger} directory in \castro.
 
@@
 
-There are a few \texttt{CASTRO} input parameters that you will typically want to control when running problems. 
+There are a few \castro\ input parameters that you will typically want to control when running problems. 
 Among the most important is whether we are doing rotation (\texttt{castro.do\_rotation}) and if so, 
@@
 \texttt{amr.checkpoint\_files\_output}, respectively; these are the simulation time differences
-between successive plotfiles and checkpoints. Note that \texttt{CASTRO} does not adjust the timestep 
+between successive plotfiles and checkpoints. Note that \castro\ does not adjust the timestep 
 to meet this criterion, so it will generate these at the first timestep that passes this interval.
@@
 In the following list we describe the parameters; you can see their current defaults by 
-examining \texttt{source/probin}. (Some parameters in there are currently beta quality and 
+examining \texttt{probin}. (Some parameters in there are currently beta quality and 
 not recommended for use, and for a couple others there will be no practical reason to change. 
@@
 
-The \texttt{tagging} namelist is described in the AMR chapter of the \texttt{CASTRO} user's guide. At present, we 
+The \texttt{tagging} namelist is described in the AMR chapter of the \castro\ user's guide. At present, we 
 only tag based on absolute values of the density and temperature.
@@
 
-The routines that generate both of these are found in \\\texttt{source/sum\_integrated\_quantities.cpp}. 
-This is a replacement of the \texttt{CASTRO} function that does something similar. We use the same 
+The routines that generate both of these are found in \\\texttt{sum\_integrated\_quantities.cpp}. 
+This is a replacement of the \castro\ function that does something similar. We use the same 
 conventions; for example, if summing over the whole grid and we encounter a region that is covered by 
@@
 The main makefile sets all of our build options (compiler, number of space dimensions, 
-equation of state, etc.) and then includes the \texttt{CASTRO} makefile, which then runs as normal.
+equation of state, etc.) and then includes the \castro\ makefile, which then runs as normal.
 If you want to adjust the build options, then simply add a line before this include. As an 
@@
 and \texttt{probin} files from the root \texttt{wdmerger} source directory; these are the runtime
-files \texttt{CASTRO} needs to determine what your problem setup will be. They contain the default 
+files \castro\ needs to determine what your problem setup will be. They contain the default 
 parameters for merger runs. It also copies any auxiliary files one might need, like the 
-table for the equation of state, as well as the actual \texttt{CASTRO} executable you just built in the 
+table for the equation of state, as well as the actual \castro\ executable you just built in the 
 \texttt{compile/} directory, and a copy of the run utilities that manage the job itself (so that the 
@@
 The \texttt{science/wdmerger\_I/} directory contains a number of example problems that we use
-to assess the reliability of \texttt{CASTRO} in performing hydrodynamics simulations 
+to assess the reliability of \castro\ in performing hydrodynamics simulations 
 with self-gravity. This section includes a list of the existing problems 
@@
 and michael.zingale@stonybrook.edu, respectively. Alan Calder and Doug Swesty at Stony Brook are 
-also co-investigators on this project. The \texttt{CASTRO} code was developed primarily at the 
+also co-investigators on this project. The \castro\ code was developed primarily at the 
 Lawrence Berkeley National Laboratory. Staff there that have contributed directly and 

### 2017-05-13 source/Prob.cpp

@@ -1,1162 +1,0 @@
-#include "Castro.H"
-#include "Castro_F.H"
-
-#include "Gravity.H"
-#include <Gravity_F.H>
-
-#include "ParmParse.H"
-
-#include "buildInfo.H"
-
-#include <fstream>
-
-int Castro::relaxation_is_done = 0;
-int Castro::problem = -1;
-int Castro::use_stopping_criterion = 1;
-
-Real Castro::mass_p = 0.0;
-Real Castro::mass_s = 0.0;
-
-Real Castro::mdot_p = 0.0;
-Real Castro::mdot_s = 0.0;
-
-Real Castro::com_p[3] = { 0.0 };
-Real Castro::com_s[3] = { 0.0 };
-
-Real Castro::vel_p[3] = { 0.0 };
-Real Castro::vel_s[3] = { 0.0 };
-
-Real Castro::rad_p[7] = { 0.0 };
-Real Castro::rad_s[7] = { 0.0 };
-
-Real Castro::vol_p[7] = { 0.0 };
-Real Castro::vol_s[7] = { 0.0 };
-
-Real Castro::rho_avg_p = 0.0;
-Real Castro::rho_avg_s = 0.0;
-
-Real Castro::t_ff_p = 0.0;
-Real Castro::t_ff_s = 0.0;
-
-Real Castro::T_global_max = 0.0;
-Real Castro::rho_global_max = 0.0;
-Real Castro::ts_te_global_max = 0.0;
-
-Real Castro::T_curr_max = 0.0;
-Real Castro::rho_curr_max = 0.0;
-Real Castro::ts_te_curr_max = 0.0;
-
-Real Castro::total_ener_array[num_previous_ener_timesteps] = { 0.0 };
-
-#ifdef DO_PROBLEM_POST_TIMESTEP
-void
-Castro::problem_post_timestep()
-{
-
-    if (level != 0) return;
-
-    int finest_level = parent->finestLevel();
-    Real time = state[State_Type].curTime();
-    Real dt = parent->dtLevel(0);
-
-    if (time == 0.0) dt = 0.0; // dtLevel returns the next timestep for t = 0, so overwrite
-
-    // Update white dwarf masses, positions, velocities, and auxiliary data.
-
-    wd_update(time, dt);
-
-    // If we are doing problem 3, which has an initial relaxation step,
-    // perform any post-timestep updates to assist with the relaxation, then
[…]

### 2017-05-13 source/Prob_nd.F90

@@ -1,234 +1,0 @@
-   subroutine PROBINIT (init,name,namlen,problo,probhi)
-
-     use problem_io_module, only: initialize_io
-     use wdmerger_util_module, only: initialize_problem
-
-     implicit none
-
-     integer :: init, namlen
-     integer :: name(namlen)
-     double precision :: problo(3), probhi(3)
-
-     call initialize_io(name, namlen)
-     call initialize_problem(init)
-
-   end subroutine PROBINIT
-
-
-   ! ::: -----------------------------------------------------------
-   ! ::: This routine is called at problem setup time and is used
-   ! ::: to initialize data on each grid.  
-   ! ::: 
-   ! ::: NOTE:  all arrays have one cell of ghost zones surrounding
-   ! :::        the grid interior.  Values in these cells need not
-   ! :::        be set here.
-   ! ::: 
-   ! ::: INPUTS/OUTPUTS:
-   ! ::: 
-   ! ::: level     => amr level of grid
-   ! ::: time      => time at which to init data             
-   ! ::: lo,hi     => index limits of grid interior (cell centered)
-   ! ::: nstate    => number of state components.  You should know
-   ! :::		   this already!
-   ! ::: state     <=  Scalar array
-   ! ::: delta     => cell size
-   ! ::: xlo,xhi   => physical locations of lower left and upper
-   ! :::              right hand corner of grid.  (does not include
-   ! :::		   ghost region).
-   ! ::: -----------------------------------------------------------
-   subroutine ca_initdata(level,time,lo,hi,nscal, &
-                          state,state_lo,state_hi, &
-                          dx,xlo,xhi)
-
-     use probdata_module
-     use wdmerger_util_module
-     use prob_params_module, only: center, dim
-     use eos_module
-     use meth_params_module, only: NVAR, URHO, UMX, UMY, UMZ, UTEMP, &
-                                   UEDEN, UEINT, UFS, do_rotation, state_in_rotating_frame
-     use network, only: nspec
-     use bl_constants_module
-     use model_parser_module, only: idens_model, itemp_model, ipres_model, ispec_model
-     use initial_model_module, only: interpolate_3d_from_1d
-     use math_module, only: cross_product
-     use castro_util_module, only: position
-     use rotation_frequency_module, only: get_omega
[…]

### 2017-05-13 source/Problem.H

@@ -1,137 +1,0 @@
-// Preprocessor directive for allowing us to do a post-initialization update.
-
-#ifndef DO_PROBLEM_POST_INIT
-#define DO_PROBLEM_POST_INIT
-#endif
-
-// Same for post-restart.
-
-#ifndef DO_PROBLEM_POST_RESTART
-#define DO_PROBLEM_POST_RESTART
-#endif
-
-// Preprocessor directive for allowing us to do a post-timestep update.
-
-#ifndef DO_PROBLEM_POST_TIMESTEP
-#define DO_PROBLEM_POST_TIMESTEP
-#endif
-
-// Routine for determining white dwarf masses, centers of mass and velocities.
-
-void wd_update(Real time, Real dt);
-
-// Volume within white dwarfs at a given density cutoff.
-
-void volInBoundary (Real time, Real& vol_p, Real& vol_s, Real rho_cutoff, bool local = false);
-
-// Calculate gravitational wave signal.
-
-void gwstrain (Real time,
-	       Real& h_plus_1, Real& h_cross_1,
-	       Real& h_plus_2, Real& h_cross_2,
-	       Real& h_plus_3, Real& h_cross_3,
-	       bool local = false);
-
-// Computes standard dot product of two three-vectors.
-
-Real dot_product(const Real a[], const Real b[]);
-
-// Computes standard cross-product of two three-vectors.
-
-void cross_product(const Real a[], const Real b[], Real c[]);
-
-// Computes norm of a three vector.
-
-Real norm(const Real a[]);
-
-// Problem post-initialization routine.
-
-void problem_post_init();
-
-// Problem post-restart routine.
-
-void problem_post_restart();
-
-// End of timestep analytics on the white dwarf properties.
-
-void problem_post_timestep();
-
-// Write out the git hashes for the various parts of the code.
-
-void writeGitHashes(std::ostream& log);
-
-// Check whether we should stop the run.
-
-void check_to_stop(Real time);
-
-// Calculate extrema on grid.
-
-void update_extrema(Real time);
-
-// Update relaxation process.
-
-void update_relaxation(Real time, Real dt);
-
-// Input parameters
-
-static int use_stopping_criterion;
-
-// Data we want to store, for interfacing with the Fortran
-
-static int relaxation_is_done;
-static int problem;
-
-// Stellar masses
-
-static Real mass_p;
-static Real mass_s;
-
-// Rate of change of stellar masses
-
-static Real mdot_p;
-static Real mdot_s;
-
[…]

### 2017-05-13 source/Problem.f90

@@ -1,1022 +1,0 @@
-! Problem-specific Fortran routines that are designed to interact with C++
-
-subroutine problem_checkpoint(int_dir_name, len) bind(C, name="problem_checkpoint")
-
-  ! called by the IO processor during checkpoint
-
-  use bl_IO_module
-  use probdata_module, only: com_P, com_S, vel_P, vel_S, mass_P, mass_S, t_ff_P, t_ff_S, &
-                             T_global_max, rho_global_max, ts_te_global_max
-  use prob_params_module, only: center
-  use meth_params_module, only: rot_period
-  use probdata_module, only: jobIsDone, signalJobIsNotDone, num_previous_ener_timesteps, total_ener_array, &
-                             relaxation_is_done
-
-  implicit none
-
-  integer :: len
-  integer :: int_dir_name(len)
-  character (len=len) :: dir
-
-  integer :: i, un
-
-  ! dir will be the string name of the checkpoint directory
-  do i = 1, len
-     dir(i:i) = char(int_dir_name(i))
-  enddo
-
-  un = unit_new()
-  open (unit=un, file=trim(dir)//"/COM", status="unknown")
-
-100 format(1x, g30.20)
-200 format(1x, g30.20, 1x, g30.20)
-300 format(1x, g30.20, 1x, g30.20, 1x, g30.20)
-
-  write (un,300) center(1), center(2), center(3)
-  write (un,200) mass_P,   mass_S
-  write (un,200) com_P(1), com_S(1)
-  write (un,200) com_P(2), com_S(2)
-  write (un,200) com_P(3), com_S(3)
-  write (un,200) vel_P(1), vel_S(1)
-  write (un,200) vel_P(2), vel_S(2)
-  write (un,200) vel_P(3), vel_S(3)
-  write (un,200) t_ff_P,   t_ff_S
-
-  close (un)
-
-
-
-  open (unit=un, file=trim(dir)//"/Rotation", status="unknown")
-
-  write (un,100) rot_period
-
-  close (un)
-
-
-
-  open (unit=un, file=trim(dir)//"/Extrema", status="unknown")
-
-  write (un,100) T_global_max
-  write (un,100) rho_global_max
-  write (un,100) ts_te_global_max
-
-  close (un)
-
-
-
-  open (unit=un, file=trim(dir)//"/Energy", status="unknown")
-
-  do i = 1, num_previous_ener_timesteps
-
-     write (un,100) total_ener_array(i)
-
-  enddo
-
[…]

### 2017-05-13 source/Problem_Derive_F.H

@@ -1,132 +1,0 @@
-extern "C"
-{
-
-  void ca_derinertialmomentumx(BL_FORT_FAB_ARG_3D(der),const int* nvar,
-			       const BL_FORT_FAB_ARG_3D(data),const int* ncomp,
-			       const int* lo, const int* hi,
-			       const int* domain_lo, const int* domain_hi,
-			       const Real* delta, const Real* xlo,
-			       const Real* time, const Real* dt, const int* bcrec, 
-			       const int* level, const int* grid_no);
-
-  void ca_derinertialmomentumy(BL_FORT_FAB_ARG_3D(der),const int* nvar,
-			       const BL_FORT_FAB_ARG_3D(data),const int* ncomp,
-			       const int* lo, const int* hi,
-			       const int* domain_lo, const int* domain_hi,
-			       const Real* delta, const Real* xlo,
-			       const Real* time, const Real* dt, const int* bcrec, 
-			       const int* level, const int* grid_no);
-
-  void ca_derinertialmomentumz(BL_FORT_FAB_ARG_3D(der),const int* nvar,
-			       const BL_FORT_FAB_ARG_3D(data),const int* ncomp,
-			       const int* lo, const int* hi,
-			       const int* domain_lo, const int* domain_hi,
-			       const Real* delta, const Real* xlo,
-			       const Real* time, const Real* dt, const int* bcrec, 
-			       const int* level, const int* grid_no);
-
-  void ca_derinertialangmomx(BL_FORT_FAB_ARG_3D(der),const int* nvar,
-			     const BL_FORT_FAB_ARG_3D(data),const int* ncomp,
-			     const int* lo, const int* hi,
-			     const int* domain_lo, const int* domain_hi,
-			     const Real* delta, const Real* xlo,
-			     const Real* time, const Real* dt, const int* bcrec, 
-			     const int* level, const int* grid_no);
-
-  void ca_derinertialangmomy(BL_FORT_FAB_ARG_3D(der),const int* nvar,
-			     const BL_FORT_FAB_ARG_3D(data),const int* ncomp,
-			     const int* lo, const int* hi,
-			     const int* domain_lo, const int* domain_hi,
-			     const Real* delta, const Real* xlo,
-			     const Real* time, const Real* dt, const int* bcrec, 
-			     const int* level, const int* grid_no);
-
-  void ca_derinertialangmomz(BL_FORT_FAB_ARG_3D(der),const int* nvar,
-			     const BL_FORT_FAB_ARG_3D(data),const int* ncomp,
[…]

### 2017-05-13 source/Problem_Derives.H

@@ -1,107 +1,0 @@
-
-#ifdef ROTATION
-    //
-    // Momemtum
-    //
-    derive_lst.add("inertial_momentum_x",IndexType::TheCellType(),1,ca_derinertialmomentumx,the_same_box);
-    derive_lst.addComponent("inertial_momentum_x",desc_lst,State_Type,Density,1);
-    derive_lst.addComponent("inertial_momentum_x",desc_lst,State_Type,Xmom,3);
-
-    derive_lst.add("inertial_momentum_y",IndexType::TheCellType(),1,ca_derinertialmomentumy,the_same_box);
-    derive_lst.addComponent("inertial_momentum_y",desc_lst,State_Type,Density,1);
-    derive_lst.addComponent("inertial_momentum_y",desc_lst,State_Type,Xmom,3);
-
-    derive_lst.add("inertial_momentum_z",IndexType::TheCellType(),1,ca_derinertialmomentumz,the_same_box);
-    derive_lst.addComponent("inertial_momentum_z",desc_lst,State_Type,Density,1);
-    derive_lst.addComponent("inertial_momentum_z",desc_lst,State_Type,Xmom,3);
-
-    //
-    // Angular momentum
-    //
-    derive_lst.add("inertial_angular_momentum_x",IndexType::TheCellType(),1,ca_derinertialangmomx,the_same_box);
-    derive_lst.addComponent("inertial_angular_momentum_x",desc_lst,State_Type,Density,1);
-    derive_lst.addComponent("inertial_angular_momentum_x",desc_lst,State_Type,Xmom,3);
-
-    derive_lst.add("inertial_angular_momentum_y",IndexType::TheCellType(),1,ca_derinertialangmomy,the_same_box);
-    derive_lst.addComponent("inertial_angular_momentum_y",desc_lst,State_Type,Density,1);
-    derive_lst.addComponent("inertial_angular_momentum_y",desc_lst,State_Type,Xmom,3);
-
-    derive_lst.add("inertial_angular_momentum_z",IndexType::TheCellType(),1,ca_derinertialangmomz,the_same_box);
-    derive_lst.addComponent("inertial_angular_momentum_z",desc_lst,State_Type,Density,1);
-    derive_lst.addComponent("inertial_angular_momentum_z",desc_lst,State_Type,Xmom,3);
-
-    //
-    // Radial momentum
-    //
[…]

### 2017-05-13 source/Problem_F.H

@@ -1,116 +1,0 @@
-extern "C"
-{
-
-  void problem_checkpoint(int * int_dir_name, int * len);
-
-  void problem_restart(int * int_dir_name, int * len);
-
-  void wdcom(BL_FORT_FAB_ARG_3D(rho), 
-	     BL_FORT_FAB_ARG_3D(xmom), 
-	     BL_FORT_FAB_ARG_3D(ymom), 
-	     BL_FORT_FAB_ARG_3D(zmom),
-	     BL_FORT_FAB_ARG_3D(pmask),
-	     BL_FORT_FAB_ARG_3D(smask),
-	     BL_FORT_FAB_ARG_3D(vol),
-	     const int* lo, const int* hi,
-	     const Real* dx, const Real* time,
-	     Real* com_p_x, Real* com_p_y, Real* com_p_z,
-	     Real* com_s_x, Real* com_s_y, Real* com_s_z,
-	     Real* vel_p_x, Real* vel_p_y, Real* vel_p_z,
-	     Real* vel_s_x, Real* vel_s_y, Real* vel_s_z,
-	     Real* m_p, Real* m_s);
-
-  void ca_volumeindensityboundary(BL_FORT_FAB_ARG_3D(rho),
-				  BL_FORT_FAB_ARG_3D(pmask),
-				  BL_FORT_FAB_ARG_3D(smask),
-				  BL_FORT_FAB_ARG_3D(vol),
-				  const int* lo, const int* hi,
-				  const Real* dx,  
-				  Real* volp, Real* vols,
-				  Real* rho_cutoff);
-
-  void get_single_star(int& flag);
-
-  void set_star_data(const Real com_p[], const Real com_s[], 
-		     const Real vel_p[], const Real vel_s[], 
-		     const Real* mass_p, const Real* mass_s,
-		     const Real* t_ff_p, const Real* t_ff_s);
-
-  void get_star_data(Real com_p[], Real com_s[], 
-		     Real vel_p[], Real vel_s[], 
-		     Real* mass_p, Real* mass_s,
-		     Real* t_ff_p, Real* t_ff_s);
-
-  void quadrupole_tensor_double_dot(BL_FORT_FAB_ARG_3D(rho),
-				    BL_FORT_FAB_ARG_3D(xmom),
-				    BL_FORT_FAB_ARG_3D(ymom),
-				    BL_FORT_FAB_ARG_3D(zmom),
[…]

### 2017-05-13 source/Rotation_frequency.F90

@@ -1,47 +1,0 @@
-module rotation_frequency_module
-
-  implicit none
-
-  private
-
-  public get_omega, get_domegadt
-
-contains
-
-  function get_omega(time) result(omega)
-
-    use meth_params_module, only: rot_period, rot_axis
-    use bl_constants_module, only: ZERO, TWO, M_PI
-    use fundamental_constants_module, only: Gconst
-
-    implicit none
-
-    double precision :: time
-    double precision :: omega(3)
-
-    ! If rot_period is less than zero, that means rotation is disabled, and so we should effectively
-    ! shut off the source term by setting omega = 0. Note that by default rot_axis == 3 for Cartesian
-    ! coordinates and rot_axis == 2 for cylindrical coordinates.
-
-    omega = ZERO
-
-    omega(rot_axis) = TWO * M_PI / rot_period
-
-  end function get_omega
-
-
-
-  function get_domegadt(time) result(domegadt)
-
-    use bl_constants_module, only: ZERO
-
-    implicit none
-
-    double precision :: time
-    double precision :: domegadt(3)
-
-    domegadt = ZERO
-
-  end function get_domegadt
-
-end module rotation_frequency_module

### 2017-05-13 source/bc_fill_nd.F90

@@ -1,281 +1,0 @@
-module bc_fill_module
-
-  implicit none
-
-  public
-
-contains
-
-  subroutine ca_hypfill(adv,adv_lo,adv_hi,domlo,domhi,delta,xlo,time,bc) &
-       bind(C, name="ca_hypfill")
-
-     use bl_constants_module, only: HALF
-     use meth_params_module, only: NVAR
-     use prob_params_module, only: center, dim
-     use probdata_module, only: fill_ambient_bc
-     use wdmerger_util_module, only: fill_ambient
-
-     implicit none
-
-     include 'bc_types.fi'
-
-     integer          :: adv_lo(3), adv_hi(3)
-     integer          :: bc(dim,2,NVAR)
-     integer          :: domlo(3), domhi(3)
-     double precision :: delta(3), xlo(3), time
-     double precision :: adv(adv_lo(1):adv_hi(1),adv_lo(2):adv_hi(2),adv_lo(3):adv_hi(3),NVAR)
-
-     integer          :: i, j, k, n
-     double precision :: loc(3)
-
-     ! First, use the generic filling routines to make sure we have valid data everywhere
-     ! on physical domain ghost cells.
-     
-     do n = 1,NVAR
-        call filcc_nd(adv(:,:,:,n),adv_lo,adv_hi,domlo,domhi,delta,xlo,bc(:,:,n))
-     enddo
-     
-     ! Override the generic routine at the physical boundaries by
-     ! setting the material to the ambient state. Note that we
-     ! don't want to do this for interior/periodic boundaries, which have bc == 0.
-
-     if (fill_ambient_bc) then
-
-        do k = adv_lo(3), adv_hi(3)
-           loc(3) = xlo(3) + dble(k - adv_lo(3) + HALF)*delta(3) - center(3)
-           do j = adv_lo(2), adv_hi(2)
-              loc(2) = xlo(2) + dble(j - adv_lo(2) + HALF)*delta(2) - center(2)
-              do i = adv_lo(1), adv_hi(1)
-                 loc(1) = xlo(1) + dble(i - adv_lo(1) + HALF)*delta(1) - center(1)
-
-                 if (i .lt. domlo(1) .and. bc(1,1,1) .ne. 0) then
-                    call fill_ambient(adv(i,j,k,:), loc, time)
[…]

### 2017-05-13 source/binary.f90

@@ -1,297 +1,0 @@
-module binary_module
-
-  ! This module contains a number of routines that are designed to
-  ! calculate generic properties of binary orbits.
-  
-  implicit none
-
-contains
-
-  ! Given the mass ratio q of two stars (assumed to be q = M_1 / M_2), 
-  ! compute the effective Roche radii of the stars, normalized to unity, 
-  ! using the approximate formula of Eggleton (1983). Optionally we can
-  ! pass in a distance scale.
-  
-  subroutine get_roche_radii(mass_ratio, r_1, r_2, a)
-
-    use bl_constants_module, only: ONE, TWO3RD, THIRD
-
-    implicit none
-
-    double precision, intent(in   ) :: mass_ratio
-    double precision, intent(inout) :: r_1, r_2
-    double precision, intent(in   ), optional :: a
-
-    double precision :: q
-    double precision :: c1, c2
-
-    double precision :: scale
-
-    if (present(a)) then
-       scale = a
-    else
-       scale = ONE
-    endif
-
-    c1 = 0.49d0
-    c2 = 0.60d0
-
-    q = mass_ratio
-
-    r_1 = scale * c1 * q**(TWO3RD) / (c2 * q**(TWO3RD) + LOG(ONE + q**(THIRD)))
-
-    q = ONE / q
-
-    r_2 = scale * c1 * q**(TWO3RD) / (c2 * q**(TWO3RD) + LOG(ONE + q**(THIRD)))
-
-  end subroutine get_roche_radii
-
-
-  
-  ! Calculate Lagrange points. In each case we give the zone index
-  ! closest to it (assuming we're on the coarse grid).
-
-  subroutine get_lagrange_points(mass_1, mass_2, com_1, com_2, &
-                                 L1, L2, L3) bind(C, name="get_lagrange_points")
-
-    use bl_constants_module
-
-    implicit none
-
-    double precision :: mass_1, mass_2
-    double precision :: com_1(3), com_2(3)
-    double precision :: L1(3), L2(3), L3(3)
-    
-    double precision :: r ! Distance from Lagrange point to primary
-    double precision :: a ! Distance between secondary and primary
-
-    double precision :: r1, r2
-
-    ! Don't try to calculate the Lagrange points if the secondary
-    ! is already gone.
-
-    if (mass_2 < ZERO) return
-
-    a = sqrt(sum((com_2 - com_1)**2))
-
[…]

### 2017-05-13 source/ext_src_nd.F90

@@ -1,145 +1,0 @@
-
-     subroutine ca_ext_src(lo,hi,&
-                           old_state,os_lo,os_hi,&
-                           new_state,ns_lo,ns_hi,&
-                           src,src_lo,src_hi,problo,dx,time,dt)
-
-       use meth_params_module,  only: NVAR, URHO, UMX, UMZ, UEDEN
-       use prob_params_module,  only: center
-       use bl_constants_module, only: ZERO, HALF, ONE, TWO
-       use probdata_module,     only: problem, relaxation_damping_timescale, radial_damping_factor, &
-                                      t_ff_P, t_ff_S, axis_1, axis_2, axis_3, relaxation_implicit
-       use castro_util_module,  only: position
-       use wdmerger_util_module, only: inertial_velocity
-#ifdef HYBRID_MOMENTUM
-       use hybrid_advection_module, only: linear_to_hybrid
-       use meth_params_module, only: UMR, UMP
-#endif
-
-       implicit none
-
-       integer          :: lo(3),hi(3)
-       integer          :: os_lo(3),os_hi(3)
-       integer          :: ns_lo(3),ns_hi(3)
-       integer          :: src_lo(3),src_hi(3)
-       double precision :: old_state(os_lo(1):os_hi(1),os_lo(2):os_hi(2),os_lo(3):os_hi(3),NVAR)
-       double precision :: new_state(ns_lo(1):ns_hi(1),ns_lo(2):ns_hi(2),ns_lo(3):ns_hi(3),NVAR)
-       double precision :: src(src_lo(1):src_hi(1),src_lo(2):src_hi(2),src_lo(3):src_hi(3),NVAR)
-       double precision :: problo(3),dx(3),time,dt
-
-       ! Local variables
-
-       double precision :: radial_damping_timescale
-       double precision :: dynamical_timescale, damping_factor
-       double precision :: loc(3), R_prp, sinTheta, cosTheta, v_rad, Sr(3)
-       integer          :: i, j, k
-       double precision :: new_mom(3), old_mom(3), rhoInv
-
-       ! Note that this function exists in a tiling region so we should only 
-       ! modify the zones between lo and hi. 
-
-       src(lo(1):hi(1),lo(2):hi(2),lo(3):hi(3),:) = ZERO
-
-       ! The relevant dynamical timescale for determining our source term timescales should be
-       ! the larger of the two WD timescales. Generally this should be the secondary, but we'll
-       ! be careful just in case.
-
[…]

### 2017-05-13 source/initial_model.f90

@@ -1,370 +1,0 @@
-! This model contains routines to:
-! Generate a 1D isothermal white dwarf in hydrostatic equilibrium
-! Interpolate a 1D WD model onto a 3D Cartesian grid
-
-module initial_model_module
-
-  use bl_types
-  use bl_constants_module
-  use bl_error_module, only: bl_error
-  use eos_module, only: eos_input_rt, eos
-  use eos_type_module, only: eos_t
-  use network, only: nspec
-  use model_parser_module, only: itemp_model, idens_model, ipres_model, ispec_model
-  use fundamental_constants_module, only: Gconst, M_solar
-  use interpolate_module, only: interpolate
-  use meth_params_module, only: small_temp
-
-  type :: initial_model
-
-     ! Physical characteristics
-
-     double precision :: mass = ZERO
-     double precision :: envelope_mass = ZERO
-     double precision :: central_density = ZERO
-     double precision :: central_temp = ZERO
-     double precision :: min_density = ZERO
-     double precision :: radius = ZERO
-
-     ! Composition
-
-     double precision :: core_comp(nspec) = ZERO
-     double precision :: envelope_comp(nspec) = ZERO
-
-     ! Model storage
-
-     double precision :: dx
-     integer          :: npts
-     double precision :: mass_tol, hse_tol
-
-     double precision, allocatable :: r(:), rl(:), rr(:)
-     double precision, allocatable :: M_enclosed(:), g(:)
-     type (eos_t), allocatable :: state(:)
-
-  end type initial_model
-  
-contains
-
-  subroutine initialize_model(model, dx, npts, mass_tol, hse_tol)
-
-    implicit none
-
-    type (initial_model) :: model
-    integer :: npts
-    double precision :: dx, mass_tol, hse_tol
-
-    integer :: i
-
-    model % dx = dx
-    model % npts = npts
-    model % mass_tol = mass_tol
-    model % hse_tol = hse_tol
-
-    allocate(model % r(npts))
-    allocate(model % rl(npts))
-    allocate(model % rr(npts))
-    allocate(model % M_enclosed(npts))
-    allocate(model % g(npts))
-    allocate(model % state(npts))
-
-    do i = 1, npts
-
-       model % rl(i) = (dble(i) - ONE)*dx
-       model % rr(i) = (dble(i)      )*dx
-       model % r(i)  = HALF*(model % rl(i) + model % rr(i))
-
-    enddo
-
[…]

### 2017-05-13 source/probdata.f90

@@ -1,293 +1,0 @@
-module probdata_module
-
-  use network, only: nspec, network_species_index
-  use eos_type_module, only: eos_t
-  use eos_module, only: eos_input_rt, eos
-  use bl_constants_module, only: ZERO, THIRD, HALF, ONE, TWO, THREE, M_PI, FOUR
-  use fundamental_constants_module, only: Gconst, M_solar, AU
-  use initial_model_module, only: initial_model
-
-  ! Initial stellar properties
-  ! Note that the envelope mass is included within the total mass of the star
-
-  double precision, save :: mass_P = ONE
-  double precision, save :: mass_S = ONE
-  double precision, save :: central_density_P = -ONE
-  double precision, save :: central_density_S = -ONE
-  double precision, save :: stellar_temp = 1.0d7
-  double precision, save :: primary_envelope_mass, secondary_envelope_mass
-  double precision, save :: primary_envelope_comp(nspec), secondary_envelope_comp(nspec)
-
-
-
-  ! Ambient medium
-
-  double precision, save :: ambient_density = 1.0d-4
-  double precision, save :: ambient_temp = 1.0d7
-  double precision, save :: ambient_comp(nspec)
-
-
-
-  ! Smallest allowed velocity on the grid
-
-  double precision, save :: smallu = ZERO
-
-
-
-  ! Parameters for interpolation from 1D model to 3D model:
-
-  ! Number of sub-grid-scale zones to use
-
-  integer, save :: nsub = 1
-
-  ! Default to interpolation that preserves temperature; otherwise, use pressure
-
-  logical, save :: interp_temp = .true.
-
-
-
-  ! Method for determining the initial problem setup.
-  !
-  ! 0 = Collision; distance determined by a multiple of the secondary WD radius
-  ! 1 = Keplerian orbit; distance determined by the rotation period
-  ! 2 = Keplerian orbit; distance set so that the secondary exactly fills its Roche lobe radius
-  ! 3 = Problem 2 with an initial relaxation step
-  ! 4 = Free-fall; distance determined by a multiple of the secondary WD radius
-
-  integer, save :: problem = 2
-
-
-
-  ! If we're automatically determining the initial distance based on the Roche lobe
-  ! radii for problem 3, this is the sizing factor we use. We set the default value
-  ! to be a large enough distance so that the system is close to stable.
-
[…]

### 2017-05-13 source/problem_derive_nd.f90

@@ -1,728 +1,0 @@
-! Derive momentum, given input vector of the grid momenta.
-
-subroutine ca_derinertialmomentumx(p,p_lo,p_hi,ncomp_p, &
-                                   u,u_lo,u_hi,ncomp_u, &
-                                   lo,hi,domlo,domhi, &
-                                   dx,xlo,time,dt,bc,level,grid_no) &
-                                   bind(C,name='ca_derinertialmomentumx')
-
-  use bl_constants_module, only: HALF
-  use wdmerger_util_module, only: inertial_velocity
-  use prob_params_module, only: center
-
-  implicit none
-
-  integer          :: p_lo(3), p_hi(3), ncomp_p ! == 1
-  integer          :: u_lo(3), u_hi(3), ncomp_u ! == 4
-  integer          :: lo(3), hi(3), domlo(3), domhi(3)
-  double precision :: p(p_lo(1):p_hi(1),p_lo(2):p_hi(2),p_lo(3):p_hi(3),ncomp_p)
-  double precision :: u(u_lo(1):u_hi(1),u_lo(2):u_hi(2),u_lo(3):u_hi(3),ncomp_u)
-  double precision :: dx(3), xlo(3), time, dt
-  integer          :: bc(3,2,ncomp_u), level, grid_no
-
-  integer          :: i, j, k
-  double precision :: loc(3), vel(3), mom(3), rho
-
-  do k = lo(3), hi(3)
-     loc(3) = xlo(3) + (dble(k - lo(3)) + HALF) * dx(3) - center(3)
-     do j = lo(2), hi(2)
-        loc(2) = xlo(2) + (dble(j - lo(2)) + HALF) * dx(2) - center(2)
-        do i = lo(1), hi(1)
-           loc(1) = xlo(1) + (dble(i - lo(1)) + HALF) * dx(1) - center(1)
-
-           rho = u(i,j,k,1)
-           vel = u(i,j,k,2:4) / rho
-           mom = rho * inertial_velocity(loc, vel, time)
-           p(i,j,k,1) = mom(1)
-
-        enddo
-     enddo
-  enddo
-
-end subroutine ca_derinertialmomentumx
-
-
-
-! Derive momentum, given input vector of the grid momenta.
-
-subroutine ca_derinertialmomentumy(p,p_lo,p_hi,ncomp_p, &
-                                   u,u_lo,u_hi,ncomp_u, &
[…]

### 2017-05-13 source/problem_io.f90

@@ -1,39 +1,0 @@
-module problem_io_module
-
-  implicit none
-
-  ! For determining if we are the I/O processor.
-  
-  logical, save :: ioproc
-
-  ! Probin file
-
-  character (len=:), allocatable, save :: probin  
-
-contains
-
-  subroutine initialize_io(name, namlen)
-    
-    implicit none
-
-    integer :: namlen, i
-    integer :: name(namlen)
-
-    integer :: is_ioprocessor
-    
-    ! Build "probin" filename -- the name of the file containing the fortin namelist.
-    
-    allocate(character(len=namlen) :: probin)
-    do i = 1, namlen
-       probin(i:i) = char(name(i))
-    enddo
-
-    ! Determine whether we are the I/O procoessor.
-    
-    call bl_pd_is_ioproc(is_ioprocessor)
-
-    ioproc = is_ioprocessor .ne. 0
-    
-  end subroutine initialize_io
-  
-end module problem_io_module

### 2017-05-13 source/problem_tagging_nd.F90

@@ -1,188 +1,0 @@
-module problem_tagging_module
-
-  implicit none
-
-  public
-
-contains
-
-  subroutine set_problem_tags(tag,tag_lo,tag_hi, &
-                              state,state_lo,state_hi, &
-                              set,clear,&
-                              lo,hi,&
-                              dx,problo,time,level) bind(C,name='set_problem_tags')
-
-    use bl_constants_module, only: ZERO, HALF, TWO
-    use meth_params_module, only: NVAR, URHO, UTEMP
-    use prob_params_module, only: center, probhi, dim, Symmetry, physbc_lo, physbc_hi, &
-                                  n_error_buf, ref_ratio, blocking_factor, domlo_level, domhi_level
-    use probdata_module, only: max_tagging_radius, &
-                               max_stellar_tagging_level, &
-                               max_temperature_tagging_level, &
-                               max_center_tagging_level, &
-                               roche_tagging_factor, &
-                               stellar_density_threshold, &
-                               temperature_tagging_threshold, &
-                               center_tagging_radius, &
-                               com_P, com_S, roche_rad_P, roche_rad_S, &
-                               problem
-
-    implicit none
-
-    integer          :: lo(3),hi(3)
-    integer          :: state_lo(3),state_hi(3)
-    integer          :: tag_lo(3),tag_hi(3)
-    double precision :: state(state_lo(1):state_hi(1), &
-                        state_lo(2):state_hi(2), &
-                        state_lo(3):state_hi(3),NVAR)
-    integer          :: tag(tag_lo(1):tag_hi(1),tag_lo(2):tag_hi(2),tag_lo(3):tag_hi(3))
-    double precision :: problo(3),dx(3),time
-    integer          :: level,set,clear
-
-    integer          :: i, j, k, n
-    double precision :: loc(3), r, r_P, r_S
-
-    logical          :: outer_boundary_test(3)
-    integer          :: boundary_buf(3), idx(3)
-
-    do k = lo(3), hi(3)
-       loc(3) = problo(3) + (dble(k) + HALF) * dx(3)
-
-       do j = lo(2), hi(2)
-          loc(2) = problo(2) + (dble(j) + HALF) * dx(2)
-
-          do i = lo(1), hi(1)
-             loc(1) = problo(1) + (dble(i) + HALF) * dx(1)
-
-             if (level < max_stellar_tagging_level) then
-
-                if (problem .eq. 0 .or. problem .eq. 4) then
-
[…]

### 2017-05-13 source/sum_integrated_quantities.cpp

@@ -1,851 +1,0 @@
-#include <cmath>
-
-#include <iomanip>
-
-#include <vector>
-
-#include <Castro.H>
-#include <Castro_F.H>
-#include <Geometry.H>
-
-#include <Problem.H>
-#include <Problem_F.H>
-
-#include <Gravity.H>
-#include <Gravity_F.H>
-
-void
-Castro::sum_integrated_quantities ()
-{
-    if (level > 0) return;
-
-    bool local_flag = true;
-
-    int finest_level  = parent->finestLevel();
-    Real time         = state[State_Type].curTime();
-    Real dt           = parent->dtLevel(0);
-
-    if (time == 0.0) dt = 0.0; // dtLevel returns the next timestep for t = 0, so overwrite
-
-    int timestep = parent->levelSteps(0);
-
-    Real mass                 = 0.0;
-    Real momentum[3]          = { 0.0 };
-    Real angular_momentum[3]  = { 0.0 };
-    Real hybrid_momentum[3]   = { 0.0 };
-    Real rho_E                = 0.0;
-    Real rho_e                = 0.0;
-    Real rho_K                = 0.0;
-    Real rho_phi              = 0.0;
-    Real rho_phirot           = 0.0;
-
-    // Total energy on the grid, including decomposition
-    // into the various components.
-
-    Real gravitational_energy = 0.0;
-    Real kinetic_energy       = 0.0;
-    Real gas_energy           = 0.0;
-    Real rotational_energy    = 0.0;
-    Real internal_energy      = 0.0;
-    Real total_energy         = 0.0;
-    Real total_E_grid         = 0.0;
-
-    // Rotation frequency.
-
-    Real omega[3] = { 0.0 };
-
-    get_omega_vec(omega, time);
-
-    // Mass transfer rate
-
-    Real mdot = 0.5 * (std::abs(mdot_p) + std::abs(mdot_s));
-
-    // Center of mass of the system.
-
-    Real com[3]       = { 0.0 };
-    Real com_vel[3]   = { 0.0 };
-
-    // Stellar masses.
-
-    Real mass_p       = 0.0;
-    Real mass_s       = 0.0;
-
-    // Distance between the WDs.
-
-    Real wd_dist[3] = { 0.0 };
-    Real wd_dist_init[3] = { 0.0 };
-
[…]

### 2017-05-13 source/wdmerger_util.f90

@@ -1,1176 +1,0 @@
-module wdmerger_util_module
-
-  use probdata_module
-
-  implicit none
-
-contains
-
-  ! This routine calls all of the other subroutines at the beginning
-  ! of a simulation to fill in the basic problem data.
-
-  subroutine initialize_problem(init_in)
-
-    use bl_error_module, only: bl_error
-    use prob_params_module, only: dim
-
-    implicit none
-
-    integer :: init_in
-
-    init = init_in
-
-    ! Read in the namelist to set problem parameters.
-
-    call read_namelist
-
-    ! Establish binary parameters and create initial models.
-
-    call binary_setup
-
-    ! Set small_pres and small_ener.
-
-    call set_small
-
-  end subroutine initialize_problem
-
-
-
-  ! This routine reads in the namelist
-
-  subroutine read_namelist
-
-    use bl_constants_module, only: ZERO
-    use meth_params_module
-    use prob_params_module, only: dim, coord_type
-    use problem_io_module, only: probin
-
-    implicit none
-
-    integer :: untin
-
-    ! Read namelist to override the module defaults.
-
-    untin = 9 
-    open(untin,file=probin,form='formatted',status='old')
-    read(untin,fortin)
-    close(unit=untin)
-
-    ! Convert masses from solar masses to grams.
-
-    mass_P = mass_P * M_solar
-    mass_S = mass_S * M_solar
-
-    max_he_wd_mass = max_he_wd_mass * M_solar
-    max_hybrid_wd_mass = max_hybrid_wd_mass * M_solar
-    max_co_wd_mass = max_co_wd_mass * M_solar
-
-    hybrid_wd_he_shell_mass = hybrid_wd_he_shell_mass * M_solar
-    co_wd_he_shell_mass = co_wd_he_shell_mass * M_solar
-
-    if (mass_S < ZERO .and. central_density_S < ZERO) single_star = .true.
-
-    ! Make sure that the primary mass is really the larger mass
-
-    call ensure_primary_mass_larger
-
-    ! We enforce that the orbital plane is (z, phi) for two-dimensional problems,
-    ! with rotation about z = 0 along the radial axis.
-
-    if (dim .eq. 2) then
-
-       if (coord_type .ne. 1) then
-          call bl_error("We only support cylindrical coordinates in two dimensions. Set coord_type == 1.")
-       endif
-
-       axis_1 = 2
-       axis_2 = 3
-       rot_axis = 1
-
-    endif
-
[…]

### 2017-07-09 papers/wdmerger_II/paper.tex
Update names, URLs

@@
 % Software names
+\newcommand{\amrex}{\texttt{AMReX}}
 \newcommand{\boxlib}{\texttt{BoxLib}}
@@
 In this paper we seek to add another view to the problem using the reactive hydrodynamics
-code \castro\ \citep{castro} and its underlying adaptive mesh refinement library \boxlib.
+code \castro\ \citep{castro} and its underlying adaptive mesh refinement library \amrex\footnote{\amrex is the successor to \boxlib.}.
 This paper is a follow-up to \citet{wdmergerI} (hereafter, Paper I), which established
@@
 All four of these networks have been ported into a form that is consistent
-with the \boxlib\ codes, in the freely available \microphysics\ code
-repository\footnote{\microphysics\ can be obtained at \url{https://github.com/BoxLib-Codes/Microphysics}.},
+with the \amrex\ codes, in the freely available \microphysics\ code
+repository\footnote{\microphysics\ can be obtained at \url{https://github.com/starkiller-astro/Microphysics}.},
 a collection of microphysical routines that are designed to be used in our
@@
 that is similar to the integrator which ships with
-the original versions of the networks mentioned above. Previous work using the \boxlib\
+the original versions of the networks mentioned above. Previous work using the \amrex\
 codes typically used the classic \vode\ integrator \citep{vode}. We do maintain a version
@@
 retry it from the starting state using that many levels of refinement.
-As mentioned in \autoref{sec:burnerhydrocoupling}, the ability to reject
+As mentioned in \autoref{sec:hydrocoupling}, the ability to reject
 a timestep and retry it is functionality that
@@
 
-\software{BoxLib\footnote{\url{https://github.com/BoxLib-Codes/BoxLib}},
-          CASTRO \citep{castro} (https://github.com/BoxLib-Codes/Castro),
-          wdmerger \citep{wdmergerI} (\url{https://github.com/BoxLib-Codes/wdmerger}),
+\software{\amrex\ (\url{https://github.com/AMReX-Codes/Amrex}),
+          \castro\ \citep{castro} (\url{https://github.com/AMReX-Astro/Castro}),
+          \wdmerger\ \citep{wdmergerI} (\url{https://github.com/AMReX-Astro/wdmerger}),
           GCC (\url{https://gcc.gnu.org/}),
           python (\url{https://www.python.org/}),
-          matplotlib \citep{matplotlib}, \url{http://matplotlib.org/}),
-          yt \citep{yt}, \url{http://yt-project.org/})}
+          matplotlib \citep{matplotlib} (\url{http://matplotlib.org/}),
+          yt \citep{yt} (\url{http://yt-project.org/})}
 

### 2017-07-09 documentation/userguide.tex
Update Microphysics URL

@@
 \href{https://github.com/AMReX-Astro/Castro}{CASTRO}. \castro\ will also need
-the \href{https://github.com/BoxLib-Codes/Microphysics}{Microphysics} repository, 
+the \href{https://github.com/starkiller-astro/Microphysics}{Microphysics} repository, 
 which stores the equation of state and nuclear network we will use. Instructions for

### 2017-07-09 papers/wdmerger_II/paper.tex
Add a description of the hydro flux limiter

@@
 \shorttitle{WD Mergers. II. Collisions}
-\shortauthors{Katz et al. (2016)}
+\shortauthors{Katz et al. (2017)}
 
@@
 The hydrodynamical equations solved in this paper were presented in \citet{wdmergerI} (hereafter, Paper I).
+Revisions to the core algorithm since then are presented in \autoref{app:hydro}.
 In this paper we focus our discussion on our implementation of the nuclear
@@
 
+\appendix
+
+\section{Hydrodynamics Algorithm Updates}
+\label{app:hydro}
+
+The core \castro\ algorithm for hydrodynamic evolution is presented in
+\citet{wdmergerI}, and is largely unchanged in the present work. This appendix
+focuses on minor revisions to the algorithm that have occurred since then.
+
+\subsection{Flux Limiting}
+\label{app:hydro:flux_limiting}
+
+Multi-dimensional hydrodynamic white dwarf simulations often have numerical artifacts
+that result from the sharp surface between the white dwarf and the ambient material
+in the simulation domain, which is difficult to resolve. A somewhat common issue, especially
+at low resolution, is negative densities that occur as a result of a hydro update.
+The \castro\ code contains a prescription for dealing with negative densities, that
+resets the negative density to be similar to nearby zones. Various choices exist
+for how to do this, such as resetting it to the original zone density before the
+update or resetting it to some linear combination of the density of nearby zones.
+The reset is problematic because the strategy is not unique and no choice is
+clearly better than the rest in all cases. Additionally, it is not specified at all
+how to reset momenta in such a case. Consequently, we desired to improve the situation
+by limiting fluxes such that negative densities could not occur, so that such a
+reset would in practice always be avoided. Our solution implements the positivity-preserving
+method of \cite{hu:2013}. A hydrodynamical update to a zone can be broken down into
+an update over every face of the zone where a flux crosses the face over the timestep.
+The central insight of the positivity-preserving method is that if the update over every
+face is positivity-preserving, then the total update must be positivity-preserving as well.
+To guarantee positivity preservation at the zone edge ${\rm i}+1/2$, the flux
+$\mathbf{F}^{n+1/2}_{{\rm i}+1/2}$ at that face is modified to become:
+\begin{equation}
+  \mathbf{F}^{n+1/2}_{{\rm i}+1/2} \rightarrow \theta_{{\rm i}+1/2} \mathbf{F}^{n+1/2}_{{\rm i}+1/2} + (1 - \theta_{{\rm i}+1/2}) \mathbf{F}^{LF}_{{\rm i}+1/2}, \label{eq:limited_flux}
+\end{equation}
+where $0 \leq \theta_{{\rm i}+1/2} \leq 1$ is a scalar, and $\mathbf{F}^{LF}_{{\rm i}+1/2}$ is the Lax-Friedrichs flux,
+\begin{equation}
+  \mathbf{F}^{LF}_{{\rm i}+1/2} = \frac{1}{2}\left[\mathbf{F}^{n}_{{\rm i}} + \mathbf{F}^{n}_{{\rm i}+1} + \text{CFL}\frac{\Delta x}{\Delta t} \frac{1}{\alpha}\left(\mathbf{U}^{n}_{{\rm i}} - \mathbf{U}^{n}_{{\rm i}+1}\right)\right],
+\end{equation}
+where $0 < \text{CFL} < 1$ is the CFL safety factor (the method is guaranteed to
+preserve positivity as long as $\text{CFL} < 1/2$), and $\alpha$ is a scalar that
+ensures multi-dimensional correctness ($\alpha = 1$ in 1D, $1/2$ in 2D, $1/3$ in 3D).
+$\mathbf{F}_{{\rm i}}$ is the flux of material evaluated at the zone center ${\rm i}$
+using the cell-centered quantities $\mathbf{U}$. The scalar $\theta_{{\rm i}+1/2}$ is
+chosen at every interface by calculating the update that would be obtained from
+\autoref{eq:limited_flux}, setting the density component equal to some small, positive
+value, and solving for the value of $\theta$ at the interface that makes
+the equality hold. In regions where the density is not at risk of going negative,
+$\theta \approx 1$ and the original hydrodynamic update is recovered.
+Further discussion, including a proof of the method, a description of multi-dimensional effects,
+and test verification problems, can be found in \cite{hu:2013}.
+
+\clearpage
+
 \bibliographystyle{../aasjournal}

### 2017-07-09 papers/wdmerger_II/paper.tex
Describe the forced post-step regrid algorithm

@@
 the added resolution will still go to the most-needed places.
-However, a significant drawback to this approach is that it
-does not turn on until after an unstable burning region has
-been generated, so this AMR criterion cannot help in the case
-where a spurious detonation begins in a single timestep on the coarse
-grid, though it will kick in immediately after that timestep
-to resolve the regions where the detonation has occurred.
-A possible remedy to be explored in future work is to add
-refinement pre-emptively in regions where the ratio $\tau_s / \tau_e$
-approaches $f_{s}$ but does not yet exceed it. Another
-possibility for this case would be to track how many levels
-would have been needed to prevent the numerical detonation,
-based on the stability criterion, and reject the timestep and
-retry it from the starting state using that many levels of refinement.
-As mentioned in \autoref{sec:hydrocoupling}, the ability to reject
-a timestep and retry it is functionality that
-we have recently added to \castro, so we may try this in future work.
+
+A potential limitation of this approach is that it does not turn on
+until \textit{after} an unstable burning region has been generated,
+because we generally only perform a regridding and level generation
+step at times when AMR levels are synchronized, such as at the beginning
+of the timestep on the coarsest level. So this AMR criterion cannot
+help in the case where a spurious detonation begins in a single timestep
+on the coarse grid, though it will kick in immediately after that
+timestep to resolve the regions where the detonation has occurred.
+This situation is common in our white dwarf collision simulations.
+We have developed a remedy to this issue. At the \textit{end} of every step
+on every level, we check whether the stability criterion has been violated.
+If so, we recognize that it is not too late to solve the problem -- we can
+still add levels of refinement in those places. We then perform the new level
+generation to cover specifically those places that violated the stability criterion.
+To ensure consistency with our algorithm, we use the \textit{old}-time values
+of the coarse fluid state to interpolate into the new fine level, even
+though it is the \textit{new}-time values that were used to guide the level
+generation. Then our algorithm evolves the timestep on the new level as normal.
+Because our algorithm is recursive, this can continue to generate
+as many new levels as are needed to satisfy the stability criterion.
+The upshot is that as long as the simulation can afford enough levels to
+satisfy the stability criterion, there never needs to be even a single timestep
+that violates the criterion.
 

### 2017-07-11 papers/wdmerger_II/paper.tex
Add comments on suppressed burn variants

@@
 conditions; instead, only QSE conditions are reached near the center of the collision, leaving a significant
-amount of silicon and sulfur on the domain that is not further processed.
+amount of silicon and sulfur on the domain that is not further processed. In order to ensure that the
+very different results we obtained were not a result of the specific choice we made for how to suppress
+the burn, we tried varying the approach discussed in \autoref{sec:burner}. First, we limited based on
+the current value of the internal energy and sound speed rather than the fixed, initial value during a burn
+(this should be a less severe limiter). This did not yield a significant difference in the outcome. Second,
+we also tried applying the limiting after the burn had completed, using the differences between the input
+and output states of the integration, rather than directly limiting the RHS in the ODE. The behavior of the
+burn was qualitatively different, in that nickel production began much earlier in the collision when the limiting
+was applied this way. Nevertheless, the qualitative final outcome of the collision -- only trace amounts
+of nickel produced in total -- was still unchanged. With this variant too, it did not matter whether the limiting
+factor was determined based on the initial value of the state or the final value of the state.
 

### 2017-07-12 papers/wdmerger_II/paper.tex
Add some commentary on resolution dependence of suppressed burn

@@
 factor was determined based on the initial value of the state or the final value of the state.
+Our default 400 km resolution is significantly lower than that of \citeauthor{kushnir:2013} (the text
+says they used resolutions up to a few km) but we have tried runs at 100 km and 50 km resolution and
+the results are qualitatively similar. We did not directly compare the results of the method at
+the same resolution as theirs, but our results suggest at least that this limiter significantly
+modifies the evolution for modest resolution simulations and may be inappropriate as a method for enabling
+simulations at those resolutions to obtain accurate results.
 

### 2017-09-09 papers/wdmerger_II/paper.tex
Add note on threshold for timestep limiter for X
Clarify the understanding of WD collision rates

@@
 merger resembles a head-on collision of the binary WDs. Whether the rate of these collisions
-may be high enough to significantly contribute to the observed SN Ia rate is uncertain,
-though \cite{katzdong:2012} suggest that this is plausible.
+may be high enough to significantly contribute to the observed SN Ia rate is debated:
+\cite{katzdong:2012} suggest that this is plausible, but \cite{hamers:2013} and
+\cite{papish:2015} suggest that the rate is relatively low. Even if the rate is low,
+if at least some SNe Ia are caused by WD collisions then they are worth understanding.
 
@@
 timestep to timestep. Similarly, given an estimate for $\dot{X_k}$, the factor $f_{X}$ 
-determines the maximum change in the mass fraction of any species. By making 
+determines the maximum change in the mass fraction of any species. To prevent species
+whose abundances are small in absolute terms from controlling the timestep, we apply
+the limiter based on $f_{X}$ only to species with abundances in a given zone greater
+than $10^{-3}$ (this abundance threshold can be changed by the user of the code). By making 
 $f_{e}$ and $f_{X}$ smaller, we can control the magnitude of the decoupling 

### 2017-09-09 papers/wdmerger_II/paper.tex
Fix the comment about whether we use retries
Update the description of the ODE integrators

@@
 In our \microphysics\ repository we provide several software options for
-solving a set of coupled stiff ODEs. For this work we use an implementation of
-the well known variable-order Richardson extrapolation method presented by \citet{stoer:1980},
-that is similar to the integrator which ships with
-the original versions of the networks mentioned above. Previous work using the \amrex\
-codes typically used the classic \vode\ integrator \citep{vode}. We do maintain a version
-of the software that is compatible with our software interfaces in the \microphysics\
-repository, but we have largely shifted to integrators which are written in modern
-Fortran as a consequence of our efforts to run our codes on GPUs. The Stoer and Bulirsch
-integrator we provide satisfies this criterion; we also provide a rewrite of the VODE
-BDF algorithm that uses modern Fortran features. The Stoer and Bulirsch algorithm
-relies on a uniform relative error tolerance for all of the ODEs in the system, which
-we set at $10^{-6}$ for all simulations described here.
+solving a set of coupled stiff ODEs. For this work our primary integrator
+is the classic \vode\ integrator \citep{vode}, of which we have a version
+compatible with our software interfaces in the \microphysics\ repository.
+We also employ an implementation of the well known variable-order Richardson
+extrapolation method presented by \citet{stoer:1980}, that is similar to the
+integrator which ships with the original versions of the networks mentioned
+above. Our experience shows that relying on a single ODE integrator for the
+variety of stiff ODEs that are present in a collision is insufficient, as we
+sometimes run into integration failures. A combination of the two integrators
+is usually sufficient, however: we first try \vode, and if that fails, we then
+try the Stoer and Bulirsch method. Since it is possible for neither method to
+succeed, we have one more backup scheme. Our specified error tolerances are
+$10^{-6}$ for the energy and temperature ODEs, and $10^{-8}$ for the
+nuclear abundance ODEs (we discuss the sensitivity to this choice in
+\autoref{sec:parameters:tolerance}). We can loosen these tolerances in an
+attempt to give the integrator an easier time. If both \vode\ and the Stoer
+and Bulirsch algorithm fail, then we loosen the tolerances by a multiplicative
+factor of 25\%, and then repeat the attempts with \vode\ and Stoer and Bulirsch.
+We keep doing this until either we have loosened the tolerances enough for one
+of the two methods to succeed, or until we have loosened the tolerances by two
+orders of magnitude, at which point we finally give up. We have found that this
+scheme is resilient enough to handle all of the simulations shown below.
 
@@
 However we have found that the benefits for this problem are
-small and thus we do not use it for the simulations here.
+small.
 
@@
 inexpensive.
+
+\subsection{Integration Error Tolerance}
+\label{sec:parameters:tolerance}
+
+%TODO: add discussion of how ODE tolerances affect the answer
 

### 2017-09-09 papers/wdmerger_II/paper.tex
Update discussion of iso7

@@
 
-\isoseven\ also has the peculiar property that after the detonation had passed through the WDs,
-the region at the central contact point did not fully stop burning, yielding a quasistatic energy
-release that added a couple of million extra timesteps to the simulation (in addition to the few
-hundred it normally takes) before we manually terminated the run. The full effect of these additional
-timesteps only amounted to a $10^{-4}$ relative change in the nickel production. By examining the behavior
-of zones near the collision point, we discovered the reason for this. The zones in question have a composition that
+\isoseven\ also has the peculiar property that after the detonation has passed through the WDs,
+the region at the central contact point does not fully stop burning, yielding a quasistatic energy
+release that can add thousands to millions of extra timesteps to the simulation (in addition to the few
+hundred it normally takes). The full effect of these additional timesteps only amounts to a trace change
+in the nickel production. By examining the behavior of zones near the collision point, we discovered the
+reason for this. The zones in question have a composition that
 is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in

### 2017-09-09 papers/wdmerger_II/paper.tex
Update description of T floor effect

@@
 not the reactions.) We also tested whether our choice for the temperature floor mattered
-by lowering it to $10^6\ \text{K}$ and $10^5\ \text{K}$, and this had no effect on nickel
-production. It does allow parts of the WDs to become colder as they move, due to numerical
-errors in the advecting flow (the Helmholtz equation of state is very insensitive to
-temperature, so even small changes in the internal energy can create very large changes
-in the implied temperature), but the amount of energy injected by the collision is so
-large that the final result is insensitive to whatever the initial temperature of the
+by lowering it to $10^6\ \text{K}$ and $10^5\ \text{K}$; this had a minor effect on nickel
+production, again on the order of a few percent. It does allow parts of the WDs to
+become colder as they move, due to numerical errors in the advecting flow (the Helmholtz
+equation of state is very insensitive to temperature, so even small changes in the internal
+energy can create very large changes in the implied temperature), but the amount of energy
+injected by the collision is so large that the final result is largely insensitive to whatever
+the initial temperature of the
 WDs were. We also checked whether there was any consequence of our choice to disable the

### 2017-09-09 papers/wdmerger_II/paper.tex
Re-title paper II, rework the abstract

@@
 %==========================================================================
-\title{White Dwarf Mergers on Adaptive Meshes\\ II. Collisions}
-
-\shorttitle{WD Mergers. II. Collisions}
+\title{White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions}
+
+\shorttitle{WD Mergers. II. Nuclear Reactions}
 \shortauthors{Katz et al. (2017)}
@@
 \begin{abstract}
-Collisions of white dwarfs are a subset of white dwarf mergers that occur
-when the stars approach each other nearly head-on. These have the potential
-for explosive nucleosynthesis, leading to a thermonuclear detonation that
-disrupts both stars, possibly causing a Type Ia supernova. Simulation of
-these events is challenging due to the rapid evolution of the nuclear burning,
-posing numerical challenges related to both spatial and temporal resolution.
-In this study we explore a number of the algorithms that affect the outcome
-of reacting hydrodynamics simulations that lead to explosive burning and
-address the question of what needs to be done to ensure that these simulations
-are properly converged. We find that achieving a converged simulation in the
-code \castro\ is non-trivial and requires demanding computational resources.
-We also discuss the quantities obtained from our simulations that bear relevance
-to observations of these events, particularly production of nickel-group
-elements and gravitational wave emission during the collision.
+Simulations of white dwarf mergers and collisions that attempt to answer the
+question of whether these events are progenitors of Type Ia supernovae must
+confront the question of how nuclear detonations form and whether observed
+detonations in a numerical simulation are realistic. We address this question
+by performing simulations of collisions of white dwarfs, a subset of white
+dwarf mergers that occur when the stars approach each other nearly head-on.
+These have the potential for explosive nucleosynthesis, leading to a thermonuclear
+detonation that disrupts both stars, possibly causing a Type Ia supernova.
+While this story is intuitively plausible, demonstrating this in a reacting
+hydrodynamics simulation is challenging due to the rapid evolution of the
+nuclear burning, posing numerical challenges related to both spatial and
+temporal resolution. In this study we explore a number of the algorithms
+that affect the outcome of reacting hydrodynamics simulations that lead
+to explosive burning and address the question of what can be done to ensure
+that these simulations are properly converged. We find that achieving a
+converged simulation in the code \castro\ is non-trivial, and it is not
+clear that we can correctly simulate a white dwarf collision and obtain
+an adequate estimate of the nucleosynthetic yield. This result calls into
+question whether it is possible to self-consistently simulate a white dwarf
+merger leading to a thermonuclear detonation.
 \end{abstract}

### 2017-09-09 papers/wdmerger_II/paper.tex
Rework the intro using our new theme

@@
 
-Binary white dwarf (WD) systems have long been thought to be potential progenitors
-of Type Ia supernovae (SNe Ia) \citep{ibentutukov:1984,webbink:1984}. Recently there
-has been a growing amount of observational evidence that double WD systems play
-an important role in explaining at least some SNe Ia \citep{maoz:2014}. These systems
-may be isolated or they may be in hierarchical triple systems, where a WD binary
-system in a tight inner orbit is gravitationally coupled to a third, outer star.
-In either case the binary WD pair can merge due to the emissions of gravitational waves,
-but in the hierarchical triple case, dynamical interactions between the WD binary and
-the outer star can prompt mergers as well \citep{thompson:2011,hamers:2013}. The mergers
-occur due to perturbations in the eccentricity of the binary orbit caused by the outer
-star, which significantly increase the gravitational wave emission of the system. In a
-subset of these mergers the eccentricity may be driven to such a high value that the
-merger resembles a head-on collision of the binary WDs. Whether the rate of these collisions
-may be high enough to significantly contribute to the observed SN Ia rate is debated:
-\cite{katzdong:2012} suggest that this is plausible, but \cite{hamers:2013} and
-\cite{papish:2015} suggest that the rate is relatively low. Even if the rate is low,
-if at least some SNe Ia are caused by WD collisions then they are worth understanding.
-
-Over the past few years a number of groups have performed simulations of WD collisions
-to understand whether they may yield astrophysical transients that look like SNe Ia
-\citep{rosswog:2009,raskin:2010,loren-aguilar:2010,hawley:2012,garcia-senz:2013,
-kushnir:2013,papish:2015,holcomb:2015}. Head-on collisions rapidly convert a significant
-amount of kinetic energy into thermal energy in a small region and thus set up the
-conditions for a thermonuclear detonation, and these previous simulations have indeed
+Binary white dwarf (WD) systems are a promising progenitor candidate for Type
+Ia supernovae (SNe Ia) for observational reasons, but from the simulation
+perspective there is still significant uncertainty regarding how nuclear
+detonations form in an exploding white dwarf. There is uncertainty in the
+model, regarding whether detonations form on the surface of a star as
+accreting material heats up when it strikes the surface, or at or near the
+center of the star as the accreted material compresses the WD. There is also
+uncertainty in the numerics: given one of these models, can we self-consistently
+demonstrate that a detonation occurs? This question has haunted both single
+WD and double WD progenitor models, because the length and time scale at which
+a detonation forms is orders of magnitude smaller than the resolution that
+typical hydrodynamic simulations can achieve. Furthemore, the approximate equations
+we use at the resolution that we can achieve are still notoriously
+difficult to solve accurately. The mere presence of a detonation
+(or lack thereof) in a simulation is therefore only weak evidence regarding
+whether a detonation would truly occur.
+
+In this study we examine the challenges associated with simulating nuclear detonations.
+This paper follows from our first WD merger methodology paper, \citet{wdmergerI}
+(hereafter, Paper I), where we deferred discussion of our nuclear reaction methodology.
+After presenting the burning methodology, we apply it to the problem of collisions of white dwarfs.
+WD collisions can occur in isolated binaries or in hierarchical triple systems, where the WD binary
+system in a tight inner orbit is gravitationally coupled to a third, outer star \citep{thompson:2011,hamers:2013}.
+Whether the rate of these collisions may be high enough to significantly contribute to
+the observed SN Ia rate is debated: \cite{katzdong:2012} suggest that this is plausible,
+but \cite{hamers:2013} and \cite{papish:2015} suggest that the rate is relatively low.
+Regardless, they are a useful vehicle for understanding the problem of nuclear burning.
+Head-on WD collisions rapidly convert a significant amount of kinetic energy into thermal
+energy in a small region and thus set up conditions ripe for a thermonuclear detonation;
+being able to correctly simulate these is in some sense a prerequisite to correct simulation
+of the more complicated general case of white dwarf mergers, where we do not initially know
+under what conditions the seed of a thermonuclear detonation is born.
+
+There have been a number of previous studies which provide useful insight into setting up
+the problem and the burning methodology \citep{rosswog:2009,raskin:2010,loren-aguilar:2010,
+hawley:2012,garcia-senz:2013,kushnir:2013,papish:2015,holcomb:2015}. These studies
 found that detonations occur and convert a large amount of carbon/oxygen material into
@@
 the nuclear network and in the evolution of the temperature), and the temporal and
-spatial resolution. Consequently they yielded estimates of production of nickel-group
-elements that varied significantly, leaving much uncertainty about how such an event
-would appear observationally and whether it bears any resemblance to a SN Ia.
-See Table 4 of \cite{garcia-senz:2013} for a summary of the outcome of many of these studies.
-
-In this paper we seek to add another view to the problem using the reactive hydrodynamics
-code \castro\ \citep{castro} and its underlying adaptive mesh refinement library \amrex\footnote{\amrex is the successor to \boxlib.}.
-This paper is a follow-up to \citet{wdmergerI} (hereafter, Paper I), which established
-the hydrodynamic methdology we are using to simulate mergers of white dwarfs.
-Paper I left out a discussion of the methodology we are using for nuclear reactions,
-and we fill out this information below in \autoref{sec:numericalmethods}. While it is
-our aim is to add information to the discussion about collisions of white dwarfs,
-we also view this work as a methodology study that asks to what extent we
-can trust the reliability of explosive thermonuclear reactions in Eulerian
-hydrodynamics simulations. As a corollary, to what extent does static and/or
-adaptive mesh refinement resolve the types of problems that occur in low resolution
-simulations? We discuss this in Section \autoref{sec:resolution}, though first
-we examine the response of the code to a number of user-specified parameters in
-\autoref{sec:parameters}. We then close with our thoughts on what our work implies
-for the broader astrophysical reacting hydrodynamics literature in \autoref{sec:conclusion}.
+spatial resolution. Consequently the detonations they formed varied in location and nature,
+and the resultant estimates of production of nickel-group elements varied significantly,
+leaving much uncertainty about how such an event would appear observationally and whether
+it bears any resemblance to a SN Ia. See Table 4 of \cite{garcia-senz:2013} for a summary
+of the outcome of many of these studies. In this paper we make the case that
+while we can identify the main problems prevent accurate simulation of the collision
+event, there is no clear path to overcoming them using the computational resources
+typically available for contemporary hydrodynamics simulations.
 
@@
 
-The hydrodynamical equations solved in this paper were presented in \citet{wdmergerI} (hereafter, Paper I).
+The hydrodynamical equations solved in this paper were presented in \citet{wdmergerI}
+(hereafter, Paper I) for the reacting hydrodynamics code \castro\ \citep{castro}.
 Revisions to the core algorithm since then are presented in \autoref{app:hydro}.
@@
 
-We implement the white dwarf collision problem in \castro\ using our \wdmerger\
-package. We load the grid in the way most previous papers have done so:
+We implement the white dwarf collision problem in \castro\ in the \wdmerger\ problem
+directory. We prepare the domain in the way most previous papers have done so:
 the white dwarf centers of mass are initially separated by a distance of four times

### 2017-09-09 papers/wdmerger_II/paper.tex
Remove discussion of the dual-energy parameter
This has no serious effect on the outcome, so it ends up being a distraction.

@@
 tracking of photodisintegration into ${}^{54}$Fe. This network was used by
-\citet{kushnir:2013}, \citet{kushnir:2014}, \citet{rosswog:2009} for their
+\citet{kushnir:2013}, \citet{kushnir:2014}, and \citet{rosswog:2009} for their
 calculations with FLASH, and \citet{papish:2015}. Finally we will also
@@
 
-One other point worth noting for the coupling of the reactions
-and the hydrodynamics relates to the dual-energy formalism used
-by \castro\ and many other hydrodynamics codes \cite{bryan:1995}.
-\castro\ evolves separately equations for the total energy $(\rho E)$ and
-for the internal energy $(\rho e)$. The former is conservative while the latter
-is not, so for accurate hydrodynamics we prefer to use the total
-energy when possible and to use the evolved internal energy variable
-only in situations where the kinetic energy dominates the contribution
-to the total energy. Indeed, for the cosmological purpose for which
-this was originally developed, \cite{ENZO} choose a value $\eta_2 = 0.1$
-so that $e$ is only updated to be equal to $E - \mathbf{v}^2/2$ if
-$e > \eta_2\, E$. In other cases the evolved value of the internal
-energy variable is preserved. However, this choice is somewhat unsafe
-for our application, because the ultimate cause of the nuclear
-burning in a white dwarf collision is the rapid conversion of
-kinetic energy from the white dwarf bulk motion to thermal energy
-as the white dwarfs slam into each other. Keeping $\eta_2$ large
-prevents this from happening, and consequently the temperature
-will never reach values high enough to generate significant
-amounts of nickel. For this paper we choose $\eta_2 = 10^{-4}$,
-which is low enough for the correct conversion of kinetic energy
-but not so low that we need to be concerned about roundoff issues
-caused by subtracting kinetic from total energy. This is also
-the value that was used by \cite{hawley:2012}.
-
 \subsection{Numerically Unstable Burning}
@@
 value (any zone that reaches below this temperature will be reset).
-
-\castro\ uses the dual energy formalism of \cite{bryan:1995,ENZO}, where we evolve equations
-for both the total energy $E$ and the internal energy $e$, using one or the other to
-determine the pressure in the hydrodynamics step (if we are using the total energy, we
-compute the corresponding internal energy by subtracting the kinetic energy $K$). The switching
-parameter is $\eta_1 = 10^{-3}$, and if $(E - K) / E < \eta_1$, we use the internal energy variable $e$
-in the equation of state call that obtains the pressure. Otherwise we use $(E - K)$. The
-idea behind this is that in high Mach number flows where the kinetic energy dominates the
-total energy, we do not want to risk generating a negative estimate of the internal energy
-due to roundoff error. Similarly, at regular intervals when we enforce consistency of the
-internal energy variable (which occurs a few times per timestep), we sync the internal energy
-variable so that it is equal to $E - K$, skipping this step only if $(E - K) / E < \eta_2$.
-We adopt by default $\eta_2 = 10^{-4}$. This choice can matter much more for reacting
-hydrodynamics than non-reacting hydrodynamics. For example, a collision can only lead to a
-detonation if there is sufficient conversion of kinetic to thermal energy. Obviously
-this happens physically, but if the parameter $\eta_2$ is set high enough then it could
-in principle interfere by not allowing the simulation to efficiently convert kinetic energy
-to thermal energy. (This does not end up mattering much for our simulations because the
-kinetic energy does not strongly dominate the internal energy for colliding WDs.) And on
-a similar note, when using this formalism it is ambiguous which variable to use ($(E - K)$
-or $e$) when calculating the initial temperature that goes into the reaction step. We
-have introduced a new parameter $\eta_3$ that uses the same logic as the first two: if
-$(E - K) / E < \eta_3$, we use $(E - K)$ in the equation of state call that obtains the
-temperature; otherwise, we use $e$. However, unlike the other two, the \castro\ default
-is $\eta_3 = 1.0$, meaning that we normally use the internal energy variable.
 
@@
 We tested a few other parameters that ended up having no serious effect on the outcome.
-The dual energy parameter $\eta_2$ described in \autoref{sec:problemsetup} can have a
-minor effect on the nickel production, yielding variations on the order of a few percent,
-but there is no clear pattern. The dual energy parameter $\eta_3$ has no effect on the
-outcome: the answer is the same for $\eta_3 = 0$ and $\eta_3 = 1$, which makes sense
-because all of the reacting regions have a low enough kinetic energy relative to their
-total energy such that the internal energy variable is always kept in sync with $(E - K)$.
-(We did not look at $\eta_1$ because that has a direct effect only on the hydrodynamics,
-not the reactions.) We also tested whether our choice for the temperature floor mattered
+The dual energy hydrodynamic formalism (discussed in Paper I) has parameters that determine
+whether we obtain the internal energy from the total energy or from a separately advected
+internal energy variable. This is usually employed when calculating the pressure during
+a hydrodynamics update, and we have also added to \castro\ a similar switch when calculating
+the initial temperature for a burn. We find that the value of the dual energy parameter has
+only a minor effect on the nickel production, yielding variations on the order of a few percent
+with no systematic trend. We also tested whether our choice for the temperature floor mattered
 by lowering it to $10^6\ \text{K}$ and $10^5\ \text{K}$; this had a minor effect on nickel

### 2017-09-09 papers/wdmerger_II/paper.tex
Remove gravitational wave signature, this is irrelevant

@@
 %==========================================================================
-% Gravitational Wave Signature
-%==========================================================================
-\section{Gravitational Wave Signature}
-\label{sec:gwsignature}
-
-Paper I established our methodology for calculating the gravitational wave signature
-of a white dwarf merger using the post-Newtonian framework of \citet{blanchet:1990}.
-While a typical merger event has a characteristic profile including a chirp and ringdown,
-white dwarf collisions are noticeably distinct in their gravitational wave signatures, and
-hence may be interesting targets for observation for future low frequency gravitational
-wave detectors such as LISA \citep{eLISA}. The strain increases monotonically for a collision until impact,
-instead of oscillating back and forth at a characteristic frequency as for mergers.
-The main caveat for these events compared to events
-such as neutron star mergers is that due to the event being inherently less energetic,
-the typical distance scale on which we could observe a white dwarf merger or collision
-is the size of our own galaxy, with extragalactic events being unlikely to be
-detected, at least with the first generation of space-based gravitational wave
-detectors.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.425]{plots/gw_signal_2D}
-  \caption{Gravitational wave strains for the head-on collision of two $0.64\ \msolar$
-           white dwarfs. The strain is normalized to a source distance of 10 kpc.
-           \label{fig:gw_signal_2D}}
-\end{figure}
-
-\autoref{fig:gw_signal_2D} is a plot of the gravitational wave strain for
-the representative case of \autoref{sec:parameters} ($M_P = 0.64\ \msolar$, $M_S = 0.64\ \msolar$, $b = 0$),
-where we disabled the normal stopping criterion and let the simulation run until $t = 20$.
-We use the convention that the absolute scale of the strain is normalized to a distance
-of 10 kpc (the strain scales inversely with the source distance). We show the $+$ and $\times$
-polarizations for observers situated along the $x$, $y$, and $z$ axes (for the 2D case,
-this corresponds to the $z$, $R$, and $\phi$ axes, respectively, in the simulation).
-The shape of the profile agrees qualitatively with the gravitational wave signatures
-predicted by \cite{loren-aguilar:2009:collisions} and \cite{garcia-senz:2013}.
-\autoref{fig:gw_signal_3D} shows the strain prediction for a 3D collision event
-with $M_P = 0.8\ \msolar$, $M_S = 0.6\ \msolar$, and $b = 0.8$. This mass combination
-was chosen because \cite{loren-aguilar:2009:collisions} used the same combination for
-gravitational wave signal prediction, but we cannot replicate the large impact parameter
-they used on our simulation grid, so the detonation looks quite different than in their case.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.425]{plots/gw_signal_3D}
-  \caption{Gravitational wave strains for the off-center collision ($b = 0.8$) of
-           $0.8\ \msolar$ and $0.6\ \msolar$ WDs. The strain is normalized to a
-           source distance of 10 kpc.
-           \label{fig:gw_signal_3D}}
-\end{figure}
-
-%==========================================================================
 % Conclusions

### 2017-09-09 papers/wdmerger_II/paper.tex
Narrow focus to a single 2D 0.64+0.64 C/O WD problem

@@
 two point-masses in free-fall towards each other coming in from infinity, such that
-the contact point is at the origin, and they approach each other along the $x$-axis.
-The composition of the WDs is as described in Paper I: low-mass WDs are helium-dominated,
-intermediate-mass WDs are carbon and oxygen dominated, and high-mass WDs contain
-oxygen, neon, and magnesium. By default the WDs approach each other head-on, with
-their respective centers lying on the $x$-axis, but we have also added an option
-to have the WDs approach each other at a non-zero impact parameter $b$, where the
-offset is perpendicular to the $x$-axis in the $y$ direction. We measure
-$b$ in units of the secondary WD's radius, so that the default head-on case
-corresponds to $b = 0$ and the case where the two WDs just graze past each other
-corresponds to $b = 1$ (in practice it happens slightly differently because
-of the WDs responding to each other's gravitational fields).
+the contact point is at the origin, and they approach each other along a coordinate axis.
+In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
+composed of equal parts carbon and oxygen by mass, for comparison to earlier studies;
+however, the code does have the ability to vary the impact parameter of the collision
+and the composition of the WDs.
 
-For all the 3D simulations to follow, we use the same coarse grid as in Paper I:
-there are 256 zones per dimension, and the width of each axis is $1.024 \times 10^{10}$
-cm, yielding a base resolution of 400 km. When we use adaptive-mesh refinement, we will
-use the refinement scheme described in \autoref{sec:unstable_burning}. This is
-acceptable because the outcome of the collision problem is chiefly a function
-of the propagation of burning front through the WDs, so for this problem one
-can take some liberties in other parts of the algorithm for the purpose of
-computational efficiency, which can always be turned back on later if a more
-accurate answer is desired. So, for example, we do not need to add refinement
-everywhere in the WDs in the interest of getting the collision dynamics to be
-slightly more accurate. Additionally, we choose to solve the Poisson equation
-for the self-gravity of the system only on the coarse grid. The gravitational
-potential and acceleration on the finer levels are interpolated from the coarse
-grid. This grants enhanced computational speed without a serious effect on the
-accuracy of the simulation (for the case where the fine grids cover the stars,
-we have checked that the more accurate gravitational forcing would only modify
-slightly the time to impact and the subsequent detonation process).
-
-We have also added to \wdmerger\ the ability to use the 2D cylindrical ($R-z$)
-coordinate system evolution in \castro. In this coordinate system, we align
-the WDs along the $z$-axis (which is analogous to our $x$-axis in the Cartesian
-evolution) at the usual distance, with the center of the WDs at $R = 0$. The
-un-simulated $\phi$ dimension then would extend the WDs through a $2\pi$ revolution.
-The nature of the axisymmetry inherent to the cylindrical coordinate system
-means that we can only run head-on collisions with $b = 0$. The coarse grid
-is the same resolution as the 3D case: the width of the $z$ axis is $1.024 \times 10^{10}$
-cm and the width of the $R$ axis is $5.012 \times 10^{9}$ cm, with twice as
-many zones along the $z$ axis as the $R$ axis so that the equal 400 km
-resolution is maintained.
+We have added to the \wdmerger\ problem setup the ability to take advantage of 
+the 2D cylindrical ($R-z$) coordinate system evolution in \castro, and we use it
+in this study, because the axisymmetry inherent to the cylindrical coordinate system
+lends itself well to the head-on collision problem. In this coordinate system, we
+align the WDs along the $z$-axis (which is analogous to the $x$-axis in Cartesian
+evolution), with the center of the WDs at $R = 0$. The un-simulated $\phi$ dimension
+then would extend the WDs through a $2\pi$ revolution. The domain has width $4 \times 10^{9}$
+cm along the $z$ axis, and is half as wide along the $R$ axis. The resolution is equal
+in both dimensions, so that there are twice as many zones in the $z$ dimension in
+a uniform grid. The default grid has 256 zones along the $z$ axis, yielding a resolution
+of 156.25 km.
 
@@
 material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition is the
-same as that of the WDs (which is equal carbon and oxygen by mass, unless
-stated otherwise). Reactions in this region are unimportant, so for
+same as that of the WDs. Reactions in this region are unimportant, so for
 computational efficiency, we disable all nuclear burning for zones that have
@@
 factor was determined based on the initial value of the state or the final value of the state.
-Our default 400 km resolution is significantly lower than that of \citeauthor{kushnir:2013} (the text
-says they used resolutions up to a few km) but we have tried runs at 100 km and 50 km resolution and
+Our default resolution is significantly lower than that of \citeauthor{kushnir:2013} (the text
+says they used resolutions up to a few km) but we have tried runs at higher resolution and
 the results are qualitatively similar. We did not directly compare the results of the method at
@@
 simulations at those resolutions to obtain accurate results.
-
-\subsection{Impact Parameter}
-\label{sec:parameters:impactparameter}
-
-As the impact parameter varies for a head-on collision, the cross-section of the contact
-zone between the two stars varies in size and geometry. In \autoref{fig:impact_parameter}, we
-see the dependence of the nickel production on the impact parameter $b$, measured in units
-of the WD radius. This test was performed using the 3D setup at our usual base resolution of
-400 km. For computational efficiency, timestep limiting based on burning was disabled.
-We sampled at $b = 0.0$, $0.01$, $0.02$, $0.03$, $0.04$, $0.05$, and then increasing
-in increments of $0.05$. Relative to $b = 0$, there is a slight increase in nickel production for small
-but non-zero $b$, which then declines with increasing $b$ except for a bump at $b = 0.5$.
-For $b \ge 0.7$, no detonation forms at all, and the merged object remains a stable C/O WD
-with $M = 1.2\ \msolar$ at least until $t = 100$.
-
-It is worth nothing that for $b = 0.0$ the nickel production is a bit smaller for the 3D simulation
-than its equivalent 2D simulation, the default case described above. There are differences both in
-the size of the timestep (which as we saw in \autoref{sec:parameters:timestepping} plays a role in
-the total nickel production) and in the discretized nature of the stars; the assumption of perfect
-azimuthal symmetry inherent to the 2D simulation breaks down for the 3D simulation, especially at
-low resolution, since the zones are rectangular.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.425]{plots/impact_parameter}
-  \caption[Nickel production dependence on the impact parameter]
-          {$^{56}$Ni generation as a function of the initial WD impact parameter for
-           a $0.64\ \msolar$ + $0.64\ \msolar$ WD collision.
-           \label{fig:impact_parameter}}
-\end{figure}
 
@@
 and varying the minimum $T$ for reactions from $2 \times 10^{7}\ \text{K}$ to
-$4 \times 10^{8}\ \text{K}$ also had no meaningful effect.
+$4 \times 10^{8}\ \text{K}$ also had no meaningful effect. Finally, we also checked
+that the 2D axisymmetric simulation matched the results of an equivalent 3D Cartesian simulation.
 
 %TODO: maybe show an image of the temperature field in the WDs before colliding to demonstrate the point about numerical error?
-
-The choice of the WD composition to be equal carbon and oxygen by mass is a common
-choice in the literature and is broadly comparable to what stellar evolution codes
-yield for WDs of this mass. Still, it is an arbitrary approximation, especially the
-insistence that the composition is uniform throughout the WD. In \autoref{table:co}
-we list the nickel production as a function of the initial carbon mass fraction.
-This can have a few percent impact on the outcome for the plausible values we tested.
-In future work we may look at the effect of a more realistic progenitor structure,
-including the effects of a helium shell on the surface of the WDs \citep{holcomb:2015}
-and a non-uniform interior structure.
-
-\input{plots/co_frac_m_P_0.64_m_S_0.64.tbl}
 

### 2017-09-09 papers/wdmerger_II/paper.tex
Remove the appendix describing the flux limiting
This should eventually be described somewhere, but it does not make sense
given the narrowed focus we have for this paper now.

@@
 (hereafter, Paper I) for the reacting hydrodynamics code \castro\ \citep{castro}.
-Revisions to the core algorithm since then are presented in \autoref{app:hydro}.
+Only minor revisions to the core hydrodynamic algorithm have since occurred.
 In this paper we focus our discussion on our implementation of the nuclear
@@
 
-\appendix
-
-\section{Hydrodynamics Algorithm Updates}
-\label{app:hydro}
-
-The core \castro\ algorithm for hydrodynamic evolution is presented in
-\citet{wdmergerI}, and is largely unchanged in the present work. This appendix
-focuses on minor revisions to the algorithm that have occurred since then.
-
-\subsection{Flux Limiting}
-\label{app:hydro:flux_limiting}
-
-Multi-dimensional hydrodynamic white dwarf simulations often have numerical artifacts
-that result from the sharp surface between the white dwarf and the ambient material
-in the simulation domain, which is difficult to resolve. A somewhat common issue, especially
-at low resolution, is negative densities that occur as a result of a hydro update.
-The \castro\ code contains a prescription for dealing with negative densities, that
-resets the negative density to be similar to nearby zones. Various choices exist
-for how to do this, such as resetting it to the original zone density before the
-update or resetting it to some linear combination of the density of nearby zones.
-The reset is problematic because the strategy is not unique and no choice is
-clearly better than the rest in all cases. Additionally, it is not specified at all
-how to reset momenta in such a case. Consequently, we desired to improve the situation
-by limiting fluxes such that negative densities could not occur, so that such a
-reset would in practice always be avoided. Our solution implements the positivity-preserving
-method of \cite{hu:2013}. A hydrodynamical update to a zone can be broken down into
-an update over every face of the zone where a flux crosses the face over the timestep.
-The central insight of the positivity-preserving method is that if the update over every
-face is positivity-preserving, then the total update must be positivity-preserving as well.
-To guarantee positivity preservation at the zone edge ${\rm i}+1/2$, the flux
-$\mathbf{F}^{n+1/2}_{{\rm i}+1/2}$ at that face is modified to become:
-\begin{equation}
-  \mathbf{F}^{n+1/2}_{{\rm i}+1/2} \rightarrow \theta_{{\rm i}+1/2} \mathbf{F}^{n+1/2}_{{\rm i}+1/2} + (1 - \theta_{{\rm i}+1/2}) \mathbf{F}^{LF}_{{\rm i}+1/2}, \label{eq:limited_flux}
-\end{equation}
-where $0 \leq \theta_{{\rm i}+1/2} \leq 1$ is a scalar, and $\mathbf{F}^{LF}_{{\rm i}+1/2}$ is the Lax-Friedrichs flux,
-\begin{equation}
-  \mathbf{F}^{LF}_{{\rm i}+1/2} = \frac{1}{2}\left[\mathbf{F}^{n}_{{\rm i}} + \mathbf{F}^{n}_{{\rm i}+1} + \text{CFL}\frac{\Delta x}{\Delta t} \frac{1}{\alpha}\left(\mathbf{U}^{n}_{{\rm i}} - \mathbf{U}^{n}_{{\rm i}+1}\right)\right],
-\end{equation}
-where $0 < \text{CFL} < 1$ is the CFL safety factor (the method is guaranteed to
-preserve positivity as long as $\text{CFL} < 1/2$), and $\alpha$ is a scalar that
-ensures multi-dimensional correctness ($\alpha = 1$ in 1D, $1/2$ in 2D, $1/3$ in 3D).
-$\mathbf{F}_{{\rm i}}$ is the flux of material evaluated at the zone center ${\rm i}$
-using the cell-centered quantities $\mathbf{U}$. The scalar $\theta_{{\rm i}+1/2}$ is
-chosen at every interface by calculating the update that would be obtained from
-\autoref{eq:limited_flux}, setting the density component equal to some small, positive
-value, and solving for the value of $\theta$ at the interface that makes
-the equality hold. In regions where the density is not at risk of going negative,
-$\theta \approx 1$ and the original hydrodynamic update is recovered.
-Further discussion, including a proof of the method, a description of multi-dimensional effects,
-and test verification problems, can be found in \cite{hu:2013}.
-
-\clearpage
-
 \bibliographystyle{../aasjournal}

### 2017-09-10 papers/wdmerger_II/paper.tex
Fix the resolution number
Clarify wording regarding iso7

@@
 in both dimensions, so that there are twice as many zones in the $z$ dimension in
-a uniform grid. The default grid has 256 zones along the $z$ axis, yielding a resolution
-of 156.25 km.
+a uniform grid. The default grid has 256 zones along the $z$ axis, yielding a
+spatial resolution of 312.5 km.
 
@@
 release that can add thousands to millions of extra timesteps to the simulation (in addition to the few
-hundred it normally takes). The full effect of these additional timesteps only amounts to a trace change
-in the nickel production. By examining the behavior of zones near the collision point, we discovered the
-reason for this. The zones in question have a composition that
+hundred it normally takes, using our designated stopping criterion). The full effect of these additional
+timesteps only amounts to a trace change in the nickel production. By examining the behavior of zones
+near the collision point, we discovered the reason for this. The zones in question have a composition that
 is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in

### 2017-09-10 papers/wdmerger_II/paper.tex
Add comment on aprox19 and aprox21 failures

@@
 
+\aproxnineteen\ and \aproxtwentyone\ both take significantly longer to evaluate at our default
+ODE tolerances than \aproxthirteen\ does, even after accounting for the increased number of species.
+This is due to a significant number of integration failures and subsequent retries (see the discussion
+in \autoref{sec:burner}). The species tolerance of $10^{-8}$ is challenging for these networks.
+
 \subsection{Burning Mode}

### 2017-09-11 papers/wdmerger_II/paper.tex
Update IACS acknowledgement

@@
 of Science of the U.S. Department of Energy under Contract
-No. DE-AC02-05CH11231.  Results in this paper were obtained using the
-high-performance LIred computing system at the Institute for Advanced
-Computational Science at Stony Brook University, which was obtained
-through the Empire State Development grant NYS \#28451.
+No. DE-AC02-05CH11231. The authors would like to thank Stony Brook
+Research Computing and Cyberinfrastructure, and the Institute for
+Advanced Computational Science at Stony Brook University for access
+to the high-performance LIred and SeaWulf computing systems, the latter
+of which was made possible by a $1.4M National Science Foundation grant (#1531492).
 

### 2017-09-11 papers/wdmerger_II/paper.tex
fix latex-isms

--- 2017-09-10 papers/wdmerger_II/paper.tex
@@
 of Science of the U.S. Department of Energy under Contract
-No. DE-AC02-05CH11231.  Results in this paper were obtained using the
-high-performance LIred computing system at the Institute for Advanced
-Computational Science at Stony Brook University, which was obtained
-through the Empire State Development grant NYS \#28451.
+No. DE-AC02-05CH11231. The authors would like to thank Stony Brook
+Research Computing and Cyberinfrastructure, and the Institute for
+Advanced Computational Science at Stony Brook University for access
+to the high-performance LIred and SeaWulf computing systems, the latter
+of which was made possible by a \$1.4M National Science Foundation grant (\#1531492).
 

### 2017-09-12 papers/wdmerger_II/paper.tex
Update timestepping section with new data

@@
 burning rate. As the timestep gets smaller, the error due to the Strang splitting
-scheme coupling the reactions and the burning also gets smaller, and we expect
+scheme coupling the reactions and the hydro also gets smaller, and we expect
 that the results become more accurate.
@@
 \autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
-production. As either $f$ becomes small, the restriction on the timestep becomes quite
-severe: for $f_{bX} = 0.01$, the minimum timestep in the simulation is smaller than
-$2 \times 10^{-9}\ \text{s}$ and the full simulation requires over 750,000 timesteps.
+production (these limiters are somewhat redundant in the sense that a given $f_{bX}$
+roughly corresponds to another (higher) $f_{be}$, as is evident by the fact that on
+the graph the curves look qualitatively similar in shape but with a horizontal offset.
+As either $f$ becomes small, the restriction on the timestep becomes quite severe:
+for $f_{be} = 0.001$, the minimum timestep in the simulation is smaller than
+$5 \times 10^{-9}\ \text{s}$ and the full simulation requires over 100,000 timesteps.
+Such a simulation would be prohibitively expensive for a high resolution 3D calculation.
 Yet it is clear that even for this steep value of the limiter, which is much stricter
 than the limiter used in, e.g., \cite{hawley:2012} and \cite{raskin:2010}, the
-simulation has not fully converged in nickel production. These limiters are somewhat
-redundant in the sense that a given $f_{bX}$ corresponds to another (higher) $f_{be}$,
-as is evident by the fact that on the graph the curves look similar in shape but with
-a horizontal offset.
+simulation has not fully converged in nickel production.
 
 We also examined the effect of the method used to enforce the limiting, using the
-four limiting modes discussed in \autoref{sec:hydrocoupling}. The results are shown
-in \autoref{table:burninglimitermode} for the case $f_{be} = 0.3$ (we did not use
-limiting with $f_{bX}$ for this test). The two modes that use the full last timestep's
-worth of information to estimate $\dot{e}$ yield a slightly larger nickel mass than the
-two modes that only use more recent information, but the difference is negligible.
-
-\input{plots/burning_limiter_mode_m_P_0.64_m_S_0.64.tbl}
+four limiting modes discussed in \autoref{sec:hydrocoupling}. The two modes that
+use the full last timestep's worth of information to estimate $\dot{e}$ (Mode 3 and Mode 4)
+yield a slightly larger nickel mass than the two modes that only use more recent information,
+but the difference is less than 10\%, and it is not clear which is preferable since
+for the value of the timestep limiter we used to compare these modes ($f_{be} = 0.3$),
+the simulation is not converged anyway.
 
@@
 For all remaining tests in this section, unless otherwise stated, we take $f_{be} = 10^2$
-and disable limiting with $f_{bX}$ by setting it to a large number. This choice is a balance
-between accuracy and efficiency: we avoid the obviously wrong answers that occur for $f_{be} > 10^3$
-while ensuring that the timesteps are long enough that the following tests are computationally
-inexpensive.
+and disable limiting with $f_{bX}$ by setting it to a large number. This choice was made
+somewhat arbitrarily for computational efficiency.
 

### 2017-09-12 science/wdmerger_II/analyze.py
Update the paper II analysis script

@@ -5,23 +5,45 @@ matplotlib.use('agg')
 from matplotlib import pyplot as plt
 import os
 
-linestyles = ['-', '--', '-.', ':']
-markers = ['o', 's', 'D', '^']
-
-
+linestyles = ['-', '--', ':', '-', '--', ':']
+markers = ['', '', '', 'o', 's', 'D']
+colors = ['b', 'g', 'r', 'c', 'm', 'k']
 
 def get_ni56(results_dir):
     """Return a list of maximum 56Ni production from all completed sub-directories in results_dir."""
 
     import wdmerger_spec_diag_analysis as spec_diag
+    import numpy as np
 
     dir_list = wdmerger.get_parameter_list(results_dir)
 
     diag_filename_list = [results_dir + '/' + directory + '/output/species_diag.out' for directory in dir_list]
 
-    ni56_arr = [np.amax(spec_diag.get_column('Mass ni56', diag_filename)) for diag_filename in diag_filename_list]
+    ni56_arr = [np.amax(spec_diag.get_column('Mass ni56', diag_filename)) - np.amin(spec_diag.get_column('Mass ni56', diag_filename))
+                for diag_filename in diag_filename_list]
 
     return ni56_arr
+
+
+
+def get_e_release(results_dir):
+    """Return a list of energy generation from all completed sub-directories in results_dir."""
+
+    import wdmerger_grid_diag_analysis as grid_diag
+    import numpy as np
+
+    dir_list = wdmerger.get_parameter_list(results_dir)
+
+    diag_filename_list = [results_dir + '/' + directory + '/output/grid_diag.out' for directory in dir_list]
+
+    e_arr = [np.amax(grid_diag.get_column('TOTAL ENERGY', diag_filename)) - np.amin(grid_diag.get_column('TOTAL ENERGY', diag_filename))
+             for diag_filename in diag_filename_list]
+
+    # Divide by 10**51
+
+    e_arr = [e / 1.0e51 for e in e_arr]
+
+    return e_arr
 
 
 
@@ -42,8 +64,16 @@ def get_diagfile(results_dir):
 
 
 
-def write_ni56_table(results_dir, out_filename, v_list, n_list, comment, caption, col1title, label):
-    """Write out a LaTeX table of 56Ni mass as a function of a given variable."""
[…]

### 2017-09-12 papers/wdmerger_II/paper.tex
Update the burning mode discussion and data

@@
 In contrast, the suppressed burning mode (modeled after \citet{kushnir:2013}) produces only about
-$5 \times 10^{-4}\ \msolar$ of $^{56}$Ni. In a suppressed burn, we ensure that at every evaluation
+$2 \times 10^{-4}\ \msolar$ of $^{56}$Ni. In a suppressed burn, we ensure that at every evaluation
 of the right-hand-side vector for the nuclear network integration, all quantities are scaled by the
@@
 the results are qualitatively similar. We did not directly compare the results of the method at
-the same resolution as theirs, but our results suggest at least that this limiter significantly
-modifies the evolution for modest resolution simulations and may be inappropriate as a method for enabling
-simulations at those resolutions to obtain accurate results.
+the same resolution as theirs, but our results indicate at least that this limiter significantly
+modifies the evolution for modest resolution simulations. The low amount of nickel generated is,
+interestingly, similar to the results at higher resolution from \autoref{sec:resolution}. However,
+this may be a coincidence rather than, as claimed by \citeauthor{kushnir:2013}, a shortcut to obtaining
+a resolved simulation. In any case, the suppressed burn limiting used here is ultimately unphysical
+and we choose not to use it in any other simulation.
 

### 2017-09-13 papers/wdmerger_II/paper.tex
Use more careful wording regarding the suppressed burn

@@
 the same resolution as theirs, but our results indicate at least that this limiter significantly
-modifies the evolution for modest resolution simulations. The low amount of nickel generated is,
-interestingly, similar to the results at higher resolution from \autoref{sec:resolution}. However,
-this may be a coincidence rather than, as claimed by \citeauthor{kushnir:2013}, a shortcut to obtaining
-a resolved simulation. In any case, the suppressed burn limiting used here is ultimately unphysical
-and we choose not to use it in any other simulation.
+modifies the evolution for modest resolution simulations. The limiter does delay the time at which
+the detonation occurs, as expected by \citeauthor{kushnir:2013}, but the conditions for generating
+significant amounts of iron-group elements do not develop. It is not clear what the relevant difference
+is between our simulation and that of \citeauthor{kushnir:2013}, but as the suppressed burn limiting
+used here is ultimately unphysical, we choose not to use it in any other simulation.
 

### 2017-09-13 papers/wdmerger_II/paper.tex
End the abstract on a more hopeful note

@@
 that these simulations are properly converged. We find that achieving a
-converged simulation in the code \castro\ is non-trivial, and it is not
-clear that we can correctly simulate a white dwarf collision and obtain
-an adequate estimate of the nucleosynthetic yield. This result calls into
-question whether it is possible to self-consistently simulate a white dwarf
-merger leading to a thermonuclear detonation.
+converged simulation in the code \castro\ is very difficult for a standard
+approach with uniform resolution. The typical timesteps taken by hydrodynamics
+codes are too large and the spatial resolution is too low, while increasing
+either to the point required to actually obtain a converged simulation is
+prohibitive. However, we have identified a criterion for adaptive mesh refinement
+that substantially improves the situation: by targeting for extra sptial resolution
+only those regions that are likely to undergo numerically unstable burning,
+we can significantly reduce the effects of numerically spurious detonations.
+This result is a possible path forward for obtaining reliable numerical
+simulations of Type Ia supernova involving thermonuclear detonations.
 \end{abstract}

### 2017-09-14 papers/wdmerger_II/paper.tex
Correct the data on what ODE tolerances we use

@@
 try the Stoer and Bulirsch method. Since it is possible for neither method to
-succeed, we have one more backup scheme. Our specified error tolerances are
-$10^{-6}$ for the energy and temperature ODEs, and $10^{-8}$ for the
-nuclear abundance ODEs (we discuss the sensitivity to this choice in
+succeed, we have one more backup scheme. Our specified (relative and absolute)
+error tolerances are uniformly $10^{-6}$ for the energy, temperature, and nuclear
+abundance ODEs (we discuss the sensitivity to this choice in
 \autoref{sec:parameters:tolerance}). We can loosen these tolerances in an

### 2017-09-19 papers/wdmerger_II/paper.tex
Clean up and simplify discussion of other parameters

@@
 however, the code does have the ability to vary the impact parameter of the collision
-and the composition of the WDs.
+and the mass and composition of the WDs.
 
@@
 
-The simulation is terminated when the total energy is positive, indicating that
-the system has become unbound due to nuclear energy release, and has been
-constant or decreasing for the last five timesteps, indicating that no further
-meaningful nuclear energy generation is occurring and material is now beginning
-to stream off the simulation domain. As all simulations in this paper are of
-WD collisions that generate enough energy to unbind the system, this stopping
-criterion is applied throughout all simulations shown here.
+We terminate the simulation at $t = 10$ seconds, as we have found that for this
+particular setup, all of the significant burning has completed by this time. At
+this point, the total energy on the domain has become positive, typically about
+$10^{51}$ ergs, and has stopped increasing. This means that the system has become
+unbound due to nuclear energy release, and that no further meaningful nuclear energy
+generation is occurring.
 
@@
 computational efficiency, we disable all nuclear burning for zones that have
-$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. The initial ambient
+$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
+the results do not depend on this choice. The initial ambient
 temperature is $10^7\ \text{K}$, which is the same as the initial temperature
 everywhere throughout the WDs, and we set the temperature floor to the same
-value (any zone that reaches below this temperature will be reset).
+value (any zone that reaches below this temperature will be reset). Setting the
+temperature floor to a lower value has only minor effects on the simulation outcome.
 
@@
 
-\subsection{Other Parameters}
-
-We tested a few other parameters that ended up having no serious effect on the outcome.
-The dual energy hydrodynamic formalism (discussed in Paper I) has parameters that determine
-whether we obtain the internal energy from the total energy or from a separately advected
-internal energy variable. This is usually employed when calculating the pressure during
-a hydrodynamics update, and we have also added to \castro\ a similar switch when calculating
-the initial temperature for a burn. We find that the value of the dual energy parameter has
-only a minor effect on the nickel production, yielding variations on the order of a few percent
-with no systematic trend. We also tested whether our choice for the temperature floor mattered
-by lowering it to $10^6\ \text{K}$ and $10^5\ \text{K}$; this had a minor effect on nickel
-production, again on the order of a few percent. It does allow parts of the WDs to
-become colder as they move, due to numerical errors in the advecting flow (the Helmholtz
-equation of state is very insensitive to temperature, so even small changes in the internal
-energy can create very large changes in the implied temperature), but the amount of energy
-injected by the collision is so large that the final result is largely insensitive to whatever
-the initial temperature of the
-WDs were. We also checked whether there was any consequence of our choice to disable the
-burning for small enough $\rho$ and $T$. Lowering the minimum $\rho$ for reactions to
-$10^0\ \text{g cm}^{-3}$ from $10^6\ \text{g cm}^{-3}$ had a negligible effect,
-and varying the minimum $T$ for reactions from $2 \times 10^{7}\ \text{K}$ to
-$4 \times 10^{8}\ \text{K}$ also had no meaningful effect. Finally, we also checked
-that the 2D axisymmetric simulation matched the results of an equivalent 3D Cartesian simulation.
-
-%TODO: maybe show an image of the temperature field in the WDs before colliding to demonstrate the point about numerical error?
-
 %==========================================================================

### 2017-09-19 papers/wdmerger_II/paper.tex
Remove nuclear network tests; these are outside the scope

@@
 will eventually get us to ${}^{56}$Ni, the nuclide responsible for the
-energy output of Type Ia supernovae. In this paper we consider four networks
-to do this, presented in order of increasing complexity. The most simple is
-\isoseven\ \citep{timmes:2000}, which includes all of the aforementioned isotopes and
-${}^{28}$Si (see also \citet{hix:1998}). ${}^{28}$Si effectively measures the
-equilibrium state of silicon-group elements, and ${}^{56}$Ni effectively measures
-the equilibrium state of iron-group elements, with the link between them governed
-by the effective loss or gain of seven $\alpha$-particles. This type of network
-was used by \citet{rosswog:2009} for their collision calculations in SPH.
-
-Next is \aproxthirteen\ \citep{timmes:1999,timmes:2000}. This includes
-all of the isotopes of \isoseven, and all of the $\alpha$-chain particles between
-silicon and nickel (${}^{32}$S, ${}^{36}$Ar, ${}^{40}$Ca, ${}^{44}$Ti, ${}^{48}$Cr,
-and ${}^{52}$Fe). This network was used by \citet{hawley:2012} and \citet{raskin:2010}.
-\citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a very similar network
-that included additionally ${}^{60}$Zn. The \aproxnineteen\ network \citep{timmes:1999}
-builds on \aproxthirteen\ by including isotopes for hydrogen burning and explicit
-tracking of photodisintegration into ${}^{54}$Fe. This network was used by
-\citet{kushnir:2013}, \citet{kushnir:2014}, and \citet{rosswog:2009} for their
-calculations with FLASH, and \citet{papish:2015}. Finally we will also
-consider \aproxtwentyone, which includes all of the above plus ${}^{56}$Cr
-and ${}^{56}$Fe and related reaction links. The primary virtue of using
-the latter two networks is that they allow us to track changes away from
-$Y_e = 0.5$.
-
-All four of these networks have been ported into a form that is consistent
-with the \amrex\ codes, in the freely available \microphysics\ code
-repository\footnote{\microphysics\ can be obtained at \url{https://github.com/starkiller-astro/Microphysics}.},
+energy output of Type Ia supernovae. In this paper we use the \aproxthirteen\
+network\ \citep{timmes:1999,timmes:2000}. This includes all thirteen $\alpha$-chain
+particles between ${}^4$He and ${}^{56}$Ni. This network was used by \citet{hawley:2012}
+and \citet{raskin:2010}. \citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a
+very similar network that additionally included ${}^{60}$Zn. This network has been
+ported into a form that is consistent with the \amrex\ codes, in the freely available
+\microphysics\ code repository
+\footnote{\microphysics\ can be obtained at \url{https://github.com/starkiller-astro/Microphysics}.},
 a collection of microphysical routines that are designed to be used in our
-hydrodynamics codes. These can be easily swapped at compile time by using the 
-appropriate makefile variable.
+hydrodynamics codes. In our \microphysics\ repository we have also ported \aproxnineteen\
+\citep{timmes:1999}, which was used in a few previous collision papers, and \aproxtwentyone.
+The isotopes added by these networks relative to \aproxthirteen\ are not particularly relevant
+to the problem of head-on white dwarf collisions, so we do not discuss results with these networks
+here (though we have verified for this problem that these larger networks do not significantly change
+the answer). We also make available \isoseven\ \citep{timmes:2000}, which skips the nuclides between
+silicon and nickel, and was used by \citet{rosswog:2009}. \isoseven\ gives significantly different
+quantitative results than \aproxthirteen\ for this problem, due to the inapplicable assumption of
+an equilibrium abundance of the unused isotopes, but the results are still qualitatively similar.
 
@@
 
-\subsection{Nuclear Network}
-\label{sec:parameters:network}
-
-Now we move to a discussion of the effect of including various isotopes
-in the nuclear reaction network. In particular we compare the \isoseven, \aproxthirteen,
-\aproxnineteen, and \aproxtwentyone\ reaction networks. In all cases we run
-the same problem setup, and the unused isotopes (everything but carbon and oxygen)
-have their mass fractions in every zone set to a negligibly small number at initialization.
-\autoref{table:networks} lists the $^{56}$Ni generation and total energy generation (determined
-by subtracting the initial energy from the final energy). The energy generated in the latter
-three cases agrees remarkably well, with a less than $0.5\%$ difference. \aproxthirteen\
-over-generates nickel at the $2\%$ level compared to the networks with more isotopes, a reasonably
-small difference. On the other hand, agreement with \isoseven\ is not very strong, as \isoseven\
-generates far too much nickel and far too little energy. Nevertheless while the quantitative
-agreement is not there, the collision looks qualitatively similar compared to the evolution with
-more complicated networks, so while \isoseven\ should not be used for predicting nucleosynthetic
-yields in production science runs, it may still be fair for use in proof-of-concept studies.
-
-\input{plots/networks_m_P_0.64_m_S_0.64.tbl}
-
-\isoseven\ also has the peculiar property that after the detonation has passed through the WDs,
-the region at the central contact point does not fully stop burning, yielding a quasistatic energy
-release that can add thousands to millions of extra timesteps to the simulation (in addition to the few
-hundred it normally takes, using our designated stopping criterion). The full effect of these additional
-timesteps only amounts to a trace change in the nickel production. By examining the behavior of zones
-near the collision point, we discovered the reason for this. The zones in question have a composition that
-is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in
-comparison to, say, \aproxthirteen, reveals that $dX/dt$ for these isotopes is off by many orders of magnitude.
-This is a consequence of the choice inherent to the network to assume that the isotopes in between $^{28}$Si
-and $^{56}$Ni are in an effective equilibrium state. So an advective update that yields even a small change
-to the abundances in the zone will very strongly knock the zone out of equilibrium with respect to the burn
-step following the advective update. This then causes the burning timestep limiter to kick in and make the
-timestep very small so that the advective changes do not cause such strong effects on the burning equilibrium.
-This effect is an argument against the use of \isoseven\ for explosive, non-hydrostatic burning problems.
-
-\aproxnineteen\ and \aproxtwentyone\ both take significantly longer to evaluate at our default
-ODE tolerances than \aproxthirteen\ does, even after accounting for the increased number of species.
-This is due to a significant number of integration failures and subsequent retries (see the discussion
-in \autoref{sec:burner}). The species tolerance of $10^{-8}$ is challenging for these networks.
-
 \subsection{Burning Mode}

### 2017-09-19 science/wdmerger_II/run.sh

@@ -296,6 +296,10 @@ problem_dir=$CASTRO_HOME/Exec/science/wdmerger
 
 DIM="2"
 
+# Use the aprox13 network for all following tests.
+
+NETWORK_DIR="aprox13"
+
 # Get the right inputs and probin files.
 
 inputs="inputs_2d"
@@ -379,7 +383,6 @@ dtnuc_X_default="1.e200"
 dxnuc_default="1.0e200"
 mass_P_default="0.64"
 mass_S_default="0.64"
-network_default="aprox13"
 limiter_mode_default="1"
 burning_mode_default="1"
 T_min_default="1.0e8"
@@ -604,27 +607,6 @@ do
 done
 
 castro_small_temp=$small_temp_default
-
-
-
-# Test the effect of the various networks.
-
-network_list="iso7 aprox13 aprox19 aprox21"
-local_compile=1
-
-for NETWORK_DIR in $network_list
-do
-
-    dir=$results_dir/networks/$NETWORK_DIR
-
-    set_run_opts
-    run
-
-done
-
-NETWORK_DIR=$network_default
-
-unset local_compile
 
 
 

### 2017-09-19 papers/wdmerger_II/paper.tex
Significant reorganization and cleanup

@@
 error tolerances are uniformly $10^{-6}$ for the energy, temperature, and nuclear
-abundance ODEs (we discuss the sensitivity to this choice in
-\autoref{sec:parameters:tolerance}). We can loosen these tolerances in an
-attempt to give the integrator an easier time. If both \vode\ and the Stoer
-and Bulirsch algorithm fail, then we loosen the tolerances by a multiplicative
-factor of 25\%, and then repeat the attempts with \vode\ and Stoer and Bulirsch.
-We keep doing this until either we have loosened the tolerances enough for one
-of the two methods to succeed, or until we have loosened the tolerances by two
-orders of magnitude, at which point we finally give up. We have found that this
-scheme is resilient enough to handle all of the simulations shown below.
+abundance ODEs. We can loosen these tolerances in an attempt to give the integrator
+an easier time. If both \vode\ and the Stoer and Bulirsch algorithm fail, then we
+loosen the tolerances by a multiplicative factor of 25\%, and then repeat the attempts
+with \vode\ and Stoer and Bulirsch. We keep doing this until either we have loosened
+the tolerances enough for one of the two methods to succeed, or until we have loosened
+the tolerances by two orders of magnitude, at which point we finally give up. We have
+found that this scheme is resilient enough to handle all of the simulations shown below.
+We have also verified that our results do not depend much on the tolerances when they
+are tighter than $10^{-4}$.
 
@@
 
+%TODO: add a ``Nonaka plot''
+
 To understand the consequences of this choice, and more broadly to 
@@
 a substantial amount of iron-group elements can be be generated,
-perhaps forming a detonation. This can have non-trivial effects
-on the total amount of iron-group elements generated over the course of
-the simulation. We have found that the addition of the 
-limiter based on changes in species functionally precludes this, and
-so the two methods can complement each other.
+perhaps forming a detonation. In practice, however, we find that
+this effect is not large enough to have a meaningful outcome, and
+the results are largely insensitive to which method of limiting
+is used.
 
@@
 stability criteria as judged from the end of the timestep.
-However we have found that the benefits for this problem are
-small.
+We have found that the benefits for this problem are, however, small.
 
@@
 we desire the most conservative version of the limiter. We discuss
-our results with the suppressed burning mode in \autoref{sec:parameters:burningmode}.
+our results with the suppressed burning mode in \autoref{sec:collisionburning}.
 Note that we have found that with this option enabled, it typically takes
@@
 in both dimensions, so that there are twice as many zones in the $z$ dimension in
-a uniform grid. The default grid has 256 zones along the $z$ axis, yielding a
-spatial resolution of 312.5 km.
+a uniform grid.
 
@@
 
+Our primary metric for this test is the amount of $^{56}$Ni generated in the collision,
+as this is the parameter most directly related to the observable quantities of interest
+for Type Ia supernovae. We collect information at the end of every timestep about the
+total amount of nickel on the grid (in solar masses), and we use the maximum value of
+this nickel mass.
+
 %==========================================================================
-% Parameter Study
+% Standard Results
 %==========================================================================
-\section{Parameter Study}
-\label{sec:parameters}
+\section{Collision Burning}
+\label{sec:collisionburning}
 
-Before turning to the effect that adaptive mesh refinement has on WD collisions,
-we will first examine the effect of a number of other code parameters. For most
-of the following we will use two-dimensional axisymmetric simulations which have
-impact parameter $b = 0$, unless otherwise specified, and we will stick to a
-moderate resolution (we use only the coarse grid described in \autoref{sec:problemsetup}).
-For all simulations in this section, we use a pair of $0.64\ \msolar$ WDs, a binary
-system that has been studied in most of the collision papers to date. This allows
-us to cheaply test how the collision responds to many effects. One limitation of
-this is that we will not find out exactly how a simulation differs in the case of a
-high-resolution 3D run when these parameters changed. Another is that we will not
-perform a full parameter study where we consider the full non-linear dependence of
-all parameters on all other parameters; here we are merely interested in qualitative
-trends that will help to understand the limitations of WD collision simulations.
-
-With those caveats in mind, let us consider the tests and the results we obtained.
-Our primary metric for the following tests will be the amount of $^{56}$Ni generated
-in the collision, as this is the parameter most directly related to the observable
-quantities of interest for Type Ia supernovae. Specifically, we collect information
-at the end of every timestep about the total amount of nickel on the grid (in solar
-masses), and we will use the maximum value of this nickel mass. When the stopping
-criterion described in \autoref{sec:problemsetup} is used, this nickel mass is generally
-constant or decreasing by the end of the run. The typical amount of nickel mass generated
-for our runs at low resolution is $0.2\ \msolar$, which is significantly lower than
-that obtained by \cite{raskin:2010} and \cite{kushnir:2013} for the same problem.
-We will discuss this in \autoref{sec:resolution}.
-
-\subsection{Timestepping}
-\label{sec:parameters:timestepping}
-
-For the standard self-heating burns that we use, the most important parameters we
-examined come from the timestep limiting scheme described in \autoref{sec:hydrocoupling}.
-The parameters $f_{be}$ (from \autoref{eq:timestep_e}) and $f_{bX}$ (from
-\autoref{eq:timestep_X}) control the size of the timestep as a function of the
-burning rate. As the timestep gets smaller, the error due to the Strang splitting
-scheme coupling the reactions and the hydro also gets smaller, and we expect
-that the results become more accurate.
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
-  \caption{Nickel production as a function of the timestep factors $f_{be}$
-           (solid blue, limiting the timestep on changes in internal energy) and
-           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
-           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
-           the timestep by at least an order of magnitude more.
-           \label{fig:timestep_factor}}
-\end{figure}
-\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
-production (these limiters are somewhat redundant in the sense that a given $f_{bX}$
-roughly corresponds to another (higher) $f_{be}$, as is evident by the fact that on
-the graph the curves look qualitatively similar in shape but with a horizontal offset.
-As either $f$ becomes small, the restriction on the timestep becomes quite severe:
-for $f_{be} = 0.001$, the minimum timestep in the simulation is smaller than
-$5 \times 10^{-9}\ \text{s}$ and the full simulation requires over 100,000 timesteps.
-Such a simulation would be prohibitively expensive for a high resolution 3D calculation.
-Yet it is clear that even for this steep value of the limiter, which is much stricter
-than the limiter used in, e.g., \cite{hawley:2012} and \cite{raskin:2010}, the
-simulation has not fully converged in nickel production.
-
-We also examined the effect of the method used to enforce the limiting, using the
-four limiting modes discussed in \autoref{sec:hydrocoupling}. The two modes that
-use the full last timestep's worth of information to estimate $\dot{e}$ (Mode 3 and Mode 4)
-yield a slightly larger nickel mass than the two modes that only use more recent information,
-but the difference is less than 10\%, and it is not clear which is preferable since
-for the value of the timestep limiter we used to compare these modes ($f_{be} = 0.3$),
-the simulation is not converged anyway.
-
-%TODO: add a ``Nonaka plot''
-
-For all remaining tests in this section, unless otherwise stated, we take $f_{be} = 10^2$
-and disable limiting with $f_{bX}$ by setting it to a large number. This choice was made
-somewhat arbitrarily for computational efficiency.
-
-\subsection{Integration Error Tolerance}
-\label{sec:parameters:tolerance}
-
-%TODO: add discussion of how ODE tolerances affect the answer
-
-\subsection{Burning Mode}
-\label{sec:parameters:burningmode}
+% Discuss the results for the case of the uniform grid with 256 zones along
+% the $z$ axis, yielding a spatial resolution of 312.5 km, with no timestep limiting.
+% Discuss the effect of the burning mode here.
 
@@
 %==========================================================================
+% Temporal Resolution
+%==========================================================================
+\section{Temporal Resolution}
+\label{sec:temporalresolution}
+
+For the standard self-heating burns that we use, the most important parameters we
+examined come from the timestep limiting scheme described in \autoref{sec:hydrocoupling}.
+The parameters $f_{be}$ (from \autoref{eq:timestep_e}) and $f_{bX}$ (from
+\autoref{eq:timestep_X}) control the size of the timestep as a function of the
+burning rate. As the timestep gets smaller, the error due to the Strang splitting
+scheme coupling the reactions and the hydro also gets smaller, and we expect
+that the results become more accurate.
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
+  \caption{Nickel production as a function of the timestep factors $f_{be}$
+           (solid blue, limiting the timestep on changes in internal energy) and
+           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
+           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
+           the timestep by at least an order of magnitude more.
+           \label{fig:timestep_factor}}
+\end{figure}
+\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
+production (these limiters are somewhat redundant in the sense that a given $f_{bX}$
+roughly corresponds to another (higher) $f_{be}$, as is evident by the fact that on
+the graph the curves look qualitatively similar in shape but with a horizontal offset.
+As either $f$ becomes small, the restriction on the timestep becomes quite severe:
+for $f_{be} = 0.001$, the minimum timestep in the simulation is smaller than
+$5 \times 10^{-9}\ \text{s}$ and the full simulation requires over 100,000 timesteps.
+Such a simulation would be prohibitively expensive for a high resolution 3D calculation.
+Yet it is clear that even for this steep value of the limiter, which is much stricter
+than the limiter used in, e.g., \cite{hawley:2012} and \cite{raskin:2010}, the
+simulation has not fully converged in nickel production.
+
+%TODO: update the above with more recent data.
+
+%==========================================================================
 % Resolution Dependence
 %==========================================================================
-\section{Resolution Dependence}
-\label{sec:resolution}
+\section{Spatial Resolution}
+\label{sec:spatialresolution}
 

### 2017-09-19 papers/wdmerger_II/paper.tex
Update the timestep limiting discussion with preliminary data

@@
 
-For the standard self-heating burns that we use, the most important parameters we
-examined come from the timestep limiting scheme described in \autoref{sec:hydrocoupling}.
+For the standard self-heating burns that we use, the timestep limiting scheme
+described in \autoref{sec:hydrocoupling} has a significant effect on the evolution.
 The parameters $f_{be}$ (from \autoref{eq:timestep_e}) and $f_{bX}$ (from
@@
 Such a simulation would be prohibitively expensive for a high resolution 3D calculation.
-Yet it is clear that even for this steep value of the limiter, which is much stricter
-than the limiter used in, e.g., \cite{hawley:2012} and \cite{raskin:2010}, the
-simulation has not fully converged in nickel production.
+It is clear, though, that the timestep effect has converged when $f_{be} \sim 0.1$. We
+agree that the use of values in this range by \cite{raskin:2010}, \cite{hawley:2012},
+and \cite{kushnir:2013} is appropriate.
 

### 2017-09-19 papers/wdmerger_II/paper.tex
Simplify the timestep limiting discussion

@@
 \begin{align}
-  \Delta t_{be} &= f_{e}\, \tau_e \label{eq:timestep_e}\\
-  \Delta t_{bX_k} &= f_{X}\, \tau_{X_k}. \label{eq:timestep_X}
+  \Delta t_{be} &= f_{be}\, \tau_e \label{eq:timestep_e}\\
+  \Delta t_{bX_k} &= f_{bX}\, \tau_{X_k}. \label{eq:timestep_X}
 \end{align}
-Given an estimate for $\dot{e}$, the factor $f_{e}$ determines by what 
+Given an estimate for $\dot{e}$, the factor $f_{be}$ determines by what 
 fraction we would like to allow the internal energy to change
 in the current timestep, under the assumption that $\dot{e}$ does not change from
-timestep to timestep. Similarly, given an estimate for $\dot{X_k}$, the factor $f_{X}$ 
+timestep to timestep. Similarly, given an estimate for $\dot{X_k}$, the factor $f_{bX}$ 
 determines the maximum change in the mass fraction of any species. To prevent species
 whose abundances are small in absolute terms from controlling the timestep, we apply
-the limiter based on $f_{X}$ only to species with abundances in a given zone greater
+the limiter based on $f_{bX}$ only to species with abundances in a given zone greater
 than $10^{-3}$ (this abundance threshold can be changed by the user of the code). By making 
-$f_{e}$ and $f_{X}$ smaller, we can control the magnitude of the decoupling 
-between the reactions and the hydro. A typical choice for $f_e$ parameter in the
-literature is in the range of 0.2 or 0.3, while to our knowledge a limiter based on
-$f_X$ has not been used by others performing these types of calculations. The sensitivity
+$f_{be}$ and $f_{bX}$ smaller, we can control the magnitude of the decoupling 
+between the reactions and the hydro. A typical choice for $f_{be}$ in the
+literature is in the range of 0.1 to 0.3, while to our knowledge a limiter based on
+$f_{bX}$ has not been used by others performing these types of calculations. The sensitivity
 of results to the value of these timestep limiters will be discussed in 
-\autoref{sec:parameters:timestepping}. The factors $f_{e}$ and $f_{X}$ can be set at runtime in \castro.
+\autoref{sec:temporalresolution}. The factors $f_{be}$ and $f_{bX}$ can be set at runtime in \castro.
 
@@
 burning timesteps across all zones. To do this, we need a method for determining 
-$\dot{e}$ and $\dot{X_k}$. A typical choice in the literature has been to set, for example,
-\begin{equation}
-  \dot{e} = \frac{e^{n} - e^{n-1}}{\Delta t^{n-1}}, \label{eq:burning_limiter_mode_4}
-\end{equation}
-where is $e^n$ is the internal energy at the start of the current timestep and
-$e^{n-1}$ is the internal energy at the start of the previous timestep. 
-The obvious analogue is used for constructing the species rate of change.
-However, there are alternative methods of constructing this derivative estimate, 
-and we have found that these different methods have measurable consequences.
-We define four separate methods for calculating the time derivative, with 
-the above being mode 4. Mode 3 is similar to mode 4 but replaces the
-denominator in \autoref{eq:burning_limiter_mode_4} with the change in 
-internal energy over the last timestep only from the nuclear reactions.
-Mode 2 is the same as mode 3 but we only use the change in internal 
-energy from the most recent nuclear burning step, that is, the second-half
-of the Strang-split burning from the last timestep (the denominator 
-then becomes $\Delta t / 2$). In mode 1, the most accurate option and 
-the current default in \castro, we evaluate the right-hand-side of the 
-burning network given the current state to explicitly obtain the 
-instantaneous value of $\dot{e}$ and $\dot{X_k}$.
+$\dot{e}$ and $\dot{X_k}$. We choose to do this with a direct evaluation of the
+right-hand-side of the burning network, given the current state at the time when
+the timestep is evaluated. Other work in the literature has typically evaluated
+this as a finite difference using the change over the last timestep. Our method
+is more accurate in principle, but in practice for this problem we find no
+meaningful difference between the limiting we propose and other methods of
+estimating the time derivatives.
+
+To be precise, we note that the timestep
+will only actually satisfy the energy criterion
+$\Delta t \leq f_e \tau_e$ and species criterion
+$\Delta t \leq f_{X_k} \tau_{X_k}$ when the estimates for
+$\dot{e}$ and $\dot{X_k}$ we generate are at least as large
+as the actual rate of change of energy and mass fractions
+over the timestep. However, this can assumption will sometimes fail
+during periods of runaway burning when the rate of change
+of these quantities is highly nonlinear. We may not want
+to neglect the errors caused by this approximation
+because they may build up over an extended period of nonlinear
+evolution and perhaps substantially change the final results.
+To this end, we have implemented a timestep retry option in
+\castro, which re-computes an advance if it violated the
+stability criteria as judged from the end of the timestep,
+using subcycled timesteps which individually satisfy the
+stability criterion to the tolerance we have asked for.
+We have found that the benefits for this problem are, however, small.
 
@@
 
-To understand the consequences of this choice, and more broadly to 
+To understand the consequences of the timestep limiting, and more broadly to 
 understand the limitations of Strang splitting, we consider the 
@@
 rather separate discrete burning and hydro steps, and the nature of 
-the burning evolution will be quite different. Furthermore, by the 
-time we return to Step 1 and estimate the next timestep size, all 
-of the burning rates will be small again, and the instantaneous 
-timestep limiter of mode 1 may actually substantially overestimate 
-the needed timestep. The other modes will see that the energy/species  
-substantially changed over the last timestep, but will still
-overestimate the needed timestep because the burning was quiescent
-for at least some portion of the last advance. Our experience has
-shown that none of these methods is flawless, and that limiting
-based on only changes in internal energy is particularly susceptible
-to this staggered burning phenomenon; silicon-group material can
-build up without changing the internal energy by a  large fraction,
-so the timestep limiter is never triggered, and then in a single step
-a substantial amount of iron-group elements can be be generated,
-perhaps forming a detonation. In practice, however, we find that
-this effect is not large enough to have a meaningful outcome, and
-the results are largely insensitive to which method of limiting
-is used.
+the burning evolution will be quite different. As the timestep becomes
+smaller, the evolution becomes smoother and the splitting error decreases.
 
-With the timestep limited the way we advocate in this paper, 
-the timesteps are generally short enough so that the errors 
-due to splitting are small. Other approaches to the coupling 
+Other approaches to the coupling 
 between reactions and hydrodynamics have been proposed in the 
@@
 and it may form the basis of further work on this subject.
-
-Now we return to a point we hinted at above. The timestep
-will only actually satisfy the energy criterion
-$\Delta t \leq f_e \tau_e$ and species criterion
-$\Delta t \leq f_{X_k} \tau_{X_k}$ when the estimates for
-$\dot{e}$ and $\dot{X_k}$ we generate are at least as large
-as the actual rate of change of energy and mass fractions
-over the timestep. However, this can assumption can fail
-during periods of runaway burning when the rate of change
-of these quantities is highly nonlinear. We may not want
-to neglect the errors caused by this approximation
-because they may build up over an extended period of nonlinear
-evolution and perhaps substantially change the final results.
-To this end, we have implemented a timestep retry option in
-\castro, which re-computes an advance if it violated the
-stability criteria as judged from the end of the timestep.
-We have found that the benefits for this problem are, however, small.
 

### 2017-09-19 papers/wdmerger_II/paper.tex
Add preliminary data for spatial resolution dependence

@@
 
-%TODO: update the above with more recent data.
-
 %==========================================================================
@@
 \label{sec:spatialresolution}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.425]{{{plots/amr}}}
+  \caption{Nickel production as a function of the simulation resolution. The solid
+           black curve connects the points at uniform resolution. For each uniform
+           resolution simulation, we include additional simulations using the adaptive
+           mesh refinement criterion described in \autoref{sec:unstable_burning}.
+           The horizontal axis is the effective number of zones per dimension on the
+           finest level, which is the number of zones per dimension that would exist
+           if the finest resolution was uniformly applied on the grid.
+           \label{fig:amr}}
+\end{figure}
 

### 2017-09-22 papers/wdmerger_II/paper.tex
NERSC and OLCF are now officially recognized facilities
Update paper II setup description

@@
 In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
-composed of equal parts carbon and oxygen by mass, for comparison to earlier studies;
-however, the code does have the ability to vary the impact parameter of the collision
-and the mass and composition of the WDs.
+composed of equal parts carbon and oxygen by mass, at an initially uniform temperature
+of $T = 10^7$ K. This choice is for comparison to earlier studies; the code does have
+the ability to vary the impact parameter of the collision and the mass and composition
+of the WDs.
 
@@
 As in Paper I, the region outside the WDs is filled with a low-density ambient
-material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition is the
+material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition and temperature is the
 same as that of the WDs. Reactions in this region are unimportant, so for
@@
 $\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
-the results do not depend on this choice. The initial ambient
-temperature is $10^7\ \text{K}$, which is the same as the initial temperature
-everywhere throughout the WDs, and we set the temperature floor to the same
-value (any zone that reaches below this temperature will be reset). Setting the
-temperature floor to a lower value has only minor effects on the simulation outcome.
+the results do not depend on this choice. The density floor for the simulation is set
+at $10^{-5}\ \text{g / cm}^{-3}$ and the temperature floor is $10^7$ K.
 
@@
 total amount of nickel on the grid (in solar masses), and we use the maximum value of
-this nickel mass.
+this nickel mass. Plotfiles of a few relevant quantities (density, temperature, velocity,
+pressure, sound speed, and the ratio $\tau_s / \tau_e$) are created every 0.01 seconds.
+The timesteps are limited to ensure that every multiple of this plotfile interval is
+reached exactly, which implies an effective maximum timestep of 0.01 seconds.
 
@@
 
+\facilities{OLCF, NERSC}
+
 \clearpage

### 2017-09-24 job_scripts/run_utils.sh
Allow overriding of job done status

@@ -2201,7 +2201,7 @@ function run {
 
       done_flag=$(is_dir_done)
 
-      if [ $done_flag -eq 0 ]; then
+      if [ $done_flag -eq 0 ] || [ "$submit_even_if_done" == "1" ]; then
 
   	  echo "Continuing job in directory $dir."
 

### 2017-09-24 papers/wdmerger_II/paper.tex
Add sponge comment to paper

@@
 
-We terminate the simulation at $t = 10$ seconds, as we have found that for this
+We terminate the simulation at $t = 9$ seconds, as we have found that for this
 particular setup, all of the significant burning has completed by this time. At
@@
 $\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
-the results do not depend on this choice. The density floor for the simulation is set
+the results do not depend on this choice. A sponge is applied to all gas at a distance
+greater than $2.5 \times 10^9$ cm from the origin, which is just larger than the
+initial location of the outer edges of the white dwarfs. The sponge damps out the
+velocity in this region (see Paper I) to prevent spurious numerical error from dragging
+down the global timestep. The density floor for the simulation is set
 at $10^{-5}\ \text{g / cm}^{-3}$ and the temperature floor is $10^7$ K.

### 2017-09-25 documentation/userguide.tex
Minor updates to the documentation

@@
   \item \texttt{AMREX\_HOME}, the location of the \amrex\ code's top-level directory
-  \item \texttt{CASTRO\_DIR}, the location of the \castro\ code's top-level directory
+  \item \texttt{CASTRO\_HOME}, the location of the \castro\ code's top-level directory
   \item \texttt{WDMERGER\_HOME}, the location of this software's top-level directory
-  \item \texttt{MICROPHYSICS\_DIR}, the location of the \texttt{Microphysics} code's top-level directory
+  \item \texttt{MICROPHYSICS\_HOME}, the location of the \texttt{Microphysics} code's top-level directory
 \end{itemize}
@@
 like \texttt{sed}. We do not guarantee that this software will work out of the box on OS X
-or other Unix derivatives; however, we generally try to avoid GNU-specific options, so that 
-trying to run this software on OS X will not be too painful. With that being said, this 
-software does heavily rely on \texttt{bash} and at present there are no plans to maintain
-compatibility with any other shell.
+or other Unix derivatives; however, we have generally tried to avoid GNU-specific options
+where possible. This software does heavily rely on \texttt{bash} and at present there are
+no plans to maintain compatibility with any other shell.
 
@@
 appropriate initial orbital distance. The center of mass of the system corresponds with the center 
-of the domain, which is the origin; the domain is $1.024 \times 10^{10}\ \text{cm}$ wide in each 
-coordinate direction, so that each dimension runs from $-5.12 \times 10^{9}\ \text{cm}$ to 
-$+5.12 \times 10^{9}\ \text{cm}$. In a rotating reference frame, we set the stars at rest. 
+of the domain, which is the origin. In a rotating reference frame, we set the stars at rest. 
 We can instead run a simulation in the inertial reference frame. To make it equivalent, we
@@
 reference frame velocity. This means that the stars will be synchronously rotating (though we
-may relax this assumption in future work).
+may relax this assumption in the future).
 
@@
   stop_time=$period
-  run $dir $nprocs $walltime
+  run
 done
@@
 
-This project was started as the Ph.D. dissertation project for Max Katz at Stony Brook University. 
-Michael Zingale is Max's primary thesis advisor. Correspondence and requests for information 
+This project was started as the Ph.D. dissertation project for Max Katz at Stony Brook University;
+Max now continues work on it, though he is no longer physically at Stony Brook.
+Michael Zingale is the other main contact at Stony Brook. Correspondence and requests for information 
 about this software should be directed to them; their e-mail addresses are maximilian.katz@stonybrook.edu

### 2017-09-25 job_scripts/run_utils.sh
Add a function for echoing to stderr

@@ -12,6 +12,9 @@ else
   script_dir="$WDMERGER_HOME/job_scripts"
 fi
 
+# Utility shell functions.
+source $script_dir/util.sh
+
 # Functions for working with inputs and probin files.
 source $script_dir/inputs.sh
 source $script_dir/probin.sh
@@ -157,7 +160,7 @@ function get_last_output {
 function get_last_checkpoint {
 
   if [ -z $1 ]; then
-      echo "No directory passed to get_last_checkpoint; exiting." >&2
+      echoerr "No directory passed to get_last_checkpoint; exiting."
       echo ""
       return
   else
@@ -406,7 +409,7 @@ function get_remaining_walltime {
   if [ ! -z $1 ]; then
       job_number=$1
   else
-      echo "Job number not given to get_remaining_walltime" >&2
+      echoerr "Job number not given to get_remaining_walltime"
       echo ""
       return
   fi
@@ -422,7 +425,7 @@ function get_remaining_walltime {
       total_time=$(showq -u $USER | grep $job_number | awk '{ print $5 }')
 
       if [ -z $total_time ]; then
-	  echo "Unable to get total_time from job submission system" >&2
+	  echoerr "Unable to get total_time from job submission system"
       fi
 
       if [ ! -z $total_time ]; then
@@ -461,7 +464,7 @@ function get_remaining_walltime {
 
   else
 
-      echo "Unknown job submission system in get_remaining_walltime" >&2
+      echoerr "Unknown job submission system in get_remaining_walltime"
 
   fi
 
@@ -1462,7 +1465,7 @@ function get_last_submitted_job {
 
   else
 
-      echo "No jobs_submitted.txt file to obtain last job number from" >&2
+      echoerr "No jobs_submitted.txt file to obtain last job number from"
 
       job_number=-1
 

### 2017-09-25 job_scripts/run_utils.sh
Speed up the checks again

@@ -532,6 +532,8 @@ function is_job_running {
       num_jobs_in_dir=$(cat $dir/jobs_submitted.txt | wc -l)
       jobs_in_directory=$(cat $dir/jobs_submitted.txt | awk '{print $1}')
 
+      get_submitted_jobs
+
       for job1 in ${job_arr[@]}
       do
 	  for job2 in $jobs_in_directory
@@ -2181,11 +2183,19 @@ function run {
   # job is already done, or if a job is
   # currently running.
 
-  done_flag=$(is_dir_done)
+  done_flag=''
   job_running_status=$(is_job_running $dir)
 
-  if [ $done_flag -ne 1 ] && [ $job_running_status -ne 1 ]; then
-      set_up_problem_dir
+  if [ $job_running_status -ne 1 ]; then
+
+      if [ "$done_flag" == '' ]; then
+          done_flag=$(is_dir_done)
+      fi
+
+      if [ $done_flag -ne 1 ]; then
+          set_up_problem_dir
+      fi
+
   fi
 
   if [ -z $nprocs ]; then
@@ -2222,6 +2232,10 @@ function run {
 
       rm -f $dir/dump_and_stop
 
+      if [ "$done_flag" == '' ]; then
+          done_flag=$(is_dir_done)
+      fi
+
       if [ $done_flag -eq 0 ] || [ "$submit_even_if_done" == "1" ]; then
 
   	  echo "Continuing job in directory $dir."
@@ -2243,20 +2257,28 @@ function run {
 
   # Optionally, the user can force a recompile from the run script.
 
-  if [ $done_flag -ne 1 ] && [ $job_running_status -ne 1 ]; then
+  if [ $job_running_status -ne 1 ]; then
 
-      if [ ! -z "$local_compile" ] && [ -z "$inputs_only" ]; then
-          if [ "$local_compile" -eq "1" ]; then
-	      compile_in_job_directory $dir
-          fi
+      if [ "$done_flag" == '' ]; then
+          done_flag=$(is_dir_done)
       fi
 
-      copy_files $dir
+      if [ $done_flag -ne 1 ]; then
 
-      if [ -z "$inputs_only" ]; then
-          if [ ! -e "$dir/$job_script" ]; then
-	      create_job_script $dir $nprocs $walltime
+          if [ ! -z "$local_compile" ] && [ -z "$inputs_only" ]; then
+              if [ "$local_compile" -eq "1" ]; then
+	          compile_in_job_directory $dir
+              fi
[…]

### 2017-09-25 papers/wdmerger_II/paper.tex
Update description of ODE tolerances

@@
 found that this scheme is resilient enough to handle all of the simulations shown below.
-We have also verified that our results do not depend much on the tolerances when they
-are tighter than $10^{-4}$.
+We have also verified that our results do not depend strongly on the tolerances when they
+are tighter than $10^{-4}$. For the typical collision problem we discuss below, the
+nucleosynthetic yield for nickel varies by less than 5\% between a tolerance of $10^{-4}$
+and a tolerance of $10^{-12}$.
 

### 2017-09-25 papers/wdmerger_II/paper.tex
Update sponge discussion

@@
 $\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
-the results do not depend on this choice. A sponge is applied to all gas at a distance
-greater than $2.5 \times 10^9$ cm from the origin, which is just larger than the
-initial location of the outer edges of the white dwarfs. The sponge damps out the
-velocity in this region (see Paper I) to prevent spurious numerical error from dragging
+the results do not depend on this choice. A sponge is applied to all gas of density
+lower than $1\ \text{g / cm}^{-3}$. The sponge damps out the velocity in these zones
+(see Paper I) to prevent spurious numerical error in low-density gas from dragging
 down the global timestep. The density floor for the simulation is set

### 2017-09-25 analysis/wdmerger.py
Remove some unnecessary warning messages

@@ -324,9 +324,10 @@ def get_last_checkpoint(directory):
 
     import os
 
+    checkpoint = ""
+
     if (not os.path.isdir(directory)):
-        print("Error: directory " + directory + " does not exist in get_last_checkpoint().")
-        return
+        return checkpoint
 
     # Doing a search this way will treat first any checkpoint files 
     # with seven digits, and then will fall back to ones with six and then five digits.
@@ -354,12 +355,9 @@ def get_last_checkpoint(directory):
     add_to_list(checkpointList, checkpointNums, directory + '/*/*chk?????')
 
     if not checkpointList or not checkpointNums:
-        print("Error: no checkpoints found in directory " + directory)
-        return
+        return checkpoint
 
     # Match up the last checkpoint number with the actual file path location. 
-
-    checkpoint = ""
 
     for chkNum in checkpointNums:
 
@@ -832,9 +830,10 @@ def is_dir_done(directory):
 
     import os
 
+    done_status = 0
+
     if not os.path.isdir(directory):
-        print("Error: directory " + directory + " does not exist in is_dir_done().")
-        return
+        return done_status
 
     if os.path.isfile(directory + '/jobIsDone'):
 
@@ -860,7 +859,6 @@ def is_dir_done(directory):
             pass
 
         if not checkpoint and not last_output:
-            print("Error: no checkpoint or stdout in directory " + directory)
             done_status = 0
             return done_status
 

### 2017-09-25 papers/wdmerger_II/paper.tex
Start writing up the AMR section

@@
 
+Now we turn to an examination of the effect of spatial resolution on the detonation.
+The first question we want to answer is: at reasonable uniform spatial resolution, is
+the simulation converged when a detonation occurs? If the answer is no, then the
+follow-up question is: can adaptive mesh refinement help achieve a converged simulation?
+
+To answer these questions we performed a series of simulations at varying base
+resolutions. In each case we ran the simulation to the point when the density
+at the collision point reached $10^7\ \text{g / cm}^{-3}$, which we will informally
+call the collision time, $t_{\rm c}$. The collision time is about $t_{\rm c} = 6.85\ \text{s}$
+at high resolution. At low resolution, the simulation is slightly different because
+of the distortion imposed by a rectangular grid representing initially spherical stars,
+so the collision time is slightly later. Then, we restarted the run from this checkpoint
+and reset the time to zero (such that $t_{\rm c} = 0$). The new simulation runs until about
+$t = 2.15\ \text{s}$, so that the full 9 seconds are simulated. During this second phase
+the detonation occurs. An estimate of the detonation time, $t_{\rm d}$, that serves well
+is the time when the peak density on the grid first reaches $5 \times 10^9\ \text{K}$.
+We measure the detonation time relative to $t_{\rm c}$. Approximately speaking, the later
+$t_{\rm d}$ is, the more nickel is produced. As time passes, more white dwarf material
+slams into the center, compressing and heating up the gas. Both effects are favorable
+to production of nickel. So, a later $t_{\rm d}$ implies that more material can be converted
+to nickel. Similarly, the speed of the detonation matters. A slower speed detonation prolongs
+the collision event, allowing more gas to be processed into nickel.
+
 \begin{figure}[ht]

### 2017-09-25 science/wdmerger_II/analyze.py
Update the collision analysis script

@@ -5,22 +5,39 @@ matplotlib.use('agg')
 from matplotlib import pyplot as plt
 import os
 
-linestyles = ['-', '--', ':', '-', '--', ':']
-markers = ['', '', '', 'o', 's', 'D']
-colors = ['b', 'g', 'r', 'c', 'm', 'k']
+linestyles = ['-', '--', ':,' ':']
+markers = ['', '', '', 'o']
 
-def get_ni56(results_dir):
+# Colors are chosen with the assistance of colorbrewer2.org,
+# using 4 qualitative classes.
+
+colors = ['#1f78b4', '#33a02c', '#a6cee3', '#b2df8a']
+
+def get_ni56(results_dir, prefix = ''):
     """Return a list of maximum 56Ni production from all completed sub-directories in results_dir."""
 
     import wdmerger_spec_diag_analysis as spec_diag
     import numpy as np
 
+    ni56_arr = []
+
     dir_list = wdmerger.get_parameter_list(results_dir)
 
+    # Strip out those directories that don't match the prefix.
+
+    if (prefix != ''):
+        dir_list = [dir for dir in dir_list if dir[0:len(prefix)] == prefix]
+
     if (dir_list == []):
-        return
+        return ni56_arr
 
     diag_filename_list = [results_dir + '/' + directory + '/output/species_diag.out' for directory in dir_list]
+
+    # Ensure all of the output files are there.
+
+    for diag_file in diag_filename_list:
+        if not os.path.isfile(diag_file):
+            return ni56_arr
 
     ni56_arr = [np.amax(wdmerger.get_column('Mass ni56', diag_filename)) - np.amin(wdmerger.get_column('Mass ni56', diag_filename))
                 for diag_filename in diag_filename_list]
@@ -197,6 +214,8 @@ def burning_limiter_e(eps_filename, results_base):
 
     if (os.path.isfile(eps_filename)):
         return
+    else:
+        print("Generating file %s" % eps_filename)
 
     results_dir = results_base + 'burning_limiter_e/'
 
@@ -216,15 +235,17 @@ def burning_limiter_e(eps_filename, results_base):
     ni56_arr = np.array([x for _,x in sorted(zip(dtnuc_list,ni56_arr))])
     dtnuc_list = sorted(dtnuc_list)
 
-    plt.plot(dtnuc_list, ni56_arr, linestyle=linestyles[0], lw = 4.0)
[…]

### 2017-10-07 papers/wdmerger_II/paper.tex
Simplify discussion of burning modes

@@
 ``suppressed'' burn for the remainder of this work. All four options
-are implemented in our burner software. The simulations shown in this work all
-use the self-heating mode unless otherwise specified.
+are implemented in our burner software.
+
+Except for the discussion later about the suppressed burn mode, the simulations
+shown in this work all use the self-heating mode. The hydrostatic and hybrid burns
+are only a slight variation on the self-heating burn in practice for this problem.
+The reason that the hydrostatic burn cannot substantially change the outcome is
+that even though the temperature is fixed during the burn, it is \textit{not}
+constant in the rest of the hydrodynamics step. And in the hydrodynamics step,
+the \castro\ algorithm performs multiple EOS calls to ensure that the temperature
+is synchronized with the internal energy at various points in the update. So in
+practice there is a temperature change due to the burn for our problem. This seems
+inevitable for any hydro code unless the hydrodynamics update includes an explicit
+equation for $T$ that is independent of what is happening for the internal energy.
+The hydrostatic burn does slightly limit the energy release during any given burn,
+but summed up over the whole simulation this is not enough to qualitatively affect
+the outcome. The hybrid burn does not affect the simulation outcome at all relative
+to the hydrostatic burn. The difference between them only applies for zones that have
+a net negative energy release from the hydrostatic burn, which occurs only after
+we have burned to NSE. The hybrid burn was mainly implemented by \citet{raskin:2010}
+for making it easier to numerically complete a burn, rather than because of the
+effects it has on the outcome of the simulation.
 
@@
 total amount of nickel on the grid (in solar masses), and we use the maximum value of
-this nickel mass. Plotfiles of a few relevant quantities (density, temperature, velocity,
-pressure, sound speed, and the ratio $\tau_s / \tau_e$) are created every 0.01 seconds.
+this nickel mass. Plotfiles of relevant quantities are created every 0.01 seconds.
 The timesteps are limited to ensure that every multiple of this plotfile interval is
-reached exactly, which implies an effective maximum timestep of 0.01 seconds.
+reached exactly, which means an effective maximum timestep of 0.01 seconds.
 
@@
 
-In \autoref{sec:burner} we observed that there are several alternatives to the traditional
-self-heating approach in a burn. These alternatives are relevant to a collision problem because,
-as pointed out by \cite{raskin:2010} and \cite{kushnir:2013}, it is possible for a typical
-zone in our simulation to release a very large amount of energy in a burn before it is cooled
+In \autoref{sec:burner} we observed that there is an alternative to the traditional
+self-heating approach in a burn, the ``suppressed'' burning mode. This alternative
+is relevant to a collision problem because, as pointed out by \cite{raskin:2010}
+and \cite{kushnir:2013}, it is possible for a typical zone in our simulation to
+release a very large amount of energy in a burn before it is cooled
 off by a subsequent hydrodynamic expansion step, leading to the numerically unstable burning
-problem described in \autoref{sec:unstable_burning}. \autoref{table:burningmode} lists nickel
-production for the four burning modes described earlier. The hydrostatic burn does produces
-slightly more nickel than the self-heating burn, consistent with the idea that by limiting the
-energy release from early carbon burning, we can forestall a detonation which is numerically
-spurious. Yet the difference is only about 5\%. The reason is that the hydrostatic burn controls
-the energy release during the burn, but \textit{not} what happens in the rest of the hydrodynamics
-step. And in the hydrodynamics step, the \castro\ algorithm performs multiple EOS calls to ensure
-that the temperature is synchronized with the internal energy at various points in the update.
-So in practice there is a temperature change due to the burn for our problem. This seems inevitable
-for any hydro code unless the hydrodynamics update includes an explicit equation for $T$ that is
-independent of what is happening for the internal energy. Similarly, the hybrid burn does not
-affect the total nickel production relative to the hydrostatic burn. The hybrid burn only changes
-the behavior for zones that have a net negative energy release from the hydrostatic burn, which
-occurs only after we have burned to NSE.
-
-\input{plots/burning_mode_m_P_0.64_m_S_0.64.tbl}
-
-In contrast, the suppressed burning mode (modeled after \citet{kushnir:2013}) produces only about
+problem described in \autoref{sec:unstable_burning}. However, the suppressed burning mode
+(modeled after \citet{kushnir:2013}) produces only about
 $2 \times 10^{-4}\ \msolar$ of $^{56}$Ni. In a suppressed burn, we ensure that at every evaluation
@@
 To answer these questions we performed a series of simulations at varying base
-resolutions. In each case we ran the simulation to the point when the density
-at the collision point reached $10^7\ \text{g / cm}^{-3}$, which we will informally
-call the collision time, $t_{\rm c}$. The collision time is about $t_{\rm c} = 6.85\ \text{s}$
+resolutions, augmented with additional runs with AMR. In each case we ran the
+simulation to the point at which unstable burning was approached. Then we stopped
+the run and created a checkpoint. Starting from that checkpoint, we did a simulation
+at varying levels of refinement based on the unstable burning criterion. We will refer
+to the refinement through the refinement factor, $r$. This factor is the jump in
+resolution between the coarse grid and the finest grid. The uniform base grid case is $r = 1$.
+Meanwhile, $r = 16$ means a uniform coarse grid plus two levels of 4x refinement.
+
+The collision time is about $t_{\rm c} = 6.85\ \text{s}$
 at high resolution. At low resolution, the simulation is slightly different because

### 2017-11-14 papers/wdmerger_II/paper.tex
Move the problem setup to be section 2

@@
 typically available for contemporary hydrodynamics simulations.
+
+%==========================================================================
+% Problem Setup
+%==========================================================================
+\section{White Dwarf Collisions}
+\label{sec:collisions}
+
+We start with a description of the problem setup and results of a typical collision
+at low resolution. These results set the stage for our numerical methodology. This enables
+a concrete discussion about the methodology using the language of how it affects the
+collision outcome. In this setting, a concrete discussion is more fruitful than an abstract one.
+
+\subsection{Problem Setup}
+\label{sec:problemsetup}
+
+We implement the white dwarf collision problem in \castro\ in the \wdmerger\ problem
+directory. We prepare the domain in the way most previous papers have done so:
+the white dwarf centers of mass are initially separated by a distance of four times
+the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
+consequently the initial distance, depends on the equation of state used, and this
+can vary somewhat depending on the relevant physics used, such as Coulomb corrections
+in the Helmholtz equation of state. However, the results do not really depend on the
+exact initial distance as long as there is enough time for the WDs to distort in
+response to tidal forces as they approach. Their initial velocity is that of
+two point-masses in free-fall towards each other coming in from infinity, such that
+the contact point is at the origin, and they approach each other along a coordinate axis.
+In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
+composed of equal parts carbon and oxygen by mass, at an initially uniform temperature
+of $T = 10^7$ K. This choice is for comparison to earlier studies; the code does have
+the ability to vary the impact parameter of the collision and the mass and composition
+of the WDs.
+
+We have added to the \wdmerger\ problem setup the ability to take advantage of 
+the 2D cylindrical ($R-z$) coordinate system evolution in \castro, and we use it
+in this study, because the axisymmetry inherent to the cylindrical coordinate system
+lends itself well to the head-on collision problem. In this coordinate system, we
+align the WDs along the $z$-axis (which is analogous to the $x$-axis in Cartesian
+evolution), with the center of the WDs at $R = 0$. The un-simulated $\phi$ dimension
+then would extend the WDs through a $2\pi$ revolution. The domain has width $4 \times 10^{9}$
+cm along the $z$ axis, and is half as wide along the $R$ axis. The resolution is equal
+in both dimensions, so that there are twice as many zones in the $z$ dimension in
+a uniform grid.
+
+We terminate the simulation at $t = 9$ seconds, as we have found that for this
+particular setup, all of the significant burning has completed by this time. At
+this point, the total energy on the domain has become positive, typically about
+$10^{51}$ ergs, and has stopped increasing. This means that the system has become
+unbound due to nuclear energy release, and that no further meaningful nuclear energy
+generation is occurring. In our discussions below, it helps to have some other
+anchoring points in time when significant events occur. We (arbitrarily) define the
+collision time, $t_c$, as the first time when the maximum density on the grid reaches
+$5 \times 10^6\ \text{g / cm}^3$. This is slightly larger than the initial maximum
+density of the two WDs, and is reached only when the WDs meet at the center of the
+domain. We define the detonation time, $t_d$, as the time when the temperature
+first reaches $5 \times 10^9\ \text{K}$.
+
+As in Paper I, the region outside the WDs is filled with a low-density ambient
+material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition and temperature is the
+same as that of the WDs. Reactions in this region are unimportant, so for
+computational efficiency, we disable all nuclear burning for zones that have
+$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
+the results do not depend on this choice. A sponge is applied to all gas of density
+lower than $1\ \text{g / cm}^{-3}$. The sponge damps out the velocity in these zones
+(see Paper I) to prevent spurious numerical error in low-density gas from dragging
+down the global timestep. The density floor for the simulation is set
+at $10^{-5}\ \text{g / cm}^{-3}$ and the temperature floor is $10^7$ K.
+
+Our primary metric for this test is the amount of $^{56}$Ni generated in the collision,
+as this is the parameter most directly related to the observable quantities of interest
+for Type Ia supernovae. We collect information at the end of every timestep about the
+total amount of nickel on the grid (in solar masses), and we use the maximum value of
+this nickel mass. Plotfiles of relevant quantities are created every 0.01 seconds.
+The timesteps are limited to ensure that every multiple of this plotfile interval is
+reached exactly, which means an effective maximum timestep of 0.01 seconds.
+
+\subsection{Baseline Results}
+\label{sec:baseline_results}
+
+Now let us examine the results for a collision of two 0.64 $\msolar$ WDs. We use
+128 zones in the radial dimension, corresponding to a spatial resolution of 312.5 km.
+Timesteps are limited only by the hydrodynamic stability constraint (and the plotfile
+interval constraint). Resolutions in the range of a few hundred km have been a common
+choice in earlier papers on white dwarf mergers and collisions. No adaptive mesh
+refinement is used. For this configuration, the collision time is $t_c = 6.47\ \text{s}$,
+and the detonation time is $t_d = 7.43\ \text{s}$. The peak density at the detonation time
+is $2.3 \times 10^7\ \text{g / cm}^3$.
 
@@
 %==========================================================================
-% Problem Setup
-%==========================================================================
-\section{Problem Setup}
-\label{sec:problemsetup}
-
-We implement the white dwarf collision problem in \castro\ in the \wdmerger\ problem
-directory. We prepare the domain in the way most previous papers have done so:
-the white dwarf centers of mass are initially separated by a distance of four times
-the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
-consequently the initial distance, depends on the equation of state used, and this
-can vary somewhat depending on the relevant physics used, such as Coulomb corrections
-in the Helmholtz equation of state. However, the results do not really depend on the
-exact initial distance as long as there is enough time for the WDs to distort in
-response to tidal forces as they approach. Their initial velocity is that of
-two point-masses in free-fall towards each other coming in from infinity, such that
-the contact point is at the origin, and they approach each other along a coordinate axis.
-In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
-composed of equal parts carbon and oxygen by mass, at an initially uniform temperature
-of $T = 10^7$ K. This choice is for comparison to earlier studies; the code does have
-the ability to vary the impact parameter of the collision and the mass and composition
-of the WDs.
-
-We have added to the \wdmerger\ problem setup the ability to take advantage of 
-the 2D cylindrical ($R-z$) coordinate system evolution in \castro, and we use it
-in this study, because the axisymmetry inherent to the cylindrical coordinate system
-lends itself well to the head-on collision problem. In this coordinate system, we
-align the WDs along the $z$-axis (which is analogous to the $x$-axis in Cartesian
-evolution), with the center of the WDs at $R = 0$. The un-simulated $\phi$ dimension
-then would extend the WDs through a $2\pi$ revolution. The domain has width $4 \times 10^{9}$
-cm along the $z$ axis, and is half as wide along the $R$ axis. The resolution is equal
-in both dimensions, so that there are twice as many zones in the $z$ dimension in
-a uniform grid.
-
-We terminate the simulation at $t = 9$ seconds, as we have found that for this
-particular setup, all of the significant burning has completed by this time. At
-this point, the total energy on the domain has become positive, typically about
-$10^{51}$ ergs, and has stopped increasing. This means that the system has become
-unbound due to nuclear energy release, and that no further meaningful nuclear energy
-generation is occurring.
-
-As in Paper I, the region outside the WDs is filled with a low-density ambient
-material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition and temperature is the
-same as that of the WDs. Reactions in this region are unimportant, so for
-computational efficiency, we disable all nuclear burning for zones that have
-$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
-the results do not depend on this choice. A sponge is applied to all gas of density
-lower than $1\ \text{g / cm}^{-3}$. The sponge damps out the velocity in these zones
-(see Paper I) to prevent spurious numerical error in low-density gas from dragging
-down the global timestep. The density floor for the simulation is set
-at $10^{-5}\ \text{g / cm}^{-3}$ and the temperature floor is $10^7$ K.
-
-Our primary metric for this test is the amount of $^{56}$Ni generated in the collision,
-as this is the parameter most directly related to the observable quantities of interest
-for Type Ia supernovae. We collect information at the end of every timestep about the
-total amount of nickel on the grid (in solar masses), and we use the maximum value of
-this nickel mass. Plotfiles of relevant quantities are created every 0.01 seconds.
-The timesteps are limited to ensure that every multiple of this plotfile interval is
-reached exactly, which means an effective maximum timestep of 0.01 seconds.
-
-%==========================================================================
 % Standard Results

### 2017-11-14 papers/wdmerger_II/paper.tex
Rework the suppressed burn section

@@
 domain. We define the detonation time, $t_d$, as the time when the temperature
-first reaches $5 \times 10^9\ \text{K}$.
+first reaches $4 \times 10^9\ \text{K}$.
 
@@
 and the detonation time is $t_d = 7.43\ \text{s}$. The peak density at the detonation time
-is $2.3 \times 10^7\ \text{g / cm}^3$.
+is $2.3 \times 10^7\ \text{g / cm}^3$. The detonation occurs at the center of the domain
+where the two WDs initially meet.
 
@@
 
-Except for the discussion later about the suppressed burn mode, the simulations
-shown in this work all use the self-heating mode. The hydrostatic and hybrid burns
-are only a slight variation on the self-heating burn in practice for this problem.
+The simulations shown in this work all use the self-heating mode. The hydrostatic
+and hybrid burns are only a slight variation on the self-heating burn in practice
+for this problem.
 The reason that the hydrostatic burn cannot substantially change the outcome is
@@
 for making it easier to numerically complete a burn, rather than because of the
-effects it has on the outcome of the simulation.
+effects it has on the outcome of the simulation. The suppressed burn has separate
+issues which we defer to \autoref{sec:unstable_burning}
 
@@
 not the only way for the numerical discretization to cause 
-severe errors in the burning. Another failure mode is when
+severe errors in the burning. Another possible failure mode is when
 the energy injection timescale
@@
 a zone faster than it can be advected away by pressure waves.
-This is obviously a problem inherent to numerically discretized
+This is of course a problem inherent only to numerically discretized
 systems as the underlying fluid equations are continuous.
@@
 temperature building up too quickly in the zone; the detonation
-is spurious in this case and should be avoided if possible.
-The goal is to ensure that the following condition holds:
+may be spurious in this case. If the following condition holds,
+we can be confident that a numerically seeded detonation has not occurred:
 \begin{equation}
@@
 sound-crossing time. \citet{kushnir:2013} choose $f_{s} = 0.1$ 
-for their simulations, and we do too (this parameter can be set 
-at runtime in \castro).
+for their simulations. \citet{kushnir:2013} implemented this
+criterion by artificially limiting the magnitude of the energy
+release after a burn.
 
-\citet{kushnir:2013} implemented this criterion by artificially 
-limiting the magnitude of the energy release after a burn. We
-too have developed an option for our burner to do this,
-the ``suppressed'' burning mode. In a suppressed burn, we limit
+To understand this choice, consider the outcome of our baseline case.
+When the two WDs collide, there is a prompt detonation
+at the contact point at the center of the domain. This detonation occurs
+when the density is relatively low because the WDs have not had much
+time to collide, and so the nucleosynthetic yield of nickel is low. If
+this is the case, the colliding WDs are not a good candidate for
+normal SNe Ia. However, if this detonation is delayed, material builds
+up at a stalled shock in the center, and the eventual detonation occurs
+on the outer edge of the stalled shock. There will then be much more
+high density material to process into nickel, and the resulting
+explosion creates an amount of nickel potentially sufficient to explain
+normal SNe Ia. The discussion thus centers on whether the prompt detonation
+is physical. \citeauthor{kushnir:2013} argues that it is not, on the basis
+that $\tau_s > \tau_e$ when this detonation occurs. We find the same thing:
+for the baseline case, the sound-crossing time is about five times longer
+than the energy injection time. This means that a detonation is guaranteed
+to occur for numerical reasons. \citeauthor{kushnir:2013} thus argue that
+the suppressed burn is justifiable, as it suppresses this prompt detonation.
+
+This argument is flawed because a physical detonation may \textit{also}
+occur this way. For example, consider a region of WD material at uniformly
+high temperature, say $5 \times 10^9\ \text{K}$, with an arbitrarily large size,
+say a cube with side length 100 km. This region will very likely detonate,
+even if it is surrounded by much cooler material. By the time the material on
+the edges can advect heat away, the material in the center will have long since
+started burning carbon, as the sound crossing time scale is sufficiently large
+compared to the energy injection time scale. This is true regardless of whether
+the size of this cube corresponds to the spatial resolution in a simulation.
+Suppression of the burn in this case is unphysical: if we have a zone matching
+these characteristics, the zone should detonate.
+
+It is true that when our resolution is low enough, there is a floor on the size
+of a hotspot, possibly making such a detonation more likely. However, this is an
+unavoidable consequence of the low resolution: it is the correct result of the
+simulation that was performed. If the results do not match what occurs at higher
+resolution, then the simulation is not converged and the results are not reliable.
+However, it may also be the case that a higher resolution simulation will yield
+similar results, for example because even at the higher resolution, the physical
+size of the hotspot stays the same. One cannot appeal to the numerical instability
+criterion alone to make a decision in this context.
+
+Furthermore, we find that \textit{all} of the detonations we have seen at low
+and moderate resolution inevitably violate the stability criterion. This is true
+even if the prompt detonation is suppressed: the delayed detonation also is
+numerically unstable. For comparison purposes, we developed an option for our
+burner to use a ``suppressed'' burning mode. In a suppressed burn, we limit
 the changes to the state so that \autoref{eq:burning_limiter_2}
-is always satisfied. To achieve this we directly multiply the
-right-hand-side vector in the integration by a constant factor $F$
+is always satisfied\footnote{To achieve this we directly multiply the
+right-hand-side vector in the integration by a factor $F$
 for all variables, where $F$ is the multiplicative factor needed to
@@
 that goes into the estimate for $\tau_e$ to be the value of the
-internal energy of the zone at the beginning of the burn. If
-instead one allowed $c_s$ and $e$ to evolve with the burn, one
-would obtain a less conservative limiter in the case of explosive
-burning, as $c_s$ and $e$ are both increasing in this case.
-As the point of the limiter is to ensure that the changes to the
-\textit{original} energy are small enough so that the following
-hydrodynamics update can advect away newly generated energy
-quickly enough to avoid a numerically seeded detonation,
-we desire the most conservative version of the limiter. We discuss
-our results with the suppressed burning mode in \autoref{sec:collisionburning}.
-Note that we have found that with this option enabled, it typically takes
-many more timesteps to complete a burn than in self-heating mode.
+internal energy of the zone at the beginning of the burn.}. When we
+evolve the baseline case with this suppressed burner, the prompt detonation
+is indeed suppressed. A later detonation does occur, but this delayed
+detonation also triggers the suppression. The result is that the detonation
+does not get to very high temperatures, and the result is again a ``failed''
+supernova that does not generate nearly enough nickel. Thus we do not
+reproduce the claim of \citeauthor{kushnir:2013} that the suppressed
+burning mode suppresses only the prompt detonation, but allows the
+delayed detonation to occur. We could find some way to turn off the
+suppression and allow the delayed detonation, but without some independent
+confirmation that the delayed detonation is physical (while the prompt
+detonation is not) this would be begging the question.
 
-Regardless of whether the suppressed burning mode works for this
-particular problem, it is not physical, so we include for
-consideration a different approach. If we insist that we
-cannot directly control the energy injection timescale, we 
-must find a way to alter the sound-crossing timescale. 
-We can achieve this by adding levels of refinement in 
-regions that do not satisfy \autoref{eq:burning_limiter_2},
-which effectively lowers $\Delta x$ and thus the
-sound-crossing time. We keep tagging zones for refinement
-based on this criterion until the criterion is satisfied
-on the finest level. Since the concern is regions that 
-may detonate, we also tag nearby zones in a buffer region
-which do not themselves satisfy the criterion,
-so that a detonation in a single timestep cannot 
-escape into non-refined regions. The width of the buffer 
-region should thus be at least as large as the number of 
-timesteps before a regridding procedure is performed.
-We choose a value of two for both the number of zones in the 
-buffer region and the number of steps in between regrids,
-for all simulations in this paper.
-
-We agree with \citet{kushnir:2013} that solving this numerical
-instability is crucial to avoiding unphysical detonations.
-A simulation that does not solve this problem will not obtain
-the correct amount and will not converge properly with resolution.
-This may justify the addition of many AMR levels to the domain
-if a correct evaluation of the burning phase is desired.
-Even if one cannot afford the full resolution required by this
-AMR criterion, and chooses to limit the number of levels to some
-predetermined maximum based on a constraint of computing time,
-the added resolution will still go to the most-needed places.
-
-A potential limitation of this approach is that it does not turn on
-until \textit{after} an unstable burning region has been generated,
-because we generally only perform a regridding and level generation
-step at times when AMR levels are synchronized, such as at the beginning
-of the timestep on the coarsest level. So this AMR criterion cannot
-help in the case where a spurious detonation begins in a single timestep
-on the coarse grid, though it will kick in immediately after that
-timestep to resolve the regions where the detonation has occurred.
-This situation is common in our white dwarf collision simulations.
-We have developed a remedy to this issue. At the \textit{end} of every step
-on every level, we check whether the stability criterion has been violated.
-If so, we recognize that it is not too late to solve the problem -- we can
-still add levels of refinement in those places. We then perform the new level
-generation to cover specifically those places that violated the stability criterion.
-To ensure consistency with our algorithm, we use the \textit{old}-time values
-of the coarse fluid state to interpolate into the new fine level, even
-though it is the \textit{new}-time values that were used to guide the level
-generation. Then our algorithm evolves the timestep on the new level as normal.
-Because our algorithm is recursive, this can continue to generate
-as many new levels as are needed to satisfy the stability criterion.
-The upshot is that as long as the simulation can afford enough levels to
-satisfy the stability criterion, there never needs to be even a single timestep
-that violates the criterion.
-
-%==========================================================================
-% Standard Results
-%==========================================================================
-\section{Collision Burning}
-\label{sec:collisionburning}
-
-% Discuss the results for the case of the uniform grid with 256 zones along
-% the $z$ axis, yielding a spatial resolution of 312.5 km, with no timestep limiting.
-% Discuss the effect of the burning mode here.
-
-In \autoref{sec:burner} we observed that there is an alternative to the traditional
-self-heating approach in a burn, the ``suppressed'' burning mode. This alternative
-is relevant to a collision problem because, as pointed out by \cite{raskin:2010}
-and \cite{kushnir:2013}, it is possible for a typical zone in our simulation to
-release a very large amount of energy in a burn before it is cooled
-off by a subsequent hydrodynamic expansion step, leading to the numerically unstable burning
-problem described in \autoref{sec:unstable_burning}. However, the suppressed burning mode
-(modeled after \citet{kushnir:2013}) produces only about
-$2 \times 10^{-4}\ \msolar$ of $^{56}$Ni. In a suppressed burn, we ensure that at every evaluation
-of the right-hand-side vector for the nuclear network integration, all quantities are scaled by the
-same factor, with this factor ensuring that the energy release is not large enough to permit a
-spurious detonation. This seems to be equivalent to the method used in \citet{kushnir:2013}. We
-reproduce their claim that this suppresses the detonation until later in the collision, after more
-material has approached the stalled shock. We also observe that, for a given resolution, the simultaneous
-detonations occur slightly outside the initial contact point, near the edges of the stalled shock. They
-establish as the main source of error in simulations that produce too little nickel. However, we find
-that the resulting detonations are not strong enough to convert a significant amount of material to NSE
-conditions; instead, only QSE conditions are reached near the center of the collision, leaving a significant
-amount of silicon and sulfur on the domain that is not further processed. In order to ensure that the
-very different results we obtained were not a result of the specific choice we made for how to suppress
-the burn, we tried varying the approach discussed in \autoref{sec:burner}. First, we limited based on
-the current value of the internal energy and sound speed rather than the fixed, initial value during a burn
-(this should be a less severe limiter). This did not yield a significant difference in the outcome. Second,
-we also tried applying the limiting after the burn had completed, using the differences between the input
-and output states of the integration, rather than directly limiting the RHS in the ODE. The behavior of the
-burn was qualitatively different, in that nickel production began much earlier in the collision when the limiting
-was applied this way. Nevertheless, the qualitative final outcome of the collision -- only trace amounts
-of nickel produced in total -- was still unchanged. With this variant too, it did not matter whether the limiting
-factor was determined based on the initial value of the state or the final value of the state.
-Our default resolution is significantly lower than that of \citeauthor{kushnir:2013} (the text
-says they used resolutions up to a few km) but we have tried runs at higher resolution and
-the results are qualitatively similar. We did not directly compare the results of the method at
-the same resolution as theirs, but our results indicate at least that this limiter significantly
-modifies the evolution for modest resolution simulations. The limiter does delay the time at which
-the detonation occurs, as expected by \citeauthor{kushnir:2013}, but the conditions for generating
-significant amounts of iron-group elements do not develop. It is not clear what the relevant difference
-is between our simulation and that of \citeauthor{kushnir:2013}, but as the suppressed burn limiting
-used here is ultimately unphysical, we choose not to use it in any other simulation.
+The upshot of this discussion is that the only way to resolve the
+question is with higher resolution simulations. The semi-analytic
+criterion provided by \cite{garg:2017} indicates that typical critical
+radii for detonation of equal C/O material are in the range 1-10 km.
+So, as pointed out by \citeauthor{kushnir:2013}, this numerical
+stability issue is mainly a concern for simulations with
+resolution of O(10 km) or worse. Detonations observed at higher
+resolution are physically plausible because the critical temperature
+gradient will be resolved. We have developed an AMR criterion
+to help achieve this resolution: we tag on regions suggested by
+the burning stability criterion, \autoref{eq:burning_limiter_2}.
+The original motivation for this choice was that by lowering
+$\Delta x$ and thus the sound-crossing time, we could reach
+the point where the detonation was occurring in a numerically stable
+region. This goal was not realized due to the reasons discussed above.
+Instead, we simply use it as a guide for where to refine. We tag
+for refinement all zones with $10^{-6} \leq f_{s} \leq 1$, where
+$f_{s}$ is the ratio of the sound crossing time to the energy injection
+time. We do not target zones for refinement with $f_{s} > 1$ since
+at that point a detonation is locked in. We refine up to a specified
+maximum resolution ratio relative to the coarse grid.
 
@@
 resolutions, augmented with additional runs with AMR. In each case we ran the
-simulation to the point at which unstable burning was approached. Then we stopped
+simulation to the point at which unstable burning was approached:
+$\tau_{s} / \tau_{e} = 10^{-6}$. Then we stopped
 the run and created a checkpoint. Starting from that checkpoint, we did a simulation

### 2017-11-14 papers/wdmerger_II/paper.tex
Update the resolution effect of the detonation time

@@
 anchoring points in time when significant events occur. We (arbitrarily) define the
-collision time, $t_c$, as the first time when the maximum density on the grid reaches
+collision time, $t_{\rm c}$, as the first time when the maximum density on the grid reaches
 $5 \times 10^6\ \text{g / cm}^3$. This is slightly larger than the initial maximum
 density of the two WDs, and is reached only when the WDs meet at the center of the
-domain. We define the detonation time, $t_d$, as the time when the temperature
+domain. We define the detonation time, $t_{\rm d}$, as the time when the temperature
 first reaches $4 \times 10^9\ \text{K}$.
@@
 choice in earlier papers on white dwarf mergers and collisions. No adaptive mesh
-refinement is used. For this configuration, the collision time is $t_c = 6.47\ \text{s}$,
-and the detonation time is $t_d = 7.43\ \text{s}$. The peak density at the detonation time
+refinement is used. For this configuration, the collision time is $t_{\rm c} = 6.47\ \text{s}$,
+and the detonation time is $t_{\rm d} = 7.43\ \text{s}$. The peak density at the detonation time
 is $2.3 \times 10^7\ \text{g / cm}^3$. The detonation occurs at the center of the domain
@@
 
-The collision time is about $t_{\rm c} = 6.85\ \text{s}$
-at high resolution. At low resolution, the simulation is slightly different because
-of the distortion imposed by a rectangular grid representing initially spherical stars,
-so the collision time is slightly later. Then, we restarted the run from this checkpoint
-and reset the time to zero (such that $t_{\rm c} = 0$). The new simulation runs until about
-$t = 2.15\ \text{s}$, so that the full 9 seconds are simulated. During this second phase
-the detonation occurs. An estimate of the detonation time, $t_{\rm d}$, that serves well
-is the time when the peak density on the grid first reaches $5 \times 10^9\ \text{K}$.
-We measure the detonation time relative to $t_{\rm c}$. Approximately speaking, the later
-$t_{\rm d}$ is, the more nickel is produced. As time passes, more white dwarf material
-slams into the center, compressing and heating up the gas. Both effects are favorable
-to production of nickel. So, a later $t_{\rm d}$ implies that more material can be converted
-to nickel. Similarly, the speed of the detonation matters. A slower speed detonation prolongs
-the collision event, allowing more gas to be processed into nickel.
+At high resolution the collision time is $t_{\rm c} = 6.41\ \text{s}$, which is similar to the
+collision time at low resolution. However, the detonation time is around $t_{\rm d} = 6.62\ \text{s}$,
+which is significantly earlier than at low resolution due to the high resolution simulations
+being better able to resolve the point of contact between the WDs. The peak density crosses
+$10^7\ \text{g / cm}^3$ significantly earlier, and this is what triggers the burning sufficient
+to generate an earlier detonation. Approximately speaking, the later $t_{\rm d}$ is, the more
+nickel is produced. As time passes, more white dwarf material slams into the center, compressing
+and heating up the gas. Both effects are favorable to production of nickel. So, a later $t_{\rm d}$
+implies that more material can be converted to nickel.
 

### 2017-11-14 papers/wdmerger_II/paper.tex
Move the timestep limiter discussion into the methodology section

@@
 
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
+  \caption{Nickel production as a function of the timestep factors $f_{be}$
+           (solid blue, limiting the timestep on changes in internal energy) and
+           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
+           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
+           the timestep by at least an order of magnitude more.
+           \label{fig:timestep_factor}}
+\end{figure}
+\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
+production for the baseline collision case. These limiters are somewhat redundant in
+the sense that a given $f_{bX}$ roughly corresponds to another (higher) $f_{be}$, as
+is evident by the fact that on the plot the curves look qualitatively similar in
+shape but with a horizontal offset. As either $f$ becomes small, the restriction on
+the timestep becomes more severe: for $f_{bX} = 0.1$, the full simulation requires
+nearly 30,000 timesteps, compared to just over 1,000 for the baseline case.
+Such a simulation would be quite expensive for a high resolution 3D calculation.
+It is clear, though, that even though the simulation is not converged when the
+timestep is not limited due to burning, the effect is small in absolute terms.
+So we agree that the use of values in this range by \cite{raskin:2010},
+\cite{hawley:2012}, and \cite{kushnir:2013} is appropriate. However, for reasons of
+computational expense we do not use the timestep limiter in the high resolution
+simulations that follow. The error due to this choice should be small, because
+the effect is not that large to begin with, and because the higher resolution
+decreases the timestep anyway due to the hydrodynamic stability constraint.
+
 Other approaches to the coupling 
@@
 %==========================================================================
-% Temporal Resolution
-%==========================================================================
-\section{Temporal Resolution}
-\label{sec:temporalresolution}
-
-For the standard self-heating burns that we use, the timestep limiting scheme
-described in \autoref{sec:hydrocoupling} has a significant effect on the evolution.
-The parameters $f_{be}$ (from \autoref{eq:timestep_e}) and $f_{bX}$ (from
-\autoref{eq:timestep_X}) control the size of the timestep as a function of the
-burning rate. As the timestep gets smaller, the error due to the Strang splitting
-scheme coupling the reactions and the hydro also gets smaller, and we expect
-that the results become more accurate.
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
-  \caption{Nickel production as a function of the timestep factors $f_{be}$
-           (solid blue, limiting the timestep on changes in internal energy) and
-           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
-           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
-           the timestep by at least an order of magnitude more.
-           \label{fig:timestep_factor}}
-\end{figure}
-\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
-production (these limiters are somewhat redundant in the sense that a given $f_{bX}$
-roughly corresponds to another (higher) $f_{be}$, as is evident by the fact that on
-the graph the curves look qualitatively similar in shape but with a horizontal offset.
-As either $f$ becomes small, the restriction on the timestep becomes quite severe:
-for $f_{be} = 0.001$, the minimum timestep in the simulation is smaller than
-$5 \times 10^{-9}\ \text{s}$ and the full simulation requires over 100,000 timesteps.
-Such a simulation would be prohibitively expensive for a high resolution 3D calculation.
-It is clear, though, that the timestep effect has converged when $f_{be} \sim 0.1$. We
-agree that the use of values in this range by \cite{raskin:2010}, \cite{hawley:2012},
-and \cite{kushnir:2013} is appropriate.
-
-%==========================================================================
 % Resolution Dependence

### 2017-12-05 papers/wdmerger_II/paper.tex
Add a clarifying comment about ending simulations early

@@
 We terminate the simulation at $t = 9$ seconds, as we have found that for this
-particular setup, all of the significant burning has completed by this time. At
+particular setup, all of the significant burning has completed by this time\footnote{In
+some simulations at high resolution we were mainly interested in seeing when the
+detonation happened, and we terminated the simulation at that point to save
+computational cycles. We show nucleosynthesis results only for those simulations
+that had completed the nuclear burning.}. At
 this point, the total energy on the domain has become positive, typically about

### 2017-12-05 papers/wdmerger_II/paper.tex
Update the burning stability comments

@@
 similar results, for example because even at the higher resolution, the physical
-size of the hotspot stays the same. One cannot appeal to the numerical instability
-criterion alone to make a decision in this context.
+size of the hotspot stays the same. Additionally, one might think that simply
+adding adaptive mesh refinement in the burning region could help, because then
+$\Delta x$ decreases, so the ratio of $\tau_s$ to $\tau_e$ should decrease.
+We developed an AMR criterion to do just this: we tag on regions suggested by
+the burning stability criterion, \autoref{eq:burning_limiter_2}. But
+if a low resolution simulation develops a very large hotspot that violates the
+stability criterion, adding refinement inside this region will likely not help,
+because the energy trapped in the center of the region may not be able to escape
+before ignition occurs. So for low resolution simulations, even adding a significant
+amount of refinement (say, refining by a factor of 64 or 128) did not change the
+propagation of a detonation.
 
-Furthermore, we find that \textit{all} of the detonations we have seen at low
-and moderate resolution inevitably violate the stability criterion. This is true
-even if the prompt detonation is suppressed: the delayed detonation also is
-numerically unstable. For comparison purposes, we developed an option for our
-burner to use a ``suppressed'' burning mode. In a suppressed burn, we limit
-the changes to the state so that \autoref{eq:burning_limiter_2}
+For the above reasons, an appeal to the numerical instability criterion alone is
+insufficient to understand whether a given ignition is real. We find that \textit{all}
+of the detonations we have seen at low and moderate resolution inevitably violate
+the stability criterion. This is true even if the prompt detonation is suppressed:
+the delayed detonation also is numerically unstable. For comparison purposes, we
+developed an option for our burner to use a ``suppressed'' burning mode. In a
+suppressed burn, we limit the changes to the state so that \autoref{eq:burning_limiter_2}
 is always satisfied\footnote{To achieve this we directly multiply the
@@
 resolution are physically plausible because the critical temperature
-gradient will be resolved. We have developed an AMR criterion
-to help achieve this resolution: we tag on regions suggested by
-the burning stability criterion, \autoref{eq:burning_limiter_2}.
-The original motivation for this choice was that by lowering
-$\Delta x$ and thus the sound-crossing time, we could reach
-the point where the detonation was occurring in a numerically stable
-region. This goal was not realized due to the reasons discussed above.
-Instead, we simply use it as a guide for where to refine. We tag
-for refinement all zones with $10^{-6} \leq f_{s} \leq 1$, where
-$f_{s}$ is the ratio of the sound crossing time to the energy injection
-time. We do not target zones for refinement with $f_{s} > 1$ since
-at that point a detonation is locked in. We refine up to a specified
-maximum resolution ratio relative to the coarse grid.
+gradient will be resolved. The high resolution needs to be achieved
+with some combination of high resolution of the collision process
+itself, and perhaps AMR based on the stability criterion as described
+above, where we tag for refinement all zones above some $f_{s}$, the
+ratio of the sound crossing time to the energy injection time. It
+cannot be solely based on the burning stability criterion for the reason
+described above, that a detonation may be spuriously locked in by
+the low resolution regardless of how much resolution we add to it.
 

### 2017-12-06 papers/wdmerger_II/paper.tex
Add cite to Seitenzahl 2009

@@
 question is with higher resolution simulations. The semi-analytic
-criterion provided by \cite{garg:2017} indicates that typical critical
-radii for detonation of equal C/O material are in the range 1-10 km.
+criterion provided by \cite{garg:2017}, and the numerical calculations
+of \cite{seitenzahl:2009}, give critical radii for spontaneous detonation.
+For the densities we are interested in ($\sim 10^7\ \text{g / cm}^3$),
+assuming equal C/O material, the critical radii are in the range 1-10 km.
 So, as pointed out by \citeauthor{kushnir:2013}, this numerical

### 2018-01-21 papers/wdmerger_II/paper.tex
Another stab at the abstract

@@
 \begin{abstract}
-Simulations of white dwarf mergers and collisions that attempt to answer the
-question of whether these events are progenitors of Type Ia supernovae must
+Simulations of white dwarf mergers that include thermonuclear reactions must
 confront the question of how nuclear detonations form and whether observed
@@
 detonation that disrupts both stars, possibly causing a Type Ia supernova.
-While this story is intuitively plausible, demonstrating this in a reacting
+While this story is intuitively plasible, demonstrating this in a reacting
 hydrodynamics simulation is challenging due to the rapid evolution of the
 nuclear burning, posing numerical challenges related to both spatial and
-temporal resolution. In this study we explore a number of the algorithms
-that affect the outcome of reacting hydrodynamics simulations that lead
-to explosive burning and address the question of what can be done to ensure
-that these simulations are properly converged. We find that achieving a
-converged simulation in the code \castro\ is very difficult for a standard
-approach with uniform resolution. The typical timesteps taken by hydrodynamics
-codes are too large and the spatial resolution is too low, while increasing
+temporal resolution. In this study we explore a number of the algorithmic choices
+that affect the outcome of reacting hydrodynamics simulations, and address
+the question of what is required to ensure that these simulations are properly
+converged. We find that achieving a converged white dwarf collision simulation
+in the code \castro\ is very difficult. The typical timesteps taken by hydrodynamics
+codes are too large, and the achievable spatial resolution is too low, while increasing
 either to the point required to actually obtain a converged simulation is
-prohibitive. However, we have identified a criterion for adaptive mesh refinement
-that substantially improves the situation: by targeting for extra sptial resolution
-only those regions that are likely to undergo numerically unstable burning,
-we can significantly reduce the effects of numerically spurious detonations.
-This result is a possible path forward for obtaining reliable numerical
-simulations of Type Ia supernova involving thermonuclear detonations.
+prohibitive. We have identified a criterion for adaptive mesh refinement
+that covers the most important regions in the simulation and allows us to
+achieve very high resolution in the burning region. However, even in our
+highest resolution simulations, we do not find a converged result, and
+detonations seem to occur stochastically. Our results point to a path forward
+for astrophysical simulations with thermonuclear detonations, but for now
+they encourage severe caution in interpreting observed detonations in previous
+and current simulations.
 \end{abstract}

### 2018-01-22 papers/wdmerger_II/paper.tex
Redefine the collision time

@@
 generation is occurring. In our discussions below, it helps to have some other
-anchoring points in time when significant events occur. We (arbitrarily) define the
-collision time, $t_{\rm c}$, as the first time when the maximum density on the grid reaches
-$5 \times 10^6\ \text{g / cm}^3$. This is slightly larger than the initial maximum
-density of the two WDs, and is reached only when the WDs meet at the center of the
-domain. We define the detonation time, $t_{\rm d}$, as the time when the temperature
-first reaches $4 \times 10^9\ \text{K}$.
+anchoring points in time when significant events occur. We arbitrarily define the
+collision time, $t_{\rm c}$, as the time when the density at the center of the grid
+first reaches $10^2\ \text{g / cm}^3$. This is reached only when the WDs meet at the
+center of the domain. We define the detonation time, $t_{\rm d}$, as the time when the
+temperature first reaches $4.5 \times 10^9\ \text{K}$. This is also an arbitrary
+threshold, but we have found empirically that this temperature is only reached in
+the collision when a detonation is formed.
 
@@
 choice in earlier papers on white dwarf mergers and collisions. No adaptive mesh
-refinement is used. For this configuration, the collision time is $t_{\rm c} = 6.47\ \text{s}$,
+refinement is used. For this configuration, the collision time is approximately $t_{\rm c} = 4.6\ \text{s}$,
 and the detonation time is $t_{\rm d} = 7.43\ \text{s}$. The peak density at the detonation time
@@
 
-At high resolution the collision time is $t_{\rm c} = 6.41\ \text{s}$, which is similar to the
+At high resolution the collision time, $t_{\rm c} = 4.6\ \text{s}$, is similar to the
 collision time at low resolution. However, the detonation time is around $t_{\rm d} = 6.62\ \text{s}$,

### 2018-01-22 papers/wdmerger_II/paper.tex
Clarify a footnote

@@
 particular setup, all of the significant burning has completed by this time\footnote{In
-some simulations at high resolution we were mainly interested in seeing when the
-detonation happened, and we terminated the simulation at that point to save
-computational cycles. We show nucleosynthesis results only for those simulations
-that had completed the nuclear burning.}. At
+some simulations we were mainly interested in seeing when the detonation happened, and
+we terminated the simulation at that point. We show nucleosynthesis results only for
+those simulations that had completed nuclear burning.}. At
 this point, the total energy on the domain has become positive, typically about

### 2018-01-22 papers/wdmerger_II/paper.tex
Clean up formatting on density units

@@
 collision time, $t_{\rm c}$, as the time when the density at the center of the grid
-first reaches $10^2\ \text{g / cm}^3$. This is reached only when the WDs meet at the
+first reaches $10^2\ \text{g\,/\,cm}^3$. This is reached only when the WDs meet at the
 center of the domain. We define the detonation time, $t_{\rm d}$, as the time when the
@@
 As in Paper I, the region outside the WDs is filled with a low-density ambient
-material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition and temperature is the
+material with $\rho = 10^{-4}\ \text{g\,/\,cm}^{3}$ whose composition and temperature is the
 same as that of the WDs. Reactions in this region are unimportant, so for
 computational efficiency, we disable all nuclear burning for zones that have
-$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
+$\rho < 10^6\ \text{g\,/\,cm}^{3}$ and $T < 10^8\ \text{K}$. We have confirmed that
 the results do not depend on this choice. A sponge is applied to all gas of density
-lower than $1\ \text{g / cm}^{-3}$. The sponge damps out the velocity in these zones
+lower than $1\ \text{g\,/\,cm}^{3}$. The sponge damps out the velocity in these zones
 (see Paper I) to prevent spurious numerical error in low-density gas from dragging
 down the global timestep. The density floor for the simulation is set
-at $10^{-5}\ \text{g / cm}^{-3}$ and the temperature floor is $10^7$ K.
+at $10^{-5}\ \text{g\,/\,cm}^{3}$ and the temperature floor is $10^7$ K.
 
@@
 and the detonation time is $t_{\rm d} = 7.43\ \text{s}$. The peak density at the detonation time
-is $2.3 \times 10^7\ \text{g / cm}^3$. The detonation occurs at the center of the domain
+is $2.3 \times 10^7\ \text{g\,/\,cm}^3$. The detonation occurs at the center of the domain
 where the two WDs initially meet.
@@
 of \cite{seitenzahl:2009}, give critical radii for spontaneous detonation.
-For the densities we are interested in ($\sim 10^7\ \text{g / cm}^3$),
+For the densities we are interested in ($\sim 10^7\ \text{g\,/\,cm}^3$),
 assuming equal C/O material, the critical radii are in the range 1-10 km.
@@
 being better able to resolve the point of contact between the WDs. The peak density crosses
-$10^7\ \text{g / cm}^3$ significantly earlier, and this is what triggers the burning sufficient
+$10^7\ \text{g\,/\,cm}^3$ significantly earlier, and this is what triggers the burning sufficient
 to generate an earlier detonation. Approximately speaking, the later $t_{\rm d}$ is, the more

### 2018-01-22 papers/wdmerger_II/paper.tex
Remove an unnecessary paper I reference

@@
 
-As in Paper I, the region outside the WDs is filled with a low-density ambient
+The region outside the WDs is filled with a low-density ambient
 material with $\rho = 10^{-4}\ \text{g\,/\,cm}^{3}$ whose composition and temperature is the

### 2018-01-22 papers/wdmerger_II/paper.tex
Clean up formatting

@@
 
-White dwarfs are mainly composed of $\alpha$-chain particles, primarily ${}^4$He,
-${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, and ${}^{24}$ Mg. Therefore the core of
+White dwarfs are primarily composed of the $\alpha$-chain particles ${}^4$He,
+${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, and ${}^{24}$Mg. Therefore the core of
 any network appropriate for modeling nuclear burning in white dwarfs will be
@@
 ported into a form that is consistent with the \amrex\ codes, in the freely available
-\microphysics\ code repository
-\footnote{\microphysics\ can be obtained at \url{https://github.com/starkiller-astro/Microphysics}.},
+\microphysics\ code repository\footnote{\microphysics\ can be obtained at
+\url{https://github.com/starkiller-astro/Microphysics}.},
 a collection of microphysical routines that are designed to be used in our
@@
 $f_{bX}$ has not been used by others performing these types of calculations. The sensitivity
-of results to the value of these timestep limiters will be discussed in 
-\autoref{sec:temporalresolution}. The factors $f_{be}$ and $f_{bX}$ can be set at runtime in \castro.
+of results to the value of these timestep limiters will be discussed later in the
+results. The factors $f_{be}$ and $f_{bX}$ can be set at runtime in \castro.
 

### 2018-01-22 papers/wdmerger_II/paper.tex
Rework the problem setup section

@@
 %==========================================================================
-\section{White Dwarf Collisions}
+\section{White Dwarf Collision Setup}
 \label{sec:collisions}
-
-We start with a description of the problem setup and results of a typical collision
-at low resolution. These results set the stage for our numerical methodology. This enables
-a concrete discussion about the methodology using the language of how it affects the
-collision outcome. In this setting, a concrete discussion is more fruitful than an abstract one.
-
-\subsection{Problem Setup}
-\label{sec:problemsetup}
 
@@
 
-Our primary metric for this test is the amount of $^{56}$Ni generated in the collision,
-as this is the parameter most directly related to the observable quantities of interest
-for Type Ia supernovae. We collect information at the end of every timestep about the
-total amount of nickel on the grid (in solar masses), and we use the maximum value of
-this nickel mass. Plotfiles of relevant quantities are created every 0.01 seconds.
+We collect information at the end of every timestep about the total amount of
+all of the species on the grid (in solar masses), and we use the mass of $^{56}$Ni
+generated as a metric for this test, as this is the parameter most directly related
+to the observable quantities of interest for Type Ia supernovae. Since, as we will see
+below, not all of our simulations detonate at a density high enough to generate substantial
+amounts of nickel, we employ the mass-weighted average atomic mass number on the domain,
+$\bar{A}$, as a more general metric describing the outcome of the simulation. A
+final relevant metric is the total amount of energy $E$ released by the detonation.
+
+Plotfiles of relevant quantities are created every 0.01 seconds.
 The timesteps are limited to ensure that every multiple of this plotfile interval is
-reached exactly, which means an effective maximum timestep of 0.01 seconds.
+reached exactly, which means an effective maximum timestep of 0.01 seconds. Timesteps
+are by default limited only by the hydrodynamic stability constraint and this plotfile
+interval constraint, except where noted otherwise.
 
-\subsection{Baseline Results}
-\label{sec:baseline_results}
-
-Now let us examine the results for a collision of two 0.64 $\msolar$ WDs. We use
-128 zones in the radial dimension, corresponding to a spatial resolution of 312.5 km.
-Timesteps are limited only by the hydrodynamic stability constraint (and the plotfile
-interval constraint). Resolutions in the range of a few hundred km have been a common
-choice in earlier papers on white dwarf mergers and collisions. No adaptive mesh
-refinement is used. For this configuration, the collision time is approximately $t_{\rm c} = 4.6\ \text{s}$,
-and the detonation time is $t_{\rm d} = 7.43\ \text{s}$. The peak density at the detonation time
-is $2.3 \times 10^7\ \text{g\,/\,cm}^3$. The detonation occurs at the center of the domain
-where the two WDs initially meet.
+For a collision of two 0.64 $\msolar$ WDs with 128 zones in the radial dimension,
+corresponding to a uniform spatial resolution of 312.5 km, the collision time is
+$t_{\rm c} = 4.16\ \text{s}$, and the detonation time is $t_{\rm d} = 7.43\ \text{s}$.
+The peak density at the detonation time is $2.3 \times 10^7\ \text{g\,/\,cm}^3$.
+Approximately $0.21\ \msolar$ of nickel and $0.46\ \msolar$ of silicon is created,
+corresponding to an $\bar{A}$ of about 29. Approximately $1.3 \times 10^{51}$
+ergs of energy are released. Simulations with resolutions in this few hundred kilometer
+range have been a common choice in earlier papers on white dwarf mergers and collisions
+due to computational constraints; our goal is to see how higher resolution runs differ.
 
@@
 
-At high resolution the collision time, $t_{\rm c} = 4.6\ \text{s}$, is similar to the
-collision time at low resolution. However, the detonation time is around $t_{\rm d} = 6.62\ \text{s}$,
-which is significantly earlier than at low resolution due to the high resolution simulations
-being better able to resolve the point of contact between the WDs. The peak density crosses
-$10^7\ \text{g\,/\,cm}^3$ significantly earlier, and this is what triggers the burning sufficient
-to generate an earlier detonation. Approximately speaking, the later $t_{\rm d}$ is, the more
+At high resolution the collision time, $t_{\rm c} = 4.6\ \text{s}$, is later than the
+collision time at low resolution, due to a sharper view of the contact point. However,
+for the same reason, the detonation time is significantly earlier at $t_{\rm d} = 6.62\ \text{s}$,
+as the peak density reaches the key density range of $10^7\ \text{g\,/\,cm}^3$ significantly earlier,
+and this is what triggers the burning sufficient to generate an earlier detonation.
+Approximately speaking, the later $t_{\rm d}$ is, the more
 nickel is produced. As time passes, more white dwarf material slams into the center, compressing

### 2018-01-23 papers/wdmerger_II/paper.tex
Simplify the nuclear network section

@@
 
-White dwarfs are primarily composed of the $\alpha$-chain particles ${}^4$He,
-${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, and ${}^{24}$Mg. Therefore the core of
-any network appropriate for modeling nuclear burning in white dwarfs will be
-these alpha chain nuclides, with the idea being that links up the $\alpha$-chain
-will eventually get us to ${}^{56}$Ni, the nuclide responsible for the
-energy output of Type Ia supernovae. In this paper we use the \aproxthirteen\
-network\ \citep{timmes:1999,timmes:2000}. This includes all thirteen $\alpha$-chain
-particles between ${}^4$He and ${}^{56}$Ni. This network was used by \citet{hawley:2012}
+In this paper we use the \aproxthirteen\ network\ \citep{timmes:1999,timmes:2000}.
+This includes all thirteen $\alpha$-chain particles between ${}^4$He and ${}^{56}$Ni.
+These $\alpha$-chain particles capture the most important pathway connecting a WD
+and its thermonuclear detonation. This network was used by \citet{hawley:2012}
 and \citet{raskin:2010}. \citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a
-very similar network that additionally included ${}^{60}$Zn. This network has been
-ported into a form that is consistent with the \amrex\ codes, in the freely available
+very similar network that additionally included ${}^{60}$Zn. The \aproxthirteen\ network has been
+ported into a form that is consistent with the \amrex\ codes in the freely available
 \microphysics\ code repository\footnote{\microphysics\ can be obtained at
@@
 the answer). We also make available \isoseven\ \citep{timmes:2000}, which skips the nuclides between
-silicon and nickel, and was used by \citet{rosswog:2009}. \isoseven\ gives significantly different
-quantitative results than \aproxthirteen\ for this problem, due to the inapplicable assumption of
-an equilibrium abundance of the unused isotopes, but the results are still qualitatively similar.
+silicon and nickel, and which was used by \citet{rosswog:2009}. \isoseven\ gives significantly different
+quantitative results than \aproxthirteen\ for this problem, due to its inapplicable assumption of
+an equilibrium abundance of the unsimulated isotopes, but the results are still qualitatively similar.
 
@@
 effects it has on the outcome of the simulation. The suppressed burn has separate
-issues which we defer to \autoref{sec:unstable_burning}
+issues which we defer to \autoref{sec:unstable_burning}.
 

### 2018-01-23 papers/wdmerger_II/paper.tex
Clean up numerical instability section

@@
 for the baseline case, the sound-crossing time is about five times longer
-than the energy injection time. This means that a detonation is guaranteed
+than the energy injection time. This means that an ignition is guaranteed
 to occur for numerical reasons. \citeauthor{kushnir:2013} thus argue that
@@
 high temperature, say $5 \times 10^9\ \text{K}$, with an arbitrarily large size,
-say a cube with side length 100 km. This region will very likely detonate,
+say a cube with side length 100 km. This region will very likely ignite,
 even if it is surrounded by much cooler material. By the time the material on
@@
 
-It is true that when our resolution is low enough, there is a floor on the size
-of a hotspot, possibly making such a detonation more likely. However, this is an
-unavoidable consequence of the low resolution: it is the correct result of the
-simulation that was performed. If the results do not match what occurs at higher
+When the resolution is low enough, there is a floor on the size of a hotspot,
+possibly making such a detonation more likely. This is an unavoidable consequence
+of the low resolution; yet, it may be the correct result of the simulation that
+was performed. That is, even if large hotspots are unphysical because in reality
+the temperature distribution would be smoother, if such a large hotspot \textit{were}
+to develop (which is the implicit assumption of a low resolution simulation), then
+it would likely detonate. If the results do not match what occurs at higher
 resolution, then the simulation is not converged and the results are not reliable.
@@
 We developed an AMR criterion to do just this: we tag on regions suggested by
-the burning stability criterion, \autoref{eq:burning_limiter_2}. But
-if a low resolution simulation develops a very large hotspot that violates the
-stability criterion, adding refinement inside this region will likely not help,
-because the energy trapped in the center of the region may not be able to escape
-before ignition occurs. So for low resolution simulations, even adding a significant
-amount of refinement (say, refining by a factor of 64 or 128) did not change the
-propagation of a detonation.
+the burning stability criterion, \autoref{eq:burning_limiter_2}. But if a low
+resolution simulation develops a very large hotspot that violates the stability
+criterion, adding refinement inside this region will likely not help, because the
+energy trapped in the center of the region may not be able to escape before ignition
+occurs. So for the simulations with resolution inside the stars that was coarser than
+10 km, even adding a significant amount of refinement (say, refining by a factor of
+64 or 128) did not change the propagation of a detonation.
 
@@
 insufficient to understand whether a given ignition is real. We find that \textit{all}
-of the detonations we have seen at low and moderate resolution inevitably violate
-the stability criterion. This is true even if the prompt detonation is suppressed:
+of the detonations we have seen at any resolution violate the stability criterion,
+even for the finest resolution simulation we performed (150 m, inside the burning region).
+This is true even if the prompt detonation is suppressed:
 the delayed detonation also is numerically unstable. For comparison purposes, we
@@
 assuming equal C/O material, the critical radii are in the range 1-10 km.
-So, as pointed out by \citeauthor{kushnir:2013}, this numerical
-stability issue is mainly a concern for simulations with
-resolution of O(10 km) or worse. Detonations observed at higher
-resolution are physically plausible because the critical temperature
-gradient will be resolved. The high resolution needs to be achieved
-with some combination of high resolution of the collision process
-itself, and perhaps AMR based on the stability criterion as described
+So this suggests that our simulation needs to do significantly better than
+this if we are likely to make headway on this issue. Detonations observed
+at higher resolution may be physically plausible because the critical
+temperature gradient can be resolved. The high resolution needs to be
+achieved with some combination of high resolution of the collision process
+itself, and AMR based on the stability criterion as described
 above, where we tag for refinement all zones above some $f_{s}$, the
-ratio of the sound crossing time to the energy injection time. It
-cannot be solely based on the burning stability criterion for the reason
-described above, that a detonation may be spuriously locked in by
-the low resolution regardless of how much resolution we add to it.
+ratio of the sound crossing time to the energy injection time. We
+cannot solely rely on a low resolution run with AMR based on the
+burning stability criterion for the reason described above, that a
+detonation may be spuriously locked in by the low resolution regardless
+of how much resolution we can practicably add to it.
 

### 2018-01-28 papers/wdmerger_II/paper.tex
Add release information
Remove specificity about the time of the restart

@@
 
-We implement the white dwarf collision problem in \castro\ in the \wdmerger\ problem
+We implement the white dwarf collision problem in \castro\footnote{The main results
+in this paper were generated using the 17.12 releases of \castro, \amrex, and
+\microphysics.} in the \wdmerger\ problem
 directory. We prepare the domain in the way most previous papers have done so:
@@
 resolutions, augmented with additional runs with AMR. In each case we ran the
-simulation to the point at which unstable burning was approached:
-$\tau_{s} / \tau_{e} = 10^{-6}$. Then we stopped
+simulation to a point just after the burning started, but before a detonation
+had occurred. Then we stopped
 the run and created a checkpoint. Starting from that checkpoint, we did a simulation

### 2018-07-10 papers/ignition/paper.tex
Update title, author list

--- 2018-01-28 papers/wdmerger_II/paper.tex
@@
 %==========================================================================
-\title{White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions}
+\title{Numerical Stability of Detonations in White Dwarf Simulations}
 
-\shorttitle{WD Mergers. II. Nuclear Reactions}
-\shortauthors{Katz et al. (2017)}
+\shorttitle{Numerical Detonations}
+\shortauthors{Katz et al. (2018)}
 
-\author{Max P. Katz\altaffilmark{1}}
-\author{Michael Zingale\altaffilmark{1}}
-\author{Alan C. Calder\altaffilmark{1,2}}
-\author{F. Douglas Swesty\altaffilmark{1}}
-\author{Ann S. Almgren\altaffilmark{3}}
-\author{Weiqun Zhang\altaffilmark{3}}
-\author{Frank X. Timmes\altaffilmark{4,5}}
+\author{Max P. Katz\altaffilmark{1,2}}
+\author{Michael Zingale\altaffilmark{2}}
 
@@
 {
-  Department of Physics and Astronomy,
-  Stony Brook University, Stony Brook, NY, 11794-3800, USA
+  NVIDIA Corporation
+  2788 San Tomas Expressway
+  Santa Clara, CA, 95051, USA
 }
@@
 {
-  Institute for Advanced Computational Sciences,
-  Stony Brook University, Stony Brook, NY, 11794-5250, USA
-}
-
-\altaffiltext{3}
-{
-  Center for Computational Sciences and Engineering,
-  Lawrence Berkeley National Laboratory, Berkeley, CA 94720
-}
-
-\altaffiltext{4}
-{
-  The Joint Institute for Nuclear Astrophysics, USA
-}
-
-\altaffiltext{5}
-{
-  School of Earth and Space Exploration, Arizona State University, Tempe, AZ, USA
+  Department of Physics and Astronomy,
+  Stony Brook University, Stony Brook, NY, 11794-3800, USA
 }

### 2018-07-10 papers/ignition/paper.tex
Rename to aastex60
Update to aastex62

@@
-\documentclass[twocolumn,numberedappendix]{../aastex6}
+\documentclass[twocolumn,numberedappendix]{../aastex62}
 
@@
 
-\author{Max P. Katz\altaffilmark{1,2}}
-\author{Michael Zingale\altaffilmark{2}}
-
-\altaffiltext{1}
+\author{Max P. Katz}
+\affiliation
 {
-  NVIDIA Corporation
-  2788 San Tomas Expressway
-  Santa Clara, CA, 95051, USA
+  NVIDIA Corporation \\
+  2788 San Tomas Expressway, Santa Clara, CA, 95051, USA
+}
+\affiliation
+{
+  Department of Physics and Astronomy, Stony Brook University, Stony Brook, NY, 11794-3800, USA
 }
 
-\altaffiltext{2}
+\author{Michael Zingale}
+\affiliation
 {
-  Department of Physics and Astronomy,
-  Stony Brook University, Stony Brook, NY, 11794-3800, USA
+  Department of Physics and Astronomy, Stony Brook University, Stony Brook, NY, 11794-3800, USA
 }

### 2018-07-10 papers/ignition/paper.tex
Very rough first cut at the new abstract

@@
 \begin{abstract}
-Simulations of white dwarf mergers that include thermonuclear reactions must
-confront the question of how nuclear detonations form and whether observed
-detonations in a numerical simulation are realistic. We address this question
-by performing simulations of collisions of white dwarfs, a subset of white
-dwarf mergers that occur when the stars approach each other nearly head-on.
-These have the potential for explosive nucleosynthesis, leading to a thermonuclear
-detonation that disrupts both stars, possibly causing a Type Ia supernova.
-While this story is intuitively plasible, demonstrating this in a reacting
-hydrodynamics simulation is challenging due to the rapid evolution of the
-nuclear burning, posing numerical challenges related to both spatial and
-temporal resolution. In this study we explore a number of the algorithmic choices
-that affect the outcome of reacting hydrodynamics simulations, and address
+Some simulations of Type Ia supernovae feature thermonuclear detonations that
+arise naturally during the evolution. It is usually not clear whether these
+detonations occur in the same time and place as they would in a simulation
+that fully resolved the burning length scale (assuming the resolved simulation
+even features a detonation). In this study we explore a number of the algorithmic
+choices that affect the outcome of reacting hydrodynamics simulations, and address
 the question of what is required to ensure that these simulations are properly
-converged. We find that achieving a converged white dwarf collision simulation
-in the code \castro\ is very difficult. The typical timesteps taken by hydrodynamics
-codes are too large, and the achievable spatial resolution is too low, while increasing
-either to the point required to actually obtain a converged simulation is
-prohibitive. We have identified a criterion for adaptive mesh refinement
-that covers the most important regions in the simulation and allows us to
-achieve very high resolution in the burning region. However, even in our
-highest resolution simulations, we do not find a converged result, and
-detonations seem to occur stochastically. Our results point to a path forward
-for astrophysical simulations with thermonuclear detonations, but for now
-they encourage severe caution in interpreting observed detonations in previous
-and current simulations.
+converged. We focus on the spatial and temporal resolution, as well as the coupling
+between the hydrodynamics and the reactions. Adaptive mesh refinement, properly
+targeted at the relevant burning regions, works well for achieving converged simulations
+on a relatively coarse grid.
 \end{abstract}

### 2018-07-10 papers/ignition/paper.tex
Start reworking the introduction

@@
 
-Binary white dwarf (WD) systems are a promising progenitor candidate for Type
-Ia supernovae (SNe Ia) for observational reasons, but from the simulation
-perspective there is still significant uncertainty regarding how nuclear
-detonations form in an exploding white dwarf. There is uncertainty in the
-model, regarding whether detonations form on the surface of a star as
-accreting material heats up when it strikes the surface, or at or near the
-center of the star as the accreted material compresses the WD. There is also
-uncertainty in the numerics: given one of these models, can we self-consistently
-demonstrate that a detonation occurs? This question has haunted both single
-WD and double WD progenitor models, because the length and time scale at which
-a detonation forms is orders of magnitude smaller than the resolution that
+Thermonuclear detonations are common to all current likely models of Type Ia
+supernovae (SNe Ia), but how they are actually generated in progenitor systems
+is still an open question. Different models predict different locations for
+the detonation (e.g. in the center of a star or on the edge) and different
+mechanisms for initiating the event (e.g. a prompt detonation in a white dwarf
+collision versus a deflagration to detonation transition in the single degenerate
+model). Common to all of the cases is a severe lack of numerical resolution in
+the location where the detonation is expected to occur. The length and time scale
+at which a detonation forms is orders of magnitude smaller than the resolution that
 typical hydrodynamic simulations can achieve. Furthemore, the approximate equations
-we use at the resolution that we can achieve are still notoriously
-difficult to solve accurately. The mere presence of a detonation
-(or lack thereof) in a simulation is therefore only weak evidence regarding
-whether a detonation would truly occur.
+we use at the resolution that we can achieve are still notoriously difficult to
+solve accurately. The mere presence of a detonation (or lack thereof) in a simulation
+is therefore only weak evidence regarding whether a detonation would truly occur.
 

### 2018-07-12 papers/ignition/paper.tex
Rework introduction

@@
 the detonation (e.g. in the center of a star or on the edge) and different
-mechanisms for initiating the event (e.g. a prompt detonation in a white dwarf
+mechanisms for initiating the event (e.g. a prompt detonation in a white dwarf (WD)
 collision versus a deflagration to detonation transition in the single degenerate
@@
 In this study we examine the challenges associated with simulating nuclear detonations.
-This paper follows from our first WD merger methodology paper, \citet{wdmergerI}
-(hereafter, Paper I), where we deferred discussion of our nuclear reaction methodology.
-After presenting the burning methodology, we apply it to the problem of collisions of white dwarfs.
-WD collisions can occur in isolated binaries or in hierarchical triple systems, where the WD binary
-system in a tight inner orbit is gravitationally coupled to a third, outer star \citep{thompson:2011,hamers:2013}.
-Whether the rate of these collisions may be high enough to significantly contribute to
-the observed SN Ia rate is debated: \cite{katzdong:2012} suggest that this is plausible,
-but \cite{hamers:2013} and \cite{papish:2015} suggest that the rate is relatively low.
-Regardless, they are a useful vehicle for understanding the problem of nuclear burning.
-Head-on WD collisions rapidly convert a significant amount of kinetic energy into thermal
-energy in a small region and thus set up conditions ripe for a thermonuclear detonation;
-being able to correctly simulate these is in some sense a prerequisite to correct simulation
-of the more complicated general case of white dwarf mergers, where we do not initially know
-under what conditions the seed of a thermonuclear detonation is born.
+The inspiration for this work comes from the literature on head-on collisions of WDs,
+which can occur, for example, in certain triple star systems \citep{thompson:2011,hamers:2013}.
+This problem is useful for understanding the problem of numerical simulation of nuclear
+burning. Head-on WD collisions rapidly convert a significant amount of kinetic energy
+into thermal energy in a small region and thus set up conditions ripe for a thermonuclear
+detonation.
 
-There have been a number of previous studies which provide useful insight into setting up
-the problem and the burning methodology \citep{rosswog:2009,raskin:2010,loren-aguilar:2010,
-hawley:2012,garcia-senz:2013,kushnir:2013,papish:2015,holcomb:2015}. These studies
-found that detonations occur and convert a large amount of carbon/oxygen material into
-iron-group elements. These papers varied significantly in the hydrodynamic methods used
-(Lagrangian versus Eulerian methods), the methodology used for coupling nuclear
-reactions to the hydrodynamics (including variation in the number of isotopes in
-the nuclear network and in the evolution of the temperature), and the temporal and
-spatial resolution. Consequently the detonations they formed varied in location and nature,
-and the resultant estimates of production of nickel-group elements varied significantly,
-leaving much uncertainty about how such an event would appear observationally and whether
-it bears any resemblance to a SN Ia. See Table 4 of \cite{garcia-senz:2013} for a summary
-of the outcome of many of these studies. In this paper we make the case that
-while we can identify the main problems prevent accurate simulation of the collision
-event, there is no clear path to overcoming them using the computational resources
-typically available for contemporary hydrodynamics simulations.
+Early studies on WD collisions \citep{rosswog:2009,raskin:2010,loren-aguilar:2010,
+hawley:2012,garcia-senz:2013} typically had effective spatial resolutions in the
+range 100--500 km, and observed detonations that convert a large amount of
+carbon/oxygen material into iron-group elements. These studies varied in methodology
+(Lagrangian versus Eulerian evolution, nuclear network used) and did not closely agree
+on the final result of the event (see Table 4 of \cite{garcia-senz:2013} for a summary).
+\cite{kushnir:2013} argued that many of these simulations featured numerically unstable
+evolution, ultimately caused by the zone size being significantly larger than the length
+scale over which detonations form. The detonation length scale can vary widely based
+on physical conditions \citep{seitenzahl:2009,garg:2017} but is generally much
+smaller than 100 km. In this paper, we seek to find what numerical methods are required
+to achieve converged detonations.
+
+\section{Numerical Instability}
+\label{sec:instability}
+
+% Discuss the Kushnir stability argument
 

### 2018-07-12 papers/ignition/paper.tex
Move methodology section after introduction

@@
 
-\section{Numerical Instability}
-\label{sec:instability}
+\section{Nuclear Burning Methodology}
+\label{sec:methodology}
 
-% Discuss the Kushnir stability argument
-
-%==========================================================================
-% Problem Setup
-%==========================================================================
-\section{White Dwarf Collision Setup}
-\label{sec:collisions}
-
-We implement the white dwarf collision problem in \castro\footnote{The main results
-in this paper were generated using the 17.12 releases of \castro, \amrex, and
-\microphysics.} in the \wdmerger\ problem
-directory. We prepare the domain in the way most previous papers have done so:
-the white dwarf centers of mass are initially separated by a distance of four times
-the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
-consequently the initial distance, depends on the equation of state used, and this
-can vary somewhat depending on the relevant physics used, such as Coulomb corrections
-in the Helmholtz equation of state. However, the results do not really depend on the
-exact initial distance as long as there is enough time for the WDs to distort in
-response to tidal forces as they approach. Their initial velocity is that of
-two point-masses in free-fall towards each other coming in from infinity, such that
-the contact point is at the origin, and they approach each other along a coordinate axis.
-In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
-composed of equal parts carbon and oxygen by mass, at an initially uniform temperature
-of $T = 10^7$ K. This choice is for comparison to earlier studies; the code does have
-the ability to vary the impact parameter of the collision and the mass and composition
-of the WDs.
-
-We have added to the \wdmerger\ problem setup the ability to take advantage of 
-the 2D cylindrical ($R-z$) coordinate system evolution in \castro, and we use it
-in this study, because the axisymmetry inherent to the cylindrical coordinate system
-lends itself well to the head-on collision problem. In this coordinate system, we
-align the WDs along the $z$-axis (which is analogous to the $x$-axis in Cartesian
-evolution), with the center of the WDs at $R = 0$. The un-simulated $\phi$ dimension
-then would extend the WDs through a $2\pi$ revolution. The domain has width $4 \times 10^{9}$
-cm along the $z$ axis, and is half as wide along the $R$ axis. The resolution is equal
-in both dimensions, so that there are twice as many zones in the $z$ dimension in
-a uniform grid.
-
-We terminate the simulation at $t = 9$ seconds, as we have found that for this
-particular setup, all of the significant burning has completed by this time\footnote{In
-some simulations we were mainly interested in seeing when the detonation happened, and
-we terminated the simulation at that point. We show nucleosynthesis results only for
-those simulations that had completed nuclear burning.}. At
-this point, the total energy on the domain has become positive, typically about
-$10^{51}$ ergs, and has stopped increasing. This means that the system has become
-unbound due to nuclear energy release, and that no further meaningful nuclear energy
-generation is occurring. In our discussions below, it helps to have some other
-anchoring points in time when significant events occur. We arbitrarily define the
-collision time, $t_{\rm c}$, as the time when the density at the center of the grid
-first reaches $10^2\ \text{g\,/\,cm}^3$. This is reached only when the WDs meet at the
-center of the domain. We define the detonation time, $t_{\rm d}$, as the time when the
-temperature first reaches $4.5 \times 10^9\ \text{K}$. This is also an arbitrary
-threshold, but we have found empirically that this temperature is only reached in
-the collision when a detonation is formed.
-
-The region outside the WDs is filled with a low-density ambient
-material with $\rho = 10^{-4}\ \text{g\,/\,cm}^{3}$ whose composition and temperature is the
-same as that of the WDs. Reactions in this region are unimportant, so for
-computational efficiency, we disable all nuclear burning for zones that have
-$\rho < 10^6\ \text{g\,/\,cm}^{3}$ and $T < 10^8\ \text{K}$. We have confirmed that
-the results do not depend on this choice. A sponge is applied to all gas of density
-lower than $1\ \text{g\,/\,cm}^{3}$. The sponge damps out the velocity in these zones
-(see Paper I) to prevent spurious numerical error in low-density gas from dragging
-down the global timestep. The density floor for the simulation is set
-at $10^{-5}\ \text{g\,/\,cm}^{3}$ and the temperature floor is $10^7$ K.
-
-We collect information at the end of every timestep about the total amount of
-all of the species on the grid (in solar masses), and we use the mass of $^{56}$Ni
-generated as a metric for this test, as this is the parameter most directly related
-to the observable quantities of interest for Type Ia supernovae. Since, as we will see
-below, not all of our simulations detonate at a density high enough to generate substantial
-amounts of nickel, we employ the mass-weighted average atomic mass number on the domain,
-$\bar{A}$, as a more general metric describing the outcome of the simulation. A
-final relevant metric is the total amount of energy $E$ released by the detonation.
-
-Plotfiles of relevant quantities are created every 0.01 seconds.
-The timesteps are limited to ensure that every multiple of this plotfile interval is
-reached exactly, which means an effective maximum timestep of 0.01 seconds. Timesteps
-are by default limited only by the hydrodynamic stability constraint and this plotfile
-interval constraint, except where noted otherwise.
-
-For a collision of two 0.64 $\msolar$ WDs with 128 zones in the radial dimension,
-corresponding to a uniform spatial resolution of 312.5 km, the collision time is
-$t_{\rm c} = 4.16\ \text{s}$, and the detonation time is $t_{\rm d} = 7.43\ \text{s}$.
-The peak density at the detonation time is $2.3 \times 10^7\ \text{g\,/\,cm}^3$.
-Approximately $0.21\ \msolar$ of nickel and $0.46\ \msolar$ of silicon is created,
-corresponding to an $\bar{A}$ of about 29. Approximately $1.3 \times 10^{51}$
-ergs of energy are released. Simulations with resolutions in this few hundred kilometer
-range have been a common choice in earlier papers on white dwarf mergers and collisions
-due to computational constraints; our goal is to see how higher resolution runs differ.
-
-%==========================================================================
-% Numerical Implementation
-%==========================================================================
-\section{Numerical Methods}
-\label{sec:numericalmethods}
-
-The hydrodynamical equations solved in this paper were presented in \citet{wdmergerI}
-(hereafter, Paper I) for the reacting hydrodynamics code \castro\ \citep{castro}.
-Only minor revisions to the core hydrodynamic algorithm have since occurred.
-In this paper we focus our discussion on our implementation of the nuclear
-reaction network and integrator. There are three main areas of concern: first,
-specifying what nuclides are in the network and what reaction rates link the
-various nuclides; second, once the system of ODEs governing the evolution of
-the nuclear species has been written down, how to couple them to ODEs governing
-the evolution of the thermodynamics, in particular the temperature; third,
-coupling the nuclear reaction updates to the evolution of the hydrodynamical
-system. We will deal with each of these in turn in the following subsections.
-
-\subsection{Nuclear Network}
-\label{sec:network}
-
-In this paper we use the \aproxthirteen\ network\ \citep{timmes:1999,timmes:2000}.
-This includes all thirteen $\alpha$-chain particles between ${}^4$He and ${}^{56}$Ni.
-These $\alpha$-chain particles capture the most important pathway connecting a WD
-and its thermonuclear detonation. This network was used by \citet{hawley:2012}
-and \citet{raskin:2010}. \citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a
-very similar network that additionally included ${}^{60}$Zn. The \aproxthirteen\ network has been
-ported into a form that is consistent with the \amrex\ codes in the freely available
-\microphysics\ code repository\footnote{\microphysics\ can be obtained at
-\url{https://github.com/starkiller-astro/Microphysics}.},
-a collection of microphysical routines that are designed to be used in our
-hydrodynamics codes. In our \microphysics\ repository we have also ported \aproxnineteen\
-\citep{timmes:1999}, which was used in a few previous collision papers, and \aproxtwentyone.
-The isotopes added by these networks relative to \aproxthirteen\ are not particularly relevant
-to the problem of head-on white dwarf collisions, so we do not discuss results with these networks
-here (though we have verified for this problem that these larger networks do not significantly change
-the answer). We also make available \isoseven\ \citep{timmes:2000}, which skips the nuclides between
-silicon and nickel, and which was used by \citet{rosswog:2009}. \isoseven\ gives significantly different
-quantitative results than \aproxthirteen\ for this problem, due to its inapplicable assumption of
-an equilibrium abundance of the unsimulated isotopes, but the results are still qualitatively similar.
+In this section we discuss several issues in reactive hydrodynamic simulations and
+how we addressed them in the \castro\ code \citep{castro}. We first describe the
+core operation of our burner module, which is implemented in the \microphysics\ code
+repository\footnote{\microphysics\ can be obtained at \url{https://github.com/starkiller-astro/Microphysics}.},
+a collection of microphysical routines that are compatible with \castro.
 
@@
 %==========================================================================
+% 1D collision test problem
+%==========================================================================
+\section{Numerical Test Problem}
+\label{sec:collisions}
+
+We implement the white dwarf collision problem in \castro\footnote{The main results
+in this paper were generated using the 17.12 releases of \castro, \amrex, and
+\microphysics.} in the \wdmerger\ problem
+directory. We prepare the domain in the way most previous papers have done so:
+the white dwarf centers of mass are initially separated by a distance of four times
+the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
+consequently the initial distance, depends on the equation of state used, and this
+can vary somewhat depending on the relevant physics used, such as Coulomb corrections
+in the Helmholtz equation of state. However, the results do not really depend on the
+exact initial distance as long as there is enough time for the WDs to distort in
+response to tidal forces as they approach. Their initial velocity is that of
+two point-masses in free-fall towards each other coming in from infinity, such that
+the contact point is at the origin, and they approach each other along a coordinate axis.
+In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
+composed of equal parts carbon and oxygen by mass, at an initially uniform temperature
+of $T = 10^7$ K. This choice is for comparison to earlier studies; the code does have
+the ability to vary the impact parameter of the collision and the mass and composition
+of the WDs.
+
+We have added to the \wdmerger\ problem setup the ability to take advantage of 
+the 2D cylindrical ($R-z$) coordinate system evolution in \castro, and we use it
+in this study, because the axisymmetry inherent to the cylindrical coordinate system
+lends itself well to the head-on collision problem. In this coordinate system, we
+align the WDs along the $z$-axis (which is analogous to the $x$-axis in Cartesian
+evolution), with the center of the WDs at $R = 0$. The un-simulated $\phi$ dimension
+then would extend the WDs through a $2\pi$ revolution. The domain has width $4 \times 10^{9}$
+cm along the $z$ axis, and is half as wide along the $R$ axis. The resolution is equal
+in both dimensions, so that there are twice as many zones in the $z$ dimension in
+a uniform grid.
+
+We terminate the simulation at $t = 9$ seconds, as we have found that for this
+particular setup, all of the significant burning has completed by this time\footnote{In
+some simulations we were mainly interested in seeing when the detonation happened, and
+we terminated the simulation at that point. We show nucleosynthesis results only for
+those simulations that had completed nuclear burning.}. At
+this point, the total energy on the domain has become positive, typically about
+$10^{51}$ ergs, and has stopped increasing. This means that the system has become
+unbound due to nuclear energy release, and that no further meaningful nuclear energy
+generation is occurring. In our discussions below, it helps to have some other
+anchoring points in time when significant events occur. We arbitrarily define the
+collision time, $t_{\rm c}$, as the time when the density at the center of the grid
+first reaches $10^2\ \text{g\,/\,cm}^3$. This is reached only when the WDs meet at the
+center of the domain. We define the detonation time, $t_{\rm d}$, as the time when the
+temperature first reaches $4.5 \times 10^9\ \text{K}$. This is also an arbitrary
+threshold, but we have found empirically that this temperature is only reached in
+the collision when a detonation is formed.
+
+The region outside the WDs is filled with a low-density ambient
+material with $\rho = 10^{-4}\ \text{g\,/\,cm}^{3}$ whose composition and temperature is the
+same as that of the WDs. Reactions in this region are unimportant, so for
+computational efficiency, we disable all nuclear burning for zones that have
+$\rho < 10^6\ \text{g\,/\,cm}^{3}$ and $T < 10^8\ \text{K}$. We have confirmed that
+the results do not depend on this choice. A sponge is applied to all gas of density
+lower than $1\ \text{g\,/\,cm}^{3}$. The sponge damps out the velocity in these zones
+(see Paper I) to prevent spurious numerical error in low-density gas from dragging
+down the global timestep. The density floor for the simulation is set
+at $10^{-5}\ \text{g\,/\,cm}^{3}$ and the temperature floor is $10^7$ K.
+
+We collect information at the end of every timestep about the total amount of
+all of the species on the grid (in solar masses), and we use the mass of $^{56}$Ni
+generated as a metric for this test, as this is the parameter most directly related
+to the observable quantities of interest for Type Ia supernovae. Since, as we will see
+below, not all of our simulations detonate at a density high enough to generate substantial
+amounts of nickel, we employ the mass-weighted average atomic mass number on the domain,
+$\bar{A}$, as a more general metric describing the outcome of the simulation. A
+final relevant metric is the total amount of energy $E$ released by the detonation.
+
+Plotfiles of relevant quantities are created every 0.01 seconds.
+The timesteps are limited to ensure that every multiple of this plotfile interval is
+reached exactly, which means an effective maximum timestep of 0.01 seconds. Timesteps
+are by default limited only by the hydrodynamic stability constraint and this plotfile
+interval constraint, except where noted otherwise.
+
+For a collision of two 0.64 $\msolar$ WDs with 128 zones in the radial dimension,
+corresponding to a uniform spatial resolution of 312.5 km, the collision time is
+$t_{\rm c} = 4.16\ \text{s}$, and the detonation time is $t_{\rm d} = 7.43\ \text{s}$.
+The peak density at the detonation time is $2.3 \times 10^7\ \text{g\,/\,cm}^3$.
+Approximately $0.21\ \msolar$ of nickel and $0.46\ \msolar$ of silicon is created,
+corresponding to an $\bar{A}$ of about 29. Approximately $1.3 \times 10^{51}$
+ergs of energy are released. Simulations with resolutions in this few hundred kilometer
+range have been a common choice in earlier papers on white dwarf mergers and collisions
+due to computational constraints; our goal is to see how higher resolution runs differ.
+
+%==========================================================================
 % Resolution Dependence

### 2018-07-13 papers/ignition/paper.tex
Move burner to its own section

@@
 \newcommand{\castro}{\texttt{CASTRO}}
+\newcommand{\maestro}{\texttt{Maestro}}
 \newcommand{\microphysics}{\texttt{Microphysics}}
@@
 \section{Nuclear Burning Methodology}
-\label{sec:methodology}
-
-In this section we discuss several issues in reactive hydrodynamic simulations and
-how we addressed them in the \castro\ code \citep{castro}. We first describe the
-core operation of our burner module, which is implemented in the \microphysics\ code
-repository\footnote{\microphysics\ can be obtained at \url{https://github.com/starkiller-astro/Microphysics}.},
-a collection of microphysical routines that are compatible with \castro.
-
-\subsection{Nuclear Burning}
 \label{sec:burner}
 
-Given a set of nuclides and the reaction links between them, we now consider
+We first describe the core operation of our burner module, which is implemented in
+the \microphysics\ code repository\footnote{\microphysics\ can be obtained at
+  \url{https://github.com/starkiller-astro/Microphysics}.}, a collection of microphysical
+routines that are compatible with \castro, \maestro, and other \amrex-based astrophysics
+codes. We review this because, even though the ideas are not new,
+it serves as an update on the state of the \castro\ code since the original release of
+the code paper, similar to the goal of \cite{wdmergerI} for the hydrodynamics and gravity
+modules.
+
+Given a set of nuclides and the reaction links between them, we consider
 how a burning step is performed in our software. The goal is to integrate the
@@
 
-We define several burning modes that determine how $T$ and $e$ are evolved
-during a nuclear burn. In a hydrostatic burn, which we call burning mode 0,
-we keep $\rho$ and $T$ fixed throughout, and use 
-the energy released at the end to compute a final temperature that is
-thermodynamically consistent with the new internal energy. By contrast,
-in a self-heating burn (mode 1), we allow the temperature to evolve in response
+For most of the problems of interest to \castro\, we employ a self-heating burn,
+where we allow the temperature to evolve in response
 to the burning (see\footnote{In the cited paper, a term based on the
@@
 \end{equation}
-(Although $T$ evolves during the burn so that the integration is physically
-accurate, as in the hydrostatic method we discard the final value
-for $T$ at the end of the burn and recompute a temperature for the zone that is
-consistent with its new internal energy.) Here $c_V$ is the specific heat at
-constant volume, which is provided by the equation of state.  During this burn,
-we can keep $c_V$ constant using its initial value, or at each step we
-can choose to re-evaluate the equation of state using the latest value of $(\rho, T)$.
-The latter is more expensive but also more accurate, and we use it in this paper.
-In practice we find that the cost is small in comparison to the more expensive
-parts of the calculation, and it can significantly speed up convergence near NSE.
-A third option (mode 2) presented by \citet{raskin:2010} is a so-called ``hybrid'' mode.
-In this mode, by default we do a hydrostatic burn. If that burn fails, or if the net
-energy change is negative, we do the burn again in self-heating mode. A final option (mode 3)
-is a burn that limits the changes due to a burn to avoid numerically unstable burning.
-This mode is discussed in \autoref{sec:unstable_burning} and we will call it a
-``suppressed'' burn for the remainder of this work. All four options
-are implemented in our burner software.
-
-The simulations shown in this work all use the self-heating mode. The hydrostatic
-and hybrid burns are only a slight variation on the self-heating burn in practice
-for this problem.
-The reason that the hydrostatic burn cannot substantially change the outcome is
-that even though the temperature is fixed during the burn, it is \textit{not}
-constant in the rest of the hydrodynamics step. And in the hydrodynamics step,
-the \castro\ algorithm performs multiple EOS calls to ensure that the temperature
-is synchronized with the internal energy at various points in the update. So in
-practice there is a temperature change due to the burn for our problem. This seems
-inevitable for any hydro code unless the hydrodynamics update includes an explicit
-equation for $T$ that is independent of what is happening for the internal energy.
-The hydrostatic burn does slightly limit the energy release during any given burn,
-but summed up over the whole simulation this is not enough to qualitatively affect
-the outcome. The hybrid burn does not affect the simulation outcome at all relative
-to the hydrostatic burn. The difference between them only applies for zones that have
-a net negative energy release from the hydrostatic burn, which occurs only after
-we have burned to NSE. The hybrid burn was mainly implemented by \citet{raskin:2010}
-for making it easier to numerically complete a burn, rather than because of the
-effects it has on the outcome of the simulation. The suppressed burn has separate
-issues which we defer to \autoref{sec:unstable_burning}.
+Here $c_V$ is the specific heat at constant volume, which
+is provided by the equation of state. During this burn, we re-evaluate $c_V$
+by calling the equation of state using the latest value of $(\rho, T)$.
+The added expense of the equation of state call (relative to keeping $c_V$ fixed
+through the burn) is small in comparison to the more expensive parts of the calculation,
+and it can significantly speed up convergence near NSE. Although $T$ evolves during the
+burn so that the computation of the reaction rates is accurate, we discard the final value
+for $T$ at the end of the burn and recompute a temperature for the zone that is consistent
+with its new internal energy according to the equation of state.
 
@@
 compatible with our software interfaces in the \microphysics\ repository.
-We also employ an implementation of the well known variable-order Richardson
-extrapolation method presented by \citet{stoer:1980}, that is similar to the
-integrator which ships with the original versions of the networks mentioned
-above. Our experience shows that relying on a single ODE integrator for the
-variety of stiff ODEs that are present in a collision is insufficient, as we
-sometimes run into integration failures. A combination of the two integrators
-is usually sufficient, however: we first try \vode, and if that fails, we then
-try the Stoer and Bulirsch method. Since it is possible for neither method to
-succeed, we have one more backup scheme. Our specified (relative and absolute)
-error tolerances are uniformly $10^{-6}$ for the energy, temperature, and nuclear
-abundance ODEs. We can loosen these tolerances in an attempt to give the integrator
-an easier time. If both \vode\ and the Stoer and Bulirsch algorithm fail, then we
-loosen the tolerances by a multiplicative factor of 25\%, and then repeat the attempts
-with \vode\ and Stoer and Bulirsch. We keep doing this until either we have loosened
-the tolerances enough for one of the two methods to succeed, or until we have loosened
-the tolerances by two orders of magnitude, at which point we finally give up. We have
-found that this scheme is resilient enough to handle all of the simulations shown below.
-We have also verified that our results do not depend strongly on the tolerances when they
-are tighter than $10^{-4}$. For the typical collision problem we discuss below, the
-nucleosynthetic yield for nickel varies by less than 5\% between a tolerance of $10^{-4}$
-and a tolerance of $10^{-12}$.
+Another option is an implementation of the well known variable-order Richardson
+extrapolation method presented by \citet{stoer:1980}.
 
-\subsection{Coupling to the Hydrodynamics}
+\section{Coupling Burning and Hydrodynamics}
 \label{sec:hydrocoupling}
@@
 We have found that the benefits for this problem are, however, small.
+
+% Sticking this here for now, needs to be updated with new retry discussion
+
+Our experience shows that relying on a single ODE integrator for the
+variety of stiff ODEs that are present in a collision is insufficient, as we
+sometimes run into integration failures. A combination of the two integrators
+is usually sufficient, however: we first try \vode, and if that fails, we then
+try the Stoer and Bulirsch method. Since it is possible for neither method to
+succeed, we have one more backup scheme. Our specified (relative and absolute)
+error tolerances are uniformly $10^{-6}$ for the energy, temperature, and nuclear
+abundance ODEs. We can loosen these tolerances in an attempt to give the integrator
+an easier time. If both \vode\ and the Stoer and Bulirsch algorithm fail, then we
+loosen the tolerances by a multiplicative factor of 25\%, and then repeat the attempts
+with \vode\ and Stoer and Bulirsch. We keep doing this until either we have loosened
+the tolerances enough for one of the two methods to succeed, or until we have loosened
+the tolerances by two orders of magnitude, at which point we finally give up. We have
+found that this scheme is resilient enough to handle all of the simulations shown below.
+We have also verified that our results do not depend strongly on the tolerances when they
+are tighter than $10^{-4}$. For the typical collision problem we discuss below, the
+nucleosynthetic yield for nickel varies by less than 5\% between a tolerance of $10^{-4}$
+and a tolerance of $10^{-12}$.
 

### 2018-07-13 papers/ignition/paper.tex
Strip reaction equations

@@
 
-\section{Nuclear Burning Methodology}
-\label{sec:burner}
-
-We first describe the core operation of our burner module, which is implemented in
-the \microphysics\ code repository\footnote{\microphysics\ can be obtained at
-  \url{https://github.com/starkiller-astro/Microphysics}.}, a collection of microphysical
-routines that are compatible with \castro, \maestro, and other \amrex-based astrophysics
-codes. We review this because, even though the ideas are not new,
-it serves as an update on the state of the \castro\ code since the original release of
-the code paper, similar to the goal of \cite{wdmergerI} for the hydrodynamics and gravity
-modules.
-
-Given a set of nuclides and the reaction links between them, we consider
-how a burning step is performed in our software. The goal is to integrate the
-vector ${\bm{Y}} = (Y_1, Y_2, \ldots, Y_n, e, T)$, where $Y_{n} = X_{n} / A_{n}$
-is the molar fraction of species $n$, with $X_n$ the mass fraction and $A_n$ the
-mass number of that species, $e$ is the energy released during the burn, and
-$T$ is the temperature. The equation describing its evolution is given by
-\begin{equation}
-  \frac{d\bm{Y}}{dt} = f(\mathbf{Y}),
-\end{equation}
-where the components of the right-hand-side for the species come from the particular
-nuclear burning network we are using. The energy $e$ of the zone
-will change when the nuclear abundances evolve, according to
-\begin{equation}
-  \frac{\partial e}{\partial t} = N_A \sum_{n} \frac{\partial Y_{n}}{\partial t} m_{n} c^2,
-\end{equation}
-where $c$ is the speed of light and $m_n$ is the mass of each nuclide.
-
-For most of the problems of interest to \castro\, we employ a self-heating burn,
-where we allow the temperature to evolve in response
-to the burning (see\footnote{In the cited paper, a term based on the
-thermodynamic chemical potential was included; we now believe
-that it is incorrect to include such a term in the burn, since it
-automatically sums to zero analytically.} \citet{maestro3}):
-\begin{equation}
-  \frac{dT}{dt} = \frac{1}{c_V}\frac{\partial e}{\partial t}
-\end{equation}
-Here $c_V$ is the specific heat at constant volume, which
-is provided by the equation of state. During this burn, we re-evaluate $c_V$
-by calling the equation of state using the latest value of $(\rho, T)$.
-The added expense of the equation of state call (relative to keeping $c_V$ fixed
-through the burn) is small in comparison to the more expensive parts of the calculation,
-and it can significantly speed up convergence near NSE. Although $T$ evolves during the
-burn so that the computation of the reaction rates is accurate, we discard the final value
-for $T$ at the end of the burn and recompute a temperature for the zone that is consistent
-with its new internal energy according to the equation of state.
-
-In our \microphysics\ repository we provide several software options for
-solving a set of coupled stiff ODEs. For this work our primary integrator
-is the classic \vode\ integrator \citep{vode}, of which we have a version
-compatible with our software interfaces in the \microphysics\ repository.
-Another option is an implementation of the well known variable-order Richardson
-extrapolation method presented by \citet{stoer:1980}.
+The numerical simulations in this paper are performed using the reacting hydrodynamics AMR
+code \castro\ \citep{castro}. \cite{wdmergerI} described updates to the core hydrodynamic
+algorithm since the code's original release. The goal of this paper is to describe several
+limitations we found when coupling the hydrodynamics module to the reactions module for detonation
+problems, and the changes we made to the \castro\ code to deal with those limitations (though
+most of the ideas are applicable to any adaptive mesh-based code). The challenge of accurately
+capturing this coupling explains most of the difficulty in obtaining converged detonation
+simulations.
 

### 2018-07-13 papers/ignition/paper.tex
Start shaping up section 2

@@
 
+\section{Coupling Burning and Hydrodynamics}
+\label{sec:hydrocoupling}
+
 The numerical simulations in this paper are performed using the reacting hydrodynamics AMR
 code \castro\ \citep{castro}. \cite{wdmergerI} described updates to the core hydrodynamic
-algorithm since the code's original release. The goal of this paper is to describe several
+algorithm since the code's original release. The goal of this section is to describe several
 limitations we found when coupling the hydrodynamics module to the reactions module for detonation
@@
 
-\section{Coupling Burning and Hydrodynamics}
-\label{sec:hydrocoupling}
+\subsection{Timestep Limiting}
+\label{sec:timestep}
 
@@
 $f_{bX}$ has not been used by others performing these types of calculations. The sensitivity
-of results to the value of these timestep limiters will be discussed later in the
+of simulations to the value of these timestep limiters will be discussed later in the
 results. The factors $f_{be}$ and $f_{bX}$ can be set at runtime in \castro.

### 2018-07-13 papers/ignition/paper.tex
Update timestep limiter

@@
 In \castro, the reactions are coupled to the hydrodynamics using Strang splitting.
-In a given timestep advance $\Delta t$, we first evolve the reactions alone through
-a time interval $\Delta t / 2$. Then, we evolve the hydrodynamics for $\Delta t$,
-and we evolve the reactions again for a further $\Delta t / 2$. The principal
-drawback of this approach is that the reactions and the hydrodynamics can become
-decoupled from each other. A common solution to this problem presented in
-the literature has been to limit the size of the timestep and thereby limit the
-extent of this decoupling \citep{raskin:2010,hawley:2012}, which we adopt here 
-and have implemented in \castro. Defining the nuclear energy injection timescale 
-$\tau_e$, and the species evolution timescale $\tau_{X_k}$,
+Consider the basic outline of a single-level advance in an advection-reaction system:
+\begin{enumerate}
+  \item Evaluate timestep $\Delta t$ for the current advance
+  \item Advance the nuclear burning network by $\Delta t / 2$
+  \item Advance the hydrodynamics by $\Delta t$
+  \item Advance the nuclear burning network by $\Delta t / 2$
+  \item Return to Step 1
+\end{enumerate}
+The principal drawback of this approach is that the reactions and the hydrodynamics
+can become decoupled from each other. For example, consider the example of a head-on
+collision of white dwarfs. Initial nuclear burning will occur at the contact point
+between the two stars. Because of the staggered updates from splitting, the evolution
+effectively progresses as a cycle between burning for $\Delta t$ and getting fresh
+material advected into the contact point by the hydro update for $\Delta t$.
+When the collision begins, $\Delta t$ is controlled by the hydrodynamic stability
+criterion, and may be large enough that it is possible for the burning advance in
+Step 4 to completely burn the freshly advected material all the way to NSE.
+Consequently the evolution is no longer a good approximation to smooth burning of
+the in-falling material but rather separate discrete burning and hydro steps, and
+the nature of the burning evolution will be quite different.
+
+A common solution to this problem presented in the literature has been to limit the
+size of the timestep and thereby limit the extent of this decoupling
+\citep{raskin:2010,hawley:2012}, which we adopt here and have implemented in \castro.
+Defining the nuclear energy injection timescale $\tau_e$, and the species evolution
+timescale $\tau_{X_k}$,
 \begin{align}
@@
 literature is in the range of 0.1 to 0.3, while to our knowledge a limiter based on
-$f_{bX}$ has not been used by others performing these types of calculations. The sensitivity
-of simulations to the value of these timestep limiters will be discussed later in the
-results. The factors $f_{be}$ and $f_{bX}$ can be set at runtime in \castro.
+$f_{bX}$ has not been used by others performing these types of calculations.
+This limiter is complementary to the limiter based on $f_{be}$, and can catch
+certain burning regimes that limiting only based on changes in energy would miss.
 
@@
 estimating the time derivatives.
+
+The size of the splitting error is a function of the advance timestep, and it
+decreases as the timestep becomes smaller. When we add AMR in the regions of
+nuclear burning, the hydrodynamic timestep automatically becomes smaller due
+to the increased resolution, which mitigates the splitting error. In fact, we
+have found that when we have sufficient refinement to achieve a converged
+simulation, the size of the hydrodynamic timestep is already small enough
+that the splitting error is minimal. Consequently this timestep limiter
+is only of great importance for low resolution simulations. Additionally,
+we found that for the test problem described later, where we are only trying
+to locate the point and time of a detonation rather than assess accurate
+nucleosynthetic yields, this timestep limiter has no observable effect,
+so we do not discuss it further in this paper.
+
+\section{Timestep Retries}
+\label{sec:retries}
 
@@
 %TODO: add a ``Nonaka plot''
-
-To understand the consequences of the timestep limiting, and more broadly to 
-understand the limitations of Strang splitting, we consider the 
-basic outline of a single-level advance in an advection-reaction system:
-\begin{enumerate}
-  \item Evaluate timestep $\Delta t$ for the current advance
-  \item Advance the nuclear burning network by $\Delta t / 2$
-  \item Advance the hydrodynamics by $\Delta t$
-  \item Advance the nuclear burning network by $\Delta t / 2$
-  \item Return to Step 1
-\end{enumerate}
-Now, consider that during a head-on collision, initial nuclear burning 
-will occur at the contact point between the two stars. Because of 
-the staggered updates from splitting, the evolution effectively progresses 
-as a cycle between burning for $\Delta t$ and getting fresh material 
-advected into the contact point by the hydro update for $\Delta t$. 
-When the collision begins, $\Delta t$ is controlled by the hydrodynamic 
-stability criterion, and may be large enough that it is possible for 
-the burning advance in Step 4 to completely burn the freshly advected 
-material all the way to NSE. Consequently the evolution is no longer 
-a good approximation to smooth burning of the in-falling material but
-rather separate discrete burning and hydro steps, and the nature of 
-the burning evolution will be quite different. As the timestep becomes
-smaller, the evolution becomes smoother and the splitting error decreases.
 

### 2018-07-17 papers/ignition/paper.tex
Section -> subsection
Put the AMR discussion before the timestep limiter discussion

@@
 
-\subsection{Timestep Limiting}
+\subsection{Timestep Size}
 \label{sec:timestep}
@@
 the in-falling material but rather separate discrete burning and hydro steps, and
-the nature of the burning evolution will be quite different.
+the nature of the burning evolution will be quite different. This can have a
+substantial effect on nucleosynthetic yields predicted by white dwarf simulations.
 
-A common solution to this problem presented in the literature has been to limit the
-size of the timestep and thereby limit the extent of this decoupling
-\citep{raskin:2010,hawley:2012}, which we adopt here and have implemented in \castro.
+The size of the splitting error is a function of the advance timestep, and it
+decreases as the timestep becomes smaller. When we add AMR in the regions of
+nuclear burning, the hydrodynamic timestep (as set by the CFL criterion)
+automatically becomes smaller due to the increased resolution, which mitigates
+the splitting error. We have found that when we have sufficient refinement to
+achieve a converged simulation, the size of the hydrodynamic timestep is already
+small enough that the splitting error is minimal. For the test problem that we
+will describe later, the operator splitting has no observable effect on the
+physical effect we care about --- the time and location of the detonation --- at
+any resolution. Consequently we do not apply any special timestep limiter based
+on the rate of burning in this paper.
+
+However, for completeness, we describe our implementation of a common approach
+presented in the literature for timestep limiting based on burning, which should
+be used when performing low resolution SN Ia simulations \citep{raskin:2010,hawley:2012}.
 Defining the nuclear energy injection timescale $\tau_e$, and the species evolution
@@
 this as a finite difference using the change over the last timestep. Our method
-is more accurate in principle, but in practice for this problem we find no
-meaningful difference between the limiting we propose and other methods of
-estimating the time derivatives.
+is more accurate in principle, but in practice we find no meaningful difference
+between the limiting we propose and other methods of estimating the time derivatives.
 
-The size of the splitting error is a function of the advance timestep, and it
-decreases as the timestep becomes smaller. When we add AMR in the regions of
-nuclear burning, the hydrodynamic timestep automatically becomes smaller due
-to the increased resolution, which mitigates the splitting error. In fact, we
-have found that when we have sufficient refinement to achieve a converged
-simulation, the size of the hydrodynamic timestep is already small enough
-that the splitting error is minimal. Consequently this timestep limiter
-is only of great importance for low resolution simulations. Additionally,
-we found that for the test problem described later, where we are only trying
-to locate the point and time of a detonation rather than assess accurate
-nucleosynthetic yields, this timestep limiter has no observable effect,
-so we do not discuss it further in this paper.
-
-\section{Timestep Retries}
+\subsection{Timestep Retries}
 \label{sec:retries}

### 2018-07-17 papers/ignition/paper.tex
Clean up the timestep retry section

@@
 
+%TODO: add a ``Nonaka plot''
+
 The size of the splitting error is a function of the advance timestep, and it
@@
 
-To be precise, we note that the timestep
-will only actually satisfy the energy criterion
-$\Delta t \leq f_e \tau_e$ and species criterion
-$\Delta t \leq f_{X_k} \tau_{X_k}$ when the estimates for
-$\dot{e}$ and $\dot{X_k}$ we generate are at least as large
-as the actual rate of change of energy and mass fractions
-over the timestep. However, this can assumption will sometimes fail
-during periods of runaway burning when the rate of change
-of these quantities is highly nonlinear. We may not want
-to neglect the errors caused by this approximation
-because they may build up over an extended period of nonlinear
-evolution and perhaps substantially change the final results.
-To this end, we have implemented a timestep retry option in
-\castro, which re-computes an advance if it violated the
-stability criteria as judged from the end of the timestep,
-using subcycled timesteps which individually satisfy the
-stability criterion to the tolerance we have asked for.
-We have found that the benefits for this problem are, however, small.
+There are many circumstances during a timestep of a hydro code where the advance
+fails. As one common example, in \castro\ timesteps can be subcycled on fine levels,
+so that if the refinement factor between levels is 4, we take four timesteps on
+the fine level for every timestep on the coarse level, each with a quarter the
+size of the coarse level timestep. Then the coarse and fine levels can be
+synchronized \citep{bergercolella:1989}, which requires that we set the fine level
+timestep at the beginning of the first fine level advance and have to use the same
+timestep for each of the four advances. But it can occur that the velocity on the
+fine level grows during these timesteps, such that during the second or later of
+the four advances, the CFL criterion is no longer satisfied.
 
-% Sticking this here for now, needs to be updated with new retry discussion
+Another example of a failure mode comes from the timestep limiting described in
+\ref{sec:timestep}. The timestep will only actually satisfy the energy criterion
+$\Delta t \leq f_e \tau_e$ and species criterion $\Delta t \leq f_{X_k} \tau_{X_k}$
+when the estimates for $\dot{e}$ and $\dot{X_k}$ we generate are at least as large
+as the actual rate of change of energy and mass fractions over the timestep. However,
+this can assumption will sometimes fail during periods of runaway burning when the
+rate of change of these quantities is highly nonlinear. We may not want to neglect
+the errors caused by this approximation because they may build up over an extended
+period of nonlinear evolution and perhaps substantially change the final results.
 
-Our experience shows that relying on a single ODE integrator for the
-variety of stiff ODEs that are present in a collision is insufficient, as we
-sometimes run into integration failures. A combination of the two integrators
-is usually sufficient, however: we first try \vode, and if that fails, we then
-try the Stoer and Bulirsch method. Since it is possible for neither method to
-succeed, we have one more backup scheme. Our specified (relative and absolute)
-error tolerances are uniformly $10^{-6}$ for the energy, temperature, and nuclear
-abundance ODEs. We can loosen these tolerances in an attempt to give the integrator
-an easier time. If both \vode\ and the Stoer and Bulirsch algorithm fail, then we
-loosen the tolerances by a multiplicative factor of 25\%, and then repeat the attempts
-with \vode\ and Stoer and Bulirsch. We keep doing this until either we have loosened
-the tolerances enough for one of the two methods to succeed, or until we have loosened
-the tolerances by two orders of magnitude, at which point we finally give up. We have
-found that this scheme is resilient enough to handle all of the simulations shown below.
-We have also verified that our results do not depend strongly on the tolerances when they
-are tighter than $10^{-4}$. For the typical collision problem we discuss below, the
-nucleosynthetic yield for nickel varies by less than 5\% between a tolerance of $10^{-4}$
-and a tolerance of $10^{-12}$.
+A third, related example comes from the burning module itself. The severe stiffness
+of the nuclear reaction ODEs means that we can sometimes run into integration failures.
+The most common cause of an integration failure is inability to reach the specified
+error tolerance within a specified maximum number of timesteps. These are rare in
+relative terms, but typically we find at least a few occur in any sufficiently long
+simulation, and we either have to deal with these or abort the simulation. In the
+\microphysics\ repository we have constructed a couple of ways of handling these
+failures internal to the burner: we can attempt to loosen the tolerance of the ODEs
+and try again; and, if that repeatedly fails, we can use a different integration method.
+However, we can be a little wiser about this and note that the failures are almost always
+ultimately related to the fact that the advance timestep is so large that the ODE
+integrator has to solve a very challenging problem. For a sufficiently small timestep,
+the size of the change that could occur during the burn would be similarly small,
+and it would be much easier for the integration to succeed.
 
-%TODO: add a ``Nonaka plot''
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
-  \caption{Nickel production as a function of the timestep factors $f_{be}$
-           (solid blue, limiting the timestep on changes in internal energy) and
-           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
-           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
-           the timestep by at least an order of magnitude more.
-           \label{fig:timestep_factor}}
-\end{figure}
-\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
-production for the baseline collision case. These limiters are somewhat redundant in
-the sense that a given $f_{bX}$ roughly corresponds to another (higher) $f_{be}$, as
-is evident by the fact that on the plot the curves look qualitatively similar in
-shape but with a horizontal offset. As either $f$ becomes small, the restriction on
-the timestep becomes more severe: for $f_{bX} = 0.1$, the full simulation requires
-nearly 30,000 timesteps, compared to just over 1,000 for the baseline case.
-Such a simulation would be quite expensive for a high resolution 3D calculation.
-It is clear, though, that even though the simulation is not converged when the
-timestep is not limited due to burning, the effect is small in absolute terms.
-So we agree that the use of values in this range by \cite{raskin:2010},
-\cite{hawley:2012}, and \cite{kushnir:2013} is appropriate. However, for reasons of
-computational expense we do not use the timestep limiter in the high resolution
-simulations that follow. The error due to this choice should be small, because
-the effect is not that large to begin with, and because the higher resolution
-decreases the timestep anyway due to the hydrodynamic stability constraint.
-
-Other approaches to the coupling 
-between reactions and hydrodynamics have been proposed in the 
-broader literature, especially iterative methods such as 
-deferred corrections that allow each of these operators to 
-feel the lagged effects of the other operators. For example,
-in the context of low Mach number flows, \cite{nonaka:2012} have
-used the method of spectral deferred corrections \citep{SDC} to
-couple their advection-diffusion-reaction equation set. In our
-context this would involve treating the full evolution equation
-for each of the state variables as an ODE with a directly coupled
-burning source term integrated at high order; the advective
-flux is evaluated at the standard second order accuracy
-and is included as an ODE source term.
-We are presently investigating such a method,
-and it may form the basis of further work on this subject.
+A natural solution to all of these problems is to be able to detect these failure
+modes, reject the proposed timestep, and subcycle the advance within a level with
+a smaller timestep. To this end, we have implemented a timestep retry option in
+\castro. If a failure is detected, the timestep is cut by a factor of 2 and we
+try again. If this does not work, the timestep is recursively cut until the
+advance succeeds. This approach neatly addresses several different types of
+problems in a single, simple framework.
 

### 2018-07-17 papers/ignition/paper.tex
Clear the results sections

@@
 Another example of a failure mode comes from the timestep limiting described in
-\ref{sec:timestep}. The timestep will only actually satisfy the energy criterion
+\autoref{sec:timestep}. The timestep will only actually satisfy the energy criterion
 $\Delta t \leq f_e \tau_e$ and species criterion $\Delta t \leq f_{X_k} \tau_{X_k}$
@@
 
-We implement the white dwarf collision problem in \castro\footnote{The main results
-in this paper were generated using the 17.12 releases of \castro, \amrex, and
-\microphysics.} in the \wdmerger\ problem
-directory. We prepare the domain in the way most previous papers have done so:
-the white dwarf centers of mass are initially separated by a distance of four times
-the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
-consequently the initial distance, depends on the equation of state used, and this
-can vary somewhat depending on the relevant physics used, such as Coulomb corrections
-in the Helmholtz equation of state. However, the results do not really depend on the
-exact initial distance as long as there is enough time for the WDs to distort in
-response to tidal forces as they approach. Their initial velocity is that of
-two point-masses in free-fall towards each other coming in from infinity, such that
-the contact point is at the origin, and they approach each other along a coordinate axis.
-In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
-composed of equal parts carbon and oxygen by mass, at an initially uniform temperature
-of $T = 10^7$ K. This choice is for comparison to earlier studies; the code does have
-the ability to vary the impact parameter of the collision and the mass and composition
-of the WDs.
-
-We have added to the \wdmerger\ problem setup the ability to take advantage of 
-the 2D cylindrical ($R-z$) coordinate system evolution in \castro, and we use it
-in this study, because the axisymmetry inherent to the cylindrical coordinate system
-lends itself well to the head-on collision problem. In this coordinate system, we
-align the WDs along the $z$-axis (which is analogous to the $x$-axis in Cartesian
-evolution), with the center of the WDs at $R = 0$. The un-simulated $\phi$ dimension
-then would extend the WDs through a $2\pi$ revolution. The domain has width $4 \times 10^{9}$
-cm along the $z$ axis, and is half as wide along the $R$ axis. The resolution is equal
-in both dimensions, so that there are twice as many zones in the $z$ dimension in
-a uniform grid.
-
-We terminate the simulation at $t = 9$ seconds, as we have found that for this
-particular setup, all of the significant burning has completed by this time\footnote{In
-some simulations we were mainly interested in seeing when the detonation happened, and
-we terminated the simulation at that point. We show nucleosynthesis results only for
-those simulations that had completed nuclear burning.}. At
-this point, the total energy on the domain has become positive, typically about
-$10^{51}$ ergs, and has stopped increasing. This means that the system has become
-unbound due to nuclear energy release, and that no further meaningful nuclear energy
-generation is occurring. In our discussions below, it helps to have some other
-anchoring points in time when significant events occur. We arbitrarily define the
-collision time, $t_{\rm c}$, as the time when the density at the center of the grid
-first reaches $10^2\ \text{g\,/\,cm}^3$. This is reached only when the WDs meet at the
-center of the domain. We define the detonation time, $t_{\rm d}$, as the time when the
-temperature first reaches $4.5 \times 10^9\ \text{K}$. This is also an arbitrary
-threshold, but we have found empirically that this temperature is only reached in
-the collision when a detonation is formed.
-
-The region outside the WDs is filled with a low-density ambient
-material with $\rho = 10^{-4}\ \text{g\,/\,cm}^{3}$ whose composition and temperature is the
-same as that of the WDs. Reactions in this region are unimportant, so for
-computational efficiency, we disable all nuclear burning for zones that have
-$\rho < 10^6\ \text{g\,/\,cm}^{3}$ and $T < 10^8\ \text{K}$. We have confirmed that
-the results do not depend on this choice. A sponge is applied to all gas of density
-lower than $1\ \text{g\,/\,cm}^{3}$. The sponge damps out the velocity in these zones
-(see Paper I) to prevent spurious numerical error in low-density gas from dragging
-down the global timestep. The density floor for the simulation is set
-at $10^{-5}\ \text{g\,/\,cm}^{3}$ and the temperature floor is $10^7$ K.
-
-We collect information at the end of every timestep about the total amount of
-all of the species on the grid (in solar masses), and we use the mass of $^{56}$Ni
-generated as a metric for this test, as this is the parameter most directly related
-to the observable quantities of interest for Type Ia supernovae. Since, as we will see
-below, not all of our simulations detonate at a density high enough to generate substantial
-amounts of nickel, we employ the mass-weighted average atomic mass number on the domain,
-$\bar{A}$, as a more general metric describing the outcome of the simulation. A
-final relevant metric is the total amount of energy $E$ released by the detonation.
-
 Plotfiles of relevant quantities are created every 0.01 seconds.
@@
 are by default limited only by the hydrodynamic stability constraint and this plotfile
-interval constraint, except where noted otherwise.
-
-For a collision of two 0.64 $\msolar$ WDs with 128 zones in the radial dimension,
-corresponding to a uniform spatial resolution of 312.5 km, the collision time is
-$t_{\rm c} = 4.16\ \text{s}$, and the detonation time is $t_{\rm d} = 7.43\ \text{s}$.
-The peak density at the detonation time is $2.3 \times 10^7\ \text{g\,/\,cm}^3$.
-Approximately $0.21\ \msolar$ of nickel and $0.46\ \msolar$ of silicon is created,
-corresponding to an $\bar{A}$ of about 29. Approximately $1.3 \times 10^{51}$
-ergs of energy are released. Simulations with resolutions in this few hundred kilometer
-range have been a common choice in earlier papers on white dwarf mergers and collisions
-due to computational constraints; our goal is to see how higher resolution runs differ.
-
-%==========================================================================
-% Resolution Dependence
-%==========================================================================
-\section{Spatial Resolution}
-\label{sec:spatialresolution}
-
-Now we turn to an examination of the effect of spatial resolution on the detonation.
-The first question we want to answer is: at reasonable uniform spatial resolution, is
-the simulation converged when a detonation occurs? If the answer is no, then the
-follow-up question is: can adaptive mesh refinement help achieve a converged simulation?
-
-To answer these questions we performed a series of simulations at varying base
-resolutions, augmented with additional runs with AMR. In each case we ran the
-simulation to a point just after the burning started, but before a detonation
-had occurred. Then we stopped
-the run and created a checkpoint. Starting from that checkpoint, we did a simulation
-at varying levels of refinement based on the unstable burning criterion. We will refer
-to the refinement through the refinement factor, $r$. This factor is the jump in
-resolution between the coarse grid and the finest grid. The uniform base grid case is $r = 1$.
-Meanwhile, $r = 16$ means a uniform coarse grid plus two levels of 4x refinement.
-
-At high resolution the collision time, $t_{\rm c} = 4.6\ \text{s}$, is later than the
-collision time at low resolution, due to a sharper view of the contact point. However,
-for the same reason, the detonation time is significantly earlier at $t_{\rm d} = 6.62\ \text{s}$,
-as the peak density reaches the key density range of $10^7\ \text{g\,/\,cm}^3$ significantly earlier,
-and this is what triggers the burning sufficient to generate an earlier detonation.
-Approximately speaking, the later $t_{\rm d}$ is, the more
-nickel is produced. As time passes, more white dwarf material slams into the center, compressing
-and heating up the gas. Both effects are favorable to production of nickel. So, a later $t_{\rm d}$
-implies that more material can be converted to nickel.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.425]{{{plots/amr}}}
-  \caption{Nickel production as a function of the simulation resolution. The solid
-           black curve connects the points at uniform resolution. For each uniform
-           resolution simulation, we include additional simulations using the adaptive
-           mesh refinement criterion described in \autoref{sec:unstable_burning}.
-           The horizontal axis is the effective number of zones per dimension on the
-           finest level, which is the number of zones per dimension that would exist
-           if the finest resolution was uniformly applied on the grid.
-           \label{fig:amr}}
-\end{figure}
+interval constraint. For simulations with resolution better than 64 km, we use a CFL
+number of 0.5. For the simulations with lower resolution, we use a CFL number of 0.01.
+This is because at low resolution, the timestep is so large that the detonation can
+form in a single timestep and represent such a large change that the burning integration
+fails (\autoref{sec:retries}). The timestep retry mechanism can handle this gracefully,
+but the subcycled timestep necessary to do so is usually so small that the full level
+timestep ends up being so long that it just makes more sense to avoid this situation
+entirely and use smaller level advance timesteps.
 
@@
 
+The authors thank Chris Malone and Don Willcox for useful discussions
+on the nature of explosive burning, and Doron Kushnir for providing
+clarification on the nature of the burning limiter used in \cite{kushnir:2013}.
+
 This research has made use of NASA's Astrophysics Data System 
-Bibliographic Services. In addition, this research has made use
-of the AstroBetter blog and wiki.
-
-\software{\amrex\ (\url{https://github.com/AMReX-Codes/Amrex}),
-          \castro\ \citep{castro} (\url{https://github.com/AMReX-Astro/Castro}),
-          \wdmerger\ \citep{wdmergerI} (\url{https://github.com/AMReX-Astro/wdmerger}),
-          GCC (\url{https://gcc.gnu.org/}),
-          python (\url{https://www.python.org/}),
-          matplotlib \citep{matplotlib} (\url{http://matplotlib.org/}),
-          yt \citep{yt} (\url{http://yt-project.org/})}
-
-\facilities{OLCF, NERSC}
+Bibliographic Services.
 

### 2018-11-22 papers/ignition/paper.tex
Update affiliation
Update abstract

@@
 {
-  NVIDIA Corporation \\
-  2788 San Tomas Expressway, Santa Clara, CA, 95051, USA
-}
-\affiliation
-{
-  Department of Physics and Astronomy, Stony Brook University, Stony Brook, NY, 11794-3800, USA
+  NVIDIA Corporation, 2788 San Tomas Expressway, Santa Clara, CA, 95051, USA
 }
@@
 Some simulations of Type Ia supernovae feature thermonuclear detonations that
-arise naturally during the evolution. It is usually not clear whether these
-detonations occur in the same time and place as they would in a simulation
-that fully resolved the burning length scale (assuming the resolved simulation
-even features a detonation). In this study we explore a number of the algorithmic
-choices that affect the outcome of reacting hydrodynamics simulations, and address
-the question of what is required to ensure that these simulations are properly
-converged. We focus on the spatial and temporal resolution, as well as the coupling
-between the hydrodynamics and the reactions. Adaptive mesh refinement, properly
-targeted at the relevant burning regions, works well for achieving converged simulations
-on a relatively coarse grid.
+arise self-consistently during the evolution. It is not clear whether these
+detonations would occur in a simulation that fully resolved the burning. In this
+study we examine a test detonation problem inspired by collisions of white dwarfs.
+This test problem demonstrates that achieving a converged thermonuclear ignition
+requires spatial resolution finer than 1 km in the burning region, and that
+the simulation as a whole likely needs a base resolution on the order of 1 km.
+Current computational resource constraints place this stringent resolution requirement
+out of reach for almost all multi-dimensional supernova simulations.
+Consequently, contemporary simulations that self-consistently demonstrate
+detonations are very likely not resolved and should not be trusted.
 \end{abstract}

### 2018-11-22 papers/ignition/paper.tex
Cleanup

@@
 is still an open question. Different models predict different locations for
-the detonation (e.g. in the center of a star or on the edge) and different
-mechanisms for initiating the event (e.g. a prompt detonation in a white dwarf (WD)
-collision versus a deflagration to detonation transition in the single degenerate
-model). Common to all of the cases is a severe lack of numerical resolution in
-the location where the detonation is expected to occur. The length and time scale
-at which a detonation forms is orders of magnitude smaller than the resolution that
-typical hydrodynamic simulations can achieve. Furthemore, the approximate equations
-we use at the resolution that we can achieve are still notoriously difficult to
-solve accurately. The mere presence of a detonation (or lack thereof) in a simulation
-is therefore only weak evidence regarding whether a detonation would truly occur.
+the detonation and different mechanisms for initiating the event. Common to all
+of the cases is a severe lack of numerical resolution in the location where the
+detonation is expected to occur. The length and time scale at which a detonation
+forms is orders of magnitude smaller than the resolution that typical multi-dimensional
+hydrodynamic simulations can achieve. The mere presence of a detonation (or lack thereof)
+in a simulation is therefore only weak evidence regarding whether a detonation would truly occur.
 
-In this study we examine the challenges associated with simulating nuclear detonations.
+In this study we examine the challenges associated with simulating thermonuclear detonations.
 The inspiration for this work comes from the literature on head-on collisions of WDs,
 which can occur, for example, in certain triple star systems \citep{thompson:2011,hamers:2013}.
-This problem is useful for understanding the problem of numerical simulation of nuclear
-burning. Head-on WD collisions rapidly convert a significant amount of kinetic energy
-into thermal energy in a small region and thus set up conditions ripe for a thermonuclear
-detonation.
+WD collisions rapidly convert a significant amount of kinetic energy into thermal energy and
+thus set up conditions ripe for a thermonuclear detonation. Collisions are interesting
+from the simulation perspective because of both this susceptibility to detonation and
+the relative ease in setting them up.
 
@@
 on physical conditions \citep{seitenzahl:2009,garg:2017} but is generally much
-smaller than 100 km. In this paper, we seek to find what numerical methods are required
-to achieve converged detonations.
+smaller than 100 km. In this paper, we attempt to find what simulation length scale
+is required to achieve converged thermonuclear ignitions.
 
-\section{Coupling Burning and Hydrodynamics}
-\label{sec:hydrocoupling}
-
-The numerical simulations in this paper are performed using the reacting hydrodynamics AMR
-code \castro\ \citep{castro}. \cite{wdmergerI} described updates to the core hydrodynamic
-algorithm since the code's original release. The goal of this section is to describe several
-limitations we found when coupling the hydrodynamics module to the reactions module for detonation
-problems, and the changes we made to the \castro\ code to deal with those limitations (though
-most of the ideas are applicable to any adaptive mesh-based code). The challenge of accurately
-capturing this coupling explains most of the difficulty in obtaining converged detonation
-simulations.
-
-\subsection{Timestep Size}
-\label{sec:timestep}
-
-In \castro, the reactions are coupled to the hydrodynamics using Strang splitting.
-Consider the basic outline of a single-level advance in an advection-reaction system:
-\begin{enumerate}
-  \item Evaluate timestep $\Delta t$ for the current advance
-  \item Advance the nuclear burning network by $\Delta t / 2$
-  \item Advance the hydrodynamics by $\Delta t$
-  \item Advance the nuclear burning network by $\Delta t / 2$
-  \item Return to Step 1
-\end{enumerate}
-The principal drawback of this approach is that the reactions and the hydrodynamics
-can become decoupled from each other. For example, consider the example of a head-on
-collision of white dwarfs. Initial nuclear burning will occur at the contact point
-between the two stars. Because of the staggered updates from splitting, the evolution
-effectively progresses as a cycle between burning for $\Delta t$ and getting fresh
-material advected into the contact point by the hydro update for $\Delta t$.
-When the collision begins, $\Delta t$ is controlled by the hydrodynamic stability
-criterion, and may be large enough that it is possible for the burning advance in
-Step 4 to completely burn the freshly advected material all the way to NSE.
-Consequently the evolution is no longer a good approximation to smooth burning of
-the in-falling material but rather separate discrete burning and hydro steps, and
-the nature of the burning evolution will be quite different. This can have a
-substantial effect on nucleosynthetic yields predicted by white dwarf simulations.
-
-%TODO: add a ``Nonaka plot''
-
-The size of the splitting error is a function of the advance timestep, and it
-decreases as the timestep becomes smaller. When we add AMR in the regions of
-nuclear burning, the hydrodynamic timestep (as set by the CFL criterion)
-automatically becomes smaller due to the increased resolution, which mitigates
-the splitting error. We have found that when we have sufficient refinement to
-achieve a converged simulation, the size of the hydrodynamic timestep is already
-small enough that the splitting error is minimal. For the test problem that we
-will describe later, the operator splitting has no observable effect on the
-physical effect we care about --- the time and location of the detonation --- at
-any resolution. Consequently we do not apply any special timestep limiter based
-on the rate of burning in this paper.
-
-However, for completeness, we describe our implementation of a common approach
-presented in the literature for timestep limiting based on burning, which should
-be used when performing low resolution SN Ia simulations \citep{raskin:2010,hawley:2012}.
-Defining the nuclear energy injection timescale $\tau_e$, and the species evolution
-timescale $\tau_{X_k}$,
-\begin{align}
-  \tau_e &\equiv \frac{e}{|\dot{e}|} \\
-  \tau_{X_k} &\equiv \frac{X_k}{|\dot{X_k}|},
-\end{align}
-where $\dot{e}$ is an estimate of the time rate of change of the internal energy
-from nuclear burning, and $\dot{X_k}$ is an estimate of the time rate of change 
-of the mass fraction of the species with index $k$, we define burning-limited 
-timesteps $\Delta t_{be}$ and $\Delta t_{bX_k}$:
-\begin{align}
-  \Delta t_{be} &= f_{be}\, \tau_e \label{eq:timestep_e}\\
-  \Delta t_{bX_k} &= f_{bX}\, \tau_{X_k}. \label{eq:timestep_X}
-\end{align}
-Given an estimate for $\dot{e}$, the factor $f_{be}$ determines by what 
-fraction we would like to allow the internal energy to change
-in the current timestep, under the assumption that $\dot{e}$ does not change from
-timestep to timestep. Similarly, given an estimate for $\dot{X_k}$, the factor $f_{bX}$ 
-determines the maximum change in the mass fraction of any species. To prevent species
-whose abundances are small in absolute terms from controlling the timestep, we apply
-the limiter based on $f_{bX}$ only to species with abundances in a given zone greater
-than $10^{-3}$ (this abundance threshold can be changed by the user of the code). By making 
-$f_{be}$ and $f_{bX}$ smaller, we can control the magnitude of the decoupling 
-between the reactions and the hydro. A typical choice for $f_{be}$ in the
-literature is in the range of 0.1 to 0.3, while to our knowledge a limiter based on
-$f_{bX}$ has not been used by others performing these types of calculations.
-This limiter is complementary to the limiter based on $f_{be}$, and can catch
-certain burning regimes that limiting only based on changes in energy would miss.
-
-At the start of each advance, we limit the size of the timestep to be the smaller
-of the minimum hydro timestep (limited by the CFL condition), and the minimum of all the
-burning timesteps across all zones. To do this, we need a method for determining 
-$\dot{e}$ and $\dot{X_k}$. We choose to do this with a direct evaluation of the
-right-hand-side of the burning network, given the current state at the time when
-the timestep is evaluated. Other work in the literature has typically evaluated
-this as a finite difference using the change over the last timestep. Our method
-is more accurate in principle, but in practice we find no meaningful difference
-between the limiting we propose and other methods of estimating the time derivatives.
-
-\subsection{Timestep Retries}
-\label{sec:retries}
-
-There are many circumstances during a timestep of a hydro code where the advance
-fails. As one common example, in \castro\ timesteps can be subcycled on fine levels,
-so that if the refinement factor between levels is 4, we take four timesteps on
-the fine level for every timestep on the coarse level, each with a quarter the
-size of the coarse level timestep. Then the coarse and fine levels can be
-synchronized \citep{bergercolella:1989}, which requires that we set the fine level
-timestep at the beginning of the first fine level advance and have to use the same
-timestep for each of the four advances. But it can occur that the velocity on the
-fine level grows during these timesteps, such that during the second or later of
-the four advances, the CFL criterion is no longer satisfied.
-
-Another example of a failure mode comes from the timestep limiting described in
-\autoref{sec:timestep}. The timestep will only actually satisfy the energy criterion
-$\Delta t \leq f_e \tau_e$ and species criterion $\Delta t \leq f_{X_k} \tau_{X_k}$
-when the estimates for $\dot{e}$ and $\dot{X_k}$ we generate are at least as large
-as the actual rate of change of energy and mass fractions over the timestep. However,
-this can assumption will sometimes fail during periods of runaway burning when the
-rate of change of these quantities is highly nonlinear. We may not want to neglect
-the errors caused by this approximation because they may build up over an extended
-period of nonlinear evolution and perhaps substantially change the final results.
-
-A third, related example comes from the burning module itself. The severe stiffness
-of the nuclear reaction ODEs means that we can sometimes run into integration failures.
-The most common cause of an integration failure is inability to reach the specified
-error tolerance within a specified maximum number of timesteps. These are rare in
-relative terms, but typically we find at least a few occur in any sufficiently long
-simulation, and we either have to deal with these or abort the simulation. In the
-\microphysics\ repository we have constructed a couple of ways of handling these
-failures internal to the burner: we can attempt to loosen the tolerance of the ODEs
-and try again; and, if that repeatedly fails, we can use a different integration method.
-However, we can be a little wiser about this and note that the failures are almost always
-ultimately related to the fact that the advance timestep is so large that the ODE
-integrator has to solve a very challenging problem. For a sufficiently small timestep,
-the size of the change that could occur during the burn would be similarly small,
-and it would be much easier for the integration to succeed.
-
-A natural solution to all of these problems is to be able to detect these failure
-modes, reject the proposed timestep, and subcycle the advance within a level with
-a smaller timestep. To this end, we have implemented a timestep retry option in
-\castro. If a failure is detected, the timestep is cut by a factor of 2 and we
-try again. If this does not work, the timestep is recursively cut until the
-advance succeeds. This approach neatly addresses several different types of
-problems in a single, simple framework.
-
-\subsection{Numerically Unstable Burning}
+\section{Numerically Unstable Burning}
 \label{sec:unstable_burning}
 
-\citet{kushnir:2013} point out that an inappropriate timestep is 
-not the only way for the numerical discretization to cause 
-severe errors in the burning. Another possible failure mode is when
-the energy injection timescale
-$\tau_e$ is shorter than the sound-crossing time $\tau_s$ in a zone.
+\citet{kushnir:2013} observe an important possible failure mode
+for reacting hydrodynamics simulations. Let us define $\tau_e = e / \dot{e}$
+as the nuclear energy injection timescale, and $\tau_s = \Delta x / c_s$
+as the sound-crossing time in a zone (where $\Delta x$ is the grid
+resolution and $c_s$ is the speed of sound).
 When the sound-crossing time is too long, energy is built up in
@@
 temperature building up too quickly in the zone; the detonation
-may be spurious in this case. If the following condition holds,
-we can be confident that a numerically seeded detonation has not occurred:
+may be spurious in this case. If $\tau_s \ll \tau_e$, we can be
+confident that a numerically seeded detonation has not occurred.
+In practice, we will quantify this as:
 \begin{equation}
@@
 \end{equation}
-The sound crossing time, $\tau_s$, is given by $\Delta x / c_s$, 
-where $c_s$ is the sound speed and $\Delta x$ is the (minimum) 
-zone width. The parameter $f_{s}$ then determines the minimum
-ratio of the nuclear energy generation timescale to the 
-sound-crossing time. \citet{kushnir:2013} choose $f_{s} = 0.1$ 
-for their simulations. \citet{kushnir:2013} implemented this
-criterion by artificially limiting the magnitude of the energy
+and require that $f_{s}$ is sufficiently smaller than one.
+\citet{kushnir:2013} state that $f_{s} = 0.1$ is a sufficient
+criterion. \citet{kushnir:2013} enforced this criterion on
+their simulations by artificially limiting the magnitude of the energy
 release after a burn.

### 2018-11-22 papers/ignition/paper.tex
Start describing the test problem

@@
 is required to achieve converged thermonuclear ignitions.
+
+%==========================================================================
+% 1D collision test problem
+%==========================================================================
+\section{Test Problem}
+\label{sec:collisions}
+
+Our test problem is inspired by \cite{kushnir:2013}, and very loosely approximates the
+conditions of two $0.64\ \msolar$ WDs colliding head-on. The simulation domain is 1D with a
+reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed (by mass)
+of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He. The fluid is relatively cold,
+$T = 10^7$ K, has density $\rho = 5 \times 10^6$ g/cm$^3$, and is traveling toward the
+origin with velocity $-2 \times 10^8$ m/s. A uniform constant gravitational acceleration
+is applied, $g = -1.1 \times 10^8$ m/s$^{2}$. This setup causes a sharp initial release
+of energy at $x = 0$, and the primary question is whether a detonation occurs promptly
+near this contact point, or occurs later (possibly at a distance from the contact point).
+The simulated domain has width $1.6384 \times 10^9$ cm, and we apply inflow boundary conditions
+that keep feeding the domain with material that has the same conditions as the initial fluid.
+Simulations are performed with the adaptive mesh refinement (AMR) code \castro\
+\citep{castro, astronum:2017}.
+
+The only parameter we vary in this study is the spatial resolution used for this problem.
+We vary both the base resolution of the grid, and also the amount of refinement allowed.
+AMR is applied on gradients of the temperature: we tag zones for refinement if the temperature
+varies by more than 50\% between two zones. Timesteps are limited only by the hydrodynamic
+stability constraint, with CFL number 0.5. Although this leads to Strang splitting error
+in the coupling of the burning and hydrodynamics for low resolution, we have verified that
+the incorrect results seen at low resolution do not meaningfully depend on the timestep constraint
+(both by applying a timestep limiter based on nuclear burning, and by using the spectral deferred
+corrections driver in \castro, which directly couples the burning and hydrodynamics). At very high
+resolution, the splitting error automatically becomes unimportant as the CFL criterion decreases
+the timestep.
 
@@
 %==========================================================================
-% 1D collision test problem
-%==========================================================================
-\section{Numerical Test Problem}
-\label{sec:collisions}
-
-Plotfiles of relevant quantities are created every 0.01 seconds.
-The timesteps are limited to ensure that every multiple of this plotfile interval is
-reached exactly, which means an effective maximum timestep of 0.01 seconds. Timesteps
-are by default limited only by the hydrodynamic stability constraint and this plotfile
-interval constraint. For simulations with resolution better than 64 km, we use a CFL
-number of 0.5. For the simulations with lower resolution, we use a CFL number of 0.01.
-This is because at low resolution, the timestep is so large that the detonation can
-form in a single timestep and represent such a large change that the burning integration
-fails (\autoref{sec:retries}). The timestep retry mechanism can handle this gracefully,
-but the subcycled timestep necessary to do so is usually so small that the full level
-timestep ends up being so long that it just makes more sense to avoid this situation
-entirely and use smaller level advance timesteps.
-
-%==========================================================================
 % Conclusions

### 2018-11-22 papers/ignition/paper.tex
Mention the network used
Add self-heating figure

@@
 Simulations are performed with the adaptive mesh refinement (AMR) code \castro\
-\citep{castro, astronum:2017}.
+\citep{castro, astronum:2017}. For the burning we use the alpha-chain nuclear network
+\texttt{aprox13}. The results we describe below do not strongly depend on the network.
 
@@
 the timestep.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.525]{{{plots/amr_ignition_self-heat}}}
+  \caption{Distance from the contact point of the ignition as a function of finest
+           spatial resolution. The solid black curve with circle markers represents
+           the simulations with uniform resolution (no AMR). The other curves represent
+           simulations with a fixed base resolution and varying amounts of AMR.
+           \label{fig:self-heat-distance}}
+\end{figure}
+
+\autoref{fig:self-heat-distance} shows our main results.
 

### 2018-11-22 papers/ignition/paper.tex
Describe stopping criterion

@@
 \texttt{aprox13}. The results we describe below do not strongly depend on the network.
+
+The simulation is terminated when the peak temperature on the domain first reaches
+$4 \times 10^9$ K, which we call a thermonuclear ignition. This stopping criterion is a
+proxy for the beginning of a detonation. Reaching this temperature does not guarantee
+that a detonation will begin, and in this study we do not directly address the question
+of whether a ignition of this kind always leads to a detonation. Rather, the main question
+we investigate here is whether this ignition is numerically converged. A converged ignition
+is a prerequisite to having a converged detonation.
 

### 2018-11-23 papers/ignition/paper.tex
Diagnostic quantities
More discussion of self-heating results

@@
 we investigate here is whether this ignition is numerically converged. A converged ignition
-is a prerequisite to having a converged detonation.
+is a prerequisite to having a converged detonation. We measure two diagnostic quantities:
+the time since the beginning of the simulation required to reach this ignition criterion,
+and the distance from the contact point of the peak temperature.
 
@@
 
-\autoref{fig:self-heat-distance} shows our main results.
+\autoref{fig:self-heat-distance} shows our main results. The lowest resolution we consider,
+256 km, is typical of the early simulations of white dwarf collisions, and demonstrates a
+prompt ignition near the contact point. As the (uniform) resolution increases, the ignition
+tends to occur earlier and nearer to the contact point. This trend is not physically meaningful:
+all simulations with resolution worse than about 1 km represent the same prompt central ignition,
+and as the resolution increases, there are grid points physically closer to the center that can ignite.
+However, when the resolution is better than 1 km, the situation changes dramatically: the prompt
+central ignition does not occur, but rather the ignition is delayed and occurs further from the contact
+point. When we have finally converged, the ignition occurs around 1000 km from the contact point,
+about 1 second after contact (contrast to less than 0.05 seconds for the simulation with 1 km resolution).
+This story contains two important lessons. First, the required resolution for a converged simulation,
+less than 100 m, is out of reach of current 3D supernova simulations. Second, the behavior for
+resolutions worse than 1 km qualitatively appears to be converged, and one could perhaps
+be misled into thinking that there was no reason to try higher resolutions, which is reason
+for caution in interpreting reacting hydrodynamics simulations.
+
+Also in \autoref{fig:self-heat-distance} we show results for a few cases where we add AMR
+using the criterion described above. The main takeaway here is that for coarse grid resolutions
+worse than 1 km, the addition of AMR yields unreliable results: the coarse grid's approximation
+of the physics must simply not be good enough for refinement to do any good. This parallels the
+fact that the trend of convergence towards the correct delayed detonation does not begin until
+resolutions better than 1 km. However, for a coarse grid resolution of 1 km, AMR does yield
+convergence to the correct answer, and at a much cheaper computational cost than the equivalent
+uniform grid simulation\marginpar{quantify this}, so AMR does add value if the resolution on the
+coarse grid is sufficiently high.
 
@@
 
+% Add discussion about how none of the conclusions depend on the microphysics
+% of how ignitions/detonations actually occur -- this paper is simply about
+% self-consistent fluid simulations, i.e. are we even correctly solving
+% the equations we wrote down, and not focusing on the harder question
+% of whether we're solving the right equations.
+
 \acknowledgments

### 2018-11-23 papers/ignition/paper.tex
Update discussion of suppressed burning limiter

@@
 as the sound-crossing time in a zone (where $\Delta x$ is the grid
-resolution and $c_s$ is the speed of sound).
-When the sound-crossing time is too long, energy is built up in
-a zone faster than it can be advected away by pressure waves.
-This is of course a problem inherent only to numerically discretized
-systems as the underlying fluid equations are continuous.
-This can lead to a numerically seeded detonation caused by the
-temperature building up too quickly in the zone; the detonation
-may be spurious in this case. If $\tau_s \ll \tau_e$, we can be
-confident that a numerically seeded detonation has not occurred.
-In practice, we will quantify this as:
+resolution and $c_s$ is the speed of sound). When the sound-crossing
+time is too long, energy is built up in a zone faster than it can be
+advected away by pressure waves. This is of course a problem inherent
+only to numerically discretized systems as the underlying fluid equations
+are continuous. This can lead to a numerically seeded detonation
+caused by the temperature building up too quickly in the zone; the
+detonation may be spurious in this case. If $\tau_s \ll \tau_e$,
+we can be confident that a numerically seeded detonation has not
+occurred. In practice, we will quantify this requirement as:
 \begin{equation}
-  \tau_s \leq f_{s}\, \tau_e \label{eq:burning_limiter_2}
+  \tau_s \leq f_{s}\, \tau_e \label{eq:burning_limiter}
 \end{equation}
@@
 \citet{kushnir:2013} state that $f_{s} = 0.1$ is a sufficient
-criterion. \citet{kushnir:2013} enforced this criterion on
-their simulations by artificially limiting the magnitude of the energy
-release after a burn.
+criterion for avoiding premature ignitions. \citeauthor{kushnir:2013}
+enforced this criterion on their simulations by artificially limiting
+the magnitude of the energy release after a burn, and claimed that
+this is resulted in more accurate WD collision simulations.
 
-To understand this choice, consider the outcome of our baseline case.
-When the two WDs collide, there is a prompt detonation
-at the contact point at the center of the domain. This detonation occurs
-when the density is relatively low because the WDs have not had much
-time to collide, and so the nucleosynthetic yield of nickel is low. If
-this is the case, the colliding WDs are not a good candidate for
-normal SNe Ia. However, if this detonation is delayed, material builds
-up at a stalled shock in the center, and the eventual detonation occurs
-on the outer edge of the stalled shock. There will then be much more
-high density material to process into nickel, and the resulting
-explosion creates an amount of nickel potentially sufficient to explain
-normal SNe Ia. The discussion thus centers on whether the prompt detonation
-is physical. \citeauthor{kushnir:2013} argues that it is not, on the basis
-that $\tau_s > \tau_e$ when this detonation occurs. We find the same thing:
-for the baseline case, the sound-crossing time is about five times longer
-than the energy injection time. This means that an ignition is guaranteed
-to occur for numerical reasons. \citeauthor{kushnir:2013} thus argue that
-the suppressed burn is justifiable, as it suppresses this prompt detonation.
+We find that for our test problem (and also the WD collisions we have simulated)
+we do observe $\tau_s > \tau_e$; typically the ratio is a factor of 2--5 at
+low resolution. This means that an ignition is very likely to occur for numerical
+reasons, regardless of whether it would occur for physical reasons. We thus
+investigate whether limiting the energy release of the burn (we will term this
+``suppressing'' the burn) is a useful technique for avoiding the prompt detonation.
 
-This argument is flawed because a physical detonation may \textit{also}
-occur this way. For example, consider a region of WD material at uniformly
+Before we examine the results, consider a flaw in the reasoning behind the limiter:
+a physical detonation may \textit{also} occur with the property that, in the detonating
+zone, $\tau_s > \tau_e$. For example, consider a region of WD material at uniformly
 high temperature, say $5 \times 10^9\ \text{K}$, with an arbitrarily large size,
@@
 similar results, for example because even at the higher resolution, the physical
-size of the hotspot stays the same. Additionally, one might think that simply
-adding adaptive mesh refinement in the burning region could help, because then
-$\Delta x$ decreases, so the ratio of $\tau_s$ to $\tau_e$ should decrease.
-We developed an AMR criterion to do just this: we tag on regions suggested by
-the burning stability criterion, \autoref{eq:burning_limiter_2}. But if a low
-resolution simulation develops a very large hotspot that violates the stability
-criterion, adding refinement inside this region will likely not help, because the
-energy trapped in the center of the region may not be able to escape before ignition
-occurs. So for the simulations with resolution inside the stars that was coarser than
-10 km, even adding a significant amount of refinement (say, refining by a factor of
-64 or 128) did not change the propagation of a detonation.
+size of the hotspot stays the same. For this reason, an appeal to the numerical
+instability criterion alone is insufficient to understand whether a given ignition
+is real.
 
-For the above reasons, an appeal to the numerical instability criterion alone is
-insufficient to understand whether a given ignition is real. We find that \textit{all}
-of the detonations we have seen at any resolution violate the stability criterion,
-even for the finest resolution simulation we performed (150 m, inside the burning region).
-This is true even if the prompt detonation is suppressed:
-the delayed detonation also is numerically unstable. For comparison purposes, we
-developed an option for our burner to use a ``suppressed'' burning mode. In a
-suppressed burn, we limit the changes to the state so that \autoref{eq:burning_limiter_2}
-is always satisfied\footnote{To achieve this we directly multiply the
-right-hand-side vector in the integration by a factor $F$
-for all variables, where $F$ is the multiplicative factor needed to
-be applied to $\dot{e}$ such that the equality in \autoref{eq:burning_limiter_2}
-holds. (If the inequality is already satisfied, then the integration
-vector is not modified.) We fix $\tau_s$ to be the value of the sound
-crossing time at the beginning of the burn (that is, we do not
-update it as the sound speed changes) and we fix the energy $e$
-that goes into the estimate for $\tau_e$ to be the value of the
-internal energy of the zone at the beginning of the burn.}. When we
-evolve the baseline case with this suppressed burner, the prompt detonation
-is indeed suppressed. A later detonation does occur, but this delayed
-detonation also triggers the suppression. The result is that the detonation
-does not get to very high temperatures, and the result is again a ``failed''
-supernova that does not generate nearly enough nickel. Thus we do not
-reproduce the claim of \citeauthor{kushnir:2013} that the suppressed
-burning mode suppresses only the prompt detonation, but allows the
-delayed detonation to occur. We could find some way to turn off the
-suppression and allow the delayed detonation, but without some independent
-confirmation that the delayed detonation is physical (while the prompt
-detonation is not) this would be begging the question.
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.30]{{{plots/amr_ignition_suppressed}}}
+  \caption{Similar to \autoref{fig:self-heat-distance}, but for simulations with the
+           suppressed burning limiter applied (\autoref{eq:burning_limiter}). Only
+           simulations without AMR are shown.
+           \label{fig:suppressed-distance}}
+\end{figure}
 
-The upshot of this discussion is that the only way to resolve the
-question is with higher resolution simulations. The semi-analytic
-criterion provided by \cite{garg:2017}, and the numerical calculations
-of \cite{seitenzahl:2009}, give critical radii for spontaneous detonation.
-For the densities we are interested in ($\sim 10^7\ \text{g\,/\,cm}^3$),
-assuming equal C/O material, the critical radii are in the range 1-10 km.
-So this suggests that our simulation needs to do significantly better than
-this if we are likely to make headway on this issue. Detonations observed
-at higher resolution may be physically plausible because the critical
-temperature gradient can be resolved. The high resolution needs to be
-achieved with some combination of high resolution of the collision process
-itself, and AMR based on the stability criterion as described
-above, where we tag for refinement all zones above some $f_{s}$, the
-ratio of the sound crossing time to the energy injection time. We
-cannot solely rely on a low resolution run with AMR based on the
-burning stability criterion for the reason described above, that a
-detonation may be spuriously locked in by the low resolution regardless
-of how much resolution we can practicably add to it.
+\autoref{fig:suppressed-distance} shows the results we obtain for our implementation of
+a ``suppressed'' burning mode. In a suppressed burn, we limit the changes to the state
+so that \autoref{eq:burning_limiter} is always satisfied. This is done by rescaling the
+energy release and species changes from a burn by a common factor such that the equality
+in \autoref{eq:burning_limiter} is satisfied. (If the inequality is already satisfied,
+then the integration vector is not modified.) We find that the suppressed burn
+generally does not yield correct results for low resolutions. The 64 km resolution
+simulation happens to yield approximately the correct ignition distance, but it does
+not occur at the right time, and in any case the incorrectness of the results at neighboring
+resolutions suggests that this is not a robust finding. Because of both the theoretical reasons
+discussed above, and this empirical finding, we do not believe that the suppressed burning
+limiter should be applied in production simulations.
 
@@
 
-\clearpage
-
 \bibliographystyle{../aasjournal}

### 2018-12-02 papers/ignition/paper.tex

@@
 
-% Add discussion about how none of the conclusions depend on the microphysics
-% of how ignitions/detonations actually occur -- this paper is simply about
-% self-consistent fluid simulations, i.e. are we even correctly solving
-% the equations we wrote down, and not focusing on the harder question
-% of whether we're solving the right equations.
+Our example detonation problem demonstrates, at least for this class of
+hydrodynamical burning problem, a grid resolution requirement much more stringent
+than 1 km. This does not represent all such burning problems: for example, we
+have tested similar setups with only carbon and oxygen for which a resolution
+in the tens of kilometers is sufficient. However, the fact that it is even
+possible for this type of burning to require a resolution better than 100 m
+should cast significant uncertainty on the results of simulations featuring
+detonations that do not meet this resolution.
+
+This study does not directly address the problem of how, in the detailed
+microphysical sense, a detonation wave actually begins to propagate, as
+we cannot resolve this length scale even in our highest resolution simulations.
+Rather, we are making the point that for simulations in which a macroscopic
+detonation wave appears self-consistently, this is only a valid numerical result
+if the resolution is sufficiently high. This convergence requirement does
+not imply that the detonation itself is physically realistic; but, it does
+imply that we are not even correctly solving the fluid equations we intend
+to solve when the convergence requirement is not met. We believe that our
+test case can be useful in the future for testing algorithmic innovations
+that hope to improve the realism of burning at low resolutions.
 

### 2018-12-02 science/ignition/analyze.py
Add a plot of ignition time

@@ -113,43 +113,55 @@ def amr_ignition(file_base, results_base, do_amr = True):
             dist_lists[i] = dist_list
             time_lists[i] = time_list
 
-    has_plot = False
+    # Make two plots: location of the ignition, and time.
 
-    # First add all plots with refinement.
+    for plot_var in ["distance", "time"]:
 
-    if do_amr:
-        plt_count = 0
-        for i, (res_list, dist_list) in enumerate(zip(res_lists, dist_lists)):
-            if len(res_list) > 1:
+        has_plot = False
+
+        if plot_var is "distance":
+            var_lists = dist_lists
+        else:
+            var_lists = time_lists
+
+        # First add all plots with refinement.
+
+        if do_amr:
+            plt_count = 0
+            for i, (res_list, var_list) in enumerate(zip(res_lists, var_lists)):
+                if len(res_list) > 1:
+                    has_plot = True
+                    lbl = '{} km base + AMR'.format(int(base_res_list[i]))
+                    plt.plot(res_list, var_list, linestyle=linestyles[plt_count], marker=markers[plt_count], lw=2.0, label=lbl)
+                    plt_count += 1
+
+        # Now add the base refinement cases.
+
+        coarse_list = []
+        r_list = []
+        for i, ncell in enumerate(ncell_list):
+            if len(var_lists[i]) > 0:
                 has_plot = True
-                lbl = '{} km base + AMR'.format(int(base_res_list[i]))
-                plt.plot(res_list, dist_list, linestyle=linestyles[plt_count], marker=markers[plt_count], lw=2.0, label=lbl)
-                plt_count += 1
+                r_list.append(res_lists[i][0])
+                coarse_list.append(var_lists[i][0])
 
-    # Now add the base refinement cases.
+        if len(coarse_list) > 0:
+            lbl = 'Uniform grid'
+            plt.plot(r_list, coarse_list, color='k', lw=2.0, label=lbl, marker='o', markersize=12)
 
-    coarse_list = []
-    r_list = []
-    for i, ncell in enumerate(ncell_list):
-        if len(dist_lists[i]) > 0:
-            has_plot = True
-            r_list.append(res_lists[i][0])
-            coarse_list.append(dist_lists[i][0])
-
-    if len(coarse_list) > 0:
-        lbl = 'Uniform grid'
[…]

### 2019-01-23 science/ignition/analyze.py
Add axis limits

@@ -138,12 +138,14 @@ def amr_ignition(file_base, results_base, do_amr = True):
     ax1.tick_params(axis='y', labelcolor=colors[0], labelsize=16)
     ax1.tick_params(axis='x', labelsize=16)
     ax1.set_ylim(bottom=0.0)
+    ax1.set_ylim(top=1100.0)
 
     ax2 = ax1.twinx()
     ax2.plot(res_list, time_list, linestyle=linestyles[1], marker=markers[1], color=colors[1], lw=2.0, label='Ignition time')
     ax2.set_ylabel(r"Ignition time (s)", fontsize=24, color=colors[1])
     ax2.tick_params(axis='y', labelcolor=colors[1], labelsize=16)
     ax2.set_ylim(bottom=0.0)
+    ax2.set_ylim(top=1.1)
 
     fig.set_size_inches(11, 8.5)
     fig.tight_layout()

### 2019-01-23 papers/ignition/paper.tex
Update plots

@@
 This test problem demonstrates that achieving a converged thermonuclear ignition
-requires spatial resolution finer than 1 km in the burning region, and that
-the simulation as a whole likely needs a base resolution on the order of 1 km.
+requires spatial resolution much finer than 1 km in the burning region, and that
+the simulation as a whole likely needs a base resolution on the order of 1 km or better.
 Current computational resource constraints place this stringent resolution requirement
@@
 Consequently, contemporary simulations that self-consistently demonstrate
-detonations are very likely not resolved and should not be trusted.
+detonations are very likely not resolved and should be treated with caution.
 \end{abstract}
@@
   \centering
-  \includegraphics[scale=0.525]{{{plots/amr_ignition_self-heat}}}
+  \includegraphics[scale=0.30]{{{plots/amr_ignition_self-heat}}}
   \caption{Distance from the contact point of the ignition as a function of finest
@@
 central ignition does not occur, but rather the ignition is delayed and occurs further from the contact
-point. When we have finally converged, the ignition occurs around 1000 km from the contact point,
+point. When we have finally converged, the ignition occurs around 900 km from the contact point,
 about 1 second after contact (contrast to less than 0.05 seconds for the simulation with 1 km resolution).

### 2019-02-05 papers/ignition/paper.tex
add facility keyword
Abstract

@@
 \begin{abstract}
-Some simulations of Type Ia supernovae feature thermonuclear detonations that
-arise self-consistently during the evolution. It is not clear whether these
-detonations would occur in a simulation that fully resolved the burning. In this
+Some simulations of Type Ia supernovae feature self-consistent thermonuclear
+detonations. However, these detonations are not meaningful if the simulations
+are not resolved, so it is important to establish an understanding of the
+requirements for achieving a numerically converged detonation. In this
 study we examine a test detonation problem inspired by collisions of white dwarfs.
 This test problem demonstrates that achieving a converged thermonuclear ignition
-requires spatial resolution much finer than 1 km in the burning region, and that
-the simulation as a whole likely needs a base resolution on the order of 1 km or better.
-Current computational resource constraints place this stringent resolution requirement
-out of reach for almost all multi-dimensional supernova simulations.
+requires spatial resolution much finer than 1 km in the burning region.
+Current computational resource constraints place this stringent resolution
+requirement out of reach for multi-dimensional supernova simulations.
 Consequently, contemporary simulations that self-consistently demonstrate
@@
 
+\facility{OLCF}
+
 \bibliographystyle{../aasjournal}

### 2019-02-05 papers/ignition/paper.tex
Final pass before initial submission

@@
 WD collisions rapidly convert a significant amount of kinetic energy into thermal energy and
-thus set up conditions ripe for a thermonuclear detonation. Collisions are interesting
-from the simulation perspective because of both this susceptibility to detonation and
-the relative ease in setting them up.
+thus set up conditions ripe for a thermonuclear detonation. Since they are easy to set up,
+they are a useful vehicle for studying the properties of detonations.
 
@@
 \texttt{aprox13}. The results we describe below do not strongly depend on the network.
+Release 18.12 of the \castro\ code was used. The \amrex\ and \microphysics\ repositories
+that \castro\ depends on were also on release 18.12. The problem is located in the
+\texttt{Exec/science/Detonation} directory, and we used the \texttt{inputs-collision} setup.
 
@@
 The only parameter we vary in this study is the spatial resolution used for this problem.
-We vary both the base resolution of the grid, and also the amount of refinement allowed.
-AMR is applied on gradients of the temperature: we tag zones for refinement if the temperature
+For low resolutions we vary only the base resolution of the grid, up to a resolution of
+0.25 km. For resolutions finer than this, we fix the base grid at a resolution of 0.25 km,
+and use AMR applied on gradients of the temperature. We tag zones for refinement if the temperature
 varies by more than 50\% between two zones. Timesteps are limited only by the hydrodynamic
@@
 corrections driver in \castro, which directly couples the burning and hydrodynamics). At very high
-resolution, the splitting error automatically becomes unimportant as the CFL criterion decreases
-the timestep.
+resolution, the splitting error tends to zero as the CFL criterion decreases the timestep.
 
@@
   \includegraphics[scale=0.30]{{{plots/amr_ignition_self-heat}}}
-  \caption{Distance from the contact point of the ignition as a function of finest
-           spatial resolution. The solid black curve with circle markers represents
-           the simulations with uniform resolution (no AMR). The other curves represent
-           simulations with a fixed base resolution and varying amounts of AMR.
+  \caption{Distance from the contact point of the ignition (solid blue), and time of the
+           ignition (dashed green), as a function of finest spatial resolution.
            \label{fig:self-heat-distance}}
@@
 central ignition does not occur, but rather the ignition is delayed and occurs further from the contact
-point. When we have finally converged, the ignition occurs around 900 km from the contact point,
+point. When we have finally (perhaps) converged, the ignition occurs around 900 km from the contact point,
 about 1 second after contact (contrast to less than 0.05 seconds for the simulation with 1 km resolution).
-This story contains two important lessons. First, the required resolution for a converged simulation,
-less than 100 m, is out of reach of current 3D supernova simulations. Second, the behavior for
-resolutions worse than 1 km qualitatively appears to be converged, and one could perhaps
-be misled into thinking that there was no reason to try higher resolutions, which is reason
-for caution in interpreting reacting hydrodynamics simulations.
-
-Also in \autoref{fig:self-heat-distance} we show results for a few cases where we add AMR
-using the criterion described above. The main takeaway here is that for coarse grid resolutions
-worse than 1 km, the addition of AMR yields unreliable results: the coarse grid's approximation
-of the physics must simply not be good enough for refinement to do any good. This parallels the
-fact that the trend of convergence towards the correct delayed detonation does not begin until
-resolutions better than 1 km. However, for a coarse grid resolution of 1 km, AMR does yield
-convergence to the correct answer, and at a much cheaper computational cost than the equivalent
-uniform grid simulation\marginpar{quantify this}, so AMR does add value if the resolution on the
-coarse grid is sufficiently high.
+This story contains two important lessons. First, the required resolution for even a qualitatively
+converged simulation, less than 100 m, is out of reach of current 3D supernova simulations.
+Second, the behavior for resolutions worse than 1 km qualitatively appears to be converged,
+and one could perhaps be misled into thinking that there was no reason to try higher resolutions,
+which is reason for caution in interpreting reacting hydrodynamics simulations.
 
@@
 investigate whether limiting the energy release of the burn (we will term this
-``suppressing'' the burn) is a useful technique for avoiding the prompt detonation.
+``suppressing'' the burn), as proposed by \citeauthor{kushnir:2013},
+is a useful technique for avoiding the prompt detonation.
 
-Before we examine the results, consider a flaw in the reasoning behind the limiter:
+Before we examine the results, consider a flaw in the application of the limiter:
 a physical detonation may \textit{also} occur with the property that, in the detonating
@@
   \caption{Similar to \autoref{fig:self-heat-distance}, but for simulations with the
-           suppressed burning limiter applied (\autoref{eq:burning_limiter}). Only
-           simulations without AMR are shown.
+           suppressed burning limiter applied (\autoref{eq:burning_limiter}).
            \label{fig:suppressed-distance}}
@@
 not occur at the right time, and in any case the incorrectness of the results at neighboring
-resolutions suggests that this is not a robust finding. Because of both the theoretical reasons
-discussed above, and this empirical finding, we do not believe that the suppressed burning
-limiter should be applied in production simulations.
+resolutions suggests that this is not a robust finding. The suppressed burning simulation
+reaches qualitative convergence at around the same 100 m resolution as the normal self-heating
+burn. Because of both the theoretical reasons discussed above, and this empirical finding that
+the burning suppression does not make low resolution simulations any more accurate, we do not
+believe that the suppressed burning limiter should be applied in production simulations.
 

### 2019-02-05 papers/ignition/submit.sh
Add submitted version

@@ -1,0 +1,31 @@
+submit_dir="submit"
+
+# Build the PDF so that we have the bibilography
+
+make clean
+make
+
+# Move all the source files to the submit directory
+
+mkdir -p $submit_dir
+
+cp -f paper.tex $submit_dir
+cp -f paper.bbl $submit_dir
+cp -f plots/*.eps $submit_dir
+cp -f ../aasjournal.bst $submit_dir
+cp -f ../aastex62.cls $submit_dir
+
+# Get rid of all references to the plots subdirectory
+
+sed -i "s/plots\///" $submit_dir/paper.tex
+
+# Get rid of all references to the parent directory
+
+sed -i "s/\.\.\///" $submit_dir/paper.tex
+
+# Replace the bibliography command with a direct input
+# of the bibliography file
+
+sed -i "s/bibliography{refs}/input{paper.bbl}/" $submit_dir/paper.tex
+
+tar -czvf $submit_dir.tar $submit_dir

### 2019-02-06 papers/ignition/paper.tex
Add trackchanges
Some tracked changes in response to review

@@
-\documentclass[twocolumn,numberedappendix]{../aastex62}
+\documentclass[twocolumn,numberedappendix,trackchanges]{../aastex62}
 
@@
 on physical conditions \citep{seitenzahl:2009,garg:2017} but is generally much
-smaller than 100 km. In this paper, we attempt to find what simulation length scale
-is required to achieve converged thermonuclear ignitions.
+smaller than 100 km.
+
+In this paper, we attempt to find what simulation length scale is required to achieve
+converged thermonuclear ignitions. \added{In our two-dimensional axisymmetric simulations
+of WD collisions, using the Eulerian reactive hydrodynamics code \castro\ \citep{castro, astronum:2017},
+the simulations are unresolved at the highest resolution we could afford to run (0.25 km).
+So, in this paper we turn to a 1D test, which allows us to test at much higher resolution.}
 
@@
 that keep feeding the domain with material that has the same conditions as the initial fluid.
-Simulations are performed with the adaptive mesh refinement (AMR) code \castro\
-\citep{castro, astronum:2017}. For the burning we use the alpha-chain nuclear network
-\texttt{aprox13}. The results we describe below do not strongly depend on the network.
+Simulations are performed with the adaptive mesh refinement (AMR) code \castro.
+For the burning we use the alpha-chain nuclear network
+\texttt{aprox13}. \replaced{The results we describe below do not strongly depend on the network.}
+{We have tried with a few (relatively small) networks, and the results we describe below hold
+for all the ones we tested.}
 Release 18.12 of the \castro\ code was used. The \amrex\ and \microphysics\ repositories
@@
 that a detonation will begin, and in this study we do not directly address the question
-of whether a ignition of this kind always leads to a detonation. Rather, the main question
-we investigate here is whether this ignition is numerically converged. A converged ignition
+of whether a ignition of this kind always leads to a detonation. \added{Nor are we commenting
+on the physics of the ignition process itself.} Rather, the main question
+we investigate here is whether this ignition is numerically converged, \added{and for this purpose
+this arbitrary stopping point is sufficient, since in a converged simulation the stopping point
+should be reached at the same time independent of resolution}. A converged ignition
 is a prerequisite to having a converged detonation. We measure two diagnostic quantities:
@@
 central ignition does not occur, but rather the ignition is delayed and occurs further from the contact
-point. When we have finally (perhaps) converged, the ignition occurs around 900 km from the contact point,
-about 1 second after contact (contrast to less than 0.05 seconds for the simulation with 1 km resolution).
-This story contains two important lessons. First, the required resolution for even a qualitatively
-converged simulation, less than 100 m, is out of reach of current 3D supernova simulations.
-Second, the behavior for resolutions worse than 1 km qualitatively appears to be converged,
-and one could perhaps be misled into thinking that there was no reason to try higher resolutions,
-which is reason for caution in interpreting reacting hydrodynamics simulations.
+point. When we have finally \replaced{(perhaps) converged}{reached the point where the curves start to flatten},
+the ignition occurs around 900 km from the contact point, about 1 second after contact (contrast to less than
+0.05 seconds for the simulation with 1 km resolution). \added{Even at this resolution, it is not clear if
+the simulation is converged. We were unable to perform higher resolution simulations to check convergence
+due to the length of time that would be required.} This story contains two important lessons. First,
+the required resolution for even a qualitatively converged simulation, less than 100 m, is out of reach
+of current 3D supernova simulations. Second, the behavior for resolutions worse than 1 km qualitatively
+appears to be converged, and one could perhaps be misled into thinking that there was no reason to
+try higher resolutions, which is reason for caution in interpreting reacting hydrodynamics simulations.
 
@@
 time is too long, energy is built up in a zone faster than it can be
-advected away by pressure waves. This is of course a problem inherent
+advected away by pressure waves. \added{This effect generalizes to
+Langrangian simulations as well, where $\tau_s$ should be understood
+as the timescale for transport of energy to a neighboring fluid element.}
+This is of course a problem inherent 
 only to numerically discretized systems as the underlying fluid equations
@@
 low resolution. This means that an ignition is very likely to occur for numerical
-reasons, regardless of whether it would occur for physical reasons. We thus
-investigate whether limiting the energy release of the burn (we will term this
+reasons, regardless of whether it would occur for physical reasons. \added{At low
+resolution, adding more resolution does not meaningfully improve the ratio of
+$\tau_s$ to $\tau_e$ at the point of ignition. The ignition timescale is so short
+that virtually all of the energy release occurs in a single timestep even if the
+timestep gets shorter due to the CFL limiter. It is only when the resolution gets
+sufficiently high that we can simultaneously resolve the energy release over multiple
+timesteps and the advection of energy across multiple zones. Even at the highest resolution
+we could achieve, about 50 cm, $\tau_s / \tau_e$ was 0.8 at ignition, which is not
+sufficiently small to be confident of numerical stability at that resolution.}
+We thus investigate whether limiting the energy release of the burn (we will term this
 ``suppressing'' the burn), as proposed by \citeauthor{kushnir:2013},
-is a useful technique for avoiding the prompt detonation.
+is a useful technique for avoiding the prompt detonation. \added{Since the limiter ensures
+the inequality in \autoref{eq:burning_limiter} holds by construction, the specific
+question to ask is whether the limiter achieves the correct answer and is converged
+in cases where the simulation would otherwise be uncorrect or unconverged.}
 
@@
 
+\added{To summarize, then, satisfying the stability criterion is neither necessary nor
+sufficient, in general, for numerical convergence. It is not necessary because of the
+counter-example provided above of a very large hotspot, and it is not sufficient, as the
+empirical example in \autoref{fig:suppressed-distance} demonstrates. Nevertheless,
+it is a very useful diagnostic metric for understanding simulations, and it is likely
+that in many real cases, violation of the stability criterion is a clear sign of
+the simulation being under-resolved.}
+
 %==========================================================================
@@
 hydrodynamical burning problem, a grid resolution requirement much more stringent
-than 1 km. This does not represent all such burning problems: for example, we
+than 1 km. This does not represent all such \added{1D} burning problems: for example, we
 have tested similar setups with only carbon and oxygen for which a resolution
-in the tens of kilometers is sufficient. However, the fact that it is even
-possible for this type of burning to require a resolution better than 100 m
-should cast significant uncertainty on the results of simulations featuring
-detonations that do not meet this resolution.
+in the tens of kilometers is sufficient. \added{However, this should not be construed
+as implying that this convergence requirement does not apply to pure C/O WDs.}
+\replaced{However}{First}, the fact that it is even possible for
+\replaced{this type of burning}{burning in white dwarf material} to require a
+resolution better than 100 m should cast significant uncertainty on the results
+of simulations featuring detonations that do not meet this resolution
+\added{(regardless of the specific conditions). Or, at least, should suggest
+that stronger demonstrations of convergence are required. Second, in our real
+multi-dimensional simulations of C/O WD collisions, we see the same effect,
+where the simulation is not converged even for resolutions up to 0.25 km. It
+simply happens to be the case that for the 1D analogue of this problem using pure
+C/O the same effect is not seen, which is not particularly surprising since the 1D
+setup and the 2D setup are very different and the former lacks any multidimensional
+hydrodynamic effects. So we found a related problem that shows a similar
+effect in 1D to test convergence at higher resolutions.}
 

### 2019-02-06 papers/ignition/paper.tex
Remove comment about network dependence

@@
 For the burning we use the alpha-chain nuclear network
-\texttt{aprox13}. \replaced{The results we describe below do not strongly depend on the network.}
-{We have tried with a few (relatively small) networks, and the results we describe below hold
-for all the ones we tested.}
+\texttt{aprox13}. \deleted{The results we describe below do not strongly depend on the network.}
 Release 18.12 of the \castro\ code was used. The \amrex\ and \microphysics\ repositories

### 2019-02-06 papers/ignition/paper.tex
Add comment about effective SPH spatial resolution

@@
 hawley:2012,garcia-senz:2013} typically had effective spatial resolutions in the
-range 100--500 km, and observed detonations that convert a large amount of
+range 100--500 km \added{for the grid codes, and 10--100 km for the SPH codes}, and observed detonations that convert a large amount of
 carbon/oxygen material into iron-group elements. These studies varied in methodology
@@
 scale over which detonations form. The detonation length scale can vary widely based
-on physical conditions \citep{seitenzahl:2009,garg:2017} but is generally much
-smaller than 100 km.
+on physical conditions \citep{seitenzahl:2009,garg:2017} but is generally \replaced{much
+smaller than 100 km}{not larger than 10 km}.
 

### 2019-02-06 papers/ignition/paper.tex
Add comment about helium in initial conditions

@@
 In this paper, we attempt to find what simulation length scale is required to achieve
-converged thermonuclear ignitions. \added{In our two-dimensional axisymmetric simulations
-of WD collisions, using the Eulerian reactive hydrodynamics code \castro\ \citep{castro, astronum:2017},
-the simulations are unresolved at the highest resolution we could afford to run (0.25 km).
-So, in this paper we turn to a 1D test, which allows us to test at much higher resolution.}
+converged thermonuclear ignitions. \added{The inspiration for this work comes from our
+simulations of WD collisions using the reactive hydrodynamics code \castro\
+\citep{castro, astronum:2017}. We have done both 2D axisymmetric and 3D simulations
+of collisions of $0.64\ \msolar$ carbon/oxygen WDs, and we were unable to achieved converged
+simulations at any resolution we could afford to run (the best was an effective zone size of
+0.25 km, using adaptive mesh refinement, for the 2D case). We were therefore forced to
+turn to 1D simulations, where we can achieve much higher resolution (at the cost, of course,
+of not being able to do a test that can be directly compared to multi-dimensional simulations).}
 
@@
 conditions of two $0.64\ \msolar$ WDs colliding head-on. The simulation domain is 1D with a
-reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed (by mass)
+reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed\footnote{
+\added{We note that the WD simulations that inspired this were pure carbon/oxygen,
+but for the setup we describe here, pure carbon/oxygen conditions do not detonate at all.
+This is not particularly surprising, since the 1D setup is a very imperfect representation
+of the real multi-dimensional case, and is missing multi-dimensional hydrodynamics that could
+substantially alter the dynamical evolution. So we added a small amount of helium to ensure
+that this setup ignites.}} (by mass)
 of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He.   The fluid is relatively cold,

### 2019-02-06 papers/ignition/paper.tex
Clean up conclusion

@@
 hydrodynamical burning problem, a grid resolution requirement much more stringent
-than 1 km. This does not represent all such \added{1D} burning problems: for example, we
+than 1 km. \replaced{This does not represent all such burning problems: for example, we
 have tested similar setups with only carbon and oxygen for which a resolution
-in the tens of kilometers is sufficient. \added{However, this should not be construed
-as implying that this convergence requirement does not apply to pure C/O WDs.}
-\replaced{However}{First}, the fact that it is even possible for
+in the tens of kilometers is sufficient.}{This test does not, of course, represent
+all possible WD burning conditions.}
+However, the fact that it is even possible for
 \replaced{this type of burning}{burning in white dwarf material} to require a
 resolution better than 100 m should cast significant uncertainty on the results
-of simulations featuring detonations that do not meet this resolution
-\added{(regardless of the specific conditions). Or, at least, should suggest
-that stronger demonstrations of convergence are required. Second, in our real
-multi-dimensional simulations of C/O WD collisions, we see the same effect,
-where the simulation is not converged even for resolutions up to 0.25 km. It
-simply happens to be the case that for the 1D analogue of this problem using pure
-C/O the same effect is not seen, which is not particularly surprising since the 1D
-setup and the 2D setup are very different and the former lacks any multidimensional
-hydrodynamic effects. So we found a related problem that shows a similar
-effect in 1D to test convergence at higher resolutions.}
+of simulations featuring detonations that do not meet this resolution,
+\added{regardless of the specific conditions). Or, at least, it should suggest
+that stronger demonstrations of convergence are required (especially bearing in mind
+our observation that the numerical instability can result in simulations that
+appear qualitatively converged when the resolution is increased by a factor of
+one or two orders of magnitude but not three orders of magnitude.}
 

### 2019-02-06 papers/ignition/paper.tex
Rename conclusion section
Add comment about timestep dependence

@@
 we could achieve, about 50 cm, $\tau_s / \tau_e$ was 0.8 at ignition, which is not
-sufficiently small to be confident of numerical stability at that resolution.}
+sufficiently small to be confident of numerical stability at that resolution. Note
+that merely decreasing the timestep (at fixed resolution) does not help here, as the
+instability criterion is, to first order, independent of the size of the timestep.}
+
 We thus investigate whether limiting the energy release of the burn (we will term this
@@
 %==========================================================================
-% Conclusions
+% Conclusion
 %==========================================================================
-\section{Conclusions and Discussion}\label{Sec:Conclusions and Discussion}
+\section{Conclusion}\label{Sec:Conclusion}
 \label{sec:conclusion}

### 2019-02-06 papers/ignition/paper.tex
Mention density at the ignition time

@@
 The simulation is terminated when the peak temperature on the domain first reaches
-$4 \times 10^9$ K, which we call a thermonuclear ignition. This stopping criterion is a
+$4 \times 10^9$ K, which we call a thermonuclear ignition (\added{for reference, the
+density at the location where the ignition occurs is approximately $1.4\times 10^7\ \text{g / cm}^3$}). This stopping criterion is a
 proxy for the beginning of a detonation. Reaching this temperature does not guarantee
@@
 sufficiently small to be confident of numerical stability at that resolution. Note
-that merely decreasing the timestep (at fixed resolution) does not help here, as the
+that merely decreasing the timestep (at fixed resolution) does not help here either, as the
 instability criterion is, to first order, independent of the size of the timestep.}

### 2019-02-06 papers/ignition/paper.tex
Remove comment

@@
 
-\added{To summarize, then, satisfying the stability criterion is neither necessary nor
-sufficient, in general, for numerical convergence. It is not necessary because of the
-counter-example provided above of a very large hotspot, and it is not sufficient, as the
-empirical example in \autoref{fig:suppressed-distance} demonstrates. Nevertheless,
-it is a very useful diagnostic metric for understanding simulations, and it is likely
-that in many real cases, violation of the stability criterion is a clear sign of
-the simulation being under-resolved.}
-
 %==========================================================================

### 2019-02-06 papers/ignition/paper.tex
Minor modification to convergence language

@@
 central ignition does not occur, but rather the ignition is delayed and occurs further from the contact
-point. When we have finally \replaced{(perhaps) converged}{reached the point where the curves start to flatten},
+point. When we have finally \replaced{(perhaps) converged}{reached the point where the curves start to flatten
+and perhaps begin to converge}, 
 the ignition occurs around 900 km from the contact point, about 1 second after contact (contrast to less than

### 2019-02-06 papers/ignition/paper.tex
Add software tag
Add figure of ts_te

@@
 \citet{kushnir:2013} observe an important possible failure mode
-for reacting hydrodynamics simulations. Let us define $\tau_e = e / \dot{e}$
-as the nuclear energy injection timescale, and $\tau_s = \Delta x / c_s$
+for reacting hydrodynamics simulations. Let us define $\tau_{\rm e} = e / \dot{e}$
+as the nuclear energy injection timescale, and $\tau_{\rm s} = \Delta x / c_{\rm s}$
 as the sound-crossing time in a zone (where $\Delta x$ is the grid
-resolution and $c_s$ is the speed of sound). When the sound-crossing
+resolution and $c_{\rm s}$ is the speed of sound). When the sound-crossing
 time is too long, energy is built up in a zone faster than it can be
 advected away by pressure waves. \added{This effect generalizes to
-Langrangian simulations as well, where $\tau_s$ should be understood
+Langrangian simulations as well, where $\tau_{\rm s}$ should be understood
 as the timescale for transport of energy to a neighboring fluid element.}
@@
 caused by the temperature building up too quickly in the zone; the
-detonation may be spurious in this case. If $\tau_s \ll \tau_e$,
+detonation may be spurious in this case. If $\tau_{\rm s} \ll \tau_{\rm e}$,
 we can be confident that a numerically seeded detonation has not
@@
 \begin{equation}
-  \tau_s \leq f_{s}\, \tau_e \label{eq:burning_limiter}
+  \tau_{\rm s} \leq f_{\rm s}\, \tau_{\rm e} \label{eq:burning_limiter}
 \end{equation}
 and require that $f_{s}$ is sufficiently smaller than one.
-\citet{kushnir:2013} state that $f_{s} = 0.1$ is a sufficient
+\citet{kushnir:2013} state that $f_{\rm s} = 0.1$ is a sufficient
 criterion for avoiding premature ignitions. \citeauthor{kushnir:2013}
@@
 We find that for our test problem (and also the WD collisions we have simulated)
-we do observe $\tau_s > \tau_e$; typically the ratio is a factor of 2--5 at
-low resolution. This means that an ignition is very likely to occur for numerical
-reasons, regardless of whether it would occur for physical reasons. \added{At low
-resolution, adding more resolution does not meaningfully improve the ratio of
+we do observe $\tau_{\rm s} > \tau_{\rm e}$; typically the ratio is a factor of 2--5 at
+low resolution (\added{see \autoref{fig:self-heat-ts_te}}). This means that an ignition is
+very likely to occur for numerical reasons, regardless of whether it would occur for physical reasons.
+\added{At low resolution, adding more resolution does not meaningfully improve the ratio of
 $\tau_s$ to $\tau_e$ at the point of ignition. The ignition timescale is so short
@@
 instability criterion is, to first order, independent of the size of the timestep.}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.30]{{{plots/amr_ignition_self-heat_ts_te}}}
+  \caption{Ratio of the sound-crossing timescale to the energy injection timescale,
+           for the simulations in \autoref{fig:self-heat-distance}.
+           \label{fig:self-heat-ts_te}}
+\end{figure}
 
@@
 \facility{OLCF}
+\software{\castro\ \citep{castro, astronum:2017},
+          \amrex\ \citep{boxlib-tiling},
+          \yt\ \citep{yt}
+          }
 

### 2019-02-06 papers/ignition/paper.tex
Add comment about remnant helium

@@
 substantially alter the dynamical evolution. So we added a small amount of helium to ensure
-that this setup ignites.}} (by mass)
+that this setup ignites. Of course, there will likely be a small amount of helium present in
+C/O white dwarfs as a remnant of the prior stellar evolution.}} (by mass)
 of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He.   The fluid is relatively cold,

### 2019-02-06 papers/ignition/paper.tex
Add last comment before resubmission

@@
 turn to 1D simulations, where we can achieve much higher resolution (at the cost, of course,
-of not being able to do a test that can be directly compared to multi-dimensional simulations).}
+of not being able to do a test that can be directly compared to multi-dimensional simulations).
+Our goal is to record our thoughts on why we believe we did not achieve convergence.}
 

### 2019-02-06 papers/ignition/submit.sh
Revised submission

@@ -1,4 +1,4 @@
-submit_dir="submit"
+submit_dir="submit2"
 
 # Build the PDF so that we have the bibilography
 

### 2019-02-23 papers/ignition/paper.tex
Remove tracked changes

@@
 hawley:2012,garcia-senz:2013} typically had effective spatial resolutions in the
-range 100--500 km \added{for the grid codes, and 10--100 km for the SPH codes}, and observed detonations that convert a large amount of
-carbon/oxygen material into iron-group elements. These studies varied in methodology
-(Lagrangian versus Eulerian evolution, nuclear network used) and did not closely agree
-on the final result of the event (see Table 4 of \cite{garcia-senz:2013} for a summary).
-\cite{kushnir:2013} argued that many of these simulations featured numerically unstable
-evolution, ultimately caused by the zone size being significantly larger than the length
-scale over which detonations form. The detonation length scale can vary widely based
-on physical conditions \citep{seitenzahl:2009,garg:2017} but is generally \replaced{much
-smaller than 100 km}{not larger than 10 km}.
+range 100--500 km for the grid codes, and 10--100 km for the SPH codes, and observed
+detonations that convert a large amount of carbon/oxygen material into iron-group elements.
+These studies varied in methodology (Lagrangian versus Eulerian evolution, nuclear network
+used) and did not closely agree on the final result of the event (see Table 4 of
+\cite{garcia-senz:2013} for a summary). \cite{kushnir:2013} argued that many of these
+simulations featured numerically unstable evolution, ultimately caused by the zone size
+being significantly larger than the length scale over which detonations form. The detonation
+length scale can vary widely based on physical conditions \citep{seitenzahl:2009,garg:2017}
+but is generally not larger than 10 km.
 
 In this paper, we attempt to find what simulation length scale is required to achieve
-converged thermonuclear ignitions. \added{The inspiration for this work comes from our
+converged thermonuclear ignitions. The inspiration for this work comes from our
 simulations of WD collisions using the reactive hydrodynamics code \castro\
@@
 of not being able to do a test that can be directly compared to multi-dimensional simulations).
-Our goal is to record our thoughts on why we believe we did not achieve convergence.}
+Our goal is to record our thoughts on why we believe we did not achieve convergence.
 
@@
 reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed\footnote{
-\added{We note that the WD simulations that inspired this were pure carbon/oxygen,
+We note that the WD simulations that inspired this were pure carbon/oxygen,
 but for the setup we describe here, pure carbon/oxygen conditions do not detonate at all.
@@
 that this setup ignites. Of course, there will likely be a small amount of helium present in
-C/O white dwarfs as a remnant of the prior stellar evolution.}} (by mass)
+C/O white dwarfs as a remnant of the prior stellar evolution.} (by mass)
 of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He.   The fluid is relatively cold,
@@
 Simulations are performed with the adaptive mesh refinement (AMR) code \castro.
-For the burning we use the alpha-chain nuclear network
-\texttt{aprox13}. \deleted{The results we describe below do not strongly depend on the network.}
+For the burning we use the alpha-chain nuclear network \texttt{aprox13}.
 Release 18.12 of the \castro\ code was used. The \amrex\ and \microphysics\ repositories
@@
 The simulation is terminated when the peak temperature on the domain first reaches
-$4 \times 10^9$ K, which we call a thermonuclear ignition (\added{for reference, the
-density at the location where the ignition occurs is approximately $1.4\times 10^7\ \text{g / cm}^3$}). This stopping criterion is a
+$4 \times 10^9$ K, which we call a thermonuclear ignition (for reference, the
+density at the location where the ignition occurs is approximately $1.4\times 10^7\ \text{g / cm}^3$). This stopping criterion is a
 proxy for the beginning of a detonation. Reaching this temperature does not guarantee
 that a detonation will begin, and in this study we do not directly address the question
-of whether a ignition of this kind always leads to a detonation. \added{Nor are we commenting
-on the physics of the ignition process itself.} Rather, the main question
-we investigate here is whether this ignition is numerically converged, \added{and for this purpose
+of whether a ignition of this kind always leads to a detonation. Nor are we commenting
+on the physics of the ignition process itself. Rather, the main question
+we investigate here is whether this ignition is numerically converged, and for this purpose
 this arbitrary stopping point is sufficient, since in a converged simulation the stopping point
-should be reached at the same time independent of resolution}. A converged ignition
+should be reached at the same time independent of resolution. A converged ignition
 is a prerequisite to having a converged detonation. We measure two diagnostic quantities:
@@
 central ignition does not occur, but rather the ignition is delayed and occurs further from the contact
-point. When we have finally \replaced{(perhaps) converged}{reached the point where the curves start to flatten
-and perhaps begin to converge}, 
+point. When we have finally reached the point where the curves start to flatten and perhaps begin to converge, 
 the ignition occurs around 900 km from the contact point, about 1 second after contact (contrast to less than
-0.05 seconds for the simulation with 1 km resolution). \added{Even at this resolution, it is not clear if
+0.05 seconds for the simulation with 1 km resolution). Even at this resolution, it is not clear if
 the simulation is converged. We were unable to perform higher resolution simulations to check convergence
-due to the length of time that would be required.} This story contains two important lessons. First,
+due to the length of time that would be required. This story contains two important lessons. First,
 the required resolution for even a qualitatively converged simulation, less than 100 m, is out of reach
@@
 time is too long, energy is built up in a zone faster than it can be
-advected away by pressure waves. \added{This effect generalizes to
+advected away by pressure waves. This effect generalizes to
 Langrangian simulations as well, where $\tau_{\rm s}$ should be understood
-as the timescale for transport of energy to a neighboring fluid element.}
+as the timescale for transport of energy to a neighboring fluid element.
 This is of course a problem inherent 
@@
 we do observe $\tau_{\rm s} > \tau_{\rm e}$; typically the ratio is a factor of 2--5 at
-low resolution (\added{see \autoref{fig:self-heat-ts_te}}). This means that an ignition is
+low resolution (see \autoref{fig:self-heat-ts_te}). This means that an ignition is
 very likely to occur for numerical reasons, regardless of whether it would occur for physical reasons.
-\added{At low resolution, adding more resolution does not meaningfully improve the ratio of
+At low resolution, adding more resolution does not meaningfully improve the ratio of
 $\tau_s$ to $\tau_e$ at the point of ignition. The ignition timescale is so short
@@
 that merely decreasing the timestep (at fixed resolution) does not help here either, as the
-instability criterion is, to first order, independent of the size of the timestep.}
+instability criterion is, to first order, independent of the size of the timestep.
 
@@
 ``suppressing'' the burn), as proposed by \citeauthor{kushnir:2013},
-is a useful technique for avoiding the prompt detonation. \added{Since the limiter ensures
+is a useful technique for avoiding the prompt detonation. Since the limiter ensures
 the inequality in \autoref{eq:burning_limiter} holds by construction, the specific
 question to ask is whether the limiter achieves the correct answer and is converged
-in cases where the simulation would otherwise be uncorrect or unconverged.}
+in cases where the simulation would otherwise be uncorrect or unconverged.
 
@@
 hydrodynamical burning problem, a grid resolution requirement much more stringent
-than 1 km. \replaced{This does not represent all such burning problems: for example, we
-have tested similar setups with only carbon and oxygen for which a resolution
-in the tens of kilometers is sufficient.}{This test does not, of course, represent
-all possible WD burning conditions.}
-However, the fact that it is even possible for
-\replaced{this type of burning}{burning in white dwarf material} to require a
+than 1 km. This test does not, of course, represent all possible WD burning conditions.
+However, the fact that it is even possible for burning in white dwarf material to require a
 resolution better than 100 m should cast significant uncertainty on the results
 of simulations featuring detonations that do not meet this resolution,
-\added{regardless of the specific conditions). Or, at least, it should suggest
+regardless of the specific conditions). Or, at least, it should suggest
 that stronger demonstrations of convergence are required (especially bearing in mind
@@
 appear qualitatively converged when the resolution is increased by a factor of
-one or two orders of magnitude but not three orders of magnitude.}
+one or two orders of magnitude but not three orders of magnitude.
 

### 2019-02-23 papers/ignition/paper.tex
Add some discussion of whether previous papers were converged

@@
 used) and did not closely agree on the final result of the event (see Table 4 of
-\cite{garcia-senz:2013} for a summary). \cite{kushnir:2013} argued that many of these
-simulations featured numerically unstable evolution, ultimately caused by the zone size
-being significantly larger than the length scale over which detonations form. The detonation
-length scale can vary widely based on physical conditions \citep{seitenzahl:2009,garg:2017}
-but is generally not larger than 10 km.
+\cite{garcia-senz:2013} for a summary).
+
+\added{There is mixed evidence for simulation convergence presented in these studies.
+\cite{raskin:2010} claim that their simulations are converged in nickel yield up to 2 million
+(constant mass) particles (Figure 11 in that paper), but the nickel yield still appears to be
+trending slightly upward with particle count. The simulations from \cite{raskin:2009} are not
+converged up to 800,000 particles (Figure 4 in that paper), where the smoothing length was kept
+constant instead of the particle mass. \cite{hawley:2012} do not achieve convergence over a
+factor of 2 in spatial resolution (Figure 8 in that paper). \cite{garcia-senz:2013} claim at
+least qualitative (though not strict absolute) convergence in Section 3.2 of their paper, but
+their convergence test is only over a factor of 2 in particle count, which is a factor of 1.26
+in spatial resolution for constant mass particles. \cite{kushnir:2013} test convergence over an
+order of magnitude in spatial resolution, and find results that appear to be reasonably well
+converged for one of the two codes used (VULCAN2D), and results that are not converged for the
+other code (FLASH). \cite{papish:2015} claim convergence in nuclear burning up to 10\% at a
+resolution of 5--10 km, but do not present specific data demonstrating this claim.
+\cite{loren-aguilar:2010} and \cite{rosswog:2009} do not present convergence studies for
+their work.}
+
+\cite{kushnir:2013} argued that many of these simulations featured numerically unstable evolution,
+ultimately caused by the zone size being significantly larger than the length scale over which
+detonations form. The detonation length scale can vary widely based on physical conditions
+\citep{seitenzahl:2009,garg:2017} but is generally not larger than 10 km. \added{\citeauthor{kushnir:2013}
+argue that this numerically unstable evolution is the primary cause of convergence difficulties.
+They further argue that it is possible to apply a burning limiter to achieve converged results,
+which was used in both their work and the simulations of \cite{papish:2015}. We will investigate
+this hypothesis in \autoref{sec:unstable_burning}.}
 

### 2019-02-23 papers/ignition/paper.tex
Rephrase last sentence in introduction

@@
 of not being able to do a test that can be directly compared to multi-dimensional simulations).
-Our goal is to record our thoughts on why we believe we did not achieve convergence.
+\replaced{Our goal is to record our thoughts on why we believe we did not achieve convergence.}
+{We believe the simulations presented below shed light onto why we and others had difficulty
+achieving convergence at the resolutions achievable in multi-dimensional WD collision simulations.}
 

### 2019-02-23 papers/ignition/paper.tex
Add a pure C/O test

@@
 reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed\footnote{
-We note that the WD simulations that inspired this were pure carbon/oxygen,
+\deleted{We note that the WD simulations that inspired this were pure carbon/oxygen,
 but for the setup we describe here, pure carbon/oxygen conditions do not detonate at all.
@@
 that this setup ignites. Of course, there will likely be a small amount of helium present in
-C/O white dwarfs as a remnant of the prior stellar evolution.} (by mass)
+C/O white dwarfs as a remnant of the prior stellar evolution.}} (by mass)
 of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He.   The fluid is relatively cold,
@@
 the simulation is converged. We were unable to perform higher resolution simulations to check convergence
-due to the length of time that would be required. This story contains two important lessons. First,
-the required resolution for even a qualitatively converged simulation, less than 100 m, is out of reach
-of current 3D supernova simulations. Second, the behavior for resolutions worse than 1 km qualitatively
-appears to be converged, and one could perhaps be misled into thinking that there was no reason to
-try higher resolutions, which is reason for caution in interpreting reacting hydrodynamics simulations.
+due to the length of time that would be required.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.30]{{{plots/amr_ignition_co}}}
+  \caption{Similar to \autoref{fig:self-heat-distance}, but with pure C/O material.
+           \label{fig:self-heat-distance-co}}
+\end{figure}
+
+\added{We also tested a similar configuration made of pure carbon/oxygen material (equal fraction by mass).
+This is closer to the configuration used in the $0.64\ \msolar$ WD collision simulations that previous
+papers have focused on. However, for the setup described above, pure carbon/oxygen conditions do not
+detonate at all. This is not particularly surprising, since the 1D setup is a very imperfect representation
+of the real multi-dimensional case, and is missing multi-dimensional hydrodynamics that could substantially
+alter the dynamical evolution. So the small amount of helium we added above ensured that the setup ignited.
+(Of course, there will likely be a small amount of helium present in C/O white dwarfs as a remnant of the prior
+stellar evolution.) However, we can prompt the C/O setup to ignite by starting the initial temperature
+at $10^9$ K instead of $10^7$ K. This loosely mimics the effect from the first test, where helium burning
+drives the temperature to the conditions necessary to begin substantial burning in C/O material. But since
+no helium is present in this case, it allows us to test whether it is easier to obtain convergence for pure
+C/O burning. The results for this case are shown in \autoref{fig:self-heat-distance-co}. In this case, the
+ignition is central at all resolutions, but the simulation is still clearly unconverged at resolutions worse
+than 100 m.}
+
+This story contains two important lessons. First, the required resolution for even a qualitatively converged
+simulation, less than 100 m, is out of reach of current 3D supernova simulations. Second, the behavior for
+resolutions worse than 1 km qualitatively appears to be converged, and one could perhaps be misled into
+thinking that there was no reason to try higher resolutions, which is reason for caution in interpreting
+reacting hydrodynamics simulations.
 
@@
 timesteps and the advection of energy across multiple zones. Even at the highest resolution
-we could achieve, about 50 cm, $\tau_s / \tau_e$ was 0.8 at ignition, which is not
+we could achieve \added{for the test including helium}, about 50 cm, $\tau_s / \tau_e$ was 0.8 at ignition, which is not
 sufficiently small to be confident of numerical stability at that resolution. Note
@@
 of simulations featuring detonations that do not meet this resolution,
-regardless of the specific conditions). Or, at least, it should suggest
-that stronger demonstrations of convergence are required (especially bearing in mind
+regardless of the specific conditions. Or, at least, it should suggest
+that stronger demonstrations of convergence are required, especially bearing in mind
 our observation that the numerical instability can result in simulations that

### 2019-02-23 papers/ignition/paper.tex
Add caveats about the implications

@@
 This story contains two important lessons. First, the required resolution for even a qualitatively converged
-simulation, less than 100 m, is out of reach of current 3D supernova simulations. Second, the behavior for
+simulation, less than 100 m, is out of reach \replaced{of current 3D supernova simulations}{for an analogous
+simulation done in 3D}. Second, the behavior for 
 resolutions worse than 1 km qualitatively appears to be converged, and one could perhaps be misled into
 thinking that there was no reason to try higher resolutions, which is reason for caution in interpreting
-reacting hydrodynamics simulations.
+reacting hydrodynamics simulations. \added{With that being said, our 1D tests are not directly comparable to
+previous multi-dimensional WD collision simulations. The 1D tests should not be substituted for understanding the
+actual convergence properties of the 2D/3D simulations, which may have different resolution requirements for
+convergence. Our tests suggest only that it is plausible that simulations at kilometer-scale (or worse) resolution
+are unconverged. This observation is, though, consistent with the situation described in \autoref{sec:introduction},
+where our 2D WD collision simulations (not shown here) are unconverged, and many of the previous collision simulations
+presented in the literature have relatively weak evidence for convergence.}
 
@@
 However, the fact that it is even possible for burning in white dwarf material to require a
-resolution better than 100 m should cast significant uncertainty on the results
+resolution better than 100 m should \deleted{cast significant uncertainty on the results
 of simulations featuring detonations that do not meet this resolution,
-regardless of the specific conditions. Or, at least, it should suggest
-that stronger demonstrations of convergence are required, especially bearing in mind
+regardless of the specific conditions. Or, at least, it should} suggest
+that stronger demonstrations of convergence are required \added{for simulations that
+do not meet this resolution}. \replaced{especially} {This is especially true} bearing in mind
 our observation that the numerical instability can result in simulations that

### 2019-02-23 papers/ignition/paper.tex
Point out the temporal behavior of the C/O test

@@
 ignition is central at all resolutions, but the simulation is still clearly unconverged at resolutions worse
-than 100 m.}
+than 100 m, as the ignition is prompt at low resolution and delayed at high resolution.}
 

### 2019-02-23 papers/ignition/paper.tex
Add caveat about the test being artificial

--- 2019-02-23 papers/ignition/paper.tex
@@
 no helium is present in this case, it allows us to test whether it is easier to obtain convergence for pure
-C/O burning. The results for this case are shown in \autoref{fig:self-heat-distance-co}. In this case, the
-ignition is central at all resolutions, but the simulation is still clearly unconverged at resolutions worse
-than 100 m.}
+C/O burning, even though the test itself is artificial. The results for this case are shown in
+\autoref{fig:self-heat-distance-co}. In this case, the ignition is central at all resolutions, but the
+simulation is still clearly unconverged at resolutions worse than 100 m, as the ignition becomes significantly
+delayed at high resolution.}
 

### 2019-02-25 papers/ignition/paper.tex
Take a pass through for style

@@
 detonations. However, these detonations are not meaningful if the simulations
-are not resolved, so it is important to establish an understanding of the
+are not resolved, so it is important to establish \deleted{an understanding of} the
 requirements for achieving a numerically converged detonation. In this
@@
 Consequently, contemporary simulations that self-consistently demonstrate
-detonations are very likely not resolved and should be treated with caution.
+detonations are \replaced{very likely not resolved}{possibly not converged}
+and should be treated with caution.
 \end{abstract}
@@
 WD collisions rapidly convert a significant amount of kinetic energy into thermal energy and
-thus set up conditions ripe for a thermonuclear detonation. Since they are easy to set up,
+thus set up conditions ripe for a thermonuclear detonation. Since they are easy to set up \added{in a simulation},
 they are a useful vehicle for studying the properties of detonations.
@@
 Early studies on WD collisions \citep{rosswog:2009,raskin:2010,loren-aguilar:2010,
-hawley:2012,garcia-senz:2013} typically had effective spatial resolutions in the
-range 100--500 km for the grid codes, and 10--100 km for the SPH codes, and observed
+hawley:2012,garcia-senz:2013} typically had effective spatial resolutions \added{in the burning region of}
+\deleted{range} 100--500 km for the grid codes, and 10--100 km for the SPH codes, and observed
 detonations that convert a large amount of carbon/oxygen material into iron-group elements.
@@
 \cite{raskin:2010} claim that their simulations are converged in nickel yield up to 2 million
-(constant mass) particles (Figure 11 in that paper), but the nickel yield still appears to be
-trending slightly upward with particle count. The simulations from \cite{raskin:2009} are not
-converged up to 800,000 particles (Figure 4 in that paper), where the smoothing length was kept
-constant instead of the particle mass. \cite{hawley:2012} do not achieve convergence over a
-factor of 2 in spatial resolution (Figure 8 in that paper). \cite{garcia-senz:2013} claim at
-least qualitative (though not strict absolute) convergence in Section 3.2 of their paper, but
-their convergence test is only over a factor of 2 in particle count, which is a factor of 1.26
-in spatial resolution for constant mass particles. \cite{kushnir:2013} test convergence over an
+(constant mass) particles, but the nickel yield still appears to be trending slightly upward
+with particle count. The earlier simulations of \cite{raskin:2009} are not converged up to
+800,000 particles, where the smoothing length was kept constant instead of the particle mass.
+\cite{hawley:2012} do not achieve convergence over a factor of 2 in spatial resolution.
+\cite{garcia-senz:2013} claim at least qualitative (though not strict absolute) convergence, but
+their convergence test is only over a factor of 2 in particle count, which is a factor of 1.3
+in spatial resolution (for constant mass particles). \cite{kushnir:2013} test convergence over an
 order of magnitude in spatial resolution, and find results that appear to be reasonably well
 converged for one of the two codes used (VULCAN2D), and results that are not converged for the
-other code (FLASH). \cite{papish:2015} claim convergence in nuclear burning up to 10\% at a
+other code used (FLASH). \cite{papish:2015} claim convergence in nuclear burning up to 10\% at a
 resolution of 5--10 km, but do not present specific data demonstrating this claim.
@@
 They further argue that it is possible to apply a burning limiter to achieve converged results,
-which was used in both their work and the simulations of \cite{papish:2015}. We will investigate
+which was used in their work and later the simulations of \cite{papish:2015}. We will investigate
 this hypothesis in \autoref{sec:unstable_burning}.}
@@
 \replaced{Our goal is to record our thoughts on why we believe we did not achieve convergence.}
-{We believe the simulations presented below shed light onto why we and others had difficulty
+{We believe the simulations presented below help show why we and others had difficulty
 achieving convergence at the resolutions achievable in multi-dimensional WD collision simulations.}
@@
 no helium is present in this case, it allows us to test whether it is easier to obtain convergence for pure
-C/O burning, even though the test itself is artificial. The results for this case are shown in
+C/O burning, even though the test itself is artificial. The only other change relative to the prior test is
+that we refined on relative temperature gradients of 25\% instead of 50\%. The results for this case are shown in
 \autoref{fig:self-heat-distance-co}. In this case, the ignition is central at all resolutions, but the
@@
 are continuous. This can lead to a numerically seeded detonation
-caused by the temperature building up too quickly in the zone; the
+caused by the temperature building up too quickly in the zone. The
 detonation may be spurious in this case. If $\tau_{\rm s} \ll \tau_{\rm e}$,
 we can be confident that a numerically seeded detonation has not
-occurred. In practice, we will quantify this requirement as:
+occurred. In practice, we \deleted{will} quantify this requirement as:
 \begin{equation}
@@
 $\tau_s$ to $\tau_e$ at the point of ignition. The ignition timescale is so short
-that virtually all of the energy release occurs in a single timestep even if the
+that \replaced{virtually}{almost} all of the energy release occurs in a single timestep even \replaced{if}{though} the
 timestep gets shorter due to the CFL limiter. It is only when the resolution gets
@@
 we could achieve \added{for the test including helium}, about 50 cm, $\tau_s / \tau_e$ was 0.8 at ignition, which is not
-sufficiently small to be confident of numerical stability at that resolution. Note
+sufficiently small to be confident of numerical stability \deleted{at that resolution}. Note
 that merely decreasing the timestep (at fixed resolution) does not help here either, as the
@@
   \includegraphics[scale=0.30]{{{plots/amr_ignition_self-heat_ts_te}}}
-  \caption{Ratio of the sound-crossing timescale to the energy injection timescale,
+  \caption{Ratio of the sound-crossing timescale to the energy injection timescale
            for the simulations in \autoref{fig:self-heat-distance}.
@@
 Suppression of the burn in this case is unphysical: if we have a zone matching
-these characteristics, the zone should detonate.
+these characteristics, the zone should \replaced{detonate}{ignite}.
 
@@
 to develop (which is the implicit assumption of a low resolution simulation), then
-it would likely detonate. If the results do not match what occurs at higher
+it would likely \replaced{detonate}{ignite}. If the results do not match what occurs at higher
 resolution, then the simulation is not converged and the results are not reliable.
@@
 
-\facility{OLCF}
+\facilities{OLCF, NERSC}
 \software{\castro\ \citep{castro, astronum:2017},

### 2019-02-27 papers/ignition/paper.tex
Cite matplotlib
Update C/O plot

@@
 \cite{garcia-senz:2013} claim at least qualitative (though not strict absolute) convergence, but
-their convergence test is only over a factor of 2 in particle count, which is a factor of 1.3
+their convergence test is only over a factor of 2 in particle count, which is a factor of $2^{1/3} = 1.3$
 in spatial resolution (for constant mass particles). \cite{kushnir:2013} test convergence over an
@@
 other code used (FLASH). \cite{papish:2015} claim convergence in nuclear burning up to 10\% at a
-resolution of 5--10 km, but do not present specific data demonstrating this claim.
-\cite{loren-aguilar:2010} and \cite{rosswog:2009} do not present convergence studies for
-their work.}
+resolution of 5--10 km, but do not present specific data demonstrating this claim or precisely define
+what is being measured. \cite{loren-aguilar:2010} and \cite{rosswog:2009} do not present convergence
+studies for their work.}
 
@@
 They further argue that it is possible to apply a burning limiter to achieve converged results,
-which was used in their work and later the simulations of \cite{papish:2015}. We will investigate
+which was used in their work and later the simulations of \cite{papish:2015}. We investigate
 this hypothesis in \autoref{sec:unstable_burning}.}
@@
 stellar evolution.) However, we can prompt the C/O setup to ignite by starting the initial temperature
-at $10^9$ K instead of $10^7$ K. This loosely mimics the effect from the first test, where helium burning
+at $10^9$ K instead of $10^7$ K. This loosely mimics the effect from the first test where helium burning
 drives the temperature to the conditions necessary to begin substantial burning in C/O material. But since
@@
           \amrex\ \citep{boxlib-tiling},
-          \yt\ \citep{yt}
+          \yt\ \citep{yt},
+          \matplotlib\ \citep{matplotlib}
           }

### 2019-02-27 papers/ignition/submit.sh
Revised draft

@@ -1,4 +1,4 @@
-submit_dir="submit2"
+submit_dir="submit3"
 
 # Build the PDF so that we have the bibilography
 

### 2019-02-27 papers/ignition/paper.tex
Add note on C/O figure
Update shorttitle
Update shortauthors
Replace track changes

@@
 
-\shorttitle{Numerical Detonations}
-\shortauthors{Katz et al. (2018)}
+\shorttitle{Detonation Stability}
+\shortauthors{Katz and Zingale (2019)}
 
@@
 detonations. However, these detonations are not meaningful if the simulations
-are not resolved, so it is important to establish \deleted{an understanding of} the
-requirements for achieving a numerically converged detonation. In this
-study we examine a test detonation problem inspired by collisions of white dwarfs.
-This test problem demonstrates that achieving a converged thermonuclear ignition
-requires spatial resolution much finer than 1 km in the burning region.
-Current computational resource constraints place this stringent resolution
-requirement out of reach for multi-dimensional supernova simulations.
-Consequently, contemporary simulations that self-consistently demonstrate
-detonations are \replaced{very likely not resolved}{possibly not converged}
+are not resolved, so it is important to establish the requirements for achieving
+a numerically converged detonation. In this study we examine a test detonation
+problem inspired by collisions of white dwarfs. This test problem demonstrates
+that achieving a converged thermonuclear ignition requires spatial resolution
+much finer than 1 km in the burning region. Current computational resource
+constraints place this stringent resolution requirement out of reach for
+multi-dimensional supernova simulations. Consequently, contemporary simulations
+that self-consistently demonstrate detonations are possibly not converged
 and should be treated with caution.
@@
 WD collisions rapidly convert a significant amount of kinetic energy into thermal energy and
-thus set up conditions ripe for a thermonuclear detonation. Since they are easy to set up \added{in a simulation},
+thus set up conditions ripe for a thermonuclear detonation. Since they are easy to set up in a simulation,
 they are a useful vehicle for studying the properties of detonations.
@@
 Early studies on WD collisions \citep{rosswog:2009,raskin:2010,loren-aguilar:2010,
-hawley:2012,garcia-senz:2013} typically had effective spatial resolutions \added{in the burning region of}
-\deleted{range} 100--500 km for the grid codes, and 10--100 km for the SPH codes, and observed
+hawley:2012,garcia-senz:2013} typically had effective spatial resolutions in the burning region of
+100--500 km for the grid codes, and 10--100 km for the SPH codes, and observed
 detonations that convert a large amount of carbon/oxygen material into iron-group elements.
@@
 
-\added{There is mixed evidence for simulation convergence presented in these studies.
+There is mixed evidence for simulation convergence presented in these studies.
 \cite{raskin:2010} claim that their simulations are converged in nickel yield up to 2 million
@@
 what is being measured. \cite{loren-aguilar:2010} and \cite{rosswog:2009} do not present convergence
-studies for their work.}
+studies for their work.
 
@@
 detonations form. The detonation length scale can vary widely based on physical conditions
-\citep{seitenzahl:2009,garg:2017} but is generally not larger than 10 km. \added{\citeauthor{kushnir:2013}
+\citep{seitenzahl:2009,garg:2017} but is generally not larger than 10 km. \citeauthor{kushnir:2013}
 argue that this numerically unstable evolution is the primary cause of convergence difficulties.
@@
 which was used in their work and later the simulations of \cite{papish:2015}. We investigate
-this hypothesis in \autoref{sec:unstable_burning}.}
+this hypothesis in \autoref{sec:unstable_burning}.
 
@@
 of not being able to do a test that can be directly compared to multi-dimensional simulations).
-\replaced{Our goal is to record our thoughts on why we believe we did not achieve convergence.}
-{We believe the simulations presented below help show why we and others had difficulty
-achieving convergence at the resolutions achievable in multi-dimensional WD collision simulations.}
+We believe the simulations presented below help show why we and others had difficulty
+achieving convergence at the resolutions achievable in multi-dimensional WD collision simulations.
 
@@
 conditions of two $0.64\ \msolar$ WDs colliding head-on. The simulation domain is 1D with a
-reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed\footnote{
-\deleted{We note that the WD simulations that inspired this were pure carbon/oxygen,
-but for the setup we describe here, pure carbon/oxygen conditions do not detonate at all.
-This is not particularly surprising, since the 1D setup is a very imperfect representation
-of the real multi-dimensional case, and is missing multi-dimensional hydrodynamics that could
-substantially alter the dynamical evolution. So we added a small amount of helium to ensure
-that this setup ignites. Of course, there will likely be a small amount of helium present in
-C/O white dwarfs as a remnant of the prior stellar evolution.}} (by mass)
+reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed (by mass)
 of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He. The fluid is relatively cold,
@@
   \caption{Similar to \autoref{fig:self-heat-distance}, but with pure C/O material.
+           Note the different vertical axis scale.
            \label{fig:self-heat-distance-co}}
@@
 
-\added{We also tested a similar configuration made of pure carbon/oxygen material (equal fraction by mass).
+We also tested a similar configuration made of pure carbon/oxygen material (equal fraction by mass).
 This is closer to the configuration used in the $0.64\ \msolar$ WD collision simulations that previous
@@
 simulation is still clearly unconverged at resolutions worse than 100 m, as the ignition becomes significantly
-delayed at high resolution.}
+delayed at high resolution.
 
 This story contains two important lessons. First, the required resolution for even a qualitatively converged
-simulation, less than 100 m, is out of reach \replaced{of current 3D supernova simulations}{for an analogous
-simulation done in 3D}. Second, the behavior for 
+simulation, less than 100 m, is out of reach for an analogous simulation done in 3D. Second, the behavior for 
 resolutions worse than 1 km qualitatively appears to be converged, and one could perhaps be misled into
 thinking that there was no reason to try higher resolutions, which is reason for caution in interpreting
-reacting hydrodynamics simulations. \added{With that being said, our 1D tests are not directly comparable to
+reacting hydrodynamics simulations. With that being said, our 1D tests are not directly comparable to
 previous multi-dimensional WD collision simulations. The 1D tests should not be substituted for understanding the
@@
 where our 2D WD collision simulations (not shown here) are unconverged, and many of the previous collision simulations
-presented in the literature have relatively weak evidence for convergence.}
+presented in the literature have relatively weak evidence for convergence.
 
@@
 we can be confident that a numerically seeded detonation has not
-occurred. In practice, we \deleted{will} quantify this requirement as:
+occurred. In practice, we quantify this requirement as:
 \begin{equation}
@@
 $\tau_s$ to $\tau_e$ at the point of ignition. The ignition timescale is so short
-that \replaced{virtually}{almost} all of the energy release occurs in a single timestep even \replaced{if}{though} the
+that almost all of the energy release occurs in a single timestep even though the
 timestep gets shorter due to the CFL limiter. It is only when the resolution gets
@@
 timesteps and the advection of energy across multiple zones. Even at the highest resolution
-we could achieve \added{for the test including helium}, about 50 cm, $\tau_s / \tau_e$ was 0.8 at ignition, which is not
-sufficiently small to be confident of numerical stability \deleted{at that resolution}. Note
+we could achieve for the test including helium, about 50 cm, $\tau_s / \tau_e$ was 0.8 at ignition, which is not
+sufficiently small to be confident of numerical stability. Note
 that merely decreasing the timestep (at fixed resolution) does not help here either, as the
@@
 Suppression of the burn in this case is unphysical: if we have a zone matching
-these characteristics, the zone should \replaced{detonate}{ignite}.
+these characteristics, the zone should ignite.
 
@@
 to develop (which is the implicit assumption of a low resolution simulation), then
-it would likely \replaced{detonate}{ignite}. If the results do not match what occurs at higher
+it would likely ignite. If the results do not match what occurs at higher
 resolution, then the simulation is not converged and the results are not reliable.
@@
 However, the fact that it is even possible for burning in white dwarf material to require a
-resolution better than 100 m should \deleted{cast significant uncertainty on the results
-of simulations featuring detonations that do not meet this resolution,
-regardless of the specific conditions. Or, at least, it should} suggest
-that stronger demonstrations of convergence are required \added{for simulations that
-do not meet this resolution}. \replaced{especially} {This is especially true} bearing in mind
-our observation that the numerical instability can result in simulations that
-appear qualitatively converged when the resolution is increased by a factor of
-one or two orders of magnitude but not three orders of magnitude.
+resolution better than 100 m should suggest that stronger demonstrations of convergence are
+required. This is especially true bearing in mind our observation that the numerical
+instability can result in simulations that appear qualitatively converged when the
+resolution is increased by a factor of one or two orders of magnitude but not three
+orders of magnitude.
 

### 2019-03-01 papers/ignition/submit.sh
arXiv version

@@ -1,4 +1,4 @@
-submit_dir="submit3"
+submit_dir="arxiv"
 
 # Build the PDF so that we have the bibilography
 

### 2019-03-04 papers/ignition/submit.sh
arXiv replacement (remove extraneous figure

@@ -1,4 +1,4 @@
-submit_dir="arxiv"
+submit_dir="arxiv2"
 
 # Build the PDF so that we have the bibilography
 
